### 2018-07-18 Paper/app.tex
..

\section{Application to $H_2$/$O_2$ reaction mechanism: high-dimensional case}
\label{sec:app}

### 2018-07-18 Paper/assa.tex
..

\section{Active subspaces and sensitivity analysis}
\label{sec:assa}

\subsection{Active subspaces}
We recall some background regarding active subspaces~\cite{Constantine15}.
\subsection{Global sensitivity measures}
Define the total Sobol' indices $T_i(f)$ and discuss their use.

For $f: \Omega \to \R$, we consider the DGSMs~\cite{SobolKucherenko09},
\[
    \nu_i(f) := \E{\left(\frac{\partial f}{\partial\xi_i}\right)^2} =
                  \int_\Omega 
                  \left(\frac{\partial f}{\partial\xi_i}\right)^2
                  \pi(\vec{\xi})d\vec{\xi}, \quad i = 1, \ldots, \Np.   
\]
Here $\pi$ is the joint PDF of $\xi$. 
Let $\mat{C} = \E{\nabla f \nabla f^T}$ and note that
\[
   \nu_i(f) = \vec{e}_i^T \mat{C} \vec{e}_i, 
\]
where $\vec{e}_i$ is the $i$th coordinate vector in $\R^\Np$.
This also shows that $\sum_{j=1}^\Np \nu_i(f) = \trace(\mat{C})$.
We consider the spectral decomposition of $\mat{C}$, given by  
$\mat{C} = \sum_{k=1}^\Np \lambda_k \vec{u}_k \vec{u}_k^T$, where 
$\lambda_i$ are the (non-negative) eigenvalues of $\mat{C}$ and
$\vec{u}_k$ are the corresponding (orthonormal) eigenvectors.
We note that,
\begin{equation}\label{equ:spectral_DGSM}
\nu_i(f) = \vec{e}_i^T \Big(\sum_{k=1}^\Np \lambda_k \vec{u}_k \vec{u}_k^T\Big) \vec{e}_i
 = \sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2, 
\end{equation}
which gives a spectral representation for the DGSMs. Truncating the summation 
leads to the notion of activity scores introduced in~\cite{Diaz16,ConstantineDiaz17}
defined as follows: 
\[
   \alpha_i(f; r) =  \sum_{k=1}^r \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2,
   \quad j = 1, \ldots, \Np, \quad r \leq \Np.
\]
The activity scores connect ideas from active subspaces and global sensitivity
analysis, and can be used to approximate DGSMs.  The following result, which
can be found in~\cite{Diaz16,ConstantineDiaz17}, quantifies the error in this
approximation. We provide a short proof for completeness. 
\begin{proposition}\label{prp:dgsm_bound} 
For $1 \leq r \leq \Np$,
\[
0 \leq \nu_i(f) - \alpha_i(f; r) \leq \lambda_{r+1}, \quad i = 1, \ldots, \Np.
\] 
\end{proposition}
\begin{proof}

%Using the spectral representation of the DGSMs and the definition of activity
%scores we clearly see:
%\[
%\alpha_i(f;r) = \sum_{k=1}^{r}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2 \leq \sum_{k=1}^{N_p}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2 = \nu_i(f), \quad \quad i = 1,\ldots,N_p, \quad r \leq N_p
%\]
%In other words
%\[
%0 \leq  \sum_{k=1}^{N_p}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2 - \sum_{k=1}^{r}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2,  \quad \quad i = 1,\ldots,N_p
%\]
%with equality if $N_p=r$.
%\newline
%We can write:
%\[
%\begin{aligned}
%\nu_i(f) = \sum_{k=1}^{N_p}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2 = \sum_{k=1}^{r}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2 + \sum_{k=r+1}^{N_p}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2  \\
%= \alpha_i(f;r) + \sum_{k=r+1}^{N_p}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2 \leq \alpha_i(f;r) + \lambda_{r+1} \sum_{k=r+1}^{N_p} \langle \vec{e}_i, \vec{u}_k \rangle^2,  \quad \quad i = 1,\ldots,N_p
%\end{aligned}
%\]
%The eigenvectors $\vec{u}_k$ are orthonormal so they all have length 1. Also note that for every $x \in \R^n$ we have \[ \norm{x}^2 = \sum_{k=1}^{n} \langle \vec{x}, \vec{u}_k \rangle^2\] This is known as Parseval's identity.
%In particular in this case $\vec{x} = \vec{e}_i \in \R^{N_p}$ so
%\[1 = \norm{\vec{e}_i}^2 = \sum_{k=r+1}^{N_p} \langle \vec{e}_i, \vec{u}_k \rangle^2\]
%Finally we write:
%\[
%\nu_i(f) \leq \alpha_i(f;r) + \lambda_{r+1}, \quad \quad i = 1,\ldots,N_p
%\]

Note that, $\nu_i(f) - \alpha(f; r)  = \sum_{k=r+1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2 \geq 0$,
which gives the first inequality. To see the upper bound, we note,
\[
   \sum_{k=r+1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2 \leq \lambda_{r+1} \sum_{k=r+1}^\Np \ip{\vec{e}_i}{\vec{u}_k}^2
   \leq \lambda_{r+1}. 
\]
The last inequality holds because 
$1 = \norm{\vec{e}_i}^2 = 
\sum_{k = 1}^\Np \ip{\vec{e}_i}{\vec{u}_k}^2 
\geq \sum_{k=r+1}^\Np \ip{\vec{e}_i}{\vec{u}_k}^2$.
\end{proof} 
The utility of this result is realized in problems with 
high-dimensional parameter spaces in which 
the eigenvalues $\lambda_i$ decay rapidly to zero; in 
such cases, this result implies that  $\nu_i(f) \approx \alpha_i(f; r)$,
where $r$ is the \emph{numerical rank} of $\mat{C}$.  This will be especially
effective if there is a large gap in the eigenvalues.

Consider a function $y = f(\xi_1, \xi_2, \ldots, \xi_\Np)$ where
$\xi_i$ are independent are either uniformly distributed or
distributed according to a Boltzman probability distribution.  
It was shown in~\cite{LamboniIoossPopelinEtAl13} that the total Sobol' 
index $T_i(f)$ can be bounded in terms of $\nu_i(f)$:
\begin{equation}\label{equ:sobol_bound}
T_i(f) \leq \frac{C_i}{\V(f)}\nu_i(f), \quad i = 1, \ldots, \Np,
\end{equation}
where for each $i$, $C_i$ is an appropriate \emph{Poincar\'{e}} constant.
This provides a strong theoretical basis for using DGSMs to identify 
unimportant inputs. Moreover, 
in~\cite{VohraAlexanderianSaftaEtAl18} we introduced the normalized screening metric
\[
   \tilde{\nu}_i(f) = \frac{C_i \nu_i(f)}{\sum_{i=1}^\Np C_i \nu_i(f)},
\]
for detecting unimportant input parameters. 
Note that if $\xi_i$ are iid, the $C_i$'s will cancel and the normalized screening
metric reduces to 
\[
    \tilde{\nu}_i(f) = \frac{\nu_i(f)}{\sum_{i=1}^\Np \nu_i(f)} 
      = \frac{\vec{e}_i^T \mat{C} \vec{e}_i}{\trace(\mat{C})} 
      = \frac{\sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2}{\sum_{k = 1}^\Np \lambda_k}.
\]
The following result is a straightforward generalization of
the result from~\cite{Diaz16}, and 
links activity scores and Sobol' indices.
\begin{proposition}
For $1 \leq r \leq \Np$, and $i = 1, \ldots, \Np$, we have
\[
       T_i(f) \leq \frac{C_i}{\V(f)}(\alpha_i(f; r) + \lambda_{r+1}).
\]
\end{proposition}
\begin{proof}
The result follows immediately from Proposition~\ref{prp:dgsm_bound} and~\eqref{equ:sobol_bound}.
\end{proof}
We can also bound the normalized DGSMs using activity scores; it is easy to see  
\[
\tilde{\nu}_i(f) \leq 
\frac{ C_i \big(\alpha_i(f; r) + \lambda_{r+1}\big)}{\sum_{i=1}^\Np C_i \alpha_i(f; r)}
=\frac{C_i \alpha_i(f; r)}{\sum_{i=1}^\Np C_i \alpha_i(f; r)} + \kappa_i \lambda_{r+1}, 
\]
whith $\kappa_i = C_i / (\sum_i C_i \alpha_i(f; r))$. 
In the case where where $\lambda_{r+1} \approx 0$, 
this motivates definition of
normalized activity scores
\[
   \tilde{\alpha}_i(f; r) =  \frac{C_i \alpha_i(f; r)}{\sum_{i=1}^\Np C_i \alpha_i(f; r)}.
\]

### 2018-07-18 Paper/intro.tex
..

\section{Introduction}
\label{sec:intro}

%Constributions:
%\begin{enumerate}
%\item Iterative computation of active subspaces for efficient surrogate modeling 
%\item Comparison of gradient-free and gradient based approaches.
%\item Perspectives on activity scores as means of approximating DGSMs.
%\item Application to challenging chemical kinetics problem. 
%\end{enumerate}
%Insert introduction here.

### 2018-07-18 Paper/cover.tex
..

\begin{center}
\textsc{Efficient UQ for chemical systems using active subspaces}

\bigskip 
\bigskip

Manav Vohra$^{1}$, Alen Alexanderian$^{2}$, Hayley Guy$^{2}$, Sankaran Mahadevan$^{1}$

\bigskip
\bigskip

\normalsize
$^1$Department of Civil and Environmental Engineering\\
Vanderbilt University\\
Nashville, TN 37235\\

\bigskip

$^2$Department of Mathematics\\
North Carolina State University\\
Raleigh, NC 27695\\

\bigskip

\end{center}

\vspace{6cm}

\begin{tabbing}
Corresponding Author: \hspace{5mm} \= Sankaran Mahadevan\\
       \>  Department of Civil and Environmental Engineering\\
       \>  Vanderbilt University\\
       \>  272 Jacobs Hall, VU Mailbox: PMB 351831 \\
       \>  Nashville, TN 37235 \\
       \> \\
Phone: \> (615) 322-3040 \\
Fax:   \> (615) 343-3773 \\
Email: \>  sankaran.mahadevan@vanderbilt.edu   \\
\\
Submitted to: \> \textit{The Journal of Chemical Physics} \\
\>  August 2018\\

\bigskip
\end{tabbing}

\clearpage

### 2018-07-18 Paper/method.tex
..

\section{Methodology}
\label{sec:method}

%% Grad-based algorithm

\bigskip
\begin{breakablealgorithm}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
  \caption{An iterative gradient-based approach for discovering the active subspace}
  \begin{algorithmic}[1]
\Require $\theta_l$, $\theta_u$, $\tau$. 
\Ensure $\Lambda$, $W$, $\alpha$. 
    \Procedure{Gradient-based}{}
	\State Draw $n_1$ random samples, $\{\bm{\xi}_k\}_{k=1}^{n_1}$ $\in$ [-1,1]
         according to $\bm{f(\xi)}$.
	\State Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_1}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_1}$
	\State $N_\text{total}$ = $n_1$ 
	\State Compute $\bm{g}^k = \nabla_{\bm{\theta}}G(\bm\theta_k)$, 
             $k=1, \ldots, N_\text{total}$.  
	\Statex\hspace{5mm} Using Finite Difference:
	\Statex\hspace{5mm} i. Assign a small increment, $d\xi$.
	\Statex\hspace{5mm} ii. Augment the set of samples with neighboring points.
	\be \{\bm{\Xi}_k\}_{k=1}^{n_1(d+1)}:~\{\bm{\xi}_k\}_{k=1}^{n_1} \cup
        \{\xi_{k,j}+d\xi\}_{j=1}^{d} \nonumber
	\ee
	\Statex\hspace{5mm} iii. Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_1(d+1)}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\Xi}_k\}_{k=1}^{n_1(d+1)}$
	\Statex\hspace{5mm} iv. Using the augmented set, $\{\bm{\theta}_k\}_{k=1}^{n_1(d+1)}$
        to compute $\bm{g}^k$. 
	\State Compute the matrix, $\mathcal{C}$ = 
        $\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^k][\bm{g}^k]^\top$
	\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$
	\State Partition the Eigenpairs: $\Lambda^{(0)}~=~ 
        \begin{bmatrix} \Lambda_1^{(0)} & \\ & \Lambda_2^{(0)} \end{bmatrix}$, 
        $W^{(0)}~=~\begin{bmatrix} w_1^{(0)} & w_2^{(0)} \end{bmatrix}$, 
        $\Lambda_1^{(0)}\in \mathbb{R}^{d\times\mathcal{S}}$
	\State Set $r$ = 0
	\Loop
		\State Set $r$ = $r$ + 1
		\State Draw $n_r$ = $\lceil \beta n_1 \rceil$ new random samples 
                $\{\bm{\xi}_k\}$ $\in$ [-1,1], $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$.
		\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.
		\State $N_\text{total}$ = $N_\text{total}$ + $n_r$ 
		\State Compute $\bm{g}^k = \nabla_{\bm{\theta}}G(\bm\theta_k)$, 
             	$k=n_{r-1}+1, \ldots, n_{r-1}+n_r$.  
		\State Compute $\mathcal{C}$ = 
        	$\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^k][\bm{g}^k]^\top$
		\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
		\State Compute $\delta w_{1,j}^{(r)}$ = 
                       \scalebox{1.25}{$\frac{\|w_{1,j}^{r} - 
                       w_{1,j}^{r-1}\|}{\|w_{1,j}^{r-1}\|}$}, 
                       $j = 1,\ldots,\mathcal{S}$.
		\If {$\max\left(\delta w_{1,j}^{(r)}\right)<\tau$}
			\State break
		\EndIf
	\EndLoop
	\State Compute $\alpha_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \Lambda_{1,j}w_{1,j}^2$,
	$i=1,\ldots,d$.

    \EndProcedure
  \end{algorithmic}
  \label{alg:screen}
\end{breakablealgorithm}
\bigskip

### 2018-07-18 Paper/old.tex
..

--- 2018-07-18 Paper/assa.tex
@@
+\documentclass[11pt]{article}
+
+\usepackage{lineno,hyperref}
+\usepackage{latexsym,amsthm,amsmath,xcolor,graphicx, amssymb, color,hyperref,booktabs,mathrsfs}
+
+%\modulolinenumbers[5]
+
+\usepackage{geometry}
+ \geometry{
+ letterpaper,
+ total={8.5in,11in},
+ left=.75in,
+ right=.75in,
+ top=.75in,
+ bottom=.75in,
+ }
+
+% Macros
+\renewcommand{\vec}[1]{{\mathchoice
+                     {\mbox{\boldmath$\displaystyle{#1}$}}
+                     {\mbox{\boldmath$\textstyle{#1}$}}
+                     {\mbox{\boldmath$\scriptstyle{#1}$}}
+                     {\mbox{\boldmath$\scriptscriptstyle{#1}$}}}}
+\newcommand{\norm}[1]{\left\| {#1} \right\|}
+\newcommand{\D}{\mathcal{D}}
+\newcommand{\R}{\mathbb{R}}
+\newcommand{\C}{\mathbb{C}}
+\newcommand{\N}{\mathbb{N}}
+\newcommand{\grad}{\nabla}
+\newcommand{\hessian}{\nabla^2}
+\newcommand{\V}{\mathbb{V}}
+\renewcommand{\S}{\mathbb{S}}
+\newcommand{\ST}{\mathbb{S}^{\text{tot}}}
+\newcommand{\ip}[2]{\langle{#1}, {#2}\rangle}
+\newcommand{\tp}{{\!\mathsf{T}}}
+\newcommand{\trace}{\mathrm{trace}}
+\newcommand{\mat}[1]{\mathbf{{#1}}}
+\newcommand{\Np}{{N_\text{p}}}
+\newcommand{\E}[1]{\mathrm{E}\left\{{#1} \right\}}
+
+\newcommand{\alennote}[1]{{\color{blue} Alen: {#1}}}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}{Lemma}[section]
+\newtheorem{proposition}{Proposition}[section]
+
+%%%%%%%%%%%%%%%%%%%%%%%
+%% Elsevier bibliography styles
+%%%%%%%%%%%%%%%%%%%%%%%
+%% To change the style, put a % in front of the second line of the current style and
+%% remove the % from the second line of the style you would like to use.
+%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Numbered
+%\bibliographystyle{model1-num-names}
+
+%% Numbered without titles
+%\bibliographystyle{model1a-num-names}
+
+%% Harvard
+%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
+
+%% Vancouver numbered
+%\usepackage{numcompress}\bibliographystyle{model3-num-names}
+
+%% Vancouver name/year
+%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
+
+%% APA style
+%\bibliographystyle{model5-names}\biboptions{authoryear}
+
+%% AMA style
+%\usepackage{numcompress}\bibliographystyle{model6-num-names}
+
+%% `Elsevier LaTeX' style
+\bibliographystyle{elsarticle-num}
+%%%%%%%%%%%%%%%%%%%%%%%
+\title{Efficient UQ for chemical systems using active 
+\author{Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan}
+subspaces}
+
+\begin{document}
+\maketitle
+
+\begin{abstract}
+Insert abstract here.
+\end{abstract}
+
+\tableofcontents
+\section{Introduction}
+
+Constributions:
+\begin{enumerate}
+\item Iterative computation of active subspaces for efficient surrogate modeling 
+\item Comparison of gradient-free and gradient based approaches.
+\item Perspectives on activity scores as means of approximating DGSMs.
+\item Application to challenging chemical kinetics problem. 
+\end{enumerate}
+Insert introduction here.
+
 \section{Active subspaces and sensitivity analysis}
-\label{sec:assa}
 
@@
 \]
+
+\section{Method}
+
+\section{Application to $H_2$/$O_2$ reaction mechanism: high-dimensional case}
+
+\alennote{Manav, can we consider a higher dimensional problem with uncertain
+activation energies as well?}
+
+\section{Conclusion}
+
+\section*{References}
+
+\bibliography{mybibfile}
+
+\begin{figure}[htbp]
+ \begin{center}
+  \includegraphics[width=0.40\textwidth]{./Figures/comp_eig}
+\hspace{1mm}
+  \includegraphics[width=0.45\textwidth]{./Figures/comp_eigv}
+  \\ \vspace{5mm}
+  \includegraphics[width=0.40\textwidth]{./Figures/comp_ssp}
+\hspace{1mm}
+  \includegraphics[width=0.45\textwidth]{./Figures/comp_as}
+  \\ \vspace{5mm}
+  \includegraphics[width=0.42\textwidth]{./Figures/ub_conv_kinetics_rich}
+\end{center}
+\end{figure}
+
+\end{document}

### 2018-07-19 Paper/method.tex
..

@@
 \Require $\theta_l$, $\theta_u$, $\tau$. 
-\Ensure $\Lambda$, $W$, $\alpha$. 
+\Ensure $\Lambda$, $W$, $\eta$. 
     \Procedure{Gradient-based}{}
@@
 	\Statex\hspace{5mm} ii. Augment the set of samples with neighboring points.
-	\be \{\bm{\Xi}_k\}_{k=1}^{n_1(d+1)}:~\{\bm{\xi}_k\}_{k=1}^{n_1} \cup
-        \{\xi_{k,j}+d\xi\}_{j=1}^{d} \nonumber
+	\be \{\bm{\Xi}_k\}_{k=1}^{n_1(N_p+1)}:~\{\bm{\xi}_k\}_{k=1}^{n_1} \cup
+        \{\xi_{k,j}+d\xi\}_{j=1}^{N_p} \nonumber
 	\ee
 	\Statex\hspace{5mm} iii. Project to the physical space:
-        $\{\bm{\theta}_k\}_{k=1}^{n_1(d+1)}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\Xi}_k\}_{k=1}^{n_1(d+1)}$
-	\Statex\hspace{5mm} iv. Using the augmented set, $\{\bm{\theta}_k\}_{k=1}^{n_1(d+1)}$
+        $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\Xi}_k\}_{k=1}^{n_1(N_p+1)}$
+	\Statex\hspace{5mm} iv. Using the augmented set, $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}$
         to compute $\bm{g}^k$. 
@@
         $W^{(0)}~=~\begin{bmatrix} w_1^{(0)} & w_2^{(0)} \end{bmatrix}$, 
-        $\Lambda_1^{(0)}\in \mathbb{R}^{d\times\mathcal{S}}$
+        $\Lambda_1^{(0)}\in \mathbb{R}^{N_p\times\mathcal{S}}$
 	\State Set $r$ = 0
@@
 	\EndLoop
-	\State Compute $\alpha_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \Lambda_{1,j}w_{1,j}^2$,
+	\State Compute $\eta_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \Lambda_{1,j}w_{1,j}^2$,
 	$i=1,\ldots,d$.
@@
   \end{algorithmic}
-  \label{alg:screen}
+  \label{alg:grad}
 \end{breakablealgorithm}
 \bigskip
+
+\bigskip
+\begin{breakablealgorithm}
+\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+\renewcommand{\algorithmicensure}{\textbf{Output:}}
+  \caption{An iterative gradient-free approach for discovering the active subspace}
+  \begin{algorithmic}[1]
+\Require $\theta_l$, $\theta_u$, $\tau$. 
+\Ensure $\Lambda$, $W$, $\eta$. 
+    \Procedure{Gradient-free}{}
+	\State Assign $\alpha$ = 3, $q$ = $N_p$+1
+	\State Set $M$ = $\lfloor\alpha q\log(N_p)\rfloor$
+	\State Draw $n_1$ random samples, $\{\bm{\xi}_k\}_{k=1}^{n_1}$ $\in$ [-1,1]
+         according to $\bm{f(\xi)}$.
+	\State Project to the physical space:
+        $\{\bm{\theta}_k\}_{k=1}^{n_1}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_1}$
+	\State $N_\text{total}$ = $n_1$ 
+	\State Compute $G(\bm\theta_k)$, $k=1, \ldots, N_\text{total}$.
+	\State Draw $M$ random samples, $\{\bm{\nu}_k\}_{k=1}^{M}$ $\in$ [-1,1]
+         according to $\bm{f(\xi)}$.
+	\State Construct the matrix, $\mathcal{C}$ using Algorithm~\ref{alg:C}.
+	\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$
+	\State Partition the Eigenpairs: $\Lambda^{(0)}~=~ 
+        \begin{bmatrix} \Lambda_1^{(0)} & \\ & \Lambda_2^{(0)} \end{bmatrix}$, 
+        $W^{(0)}~=~\begin{bmatrix} w_1^{(0)} & w_2^{(0)} \end{bmatrix}$, 
+        $\Lambda_1^{(0)}\in \mathbb{R}^{d\times\mathcal{S}}$
+	\State Set $r$ = 0
+	\Loop
+		\State Set $r$ = $r$ + 1
+		\State Draw $n_r$ = $\lceil \beta n_1 \rceil$ new random samples 
+                $\{\bm{\xi}_k\}$ $\in$ [-1,1], $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$.
+		\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.
+		\State $N_\text{total}$ = $N_\text{total}$ + $n_r$ 
+		\State Compute $G(\bm\theta_k)$, $k=n_{r-1}+1, \ldots, n_{r-1}+n_r$.  
+		\State Construct the matrix, $\mathcal{C}$ using Algorithm~\ref{alg:C}.
+		\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
+		\State Compute $\delta w_{1,j}^{(r)}$ = 
+                       \scalebox{1.25}{$\frac{\|w_{1,j}^{r} - 
+                       w_{1,j}^{r-1}\|}{\|w_{1,j}^{r-1}\|}$}, 
+                       $j = 1,\ldots,\mathcal{S}$.
+		\If {$\max\left(\delta w_{1,j}^{(r)}\right)<\tau$}
+			\State break
+		\EndIf
+	\EndLoop
+	\State Compute $\eta_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \Lambda_{1,j}w_{1,j}^2$,
+	$i=1,\ldots,d$.
+    \EndProcedure
+  \end{algorithmic}
+  \label{alg:free}
+\end{breakablealgorithm}
+\bigskip
+
+\bigskip
+\begin{breakablealgorithm}
+\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+\renewcommand{\algorithmicensure}{\textbf{Output:}}
+  \caption{Algorithm for constructing the matrix, $\mathcal{C}$}
+  \begin{algorithmic}[1]
+\Require $N_\text{total}$, $M$, $\{G(\bm\theta_k)\}_{k=1}^{N_\text{total}}$, 
+$\{\bm{\xi}_k\}_{k=1}^{N_\text{total}}$,
+         $\{\bm{\nu}_k\}_{k=1}^{M}$. 
+\Ensure $\mathcal{C}$. 
+    \Procedure{Matrix $\mathcal{C}$}{}
+	\State Set $p$ = $N_\text{total}$ - 1
+	\State Initialize $\mathcal{D}$ = $[0,\ldots,0]^\top$, $\mathcal{D}\in
+                            \mathbb{R}^{N_\text{total}\times 1}$
+	\For{$i$ = 1 to $M$}
+	\For {$j$ = 1 to $N_\text{total}$}
+	\State $\mathcal{D}(j)$ = 0
+	\For {$k$ = 1 to $d$}
+	\Statex\hspace{20mm} $\mathcal{D}(j) = \mathcal{D}(j) + \|\nu(i,k) - \xi(j,k)\|$
+	\EndFor
+	\EndFor
+	\State $[z,index(i,:)]$ = sort($\mathcal{D}$)
+	\For {$j$ = 1 to $p$}
+	\State $ip = (i-1)p + j$
+	\State $points(ip,:) = \xi(index(i,j),:)$
+	\EndFor
+	\EndFor
+	\For{$np$ = 1 to $M$}
+	\State $A$ = [1 $points((np-1)p+1,:)$]
+	\For{$i$ = $(np-1)p+2$ to ($np\ast p$)}
+	\State $A=[A; 1~points(i,:)]$
+	\EndFor
+	\State $B = G(index(np,1))$
+	\For{$i$ = 2 to $p$}
+	\State $B = [B;~G(index(np,i))]$
+	\EndFor
+	\State $z = A\setminus B$
+	\If {$np$ == 1}
+	\State $\mathcal{B} = z(2:N_p+1)$	
+	\Else 
+	\State $\mathcal{B} = [\mathcal{B}~z(2:N_p+1)]$
+	\EndIf
+	\EndFor
+	\State $\mathcal{C}$ = 0
+	\For {$i$ = 1 to $M$}
+	\State $z$ = $\mathcal{B}(:,i)$
+	\State $\mathcal{C}=\mathcal{C}+zz^\top$
+	\EndFor
+	\State $\mathcal{C}=\mathcal{C}/M$
+    \EndProcedure
+  \end{algorithmic}
+  \label{alg:C}
+\end{breakablealgorithm}
+\bigskip

### 2018-07-20 Paper/intro.tex
..

@@
 %Insert introduction here.
+
+% Outline
+1. Significance of the H2/O2 reaction. Include the mechanism and a discussion on why
+   is it important to study the kinetics? 
+2. Literature survey of studies aimed at quantifying the uncertainty associated with
+reaction kinetics (including my own work) and specifically for the H2/O2 reaction.
+3. Focus of this work and associated challenges (include a table).
+4. Key contributions.
+5. Organization

### 2018-07-21 Paper/intro.tex
..

--- 2018-07-18 Paper/intro.tex
@@
 %Insert introduction here.
+
+% Outline
+%1. Significance of the H2/O2 reaction. Include the mechanism and a discussion on why
+%   is it important to study the kinetics? 
+%2. Literature survey of studies aimed at quantifying the uncertainty associated with
+%reaction kinetics (including my own work) and specifically for the H2/O2 reaction.
+%3. Focus of this work and associated challenges (include a table).
+%4. Key contributions.
+%5. Organization
+
+Time evolution of a chemically reacting system is largely dependent upon rate constants
+associated with individual reactions. The rate constants are typically assumed to exhibit
+a certain correlation with temperature (e.g. Arrhenius-type). Hence, accurate specification
+of the rate-controlling parameters is critical to the fidelity of simulations. However,
+in practical applications, the parameters are either specified using expert knowledge or
+estimated based on a regression-fit to a set of sparse and noisy data. Consequently, the
+underlying uncertainty associated with the parameters and therefore the simulation-prediction
+is typically ignored. Vohra et al. calibrated the Arrhenius diffusion parameters to quantify
+the uncertainty in mixing rates in Zr/Al reactive multilayers using a Bayesian framework
+in~\cite{Vohra:2014} and~\cite{Vohra:2017}. The framework accounted for prior knowledge,
+measurement noise, and the discrepancy between model and experiments. Bayesian calibration
+was accelerated in these efforts using a polynomial chaos surrogate for the model. However,
+the number of uncertain parameters for calibration were limited to two in both cases. In
+higher dimensions, the construction of surrogate as well as the calibration process become
+remarkably challenging.

### 2018-07-21 Paper/assa.tex
..

@@
 \subsection{Active subspaces}
-We recall some background regarding active subspaces~\cite{Constantine15}.
+\label{sub:ac}
+We recall some background regarding active subspaces~\cite{Constantine:2015}.
+
 \subsection{Global sensitivity measures}
+\label{sub:gsa}
 Define the total Sobol' indices $T_i(f)$ and discuss their use.
 
-For $f: \Omega \to \R$, we consider the DGSMs~\cite{SobolKucherenko09},
+For $f: \Omega \to \R$, we consider the DGSMs~\cite{Sobol:2009},
 \[
@@
 which gives a spectral representation for the DGSMs. Truncating the summation 
-leads to the notion of activity scores introduced in~\cite{Diaz16,ConstantineDiaz17}
+leads to the notion of activity scores introduced in~\cite{Diaz:2016,Constantine:2017}
 defined as follows: 
@@
 analysis, and can be used to approximate DGSMs.  The following result, which
-can be found in~\cite{Diaz16,ConstantineDiaz17}, quantifies the error in this
+can be found in~\cite{Diaz:2016,Constantine:2017}, quantifies the error in this
 approximation. We provide a short proof for completeness. 
@@
 distributed according to a Boltzman probability distribution.  
-It was shown in~\cite{LamboniIoossPopelinEtAl13} that the total Sobol' 
+It was shown in~\cite{Lamboni:2013} that the total Sobol' 
 index $T_i(f)$ can be bounded in terms of $\nu_i(f)$:
@@
 unimportant inputs. Moreover, 
-in~\cite{VohraAlexanderianSaftaEtAl18} we introduced the normalized screening metric
+in~\cite{Vohra:2018} we introduced the normalized screening metric
 \[
@@
 The following result is a straightforward generalization of
-the result from~\cite{Diaz16}, and 
+the result from~\cite{Diaz:2016}, and 
 links activity scores and Sobol' indices.

### 2018-07-21 Paper/intro.tex
..

@@
 measurement noise, and the discrepancy between model and experiments. Bayesian calibration
-was accelerated in these efforts using a polynomial chaos surrogate for the model. However,
+was accelerated in these efforts using a polynomial chaos~\cite{Ghanem:2003, Xiu:2002} 
+surrogate for the model. However,
 the number of uncertain parameters for calibration were limited to two in both cases. In
 higher dimensions, the construction of surrogate as well as the calibration process become
-remarkably challenging.
+remarkably challenging due to increase in computational effort.
+
+In this paper, we focus our efforts on mitigating the computational challenge posed by uncertainty
+quantification (UQ) for applications involving a large number of chemical reactions. Specifically, we
+accomplish this by implementing an efficient approach that aims to reduce the dimensionality
+of the problem by computing the so-called active subspace~\cite{Constantine:2015}. 
+The active subspace, discussed later in~\ref{sub:ac}, essentially constitutes the dominant
+eigenspace of a matrix derived from the gradient information of the model output with respect
+to an input. The high-dimensional application problem considered in this work is the H$_2$/O$_2$
+reaction mechanism~\cite{Yetter:1991}. This mechanism has been gaining a lot of attention in the
+recent years as a potential source of clean energy for locomotive applications~\cite{Das:1996}.   
+The mechanism involves 19 reactions including chain reactions, dissociation/recombination reactions, and
+formation and consumption of intermediate species:
+
+\begin{table}[htbp]
+\renewcommand*{\arraystretch}{0.9}
+\begin{center}
+\begin{tabular}{llll}
+\toprule
+Reaction \#     & Reaction &&\\
+\bottomrule
+$\mathcal{R}_1$ & H + O$_2$          & $\rightleftharpoons$ & O + OH \\
+$\mathcal{R}_2$ & O + H$_2$          & $\rightleftharpoons$ & H + OH \\
+$\mathcal{R}_3$ & H$_2$ + OH         & $\rightleftharpoons$ & H$_2$O + H \\
+$\mathcal{R}_4$ & OH + OH            & $\rightleftharpoons$ & O + H$_2$O \\
+$\mathcal{R}_5$ & H$_2$ + M          & $\rightleftharpoons$ & H + H + M \\
+$\mathcal{R}_6$ & O + O + M          & $\rightleftharpoons$ & O$_2$ + M \\
+$\mathcal{R}_7$ & O + H + M          & $\rightleftharpoons$ & OH + M \\
+$\mathcal{R}_8$ & H + OH +M          & $\rightleftharpoons$ & H$_2$O + M \\
+$\mathcal{R}_9$ & H + O$_2$ + M      & $\rightleftharpoons$ & HO$_2$ + M \\
+$\mathcal{R}_{10}$ & HO$_2$ + H      & $\rightleftharpoons$ & H$_2$ + O$_2$ \\
+$\mathcal{R}_{11}$ & HO$_2$ + H      & $\rightleftharpoons$ & OH + OH \\
+$\mathcal{R}_{12}$ & HO$_2$ + O      & $\rightleftharpoons$ & O$_2$ + OH \\
+$\mathcal{R}_{13}$ & HO$_2$ + OH     & $\rightleftharpoons$ & H$_2$O + O$_2$ \\
+$\mathcal{R}_{14}$ & HO$_2$ + HO$_2$ & $\rightleftharpoons$ & H$_2$O$_2$ + O$_2$ \\
+$\mathcal{R}_{15}$ & H$_2$O$_2$ + M  & $\rightleftharpoons$ & OH + OH + M \\
+$\mathcal{R}_{16}$ & H$_2$O$_2$ + H  & $\rightleftharpoons$ & H$_2$O + OH \\
+$\mathcal{R}_{17}$ & H$_2$O$_2$ + H  & $\rightleftharpoons$ & HO$_2$ + H$_2$ \\
+$\mathcal{R}_{18}$ & H$_2$O$_2$ + O  & $\rightleftharpoons$ & OH + HO$_2$ \\
+$\mathcal{R}_{19}$ & H$_2$O$_2$ + OH & $\rightleftharpoons$ & HO$_2$ + H$_2$O \\
+\bottomrule
+\end{tabular}
+\end{center}
+\caption{Reaction mechanism for H$_2$/O$_2$ from~\cite{Yetter:1991}}.
+\label{tab:kinetics}
+\end{table}
+
+For each reaction in table~\ref{tab:kinetics}, the reaction rate is assumed to show an Arrhenius
+correlation with temperature:
+%
+\be
+k_i(T) = A_iT^{n_i}\exp(-E_{a,i}/RT), 
+\label{eq:rate}
+\ee
+%
+where $A_i$ is the pre-exponent, $n_i$ is the index of $T$, $E_{a,i}$ is the
+activation energy corresponding to the $i^{th}$ reaction, and $R$ is the
+universal gas constant.  The global reaction associated with the H$_2$/O$_2$ mechanism can
+be considered as follows:
+%
+\be
+2\text{H}_2 + \text{O}_2 \rightarrow 2\text{H}_2\text{O}.
+\label{eq:global}
+\ee 
+%
+The equivalence ration ($\phi$) is given as follows:
+%
+\be
+\phi = \frac{(M_{\text{H}_2}/M_{\text{O}_2})_\text{obs}}{(M_{\text{H}_2}/M_{\text{O}_2})_\text{st}}.
+\label{eq:phi}
+\ee
+%
+where the numerator on the right-hand-side denotes the fuel (H$_2$)-oxidizer (O$_2$) ratio at given
+condition to the same quantity under stoichiometric conditions. In this study, computations were performed
+at fuel-rich conditions, $\phi$ = 2.0. Homogeneous ignition at constant pressure is simulated using the
+TChem software package~\cite{Safta:2011}, and the time required for the rate of temperature increase
+to exceed a given threshold, regarded as `ignition delay' is recorded.
+
+We are interested in investigating the impact of uncertainty in the rate-controlling parameters, 
+pre-exponent~($A_i$) and the activation energy~($E_{a,i}$) on ignition delay. The $A_i$'s are
+considered to be uniformly distributed in the interval: $[0.9A_i^\ast, 1.1A_i^\ast]$.
+The $E_{a,i}$'s for all reactions except $\mathcal{R}_6$ -- $\mathcal{R}_9$ and $\mathcal{R}_{13}$
+are considered to be uncertain and uniformly distributed in the interval: $[0.99E_{a,i}^\ast, 1.01E_{a,i}^\ast]$.
+$A_i^\ast$ and $E_{a,i}^\ast$ are nominal estimates corresponding to the $i^\text{th}$ reaction
+as provided in~\cite{Yetter:1991}. Hence, the total number of uncertain parameters is 33 which makes the
+present problem reasonably high-dimensional involving a significantly large amount of computational
+effort. To reduce the effort, we explore two strategies: one involving the use of gradients (grad-based)
+and the other involving local linear approximation which does require gradient computation (grad-free).
+These strategies are based on algorithms 1.1 and 1.2 provided  in~\cite{Constantine:2015}. The amount of 
+computational effort associated with the grad-free approach is significantly smaller than the grad-based
+approach. In section~\ref{sec:method}, we implement the two strategies to a simpler problem accounting for the
+uncertainty only in the pre-exponents ($A_i$'s) of the rate constants in the H$_2$/O$_2$ reaction mechanism.
+The two strategies are observed to yield consistent results which encourages the use of the grad-free approach
+for discovering the active subspace especially in high-dimensional problems. In section~\ref{sec:app}, we present
+our findings based on the implementation of the grad-free approach to the original 33-dimensional application.
+
+The dominant eigenvectors that constitute the active subspace correspond to the important directions in the
+parameter space. In other words, the variability in the model output is predominantly observed along these
+directions. Moreover, the components of these eigenvectors can be used to estimate the so-called 
+activity scores~\cite{Diaz:2016,Constantine:2017} which provide an insight into the relative importance
+of the uncertain parameters. In 
+section~\ref{sub:gsa}, we demonstrate that the activity scores could be used to approximate the 
+derivative-based global sensitivity measures (DGSMs)~\cite{Sobol:2009}, and establish a relationship between
+the Sobol' total effect index, the DGSMs, and the activity scores.

### 2018-07-25 Paper/intro.tex
..

@@
 the uncertainty in mixing rates in Zr/Al reactive multilayers using a Bayesian framework
-in~\cite{Vohra:2014} and~\cite{Vohra:2017}. The framework accounted for prior knowledge,
-measurement noise, and the discrepancy between model and experiments. Bayesian calibration
-was accelerated in these efforts using a polynomial chaos~\cite{Ghanem:2003, Xiu:2002} 
-surrogate for the model. However,
+in~\cite{Vohra:2014}. The framework accounted for prior knowledge,
+measurement noise, and the discrepancy between model and experiments. Additionally, 
+in~\cite{Vohra:2017}, a design strategy was implemented to determine optimal conditions
+for experiments in order to enhance the information gain in the uncertain diffusion parameters.
+Bayesian calibration was accelerated in these efforts using a polynomial 
+chaos~\cite{Ghanem:2003, Xiu:2002} surrogate for the model. However,
 the number of uncertain parameters for calibration were limited to two in both cases. In
@@
 eigenspace of a matrix derived from the gradient information of the model output with respect
-to an input. The high-dimensional application problem considered in this work is the H$_2$/O$_2$
+to an input. The dominant eigenspace essentially comprises `important' directions in the parameter
+space that predominantly capture the variability in the model output due to uncertainty in the
+inputs.
+
+The high-dimensional application problem considered in this work is the H$_2$/O$_2$
 reaction mechanism~\cite{Yetter:1991}. This mechanism has been gaining a lot of attention in the
@@
 The mechanism involves 19 reactions including chain reactions, dissociation/recombination reactions, and
-formation and consumption of intermediate species:
+formation and consumption of intermediate species. For each reaction in table~\ref{tab:kinetics}, 
+the reaction rate is assumed to show an Arrhenius correlation with temperature:
+%
+\be
+k_i(T) = A_iT^{n_i}\exp(-E_{a,i}/RT), 
+\label{eq:rate}
+\ee
+%
+where $A_i$ is the pre-exponent, $n_i$ is the index of $T$, $E_{a,i}$ is the
+activation energy corresponding to the $i^{th}$ reaction, and $R$ is the
+universal gas constant.
 
@@
 
-For each reaction in table~\ref{tab:kinetics}, the reaction rate is assumed to show an Arrhenius
-correlation with temperature:
-%
-\be
-k_i(T) = A_iT^{n_i}\exp(-E_{a,i}/RT), 
-\label{eq:rate}
-\ee
-%
-where $A_i$ is the pre-exponent, $n_i$ is the index of $T$, $E_{a,i}$ is the
-activation energy corresponding to the $i^{th}$ reaction, and $R$ is the
-universal gas constant.  The global reaction associated with the H$_2$/O$_2$ mechanism can
+The global reaction associated with the H$_2$/O$_2$ mechanism can
 be considered as follows:
@@
 
-The dominant eigenvectors that constitute the active subspace correspond to the important directions in the
-parameter space. In other words, the variability in the model output is predominantly observed along these
-directions. Moreover, the components of these eigenvectors can be used to estimate the so-called 
+Components of the dominant eigenvectors can be used to estimate the so-called 
 activity scores~\cite{Diaz:2016,Constantine:2017} which provide an insight into the relative importance
@@
 derivative-based global sensitivity measures (DGSMs)~\cite{Sobol:2009}, and establish a relationship between
-the Sobol' total effect index, the DGSMs, and the activity scores.
+the Sobol' total effect index, the DGSMs, and the activity scores. Below, we summarize key contributions of
+the paper and its organization.
+
+\paragraph{Contributions}
+Main contributions of this paper are as follows: (1) A perspective on activity scores as means for approximating
+the DGSMs is provided. (2) A general relationship between DGSMs, activity scores, and the total Sobol' index 
+is established. (3) Iterative gradient-based and gradient-free approaches for computing the active subspace are
+ demonstrated. The two approaches are shown to be consistent and yield computational gains compared to the 
+ conventional approach. As mentioned above, the gains are greater in the case of grad-free approach. (4) The
+ grad-free approach is implemented for discovering an active subspace in a 33-dimensional parameter space 
+ for the H$_2$/O$_2$ reaction kinetics application.
+
+ \paragraph{Organization}
+ The paper is organized as follows. In Section~\ref{sub:ac}, a brief theoretical background on
+ the active subspace methodology is provided. In Section~\ref{sub:gsa}, it is shown that the activity
+ scores provide a reasonable approximation to the DGSMs especially in a high-dimensional setting. 
+ Additionally, a relationship between the three global sensitivity measures, namely, the activity scores,
+ DGSMs, and the total Sobol' indices is established. In Section~\ref{sec:method}, a systematic framework
+ for computing the active subspace using the grad-based, and the grad-free approaches is presented. 
+The two approaches are implemented to the H$_2$/O$_2$ reaction kinetics problem whereby only
+the $A_i$'s are considered as uncertain. Our findings illustrate the consistency in the results based on
+the two approaches thereby encouraging the use of the grad-free approach. In Section~\ref{sec:app},
+we further explore the suitability of the grad-free approach by implementing it to a higher-dimensional
+H$_2$/O$_2$ reaction kinetics problem by accounting for the uncertainty in $E_{a,i}$'s as well. 
+Finally, we draw important conclusions based on the various aspects of this work in Section~\ref{sec:conc}.

### 2018-07-25 Paper/assa.tex
..

@@
 \label{sec:assa}
+In this section, we provide key elements of the active subspace methodology in~\ref{sub:ac}, and
+provide theoretical connections between the global sensitivity measures: DGSMs, activity scores,
+and the total Sobol index in~\ref{sub:gsa}.
 
@@
 \label{sub:ac}
-We recall some background regarding active subspaces~\cite{Constantine:2015}.
+%We recall some background regarding active subspaces~\cite{Constantine:2015}.
+As mentioned earlier, the active subspace is low-dimensional linear subspace that constitutes
+important directions in a model's input space~\cite{Constantine:2015}. The effective variability
+in the model output due to uncertain inputs is predominantly captured along these directions.
+Mathematically, these directions are essentially the eigenvectors of a positive semi-definite
+matrix, $\mat{C}$, defined as follows:
+%
+\be
+\mat{C} = \int_\Omega (\nabla_{\bm{\xi}}f)(\nabla_{\bm{\xi}}f)^\top dP_{\bm{\xi}}
+\label{eq:C}
+\ee
+%
+where $\Omega$:~$[-1,1]^{N_p}$ is the domain of $\bm{\xi}$, and $N_p$ is the number of uncertain
+inputs. $dP_{\bm{\xi}}$ = $p(\bm{\xi})d\bm{\xi}$, where $p(\bm{\xi})$ is the joint probability
+distribution of $\bm{\xi}$. Note that the function, $f$ is considered to be continuous and differentiable
+in the considered parameter domain. Hence, it is possible that a given $f(\bm{\xi})$ might not admit an
+active subspace. However, it is of remarkable interest to investigate if one exists since the subspace could
+be exploited to reduce the dimensionality of the problem and hence the associated computational effort.
+$\nabla_{\bm{\xi}}f$ denotes a vector with individual components
+being partial derivatives of $f$ with respect to the $i^\text{th}$ input, $\xi_i$. Since $\mat{C}$ is symmetric and
+positive semi-definite, it admits a real eigenvalue decomposition:
+%
+\be
+\mat{C} = \bm{W}\bm{\Lambda}\bm{W}^\top
+\ee
+%
+where $\bm{\Lambda}$ = diag($\lambda_1,\ldots,\lambda_{N_p}$) with $\lambda_i$'s
+ sorted in descending order, $i:\{1,2,\ldots,N_p\}$. The eigenpairs are partitioned about
+ the $i^{\text{th}}$ eigenvalue such that the ratio, $\left(\frac{\lambda_i}{\lambda_{i+1}}\right)$ $\gg$ 1
+ as follows:
+ %
+ \be
+ \bm{W} = [\bm{W}_1~\bm{W}_2],~~\bm{\Lambda} = \begin{bmatrix}\bm{\Lambda}_1 & \\  &
+  \bm{\Lambda}_2 \\
+\end{bmatrix}
+ \ee
+ %
+ where $\bm{W}_1$ comprises the dominant eigenspace of $\mat{C}$ regarded as the active subspace,
+ and $\bm{\Lambda}_1$ is the corresponding set of eigenvalues. The active subspace is used to 
+ construct the so-called active variables, $\bm{w}_1^\top \bm{\xi}$ and $f(\bm{\xi}(\theta))$ is transformed
+ as $G(\bm{w}_1^\top \bm{\xi})$. In practice, the set of samples, $\bm{\xi}_i$'s are projected to the
+ physical space as $\bm{\theta}_i$'s at which the model output, $f(\bm{\xi}(\theta))$ is computed. 
+ The matrix, $\mat{C}$ is typically approximated using pseudo-random sampling techniques such 
+ as Monte Carlo and Latin hypercube sampling. The integral in~\eqref{eq:C} is replaced with a
+ summation as follows:
+ %
+ \be
+ \mat{C}\approx \hat{\mat{C}} = \frac{1}{N}\sum\limits_{i=1}^{N} 
+ (\nabla_{\bm{\xi}}f(\bm{\xi}_i))(\nabla_{\bm{\xi}}f(\bm{\xi}_i))^\top
+ = \hat{\bm{W}}\hat{\bm{\Lambda}}\hat{\bm{W}}^\top
+ \ee
+ %
+ It has been shown that the accuracy of approximated dominant eigenspace, $\hat{\bm{W}}_1$ is inversely
+ proportional to the difference between the smallest eigenvalue in $\hat{\bm{\Lambda}}_1$ and the
+ largest eigenvalue in $\hat{\bm{\Lambda}}_2$~\cite{Constantine:2014}.
+
+ Components of the eigenvectors
+ in the active subspace could be used for estimating the so-called activity scores as a measure for
+ global sensitivity and also be used for approximating the DGSMs as discussed in the following section.
 
@@
 \label{sub:gsa}
-Define the total Sobol' indices $T_i(f)$ and discuss their use.
+The total Sobol' index ($T_i(f)$) of a model output, $f(\theta_i)$ quantifies the total contribution of the
+input, $\theta_i$ to the variance of the output. Mathematically, this can be expressed as follows:
+%
+\be
+T_i(f) = 1 - 
+\frac{\V[\mathbb{E}(f|\bm{\theta}_{\sim i})]}{\V(f)},
+\label{eq:total}
+\ee
+%
+where $\bm{\theta}_{\sim i}$ is a vector of all but $i^\text{th}$ uncertain input, and $\V$ denotes the variance.
+Determining converged estimates of $T_i(f)$ typically involves tens of thousands of model runs and is therefore
+prohibitive in the case of compute-intensive applications. The DGSMs~\cite{Sobol:2009}
+provide a means for approximating the upper bound on $T_i(f)$ with fewer computations as 
+discussed in~\cite{Vohra:2018}. 
+\smallskip
 
-For $f: \Omega \to \R$, we consider the DGSMs~\cite{Sobol:2009},
+For $f: \Omega \to \R$, we consider the DGSMs,
 \[
@@
 Here $\pi$ is the joint PDF of $\xi$. 
-Let $\mat{C} = \E{\nabla f \nabla f^T}$ and note that
+Let $\mat{C} = \E{\nabla f \nabla f^\top}$ and note that
 \[
-   \nu_i(f) = \vec{e}_i^T \mat{C} \vec{e}_i, 
+   \nu_i(f) = \vec{e}_i^\top \mat{C} \vec{e}_i, 
 \]
@@
 We consider the spectral decomposition of $\mat{C}$, given by  
-$\mat{C} = \sum_{k=1}^\Np \lambda_k \vec{u}_k \vec{u}_k^T$, where 
+$\mat{C} = \sum_{k=1}^\Np \lambda_k \vec{u}_k \vec{u}_k^\top$, where 
 $\lambda_i$ are the (non-negative) eigenvalues of $\mat{C}$ and
@@
 \begin{equation}\label{equ:spectral_DGSM}
-\nu_i(f) = \vec{e}_i^T \Big(\sum_{k=1}^\Np \lambda_k \vec{u}_k \vec{u}_k^T\Big) \vec{e}_i
+\nu_i(f) = \vec{e}_i^T \Big(\sum_{k=1}^\Np \lambda_k \vec{u}_k \vec{u}_k^\top\Big) \vec{e}_i
  = \sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2, 

### 2018-07-25 Paper/intro.tex
..

@@
 computational effort associated with the grad-free approach is significantly smaller than the grad-based
-approach. In section~\ref{sec:method}, we implement the two strategies to a simpler problem accounting for the
-uncertainty only in the pre-exponents ($A_i$'s) of the rate constants in the H$_2$/O$_2$ reaction mechanism.
-The two strategies are observed to yield consistent results which encourages the use of the grad-free approach
-for discovering the active subspace especially in high-dimensional problems. In section~\ref{sec:app}, we present
-our findings based on the implementation of the grad-free approach to the original 33-dimensional application.
+approach as discussed later in section~\ref{sec:method}.
 
@@
 is established. (3) Iterative gradient-based and gradient-free approaches for computing the active subspace are
- demonstrated. The two approaches are shown to be consistent and yield computational gains compared to the 
+ demonstrated. The two approaches are shown to be consistent and yield greater
+computational gains compared to the 
  conventional approach. As mentioned above, the gains are greater in the case of grad-free approach. (4) The

### 2018-07-26 Paper/method.tex
..

@@
 \label{sec:method}
+
+In this section, we present the computational strategy used for computing the active subspace in the case of
+H$_2$/O$_2$ reaction kinetics with uncertain pre-exponent of individual reaction rates provided in
+Table~\ref{tab:kinetics}. Specifically, we employ two strategies: One that involves computation of model
+gradients in order to construct the matrix (grad-based), $\mat{C}$ in~\eqref{eq:chat}, and the other that involves a local 
+linear approximation of the model output to circumvent the computational cost associated with gradient
+computation (grad-free). The grad-based approach is based on Algorithm 1.1, and the grad-free approach is
+based on Algorithm 1.2 in~\cite{Constantine:2015}. However, in this work, we have implemented the
+two strategies in an iterative manner and have demonstrated potential for additional savings in both cases.
+Additionally, for the present application, our findings based on the two strategies are observed to be consistent.
+Since, the grad-free approach is relatively more efficient, this observation encourages its use for high-dimensional
+applications as discussed later in Section~\ref{sec:app}. Note however, that the model output is required to be
+differentiable throughout the input domain in both cases.
+
+\subsection{Grad-based approach}
+\label{sub:grad}
+
+As discussed earlier, in the grad-based approach, the elements of $\mat{C}$ are estimated using model gradients.
+ In situations where the gradient is not available analytically, one could consider using
+finite difference and other relatively more efficient techniques such as automatic differentiation~\cite{Kiparissides:2009} 
+and adjoint-based gradient computation~\cite{Jameson:1988,Gunzburger:2003,Borzi:2011,Alexanderian:2017}.
+Model evaluations at neighbouring points are required if finite difference is used. Hence, for 
+$N$ samples in a $d$-dimensional parameter space, $N(d+1)$ model evaluations are needed.
+
+We begin by evaluating the gradient of the model output, $\nabla_{\bm{\xi}}f$, at an initial set of $n_1$ samples. 
+For this purpose, $\bm{\xi}_i$'s are projected to the physical space as $\bm{\theta}_i$'s. Using the gradient
+evaluations, the matrix, $\mat{C}$ is computed. Eigenvalue decomposition of $\mat{C}$ yields an initial
+estimate of the dominant eigenspace, $\bm{W}_1$ and corresponding set of eigenvalues, $\bm{\Lambda}_1$.
+At each subsequent iteration, model evaluations are generated at a new set of $n_s$ samples. The new set
+of gradient evaluations are augmented with the available set to re-construct $\mat{C}$ followed by its eigenvalue
+decomposition. The relative
+change in the norm of the difference in individual components of the dominant eigenvectors between subsequent 
+iterations is evaluated. The process is terminated and the resulting eigenspace is considered to be converged once the
+relative change, $\delta w_{1,j}^{(r)}$, is smaller a given tolerance, $\tau$. A regression fit to
+ $G(\bm{w}_1^\top\bm{\xi})$ is used as a surrogate to characterize and quantify the uncertainty in the model
+ output. Moreover, the components of the dominant eigenvectors are used to compute the activity scores which
+ provide an insight into the relative importance of the uncertain parameters. The sequence of steps for the grad-based
+ approach are provided in Algorithm~\ref{alg:grad}.
 
@@
 
+In order to motivate the iterative strategy used in this work, we implement Algorithm~\ref{alg:grad} to the
+19-dimensional H$_2$/O$_2$ reaction kinetics problem with uncertain $A_i$'s. For the purpose of
+verification, $\mat{C}$ is initially constructed using a large set of samples ($N$~=~1000) in the input
+domain and the corresponding active subspace is computed. In Figure~\ref{fig:eig_comp}(a), we illustrate
+the comparison of the resulting eigenvalue spectrum ($\lambda_1 - \lambda_{19}$) with the same using a much 
+smaller set of samples, $n$~=~$\{20,40,80,120\}$. Additionally, in Figure~\ref{fig:eig_comp}(b), we compare
+relative L-2 norm of the difference between the first five eigenvectors of $\mat{C}$.
+%
+\begin{figure}[htbp]
+ \begin{center}
+  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp}
+   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1_5}
+\caption{Left: A comparison of the eigenvalue spectrum using $n$ = $\{20,40,80,120\}$ samples with that
+obtained using a much larger sample size, $N$~=~1000. Right: Relative L-2 norm of the difference between
+individual components of the first five eigenvectors of $\mat{C}$, constructed using $N$~=~1000 samples,
+and using a smaller sample size, $n$ ranging from 20 to 240.} 
+\label{fig:eig_comp}
+\end{center}
+\end{figure}
+%
+Interestingly, we observe that the dominant eigenvalues~($\lambda_1 - \lambda_4$) are approximated 
+reasonably well with just 20 samples. As expected, the accuracy of higher-order eigenvalues is observed
+to improve with the sample size. Similarly, the relative error associated with the dominant eigenvectors
+($i$ = 1,2) is found to range from $\mathcal{O}(10^{-2} - 10^{-1})$ and is relatively much smaller than 
+the higher-order eigenvalues. In this case, since $\left(\frac{\lambda_1}{\lambda_2}\right)\gg 1$, we have
+a 1-dimensional active subspace. Hence, the first eigenvector sufficiently captures the uncertainty in
+the  model output. Therefore, a sample size of 20 is adequate for computing the active subspace in this
+case. The iterative strategy therefore offers a significant potential for computational advantage. 
+However, estimating the gradient of the model output imposes a significant computational burden.
+Hence, we explore the applicability of a so-called gradient-free approach as mentioned earlier, in the
+following section.
+
+\subsection{Grad-free approach}
+\label{sub:gradfree}
+
 \bigskip
@@
          according to $\bm{f(\xi)}$.
-	\State Construct the matrix, $\mathcal{C}$ using Algorithm~\ref{alg:C}.
+	\State Construct the matrix, $\mathcal{C}$.
 	\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$
@@
 		\State Compute $G(\bm\theta_k)$, $k=n_{r-1}+1, \ldots, n_{r-1}+n_r$.  
-		\State Construct the matrix, $\mathcal{C}$ using Algorithm~\ref{alg:C}.
+		\State Construct the matrix, $\mathcal{C}$.
 		\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
@@
 
-\bigskip
-\begin{breakablealgorithm}
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
-  \caption{Algorithm for constructing the matrix, $\mathcal{C}$}
-  \begin{algorithmic}[1]
-\Require $N_\text{total}$, $M$, $\{G(\bm\theta_k)\}_{k=1}^{N_\text{total}}$, 
-$\{\bm{\xi}_k\}_{k=1}^{N_\text{total}}$,
-         $\{\bm{\nu}_k\}_{k=1}^{M}$. 
-\Ensure $\mathcal{C}$. 
-    \Procedure{Matrix $\mathcal{C}$}{}
-	\State Set $p$ = $N_\text{total}$ - 1
-	\State Initialize $\mathcal{D}$ = $[0,\ldots,0]^\top$, $\mathcal{D}\in
-                            \mathbb{R}^{N_\text{total}\times 1}$
-	\For{$i$ = 1 to $M$}
-	\For {$j$ = 1 to $N_\text{total}$}
-	\State $\mathcal{D}(j)$ = 0
-	\For {$k$ = 1 to $d$}
-	\Statex\hspace{20mm} $\mathcal{D}(j) = \mathcal{D}(j) + \|\nu(i,k) - \xi(j,k)\|$
-	\EndFor
-	\EndFor
-	\State $[z,index(i,:)]$ = sort($\mathcal{D}$)
-	\For {$j$ = 1 to $p$}
-	\State $ip = (i-1)p + j$
-	\State $points(ip,:) = \xi(index(i,j),:)$
-	\EndFor
-	\EndFor
-	\For{$np$ = 1 to $M$}
-	\State $A$ = [1 $points((np-1)p+1,:)$]
-	\For{$i$ = $(np-1)p+2$ to ($np\ast p$)}
-	\State $A=[A; 1~points(i,:)]$
-	\EndFor
-	\State $B = G(index(np,1))$
-	\For{$i$ = 2 to $p$}
-	\State $B = [B;~G(index(np,i))]$
-	\EndFor
-	\State $z = A\setminus B$
-	\If {$np$ == 1}
-	\State $\mathcal{B} = z(2:N_p+1)$	
-	\Else 
-	\State $\mathcal{B} = [\mathcal{B}~z(2:N_p+1)]$
-	\EndIf
-	\EndFor
-	\State $\mathcal{C}$ = 0
-	\For {$i$ = 1 to $M$}
-	\State $z$ = $\mathcal{B}(:,i)$
-	\State $\mathcal{C}=\mathcal{C}+zz^\top$
-	\EndFor
-	\State $\mathcal{C}=\mathcal{C}/M$
-    \EndProcedure
-  \end{algorithmic}
-  \label{alg:C}
-\end{breakablealgorithm}
-\bigskip
+%\bigskip
+%\begin{breakablealgorithm}
+%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+%\renewcommand{\algorithmicensure}{\textbf{Output:}}
+%  \caption{Algorithm for constructing the matrix, $\mathcal{C}$}
+%  \begin{algorithmic}[1]
+%\Require $N_\text{total}$, $M$, $\{G(\bm\theta_k)\}_{k=1}^{N_\text{total}}$, 
+%$\{\bm{\xi}_k\}_{k=1}^{N_\text{total}}$,
+%         $\{\bm{\nu}_k\}_{k=1}^{M}$. 
+%\Ensure $\mathcal{C}$. 
+%    \Procedure{Matrix $\mathcal{C}$}{}
+%	\State Set $p$ = $N_\text{total}$ - 1
+%	\State Initialize $\mathcal{D}$ = $[0,\ldots,0]^\top$, $\mathcal{D}\in
+%                            \mathbb{R}^{N_\text{total}\times 1}$
+%	\For{$i$ = 1 to $M$}
+%	\For {$j$ = 1 to $N_\text{total}$}
+%	\State $\mathcal{D}(j)$ = 0
+%	\For {$k$ = 1 to $d$}
+%	\Statex\hspace{20mm} $\mathcal{D}(j) = \mathcal{D}(j) + \|\nu(i,k) - \xi(j,k)\|$
+%	\EndFor
+%	\EndFor
+%	\State $[z,index(i,:)]$ = sort($\mathcal{D}$)
+%	\For {$j$ = 1 to $p$}
+%	\State $ip = (i-1)p + j$
+%	\State $points(ip,:) = \xi(index(i,j),:)$
+%	\EndFor
+%	\EndFor
+%	\For{$np$ = 1 to $M$}
+%	\State $A$ = [1 $points((np-1)p+1,:)$]
+%	\For{$i$ = $(np-1)p+2$ to ($np\ast p$)}
+%	\State $A=[A; 1~points(i,:)]$
+%	\EndFor
+%	\State $B = G(index(np,1))$
+%	\For{$i$ = 2 to $p$}
+%	\State $B = [B;~G(index(np,i))]$
+%	\EndFor
+%	\State $z = A\setminus B$
+%	\If {$np$ == 1}
+%	\State $\mathcal{B} = z(2:N_p+1)$	
+%	\Else 
+%	\State $\mathcal{B} = [\mathcal{B}~z(2:N_p+1)]$
+%	\EndIf
+%	\EndFor
+%	\State $\mathcal{C}$ = 0
+%	\For {$i$ = 1 to $M$}
+%	\State $z$ = $\mathcal{B}(:,i)$
+%	\State $\mathcal{C}=\mathcal{C}+zz^\top$
+%	\EndFor
+%	\State $\mathcal{C}=\mathcal{C}/M$
+%    \EndProcedure
+%  \end{algorithmic}
+%  \label{alg:C}
+%\end{breakablealgorithm}
+%\bigskip

### 2018-07-27 Paper/method.tex
..

@@
 Table~\ref{tab:kinetics}. Specifically, we employ two strategies: One that involves computation of model
-gradients in order to construct the matrix (grad-based), $\mat{C}$ in~\eqref{eq:chat}, and the other that involves a local 
+gradients in order to construct the matrix (grad-based), $\hat{\mat{C}}$ in~\eqref{eq:chat}, and the other that involves a local 
 linear approximation of the model output to circumvent the computational cost associated with gradient
@@
 
-As discussed earlier, in the grad-based approach, the elements of $\mat{C}$ are estimated using model gradients.
+As discussed earlier, in the grad-based approach, the elements of $\hat{\mat{C}}$ are estimated using model gradients.
  In situations where the gradient is not available analytically, one could consider using
@@
 For this purpose, $\bm{\xi}_i$'s are projected to the physical space as $\bm{\theta}_i$'s. Using the gradient
-evaluations, the matrix, $\mat{C}$ is computed. Eigenvalue decomposition of $\mat{C}$ yields an initial
-estimate of the dominant eigenspace, $\bm{W}_1$ and corresponding set of eigenvalues, $\bm{\Lambda}_1$.
+evaluations, the matrix, $\hat{\mat{C}}$ is computed. Eigenvalue decomposition of $\hat{\mat{C}}$ yields an initial
+estimate of the dominant eigenspace, $\hat{\bm{W}_1}$ and corresponding set of eigenvalues, $\hat{\bm{\Lambda}_1}$.
 At each subsequent iteration, model evaluations are generated at a new set of $n_s$ samples. The new set
-of gradient evaluations are augmented with the available set to re-construct $\mat{C}$ followed by its eigenvalue
+of gradient evaluations are augmented with the available set to re-construct $\hat{\mat{C}}$ followed by its eigenvalue
 decomposition. The relative
-change in the norm of the difference in individual components of the dominant eigenvectors between subsequent 
-iterations is evaluated. The process is terminated and the resulting eigenspace is considered to be converged once the
-relative change, $\delta w_{1,j}^{(r)}$, is smaller a given tolerance, $\tau$. A regression fit to
+change in the norm of the difference in squared value of individual components of the dominant eigenvectors between subsequent 
+iterations is evaluated. The process is terminated and the resulting eigenspace is considered to be converged once
+highest relative change at iteration $r$, $\delta w_{1,j}^{(r)}$, is smaller a given tolerance, $\tau$. A regression fit to
  $G(\bm{w}_1^\top\bm{\xi})$ is used as a surrogate to characterize and quantify the uncertainty in the model
@@
 		\State Compute $\delta w_{1,j}^{(r)}$ = 
-                       \scalebox{1.25}{$\frac{\|w_{1,j}^{r} - 
-                       w_{1,j}^{r-1}\|}{\|w_{1,j}^{r-1}\|}$}, 
+                       \scalebox{1.25}{$\frac{\|(w_{1,j}^{r})^2 - 
+                       (w_{1,j}^{r-1})^2\|}{\|(w_{1,j}^{r-1})^2\|}$}, 
                        $j = 1,\ldots,\mathcal{S}$.
@@
 19-dimensional H$_2$/O$_2$ reaction kinetics problem with uncertain $A_i$'s. For the purpose of
-verification, $\mat{C}$ is initially constructed using a large set of samples ($N$~=~1000) in the input
-domain and the corresponding active subspace is computed. In Figure~\ref{fig:eig_comp}(a), we illustrate
+verification, $\hat{\mat{C}}$ is initially constructed using a large set of samples ($N$~=~1000) in the input
+domain and the corresponding active subspace is computed. In Figure~\ref{fig:eig_comp} (left), we illustrate
 the comparison of the resulting eigenvalue spectrum ($\lambda_1 - \lambda_{19}$) with the same using a much 
-smaller set of samples, $n$~=~$\{20,40,80,120\}$. Additionally, in Figure~\ref{fig:eig_comp}(b), we compare
-relative L-2 norm of the difference between the first five eigenvectors of $\mat{C}$.
+smaller set of samples, $n$~=~$\{20,40,80,120\}$. Additionally, in Figure~\ref{fig:eig_comp} (right), we compare
+relative L-2 norm of the difference between the first five eigenvectors of $\hat{\mat{C}}$.
 %
@@
 %
-Interestingly, we observe that the dominant eigenvalues~($\lambda_1 - \lambda_4$) are approximated 
+Interestingly, we observe that the dominant eigenvalues ($\lambda_1 - \lambda_4$) are approximated 
 reasonably well with just 20 samples. As expected, the accuracy of higher-order eigenvalues is observed
@@
 
+As mentioned earlier, the grad-free approach does not involve computation of the gradient of the model
+output. Instead, a least-squares fit is performed to the local linear regression-based approximation
+to the model output. While the grad-free approach offers a significant potential for additional 
+computational savings, its ability to reasonably approximate the active subspace 
+has not been investigated for a wide variety of problems~\cite{Constantine:2015}. Nevertheless, we are motivated by the
+underlying potential of this approach in reducing the computational effort associated with the
+active subspace computation. Moreover, this effort could help advance this approach and further motivate
+its usage for a wide variety of high-dimensional applications in the future.
+
+First, we implement the grad-free approach to the 19-dimensional problem and compare our findings with
+those obtained using the grad-based approach. Later, in
+Section~\ref{sec:app}, we compute the active subspace for a 33-dimensional application using the grad-free
+approach. Once again, the strategy is implemented in an iterative manner as shown in~\ref{sub:grad}. 
+It is observed that the results from the grad-free approach are consistent with the grad-based approach
+as discussed further below.
+
+In the grad-free approach, the model output is first computed at an initial set of $n_1$ samples. An independent
+set of $M$ samples is also drawn according to the joint probability distribution of the inputs. For each sample
+in $M$, $p$ nearest neighbours in the $n_1$ set are identified. A least-squares fit to model evaluations at
+the $p$ samples is performed. The slope vector ($\bm{b}_i$) of the regression-fit is stored separately and the process is
+repeated for each sample in $M$. The symmetric and positive semi-definite matrix, $\hat{\mat{C}}$ is hence constructed
+using the slope vectors as follows:
+%
+\be
+\hat{\mat{C}} = \frac{1}{M}\sum\limits_{i=1}^{M}\bm{b}_i\bm{b}_i^\top = \hat{\bm{W}}\hat{\bm{\Lambda}}\hat{\bm{W}}^\top
+\ee
+%
+Eigenvalue decomposition of $\hat{\mat{C}}$ yields an initial approximation to the
+active subspace. The aforementioned sequence of steps is repeated until $\delta w_{1,j}^{(r)}$ as mentioned
+earlier in~\ref{sub:grad} is found to be smaller than $\tau$. An outline of the grad-free approach is
+provided in Algorithm~\ref{alg:free1}, and the specific sequence of steps pertaining to the construction of
+$\hat{\mat{C}}$ is provided in Algorithm~\ref{alg:free2}.
+
 \bigskip
@@
   \begin{algorithmic}[1]
-\Require $\theta_l$, $\theta_u$, $\tau$. 
+\Require $\theta_l$, $\theta_u$, $\alpha$, $M$, $\tau$. 
 \Ensure $\Lambda$, $W$, $\eta$. 
     \Procedure{Gradient-free}{}
-	\State Assign $\alpha$ = 3, $q$ = $N_p$+1
-	\State Set $M$ = $\lfloor\alpha q\log(N_p)\rfloor$
 	\State Draw $n_1$ random samples, $\{\bm{\xi}_k\}_{k=1}^{n_1}$ $\in$ [-1,1]
@@
   \end{algorithmic}
-  \label{alg:free}
+  \label{alg:free1}
 \end{breakablealgorithm}
@@
 
-%\bigskip
-%\begin{breakablealgorithm}
+\bigskip
+\begin{breakablealgorithm}
 %\renewcommand{\algorithmicrequire}{\textbf{Input:}}
 %\renewcommand{\algorithmicensure}{\textbf{Output:}}
-%  \caption{Algorithm for constructing the matrix, $\mathcal{C}$}
-%  \begin{algorithmic}[1]
-%\Require $N_\text{total}$, $M$, $\{G(\bm\theta_k)\}_{k=1}^{N_\text{total}}$, 
-%$\{\bm{\xi}_k\}_{k=1}^{N_\text{total}}$,
-%         $\{\bm{\nu}_k\}_{k=1}^{M}$. 
-%\Ensure $\mathcal{C}$. 
-%    \Procedure{Matrix $\mathcal{C}$}{}
-%	\State Set $p$ = $N_\text{total}$ - 1
-%	\State Initialize $\mathcal{D}$ = $[0,\ldots,0]^\top$, $\mathcal{D}\in
-%                            \mathbb{R}^{N_\text{total}\times 1}$
-%	\For{$i$ = 1 to $M$}
-%	\For {$j$ = 1 to $N_\text{total}$}
-%	\State $\mathcal{D}(j)$ = 0
-%	\For {$k$ = 1 to $d$}
-%	\Statex\hspace{20mm} $\mathcal{D}(j) = \mathcal{D}(j) + \|\nu(i,k) - \xi(j,k)\|$
-%	\EndFor
-%	\EndFor
-%	\State $[z,index(i,:)]$ = sort($\mathcal{D}$)
-%	\For {$j$ = 1 to $p$}
-%	\State $ip = (i-1)p + j$
-%	\State $points(ip,:) = \xi(index(i,j),:)$
-%	\EndFor
-%	\EndFor
-%	\For{$np$ = 1 to $M$}
-%	\State $A$ = [1 $points((np-1)p+1,:)$]
-%	\For{$i$ = $(np-1)p+2$ to ($np\ast p$)}
-%	\State $A=[A; 1~points(i,:)]$
-%	\EndFor
-%	\State $B = G(index(np,1))$
-%	\For{$i$ = 2 to $p$}
-%	\State $B = [B;~G(index(np,i))]$
-%	\EndFor
-%	\State $z = A\setminus B$
-%	\If {$np$ == 1}
-%	\State $\mathcal{B} = z(2:N_p+1)$	
-%	\Else 
-%	\State $\mathcal{B} = [\mathcal{B}~z(2:N_p+1)]$
-%	\EndIf
-%	\EndFor
-%	\State $\mathcal{C}$ = 0
-%	\For {$i$ = 1 to $M$}
-%	\State $z$ = $\mathcal{B}(:,i)$
-%	\State $\mathcal{C}=\mathcal{C}+zz^\top$
-%	\EndFor
-%	\State $\mathcal{C}=\mathcal{C}/M$
-%    \EndProcedure
-%  \end{algorithmic}
-%  \label{alg:C}
-%\end{breakablealgorithm}
-%\bigskip
+  \caption{Algorithm for constructing the matrix, $\mat{C}$}
+  \begin{algorithmic}[1]
+	\Procedure{Local Linear Approximation}{}
+	\EndProcedure
+  \end{algorithmic}
+  \label{alg:free2}
+\end{breakablealgorithm}
+\bigskip
+
+The grad-free approach when implemented to the 19-dimensional H$_2$/O$_2$ reaction kinetics problem
+with uncertain pre-exponents ($A_i$'s) also yielded a 1-dimensional active subspace. In Figure~\ref{fig:comp} (left),
+we provide a comparison of the components of the dominant eigenvector, obtained using the two approaches.
+Additionally, the quantity, $\delta w_{1,j}^{(r)}$ is also plotted in
+Figure~\ref{fig:comp} (right) to illustrate convergence characteristics for the grad-free approach.
+%
+\begin{figure}[htbp]
+ \begin{center}
+  \includegraphics[width=0.8\textwidth]{./Figures/eigv6}
+\caption{Left: An illustrative comparison of individual components of the converged dominant eigenvector obtained
+using the two approaches discussed in~\ref{sub:grad} and~\ref{sub:gradfree}. Right: The quantity,  $\delta w_{1,j}^{(r)}$
+is plotted for successive iterations to illustrate the convergence behavior.}
+\label{fig:comp}
+\end{center}
+\end{figure}
+%
+Individual components of the dominant eigenvector from the two approaches are observed to be in excellent
+agreement with each other. Moreover, as expected,  $\delta w_{1,j}^{(r)}$ is observed to decrease with iterations, and
+the eigenvector is considered to have converged if its value is found to be smaller than 0.1. Note that only 30 samples
+were found to be sufficient for approximating the active subspace in the 19-dimensional input domain.
+
+As mentioned earlier, the model output $f(\bm{\xi}(\bm{\theta}))$ predominantly varies in the active subspace. Hence, 
+$f(\bm{\xi}(\bm{\theta}))$ can be transformed as a function of the active variable as $G(\bm{w}_1^\top\bm{\xi})$ in
+the subspace.

### 2018-07-27 Paper/assa.tex
..

@@
  \be
- \bm{W} = [\bm{W}_1~\bm{W}_2],~~\bm{\Lambda} = \begin{bmatrix}\bm{\Lambda}_1 & \\  &
+ \bm{W} = [\bm{w}_1~\bm{w}_2],~~\bm{\Lambda} = \begin{bmatrix}\bm{\Lambda}_1 & \\  &
   \bm{\Lambda}_2 \\
@@
  %
- where $\bm{W}_1$ comprises the dominant eigenspace of $\mat{C}$ regarded as the active subspace,
+ where $\bm{w}_1$ comprises the dominant eigenspace of $\mat{C}$ regarded as the active subspace,
  and $\bm{\Lambda}_1$ is the corresponding set of eigenvalues. The active subspace is used to 
@@
  = \hat{\bm{W}}\hat{\bm{\Lambda}}\hat{\bm{W}}^\top
+\label{eq:chat}
  \ee
  %
- It has been shown that the accuracy of approximated dominant eigenspace, $\hat{\bm{W}}_1$ is inversely
+ It has been shown that the accuracy of approximated dominant eigenspace, $\hat{\bm{w}}_1$ is inversely
  proportional to the difference between the smallest eigenvalue in $\hat{\bm{\Lambda}}_1$ and the

### 2018-07-27 Paper/method.tex
..

@@
 evaluations, the matrix, $\hat{\mat{C}}$ is computed. Eigenvalue decomposition of $\hat{\mat{C}}$ yields an initial
-estimate of the dominant eigenspace, $\hat{\bm{W}_1}$ and corresponding set of eigenvalues, $\hat{\bm{\Lambda}_1}$.
+estimate of the dominant eigenspace, $\hat{\bm{w}_1}$ and corresponding set of eigenvalues, $\hat{\bm{\Lambda}_1}$.
 At each subsequent iteration, model evaluations are generated at a new set of $n_s$ samples. The new set
@@
 	\EndLoop
-	\State Compute $\eta_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \Lambda_{1,j}w_{1,j}^2$,
+	\State Compute $\alpha_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \Lambda_{1_{i,j}}w_{1_{i,j}}^2$,
 	$i=1,\ldots,d$.
@@
   \begin{algorithmic}[1]
-\Require $\theta_l$, $\theta_u$, $\alpha$, $M$, $\tau$. 
+\Require $\theta_l$, $\theta_u$, $M$, $\tau$. 
 \Ensure $\Lambda$, $W$, $\eta$. 
@@
 $f(\bm{\xi}(\bm{\theta}))$ can be transformed as a function of the active variable as $G(\bm{w}_1^\top\bm{\xi})$ in
-the subspace.
+the subspace. The plot of $G$ versus $\bm{w}_1^\top\bm{\xi}$, regarded as the \textit{sufficient summary plot} (SSP) is 
+compared for the two approaches in Figure~\ref{fig:comp_ssp}.
+%
+\begin{figure}[htbp]
+ \begin{center}
+  \includegraphics[width=0.48\textwidth]{./Figures/comp_ssp}
+\caption{An illustrative comparison of the SSPs generated using the grad-based and the grad-free approaches.}
+\label{fig:comp_ssp}
+\end{center}
+\end{figure}
+%
+The two SSPs are observed to be in excellent agreement with each other. Moreover, it is interesting to note
+that the variability in ignition
+delay based on the considered prior marginals for $A_i$'s is observed to be approximately linear in the
+active subspace; consistent with the fact the it is 1-dimensional.
+
+We extend the comparative assessment of the two approaches by estimating the activity score for individual
+uncertain parameters ($\alpha_i$) using
+the components of the dominant eigenvector in the following equation:
+%
+\be
+\alpha_i(n) = \sum\limits_{j=1}^{\mathcal{S}}\Lambda_{1_{i,j}}w_{1_{i,j}}^2,~i = 1,\ldots,N_p
+\ee
+%
+where $n$ is the sample size. The activity scores for the 19 uncertain pre-exponents ($A_i$'s), estimated using
+the two approaches are plotted in Figure~\ref{fig:comp_as} (left). Additionally, for the purpose of verification, we 
+compare these estimates with those obtained using the DGSMs~\cite{Vohra:2018} in 
+Figure~\ref{fig:comp_as} (right). 
+%
+\begin{figure}[htbp]
+ \begin{center}
+  \includegraphics[width=0.45\textwidth]{./Figures/comp_as}
+  \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}
+\caption{Left: A bar-graph of activity scores ($\alpha_i$'s) for the 19 uncertain pre-exponents ($A_i$'s).
+Right: A plot of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$), defined as a normalized product of the
+Poincare constant and the DGSM ($\mu_i$) (adapted from~\cite{Vohra:2018}).}
+\label{fig:comp_as}
+\end{center}
+\end{figure}
+%
+The activity scores estimated using the two approaches agree favourably with each other as well as those
+based on the screening metric involving the DGSMs from~\cite{Vohra:2018}. It is observed that the uncertainty
+associated with the ignition delay is largely due to the uncertainty in $A_9$ and $A_1$. Sensitivity towards
+$A_{15}$ and $A_{17}$ is also found to be significant.
+
+The above comparisons establish consistency between the two approaches. Note that the computational
+effort in the case of grad-free approach is significantly smaller. Hence, we are encouraged to use the
+grad-free approach for the 33-dimensional application in the next section.

### 2018-07-27 Paper/app.tex
..

@@
-\section{Application to $H_2$/$O_2$ reaction mechanism: high-dimensional case}
+\section{$\text{H}_2$/$\text{O}_2$ reaction kinetics: high-dimensional case}
 \label{sec:app}
+
+For the high-dimensional case, we aim to investigate the impact of the uncertain
+pre-exponents ($A_i$'s) as well as the activation energies ($E_{a,i}$'s) on ignition 
+delay during the H$_2$/O$_2$ reaction. The $A_i$'s are considered to be uniformly
+distributed in the interval, $[0.9A_i^\ast, 1.1A_i^\ast]$. The $E_{a,i}$'s for all
+reactions except $\mathcal{R}_6$ -- $\mathcal{R}_9$ and $\mathcal{R}_{13}$
+are considered to be uncertain and uniformly distributed in the interval: 
+$[0.99E_{a,i}^\ast, 1.01E_{a,i}^\ast]$. The nominal values, $A_i^\ast$ and $E_{a,i}^\ast$
+corresponding to the different reaction rates are provided in~\cite{Yetter:1991}.
+
+The grad-free approach discussed earlier in~\ref{sub:gradfree} was used to compute the
+active subspace. Using the iterative procedure, the convergence of the eigenvectors
+was examined by tracking $\delta w_{1,j}^{(i)}$. In Figure~\ref{fig:conv_app} (left),
+components of the converged eigenvector based on 50 samples using $\tau$ = 0.2 are plotted.
+To illustrate convergence, we compare them with components of the eigenvector obtained
+using 100 samples (twice as many). The variation of maximum $\delta w_{1,j}^{(i)}$ between
+successive iterations during the course of the iterative procedure is illustrated in 
+Figure~\ref{fig:conv_app} (right). 
+%
+\begin{figure}[htbp]
+ \begin{center}
+  \includegraphics[width=0.8\textwidth]{./Figures/eigv10}
+\caption{Left: An illustrative comparison of individual components of the converged dominant eigenvector obtained
+using the two approaches discussed in~\ref{sub:grad} and~\ref{sub:gradfree}. Right: The quantity,  $\delta w_{1,j}^{(r)}$
+is plotted for successive iterations to illustrate the convergence behavior.}
+\label{fig:conv_app}
+\end{center}
+\end{figure}
+%

### 2018-07-27 Paper/intro.tex
..

@@
 pre-exponent~($A_i$) and the activation energy~($E_{a,i}$) on ignition delay. The $A_i$'s are
-considered to be uniformly distributed in the interval: $[0.9A_i^\ast, 1.1A_i^\ast]$.
-The $E_{a,i}$'s for all reactions except $\mathcal{R}_6$ -- $\mathcal{R}_9$ and $\mathcal{R}_{13}$
-are considered to be uncertain and uniformly distributed in the interval: $[0.99E_{a,i}^\ast, 1.01E_{a,i}^\ast]$.
-$A_i^\ast$ and $E_{a,i}^\ast$ are nominal estimates corresponding to the $i^\text{th}$ reaction
-as provided in~\cite{Yetter:1991}. Hence, the total number of uncertain parameters is 33 which makes the
+considered to be uniformly distributed about their nominal estimates provided in~\cite{Yetter:1991}.
+The total number of uncertain inputs is 33 which makes the
 present problem reasonably high-dimensional involving a significantly large amount of computational
@@
 and the other involving local linear approximation which does require gradient computation (grad-free).
-These strategies are based on algorithms 1.1 and 1.2 provided  in~\cite{Constantine:2015}. The amount of 
-computational effort associated with the grad-free approach is significantly smaller than the grad-based
-approach as discussed later in section~\ref{sec:method}.
+These strategies are based on algorithms 1.1 and 1.2 provided  in~\cite{Constantine:2015}. However, it
+must be noted that in this work, an iterative approach is adopted for implementing the two strategies.
+Consequently, the amount of computational effort appears to have decreased, and is found to be
+relatively less in the case of grad-free approach as discussed later in section~\ref{sec:method}.
 

### 2018-07-29 Paper/app.tex
..

@@
 %
+The dominant eigenvector is observed to converge with only 50 samples in the 33-dimensional space which is remarkably efficient.
+
+In Figure~\ref{fig:hd}, we illustrate the resulting eigenvalue spectrum and the sufficient summary 
+plot.
+%
+\begin{figure}[htbp]
+ \begin{center}
+   \includegraphics[width=0.45\textwidth]{./Figures/eig_33D}
+   \includegraphics[width=0.45\textwidth]{./Figures/ssp_33D}
+\caption{Left: Eigenvalue spectrum of $\hat{\mat{C}}$. Right: Sufficient summary plot indicating the
+existence of a potential 1-dimensional active subspace.} 
+\label{fig:hd}
+\end{center}
+\end{figure}
+%
+The second eigenvalue is observed to be roughly four orders of magnitude smaller than the first 
+indicating the existence of a 1-dimensional active subspace. This is further confirmed by the sufficient
+summary plot which shows that the variability in $G$ can be approximated using a 1-dimensional regression
+fit which could be used its a surrogate. We will verify the accuracy of the surrogate later in~\ref{sub:verify}.
+
+Components of the dominant eigenvector in Figure~\ref{fig:conv_app} (left) are utilized for estimating the
+activity scores, defined earlier in~\ref{eq:as}. As discussed earlier, the activity scores can be regarded as a 
+measure for global sensitivity of the uncertain parameters. The sensitivity of ignition delay towards the 
+33 uncertain parameters considered in this problem is plotted below in Figure~\ref{fig:as_33D}.
+%
+\begin{figure}[htbp]
+ \begin{center}
+  \includegraphics[width=0.45\textwidth]{./Figures/as_A_33D}
+    \includegraphics[width=0.45\textwidth]{./Figures/as_E_33D}
+\caption{A bar-graph illustrating individual activity scores for the uncertain $A_i$' (left) and $E_{a,i}$'s (right).}
+\label{fig:as_33D}
+\end{center}
+\end{figure}
+%
+From the above plot, it can be inferred that the ignition delay is predominantly sensitivity towards $A_1$, $A_9$, and
+$E_{a,1}$. Furthermore, it is moderately sensitivity towards $A_{15}$ and $A_{17}$ while sensitivity towards remaining
+inputs is found to be low or negligible. Note that the sensitivity results for the pre-exponents is found to be consistent
+with the 19-dimensional case presented in~\ref{sec:method}. This observation can also be exploited for reducing the 
+dimensionality of the problem from 33 to 3-5 depending upon the required level of accuracy in terms of characterizing 
+the uncertainty in the model output. The sensitivity-driven dimension reduction for uncertainty quantification has been
+discussed in our earlier effort~\cite{Vohra:2018}. In this work, however, we focus on dimension reduction using the active 
+subspace approach which is shown to yield even greater scope for dimension reduction i.e. from 33 to 1. We verify
+the accuracy of the 1-d regression fit to $G(\bm{w}_1^\top\bm{\xi})$ in the following section.
+
+\subsection{Surrogate Verification}
+\label{sub:verify}
+
+The 1-dimensional surrogate ($\mathcal{Y}$) is investigated for its accuracy as well as the ability to capture the 
+uncertainty in the
+model output in two ways. Firstly, we estimate the relative L-2 norm of the discrepancy~($\varepsilon_d$)
+between estimates of ignition delay in the case of H$_2$/O$_2$ reaction, obtained using the 
+model output and the surrogate in the following equation:
+%
+\be
+\varepsilon_d = \frac{\|f(\bm{\theta}) - \mathcal{Y}(\bm{\xi})\|}{\|f(\bm{\theta})\|}
+\ee
+%
+The relative norm, $\varepsilon_d$ was estimated to be 8.92$\times10^{-2}$ based on a validation test data
+comprising 1000 model evaluations. Hence, in the norm-sense, it can be said that the 1-dimensional surrogate is 
+reasonably accurate.
+
+Secondly, we verify the accuracy of the surrogate in a probabilistic setting. In particular, we compare a histogram
+plot of the model evaluations for ignition delay in the validation set with a PDF of the same using the 1-dimensional
+surrogate as shown in Figure~\ref{fig:pdf_33D}.
+%
+\begin{figure}[htbp]
+ \begin{center}
+  \includegraphics[width=0.45\textwidth]{./Figures/pdf_comp_id_1e4}
+\caption{A comparison of the histogram plot based on model evaluations and the PDF based on surrogate
+predictions for ignition delay. The two plots are based on samples in the validation set.}
+\label{fig:pdf_33D}
+\end{center}
+\end{figure}
+%
+In the above comparison, it is observed that the peak of the PDF generated using the 1-dimensional surrogate in the
+active subspace is consistent with the modal estimate based on the histogram. Moreover, the spread in the ignition
+ delay is reasonably captured by the PDF. Hence, the surrogate is also verified in a probabilistic sense.

### 2018-07-29 Paper/assa.tex
..

@@
 $\xi_i$ are independent are either uniformly distributed or
-distributed according to a Boltzman probability distribution.  
+distributed according to a Boltzmann probability distribution.  
 It was shown in~\cite{Lamboni:2013} that the total Sobol' 
@@
 \]
-whith $\kappa_i = C_i / (\sum_i C_i \alpha_i(f; r))$. 
+with $\kappa_i = C_i / (\sum_i C_i \alpha_i(f; r))$. 
 In the case where where $\lambda_{r+1} \approx 0$, 

### 2018-07-29 Paper/conc.tex
..

\section{Summary and Discussion}
\label{sec:conc}

%%Brief summary
%-Mathematical connections between Sobol, DGSMs, and activity scores exist and could
%be exploited to perform GSA with reduced effort.
%-Iterative implementation yields advantage
%-Consistency between the two approaches with DGSM as well as among themselves
%encourages the use of grad-free.
%-Grad-free was found to be suitable for a high-dimensional problem.
%-Based on the results presented, the following parameter were found to be 
%significant.
%
%discussion
%- Methodology is agnostic to the choice of mechanism and could be applied to other
%systems.
%- However, constraints exist in terms of differentiability of the QoI. Not all
%input/output relationships exhibit an active subspace. However, it is important to
%check.
%- Activity scores could be used to identify key reactions and thereby construct
%reduced order mechanisms for reducing computational effort and investigating the
%inadequacy associated with rate laws.

 In this work, we focused on the uncertainty associated with the rate-controlling parameters in
 the H$_2$/O$_2$ reaction mechanism and its impact on ignition delay predictions. The mechanism
 involves 19 different reactions and in each case, the reaction rate depends upon the choice of
 a pre-exponent and an activation energy. Hence, in theory, the evolution of the chemical system
 depends upon 38 inputs which is reasonably high-dimensional. In such case, conventional means
 for uncertainty quantification such as those involving a surrogate model construction as well as
 sensitivity analysis would be computationally challenging and even prohibitive in some cases. 
 Hence, we focused our efforts on reducing the dimensionality of the problem by identifying important
 directions (active subspaces) in the parameter space such that the variability in the model output is 
 predominantly captured along these directions. Additionally, we demonstrate that the activity scores,
 computed using the components of the dominant eigenvectors can be used to approximate the
 DGSMs. Furthermore, we establish mathematical linkages between the different global sensitivity measures:
 activity scores, DGSMs, and total Sobol' index which could be exploited to reduce computational effort 
 associated with global sensitivity analysis.

Two computational strategies were explored to compute the active subspace for the H$_2$/O$_2$
kinetics application in this work. The first strategy referred to as grad-based requires computation of gradient of the 
model output with respect to the input, whereas, the second strategy, referred to as grad-free implements a local
linear approximation to the model output in order to estimate the gradient. The two strategies are shown to yield
consistent results for the 19-dimensional problem whereby only the pre-exponents are considered to be uncertain.
Additionally, the activity scores are also shown to be consistent with the screening metric estimates based on
DGSMs in~\cite{Vohra:2018}. Note that an iterative procedure is adopted for implementing the two strategies
for enhancing computational efficiency. 
Motivated by these findings, the grad-free approach is used for computing the active subspace in the high-dimensional
case involving 33 uncertain parameters. Based on the prior marginals of the uncertain parameters, our results
indicate that the uncertainty in the ignition delay is largely dependent on $A_1$, $A_9$, and $E_{a,1}$, although
 contributions from $A_{15}$ and $A_{17}$ are also found to be significant.

 The authors would like to highlight that the computational strategy presented in this work is agnostic to the choice
 of the chemical system and could thus be easily adapted for other systems. However, it is important to note that
 its applicability depends upon the nature of the relationship between the model output and the inputs. Specifically,
 the model output is required to be differentiable throughout the considered input domain to be able to evaluate
 the active subspace with reasonable accuracy. Hence, it is a realistic possibility that an active subspace might not
 exist for a given input-output relationship. While the active subspace could be exploited for a forward propagation of
 the uncertainty from inputs to the output, the low-dimensional surrogate could further be used to accelerate 
 calibration of the rate-controlling parameters in a Bayesian setting. Moreover, the activity scores could help
 identify important stages in a complex reaction mechanism and thereby guide development of a reduced-order
 surrogate. Furthermore, dimension reduction using active subspaces could be useful in studies aimed at
 investigating the inadequacy or model error associated with the reaction rate laws for such systems.

### 2018-07-29 Paper/method.tex
..

@@
 and adjoint-based gradient computation~\cite{Jameson:1988,Gunzburger:2003,Borzi:2011,Alexanderian:2017}.
-Model evaluations at neighbouring points are required if finite difference is used. Hence, for 
+Model evaluations at neighboring points are required if finite difference is used. Hence, for 
 $N$ samples in a $d$-dimensional parameter space, $N(d+1)$ model evaluations are needed.
@@
  $G(\bm{w}_1^\top\bm{\xi})$ is used as a surrogate to characterize and quantify the uncertainty in the model
- output. Moreover, the components of the dominant eigenvectors are used to compute the activity scores which
- provide an insight into the relative importance of the uncertain parameters. The sequence of steps for the grad-based
- approach are provided in Algorithm~\ref{alg:grad}.
+ output. Moreover, the components of the dominant eigenvectors are used to compute the activity scores ($\alpha_i$'s)
+ which provide an insight into the relative importance of the uncertain parameters:
+ %
+ \be
+\alpha_i(n) = \sum\limits_{j=1}^{\mathcal{S}}\Lambda_{1_{i,j}}w_{1_{i,j}}^2,~i = 1,\ldots,N_p
+\label{eq:as}
+\ee
+%
+where $n$ is the sample size. The sequence of steps for the grad-based approach are provided in 
+Algorithm~\ref{alg:grad}.
 
@@
 \Require $\theta_l$, $\theta_u$, $\tau$. 
-\Ensure $\Lambda$, $W$, $\eta$. 
+\Ensure $\bm{\hat{\Lambda}}$, $\bm{\hat{W}}$, $\bm{\alpha}$. 
     \Procedure{Gradient-based}{}
@@
 	\State Project to the physical space:
-        $\{\bm{\theta}_k\}_{k=1}^{n_1}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_1}$
+        $\{\bm{\theta}_k\}_{k=1}^{n_1}=\theta_l+0.5(\theta_u-\theta_l)(\{\bm{\xi}_k\}_{k=1}^{n_1}$+1)
 	\State $N_\text{total}$ = $n_1$ 
@@
 	\Statex\hspace{5mm} iii. Project to the physical space:
-        $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\Xi}_k\}_{k=1}^{n_1(N_p+1)}$
+        $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}=\theta_l+0.5(\theta_u-\theta_l)(\{\bm{\Xi}_k\}_{k=1}^{n_1(N_p+1)}$+1)
 	\Statex\hspace{5mm} iv. Using the augmented set, $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}$
-        to compute $\bm{g}^k$. 
-	\State Compute the matrix, $\mathcal{C}$ = 
+         compute $\bm{g}^k$. 
+	\State Compute the matrix, $\hat{\mat{C}}$ = 
         $\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^k][\bm{g}^k]^\top$
-	\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$
-	\State Partition the Eigenpairs: $\Lambda^{(0)}~=~ 
-        \begin{bmatrix} \Lambda_1^{(0)} & \\ & \Lambda_2^{(0)} \end{bmatrix}$, 
-        $W^{(0)}~=~\begin{bmatrix} w_1^{(0)} & w_2^{(0)} \end{bmatrix}$, 
-        $\Lambda_1^{(0)}\in \mathbb{R}^{N_p\times\mathcal{S}}$
+	\State Eigenvalue decomposition, $\hat{\mat{C}}$ = $\bm{\hat{W}^{(0)}}\bm{\hat{\Lambda}^{(0)}} 
+	\bm{\hat{W}^{(0)\top}}$
+	\State Partition the Eigenpairs: $\bm{\hat{\Lambda}^{(0)}}~=~ 
+        \begin{bmatrix} \bm{\Lambda_1^{(0)}} & \\ & \bm{\Lambda_2^{(0)}} \end{bmatrix}$, 
+        $\bm{\hat{W}^{(0)}}~=~\begin{bmatrix} \bm{w_1^{(0)}} & \bm{w_2^{(0)}} \end{bmatrix}$, 
+        $\bm{\Lambda_1^{(0)}}\in \mathbb{R}^{N_p\times\mathcal{S}}$
 	\State Set $r$ = 0
@@
              	$k=n_{r-1}+1, \ldots, n_{r-1}+n_r$.  
-		\State Compute $\mathcal{C}$ = 
+		\State Compute $\mat{C}$ = 
         	$\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^k][\bm{g}^k]^\top$
-		\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
-		\State Compute $\delta w_{1,j}^{(r)}$ = 
-                       \scalebox{1.25}{$\frac{\|(w_{1,j}^{r})^2 - 
-                       (w_{1,j}^{r-1})^2\|}{\|(w_{1,j}^{r-1})^2\|}$}, 
+		\State Eigenvalue decomposition, $\hat{\mat{C}}$ = $\bm{\hat{W}^{(r)}}\bm{\hat{\Lambda}^{(r)}} 
+		\bm{\hat{W}^{(r)\top}$}
+		\State Compute $\bm{\delta w_{1,j}^{(r)}}$ = 
+                       \scalebox{1.25}{$\frac{\|(\bm{w_{1,j}^{r}})^2 - 
+                       (\bm{w_{1,j}^{r-1}})^2\|}{\|(\bm{w_{1,j}^{r-1}})^2\|}$}, 
                        $j = 1,\ldots,\mathcal{S}$.
-		\If {$\max\left(\delta w_{1,j}^{(r)}\right)<\tau$}
+		\If {$\max\left(\bm{\delta w_{1,j}^{(r)}}\right)<\tau$}
 			\State break
@@
 	\EndLoop
-	\State Compute $\alpha_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \Lambda_{1_{i,j}}w_{1_{i,j}}^2$,
+	\State Compute $\alpha_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \hat{\Lambda}_{1_{i,j}}w_{1_{i,j}}^2$,
 	$i=1,\ldots,d$.
@@
  \begin{center}
-  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp}
-   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1_5}
+  \includegraphics[width=0.42\textwidth]{./Figures/eig_comp}
+   \includegraphics[width=0.45\textwidth]{./Figures/err_eigv_1_5}
 \caption{Left: A comparison of the eigenvalue spectrum using $n$ = $\{20,40,80,120\}$ samples with that
@@
 set of $M$ samples is also drawn according to the joint probability distribution of the inputs. For each sample
-in $M$, $p$ nearest neighbours in the $n_1$ set are identified. A least-squares fit to model evaluations at
+in $M$, $p$ nearest neighbors in the $n_1$ set are identified. A least-squares fit to model evaluations at
 the $p$ samples is performed. The slope vector ($\bm{b}_i$) of the regression-fit is stored separately and the process is
@@
 \Require $\theta_l$, $\theta_u$, $M$, $\tau$. 
-\Ensure $\Lambda$, $W$, $\eta$. 
+\Ensure $\bm{\hat{\Lambda}}$, $\bm{\hat{W}}$, $\bm{\alpha}$. 
     \Procedure{Gradient-free}{}
@@
 	\State Project to the physical space:
-        $\{\bm{\theta}_k\}_{k=1}^{n_1}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_1}$
+        $\{\bm{\theta}_k\}_{k=1}^{n_1}=\theta_l+0.5(\theta_u-\theta_l)(\{\bm{\xi}_k\}_{k=1}^{n_1}$+1)
 	\State $N_\text{total}$ = $n_1$ 
 	\State Compute $G(\bm\theta_k)$, $k=1, \ldots, N_\text{total}$.
-	\State Draw $M$ random samples, $\{\bm{\nu}_k\}_{k=1}^{M}$ $\in$ [-1,1]
+	\State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$ $\in$ [-1,1]
          according to $\bm{f(\xi)}$.
-	\State Construct the matrix, $\mathcal{C}$.
-	\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$
-	\State Partition the Eigenpairs: $\Lambda^{(0)}~=~ 
-        \begin{bmatrix} \Lambda_1^{(0)} & \\ & \Lambda_2^{(0)} \end{bmatrix}$, 
-        $W^{(0)}~=~\begin{bmatrix} w_1^{(0)} & w_2^{(0)} \end{bmatrix}$, 
-        $\Lambda_1^{(0)}\in \mathbb{R}^{d\times\mathcal{S}}$
+	\State Construct the matrix, $\hat{\mat{C}}$ using Algorithm~\ref{alg:free2}.
+	\State Eigenvalue decomposition, $\hat{\mat{C}}$ = 
+	$\bm{\hat{W}^{(0)}}\bm{\hat{\Lambda}^{(0)}} \bm{\hat{W}^{(0)\top}}$
+	\State Partition the Eigenpairs: $\bm{\hat{\Lambda}^{(0)}}~=~ 
+        \begin{bmatrix} \bm{\Lambda_1^{(0)}} & \\ & \bm{\Lambda_2^{(0)}} \end{bmatrix}$, 
+        $\bm{\hat{W}^{(0)}}~=~\begin{bmatrix} \bm{w_1^{(0)}} & \bm{w_2^{(0)}} \end{bmatrix}$, 
+        $\bm{\Lambda_1^{(0)}}\in \mathbb{R}^{d\times\mathcal{S}}$
 	\State Set $r$ = 0
@@
 		\State Compute $G(\bm\theta_k)$, $k=n_{r-1}+1, \ldots, n_{r-1}+n_r$.  
-		\State Construct the matrix, $\mathcal{C}$.
-		\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
-		\State Compute $\delta w_{1,j}^{(r)}$ = 
-                       \scalebox{1.25}{$\frac{\|(w_{1,j}^{r})^2 - 
-                       (w_{1,j}^{r-1})^2\|}{\|(w_{1,j}^{r-1})^2\|}$}, 
+		\State Construct the matrix, $\hat{\mat{C}}$ using Algorithm~\ref{alg:free2}.
+		\State Eigenvalue decomposition, $\hat{\mat{C}}$ = $\bm{\hat{W}^{(r)}}\bm{\hat{\Lambda}^{(r)}}
+		 \bm{\hat{W}^{(r)\top}}$
+		\State Compute $\bm{\delta w_{1,j}^{(r)}}$ = 
+                       \scalebox{1.25}{$\frac{\|(\bm{w_{1,j}^{r}})^2 - 
+                       (\bm{w_{1,j}^{r-1}})^2\|}{\|(\bm{w_{1,j}^{r-1}})^2\|}$}, 
                        $j = 1,\ldots,\mathcal{S}$.
-		\If {$\max\left(\delta w_{1,j}^{(r)}\right)<\tau$}
+		\If {$\max\left(\bm{\delta w_{1,j}^{(r)}}\right)<\tau$}
 			\State break
@@
 	\Procedure{Local Linear Approximation}{}
+	\State Draw $N$ random samples, $\{\bm{\xi}_k\}_{k=1}^{N}$ $\in$ [-1,1]
+	according to $\bm{f(\xi)}$.
+	\State Project to the physical space:
+	$\{\bm{\theta}_k\}_{k=1}^{N}=\theta_l+0.5(\theta_u-\theta_l)(\{\bm{\xi}_k\}_{k=1}^{N}$+1)
+	\State Compute $G(\bm\theta_k)$, $k=1, \ldots, N$.
+	\State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$ $\in$ [-1,1]
+	according to $\bm{f(\xi)}$.
+	\State Choose an integer $p \leq N$ 
+	\State For each $i=1, \ldots, M$, compute 
+	\[
+	\begin{aligned}
+	\Phi_i &= \{ p \text{ nearest points in } \{\bm{\xi}_k\}_{k=1}^{N} \text{ to } \bm{\nu}_i\}\\
+	\vspace{-2mm}
+	\Psi_i &= \text{subset of } G(\bm\theta_k) \text{ corresponding to the points in } \Phi_i
+	\end{aligned}
+	\]
+	\State Compute linear least square fits 
+	 $G(\bm\theta_k) \approx c_i + \bm{b}_i^T\bm{\xi}_k$, \  $\bm{\xi}_k \in \Phi_i, \ G(\bm\theta_k) \in \Psi_i$
+	\State Compute the matrix 
+	\[
+	\hat{\mat{C}} \approx \frac{1}{M} \sum_{i=1}^{M} \bm{b}_i \bm{b}_i^\top = 
+	{\hat{\bm{W}}}{\hat{\bm{\Lambda}}}{\hat{\bm{W}}}^\top
+	\]
 	\EndProcedure
@@
  \begin{center}
-  \includegraphics[width=0.48\textwidth]{./Figures/comp_ssp}
+  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp}
 \caption{An illustrative comparison of the SSPs generated using the grad-based and the grad-free approaches.}
@@
 We extend the comparative assessment of the two approaches by estimating the activity score for individual
-uncertain parameters ($\alpha_i$) using
-the components of the dominant eigenvector in the following equation:
-%
-\be
-\alpha_i(n) = \sum\limits_{j=1}^{\mathcal{S}}\Lambda_{1_{i,j}}w_{1_{i,j}}^2,~i = 1,\ldots,N_p
-\ee
-%
-where $n$ is the sample size. The activity scores for the 19 uncertain pre-exponents ($A_i$'s), estimated using
-the two approaches are plotted in Figure~\ref{fig:comp_as} (left). Additionally, for the purpose of verification, we 
+uncertain parameters ($\alpha_i$) using~\eqref{eq:as}.
+The activity scores for the 19 uncertain pre-exponents ($A_i$'s) are plotted in Figure~\ref{fig:comp_as} (left). 
+Additionally, for the purpose of verification, we 
 compare these estimates with those obtained using the DGSMs~\cite{Vohra:2018} in 
@@
  \begin{center}
-  \includegraphics[width=0.45\textwidth]{./Figures/comp_as}
-  \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}
+  \includegraphics[width=0.42\textwidth]{./Figures/comp_as}
+  \includegraphics[width=0.42\textwidth]{./Figures/ub_conv_kinetics_rich}
 \caption{Left: A bar-graph of activity scores ($\alpha_i$'s) for the 19 uncertain pre-exponents ($A_i$'s).
@@
 %
-The activity scores estimated using the two approaches agree favourably with each other as well as those
+The activity scores estimated using the two approaches agree favorably with each other as well as those
 based on the screening metric involving the DGSMs from~\cite{Vohra:2018}. It is observed that the uncertainty

### 2018-07-30 Paper/assa.tex
minor ...

@@
 \label{sec:assa}
-In this section, we provide key elements of the active subspace methodology in~\ref{sub:ac}, and
-provide theoretical connections between the global sensitivity measures: DGSMs, activity scores,
-and the total Sobol index in~\ref{sub:gsa}.
+In this section, we provide key elements of the active subspace methodology
+and provide theoretical connections between the global
+sensitivity measures (DGSMs), activity scores, and the total Sobol' indices.
 
@@
 \label{sub:ac}
-%We recall some background regarding active subspaces~\cite{Constantine:2015}.
-As mentioned earlier, the active subspace is low-dimensional linear subspace that constitutes
-important directions in a model's input space~\cite{Constantine:2015}. The effective variability
-in the model output due to uncertain inputs is predominantly captured along these directions.
-Mathematically, these directions are essentially the eigenvectors of a positive semi-definite
-matrix, $\mat{C}$, defined as follows:
+As mentioned earlier, the active subspace is low-dimensional subspace
+that constitutes important directions in a model's input
+space~\cite{Constantine:2015}. The effective variability in the model output
+due to uncertain inputs is predominantly captured along these directions.
+Mathematically, these directions are the eigenvectors of a positive
+semi-definite matrix, $\mat{C}$, defined as follows:
 %
@@
 %
-where $\Omega$:~$[-1,1]^{N_p}$ is the domain of $\bm{\xi}$, and $N_p$ is the number of uncertain
-inputs. $dP_{\bm{\xi}}$ = $p(\bm{\xi})d\bm{\xi}$, where $p(\bm{\xi})$ is the joint probability
-distribution of $\bm{\xi}$. Note that the function, $f$ is considered to be continuous and differentiable
-in the considered parameter domain. Hence, it is possible that a given $f(\bm{\xi})$ might not admit an
-active subspace. However, it is of remarkable interest to investigate if one exists since the subspace could
-be exploited to reduce the dimensionality of the problem and hence the associated computational effort.
-$\nabla_{\bm{\xi}}f$ denotes a vector with individual components
-being partial derivatives of $f$ with respect to the $i^\text{th}$ input, $\xi_i$. Since $\mat{C}$ is symmetric and
-positive semi-definite, it admits a real eigenvalue decomposition:
+where $\Omega$:~$[-1,1]^{N_p}$ is the domain of $\bm{\xi}$, and $N_p$ is the
+number of uncertain inputs. \alennote{Why do we limit to $[-1, 1]^\Np$?} 
+$dP_{\bm{\xi}}$ = $p(\bm{\xi})d\bm{\xi}$, where $p(\bm{\xi})$ is the joint probability
+distribution of $\bm{\xi}$. Note that the function, $f$ is considered to be continuous 
+and differentiable in the considered parameter domain. 
+%Hence, it is possible that a given $f(\bm{\xi})$ might not admit an active
+%subspace. However, it is of remarkable interest to investigate if one exists
+%since the subspace could be exploited to reduce the dimensionality of the
+%problem and hence the associated computational effort.
+Here 
+$\nabla_{\bm{\xi}}f$ denotes the gradient vector. 
+%with individual components
+%being partial derivatives of $f$ with respect to the $i^\text{th}$ input, $\xi_i$. 
+Since $\mat{C}$ is symmetric and
+positive semi-definite, it admits a spectral decomposition:
 %
@@
 %
-where $\bm{\Lambda}$ = diag($\lambda_1,\ldots,\lambda_{N_p}$) with $\lambda_i$'s
- sorted in descending order, $i:\{1,2,\ldots,N_p\}$. The eigenpairs are partitioned about
- the $i^{\text{th}}$ eigenvalue such that the ratio, $\left(\frac{\lambda_i}{\lambda_{i+1}}\right)$ $\gg$ 1
- as follows:
- %
+where $\bm{\Lambda}$ = diag($\lambda_1,\ldots,\lambda_{N_p}$) with
+$\lambda_i$'s sorted in descending order
+\[
+     \lambda_1 \geq \lambda_2 \geq \cdot \geq \lambda_\Np \geq 0.
+\] 
+The eigenpairs are partitioned about the $i^{\text{th}}$ eigenvalue such that
+the ratio, $\left(\frac{\lambda_i}{\lambda_{i+1}}\right)$ $\gg$ 1 as follows:
 \be
@@
  %
- where $\bm{w}_1$ comprises the dominant eigenspace of $\mat{C}$ regarded as the active subspace,
- and $\bm{\Lambda}_1$ is the corresponding set of eigenvalues. The active subspace is used to 
- construct the so-called active variables, $\bm{w}_1^\top \bm{\xi}$ and $f(\bm{\xi}(\theta))$ is transformed
- as $G(\bm{w}_1^\top \bm{\xi})$. In practice, the set of samples, $\bm{\xi}_i$'s are projected to the
- physical space as $\bm{\theta}_i$'s at which the model output, $f(\bm{\xi}(\theta))$ is computed. 
- The matrix, $\mat{C}$ is typically approximated using pseudo-random sampling techniques such 
- as Monte Carlo and Latin hypercube sampling. The integral in~\eqref{eq:C} is replaced with a
+ where $\bm{w}_1$ comprises the dominant eigenspace of $\mat{C}$ regarded as
+the active subspace, and $\bm{\Lambda}_1$ is the corresponding set of
+eigenvalues. The active subspace is used to construct the so-called active
+variables, $\bm{w}_1^\top \bm{\xi}$ and $f(\bm{\xi}(\theta))$ is transformed as
+$G(\bm{w}_1^\top \bm{\xi})$. In practice, the set of samples, $\bm{\xi}_i$'s
+are projected to the physical space as $\bm{\theta}_i$'s at which the model
+output, $f(\bm{\xi}(\theta))$ is computed.  The matrix, $\mat{C}$ is typically
+approximated using pseudo-random sampling techniques such as Monte Carlo and
+Latin hypercube sampling. The integral in~\eqref{eq:C} is replaced with a
 summation as follows:
@@
  %
- It has been shown that the accuracy of approximated dominant eigenspace, $\hat{\bm{w}}_1$ is inversely
- proportional to the difference between the smallest eigenvalue in $\hat{\bm{\Lambda}}_1$ and the
- largest eigenvalue in $\hat{\bm{\Lambda}}_2$~\cite{Constantine:2014}.
-
- Components of the eigenvectors
- in the active subspace could be used for estimating the so-called activity scores as a measure for
- global sensitivity and also be used for approximating the DGSMs as discussed in the following section.
+%%It has been shown that the accuracy of approximated dominant eigenspace,
+%%$\hat{\bm{w}}_1$ is inversely proportional to the difference between the
+%%smallest eigenvalue in $\hat{\bm{\Lambda}}_1$ and the largest eigenvalue in
+%%$\hat{\bm{\Lambda}}_2$~\cite{Constantine:2014}. 
+%%Components of the eigenvectors in the active subspace could be used for
+%%estimating the so-called activity scores as a measure for global sensitivity
+%%and also be used for approximating the DGSMs as discussed in the following
+%%section.
 
@@
 \label{sub:gsa}
-The total Sobol' index ($T_i(f)$) of a model output, $f(\theta_i)$ quantifies the total contribution of the
-input, $\theta_i$ to the variance of the output. Mathematically, this can be expressed as follows:
+The total Sobol' index ($T_i(f)$) of a model output, $f(\vec\theta)$ 
+quantifies the total contribution of the
+input, $\theta_i$ to the variance of the output.\alennote{references} 
+Mathematically, this can be expressed as follows:
 %
@@
 %
-where $\bm{\theta}_{\sim i}$ is a vector of all but $i^\text{th}$ uncertain input, and $\V$ denotes the variance.
-Determining converged estimates of $T_i(f)$ typically involves tens of thousands of model runs and is therefore
-prohibitive in the case of compute-intensive applications. The DGSMs~\cite{Sobol:2009}
-provide a means for approximating the upper bound on $T_i(f)$ with fewer computations as 
-discussed in~\cite{Vohra:2018}. 
-\smallskip
+where $\bm{\theta}_{\sim i}$ is a vector of all but $i^\text{th}$ uncertain
+input, and $\V$ denotes the variance (~\alennote{Explain this more clearly.}).
+Determining accurate estimates of $T_i(f)$ typically involves tens of
+thousands of model runs and is therefore prohibitive in the case of
+compute-intensive applications. The DGSMs~\cite{Sobol:2009} provide a means for
+approximating the upper bound on $T_i(f)$ with fewer computations; see 
+also~\cite{Vohra:2018}.
 
@@
 $\mat{C} = \sum_{k=1}^\Np \lambda_k \vec{u}_k \vec{u}_k^\top$, where 
-$\lambda_i$ are the (non-negative) eigenvalues of $\mat{C}$ and
+$\lambda_i$ are the (non-negative) eigenvalues of $\mat{C}$, in descending
+order, and
 $\vec{u}_k$ are the corresponding (orthonormal) eigenvectors.
@@
 \end{equation}
-which gives a spectral representation for the DGSMs. Truncating the summation 
-leads to the notion of activity scores introduced in~\cite{Diaz:2016,Constantine:2017}
-defined as follows: 
+which gives a spectral representation for the DGSMs. 
+In the case the eigenvalues decay rapidly to zero, we can obtain
+accurate approximations of $\nu_i(f)$ by Truncating the summation: 
+\newcommand{\act}[3]{\nu_{{#2},{#3}}({#1})}
+\newcommand{\actt}[3]{\tilde{\nu}_{{#2},{#3}}({#1})}
 \[
-   \alpha_i(f; r) =  \sum_{k=1}^r \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2,
+   \act{f}{i}{r} =  \sum_{k=1}^r \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2,
    \quad j = 1, \ldots, \Np, \quad r \leq \Np.
 \]
-The activity scores connect ideas from active subspaces and global sensitivity
-analysis, and can be used to approximate DGSMs.  The following result, which
-can be found in~\cite{Diaz:2016,Constantine:2017}, quantifies the error in this
+The quantities $\act{f}{i}{r}$ are called activity scores
+in~\cite{Diaz:2016,Constantine:2017}.
+%The activity scores connect ideas from active subspaces and global sensitivity
+%analysis, and can be used to approximate DGSMs.  
+The following result, which
+can also be found in~\cite{Diaz:2016,Constantine:2017}, quantifies the error in this
 approximation. We provide a short proof for completeness. 
@@
 \[
-0 \leq \nu_i(f) - \alpha_i(f; r) \leq \lambda_{r+1}, \quad i = 1, \ldots, \Np.
+0 \leq \nu_i(f) - \act{f}{i}{r} \leq \lambda_{r+1}, \quad i = 1, \ldots, \Np.
 \] 
@@
 
-Note that, $\nu_i(f) - \alpha(f; r)  = \sum_{k=r+1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2 \geq 0$,
+Note that, $\nu_i(f) - \act{f}{i}{r}= \sum_{k=r+1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2 \geq 0$,
 which gives the first inequality. To see the upper bound, we note,
@@
 The last inequality holds because 
-$1 = \norm{\vec{e}_i}^2 = 
+$1 = \|\vec{e}_i\|_2^2 = 
 \sum_{k = 1}^\Np \ip{\vec{e}_i}{\vec{u}_k}^2 
@@
 the eigenvalues $\lambda_i$ decay rapidly to zero; in 
-such cases, this result implies that  $\nu_i(f) \approx \alpha_i(f; r)$,
+such cases, this result implies that  $\nu_i(f) \approx \act{f}{i}{r}$,
 where $r$ is the \emph{numerical rank} of $\mat{C}$.  This will be especially
@@
 $\xi_i$ are independent are either uniformly distributed or
-distributed according to a Boltzmann probability distribution.  
+distributed according to a Boltzmann probability distribution~\alennote{Explain what this
+is}.  
 It was shown in~\cite{Lamboni:2013} that the total Sobol' 
@@
 \]
-The following result is a straightforward generalization of
-the result from~\cite{Diaz:2016}, and 
-links activity scores and Sobol' indices.
-\begin{proposition}
-For $1 \leq r \leq \Np$, and $i = 1, \ldots, \Np$, we have
+Using Proposition~\ref{prp:dgsm_bound} and~\eqref{equ:sobol_bound}, 
+we obtain the following useful inequality, which
+links activity scores and Sobol' indices:
+for $1 \leq r \leq \Np$, and $i = 1, \ldots, \Np$, we have
 \[
-       T_i(f) \leq \frac{C_i}{\V(f)}(\alpha_i(f; r) + \lambda_{r+1}).
+       T_i(f) \leq \frac{C_i}{\V(f)}(\act{f}{i}{r} + \lambda_{r+1}).
 \]
-\end{proposition}
-\begin{proof}
-The result follows immediately from Proposition~\ref{prp:dgsm_bound} and~\eqref{equ:sobol_bound}.
-\end{proof}
+This was noted in~\cite{Diaz:2016} for the specific case of all $\xi_i$'s being
+iid $U(0, 1)$, in which case $C_i = 1/\pi^2$, $i = 1, \ldots, \Np$.
+
 We can also bound the normalized DGSMs using activity scores; it is easy to see  
@@
 \tilde{\nu}_i(f) \leq 
-\frac{ C_i \big(\alpha_i(f; r) + \lambda_{r+1}\big)}{\sum_{i=1}^\Np C_i \alpha_i(f; r)}
-=\frac{C_i \alpha_i(f; r)}{\sum_{i=1}^\Np C_i \alpha_i(f; r)} + \kappa_i \lambda_{r+1}, 
+\frac{ C_i \big(\act{f}{i}{r} + \lambda_{r+1}\big)}{\sum_{i=1}^\Np C_i \act{f}{i}{r}}
+=\frac{C_i \act{f}{i}{r}}{\sum_{i=1}^\Np C_i \act{f}{i}{r}} + \kappa_i \lambda_{r+1}, 
 \]
-with $\kappa_i = C_i / (\sum_i C_i \alpha_i(f; r))$. 
+with $\kappa_i = C_i / (\sum_i C_i \act{f}{i}{r})$. 
 In the case where where $\lambda_{r+1} \approx 0$, 
@@
 \[
-   \tilde{\alpha}_i(f; r) =  \frac{C_i \alpha_i(f; r)}{\sum_{i=1}^\Np C_i \alpha_i(f; r)}.
+   \actt{f}{i}{r} =  \frac{C_i \act{f}{i}{r}}{\sum_{i=1}^\Np C_i \act{f}{i}{r}}.
 \]

### 2018-07-30 Paper/assa.tex
comments ...

@@
 \label{sec:assa}
-In this section, we provide key elements of the active subspace methodology
-and provide theoretical connections between the global
-sensitivity measures (DGSMs), activity scores, and the total Sobol' indices.
+%In this section, we provide key elements of the active subspace methodology and
+%provide theoretical connections between the global sensitivity measures
+%(DGSMs), activity scores, and the total Sobol' indices.
 
@@
 \label{sub:ac}
-As mentioned earlier, the active subspace is low-dimensional subspace
+As mentioned in the introduction, the active subspace is low-dimensional subspace
 that constitutes important directions in a model's input
@@
 Mathematically, these directions are the eigenvectors of a positive
-semi-definite matrix, $\mat{C}$, defined as follows:
+semi-definite matrix, $\mat{C}$, defined as follows: (\alennote{Need to 
+say what is the function $f$, first})
 %
@@
 \[
-     \lambda_1 \geq \lambda_2 \geq \cdot \geq \lambda_\Np \geq 0.
+     \lambda_1 \geq \lambda_2 \geq \cdots \geq \lambda_\Np \geq 0.
 \] 
@@
 the active subspace, and $\bm{\Lambda}_1$ is the corresponding set of
-eigenvalues. The active subspace is used to construct the so-called active
+eigenvalues. 
+\alennote{$\vec{w}_1$ is a supposedly the matrix whose columns are 
+the dominant eigenvectors. Denoting a matrix by a notation typically 
+used for a vector is confusing. Use $\mat{W}_1$ instead?}
+
+\alennote{The rest of this paragraph needs to be rewritten, as it is hard to 
+see what is going on; what is $\theta$? I suggest, mentioning $\vec\theta$ 
+and $\vec\xi$ earlier, and from then on make everything function of $\vec\xi$.
+Explain the remaining ideas more clearly.}
+The active subspace is used to construct the so-called active
 variables, $\bm{w}_1^\top \bm{\xi}$ and $f(\bm{\xi}(\theta))$ is transformed as

### 2018-07-30 Paper/intro.tex
comments ...

@@
 
-Time evolution of a chemically reacting system is largely dependent upon rate constants
-associated with individual reactions. The rate constants are typically assumed to exhibit
-a certain correlation with temperature (e.g. Arrhenius-type). Hence, accurate specification
-of the rate-controlling parameters is critical to the fidelity of simulations. However,
-in practical applications, the parameters are either specified using expert knowledge or
-estimated based on a regression-fit to a set of sparse and noisy data. Consequently, the
-underlying uncertainty associated with the parameters and therefore the simulation-prediction
-is typically ignored. Vohra et al. calibrated the Arrhenius diffusion parameters to quantify
-the uncertainty in mixing rates in Zr/Al reactive multilayers using a Bayesian framework
-in~\cite{Vohra:2014}. The framework accounted for prior knowledge,
-measurement noise, and the discrepancy between model and experiments. Additionally, 
-in~\cite{Vohra:2017}, a design strategy was implemented to determine optimal conditions
-for experiments in order to enhance the information gain in the uncertain diffusion parameters.
-Bayesian calibration was accelerated in these efforts using a polynomial 
-chaos~\cite{Ghanem:2003, Xiu:2002} surrogate for the model. However,
-the number of uncertain parameters for calibration were limited to two in both cases. In
-higher dimensions, the construction of surrogate as well as the calibration process become
-remarkably challenging due to increase in computational effort.
+Time evolution of a chemically reacting system is largely dependent upon rate
+constants associated with individual reactions. The rate constants are
+typically assumed to exhibit a certain correlation with temperature (e.g.
+Arrhenius-type). Hence, accurate specification of the rate-controlling
+parameters is critical to the fidelity of simulations. However, in practical
+applications, the parameters are either specified using expert knowledge or
+estimated based on a regression-fit to a set of sparse and noisy data.
+(\alennote{Any references for this?})
+Consequently, the underlying uncertainty associated with the parameters and
+therefore the simulation is typically ignored.
 
-In this paper, we focus our efforts on mitigating the computational challenge posed by uncertainty
-quantification (UQ) for applications involving a large number of chemical reactions. Specifically, we
-accomplish this by implementing an efficient approach that aims to reduce the dimensionality
-of the problem by computing the so-called active subspace~\cite{Constantine:2015}. 
-The active subspace, discussed later in~\ref{sub:ac}, essentially constitutes the dominant
-eigenspace of a matrix derived from the gradient information of the model output with respect
-to an input. The dominant eigenspace essentially comprises `important' directions in the parameter
-space that predominantly capture the variability in the model output due to uncertainty in the
-inputs.
+Vohra et al.~calibrated the Arrhenius diffusion parameters to quantify the
+uncertainty in mixing rates in Zr/Al reactive multilayers using a Bayesian
+framework in~\cite{Vohra:2014}. The framework accounted for prior knowledge,
+measurement noise, and the discrepancy between model and experiments.
+Additionally, in~\cite{Vohra:2017}, a design strategy was implemented to
+determine optimal conditions for experiments in order to enhance the
+information gain in the uncertain diffusion parameters.  Bayesian calibration
+was accelerated in these efforts using a polynomial chaos~\cite{Ghanem:2003,
+Xiu:2002} surrogate for the model. However, the number of uncertain parameters
+for calibration were limited to two in both cases. In higher dimensions, the
+construction of surrogate as well as the calibration process become remarkably
+challenging due to increase in computational effort.~\alennote{I don't see why
+these references are being cited here. Are these cited as examples of efforts
+in calibrating chemical systems?}
 
-The high-dimensional application problem considered in this work is the H$_2$/O$_2$
-reaction mechanism~\cite{Yetter:1991}. This mechanism has been gaining a lot of attention in the
-recent years as a potential source of clean energy for locomotive applications~\cite{Das:1996}.   
-The mechanism involves 19 reactions including chain reactions, dissociation/recombination reactions, and
-formation and consumption of intermediate species. For each reaction in table~\ref{tab:kinetics}, 
-the reaction rate is assumed to show an Arrhenius correlation with temperature:
+In this paper, we focus on mitigating the computational challenge
+posed by uncertainty quantification (UQ) for applications involving a large
+number of chemical reactions. We accomplish this by implementing
+an efficient approach that aims to reduce the dimensionality of the problem. This is
+done by
+computing the so-called active subspace~\cite{Constantine:2015}.  The active
+subspace, discussed later in~\ref{sub:ac}, constitutes the dominant
+eigenspace of a matrix derived from the gradient information of the model
+output with respect to an input. The dominant eigenspace essentially comprises
+``important'' directions in the parameter space that predominantly capture the
+variability in the model output due to uncertainty in the inputs.
+
+The application problem considered in this work is the
+H$_2$/O$_2$ reaction mechanism~\cite{Yetter:1991}. This mechanism has been
+gaining a lot of attention in the recent years as a potential source of clean
+energy for locomotive applications~\cite{Das:1996}. (\alennote{you say recent years, but
+the reference is from 1996, are there more recent papers you can refer to?}.   
+The mechanism involves 19
+reactions including chain reactions, dissociation/recombination reactions, and
+formation and consumption of intermediate species. For each reaction in
+table~\ref{tab:kinetics}, the reaction rate is assumed to show an Arrhenius
+correlation with temperature:
 %
@@
 be considered as follows:
-%
 \be
@@
 \ee 
-%
-The equivalence ration ($\phi$) is given as follows:
+The equivalence ratio ($\phi$) is given as follows:
 %
@@
 %
-where the numerator on the right-hand-side denotes the fuel (H$_2$)-oxidizer (O$_2$) ratio at given
-condition to the same quantity under stoichiometric conditions. In this study, computations were performed
-at fuel-rich conditions, $\phi$ = 2.0. Homogeneous ignition at constant pressure is simulated using the
-TChem software package~\cite{Safta:2011}, and the time required for the rate of temperature increase
-to exceed a given threshold, regarded as `ignition delay' is recorded.
+where the numerator on the right-hand-side denotes the fuel (H$_2$)-oxidizer
+(O$_2$) ratio at given condition to the same quantity under stoichiometric
+conditions. In this study, computations were performed at fuel-rich conditions,
+$\phi$ = 2.0. Homogeneous ignition at constant pressure is simulated using the
+TChem software package~\cite{Safta:2011}, and the time required for the rate of
+temperature increase to exceed a given threshold, regarded as \emph{ignition delay}
+is recorded.
 
-We are interested in investigating the impact of uncertainty in the rate-controlling parameters, 
-pre-exponent~($A_i$) and the activation energy~($E_{a,i}$) on ignition delay. The $A_i$'s are
-considered to be uniformly distributed about their nominal estimates provided in~\cite{Yetter:1991}.
-The total number of uncertain inputs is 33 which makes the
-present problem reasonably high-dimensional involving a significantly large amount of computational
-effort. To reduce the effort, we explore two strategies: one involving the use of gradients (grad-based)
-and the other involving local linear approximation which does require gradient computation (grad-free).
-These strategies are based on algorithms 1.1 and 1.2 provided  in~\cite{Constantine:2015}. However, it
-must be noted that in this work, an iterative approach is adopted for implementing the two strategies.
-Consequently, the amount of computational effort appears to have decreased, and is found to be
-relatively less in the case of grad-free approach as discussed later in section~\ref{sec:method}.
+We are interested in investigating the impact of uncertainty in the
+rate-controlling parameters, pre-exponent~($A_i$) and the activation
+energy~($E_{a,i}$) on ignition delay. The $A_i$'s are considered to be
+uniformly distributed about their nominal estimates provided
+in~\cite{Yetter:1991}.  The total number of uncertain inputs is 33 which makes
+the present problem computationally challenging. 
+To reduce the effort, we explore two
+strategies: 
+a gradient-based approach and  
+a gradient-free approach.
+%involving local linear approximation which does notrequire gradient computation
+%(grad-free).  
+\alennote{Thorough the paper 
+change grad-based to gradient-based, and similarly for grad-free}. 
+%
+These strategies are based on algorithms 1.1 and 1.2 
+in~\cite{Constantine:2015}. In this work, an
+iterative approach is adopted for implementing the two strategies.
+Consequently, the amount of computational effort appears to have decreased
+\alennote{This will be unclear to readers}, and
+is found to be relatively less in the case of grad-free approach as discussed
+later in section~\ref{sec:method}.
 
+\alennote{This paragraph will not be meaningful to readers, especially those not familiar
+with active subspaces; Need to explain DGSMs, their importance, 
+and say something general about our
+approach for approximating them.}
 Components of the dominant eigenvectors can be used to estimate the so-called
-activity scores~\cite{Diaz:2016,Constantine:2017} which provide an insight into the relative importance
-of the uncertain parameters. In 
-section~\ref{sub:gsa}, we demonstrate that the activity scores could be used to approximate the 
-derivative-based global sensitivity measures (DGSMs)~\cite{Sobol:2009}, and establish a relationship between
-the Sobol' total effect index, the DGSMs, and the activity scores. Below, we summarize key contributions of
-the paper and its organization.
+activity scores~\cite{Diaz:2016,Constantine:2017} which provide an insight into
+the relative importance of the uncertain parameters. In section~\ref{sub:gsa},
+we demonstrate that the activity scores could be used to approximate the
+derivative-based global sensitivity measures (DGSMs)~\cite{Sobol:2009}, and
+establish a relationship between the Sobol' total effect index, the DGSMs, and
+the activity scores.
+
+%Below, we summarize key contributions of the paper and its
+%organization.
 
 \paragraph{Contributions}
-Main contributions of this paper are as follows: (1) A perspective on activity scores as means for approximating
-the DGSMs is provided. (2) A general relationship between DGSMs, activity scores, and the total Sobol' index 
-is established. (3) Iterative gradient-based and gradient-free approaches for computing the active subspace are
- demonstrated. The two approaches are shown to be consistent and yield greater
-computational gains compared to the 
- conventional approach. As mentioned above, the gains are greater in the case of grad-free approach. (4) The
- grad-free approach is implemented for discovering an active subspace in a 33-dimensional parameter space 
- for the H$_2$/O$_2$ reaction kinetics application.
+Main contributions of this paper are as follows: (1) A perspective on activity
+scores as means for approximating the DGSMs is provided. (2) A general
+relationship between DGSMs, activity scores, and the total Sobol' index is
+established. (3) Iterative gradient-based and gradient-free approaches for
+computing the active subspace are demonstrated. The two approaches are shown to
+be consistent and yield greater computational gains compared to the
+conventional approach. As mentioned above, the gains are greater in the case of
+grad-free approach. (4) The grad-free approach is implemented for discovering
+an active subspace in a 33-dimensional parameter space for the H$_2$/O$_2$
+reaction kinetics application. 
+\alennote{This says too much. Focus on the actual novel 
+contributions; also, the application should be made more prominent.}
 

### 2018-07-30 Paper/method.tex
Add files via upload

--- 2018-07-27 Paper/method.tex
@@
 	\State Draw $n_1$ random samples, $\{\bm{\xi}_k\}_{k=1}^{n_1}$ $\in$ [-1,1]
-         according to $\bm{f(\xi)}$.
+         according to $\bm{f(\xi)}$. (\hayleynote{Just use N instead then we can get rid of line 4?})
 	\State Project to the physical space:
@@
              $k=1, \ldots, N_\text{total}$.  
-	\Statex\hspace{5mm} Using Finite Difference:
+	\Statex\hspace{5mm} Using Finite Difference: (\hayleynote{I don't think it is necessary to include details on how to calculate the gradient.})
 	\Statex\hspace{5mm} i. Assign a small increment, $d\xi$.
@@
 	\Statex\hspace{5mm} iv. Using the augmented set, $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}$
-        to compute $\bm{g}^k$. 
-	\State Compute the matrix, $\mathcal{C}$ = 
+         compute $\bm{g}^k$. 
+	\State Compute the matrix, $\mat{C}$ = 
         $\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^k][\bm{g}^k]^\top$
-	\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$
+	\State Eigenvalue decomposition, $\mat{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$
 	\State Partition the Eigenpairs: $\Lambda^{(0)}~=~ 
@@
         $W^{(0)}~=~\begin{bmatrix} w_1^{(0)} & w_2^{(0)} \end{bmatrix}$, 
-        $\Lambda_1^{(0)}\in \mathbb{R}^{N_p\times\mathcal{S}}$
+        $\Lambda_1^{(0)}\in \mathbb{R}^{N_p\times\mathcal{S}}$ (\hayleynote{I think using W1 here would be more clear. small letters make it seem like we are talking about one eigenvector})
 	\State Set $r$ = 0
@@
              	$k=n_{r-1}+1, \ldots, n_{r-1}+n_r$.  
-		\State Compute $\mathcal{C}$ = 
+		\State Compute $\mat{C}$ = 
         	$\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^k][\bm{g}^k]^\top$
-		\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
+		\State Eigenvalue decomposition, $\mat{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
 		\State Compute $\delta w_{1,j}^{(r)}$ = 
@@
 	\State Compute $G(\bm\theta_k)$, $k=1, \ldots, N_\text{total}$.
-	\State Draw $M$ random samples, $\{\bm{\nu}_k\}_{k=1}^{M}$ $\in$ [-1,1]
+	\State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$ $\in$ [-1,1]
          according to $\bm{f(\xi)}$.
-	\State Construct the matrix, $\mathcal{C}$.
-	\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$
+	\State Construct the matrix, $\mat{C}$.
+	\State Eigenvalue decomposition, $\mat{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$
 	\State Partition the Eigenpairs: $\Lambda^{(0)}~=~ 
@@
 		\State Compute $G(\bm\theta_k)$, $k=n_{r-1}+1, \ldots, n_{r-1}+n_r$.  
-		\State Construct the matrix, $\mathcal{C}$.
-		\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
+		\State Construct the matrix, $\mat{C}$.
+		\State Eigenvalue decomposition, $\mat{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
 		\State Compute $\delta w_{1,j}^{(r)}$ = 
@@
 	\Procedure{Local Linear Approximation}{}
+	\State Draw $N$ random samples, $\{\bm{\xi}_k\}_{k=1}^{N}$ $\in$ [-1,1]
+	according to $\bm{f(\xi)}$.
+	\State Project to the physical space:
+	$\{\bm{\theta}_k\}_{k=1}^{N}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{N}$
+	\State Compute $G(\bm\theta_k)$, $k=1, \ldots, N$.
+	\State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$ $\in$ [-1,1]
+	according to $\bm{f(\xi)}$.
+	\State Choose an integer $p \leq N$ 
+	\State For each $i=1, \ldots, M$, compute 
+	\[
+	\begin{aligned}
+	\Phi_i &= \{ p \text{ nearest points in } \{\bm{\xi}_k\}_{k=1}^{N} \text{ to } \bm{\nu}_i\}\\
+	\vspace{-2mm}
+	\Psi_i &= \text{subset of } G(\bm\theta_k) \text{ corresponding to the points in } \Phi_i
+	\end{aligned}
+	\]
+	\State Compute linear least square fits 
+	 $G(\bm\theta_k) \approx c_i + b_i^T\bm{\xi}_k$, \  $\bm{\xi}_k \in \Phi_i, \ G(\bm\theta_k) \in \Psi_i$
+	\State Compute the matrix 
+	\[
+	\mat{C} \approx \frac{1}{M} \sum_{i=1}^{M} b_i b_i^T = {W}{\Lambda}{W}^T
+	\]
 	\EndProcedure

### 2018-07-30 Paper/method.tex

@@
 
-\subsection{Grad-based approach}
+\subsection{Grad-based approach} (\hayleynote{Better to use gradient rather than grad})
 \label{sub:grad}
@@
 		\State Draw $n_r$ = $\lceil \beta n_1 \rceil$ new random samples 
-                $\{\bm{\xi}_k\}$ $\in$ [-1,1], $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$.
+                $\{\bm{\xi}_k\}$ $\in$ [-1,1], $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$. (\hayleynote{I am unsure what $\beta$ is? Maybe it has been defined somewhere else?})
+
+                (\hayleynote{We could reduce words below by saying something like repeat steps 3-8 above?})
 		\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.
@@
 \renewcommand{\algorithmicensure}{\textbf{Output:}}
-  \caption{An iterative gradient-free approach for discovering the active subspace}
+  \caption{An iterative gradient-free approach for discovering the active subspace} (\hayleynote{We can alter this the same way as we did for Algorithm 1?})
   \begin{algorithmic}[1]

### 2018-07-31 Paper/intro.tex
..

@@
 applications, the parameters are either specified using expert knowledge or
-estimated based on a regression-fit to a set of sparse and noisy data.
-(\alennote{Any references for this?})
+estimated based on a regression-fit to a set of sparse and noisy
+data~\cite{Burnham:1987, Burnham:1988, Vohra:2011, Sarathy:2012}.
 Consequently, the underlying uncertainty associated with the parameters and
-therefore the simulation is typically ignored.
-
-Vohra et al.~calibrated the Arrhenius diffusion parameters to quantify the
-uncertainty in mixing rates in Zr/Al reactive multilayers using a Bayesian
-framework in~\cite{Vohra:2014}. The framework accounted for prior knowledge,
-measurement noise, and the discrepancy between model and experiments.
-Additionally, in~\cite{Vohra:2017}, a design strategy was implemented to
-determine optimal conditions for experiments in order to enhance the
-information gain in the uncertain diffusion parameters.  Bayesian calibration
-was accelerated in these efforts using a polynomial chaos~\cite{Ghanem:2003,
-Xiu:2002} surrogate for the model. However, the number of uncertain parameters
-for calibration were limited to two in both cases. In higher dimensions, the
-construction of surrogate as well as the calibration process become remarkably
-challenging due to increase in computational effort.~\alennote{I don't see why
-these references are being cited here. Are these cited as examples of efforts
-in calibrating chemical systems?}
-
-In this paper, we focus on mitigating the computational challenge
-posed by uncertainty quantification (UQ) for applications involving a large
-number of chemical reactions. We accomplish this by implementing
-an efficient approach that aims to reduce the dimensionality of the problem. This is
-done by
-computing the so-called active subspace~\cite{Constantine:2015}.  The active
-subspace, discussed later in~\ref{sub:ac}, constitutes the dominant
+therefore the simulation is typically ignored. Nevertheless, the inclusion of
+uncertainty due to corrupted calibration data, parametric uncertainty,
+and the discrepancy between predictions and experiments
+due to model inadequacy has gained prominence in recent
+years~\cite{Vohra:2014, Vohra:2017, Morrison:2018, Hantouche:2018}. 
+However, in complex mechanisms involving a large number of reactions,
+characterizing the propagation of uncertainty from a large set of inputs to
+the model output becomes remarkably challenging due to associated
+computational effort. 
+A major focus of this paper pertains to the implementation of a robust
+framework that aims to identify `important' directions in the input space
+that predominantly captures the variability in the model output. These
+directions, referred to as the \textit{active subspaces}~\cite{Constantine:2015},
+constitute the dominant
 eigenspace of a matrix derived from the gradient information of the model
-output with respect to an input. The dominant eigenspace essentially comprises
-``important'' directions in the parameter space that predominantly capture the
-variability in the model output due to uncertainty in the inputs.
+output with respect to an input. The active subspace methodology thus
+focuses on reducing the dimensionality of the problem.
 

### 2018-07-31 Paper/intro.tex
..

@@
 The application problem considered in this work is the
-H$_2$/O$_2$ reaction mechanism~\cite{Yetter:1991}. This mechanism has been
-gaining a lot of attention in the recent years as a potential source of clean
-energy for locomotive applications~\cite{Das:1996}. (\alennote{you say recent years, but
-the reference is from 1996, are there more recent papers you can refer to?}.   
+H$_2$/O$_2$ reaction mechanism~\cite{Yetter:1991}. This mechanism has gained
+a lot of attention as a potential source of clean energy
+locomotive applications~\cite{Das:1996}, and more recently in fuel cells~\cite{Loges:2008}. 
 The mechanism involves 19

### 2018-07-31 Paper/method.tex
-m '..'

@@
              $k=1, \ldots, N_\text{total}$.
+    \iffalse  
 	\Statex\hspace{5mm} Using Finite Difference: (\hayleynote{I don't think it is necessary to include details on how to calculate the gradient.})
@@
          compute $\bm{g}^k$.
+         \fi 
 	\State Compute the matrix, $\mat{C}$ = 
@@
         \begin{bmatrix} \Lambda_1^{(0)} & \\ & \Lambda_2^{(0)} \end{bmatrix}$, 
-        $W^{(0)}~=~\begin{bmatrix} w_1^{(0)} & w_2^{(0)} \end{bmatrix}$, 
+        $W^{(0)}~=~\begin{bmatrix} W_1^{(0)} & W_2^{(0)} \end{bmatrix}$, 
         $\Lambda_1^{(0)}\in \mathbb{R}^{N_p\times\mathcal{S}}$ (\hayleynote{I think using W1 here would be more clear. small letters make it seem like we are talking about one eigenvector})
@@
         \begin{bmatrix} \Lambda_1^{(0)} & \\ & \Lambda_2^{(0)} \end{bmatrix}$, 
-        $W^{(0)}~=~\begin{bmatrix} w_1^{(0)} & w_2^{(0)} \end{bmatrix}$, 
+        $W^{(0)}~=~\begin{bmatrix} W_1^{(0)} & W_2^{(0)} \end{bmatrix}$, 
         $\Lambda_1^{(0)}\in \mathbb{R}^{d\times\mathcal{S}}$

### 2018-07-31 Paper/method.tex
-m 'changes'

--- 2018-07-30 Paper/method.tex
@@
 	\State Draw $n_1$ random samples, $\{\bm{\xi}_k\}_{k=1}^{n_1}$ $\in$ [-1,1]
-         according to $\bm{f(\xi)}$. (\hayleynote{Just use N instead then we can get rid of line 4?})
+         according to $\bm{f(\xi)}$. 
+    \State Set $N_\text{total}$ = $n_1$ 
 	\State Project to the physical space:
         $\{\bm{\theta}_k\}_{k=1}^{n_1}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_1}$
-	\State $N_\text{total}$ = $n_1$ 
-	\State Compute $\bm{g}^k = \nabla_{\bm{\theta}}G(\bm\theta_k)$, 
-             $k=1, \ldots, N_\text{total}$.  
-	\Statex\hspace{5mm} Using Finite Difference: (\hayleynote{I don't think it is necessary to include details on how to calculate the gradient.})
+	\State For each $k=1, \ldots, N_\text{total}$, compute $G(\bm\theta_k)$ and the gradient $\bm{g}^k = \nabla_{\bm{\theta}}G(\bm\theta_k)$
+    \iffalse  
+	\Statex\hspace{5mm} Using Finite Difference: 
 	\Statex\hspace{5mm} i. Assign a small increment, $d\xi$.
@@
          compute $\bm{g}^k$.
-	\State Compute the matrix, $\mat{C}$ = 
-        $\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^k][\bm{g}^k]^\top$
-	\State Eigenvalue decomposition, $\mat{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$
-	\State Partition the Eigenpairs: $\Lambda^{(0)}~=~ 
+         \fi 
+	\State Compute $\mat{C}$ and its eigenvalue decomposition 
+		$\mat{C}$= $\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^k][\bm{g}^k]^\top = W^{(0)}\Lambda^{(0)} W^{(0)\top}$
+	%\State Eigenvalue decomposition, $\mat{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$%
+	\State Partition the eigenpairs: $\Lambda^{(0)}~=~ 
         \begin{bmatrix} \Lambda_1^{(0)} & \\ & \Lambda_2^{(0)} \end{bmatrix}$, 
-        $W^{(0)}~=~\begin{bmatrix} w_1^{(0)} & w_2^{(0)} \end{bmatrix}$, 
-        $\Lambda_1^{(0)}\in \mathbb{R}^{N_p\times\mathcal{S}}$ (\hayleynote{I think using W1 here would be more clear. small letters make it seem like we are talking about one eigenvector})
+        $W^{(0)}~=~\begin{bmatrix} W_1^{(0)} & W_2^{(0)} \end{bmatrix}$, 
+        $\Lambda_1^{(0)}\in \mathbb{R}^{N_p\times\mathcal{S}}$ \hayleynote{What is S?}
+
 	\State Set $r$ = 0
@@
 		\State Draw $n_r$ = $\lceil \beta n_1 \rceil$ new random samples 
-                $\{\bm{\xi}_k\}$ $\in$ [-1,1], $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$.
-		\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.
-		\State $N_\text{total}$ = $N_\text{total}$ + $n_r$ 
+                $\{\bm{\xi}_k\}$ $\in$ [-1,1], $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$. (\hayleynote{I am unsure what $\beta$ is? Maybe it has been defined somewhere else?})
+
+	%	\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.%
+		\State Set $N_\text{total}$ = $N_\text{total}$ + $n_r$ 
+		\iffalse
 		\State Compute $\bm{g}^k = \nabla_{\bm{\theta}}G(\bm\theta_k)$, 
@@
 		\State Eigenvalue decomposition, $\mat{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
-		\State Compute $\delta w_{1,j}^{(r)}$ = 
-                       \scalebox{1.25}{$\frac{\|(w_{1,j}^{r})^2 - 
-                       (w_{1,j}^{r-1})^2\|}{\|(w_{1,j}^{r-1})^2\|}$}, 
+		\fi
+		\State Repeat steps 4-7 above
+		\State Compute $\delta W_{1,j}^{(r)}$ = 
+                       \scalebox{1.25}{$\frac{\|(W_{1,j}^{r})^2 - 
+                       (W_{1,j}^{r-1})^2\|}{\|(W_{1,j}^{r-1})^2\|}$}, 
                        $j = 1,\ldots,\mathcal{S}$.
-		\If {$\max\left(\delta w_{1,j}^{(r)}\right)<\tau$}
+		\If {$\max\left(\delta W_{1,j}^{(r)}\right)<\tau$}
 			\State break
@@
     \Procedure{Gradient-free}{}
-	\State Draw $n_1$ random samples, $\{\bm{\xi}_k\}_{k=1}^{n_1}$ $\in$ [-1,1]
+    \State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$ $\in$ [-1,1]
     according to $\bm{f(\xi)}$.
+    \State Set $r$ = 0
+	\State Draw $n_r$ random samples, $\{\bm{\xi}_k\}_{k=1}^{n_1}$ $\in$ [-1,1]
+         according to $\bm{f(\xi)}$. Set $N_\text{total}$ = $n_r$ 
 	\State Project to the physical space:
         $\{\bm{\theta}_k\}_{k=1}^{n_1}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_1}$
-	\State $N_\text{total}$ = $n_1$ 
 	\State Compute $G(\bm\theta_k)$, $k=1, \ldots, N_\text{total}$.
-	\State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$ $\in$ [-1,1]
-         according to $\bm{f(\xi)}$.
-	\State Construct the matrix, $\mat{C}$.
-	\State Eigenvalue decomposition, $\mat{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$
-	\State Partition the Eigenpairs: $\Lambda^{(0)}~=~ 
-        \begin{bmatrix} \Lambda_1^{(0)} & \\ & \Lambda_2^{(0)} \end{bmatrix}$, 
-        $W^{(0)}~=~\begin{bmatrix} w_1^{(0)} & w_2^{(0)} \end{bmatrix}$, 
-        $\Lambda_1^{(0)}\in \mathbb{R}^{d\times\mathcal{S}}$
-	\State Set $r$ = 0
+	\State Construct the matrix $\mat{C}$.
+	\State Eigenvalue decomposition, $\mat{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
+	\State Partition the Eigenpairs: $\Lambda^{(r)}~=~ 
+        \begin{bmatrix} \Lambda_1^{(r)} & \\ & \Lambda_2^{(r)} \end{bmatrix}$, 
+        $W^{(r)}~=~\begin{bmatrix} W_1^{(r)} & W_2^{(r)} \end{bmatrix}$, 
+        $\Lambda_1^{(r)}\in \mathbb{R}^{d\times\mathcal{S}}$
 	\Loop
 		\State Set $r$ = $r$ + 1
-		\State Draw $n_r$ = $\lceil \beta n_1 \rceil$ new random samples 
-                $\{\bm{\xi}_k\}$ $\in$ [-1,1], $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$.
-		\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.
+		\State Draw $n_r$ \iffalse = $\lceil \beta n_1 \rceil$ \fi new random samples 
+                $\{\bm{\xi}_k\}_{k=1}^{n_r}$ $\in$ [-1,1] \iffalse $\beta = 2,\ldots $ $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$. \fi
         \State $N_\text{total}$ = $N_\text{total}$ + $n_r$   
+        \State Repeat steps 5-9  
+		\iffalse \State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.
+
 		\State Compute $G(\bm\theta_k)$, $k=n_{r-1}+1, \ldots, n_{r-1}+n_r$.  
@@
 		\State Eigenvalue decomposition, $\mat{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
-		\State Compute $\delta w_{1,j}^{(r)}$ = 
-                       \scalebox{1.25}{$\frac{\|(w_{1,j}^{r})^2 - 
-                       (w_{1,j}^{r-1})^2\|}{\|(w_{1,j}^{r-1})^2\|}$}, 
+		\fi
+		\State Compute $\delta W_{1,j}^{(r)}$ = 
+                       \scalebox{1.25}{$\frac{\|(W_{1,j}^{r})^2 - 
+                       (W_{1,j}^{r-1})^2\|}{\|(W_{1,j}^{r-1})^2\|}$}, 
                        $j = 1,\ldots,\mathcal{S}$.
-		\If {$\max\left(\delta w_{1,j}^{(r)}\right)<\tau$}
+		\If {$\max\left(\delta W_{1,j}^{(r)}\right)<\tau$}
 			\State break
@@
 	\EndLoop
-	\State Compute $\alpha_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \Lambda_{1_{i,j}}w_{1_{i,j}}^2$,
+	\State Compute $\alpha_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \Lambda_{1_{i,j}}W_{1_{i,j}}^2$,
 	$i=1,\ldots,d$.

### 2018-07-31 Paper/method.tex
-m 'changes'

@@
 \Require $\theta_l$, $\theta_u$, $\tau$. 
-\Ensure $\Lambda$, $W$, $\eta$. 
+\Ensure $\Lambda$, $W$, $\alpha$ %$\eta$. %
     \Procedure{Gradient-based}{}
-	\State Draw $n_1$ random samples, $\{\bm{\xi}_k\}_{k=1}^{n_1}$ $\in$ [-1,1]
+    \State Set $r$ = 0
+	\State Draw $n_r$ random samples, $\{\bm{\xi}_k\}_{k=1}^{n_1}$ $\in$ [-1,1]
          according to $\bm{f(\xi)}$. 
-    \State Set $N_\text{total}$ = $n_1$ 
+    \State Set $N_\text{total}$ = $n_r$ 
 	\State Project to the physical space:
-        $\{\bm{\theta}_k\}_{k=1}^{n_1}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_1}$
+        $\{\bm{\theta}_k\}_{k=1}^{n_r}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_r}$
 	\State For each $k=1, \ldots, N_\text{total}$, compute $G(\bm\theta_k)$ and the gradient $\bm{g}^k = \nabla_{\bm{\theta}}G(\bm\theta_k)$
@@
 	\State Compute $\mat{C}$ and its eigenvalue decomposition 
-		$\mat{C}$= $\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^k][\bm{g}^k]^\top = W^{(0)}\Lambda^{(0)} W^{(0)\top}$
+		$\mat{C}$= $\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^k][\bm{g}^k]^\top = W^{(r)}\Lambda^{(r)} W^{(r)\top}$
 	%\State Eigenvalue decomposition, $\mat{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$%
-	\State Partition the eigenpairs: $\Lambda^{(0)}~=~ 
-        \begin{bmatrix} \Lambda_1^{(0)} & \\ & \Lambda_2^{(0)} \end{bmatrix}$, 
-        $W^{(0)}~=~\begin{bmatrix} W_1^{(0)} & W_2^{(0)} \end{bmatrix}$, 
-        $\Lambda_1^{(0)}\in \mathbb{R}^{N_p\times\mathcal{S}}$ \hayleynote{What is S?}
+	\State Partition the eigenpairs: $\Lambda^{(r)}~=~ 
+        \begin{bmatrix} \Lambda_1^{(r)} & \\ & \Lambda_2^{(r)} \end{bmatrix}$, 
+        $W^{(r)}~=~\begin{bmatrix} W_1^{(r)} & W_2^{(r)} \end{bmatrix}$, 
+        $\Lambda_1^{(r)}\in \mathbb{R}^{N_p\times\mathcal{S}}$ \hayleynote{What is S?}
 
-	\State Set $r$ = 0
 	\Loop
 		\State Set $r$ = $r$ + 1
-		\State Draw $n_r$ = $\lceil \beta n_1 \rceil$ new random samples 
-                $\{\bm{\xi}_k\}$ $\in$ [-1,1], $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$. (\hayleynote{I am unsure what $\beta$ is? Maybe it has been defined somewhere else?})
+		\State Draw $n_r =  \beta n_{r-1}$  new random samples 
+                $\{\bm{\xi}_k\}_{k=1}^{n_r}$ $\in$ [-1,1], $\beta =??$ \iffalse $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$. \fi
 
@@
 		\fi
-		\State Repeat steps 4-7 above
+		\State Repeat steps 5-8 above
 		\State Compute $\delta W_{1,j}^{(r)}$ = 
@@
 \Require $\theta_l$, $\theta_u$, $M$, $\tau$. 
-\Ensure $\Lambda$, $W$, $\eta$. 
+\Ensure $\Lambda$, $W$, $\alpha$ %$\eta$. 
     \Procedure{Gradient-free}{}
-    \State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$ $\in$ [-1,1]
-    according to $\bm{f(\xi)}$.
     \State Set $r$ = 0
@@
 	\State Project to the physical space:
-        $\{\bm{\theta}_k\}_{k=1}^{n_1}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_1}$
+        $\{\bm{\theta}_k\}_{k=1}^{n_r}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_r}$
+     \State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$ $\in$ [-1,1]
+        according to $\bm{f(\xi)}$.
 	\State Compute $G(\bm\theta_k)$, $k=1, \ldots, N_\text{total}$.
@@
 		\State Set $r$ = $r$ + 1
-		\State Draw $n_r$ \iffalse = $\lceil \beta n_1 \rceil$ \fi new random samples 
-                $\{\bm{\xi}_k\}_{k=1}^{n_r}$ $\in$ [-1,1] \iffalse $\beta = 2,\ldots $ $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$. \fi
+		\State Draw $n_r =  \beta n_{r-1}$  new random samples 
+		$\{\bm{\xi}_k\}_{k=1}^{n_r}$ $\in$ [-1,1], $\beta =??$ \iffalse $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$. \fi
         \State $N_\text{total}$ = $N_\text{total}$ + $n_r$   
-        \State Repeat steps 5-9  
+        \State Repeat steps 4-9  
 		\iffalse \State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.
@@
   \begin{algorithmic}[1]
-	\Procedure{Local Linear Approximation}{}
+	\Procedure{Local Linear Approximation}{} \hayleynote{We could get rid of this algorithm completely by putting it in with algorithm 2?}
 	\State Draw $N$ random samples, $\{\bm{\xi}_k\}_{k=1}^{N}$ $\in$ [-1,1]

### 2018-08-01 Paper/method.tex
-m 'changes'
-m '..'

@@
     \Procedure{Gradient-based}{}
-    \State Set $r$ = 0
-	\State Draw $n_r$ random samples, $\{\bm{\xi}_k\}_{k=1}^{n_1}$ $\in$ [-1,1]
-         according to $\bm{f(\xi)}$. 
-    \State Set $N_\text{total}$ = $n_r$ 
-	\State Project to the physical space:
-        $\{\bm{\theta}_k\}_{k=1}^{n_r}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_r}$
-	\State For each $k=1, \ldots, N_\text{total}$, compute $G(\bm\theta_k)$ and the gradient $\bm{g}^k = \nabla_{\bm{\theta}}G(\bm\theta_k)$
+    \State Set $k$ = 0
+	\State Draw $n_k$ random samples, $\{\bm{\xi}_i\}_{i=1}^{n_k}$ 
+         according to $\pi$. 
+    \State Set $N_\text{total}$ = $n_k$ 
+	\iffalse \State Project to the physical space:
+        $\{\bm{\theta}_k\}_{k=1}^{n_r}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_r}$ \fi
+	\State For each $i=1, \ldots, N_\text{total}$, compute $f(\bm{\xi}_i)$ and the gradient $\bm{g}^i = \nabla_{\bm{\xi}}f(\bm{\xi}_i)$
     \iffalse  
@@
 	\Statex\hspace{5mm} iv. Using the augmented set, $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}$
-         compute $\bm{g}^k$.
+         compute $\bm{g}^i$.
          \fi 
 	\State Compute $\mat{C}$ and its eigenvalue decomposition 
-		$\mat{C}$= $\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^k][\bm{g}^k]^\top = W^{(r)}\Lambda^{(r)} W^{(r)\top}$
+		$\mat{C}$= $\frac{1}{N_\text{total}}\sum\limits_{i=1}^{N_\text{total}}[\bm{g}^i][\bm{g}^i]^\top = W^{(k)}\Lambda^{(k)} W^{(k)\top}$
 	%\State Eigenvalue decomposition, $\mat{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$%
-	\State Partition the eigenpairs: $\Lambda^{(r)}~=~ 
-        \begin{bmatrix} \Lambda_1^{(r)} & \\ & \Lambda_2^{(r)} \end{bmatrix}$, 
-        $W^{(r)}~=~\begin{bmatrix} W_1^{(r)} & W_2^{(r)} \end{bmatrix}$, 
-        $\Lambda_1^{(r)}\in \mathbb{R}^{N_p\times\mathcal{S}}$ \hayleynote{What is S?}
-
+	\State Partition: $\Lambda^{(k)}=
+        \begin{bmatrix} \Lambda_1^{(k)} & \\ & \Lambda_2^{(k)} \end{bmatrix}$, 
+        $W^{(k)}=\begin{bmatrix} W_1^{(k)} & W_2^{(k)} \end{bmatrix}$, 
+        $\Lambda_1^{(k)}\in \mathbb{R}^{N_p\times\mathcal{S}}$
 	\Loop
-		\State Set $r$ = $r$ + 1
-		\State Draw $n_r =  \beta n_{r-1}$  new random samples 
-                $\{\bm{\xi}_k\}_{k=1}^{n_r}$ $\in$ [-1,1], $\beta =??$ \iffalse $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$. \fi
+		\State Set $k$ = $k$ + 1
+		\State Draw $n_k =  \beta n_{k-1}$  new random samples 
+                $\{\bm{\xi}_i\}_{i=1}^{n_k}$  $\beta =??$  $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$.
 
 	%	\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.%
-		\State Set $N_\text{total}$ = $N_\text{total}$ + $n_r$ 
-		\iffalse
-		\State Compute $\bm{g}^k = \nabla_{\bm{\theta}}G(\bm\theta_k)$, 
+		\State Set $N_\text{total}$ = $N_\text{total}$ + $n_k$ 
+		\State Compute $\bm{g}^k = \nabla_{\bm{\xi}_i}f(\bm{\xi}_i)$, 
              	$k=n_{r-1}+1, \ldots, n_{r-1}+n_r$.  
@@
 		\State Eigenvalue decomposition, $\mat{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
-		\fi
 		\State Repeat steps 5-8 above
@@
 	\State Compute $G(\bm\theta_k)$, $k=1, \ldots, N_\text{total}$.
-	\State Construct the matrix $\mat{C}$.
-	\State Eigenvalue decomposition, $\mat{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
+	\State Choose an integer $p \leq N$ 
+	\State For each $i=1, \ldots, M$, compute 
+	\[
+	\begin{aligned}
+	\Phi_i &= \{ p \text{ nearest points in } \{\bm{\xi}_k\}_{k=1}^{N} \text{ to } \bm{\nu}_i\}\\
+	\vspace{-2mm}
+	\Psi_i &= \text{subset of } G(\bm\theta_k) \text{ corresponding to the points in } \Phi_i
+	\end{aligned}
+	\]
+	\State Compute linear least square fits 
+	$G(\bm\theta_k) \approx c_i + b_i^T\bm{\xi}_k$, \  $\bm{\xi}_k \in \Phi_i, \ G(\bm\theta_k) \in \Psi_i$
+	\State Compute the matrix 
+	\[
+	\mat{C} \approx \frac{1}{M} \sum_{i=1}^{M} b_i b_i^T = {W}{\Lambda}{W}^T
+	\]
+	%\State Eigenvalue decomposition, $\mat{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
 	\State Partition the Eigenpairs: $\Lambda^{(r)}~=~ 
@@
         \State $N_\text{total}$ = $N_\text{total}$ + $n_r$   
-        \State Repeat steps 4-9  
+        \State Repeat steps 4-11  
 		\iffalse \State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.
@@
   \begin{algorithmic}[1]
-	\Procedure{Local Linear Approximation}{} \hayleynote{We could get rid of this algorithm completely by putting it in with algorithm 2?}
-	\State Draw $N$ random samples, $\{\bm{\xi}_k\}_{k=1}^{N}$ $\in$ [-1,1]
-	according to $\bm{f(\xi)}$.
-	\State Project to the physical space:
-	$\{\bm{\theta}_k\}_{k=1}^{N}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{N}$
-	\State Compute $G(\bm\theta_k)$, $k=1, \ldots, N$.
-	\State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$ $\in$ [-1,1]
-	according to $\bm{f(\xi)}$.
+	\Procedure{Local Linear Approximation}{} 
+	\State Draw $N$ random samples, $\{\bm{\xi}_j\}_{j=1}^{N}$ 
+	according to $\pi$.
+	\iffalse \State Project to the physical space:
+	$\{\bm{\theta}_k\}_{k=1}^{N}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{N}$ \fi
+	\State Compute $f(\bm\xi_j)$, $j=1, \ldots, N$.
+	\State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$
+	according to $\pi$.
 	\State Choose an integer $p \leq N$ 
@@
 	\begin{aligned}
-	\Phi_i &= \{ p \text{ nearest points in } \{\bm{\xi}_k\}_{k=1}^{N} \text{ to } \bm{\nu}_i\}\\
+	\Phi_i &= \{ p \text{ nearest points in } \{\bm{\xi}_j\}_{j=1}^{N} \text{ to } \bm{\nu}_i\}\\
 	\vspace{-2mm}
-	\Psi_i &= \text{subset of } G(\bm\theta_k) \text{ corresponding to the points in } \Phi_i
+	\Psi_i &= \text{subset of } f(\bm\xi_j) \text{ corresponding to the points in } \Phi_i
 	\end{aligned}
@@
 	\State Compute linear least square fits 
-	 $G(\bm\theta_k) \approx c_i + b_i^T\bm{\xi}_k$, \  $\bm{\xi}_k \in \Phi_i, \ G(\bm\theta_k) \in \Psi_i$
+	 $f(\bm\xi_j) \approx c_i + b_i^T\bm{\xi}_j$, \  $\bm{\xi}_j \in \Phi_i, \ f(\bm\xi_j) \in \Psi_i$
 	\State Compute the matrix 

### 2018-08-02 Paper/assa.tex
...

@@
 \subsection{Global sensitivity measures}
+
 \label{sub:gsa}
-The total Sobol' index ($T_i(f)$) of a model output, $f(\vec\theta)$ 
-quantifies the total contribution of the
-input, $\theta_i$ to the variance of the output.\alennote{references} 
-Mathematically, this can be expressed as follows:
+Consider a function $f(\xi_1, \xi_2, \ldots, \xi_\Np)$. 
+Whereas the active subspace framework described above not make any assumptions
+about statistical independence of the input parameters $\xi_i$, $i = 1, \ldots, \Np$, 
+the classical 
+framework of variance based sensitivity analysis~\cite{MMM} assumes the input
+parameters are statistically independent. While extensions to the cases 
+of correlated inputs exist, we limit the discussion in this section to the
+case of random inputs that are statistically independent and are 
+either uniformly distributed or
+distributed according to a Boltzmann probability distribution~\alennote{Explain what this
+is, and give a reference}.
+
+The total Sobol' index ($T_i(f)$) of a model output, $f(\vec\theta)$ quantifies
+the total contribution of the input, $\theta_i$ to the variance of the
+output.\alennote{references} Mathematically, this can be expressed as follows:
 %
@@
 
-Consider a function $y = f(\xi_1, \xi_2, \ldots, \xi_\Np)$ where
-$\xi_i$ are independent are either uniformly distributed or
-distributed according to a Boltzmann probability distribution~\alennote{Explain what this
-is}.  
 It was shown in~\cite{Lamboni:2013} that the total Sobol' 
@@
 \]
-Using Proposition~\ref{prp:dgsm_bound} and~\eqref{equ:sobol_bound}, 
-we obtain the following useful inequality, which
-links activity scores and Sobol' indices:
-for $1 \leq r \leq \Np$, and $i = 1, \ldots, \Np$, we have
+\newcommand{\I}{\mathcal{I}}
+Combining Proposition~\ref{prp:dgsm_bound} and~\eqref{equ:sobol_bound}, shows
+an interesting link between the activity scores and total Sobol' indices.
+Specifically, by computing the activity scores, we can indetify unimportant
+parameters.
+
+One can attempt to reduce parameter dimension by fixing
+unimportant parameters at nominal values.
+
+%Let $\I \subset \{1, \ldots,
+%Np\}$ be an index set that indexes the unimportant parameters in the parameter
+%vector $\vec{\xi}$.  Letting $|\I|$ denote the number of elements of $\I$, 
+%we denote by $\vec{z} \in \R^{|\I|}$ a generic vector of nominal values for
+%the unimportant parameters $\vec\xi_\I$, and let $\vec{y}$ be a generic vector of
+%important variables, $\vec\xi_{\I^c}$,
+%where $\I^c$ denotes the complement of $\I$ in $\{1, \ldots, \Np\}$. 
+Consider the reduced model
+The \emph{reduced} model, obtained by fixing unimportant variables is given by
+\newcommand{\xin}{\vec{\xi}^{(\sim i)}}
 \[
-       T_i(f) \leq \frac{C_i}{\V(f)}(\act{f}{i}{r} + \lambda_{r+1}).
+f_{z}(\vec\xi) = f(\xin, z)
 \] 
-This was noted in~\cite{Diaz:2016} for the specific case of all $\xi_i$'s being
-iid $U(0, 1)$, in which case $C_i = 1/\pi^2$, $i = 1, \ldots, \Np$.
+where $(\vec{y}, \vec{z})$ should be understood as a vector $\vec\xi$ with
+$\vec\xi_{I^c}$ set to $\vec{y}$ and $\vec\xi_{\I}$ set to $\vec{z}$.
+The following result quantifies the approximation error when using $f_\vec{z}$
+in place of $f$:
+\begin{theorem}
+Consider the relative error
+\[
+\delta(z) =
+\frac{ \int_\Omega \big( f(\vec\xi) - f_{z}(\vec\xi)\big)^2 \, \mu(d\vec\xi) }
+          {\int_\Omega \big( f(\vec\xi) )^2 \, \mu(d\vec\xi)}.
+\] 
+We have
+\[
+          \Ez{ \delta(z)} \leq \frac{2 C_i}{\V(f)}\big(\act{f}{i}{r} + \lambda_{r+1}\big).
+\]
+\end{theorem}
+%
+%\[
+%       T_i(f) \leq \frac{C_i}{\V(f)}(\act{f}{i}{r} + \lambda_{r+1}).
+%\]
+%we obtain the following useful inequality, which
+%links activity scores and Sobol' indices:
+%for $1 \leq r \leq \Np$, and $i = 1, \ldots, \Np$, we have
+%\[
+%       T_i(f) \leq \frac{C_i}{\V(f)}(\act{f}{i}{r} + \lambda_{r+1}).
+%\]
+%This was noted in~\cite{Diaz:2016} for the specific case of all $\xi_i$'s being
+%iid $U(0, 1)$, in which case $C_i = 1/\pi^2$, $i = 1, \ldots, \Np$.
 

### 2018-08-02 Paper/assa.tex
added a result on fixing unimportant params ...

@@
 \]
-\newcommand{\I}{\mathcal{I}}
-Combining Proposition~\ref{prp:dgsm_bound} and~\eqref{equ:sobol_bound}, shows
-an interesting link between the activity scores and total Sobol' indices.
-Specifically, by computing the activity scores, we can indetify unimportant
-parameters.
-
-One can attempt to reduce parameter dimension by fixing
-unimportant parameters at nominal values.
-
-%Let $\I \subset \{1, \ldots,
-%Np\}$ be an index set that indexes the unimportant parameters in the parameter
-%vector $\vec{\xi}$.  Letting $|\I|$ denote the number of elements of $\I$, 
-%we denote by $\vec{z} \in \R^{|\I|}$ a generic vector of nominal values for
-%the unimportant parameters $\vec\xi_\I$, and let $\vec{y}$ be a generic vector of
-%important variables, $\vec\xi_{\I^c}$,
-%where $\I^c$ denotes the complement of $\I$ in $\{1, \ldots, \Np\}$. 
-Consider the reduced model
-The \emph{reduced} model, obtained by fixing unimportant variables is given by
-\newcommand{\xin}{\vec{\xi}^{(\sim i)}}
-\[
-f_{z}(\vec\xi) = f(\xin, z)
-\] 
-where $(\vec{y}, \vec{z})$ should be understood as a vector $\vec\xi$ with
-$\vec\xi_{I^c}$ set to $\vec{y}$ and $\vec\xi_{\I}$ set to $\vec{z}$.
-The following result quantifies the approximation error when using $f_\vec{z}$
-in place of $f$:
-\begin{theorem}
-Consider the relative error
-\[
-\delta(z) =
-\frac{ \int_\Omega \big( f(\vec\xi) - f_{z}(\vec\xi)\big)^2 \, \mu(d\vec\xi) }
-          {\int_\Omega \big( f(\vec\xi) )^2 \, \mu(d\vec\xi)}.
-\] 
-We have
-\[
-          \Ez{ \delta(z)} \leq \frac{2 C_i}{\V(f)}\big(\act{f}{i}{r} + \lambda_{r+1}\big).
-\]
-\end{theorem}
-%
-%\[
-%       T_i(f) \leq \frac{C_i}{\V(f)}(\act{f}{i}{r} + \lambda_{r+1}).
-%\]
-%we obtain the following useful inequality, which
-%links activity scores and Sobol' indices:
-%for $1 \leq r \leq \Np$, and $i = 1, \ldots, \Np$, we have
-%\[
-%       T_i(f) \leq \frac{C_i}{\V(f)}(\act{f}{i}{r} + \lambda_{r+1}).
-%\]
-%This was noted in~\cite{Diaz:2016} for the specific case of all $\xi_i$'s being
-%iid $U(0, 1)$, in which case $C_i = 1/\pi^2$, $i = 1, \ldots, \Np$.
 
@@
 \]
+
+Combining Proposition~\ref{prp:dgsm_bound} and~\eqref{equ:sobol_bound}, shows
+an interesting link between the activity scores and total Sobol' indices.
+Specifically, by computing the activity scores, we can indetify unimportant
+parameters.  
+One can attempt to reduce parameter dimension by fixing
+unimportant parameters at nominal values. Suppose $\xi_i$ is
+deemed unimportant as a result of a global sensitivity analysis
+based on computing the activity scores. We want to understand
+what is the approximation error that occurs once we fix $\xi_i$ 
+at a nominal value.
+%Let $\I \subset \{1, \ldots,
+%Np\}$ be an index set that indexes the unimportant parameters in the parameter
+%vector $\vec{\xi}$.  Letting $|\I|$ denote the number of elements of $\I$, 
+%we denote by $\vec{z} \in \R^{|\I|}$ a generic vector of nominal values for
+%the unimportant parameters $\vec\xi_\I$, and let $\vec{y}$ be a generic vector of
+%important variables, $\vec\xi_{\I^c}$,
+%where $\I^c$ denotes the complement of $\I$ in $\{1, \ldots, \Np\}$. 
+%For a vector $\xi$, we let $\xin$ denote the vector obtained by removing
+%$\xi_i$.
+
+Let $\vec\xi$ be given and let $z$ be a nominal value for $\xi_i$;  we define 
+$\vec{y}^z(\vec\xi)$ as the vector with entries $y^z_j = \xi_j$ for $j \neq i$
+and $y^z_i = z_i$.
+We consider the \emph{reduced} model, 
+obtained by fixing $\xi_i$ at a nominal value $z \in \R$: 
+\[
+f^{z}(\vec\xi) = f(\vec{y}^z(\vec\xi))
+\] 
+We consider the following error indicator.
+\[
+\delta(z) =
+\frac{ \int_\Omega \big( f(\vec\xi) - f^{z}(\vec\xi)\big)^2 \, \mu(d\vec\xi) }
+          {\int_\Omega f(\vec\xi)^2 \, \mu(d\vec\xi)}.
+\]
+
+\begin{theorem}
+We have $\Ez{ \delta(z)} \leq 2C_i\big(\act{f}{i}{r} + \lambda_{r+1}\big)/{\V(f)} $.
+\end{theorem}
+%
+%\[
+%       T_i(f) \leq \frac{C_i}{\V(f)}(\act{f}{i}{r} + \lambda_{r+1}).
+%\]
+%we obtain the following useful inequality, which
+%links activity scores and Sobol' indices:
+%for $1 \leq r \leq \Np$, and $i = 1, \ldots, \Np$, we have
+%\[
+%       T_i(f) \leq \frac{C_i}{\V(f)}(\act{f}{i}{r} + \lambda_{r+1}).
+%\]
+%This was noted in~\cite{Diaz:2016} for the specific case of all $\xi_i$'s being
+%iid $U(0, 1)$, in which case $C_i = 1/\pi^2$, $i = 1, \ldots, \Np$.

### 2018-08-04 Paper/assa.tex
..

@@
 due to uncertain inputs is predominantly captured along these directions.
-Mathematically, these directions are the eigenvectors of a positive
-semi-definite matrix, $\mat{C}$, defined as follows: (\alennote{Need to 
-say what is the function $f$, first})
+Mathematically, for a given model output, $f(\bm{\xi})$, and a set of inputs, $\bm{\xi}$,
+these directions are the eigenvectors of a positive
+semi-definite matrix, $\mat{C}$, defined as follows: 
 %
@@
 %
-where $\Omega$:~$[-1,1]^{N_p}$ is the domain of $\bm{\xi}$, and $N_p$ is the
-number of uncertain inputs. \alennote{Why do we limit to $[-1, 1]^\Np$?} 
+where $\Omega\in\mathbb{R}^{N_p}$ is the domain of $\bm{\xi}$, and $N_p$ is the
+number of uncertain inputs. 
 $dP_{\bm{\xi}}$ = $p(\bm{\xi})d\bm{\xi}$, where $p(\bm{\xi})$ is the joint probability
@@
 \be
-\mat{C} = \bm{W}\bm{\Lambda}\bm{W}^\top
+\mat{C} = \mat{W}\mat{\Lambda}\mat{W}^\top
 \ee
 %
-where $\bm{\Lambda}$ = diag($\lambda_1,\ldots,\lambda_{N_p}$) with
+where $\mat{\Lambda}$ = diag($\lambda_1,\ldots,\lambda_{N_p}$) with
 $\lambda_i$'s sorted in descending order
@@
 \be
- \bm{W} = [\bm{w}_1~\bm{w}_2],~~\bm{\Lambda} = \begin{bmatrix}\bm{\Lambda}_1 & \\  &
-  \bm{\Lambda}_2 \\
+ \mat{W} = [\mat{W}_1~\mat{W}_2],~~\mat{\Lambda} = \begin{bmatrix}\mat{\Lambda}_1 & \\  &
+  \mat{\Lambda}_2 \\
 \end{bmatrix}
@@
  %
- where $\bm{w}_1$ comprises the dominant eigenspace of $\mat{C}$ regarded as
-the active subspace, and $\bm{\Lambda}_1$ is the corresponding set of
+ where $\mat{W}_1$ comprises the dominant eigenspace of $\mat{C}$ regarded as
+the active subspace, and $\mat{\Lambda}_1$ is the corresponding set of
 eigenvalues.
-\alennote{$\vec{w}_1$ is a supposedly the matrix whose columns are 
-the dominant eigenvectors. Denoting a matrix by a notation typically 
-used for a vector is confusing. Use $\mat{W}_1$ instead?}
 
@@
 The active subspace is used to construct the so-called active
-variables, $\bm{w}_1^\top \bm{\xi}$ and $f(\bm{\xi}(\theta))$ is transformed as
-$G(\bm{w}_1^\top \bm{\xi})$. In practice, the set of samples, $\bm{\xi}_i$'s
+variables, $\mat{W}_1^\top \bm{\xi}$ and $f(\bm{\xi}(\theta))$ is transformed as
+$G(\mat{W}_1^\top \bm{\xi})$. In practice, the set of samples, $\bm{\xi}_i$'s
 are projected to the physical space as $\bm{\theta}_i$'s at which the model
@@
  (\nabla_{\bm{\xi}}f(\bm{\xi}_i))(\nabla_{\bm{\xi}}f(\bm{\xi}_i))^\top
- = \hat{\bm{W}}\hat{\bm{\Lambda}}\hat{\bm{W}}^\top
+ = \hat{\mat{W}}\hat{\mat{\Lambda}}\hat{\mat{W}}^\top
 \label{eq:chat}

### 2018-08-04 Paper/conc.tex
..

@@
 %-Consistency between the two approaches with DGSM as well as among themselves
-%encourages the use of grad-free.
+%encourages the use of gradient-free.
 %-Grad-free was found to be suitable for a high-dimensional problem.
@@
 Two computational strategies were explored to compute the active subspace for the H$_2$/O$_2$
-kinetics application in this work. The first strategy referred to as grad-based requires computation of gradient of the 
-model output with respect to the input, whereas, the second strategy, referred to as grad-free implements a local
+kinetics application in this work. The first strategy referred to as gradient-based requires computation of gradient of the 
+model output with respect to the input, whereas, the second strategy, referred to as gradient-free implements a local
 linear approximation to the model output in order to estimate the gradient. The two strategies are shown to yield
@@
 for enhancing computational efficiency. 
-Motivated by these findings, the grad-free approach is used for computing the active subspace in the high-dimensional
+Motivated by these findings, the gradient-free approach is used for computing the active subspace in the high-dimensional
 case involving 33 uncertain parameters. Based on the prior marginals of the uncertain parameters, our results

### 2018-08-04 Paper/intro.tex
..

@@
 in~\cite{Yetter:1991}.  The total number of uncertain inputs is 33 which makes
-the present problem computationally challenging. 
-To reduce the effort, we explore two
-strategies: 
-a gradient-based approach and  
-a gradient-free approach.
-%involving local linear approximation which does notrequire gradient computation
-%(grad-free).  
-\alennote{Thorough the paper 
-change grad-based to gradient-based, and similarly for grad-free}. 
-%
-These strategies are based on algorithms 1.1 and 1.2 
-in~\cite{Constantine:2015}. In this work, an
-iterative approach is adopted for implementing the two strategies.
-Consequently, the amount of computational effort appears to have decreased
-\alennote{This will be unclear to readers}, and
-is found to be relatively less in the case of grad-free approach as discussed
-later in section~\ref{sec:method}.
+the present problem high-dimensional and computationally challenging. 
+For the purpose of dimension reduction, the active subspace is evaluated using
+two computational strategies: a gradient-based approach and  
+a gradient-free approach. These strategies are based on algorithms 1.1 and 1.2 
+in~\cite{Constantine:2015}. As mentioned earlier, the gradient information of the model
+output is needed in both cases. However, in the case of gradient-free approach,
+the gradient is approximated using a linear regression-fit to the model output.
+Hence, the amount of computational effort is reduced in this case. Both strategies
+are implemented in an iterative manner i.e. a new set of model evaluations are
+generated at each iteration until convergence is accomplished as discussed
+later in Section~\ref{sec:method}.
 
-\alennote{This paragraph will not be meaningful to readers, especially those not familiar
-with active subspaces; Need to explain DGSMs, their importance, 
-and say something general about our
-approach for approximating them.}
-Components of the dominant eigenvectors can be used to estimate the so-called
-activity scores~\cite{Diaz:2016,Constantine:2017} which provide an insight into
-the relative importance of the uncertain parameters. In section~\ref{sub:gsa},
-we demonstrate that the activity scores could be used to approximate the
-derivative-based global sensitivity measures (DGSMs)~\cite{Sobol:2009}, and
-establish a relationship between the Sobol' total effect index, the DGSMs, and
-the activity scores.
+An alternate approach to dimension reduction involves computing the global sensitivity
+measures associated with uncertain inputs of a model. Uncertainty along the inputs with
+a relatively small estimate of the sensitivity measure can be ignored for the purpose of
+reducing the computational effort associated with uncertainty quantification (UQ). The
+Sobol' indices are commonly used for this purpose~\cite{Sobol:2001}. However, obtaining
+converged estimates of the Sobol' indices typically requires tens of thousands of model
+runs and is therefore impractical in situations involving intensive model simulations.
+Recently, the derivative-based global sensitivity measures~(DGSMs) have been 
+developed to approximate the upper bound on the Sobol' indies with much fewer
+computations~\cite{Sobol:2009, Lamboni:2013}. In Section~\ref{sec:assa}, we show
+that the DGSMs can in fact be approximated using the components of the eigenvectors
+in the active subspace.
 
-%Below, we summarize key contributions of the paper and its
-%organization.
+Main contributions of this paper are as follows: (1) Mathematical linkages between the
+global sensitivity measures: Total Sobol index, DGSMs, and the activity scores are
+established; (2) An efficient gradient-free iterative approach based on local linear
+approximation of the model output is shown to yield consistent results with the
+gradient-based approach. The two approaches were implemented to compute
+the active subspace for a 19-dimensional H$_2$/O$_2$ kinetics problem with 
+uncertain pre-exponents in the rate law for individual reactions. (3) The gradient-free
+approach is finally implemented to discover an active subspace for the 33-dimensional
+H$_2$/O$_2$ reaction kinetics application in which case the pre-exponents as well
+as the activation energies are considered to be uncertain.
 
-\paragraph{Contributions}
-Main contributions of this paper are as follows: (1) A perspective on activity
-scores as means for approximating the DGSMs is provided. (2) A general
-relationship between DGSMs, activity scores, and the total Sobol' index is
-established. (3) Iterative gradient-based and gradient-free approaches for
-computing the active subspace are demonstrated. The two approaches are shown to
-be consistent and yield greater computational gains compared to the
-conventional approach. As mentioned above, the gains are greater in the case of
-grad-free approach. (4) The grad-free approach is implemented for discovering
-an active subspace in a 33-dimensional parameter space for the H$_2$/O$_2$
-reaction kinetics application. 
-\alennote{This says too much. Focus on the actual novel 
-contributions; also, the application should be made more prominent.}
-
- \paragraph{Organization}
- The paper is organized as follows. In Section~\ref{sub:ac}, a brief theoretical background on
+ The remaining of this paper is organized as follows. In Section~\ref{sub:ac}, a brief theoretical background on
  the active subspace methodology is provided. In Section~\ref{sub:gsa}, it is shown that the activity
@@
  DGSMs, and the total Sobol' indices is established. In Section~\ref{sec:method}, a systematic framework
- for computing the active subspace using the grad-based, and the grad-free approaches is presented. 
+ for computing the active subspace using the gradient-based, and the gradient-free approaches is presented. 
 The two approaches are implemented to the H$_2$/O$_2$ reaction kinetics problem whereby only
 the $A_i$'s are considered as uncertain. Our findings illustrate the consistency in the results based on
-the two approaches thereby encouraging the use of the grad-free approach. In Section~\ref{sec:app},
-we further explore the suitability of the grad-free approach by implementing it to a higher-dimensional
-H$_2$/O$_2$ reaction kinetics problem by accounting for the uncertainty in $E_{a,i}$'s as well. 
+the two approaches thereby encouraging the use of the gradient-free approach. In Section~\ref{sec:app},
+the gradient-free approach is implemented to the 33-dimensional
+H$_2$/O$_2$ reaction kinetics application by accounting for the uncertainty in $E_{a,i}$'s as well. The
+resulting active subspace is found to be 1-dimensional indicating enormous potential for reduction in
+computational effort pertaining to UQ, in this case. 
 Finally, we draw important conclusions based on the various aspects of this work in Section~\ref{sec:conc}.

### 2018-08-04 Paper/method.tex
..

@@
 Table~\ref{tab:kinetics}. Specifically, we employ two strategies: One that involves computation of model
-gradients in order to construct the matrix (grad-based), $\hat{\mat{C}}$ in~\eqref{eq:chat}, and the other that involves a local 
+gradients in order to construct the matrix (gradient-based), $\hat{\mat{C}}$ in~\eqref{eq:chat}, and the other that involves a local 
 linear approximation of the model output to circumvent the computational cost associated with gradient
-computation (grad-free). The grad-based approach is based on Algorithm 1.1, and the grad-free approach is
+computation (gradient-free). The gradient-based approach is based on Algorithm 1.1, and the gradient-free approach is
 based on Algorithm 1.2 in~\cite{Constantine:2015}. However, in this work, we have implemented the
@@
 Additionally, for the present application, our findings based on the two strategies are observed to be consistent.
-Since, the grad-free approach is relatively more efficient, this observation encourages its use for high-dimensional
+Since, the gradient-free approach is relatively more efficient, this observation encourages its use for high-dimensional
 applications as discussed later in Section~\ref{sec:app}. Note however, that the model output is required to be
@@
 
-\subsection{Grad-based approach} (\hayleynote{Better to use gradient rather than grad})
+\subsection{Gradient-based approach}
 \label{sub:grad}
 
-As discussed earlier, in the grad-based approach, the elements of $\hat{\mat{C}}$ are estimated using model gradients.
+As discussed earlier, in the gradient-based approach, the elements of $\hat{\mat{C}}$ are estimated using model gradients.
  In situations where the gradient is not available analytically, one could consider using
@@
 evaluations, the matrix, $\hat{\mat{C}}$ is computed. Eigenvalue decomposition of $\hat{\mat{C}}$ yields an initial
-estimate of the dominant eigenspace, $\hat{\bm{w}_1}$ and corresponding set of eigenvalues, $\hat{\bm{\Lambda}_1}$.
+estimate of the dominant eigenspace, $\hat{\mat{W}}_1$ and corresponding set of eigenvalues, $\hat{\mat{\Lambda}}_1$.
 At each subsequent iteration, model evaluations are generated at a new set of $n_s$ samples. The new set
@@
 iterations is evaluated. The process is terminated and the resulting eigenspace is considered to be converged once
-highest relative change at iteration $r$, $\delta w_{1,j}^{(r)}$, is smaller a given tolerance, $\tau$. A regression fit to
- $G(\bm{w}_1^\top\bm{\xi})$ is used as a surrogate to characterize and quantify the uncertainty in the model
- output. Moreover, the components of the dominant eigenvectors are used to compute the activity scores which
- provide an insight into the relative importance of the uncertain parameters. The sequence of steps for the grad-based
- approach are provided in Algorithm~\ref{alg:grad}.
+highest relative change at iteration $k$, $\delta \hat{\mat{W}}_{1,j}^{(k)}$, is smaller a given tolerance, $\tau$. 
+A regression fit to
+ $G(\hat{\mat{W}}_1^\top\bm{\xi})$ is used as a surrogate to characterize and quantify the uncertainty in the model
+ output. Moreover, the components of the dominant eigenvectors are used to compute the activity scores,
+ $\bm{\nu}_r(f)$, which provide an insight into the relative importance of the uncertain inputs. Note that the
+ index, $r$, corresponds to the number of dominant eigenvectors in $\hat{\mat{W}}_1$. The sequence of 
+ steps for the gradient-based approach are provided in Algorithm~\ref{alg:grad}.
 
@@
 \Require $\theta_l$, $\theta_u$, $\tau$. 
-\Ensure $\Lambda$, $W$, $\alpha$ %$\eta$. %
+\Ensure $\hat{\mat{\Lambda}}$, $\hat{\mat{W}}$, $\bm{\nu}_r(f)$ %$\eta$. %
     \Procedure{Gradient-based}{}
@@
          \fi 
-	\State Compute $\mat{C}$ and its eigenvalue decomposition 
-		$\mat{C}$= $\frac{1}{N_\text{total}}\sum\limits_{i=1}^{N_\text{total}}[\bm{g}^i][\bm{g}^i]^\top = W^{(k)}\Lambda^{(k)} W^{(k)\top}$
+	\State Compute $\hat{\mat{C}}$ and its eigenvalue decomposition 
+		$\hat{\mat{C}}$= $\frac{1}{N_\text{total}}\sum\limits_{i=1}^{N_\text{total}}[\bm{g}^i][\bm{g}^i]^\top$ = 
+		$\hat{\mat{W}}^{(k)}\hat{\mat{\Lambda}}^{(k)} \hat{\mat{W}}^{(k)\top}$
 	%\State Eigenvalue decomposition, $\mat{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$%
-	\State Partition: $\Lambda^{(k)}=
-        \begin{bmatrix} \Lambda_1^{(k)} & \\ & \Lambda_2^{(k)} \end{bmatrix}$, 
-        $W^{(k)}=\begin{bmatrix} W_1^{(k)} & W_2^{(k)} \end{bmatrix}$, 
-        $\Lambda_1^{(k)}\in \mathbb{R}^{N_p\times\mathcal{S}}$
+	\State Partition: $\hat{\mat{\Lambda}}^{(k)}=
+        \begin{bmatrix} \hat{\mat{\Lambda}}_1^{(k)} & \\ & \hat{\mat{\Lambda}}_2^{(k)} \end{bmatrix}$, 
+        $\hat{\mat{W}}^{(k)}=\begin{bmatrix} \hat{\mat{W}}_1^{(k)} & \hat{\mat{W}}_2^{(k)} \end{bmatrix}$, 
+        $\hat{\mat{\Lambda}}_1^{(k)}\in \mathbb{R}^{N_p\times r}$
 	\Loop
 		\State Set $k$ = $k$ + 1
-		\State Draw $n_k =  \beta n_{k-1}$  new random samples 
-                $\{\bm{\xi}_i\}_{i=1}^{n_k}$  $\beta =??$  $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$.
+		\State Draw $n_k =  \lceil\beta n_{k-1}\rceil$  new random samples 
+                $\{\bm{\xi}_i\}_{i=1}^{n_k}$  $\beta\in[0,1]$  $k = n_{k-1}+1,\ldots,n_{k-1}+n_k$.
 
@@
 		\State Set $N_\text{total}$ = $N_\text{total}$ + $n_k$ 
-		\State Compute $\bm{g}^k = \nabla_{\bm{\xi}_i}f(\bm{\xi}_i)$, 
-             	$k=n_{r-1}+1, \ldots, n_{r-1}+n_r$.  
-		\State Compute $\mat{C}$ = 
-        	$\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^k][\bm{g}^k]^\top$
-		\State Eigenvalue decomposition, $\mat{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
-		\State Repeat steps 5-8 above
-		\State Compute $\delta W_{1,j}^{(r)}$ = 
-                       \scalebox{1.25}{$\frac{\|(W_{1,j}^{r})^2 - 
-                       (W_{1,j}^{r-1})^2\|}{\|(W_{1,j}^{r-1})^2\|}$}, 
-                       $j = 1,\ldots,\mathcal{S}$.
-		\If {$\max\left(\delta W_{1,j}^{(r)}\right)<\tau$}
+		\State Compute $\bm{g}^i = \nabla_{\bm{\xi}_i}f(\bm{\xi}_i)$, 
+             	$i=n_{k-1}+1, \ldots, n_{k-1}+n_k$.  
+		\State Compute $\hat{\mat{C}}$ = 
+        	$\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^i][\bm{g}^i]^\top$
+		\State Eigenvalue decomposition, $\hat{\mat{C}}$ = $\hat{\mat{W}}^{(k)}\hat{\mat{\Lambda}}^{(k)}
+		 \hat{\mat{W}}^{(k)\top}$
+		\State Compute $\delta \hat{\mat{W}}_{1,j}^{(k)}$ = 
+                       \scalebox{1.25}{$\frac{\|(\hat{\mat{W}}_{1,j}^{k})^2 - 
+                       (\hat{\mat{W}}_{1,j}^{k-1})^2\|}{\|(\hat{\mat{W}}_{1,j}^{k-1})^2\|}$}, 
+                       $j = 1,\ldots,r$.
+		\If {$\max\left(\delta \hat{\mat{W}}_{1,j}^{(k)}\right)<\tau$}
 			\State break
@@
 	\EndLoop
-	\State Compute $\alpha_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \Lambda_{1_{i,j}}W_{1_{i,j}}^2$,
-	$i=1,\ldots,d$.
+	\State Compute $\nu_{i,r}(f) = \sum\limits_{j=1}^{r} \hat{\mat{\Lambda}}_{i,j}\hat{\mat{W}}_{i,j}^2$,
+	$i=1,\ldots,N_p$.
 
@@
 
-\subsection{Grad-free approach}
+\subsection{Gradient-free approach}
 \label{sub:gradfree}
 
-As mentioned earlier, the grad-free approach does not involve computation of the gradient of the model
+As mentioned earlier, the gradient-free approach does not involve computation of the gradient of the model
 output. Instead, a least-squares fit is performed to the local linear regression-based approximation
-to the model output. While the grad-free approach offers a significant potential for additional 
+to the model output. While the gradient-free approach offers a significant potential for additional 
 computational savings, its ability to reasonably approximate the active subspace 
@@
 
-First, we implement the grad-free approach to the 19-dimensional problem and compare our findings with
-those obtained using the grad-based approach. Later, in
-Section~\ref{sec:app}, we compute the active subspace for a 33-dimensional application using the grad-free
+First, we implement the gradient-free approach to the 19-dimensional problem and compare our findings with
+those obtained using the gradient-based approach. Later, in
+Section~\ref{sec:app}, we compute the active subspace for a 33-dimensional application using the gradient-free
 approach. Once again, the strategy is implemented in an iterative manner as shown in~\ref{sub:grad}. 
-It is observed that the results from the grad-free approach are consistent with the grad-based approach
+It is observed that the results from the gradient-free approach are consistent with the gradient-based approach
 as discussed further below.
 
-In the grad-free approach, the model output is first computed at an initial set of $n_1$ samples. An independent
+In the gradient-free approach, the model output is first computed at an initial set of $n_1$ samples. An independent
 set of $M$ samples is also drawn according to the joint probability distribution of the inputs. For each sample
@@
 active subspace. The aforementioned sequence of steps is repeated until $\delta w_{1,j}^{(r)}$ as mentioned
-earlier in~\ref{sub:grad} is found to be smaller than $\tau$. An outline of the grad-free approach is
+earlier in~\ref{sub:grad} is found to be smaller than $\tau$. An outline of the gradient-free approach is
 provided in Algorithm~\ref{alg:free1}, and the specific sequence of steps pertaining to the construction of
@@
     \Procedure{Gradient-free}{}
-    \State Set $r$ = 0
-	\State Draw $n_r$ random samples, $\{\bm{\xi}_k\}_{k=1}^{n_1}$ $\in$ [-1,1]
-         according to $\bm{f(\xi)}$. Set $N_\text{total}$ = $n_r$ 
-	\State Project to the physical space:
-        $\{\bm{\theta}_k\}_{k=1}^{n_r}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_r}$
-     \State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$ $\in$ [-1,1]
-        according to $\bm{f(\xi)}$.
-	\State Compute $G(\bm\theta_k)$, $k=1, \ldots, N_\text{total}$.
-	\State Choose an integer $p \leq N$ 
-	\State For each $i=1, \ldots, M$, compute 
-	\[
-	\begin{aligned}
-	\Phi_i &= \{ p \text{ nearest points in } \{\bm{\xi}_k\}_{k=1}^{N} \text{ to } \bm{\nu}_i\}\\
-	\vspace{-2mm}
-	\Psi_i &= \text{subset of } G(\bm\theta_k) \text{ corresponding to the points in } \Phi_i
-	\end{aligned}
-	\]
-	\State Compute linear least square fits 
-	$G(\bm\theta_k) \approx c_i + b_i^T\bm{\xi}_k$, \  $\bm{\xi}_k \in \Phi_i, \ G(\bm\theta_k) \in \Psi_i$
-	\State Compute the matrix 
-	\[
-	\mat{C} \approx \frac{1}{M} \sum_{i=1}^{M} b_i b_i^T = {W}{\Lambda}{W}^T
-	\]
+    \State Set $k$ = 0
+	\State Draw $n_k$ random samples, $\{\bm{\xi}_j\}_{j=1}^{n_k}$
+	\State Set $N_\text{total}$ = $n_k$ 
+	\iffalse \State Project to the physical space:
+        $\{\bm{\theta}_k\}_{i=1}^{n_k}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_r}$
+        \fi
+     \State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$ 
+	\State Compute $f(\bm\xi_j)$, $j=1, \ldots, N_\text{total}$.
+	\State Compute the matrix $\mathbb{C}$ and its eigenvalue decomposition using Algorithm 3
 	%\State Eigenvalue decomposition, $\mat{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
-	\State Partition the Eigenpairs: $\Lambda^{(r)}~=~ 
-        \begin{bmatrix} \Lambda_1^{(r)} & \\ & \Lambda_2^{(r)} \end{bmatrix}$, 
-        $W^{(r)}~=~\begin{bmatrix} W_1^{(r)} & W_2^{(r)} \end{bmatrix}$, 
-        $\Lambda_1^{(r)}\in \mathbb{R}^{d\times\mathcal{S}}$
+	\State Partition: $\Lambda^{(k)}~=~ 
+        \begin{bmatrix} \Lambda_1^{(k)} & \\ & \Lambda_2^{(k)} \end{bmatrix}$, 
+        $W^{(k)}~=~\begin{bmatrix} W_1^{(k)} & W_2^{(k)} \end{bmatrix}$, 
+        $\Lambda_1^{(k)}\in \mathbb{R}^{d\times\mathcal{S}}$
 	\Loop
-		\State Set $r$ = $r$ + 1
-		\State Draw $n_r =  \beta n_{r-1}$  new random samples 
-		$\{\bm{\xi}_k\}_{k=1}^{n_r}$ $\in$ [-1,1], $\beta =??$ \iffalse $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$. \fi
-        \State $N_\text{total}$ = $N_\text{total}$ + $n_r$   
-        \State Repeat steps 4-11  
-		\iffalse \State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.
+		\State Set $k$ = $k$ + 1
+		\State Draw $n_k =  \beta n_{k-1}$  new random samples 
+		$\{\bm{\xi}_j\}_{j=1}^{n_k}$  $\beta =??$  $j = n_{k-1}+1,\ldots,n_{k-1}+n_k$. 
+        \State $N_\text{total}$ = $N_\text{total}$ + $n_k$    
+	\iffalse	\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$. \fi
 
-		\State Compute $G(\bm\theta_k)$, $k=n_{r-1}+1, \ldots, n_{r-1}+n_r$.  
-		\State Construct the matrix, $\mat{C}$.
-		\State Eigenvalue decomposition, $\mat{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
-		\fi
-		\State Compute $\delta W_{1,j}^{(r)}$ = 
-                       \scalebox{1.25}{$\frac{\|(W_{1,j}^{r})^2 - 
-                       (W_{1,j}^{r-1})^2\|}{\|(W_{1,j}^{r-1})^2\|}$}, 
+		\State Compute $f(\bm\xi_j)$, $j=n_{k-1}+1, \ldots, n_{k-1}+n_k$.  
+		\State Construct the matrix, $\mat{C}$ and its eigenvalue decomposition
+		\State Compute $\delta W_{1,j}^{(k)}$ = 
+                       \scalebox{1.25}{$\frac{\|(W_{1,j}^{k})^2 - 
+                       (W_{1,j}^{k-1})^2\|}{\|(W_{1,j}^{k-1})^2\|}$}, 
                        $j = 1,\ldots,\mathcal{S}$.
-		\If {$\max\left(\delta W_{1,j}^{(r)}\right)<\tau$}
+		\If {$\max\left(\delta W_{1,j}^{(k)}\right)<\tau$}
 			\State break
@@
 
-The grad-free approach when implemented to the 19-dimensional H$_2$/O$_2$ reaction kinetics problem
+The gradient-free approach when implemented to the 19-dimensional H$_2$/O$_2$ reaction kinetics problem
 with uncertain pre-exponents ($A_i$'s) also yielded a 1-dimensional active subspace. In Figure~\ref{fig:comp} (left),
@@
 Additionally, the quantity, $\delta w_{1,j}^{(r)}$ is also plotted in
-Figure~\ref{fig:comp} (right) to illustrate convergence characteristics for the grad-free approach.
+Figure~\ref{fig:comp} (right) to illustrate convergence characteristics for the gradient-free approach.
 %
@@
   \includegraphics[width=0.48\textwidth]{./Figures/comp_ssp}
-\caption{An illustrative comparison of the SSPs generated using the grad-based and the grad-free approaches.}
+\caption{An illustrative comparison of the SSPs generated using the gradient-based and the gradient-free approaches.}
 \label{fig:comp_ssp}
@@
 The above comparisons establish consistency between the two approaches. Note that the computational
-effort in the case of grad-free approach is significantly smaller. Hence, we are encouraged to use the
-grad-free approach for the 33-dimensional application in the next section.
+effort in the case of gradient-free approach is significantly smaller. Hence, we are encouraged to use the
+gradient-free approach for the 33-dimensional application in the next section.

### 2018-08-04 Paper/app.tex
..

@@
 
-The grad-free approach discussed earlier in~\ref{sub:gradfree} was used to compute the
+The gradient-free approach discussed earlier in~\ref{sub:gradfree} was used to compute the
 active subspace. Using the iterative procedure, the convergence of the eigenvectors
-was examined by tracking $\delta w_{1,j}^{(i)}$. In Figure~\ref{fig:conv_app} (left),
+was examined by tracking $\delta \hat{\mat{W}}_{1,j}^{(i)}$. In Figure~\ref{fig:conv_app} (left),
 components of the converged eigenvector based on 50 samples using $\tau$ = 0.2 are plotted.
@@
 \caption{Left: An illustrative comparison of individual components of the converged dominant eigenvector obtained
-using the two approaches discussed in~\ref{sub:grad} and~\ref{sub:gradfree}. Right: The quantity,  $\delta w_{1,j}^{(r)}$
+using the two approaches discussed in~\ref{sub:grad} and~\ref{sub:gradfree}. Right: The quantity,  $\delta \hat{\mat{W}}_{1,j}^{(i)}$
 is plotted for successive iterations to illustrate the convergence behavior.}
@@
 summary plot which shows that the variability in $G$ can be approximated using a 1-dimensional regression
-fit which could be used its a surrogate. We will verify the accuracy of the surrogate later in~\ref{sub:verify}.
+fit which could be used as a surrogate. We will verify the accuracy of the surrogate later in~\ref{sub:verify}.
 
-Components of the dominant eigenvector in Figure~\ref{fig:conv_app} (left) are utilized for estimating the
-activity scores, defined earlier in~\ref{eq:as}. As discussed earlier, the activity scores can be regarded as a 
-measure for global sensitivity of the uncertain parameters. The sensitivity of ignition delay towards the 
-33 uncertain parameters considered in this problem is plotted below in Figure~\ref{fig:as_33D}.
+Components of the dominant eigenvector in Figure~\ref{fig:conv_app} (left) are used for estimating the
+activity scores, plotted below in Figure~\ref{fig:as_33D}.
 %
@@
 subspace approach which is shown to yield even greater scope for dimension reduction i.e. from 33 to 1. We verify
-the accuracy of the 1-d regression fit to $G(\bm{w}_1^\top\bm{\xi})$ in the following section.
+the accuracy of the 1-d regression fit to $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in the following section.
 
@@
 \be
-\varepsilon_d = \frac{\|f(\bm{\theta}) - \mathcal{Y}(\bm{\xi})\|}{\|f(\bm{\theta})\|}
+\varepsilon_d = \frac{\|f(\bm{\theta}) - \mathcal{Y}(\bm{\xi})\|_2}{\|f(\bm{\theta})\|_2}
 \ee

### 2018-08-04 Paper/assa.tex
..

@@
 that constitutes important directions in a model's input
-space~\cite{Constantine:2015}. The effective variability in the model output
-due to uncertain inputs is predominantly captured along these directions.
-Mathematically, for a given model output, $f(\bm{\xi})$, and a set of inputs, $\bm{\xi}$,
+space~\cite{Constantine:2015}. The effective variability in the model output, $f$
+due to uncertain inputs, $\bm{\theta}$ in the physical space, is predominantly captured
+along these directions. Each component of $\bm{\theta}$
+can be mapped to a corresponding component of $\bm{\xi}$ in $\Omega\in\mathbb{R}^{N_p}$
+($N_p$: number of uncertain inputs)
+using a linear transformation.  Hence, from this point onwards, we denote $f$ as $f(\bm{\xi})$.
+Mathematically, for a given $f$, and a set of inputs, $\bm{\xi}$,
 these directions are the eigenvectors of a positive
@@
 %
-where $\Omega\in\mathbb{R}^{N_p}$ is the domain of $\bm{\xi}$, and $N_p$ is the
-number of uncertain inputs. 
+where 
 $dP_{\bm{\xi}}$ = $p(\bm{\xi})d\bm{\xi}$, where $p(\bm{\xi})$ is the joint probability
 distribution of $\bm{\xi}$. Note that the function, $f$ is considered to be continuous 
-and differentiable in the considered parameter domain. 
-%Hence, it is possible that a given $f(\bm{\xi})$ might not admit an active
+and differentiable in the physical space, $\bm{\theta}$. %Hence, it is possible that a given $f(\bm{\xi})$ might not admit an active
 %subspace. However, it is of remarkable interest to investigate if one exists

### 2018-08-04 Paper/method.tex
..

@@
 	\State Draw $n_k$ random samples, $\{\bm{\xi}_i\}_{i=1}^{n_k}$ 
-         according to $\pi$. 
+         according to $\pi_{\bm{\xi}}$. 
     \State Set $N_\text{total}$ = $n_k$ 
@@
 %
-Eigenvalue decomposition of $\hat{\mat{C}}$ yields an initial approximation to the
-active subspace. The aforementioned sequence of steps is repeated until $\delta w_{1,j}^{(r)}$ as mentioned
-earlier in~\ref{sub:grad} is found to be smaller than $\tau$. An outline of the gradient-free approach is
-provided in Algorithm~\ref{alg:free1}, and the specific sequence of steps pertaining to the construction of
-$\hat{\mat{C}}$ is provided in Algorithm~\ref{alg:free2}.
+Hence, the gradient-free approach essentially differs in the procedure for constructing the matrix, $\hat{\mat{C}}$.
+The active subspace computation involves the same sequence of steps as outlined previously for the gradient-based 
+approach in Algorithm~\ref{alg:grad}. The procedure for constructing $\hat{\mat{C}}$ using local linear approximation
+of the model output is provided in Algorithm~\ref{alg:free2}.
 
+%\bigskip
+%\begin{breakablealgorithm}
+%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+%\renewcommand{\algorithmicensure}{\textbf{Output:}}
+%  \caption{An iterative gradient-free approach for discovering the active subspace}
+%  \begin{algorithmic}[1]
+%\Require $\theta_l$, $\theta_u$, $M$, $\tau$. 
+%\Ensure $\Lambda$, $W$, $\alpha$ %$\eta$. 
+%    \Procedure{Gradient-free}{}
+%    \State Set $k$ = 0
+%	\State Draw $n_k$ random samples, $\{\bm{\xi}_j\}_{j=1}^{n_k}$
+%	\State Set $N_\text{total}$ = $n_k$ 
+%	\iffalse \State Project to the physical space:
+%        $\{\bm{\theta}_k\}_{i=1}^{n_k}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_r}$
+%        \fi
+%     \State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$ 
+%	\State Compute $f(\bm\xi_j)$, $j=1, \ldots, N_\text{total}$.
+%	\State Compute the matrix $\mathbb{C}$ and its eigenvalue decomposition using Algorithm 3
+%	%\State Eigenvalue decomposition, $\mat{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
+%	\State Partition: $\Lambda^{(k)}~=~ 
+%        \begin{bmatrix} \Lambda_1^{(k)} & \\ & \Lambda_2^{(k)} \end{bmatrix}$, 
+%        $W^{(k)}~=~\begin{bmatrix} W_1^{(k)} & W_2^{(k)} \end{bmatrix}$, 
+%        $\Lambda_1^{(k)}\in \mathbb{R}^{d\times\mathcal{S}}$
+%	\Loop
+%		\State Set $k$ = $k$ + 1
+%		\State Draw $n_k =  \beta n_{k-1}$  new random samples 
+%		$\{\bm{\xi}_j\}_{j=1}^{n_k}$  $\beta =??$  $j = n_{k-1}+1,\ldots,n_{k-1}+n_k$. 
+%        \State $N_\text{total}$ = $N_\text{total}$ + $n_k$    
+%	\iffalse	\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$. \fi
+%		 
+%		\State Compute $f(\bm\xi_j)$, $j=n_{k-1}+1, \ldots, n_{k-1}+n_k$.  
+%		\State Construct the matrix, $\mat{C}$ and its eigenvalue decomposition
+%		\State Compute $\delta W_{1,j}^{(k)}$ = 
+%                       \scalebox{1.25}{$\frac{\|(W_{1,j}^{k})^2 - 
+%                       (W_{1,j}^{k-1})^2\|}{\|(W_{1,j}^{k-1})^2\|}$}, 
+%                       $j = 1,\ldots,\mathcal{S}$.
+%		\If {$\max\left(\delta W_{1,j}^{(k)}\right)<\tau$}
+%			\State break
+%		\EndIf
+%	\EndLoop
+%	\State Compute $\alpha_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \Lambda_{1_{i,j}}W_{1_{i,j}}^2$,
+%	$i=1,\ldots,d$.
+%    \EndProcedure
+%  \end{algorithmic}
+%  \label{alg:free1}
+%\end{breakablealgorithm}
+%\bigskip
+%
 \bigskip
@@
 \renewcommand{\algorithmicensure}{\textbf{Output:}}
-  \caption{An iterative gradient-free approach for discovering the active subspace}
-  \begin{algorithmic}[1]
-\Require $\theta_l$, $\theta_u$, $M$, $\tau$. 
-\Ensure $\Lambda$, $W$, $\alpha$ %$\eta$. 
-    \Procedure{Gradient-free}{}
-    \State Set $k$ = 0
-	\State Draw $n_k$ random samples, $\{\bm{\xi}_j\}_{j=1}^{n_k}$
-	\State Set $N_\text{total}$ = $n_k$ 
-	\iffalse \State Project to the physical space:
-        $\{\bm{\theta}_k\}_{i=1}^{n_k}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_r}$
-        \fi
-     \State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$ 
-	\State Compute $f(\bm\xi_j)$, $j=1, \ldots, N_\text{total}$.
-	\State Compute the matrix $\mathbb{C}$ and its eigenvalue decomposition using Algorithm 3
-	%\State Eigenvalue decomposition, $\mat{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
-	\State Partition: $\Lambda^{(k)}~=~ 
-        \begin{bmatrix} \Lambda_1^{(k)} & \\ & \Lambda_2^{(k)} \end{bmatrix}$, 
-        $W^{(k)}~=~\begin{bmatrix} W_1^{(k)} & W_2^{(k)} \end{bmatrix}$, 
-        $\Lambda_1^{(k)}\in \mathbb{R}^{d\times\mathcal{S}}$
-	\Loop
-		\State Set $k$ = $k$ + 1
-		\State Draw $n_k =  \beta n_{k-1}$  new random samples 
-		$\{\bm{\xi}_j\}_{j=1}^{n_k}$  $\beta =??$  $j = n_{k-1}+1,\ldots,n_{k-1}+n_k$. 
-        \State $N_\text{total}$ = $N_\text{total}$ + $n_k$    
-	\iffalse	\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$. \fi
-
-		\State Compute $f(\bm\xi_j)$, $j=n_{k-1}+1, \ldots, n_{k-1}+n_k$.  
-		\State Construct the matrix, $\mat{C}$ and its eigenvalue decomposition
-		\State Compute $\delta W_{1,j}^{(k)}$ = 
-                       \scalebox{1.25}{$\frac{\|(W_{1,j}^{k})^2 - 
-                       (W_{1,j}^{k-1})^2\|}{\|(W_{1,j}^{k-1})^2\|}$}, 
-                       $j = 1,\ldots,\mathcal{S}$.
-		\If {$\max\left(\delta W_{1,j}^{(k)}\right)<\tau$}
-			\State break
-		\EndIf
-	\EndLoop
-	\State Compute $\alpha_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \Lambda_{1_{i,j}}W_{1_{i,j}}^2$,
-	$i=1,\ldots,d$.
-    \EndProcedure
-  \end{algorithmic}
-  \label{alg:free1}
-\end{breakablealgorithm}
-\bigskip
-
-\bigskip
-\begin{breakablealgorithm}
-%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-%\renewcommand{\algorithmicensure}{\textbf{Output:}}
   \caption{Algorithm for constructing the matrix, $\mat{C}$}
@@
 	\State Draw $N$ random samples, $\{\bm{\xi}_j\}_{j=1}^{N}$ 
-	according to $\pi$.
+	according to $\pi_{\bm{\xi}}$.
 	\iffalse \State Project to the physical space:
@@
 	\State Compute $f(\bm\xi_j)$, $j=1, \ldots, N$.
-	\State Draw $M$ random samples, $\{\bm{\nu}_i\}_{i=1}^{M}$
-	according to $\pi$.
+	\State Draw $M$ random samples, $\{\bm{\zeta}_i\}_{i=1}^{M}$
+	according to $\pi_{\bm{\xi}}$.
 	\State Choose an integer $p \leq N$ 
@@
 	\begin{aligned}
-	\Phi_i &= \{ p \text{ nearest points in } \{\bm{\xi}_j\}_{j=1}^{N} \text{ to } \bm{\nu}_i\}\\
+	\Phi_i &= \{ p \text{ nearest points in } \{\bm{\xi}_j\}_{j=1}^{N} \text{ to } \bm{\zeta}_i\}\\
 	\vspace{-2mm}
@@
 	\[
-	\mat{C} \approx \frac{1}{M} \sum_{i=1}^{M} b_i b_i^T = {W}{\Lambda}{W}^T
+	\hat{\mat{C}} \approx \frac{1}{M} \sum_{i=1}^{M} b_i b_i^T = \hat{\mat{W}}\hat{\mat{\Lambda}}\hat{\mat{W}}^\top
 	\]
@@
 
-As mentioned earlier, the model output $f(\bm{\xi}(\bm{\theta}))$ predominantly varies in the active subspace. Hence, 
-$f(\bm{\xi}(\bm{\theta}))$ can be transformed as a function of the active variable as $G(\bm{w}_1^\top\bm{\xi})$ in
-the subspace. The plot of $G$ versus $\bm{w}_1^\top\bm{\xi}$, regarded as the \textit{sufficient summary plot} (SSP) is 
+As mentioned earlier, the model output $f(\bm{\xi})$ predominantly varies in the active subspace. Hence, 
+$f(\bm{\xi})$ can be transformed as a function of the active variable as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
+the subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, regarded as the \textit{sufficient summary plot} (SSP) is 
 compared for the two approaches in Figure~\ref{fig:comp_ssp}.
@@
 uncertain parameters ($\alpha_i$) using
-the components of the dominant eigenvector in the following equation:
-%
-\be
-\alpha_i(n) = \sum\limits_{j=1}^{\mathcal{S}}\Lambda_{1_{i,j}}w_{1_{i,j}}^2,~i = 1,\ldots,N_p
-\ee
-%
-where $n$ is the sample size. The activity scores for the 19 uncertain pre-exponents ($A_i$'s), estimated using
+the components of the dominant eigenvector as shown in Algorithm~\ref{alg:grad}. The activity scores for the 19 
+uncertain pre-exponents ($A_i$'s), estimated using
 the two approaches are plotted in Figure~\ref{fig:comp_as} (left). Additionally, for the purpose of verification, we 

### 2018-08-04 Paper/assa.tex
..

@@
 along these directions. Each component of $\bm{\theta}$
-can be mapped to a corresponding component of $\bm{\xi}$ in $\Omega\in\mathbb{R}^{N_p}$
+is mapped to a corresponding value in $\Omega\in\mathbb{R}^{N_p}$
 ($N_p$: number of uncertain inputs)
-using a linear transformation.  Hence, from this point onwards, we denote $f$ as $f(\bm{\xi})$.
+using a linear transformation.  The set of projected values in $\Omega$ is denoted as $\bm{\xi}$.
+Hence, from this point onwards, we denote $f$ as $f(\bm{\xi})$.
 Mathematically, for a given $f$, and a set of inputs, $\bm{\xi}$,

### 2018-08-05 Paper/assa.tex
..

@@
 ($N_p$: number of uncertain inputs)
-using a linear transformation.  The set of projected values in $\Omega$ is denoted as $\bm{\xi}$.
+using a linear transformation.  The set of canonical random variables, corresponding
+to $\theta_i$'s, in $\Omega$ is referred to as $\bm{\xi}$.
 Hence, from this point onwards, we denote $f$ as $f(\bm{\xi})$.
@@
 the active subspace, and $\mat{\Lambda}_1$ is the corresponding set of
-eigenvalues.
-
-\alennote{The rest of this paragraph needs to be rewritten, as it is hard to 
-see what is going on; what is $\theta$? I suggest, mentioning $\vec\theta$ 
-and $\vec\xi$ earlier, and from then on make everything function of $\vec\xi$.
-Explain the remaining ideas more clearly.}
-The active subspace is used to construct the so-called active
-variables, $\mat{W}_1^\top \bm{\xi}$ and $f(\bm{\xi}(\theta))$ is transformed as
-$G(\mat{W}_1^\top \bm{\xi})$. In practice, the set of samples, $\bm{\xi}_i$'s
-are projected to the physical space as $\bm{\theta}_i$'s at which the model
-output, $f(\bm{\xi}(\theta))$ is computed.  The matrix, $\mat{C}$ is typically
+eigenvalues. Once the active subspace is computed, dimension reduction is
+accomplished by transforming the set of canonical random variables, $\bm{\xi}$
+into a reduced set, $\mat{W}_1^\top \bm{\xi}$, referred to as the set of active
+variables. The number of active variables is thus equal to the number
+of dominant eigenvectors. The model output in the original space, $f(\bm{\xi}(\theta))$,
+is expressed in terms of the active variables as $G(\mat{W}_1^\top \bm{\xi})$.
+In practice, the matrix, $\mat{C}$ is typically
 approximated using pseudo-random sampling techniques such as Monte Carlo and
@@
  %
+Clearly, the amount of computational effort associated with constructing the matrix,
+$\hat{\mat{C}}$ scales with the number of samples, $N$. Hence, an iterative
+computational approach is adopted in this work to ensure that $N$ is sufficiently
+large so that the dominant eigenvectors in $\hat{\mat{W}}_1$ are approximated
+with reasonable accuracy as discussed later in Section~\ref{sec:method}.
+
 %%It has been shown that the accuracy of approximated dominant eigenspace,
@@
 the classical 
-framework of variance based sensitivity analysis~\cite{MMM} assumes the input
+framework of variance based sensitivity analysis~\cite{Sobol:2001, Saltelli:2010} assumes the input
 parameters are statistically independent. While extensions to the cases 
@@
 the total contribution of the input, $\theta_i$ to the variance of the
-output.\alennote{references} Mathematically, this can be expressed as follows:
+output~\cite{Sobol:2001}. Mathematically, this can be expressed as follows:
 %
@@
 where $\bm{\theta}_{\sim i}$ is a vector of all but $i^\text{th}$ uncertain
-input, and $\V$ denotes the variance (~\alennote{Explain this more clearly.}).
+input, and $\V$ denotes the variance. The total Sobol' index thus accounts
+for the contribution of a given input to the variability in the output by itself
+as well as due to its interaction or coupling with other inputs. 
 Determining accurate estimates of $T_i(f)$ typically involves tens of

### 2018-08-05 Paper/method.tex
..

@@
 obtained using a much larger sample size, $N$~=~1000. Right: Relative L-2 norm of the difference between
-individual components of the first five eigenvectors of $\mat{C}$, constructed using $N$~=~1000 samples,
-and using a smaller sample size, $n$ ranging from 20 to 240.} 
+the square of individual components of the first five eigenvectors of $\mat{C}$, constructed using 
+$N$~=~1000 samples ($w_{i,N}$), and using a smaller sample size, $n$ ranging from 20 to 240 ($w_{i,n}$).} 
 \label{fig:eig_comp}
@@
 ($i$ = 1,2) is found to range from $\mathcal{O}(10^{-2} - 10^{-1})$ and is relatively much smaller than 
-the higher-order eigenvalues. In this case, since $\left(\frac{\lambda_1}{\lambda_2}\right)\gg 1$, we have
+the higher-order eigenvectors. In this case, since $\left(\frac{\lambda_1}{\lambda_2}\right)\gg 1$, we have
 a 1-dimensional active subspace. Hence, the first eigenvector sufficiently captures the uncertainty in

### 2018-08-05 Paper/method.tex
..

@@
 domain and the corresponding active subspace is computed. In Figure~\ref{fig:eig_comp} (left), we illustrate
-the comparison of the resulting eigenvalue spectrum ($\lambda_1 - \lambda_{19}$) with the same using a much 
+the comparison of the resulting normalized eigenvalue spectrum ($\lambda_1 - \lambda_{19}$) with the same using a much 
 smaller set of samples, $n$~=~$\{20,40,80,120\}$. Additionally, in Figure~\ref{fig:eig_comp} (right), we compare
@@
    \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1_5}
-\caption{Left: A comparison of the eigenvalue spectrum using $n$ = $\{20,40,80,120\}$ samples with that
+\caption{Left: A comparison of the normalized eigenvalue spectrum using $n$ = $\{20,40,80,120\}$ samples with that
 obtained using a much larger sample size, $N$~=~1000. Right: Relative L-2 norm of the difference between

### 2018-08-05 Paper/assa.tex
...

@@
 is expressed in terms of the active variables as $G(\mat{W}_1^\top \bm{\xi})$.
+\alennote{This is too fast, explain a bit more, and refer to 
+Paul Constantine's book. I think we should follow the explanations
+in Chapter 4 of Paul's book; see equations (4.1)--(4.3) and algorithm 4.1
+for surrogate model construction on active subspace; these should be
+mentioned here.}
 In practice, the matrix, $\mat{C}$ is typically

### 2018-08-05 Paper/method.tex
...

@@
 
-In this section, we present the computational strategy used for computing the active subspace in the case of
-H$_2$/O$_2$ reaction kinetics with uncertain pre-exponent of individual reaction rates provided in
-Table~\ref{tab:kinetics}. Specifically, we employ two strategies: One that involves computation of model
-gradients in order to construct the matrix (gradient-based), $\hat{\mat{C}}$ in~\eqref{eq:chat}, and the other that involves a local 
-linear approximation of the model output to circumvent the computational cost associated with gradient
-computation (gradient-free). The gradient-based approach is based on Algorithm 1.1, and the gradient-free approach is
-based on Algorithm 1.2 in~\cite{Constantine:2015}. However, in this work, we have implemented the
-two strategies in an iterative manner and have demonstrated potential for additional savings in both cases.
-Additionally, for the present application, our findings based on the two strategies are observed to be consistent.
-Since, the gradient-free approach is relatively more efficient, this observation encourages its use for high-dimensional
-applications as discussed later in Section~\ref{sec:app}. Note however, that the model output is required to be
-differentiable throughout the input domain in both cases.
+In this section, we present the computational strategy used for computing the
+active subspace in the case of H$_2$/O$_2$ reaction kinetics with uncertain
+pre-exponent of individual reaction rates provided in Table~\ref{tab:kinetics}.
+Specifically, we employ two strategies: 
+A gradient-based strategy that involves computation of model
+gradients in order to construct the matrix $\hat{\mat{C}}$
+in~\eqref{eq:chat}, and a gradient-free approach 
+that involves a (\alennote{should we say 
+regression-based?}) local linear approximation of
+the model output to circumvent the computational cost associated with gradient
+computation. The gradient-based approach is based on Algorithm
+1.1, and the gradient-free approach is based on Algorithm 1.2
+in~\cite{Constantine:2015}. In this work, we implement the two
+strategies in an iterative manner and demonstrate potential for
+additional savings in both cases.  For the present application,
+our findings based on the two strategies are observed to be consistent.  
+Since,
+the gradient-free approach is more efficient, this observation
+encourages its use for high-dimensional applications as discussed later in
+Section~\ref{sec:app}. 
+%Note however, that the model output is required to be
+%differentiable throughout the input domain in both cases.
 
@@
 
-As discussed earlier, in the gradient-based approach, the elements of $\hat{\mat{C}}$ are estimated using model gradients.
- In situations where the gradient is not available analytically, one could consider using
-finite difference and other relatively more efficient techniques such as automatic differentiation~\cite{Kiparissides:2009} 
-and adjoint-based gradient computation~\cite{Jameson:1988,Gunzburger:2003,Borzi:2011,Alexanderian:2017}.
-Model evaluations at neighbouring points are required if finite difference is used. Hence, for 
-$N$ samples in a $d$-dimensional parameter space, $N(d+1)$ model evaluations are needed.
+As discussed earlier, in the gradient-based approach, the elements of
+$\hat{\mat{C}}$ are estimated using model gradients.  In situations where the
+gradient is not available analytically, one could consider using finite
+difference and other relatively more efficient techniques such as automatic
+differentiation~\cite{Kiparissides:2009} and adjoint-based gradient
+computation~\cite{Jameson:1988,Gunzburger:2003,Borzi:2011,Alexanderian:2017}.
+Model evaluations at neighbouring points are required if finite difference is
+used. Hence, for $N$ samples in a $d$-dimensional parameter space, $N(d+1)$
+model evaluations are needed.
 
-We begin by evaluating the gradient of the model output, $\nabla_{\bm{\xi}}f$, at an initial set of $n_1$ samples. 
-For this purpose, $\bm{\xi}_i$'s are projected to the physical space as $\bm{\theta}_i$'s. Using the gradient
-evaluations, the matrix, $\hat{\mat{C}}$ is computed. Eigenvalue decomposition of $\hat{\mat{C}}$ yields an initial
-estimate of the dominant eigenspace, $\hat{\mat{W}}_1$ and corresponding set of eigenvalues, $\hat{\mat{\Lambda}}_1$.
-At each subsequent iteration, model evaluations are generated at a new set of $n_s$ samples. The new set
-of gradient evaluations are augmented with the available set to re-construct $\hat{\mat{C}}$ followed by its eigenvalue
-decomposition. The relative
-change in the norm of the difference in squared value of individual components of the dominant eigenvectors between subsequent 
-iterations is evaluated. The process is terminated and the resulting eigenspace is considered to be converged once
-highest relative change at iteration $k$, $\delta \hat{\mat{W}}_{1,j}^{(k)}$, is smaller a given tolerance, $\tau$. 
-A regression fit to
- $G(\hat{\mat{W}}_1^\top\bm{\xi})$ is used as a surrogate to characterize and quantify the uncertainty in the model
- output. Moreover, the components of the dominant eigenvectors are used to compute the activity scores,
- $\bm{\nu}_r(f)$, which provide an insight into the relative importance of the uncertain inputs. Note that the
- index, $r$, corresponds to the number of dominant eigenvectors in $\hat{\mat{W}}_1$. The sequence of 
- steps for the gradient-based approach are provided in Algorithm~\ref{alg:grad}.
+We begin by evaluating the gradient of the model output, $\nabla_{\bm{\xi}}f$,
+at an initial set of $n_1$ samples.  For this purpose, $\bm{\xi}_i$'s are
+projected to the physical space as $\bm{\theta}_i$'s. 
+\alennote{The mention of $\bm{\theta}$ is redundant. 
+Say what $\bm{\xi}_i$, $i = 1, \ldots, n_1$ are before using them.} 
+Using the gradient
+evaluations, the matrix, $\hat{\mat{C}}$ is computed. Eigenvalue decomposition
+of $\hat{\mat{C}}$ yields an initial estimate of the dominant eigenspace,
+$\hat{\mat{W}}_1$ and corresponding set of eigenvalues,
+$\hat{\mat{\Lambda}}_1$.  At each subsequent iteration, model evaluations are
+generated at a new set of $n_s$ samples. The new set of gradient evaluations
+are augmented with the available set to re-construct $\hat{\mat{C}}$ followed
+by its eigenvalue decomposition. The relative change in the norm of the
+difference in squared value of individual components of the dominant
+eigenvectors between subsequent iterations is evaluated. The process is
+terminated and the resulting eigenspace is considered to be converged once
+highest relative change at iteration $k$, $\delta \hat{\mat{W}}_{1,j}^{(k)}$,
+is smaller than a given tolerance, $\tau$.  A regression fit to
+$G(\hat{\mat{W}}_1^\top\bm{\xi})$ is used as a surrogate to characterize and
+quantify the uncertainty in the model output. Moreover, the components of the
+dominant eigenvectors are used to compute the activity scores, $\bm{\nu}_r(f)$,
+which provide insight into the relative importance of the uncertain inputs.
+Note that the index, $r$, corresponds to the number of dominant eigenvectors in
+$\hat{\mat{W}}_1$. The sequence of steps for the gradient-based approach are
+provided in Algorithm~\ref{alg:grad}.
 
@@
 
-In order to motivate the iterative strategy used in this work, we implement Algorithm~\ref{alg:grad} to the
-19-dimensional H$_2$/O$_2$ reaction kinetics problem with uncertain $A_i$'s. For the purpose of
-verification, $\hat{\mat{C}}$ is initially constructed using a large set of samples ($N$~=~1000) in the input
-domain and the corresponding active subspace is computed. In Figure~\ref{fig:eig_comp} (left), we illustrate
-the comparison of the resulting normalized eigenvalue spectrum ($\lambda_1 - \lambda_{19}$) with the same using a much 
-smaller set of samples, $n$~=~$\{20,40,80,120\}$. Additionally, in Figure~\ref{fig:eig_comp} (right), we compare
-relative L-2 norm of the difference between the first five eigenvectors of $\hat{\mat{C}}$.
+To motivate our iterative strategy, we implement
+Algorithm~\ref{alg:grad} to the 19-dimensional H$_2$/O$_2$ reaction kinetics
+problem with uncertain $A_i$'s. For the purpose of verification,
+$\hat{\mat{C}}$ is initially constructed using a large set of samples
+($N$~=~1000) in the input domain and the corresponding active subspace is
+computed. In Figure~\ref{fig:eig_comp}~(left), we illustrate the comparison of
+the resulting normalized eigenvalue spectrum ($\lambda_1, \ldots, 
+\lambda_{19}$) with the same using a much 
+smaller set of samples, $n$~=~$\{20,40,80,120\}$.
+Additionally, in Figure~\ref{fig:eig_comp} (right), we compare relative $L_2$
+norm of the difference between the first five eigenvectors of $\hat{\mat{C}}$.
+\alennote{This last sentence is unclear.}
+
 %
@@
 \caption{Left: A comparison of the normalized eigenvalue spectrum using $n$ = $\{20,40,80,120\}$ samples with that
-obtained using a much larger sample size, $N$~=~1000. Right: Relative L-2 norm of the difference between
+obtained using a much larger sample size, $N$~=~1000. Right: Relative $L_2$ 
+norm of the difference between
 the square of individual components of the first five eigenvectors of $\mat{C}$, constructed using 
@@
 %
-Interestingly, we observe that the dominant eigenvalues ($\lambda_1 - \lambda_4$) are approximated 
+We observe that the dominant eigenvalues, $\lambda_1, \ldots, \lambda_4$, 
+are approximated 
 reasonably well with just 20 samples. As expected, the accuracy of higher-order eigenvalues is observed
@@
 ($i$ = 1,2) is found to range from $\mathcal{O}(10^{-2} - 10^{-1})$ and is relatively much smaller than 
-the higher-order eigenvectors. In this case, since $\left(\frac{\lambda_1}{\lambda_2}\right)\gg 1$, we have
-a 1-dimensional active subspace. Hence, the first eigenvector sufficiently captures the uncertainty in
+the higher-order eigenvectors. In this case, since 
+$\left(\frac{\lambda_1}{\lambda_2}\right)\gg 1$, we expect 
+a 1-dimensional active subspace. Hence, the first eigenvector 
+sufficiently captures the uncertainty in
 the  model output. Therefore, a sample size of 20 is adequate for computing the active subspace in this
@@
 However, estimating the gradient of the model output imposes a significant computational burden.
-Hence, we explore the applicability of a so-called gradient-free approach as mentioned earlier, in the
-following section.
+Hence, we next explore the applicability of a so-called gradient-free approach
+as mentioned earlier.
 
@@
 
-As mentioned earlier, the gradient-free approach does not involve computation of the gradient of the model
-output. Instead, a least-squares fit is performed to the local linear regression-based approximation
-to the model output. While the gradient-free approach offers a significant potential for additional 
-computational savings, its ability to reasonably approximate the active subspace 
-has not been investigated for a wide variety of problems~\cite{Constantine:2015}. Nevertheless, we are motivated by the
-underlying potential of this approach in reducing the computational effort associated with the
-active subspace computation. Moreover, this effort could help advance this approach and further motivate
-its usage for a wide variety of high-dimensional applications in the future.
+\alennote{I don't see the point of this paragraph; part of it that 
+explains the gradient-free approach is unclear, and the part 
+that seems to motivate its use is something to mention in the introduction.
+Just say what the gradient-free approach is about, and say that we 
+follow an iterative strategies as before.}
+As mentioned earlier, the gradient-free approach does not involve computation
+of the gradient of the model output. Instead, a least-squares fit is performed
+to the local linear regression-based approximation to the model output. While
+the gradient-free approach offers a significant potential for additional
+computational savings, its ability to reasonably approximate the active
+subspace has not been investigated for a wide variety of
+problems~\cite{Constantine:2015}. We are motivated by the
+potential of this approach in reducing the computational effort
+associated with the active subspace computation. Moreover, this effort could
+help advance this approach and further motivate its usage for a wide variety of
+high-dimensional applications in the future.
 
@@
 active subspace; consistent with the fact the it is 1-dimensional.
+\alennote{The last sentence is incomplete; also clarify that ``it'' refers to 
+the active subspace.}
 
-We extend the comparative assessment of the two approaches by estimating the activity score for individual
-uncertain parameters ($\alpha_i$) using
-the components of the dominant eigenvector as shown in Algorithm~\ref{alg:grad}. The activity scores for the 19 
-uncertain pre-exponents ($A_i$'s), estimated using
-the two approaches are plotted in Figure~\ref{fig:comp_as} (left). Additionally, for the purpose of verification, we 
-compare these estimates with those obtained using the DGSMs~\cite{Vohra:2018} in 
+We complement the comparative assessment of the two approaches by estimating
+the activity scores for individual uncertain parameters ($\alpha_i$) using the
+components of the dominant eigenvector as shown in Algorithm~\ref{alg:grad}.
+The activity scores (\alennote{relative activity scores?}) for the 19 uncertain pre-exponents ($A_i$'s), estimated
+using the gradient-based and gradient-free approaches are plotted in
+Figure~\ref{fig:comp_as} (left).  Additionally, for the purpose of
+verification, we compare these estimates with those obtained using the
+corresponding DGSMs, as reported in~\cite{Vohra:2018}, in
 Figure~\ref{fig:comp_as} (right).
+
 %
@@
 
-The above comparisons establish consistency between the two approaches. Note that the computational
-effort in the case of gradient-free approach is significantly smaller. Hence, we are encouraged to use the
-gradient-free approach for the 33-dimensional application in the next section.
+The above comparisons establish consistency between gradient-based and
+gradient-free approaches. Due to its reduced computational overhead, the
+gradient-free approach is our method of choice for the 33-dimensional problem
+in the next section.

### 2018-08-05 Paper/method.tex
sec 3, edits and comments ...

@@
 in~\eqref{eq:chat}, and a gradient-free approach 
-that involves a (\alennote{should we say 
-regression-based?}) local linear approximation of
+that involves a 
+local linear approximation of
 the model output to circumvent the computational cost associated with gradient
-computation. The gradient-based approach is based on Algorithm
+computation. 
+(\alennote{should we say regression-based local linear approximation? 
+When I hear local linear approximation I think of Taylor expansion, and
+probably many readers will do the same; so saying gradient-free
+local linear approximation might be confusing to some people.}) 
+The gradient-based approach is based on Algorithm
 1.1, and the gradient-free approach is based on Algorithm 1.2
@@
 subspace has not been investigated for a wide variety of
-problems~\cite{Constantine:2015}. We are motivated by the
+problems~\cite{Constantine:2015}.~\alennote{I would not say this. 
+If Paul (or a student of his) 
+reviews this article, he will probably disagree. Remember that 
+book was from several years ago. Instead, say that we 
+demonstrate its utility for efficient response surface construction in 
+chemical kinetics modeling.} We are motivated by the
 potential of this approach in reducing the computational effort

### 2018-08-05 Paper/method.tex
..

--- 2018-08-05 Paper/method.tex
@@
 we provide a comparison of the components of the dominant eigenvector, obtained using the two approaches.
-Additionally, the quantity, $\delta w_{1,j}^{(r)}$ is also plotted in
+Additionally, the quantity, $\delta \hat{\mat{W}}_{1,j}^{(r)}$ is also plotted in
 Figure~\ref{fig:comp} (right) to illustrate convergence characteristics for the gradient-free approach.
@@
 \caption{Left: An illustrative comparison of individual components of the converged dominant eigenvector obtained
-using the two approaches discussed in~\ref{sub:grad} and~\ref{sub:gradfree}. Right: The quantity,  $\delta w_{1,j}^{(r)}$
+using the two approaches discussed in~\ref{sub:grad} and~\ref{sub:gradfree}. Right: The quantity,  $\delta \hat{\mat{W}}_{1,j}^{(r)}$
 is plotted for successive iterations to illustrate the convergence behavior.}
@@
 Individual components of the dominant eigenvector from the two approaches are observed to be in excellent
-agreement with each other. Moreover, as expected,  $\delta w_{1,j}^{(r)}$ is observed to decrease with iterations, and
+agreement with each other. Moreover, as expected,  $\delta \hat{\mat{W}}_{1,j}^{(r)}$ is observed to decrease with iterations, and
 the eigenvector is considered to have converged if its value is found to be smaller than 0.1. Note that only 30 samples

### 2018-08-05 Paper/app.tex
..

@@
 
-In Figure~\ref{fig:hd}, we illustrate the resulting eigenvalue spectrum and the sufficient summary 
-plot.
+In Figure~\ref{fig:hd}, we illustrate the resulting eigenvalue spectrum and the SSP.
 %
@@
    \includegraphics[width=0.45\textwidth]{./Figures/ssp_33D}
-\caption{Left: Eigenvalue spectrum of $\hat{\mat{C}}$. Right: Sufficient summary plot indicating the
-existence of a potential 1-dimensional active subspace.} 
+\caption{Left: Eigenvalue spectrum of $\hat{\mat{C}}$. Right: SSP of the computed active subspace, and the
+corresponding 1-D linear fit.} 
 \label{fig:hd}
@@
 The second eigenvalue is observed to be roughly four orders of magnitude smaller than the first 
-indicating the existence of a 1-dimensional active subspace. This is further confirmed by the sufficient
-summary plot which shows that the variability in $G$ can be approximated using a 1-dimensional regression
-fit which could be used as a surrogate. We will verify the accuracy of the surrogate later in~\ref{sub:verify}.
+indicating the existence of a 1-dimensional active subspace. This is further confirmed by the SSP
+ which shows that the variability in $G$ can be approximated using a 1-dimensional regression
+fit ($\tilde{G}$) which could be used as a surrogate. A linear, quadratic, and a cubic fit to the SSP was performed.
+The linear fit illustrated in Figure~\ref{fig:hd} (right) was found to be the most accurate. In~\ref{sub:verify},
+we investigate the accuracy of the 1D surrogate based on the linear fit by estimating the relative L-2
+norm of its discrepancy with model evaluations at an independent set of samples, and by comparing the
+respective probability distributions.
 
@@
 
-The 1-dimensional surrogate ($\mathcal{Y}$) is investigated for its accuracy as well as the ability to capture the 
+The 1-dimensional surrogate ($\tilde{G}$) is investigated for its accuracy as well as the ability to capture the 
 uncertainty in the
@@
 \be
-\varepsilon_d = \frac{\|f(\bm{\theta}) - \mathcal{Y}(\bm{\xi})\|_2}{\|f(\bm{\theta})\|_2}
+\varepsilon_d = \frac{\|G(\bm{\xi}) - \tilde{G}(\bm{\xi})\|_2}{\|G(\bm{\xi})\|_2}
 \ee
 %
-The relative norm, $\varepsilon_d$ was estimated to be 8.92$\times10^{-2}$ based on a validation test data
+The relative norm, $\varepsilon_d$ was estimated to be 8.50$\times10^{-2}$ based on a validation test data
 comprising 1000 model evaluations. Hence, in the norm-sense, it can be said that the 1-dimensional surrogate is 

### 2018-08-05 Paper/assa.tex
boltzmann ...

@@
 either uniformly distributed or
-distributed according to a Boltzmann probability distribution~\alennote{Explain what this
-is, and give a reference}.
+distributed according to a Boltzmann probability distribution.
+Note that a measure $\mu$
+on $\R$ is referred to as a Boltzmann measure if it is 
+absolutely continuous with respect to the Lebesgue measure  
+and admits a density  of the form $\pi(x) = C \exp\{-V(x)\}$,
+where $V$ is a continuous function and $C$ a normalization 
+constant~\cite{Lamboni:2013}.
 
@@
 \end{equation}
-where for each $i$, $C_i$ is an appropriate \emph{Poincar\'{e}} constant.
+where for each $i$, $C_i$ is an appropriate \emph{Poincar\'{e}} constant
+that depends on the distribution of $\xi_i$.
 This provides a strong theoretical basis for using DGSMs to identify 

### 2018-08-06 Paper/app.tex
..

@@
 \caption{Left: An illustrative comparison of individual components of the converged dominant eigenvector obtained
-using the two approaches discussed in~\ref{sub:grad} and~\ref{sub:gradfree}. Right: The quantity,  $\delta \hat{\mat{W}}_{1,j}^{(i)}$
+using the two approaches discussed in~\ref{sub:grad} and~\ref{sub:gradfree}. Right: The quantity,  
+$\delta \hat{\mat{W}}_{1,j}^{(i)}$
 is plotted for successive iterations to illustrate the convergence behavior.}
@@
 %
-The dominant eigenvector is observed to converge with only 50 samples in the 33-dimensional space which is remarkably efficient.
+The dominant eigenvector is observed to converge with only 50 samples in the 33-dimensional space which
+is remarkably efficient.
 
@@
 %
-The relative norm, $\varepsilon_d$ was estimated to be 8.50$\times10^{-2}$ based on a validation test data
-comprising 1000 model evaluations. Hence, in the norm-sense, it can be said that the 1-dimensional surrogate is 
+The relative norm, $\varepsilon_d$ was estimated to be 8.50$\times10^{-2}$ based on a cross-validation test data
+comprising 10$^{4}$ model evaluations. Hence, in the norm-sense, it can be said that the 1-dimensional surrogate is 
 reasonably accurate.
@@
 Secondly, we verify the accuracy of the surrogate in a probabilistic setting. In particular, we compare a histogram
-plot of the model evaluations for ignition delay in the validation set with a PDF of the same using the 1-dimensional
-surrogate as shown in Figure~\ref{fig:pdf_33D}.
+plot of the model evaluations (in the cross-validation set) with a PDF evaluated on the same set of samples
+using the surrogate, $\tilde{G}$ in Figure~\ref{fig:pdf_33D}. The PDF is evaluated using surrogates constructed for
+a 1-dimensional as well as a 2-dimensional active subspace. 
 %
@@
 %
-In the above comparison, it is observed that the peak of the PDF generated using the 1-dimensional surrogate in the
-active subspace is consistent with the modal estimate based on the histogram. Moreover, the spread in the ignition
- delay is reasonably captured by the PDF. Hence, the surrogate is also verified in a probabilistic sense.
+In the above comparison, it is observed that the two PDFs are almost identical, thereby confirming a 1-dimensional
+active subspace. The modal estimate from either PDF is found to be in agreement with that
+based on the histogram plot. Moreover, the spread in the ignition
+delay is reasonably captured by the PDFs.
+
+Based on these observations, it can be said that a 1-dimensional surrogate can be used to reasonably approximate the
+uncertainty in the ignition delay due to uncertain rate-controlling parameters used in the H$_2$/O$_2$ reaction
+mechanism.

### 2018-08-06 Paper/assa.tex
..

--- 2018-08-05 Paper/assa.tex
@@
 distribution of $\bm{\xi}$. Note that the function, $f$ is considered to be continuous 
-and differentiable in the physical space, $\bm{\theta}$. %Hence, it is possible that a given $f(\bm{\xi})$ might not admit an active
+and differentiable in the physical space, $\bm{\theta}$. 
+%Hence, it is possible that a given $f(\bm{\xi})$ might not admit an active
 %subspace. However, it is of remarkable interest to investigate if one exists
@@
 variables. The number of active variables is thus equal to the number
-of dominant eigenvectors. The model output in the original space, $f(\bm{\xi}(\theta))$,
+of dominant eigenvectors. The model output in the original space, $f(\bm{\xi})$,
 is expressed in terms of the active variables as $G(\mat{W}_1^\top \bm{\xi})$.
+Hence, $f$ is essentially evaluated at the projection of $\bm{\xi}$ on the
+column space of $\mat{W}_1$. We could thus confine UQ to the inputs in the
+active subspace whose dimension is typically much smaller (in applications that
+admit such a subspace) than the dimension of the original input space. To 
+expedite UQ, one could easily fit a regression-surface to $G$ using the 
+following sequence of steps outlined in~\cite{Constantine:2015} (see chapter 4).
+\begin{enumerate}
+\item Consider a set of input-output pairs, $\{\bm{\xi}_i, f(\bm{\xi}_i)\}$.
+\item For each $\bm{\xi}_i$, compute $\mat{W}_1^\top\bm{\xi}_i$. Note that
+ $G(\mat{W}_1^\top \bm{\xi})$ $\approx$ $f(\bm{\xi}_i)$.
+\item Fit a regression surface, $\tilde{G}(\mat{W}_1^\top\bm{\xi}_i)$ $\approx$ 
+$G(\mat{W}_1^\top \bm{\xi})$.
+\item Overall approximation, $f(\bm{\xi})$ $\approx$ $\tilde{G}(\mat{W}_1^\top\bm{\xi})$.
+\end{enumerate}
+
 In practice, the matrix, $\mat{C}$ is typically
@@
 an interesting link between the activity scores and total Sobol' indices.
-Specifically, by computing the activity scores, we can indetify unimportant
+Specifically, by computing the activity scores, we can identify unimportant
 parameters.  

### 2018-08-06 Paper/method.tex
.
..

--- 2018-08-05 Paper/method.tex
@@
 in~\eqref{eq:chat}, and a gradient-free approach 
-that involves a (\alennote{should we say 
-regression-based?}) local linear approximation of
+that involves a regression-based local linear approximation of
 the model output to circumvent the computational cost associated with gradient
-computation. The gradient-based approach is based on Algorithm
+computation. 
+The gradient-based approach is based on Algorithm
 1.1, and the gradient-free approach is based on Algorithm 1.2
@@
 computation~\cite{Jameson:1988,Gunzburger:2003,Borzi:2011,Alexanderian:2017}.
-Model evaluations at neighbouring points are required if finite difference is
+Model evaluations at neighboring points are required if finite difference is
 used. Hence, for $N$ samples in a $d$-dimensional parameter space, $N(d+1)$
@@
 We begin by evaluating the gradient of the model output, $\nabla_{\bm{\xi}}f$,
-at an initial set of $n_1$ samples.  For this purpose, $\bm{\xi}_i$'s are
-projected to the physical space as $\bm{\theta}_i$'s. 
-\alennote{The mention of $\bm{\theta}$ is redundant. 
-Say what $\bm{\xi}_i$, $i = 1, \ldots, n_1$ are before using them.} 
+at an initial set of $n_0$ samples denoted by $\bm{\xi}_i$, $i$ = $1,\ldots,n_0$.
 Using the gradient
@@
 of $\hat{\mat{C}}$ yields an initial estimate of the dominant eigenspace,
-$\hat{\mat{W}}_1$ and corresponding set of eigenvalues,
-$\hat{\mat{\Lambda}}_1$.  At each subsequent iteration, model evaluations are
-generated at a new set of $n_s$ samples. The new set of gradient evaluations
+$\hat{\mat{W}}_1$ and a diagonal matrix, 
+$\hat{\mat{\Lambda}}_1$ that comprises the eigenvalues along the diagonal.
+ At each subsequent iteration, model evaluations are
+generated at a new set of $n_k$ samples. The new set of gradient evaluations
 are augmented with the available set to re-construct $\hat{\mat{C}}$ followed
@@
 terminated and the resulting eigenspace is considered to be converged once
-highest relative change at iteration $k$, $\delta \hat{\mat{W}}_{1,j}^{(k)}$,
+the maximum relative change at iteration $k$, $\delta \hat{\mat{W}}_{1,j}^{(k)}$
+($j$ is used as an index for the eigenvectors),
 is smaller than a given tolerance, $\tau$.  A regression fit to
@@
 smaller set of samples, $n$~=~$\{20,40,80,120\}$.
-Additionally, in Figure~\ref{fig:eig_comp} (right), we compare relative $L_2$
-norm of the difference between the first five eigenvectors of $\hat{\mat{C}}$.
-\alennote{This last sentence is unclear.}
+Additionally, we compare relative L-2
+norm of the difference between the squared value of corresponding components
+of the eigenvectors, obtained using the large set, $N$, and the smaller sets, $n$.
+In  Figure~\ref{fig:eig_comp} (right), we plot the relative L-2 error norm for the
+first five eigenvectors.
 
@@
 
-\alennote{I don't see the point of this paragraph; part of it that 
-explains the gradient-free approach is unclear, and the part 
-that seems to motivate its use is something to mention in the introduction.
-Just say what the gradient-free approach is about, and say that we 
-follow an iterative strategies as before.}
-As mentioned earlier, the gradient-free approach does not involve computation
-of the gradient of the model output. Instead, a least-squares fit is performed
-to the local linear regression-based approximation to the model output. While
-the gradient-free approach offers a significant potential for additional
-computational savings, its ability to reasonably approximate the active
-subspace has not been investigated for a wide variety of
-problems~\cite{Constantine:2015}. We are motivated by the
-potential of this approach in reducing the computational effort
-associated with the active subspace computation. Moreover, this effort could
-help advance this approach and further motivate its usage for a wide variety of
-high-dimensional applications in the future.
+The gradient-free approach exploits a regression-based local linear
+approximation to the model output for estimating the gradients, $\bm{g}^i$
+in Algorithm~\ref{alg:grad}. Once again, we use an iterative strategy for
+computing the active subspace. The utility of this approach is demonstrated
+by constructing an efficient response surface for the 19-dimensional 
+H$_2$/O$_2$ reaction kinetics problem considered in~\ref{sub:grad}.
+Moreover, we compare our findings with
+those obtained using the gradient-based approach to assess consistency
+between the two approaches. Motivated by our findings, we compute the active 
+subspace for a 33-dimensional H$_2$/O$_2$ reaction kinetics application using the gradient-free
+approach in Section~\ref{sec:app}.
 
-First, we implement the gradient-free approach to the 19-dimensional problem and compare our findings with
-those obtained using the gradient-based approach. Later, in
-Section~\ref{sec:app}, we compute the active subspace for a 33-dimensional application using the gradient-free
-approach. Once again, the strategy is implemented in an iterative manner as shown in~\ref{sub:grad}. 
-It is observed that the results from the gradient-free approach are consistent with the gradient-based approach
-as discussed further below.
-
-In the gradient-free approach, the model output is first computed at an initial set of $n_1$ samples. An independent
-set of $M$ samples is also drawn according to the joint probability distribution of the inputs. For each sample
-in $M$, $p$ nearest neighbours in the $n_1$ set are identified. A least-squares fit to model evaluations at
-the $p$ samples is performed. The slope vector ($\bm{b}_i$) of the regression-fit is stored separately and the process is
-repeated for each sample in $M$. The symmetric and positive semi-definite matrix, $\hat{\mat{C}}$ is hence constructed
-using the slope vectors as follows:
+In the gradient-free approach, the model output is first computed at an initial set of $n_0$ samples. 
+An independent set of $M$ samples is also drawn according to the joint probability distribution of the inputs.
+For each sample in $M$, $p$ nearest neighbors in the $n_0$ set are identified. A least-squares fit to model
+evaluations at the $p$ samples is performed. The slope vector ($\bm{b}_i$) of the regression-fit is stored
+separately and the process is repeated for each sample in $M$. The symmetric and positive semi-definite matrix,
+$\hat{\mat{C}}$ is constructed using the slope vectors as follows:
 %
@@
 The active subspace computation involves the same sequence of steps as outlined previously for the gradient-based 
-approach in Algorithm~\ref{alg:grad}. The procedure for constructing $\hat{\mat{C}}$ using local linear approximation
+approach in Algorithm~\ref{alg:grad}. The procedure for constructing $\hat{\mat{C}}$ using a regression-based
+local linear approximation
 of the model output is provided in Algorithm~\ref{alg:free2}.
@@
 delay based on the considered prior marginals for $A_i$'s is observed to be approximately linear in the
-active subspace; consistent with the fact the it is 1-dimensional.
-\alennote{The last sentence is incomplete; also clarify that ``it'' refers to 
-the active subspace.}
+active subspace. This observation indicates the the active subspace is 1-dimensional in this case.
 
 We complement the comparative assessment of the two approaches by estimating
-the activity scores for individual uncertain parameters ($\alpha_i$) using the
-components of the dominant eigenvector as shown in Algorithm~\ref{alg:grad}.
-The activity scores (\alennote{relative activity scores?}) for the 19 uncertain pre-exponents ($A_i$'s), estimated
+the activity scores for individual uncertain parameters ($\nu_{i,r}$; $r$=1 since the
+active subspace is 1-dimensional)
+using the components of the dominant eigenvector as shown in Algorithm~\ref{alg:grad}.
+The activity scores for the 19 uncertain pre-exponents ($A_i$'s), estimated
 using the gradient-based and gradient-free approaches are plotted in
@@
   \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}
-\caption{Left: A bar-graph of activity scores ($\alpha_i$'s) for the 19 uncertain pre-exponents ($A_i$'s).
+\caption{Left: A bar-graph of activity scores ($\nu_{i,r}$'s) for the 19 uncertain pre-exponents ($A_i$'s); $r$
+denotes the number of eigenvectors in the active subspace.
 Right: A plot of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$), defined as a normalized product of the

### 2018-08-06 Paper/assa.tex
..

@@
 \label{sub:ac}
-As mentioned in the introduction, the active subspace is low-dimensional subspace
+As mentioned in the introduction, the active subspace is a low-dimensional subspace
 that constitutes important directions in a model's input
@@
 $dP_{\bm{\xi}}$ = $p(\bm{\xi})d\bm{\xi}$, where $p(\bm{\xi})$ is the joint probability
-distribution of $\bm{\xi}$. Note that the function, $f$ is considered to be continuous 
-and differentiable in the physical space, $\bm{\theta}$. 
+distribution of $\bm{\xi}$. Hence, $f$ is considered to be continuous 
+and differentiable in the input domain. 
 %Hence, it is possible that a given $f(\bm{\xi})$ might not admit an active
@@
 %problem and hence the associated computational effort.
-Here 
+Here,
 $\nabla_{\bm{\xi}}f$ denotes the gradient vector. 
@@
 of dominant eigenvectors. The model output in the original space, $f(\bm{\xi})$,
-is expressed in terms of the active variables as $G(\mat{W}_1^\top \bm{\xi})$.
+is approximated  as $G(\mat{W}_1^\top \bm{\xi})$ in the active subspace.
 Hence, $f$ is essentially evaluated at the projection of $\bm{\xi}$ on the
@@
 active subspace whose dimension is typically much smaller (in applications that
-admit such a subspace) than the dimension of the original input space. To 
+admit such a subspace) than the dimension of the original input space. To further
 expedite UQ, one could easily fit a regression-surface to $G$ using the 

### 2018-08-06 Paper/assa.tex
started theorem 2.1

--- 2018-08-06 Paper/assa.tex
@@
 %iid $U(0, 1)$, in which case $C_i = 1/\pi^2$, $i = 1, \ldots, \Np$.
+\begin{proof} \hayleynote{(NOT FINISHED YET, NEED TO SORT OUT INDEXING TO BE IN LINE WITH THE PAPER)}
+First we show the numerator 
+\[\int_\Omega \big( f(\vec\xi) - f^{z}(\vec\xi)\big)^2 \, \mu(d\vec\xi) = 2\V_{U^c}(f) \]
+Using the ANOVA decomposition we write:
+\[ f(\vec\xi) = f_0 + f_1(\vec\xi_{U}) + f_2(\vec\xi_{U^C}) + f_{12}(\vec\xi_U,\vec\xi_{U^C})\]
+Then we can simplify the numerator as follows:
+\[ \int_{\Omega} \big( f_2(\vec\xi_{U^C}) + f_{12}(\vec\xi_{U},\vec\xi_{U^C}) - f_2(z) -f_{12}(\vec\xi_{U},z)\big)^2\mu(d\vec\xi) \]
+Using the properties of ANOVA:
+\[
+\int_{\Omega_{U^c}} f_2(\vec\xi_{{U^C}}) \,\mu(d\vec\xi_{U^C}) 
+= \int_{\Omega_{U^C}} f_{12}(\vec\xi_U, \vec\xi_{U^C})\,\mu(d\vec\xi_{U^C}) 
+= \int_{\Omega_U} f_{12}(\vec\xi_U, \vec\xi_{U^C})\,\mu(d\vec\xi_U) = 0,
+\]
+For a fixed $z \in \Omega_{U^C}$ we can simplify the numerator further:
+\[ \int_{\Omega}  f_2^2(\vec\xi_{U^C}) + f_{12}^2(\vec\xi_{U},\vec\xi_{U^C}) - f_2^2(z) -f_{12}^2(\vec\xi_{U},z) \mu(d\vec\xi) = \]
+\[ \V_{U^C}(f) + \V_{U^C,U}(f) + f_2^2(z) + \int_{\Omega_U} f_{12}^2(\vec\xi_U, z) \mu(d\vec\xi_U),\]
+Taking the expectation of the above expression over $\Omega_{U^c}$ gives 
+desired result:
+\[\V_{U^C}(f) + \V_{U^C,U}(f) + \int_{\Omega_{U^C}} f_2^2(z) \, \mu(dz)
++ \int_{\Omega_{U^C}}\int_{\Omega_U} f_{12}^2(\vec\xi_U,z)\,\mu (d\vec\xi_U)
+\mu (dz) \]
+\[= \V_{U^C}(f) + \V_{U,U^C}(f) + \V_{U^C}(f) + \V_{U,U^c}(f) 
+= 2 \V_{U^c}^\text{tot}(f) \]
+Using the definition of variance we rewrite the the denominator as:
+\[
+ \int_\Omega f(\vec\xi)^2\, \mu(d\vec\xi) 
+=  \V(f) + 
+\Big(\int_\Omega f(\vec\xi)\mu(d\vec\xi)\Big)^2 
+\geq   \V(f) 
+\]	
+Combining by parts we have:
+\[ \Ez{ \delta(z)} = 2T_{U^C}(f) \]
+Using (10) above yields:
+\[2T_{U^C}(f) \leq \frac{C_{U^C}}{\V(f)}\nu_{U^C}(f) \]
+Finally using Proposition 2.1 gives us the result:
+\[2T_{U^C}(f) \leq 2\frac{C_{U^C}}{\V(f)}\nu_{U^C}(f) \leq 2\frac{C_{U^C}}{\V(f)}(\nu_{{U^C},r}(f) + \lambda_{r+1} ) \]
+\end{proof}

### 2018-08-06 Paper/intro.tex

@@
 H$_2$/O$_2$ reaction mechanism~\cite{Yetter:1991}. This mechanism has gained
-a lot of attention as a potential source of clean energy
+a lot of attention as a potential source of clean energy for
 locomotive applications~\cite{Das:1996}, and more recently in fuel cells~\cite{Loges:2008}. 
@@
 %
-where the numerator on the right-hand-side denotes the fuel (H$_2$)-oxidizer
-(O$_2$) ratio at given condition to the same quantity under stoichiometric
+where the numerator on the right-hand-side denotes the ratio of the fuel (H$_2$)
+and oxidizer (O$_2$) at a given condition to the same quantity under stoichiometric
 conditions. In this study, computations were performed at fuel-rich conditions,
@@
 rate-controlling parameters, pre-exponent~($A_i$) and the activation
-energy~($E_{a,i}$) on ignition delay. The $A_i$'s are considered to be
+energy~($E_{a,i}$) on the ignition delay. The $A_i$'s are considered to be
 uniformly distributed about their nominal estimates provided
@@
 the present problem high-dimensional and computationally challenging. 
-For the purpose of dimension reduction, the active subspace is evaluated using
+To address this challenge, we focus our efforts on reducing the dimensionality
+of the problem by computing the active subspace using
 two computational strategies: a gradient-based approach and  
@@
 are implemented in an iterative manner i.e. a new set of model evaluations are
-generated at each iteration until convergence is accomplished as discussed
-later in Section~\ref{sec:method}.
+generated, and gradient evaluations are performed at each iteration until 
+a reasonably accurate approximation to the active subspace is achieved, as
+discussed later in Section~\ref{sec:method}.
 
 An alternate approach to dimension reduction involves computing the global sensitivity
-measures associated with uncertain inputs of a model. Uncertainty along the inputs with
-a relatively small estimate of the sensitivity measure can be ignored for the purpose of
-reducing the computational effort associated with uncertainty quantification (UQ). The
-Sobol' indices are commonly used for this purpose~\cite{Sobol:2001}. However, obtaining
+measures associated with uncertain inputs of a model. Depending upon the estimates of
+the sensitivity measures, only the important inputs are varied for the purpose of
+uncertainty quantification (UQ). The
+Sobol' index is a commonly used global sensitivity measure~\cite{Sobol:2001}. However, obtaining
 converged estimates of the Sobol' indices typically requires tens of thousands of model
 runs and is therefore impractical in situations involving intensive model simulations.
-Recently, the derivative-based global sensitivity measures~(DGSMs) have been 
-developed to approximate the upper bound on the Sobol' indies with much fewer
+Recently, the derivative-based global sensitivity measures~(DGSMs) were 
+developed to approximate the upper bound on the Sobol' indices with much fewer
 computations~\cite{Sobol:2009, Lamboni:2013}. In Section~\ref{sec:assa}, we show
 that the DGSMs can in fact be approximated using the components of the eigenvectors
-in the active subspace.
+in the active subspace by estimating the so-called activity 
+scores~\cite{Diaz:2016,Constantine:2017}.
 
 Main contributions of this paper are as follows: (1) Mathematical linkages between the
-global sensitivity measures: Total Sobol index, DGSMs, and the activity scores are
-established; (2) An efficient gradient-free iterative approach based on local linear
+global sensitivity measures: Total Sobol' index, DGSMs, and the activity scores are
+established; (2) An efficient gradient-free iterative approach using regression-based local linear
 approximation of the model output is shown to yield consistent results with the
@@
 
- The remaining of this paper is organized as follows. In Section~\ref{sub:ac}, a brief theoretical background on
- the active subspace methodology is provided. In Section~\ref{sub:gsa}, it is shown that the activity
+ The remaining of this paper is organized as follows. In section~\ref{sub:ac}, a brief theoretical background on
+ the active subspace methodology is provided. In section~\ref{sub:gsa}, it is shown that the activity
  scores provide a reasonable approximation to the DGSMs especially in a high-dimensional setting. 
  Additionally, a relationship between the three global sensitivity measures, namely, the activity scores,
- DGSMs, and the total Sobol' indices is established. In Section~\ref{sec:method}, a systematic framework
+ DGSMs, and the total Sobol' indices is established. In section~\ref{sec:method}, a systematic framework
  for computing the active subspace using the gradient-based, and the gradient-free approaches is presented. 
@@
 the $A_i$'s are considered as uncertain. Our findings illustrate the consistency in the results based on
-the two approaches thereby encouraging the use of the gradient-free approach. In Section~\ref{sec:app},
+the two approaches thereby encouraging the use of the gradient-free approach in high-dimensional settings.
+ In section~\ref{sec:app},
 the gradient-free approach is implemented to the 33-dimensional
@@
 computational effort pertaining to UQ, in this case. 
-Finally, we draw important conclusions based on the various aspects of this work in Section~\ref{sec:conc}.
+Finally, we summarize the results and include a discussion on this work in section~\ref{sec:conc}.

### 2018-08-06 Paper/assa.tex
proof theorem 2.1 attempt

--- 2018-08-06 Paper/assa.tex
@@
 %iid $U(0, 1)$, in which case $C_i = 1/\pi^2$, $i = 1, \ldots, \Np$.
+\begin{proof} \hayleynote{(NOT FINISHED YET, NEED TO SORT OUT INDEXING TO BE IN LINE WITH THE PAPER)}
+First we show the numerator 
+\[\int_\Omega \big( f(\vec\xi) - f^{z}(\vec\xi)\big)^2 \, \mu(d\vec\xi) = 2\V_{i}(f)  \]
+Using the ANOVA decomposition we can write:
+\[ f(\vec\xi) = f_0 + f_1(\vec\xi_j) + f_2(\vec\xi_{i}) + f_{12}(\vec\xi_j,\vec\xi_{i}), \ j \neq i\]
+Then we can simplify the numerator as follows:
+\[ \int_{\Omega} \big( f_2(\vec\xi_{i}) + f_{12}(\vec\xi_{j},\vec\xi_{i}) - f_2(z) -f_{12}(\vec\xi_{j},z)\big)^2\mu(d\vec\xi), \ j \neq i \]
+Using the properties of ANOVA:
+\[
+\int_{\Omega_{i}} f_2(\vec\xi_{i}) \,\mu(d\vec\xi_{i}) 
+= \int_{\Omega_{i}} f_{12}(\vec\xi_j, \vec\xi_{i})\,\mu(d\vec\xi_{i}) 
+= \int_{\Omega_j} f_{12}(\vec\xi_j, \vec\xi_{i})\,\mu(d\vec\xi_j) = 0, \ j \neq i
+\]
+For a fixed $z \in \Omega_{i}$ we can simplify the numerator further:
+\[ \int_{\Omega}  f_2^2(\vec\xi_{i}) + f_{12}^2(\vec\xi_{j},\vec\xi_{i}) - f_2^2(z) -f_{12}^2(\vec\xi_{j},z) \mu(d\vec\xi) = \]
+\[ \V_{i}(f) + \V_{i,j}(f) + f_2^2(z) + \int_{\Omega_j} f_{12}^2(\vec\xi_j, z) \mu(d\vec\xi_j), \ j \neq i\]
+Taking the expectation of the above expression over $\Omega_{i}$ gives 
+desired result:
+\[\V_{i}(f) + \V_{i,j}(f) + \int_{\Omega_{i}} f_2^2(z) \, \mu(dz)
++ \int_{\Omega_{i}}\int_{\Omega_j} f_{12}^2(\vec\xi_j,z)\,\mu (d\vec\xi_j)
+\mu (dz), \ j \neq i \]
+\[= \V_{i}(f) + \V_{j,i}(f) + \V_{i}(f) + \V_{j,i}(f) 
+= 2 \V_{i}^\text{tot}(f), \ j \neq i \]
+Using the definition of variance we rewrite the the denominator as:
+\[
+ \int_\Omega f(\vec\xi)^2\, \mu(d\vec\xi) 
+=  \V(f) + 
+\Big(\int_\Omega f(\vec\xi)\mu(d\vec\xi)\Big)^2 
+\geq   \V(f) 
+\]	
+Combining the two parts we have:
+\[ \Ez{ \delta(z)} = 2T_{i}(f) \]
+Using (10) above yields:
+\[2T_{i}(f) \leq 2\frac{C_{i}}{\V(f)}\nu_{i}(f) \]
+Finally using Proposition 2.1 gives us the result:
+\[2T_{i}(f) \leq 2\frac{C_{i}}{\V(f)}\nu_{i}(f) \leq 2\frac{C_{i}}{\V(f)}(\nu_{{i},r}(f) + \lambda_{r+1} ) \]
+\end{proof}

### 2018-08-06 Paper/assa.tex
..

--- 2018-08-06 Paper/assa.tex
@@
 \label{sub:gsa}
-Consider a function $f(\xi_1, \xi_2, \ldots, \xi_\Np)$. 
-Whereas the active subspace framework described above not make any assumptions
-about statistical independence of the input parameters $\xi_i$, $i = 1, \ldots, \Np$, 
+Consider a function: $f(\xi_1, \xi_2, \ldots, \xi_\Np)$. 
+While the active subspace framework described above does not make any assumptions
+about statistical independence of the inputs $\xi_i$, $i = 1, \ldots, \Np$, 
 the classical 
-framework of variance based sensitivity analysis~\cite{Sobol:2001, Saltelli:2010} assumes the input
-parameters are statistically independent. While extensions to the cases 
-of correlated inputs exist, we limit the discussion in this section to the
+framework of variance based sensitivity analysis~\cite{Sobol:2001, Saltelli:2010} 
+assumes that the inputs
+are statistically independent. While extensions to the cases 
+of correlated inputs exist~\cite{Borgonovo:2007,Li:2010,Jacques:2006,Xu:2007},
+ we limit the discussion in this section to the
 case of random inputs that are statistically independent and are 
@@
 \]
-for detecting unimportant input parameters. 
+for detecting unimportant inputs. 
 Note that if $\xi_i$ are iid, the $C_i$'s will cancel and the normalized screening
@@
 an interesting link between the activity scores and total Sobol' indices.
-Specifically, by computing the activity scores, we can identify unimportant
-parameters.  
+Specifically, by computing the activity scores, we can identify the unimportant
+inputs.  
 One can attempt to reduce parameter dimension by fixing
-unimportant parameters at nominal values. Suppose $\xi_i$ is
+unimportant inputs at nominal values. Suppose $\xi_i$ is
 deemed unimportant as a result of a global sensitivity analysis
-based on computing the activity scores. We want to understand
-what is the approximation error that occurs once we fix $\xi_i$ 
-at a nominal value.
+based on the activity scores; we want to estimate
+the approximation error that occurs once $\xi_i$ 
+is fixed at its nominal value.
 %Let $\I \subset \{1, \ldots,
@@
 %iid $U(0, 1)$, in which case $C_i = 1/\pi^2$, $i = 1, \ldots, \Np$.
-\begin{proof} \hayleynote{(NOT FINISHED YET, NEED TO SORT OUT INDEXING TO BE IN LINE WITH THE PAPER)}
-First we show the numerator 
-\[\int_\Omega \big( f(\vec\xi) - f^{z}(\vec\xi)\big)^2 \, \mu(d\vec\xi) = 2\V_{U^c}(f) \]
-Using the ANOVA decomposition we write:
-\[ f(\vec\xi) = f_0 + f_1(\vec\xi_{U}) + f_2(\vec\xi_{U^C}) + f_{12}(\vec\xi_U,\vec\xi_{U^C})\]
-Then we can simplify the numerator as follows:
-\[ \int_{\Omega} \big( f_2(\vec\xi_{U^C}) + f_{12}(\vec\xi_{U},\vec\xi_{U^C}) - f_2(z) -f_{12}(\vec\xi_{U},z)\big)^2\mu(d\vec\xi) \]
-Using the properties of ANOVA:
-\[
-\int_{\Omega_{U^c}} f_2(\vec\xi_{{U^C}}) \,\mu(d\vec\xi_{U^C}) 
-= \int_{\Omega_{U^C}} f_{12}(\vec\xi_U, \vec\xi_{U^C})\,\mu(d\vec\xi_{U^C}) 
-= \int_{\Omega_U} f_{12}(\vec\xi_U, \vec\xi_{U^C})\,\mu(d\vec\xi_U) = 0,
-\]
-For a fixed $z \in \Omega_{U^C}$ we can simplify the numerator further:
-\[ \int_{\Omega}  f_2^2(\vec\xi_{U^C}) + f_{12}^2(\vec\xi_{U},\vec\xi_{U^C}) - f_2^2(z) -f_{12}^2(\vec\xi_{U},z) \mu(d\vec\xi) = \]
-\[ \V_{U^C}(f) + \V_{U^C,U}(f) + f_2^2(z) + \int_{\Omega_U} f_{12}^2(\vec\xi_U, z) \mu(d\vec\xi_U),\]
-Taking the expectation of the above expression over $\Omega_{U^c}$ gives 
-desired result:
-\[\V_{U^C}(f) + \V_{U^C,U}(f) + \int_{\Omega_{U^C}} f_2^2(z) \, \mu(dz)
-+ \int_{\Omega_{U^C}}\int_{\Omega_U} f_{12}^2(\vec\xi_U,z)\,\mu (d\vec\xi_U)
-\mu (dz) \]
-\[= \V_{U^C}(f) + \V_{U,U^C}(f) + \V_{U^C}(f) + \V_{U,U^c}(f) 
-= 2 \V_{U^c}^\text{tot}(f) \]
-Using the definition of variance we rewrite the the denominator as:
-\[
- \int_\Omega f(\vec\xi)^2\, \mu(d\vec\xi) 
-=  \V(f) + 
-\Big(\int_\Omega f(\vec\xi)\mu(d\vec\xi)\Big)^2 
-\geq   \V(f) 
-\]	
-Combining by parts we have:
-\[ \Ez{ \delta(z)} = 2T_{U^C}(f) \]
-Using (10) above yields:
-\[2T_{U^C}(f) \leq \frac{C_{U^C}}{\V(f)}\nu_{U^C}(f) \]
-Finally using Proposition 2.1 gives us the result:
-\[2T_{U^C}(f) \leq 2\frac{C_{U^C}}{\V(f)}\nu_{U^C}(f) \leq 2\frac{C_{U^C}}{\V(f)}(\nu_{{U^C},r}(f) + \lambda_{r+1} ) \]
-\end{proof}

### 2018-08-06 Paper/conc.tex
..

@@
 Motivated by these findings, the gradient-free approach is used for computing the active subspace in the high-dimensional
-case involving 33 uncertain parameters. Based on the prior marginals of the uncertain parameters, our results
+case involving 33 uncertain inputs. Based on the prior marginals of the uncertain inputs, our results
 indicate that the uncertainty in the ignition delay is largely dependent on $A_1$, $A_9$, and $E_{a,1}$, although

### 2018-08-06 Paper/method.tex
..

@@
 of $\hat{\mat{C}}$ yields an initial estimate of the dominant eigenspace,
-$\hat{\mat{W}}_1$ and a diagonal matrix, 
-$\hat{\mat{\Lambda}}_1$ that comprises the eigenvalues along the diagonal.
+$\hat{\mat{W}}_1$ and the set of corresponding eigenvalues, 
+$\hat{\mat{\Lambda}}_1$.
  At each subsequent iteration, model evaluations are
@@
 We complement the comparative assessment of the two approaches by estimating
-the activity scores for individual uncertain parameters ($\nu_{i,r}$; $r$=1 since the
+the activity scores for individual uncertain inputs ($\nu_{i,r}$; $r$=1 since the
 active subspace is 1-dimensional)

### 2018-08-07 Paper/assa.tex
..

@@
 are statistically independent. While extensions to the cases 
-of correlated inputs exist~\cite{Borgonovo:2007,Li:2010,Jacques:2006,Xu:2007},
+of correlated inputs exist~\cite{Borgonovo:2007,Li:2010,Jacques:2006,Xu:2007,Hart:2017},
  we limit the discussion in this section to the

### 2018-08-07 Paper/method.tex
..

@@
 norm of the difference between
-the square of individual components of the first five eigenvectors of $\mat{C}$, constructed using 
-$N$~=~1000 samples ($w_{i,N}$), and using a smaller sample size, $n$ ranging from 20 to 240 ($w_{i,n}$).} 
+the square of individual components of the eigenvector evaluated using $N$=1000 and $n$ = $\{20,40,80,120\}$ samples.
+The comparison is illustrated for the first five eigenvectors.} 
 \label{fig:eig_comp}

### 2018-08-08 Paper/app.tex
..

@@
 In the above comparison, it is observed that the two PDFs are almost identical, thereby confirming a 1-dimensional
-active subspace. The modal estimate from either PDF is found to be in agreement with that
-based on the histogram plot. Moreover, the spread in the ignition
-delay is reasonably captured by the PDFs.
+active subspace. The modal estimate from either PDF is observed to be in agreement with that
+based on the histogram plot. The mean value based on the set of model evaluations used for verification as well
+as the 1-dimensional surrogate was found to be 0.133, whereas, the standard deviations were found to be 0.0198 and 0.0160
+respectively. Hence, while the surrogate yields an accurate estimation of the mean, the standard deviation is
+under-estimated. This can be understood from the SSP in Figure~\ref{fig:hd} (right) which shows that the model
+evaluations do not exactly lie on the linear 1-dimensional surrogate used to capture them. Therefore, the
+1-dimensional surrogate can be used to estimate the mean as well as the modal estimate of the ignition delay in
+the presence of uncertainty in the 33 parameter considered in this application. The uncertainty associated with the
+ignition delay though reasonably close to the true value, is under-estimated.
 
-Based on these observations, it can be said that a 1-dimensional surrogate can be used to reasonably approximate the
-uncertainty in the ignition delay due to uncertain rate-controlling parameters used in the H$_2$/O$_2$ reaction
-mechanism.
+%\subsection{Surrogate-induced risk analysis}

### 2018-08-08 Paper/assa.tex
minor
sec 2.2 ...

--- 2018-08-06 Paper/assa.tex
@@
 \label{sub:gsa}
-Consider a function: $f(\xi_1, \xi_2, \ldots, \xi_\Np)$. 
+Consider a function $f = f(\xi_1, \xi_2, \ldots, \xi_\Np)$. 
 While the active subspace framework described above does not make any assumptions
@@
 \[
-\delta(z) =
+\mathcal{E}(z) =
 \frac{ \int_\Omega \big( f(\vec\xi) - f^{z}(\vec\xi)\big)^2 \, \mu(d\vec\xi) }
@@
 \] 
-
 \begin{theorem}
-We have $\Ez{ \delta(z)} \leq 2C_i\big(\act{f}{i}{r} + \lambda_{r+1}\big)/{\V(f)} $.
+We have $\Ez{ \mathcal{E}(z)} \leq 2C_i\big(\act{f}{i}{r} + \lambda_{r+1}\big)/{\V(f)} $.
 \end{theorem}
@@
 %iid $U(0, 1)$, in which case $C_i = 1/\pi^2$, $i = 1, \ldots, \Np$.
+\begin{proof} 
+Note that, since 
+$\int_\Omega f(\vec\xi)^2 \, \mu(d\vec\xi) = \V(f) + 
+\left(\int_\Omega f(\vec\xi) \, \mu(d\vec\xi)\right)^2 \geq \V(f)$, we have
+\[
+\Ez{ \mathcal{E}(z)} \leq \frac{1}{\V(f)} \Ez{ 
+\int_\Omega \big( f(\vec\xi) - f^{z}(\vec\xi)\big)^2 \, \mu(d\vec\xi)}
+= 2 T_i(f), 
+\]
+where the equality can be shown using arguments similar to the proof of the main result 
+in~\cite{SobolTarantolaGatelliEtAl07}. Using this, along with~\eqref{equ:sobol_bound} and
+Proposition~\ref{prp:dgsm_bound}, we have 
+\[
+\Ez{ \mathcal{E}(z)} \leq 
+\frac{2C_i}{\V(f)}\nu_i(f)
+\leq 
+\frac{2C_i}{\V(f)}\big[\act{f}{i}{r} + \lambda_{r+1}\big]. \qedhere
+\]
+\end{proof}

### 2018-08-10 Paper/assa.tex
very minor changes

@@
 \be
-\mat{C} = \int_\Omega (\nabla_{\bm{\xi}}f)(\nabla_{\bm{\xi}}f)^\top dP_{\bm{\xi}}
+\mat{C} = \int_\Omega (\nabla_{\bm{\xi}}f)(\nabla_{\bm{\xi}}f)^\top d\Pi_{\bm{\xi}}
 \label{eq:C}
@@
 where 
-$dP_{\bm{\xi}}$ = $p(\bm{\xi})d\bm{\xi}$, where $p(\bm{\xi})$ is the joint probability
+$d\Pi_{\bm{\xi}}$ = $\pi(\bm{\xi})d\bm{\xi}$, where $\pi(\bm{\xi})$ is the joint probability
 distribution of $\bm{\xi}$. Hence, $f$ is considered to be continuous 
@@
 into a reduced set, $\mat{W}_1^\top \bm{\xi}$, referred to as the set of active
-variables. The number of active variables is thus equal to the number
+variables. The number of active variables is equal to the number
 of dominant eigenvectors. The model output in the original space, $f(\bm{\xi})$,
@@
 Hence, $f$ is essentially evaluated at the projection of $\bm{\xi}$ on the
-column space of $\mat{W}_1$. We could thus confine UQ to the inputs in the
+column space of $\mat{W}_1$. We could confine UQ to the inputs in the
 active subspace whose dimension is typically much smaller (in applications that
@@
 
-The total Sobol' index ($T_i(f)$) of a model output, $f(\vec\theta)$ quantifies
-the total contribution of the input, $\theta_i$ to the variance of the
+The total Sobol' index ($T_i(f)$) of a model output, $f(\vec\xi)$ quantifies
+the total contribution of the input, $\xi_i$ to the variance of the
 output~\cite{Sobol:2001}. Mathematically, this can be expressed as follows:
@@
 T_i(f) = 1 - 
-\frac{\V[\mathbb{E}(f|\bm{\theta}_{\sim i})]}{\V(f)},
+\frac{\V[\mathbb{E}(f|\bm{\xi}_{\sim i})]}{\V(f)},
 \label{eq:total}
@@
 %
-where $\bm{\theta}_{\sim i}$ is a vector of all but $i^\text{th}$ uncertain
-input, and $\V$ denotes the variance. The total Sobol' index thus accounts
+where $\bm{\xi}_{\sim i}$ is a vector of all the inputs $\xi_i$ except the  $i^\text{th}$ entry, and $\V$ denotes the variance. The total Sobol' index accounts
 for the contribution of a given input to the variability in the output by itself
@@
 which gives a spectral representation for the DGSMs. 
-In the case the eigenvalues decay rapidly to zero, we can obtain
+In the case where the eigenvalues decay rapidly to zero, we can obtain
 accurate approximations of $\nu_i(f)$ by Truncating the summation: 

### 2018-08-13 Paper/intro.tex
..

@@
 output with respect to an input. The active subspace methodology thus
-focuses on reducing the dimensionality of the problem.
+focuses on reducing the dimensionality of the problem, and hence the 
+computational effort associated with uncertainty propagation.
 
@@
 a lot of attention as a potential source of clean energy for
-locomotive applications~\cite{Das:1996}, and more recently in fuel cells~\cite{Loges:2008}. 
+locomotive applications~\cite{Das:1996}, and more recently in fuel
+ cells~\cite{Loges:2008,Cosnier:2016}. 
 The mechanism involves 19
@@
 rate-controlling parameters, pre-exponent~($A_i$) and the activation
-energy~($E_{a,i}$) on the ignition delay. The $A_i$'s are considered to be
-uniformly distributed about their nominal estimates provided
+energy~($E_{a,i}$) on the ignition delay. The $A_i$'s associated with all
+reactions and the $E_{a,i}$'s with non-zero nominal estimates
+are considered to be uniformly distributed about their nominal estimates provided
 in~\cite{Yetter:1991}.  The total number of uncertain inputs is 33 which makes
@@
 To address this challenge, we focus our efforts on reducing the dimensionality
-of the problem by computing the active subspace using
-two computational strategies: a gradient-based approach and  
-a gradient-free approach. These strategies are based on algorithms 1.1 and 1.2 
-in~\cite{Constantine:2015}. As mentioned earlier, the gradient information of the model
-output is needed in both cases. However, in the case of gradient-free approach,
-the gradient is approximated using a linear regression-fit to the model output.
-Hence, the amount of computational effort is reduced in this case. Both strategies
+of the problem by computing the aforementioned active subspace using
+an approach that involves computation of gradients of a model output with
+respect to individual parameters. Estimating the
+gradients can however be computationally challenging in high-dimensional settings. 
+Hence, in this work, we further explore the suitability of a gradient-free
+approach that involves a regression-based local linear approximation of
+the model output. It must be noted that since the gradient-free approach
+relies on approximating the model gradients, it is expected to be less
+robust than the gradient-based approach. The two approaches are based on
+algorithms 1.1 and 1.2 in~\cite{Constantine:2015}. Both strategies
 are implemented in an iterative manner i.e. a new set of model evaluations are
@@
 converged estimates of the Sobol' indices typically requires tens of thousands of model
-runs and is therefore impractical in situations involving intensive model simulations.
+runs and is therefore impractical in situations involving intensive model simulations and/or a
+large number of inputs.
 Recently, the derivative-based global sensitivity measures~(DGSMs) were 
@@
 global sensitivity measures: Total Sobol' index, DGSMs, and the activity scores are
-established; (2) An efficient gradient-free iterative approach using regression-based local linear
-approximation of the model output is shown to yield consistent results with the
-gradient-based approach. The two approaches were implemented to compute
-the active subspace for a 19-dimensional H$_2$/O$_2$ kinetics problem with 
-uncertain pre-exponents in the rate law for individual reactions. (3) The gradient-free
-approach is finally implemented to discover an active subspace for the 33-dimensional
-H$_2$/O$_2$ reaction kinetics application in which case the pre-exponents as well
-as the activation energies are considered to be uncertain.
+established; (2) An initial comparison of the active subspace and the global sensitivity
+measures, computed using the gradient-based and the gradient-free approaches in
+a 19-dimensional H$_2$/O$_2$ kinetics problem is provided. In this case, only the $A_i$'s are
+considered to be uncertain. (3) Active subspace discovery in the high-dimensional
+H$_2$/O$_2$ kinetics problem involving 33 uncertain inputs.
 
@@
  DGSMs, and the total Sobol' indices is established. In section~\ref{sec:method}, a systematic framework
- for computing the active subspace using the gradient-based, and the gradient-free approaches is presented. 
-The two approaches are implemented to the H$_2$/O$_2$ reaction kinetics problem whereby only
-the $A_i$'s are considered as uncertain. Our findings illustrate the consistency in the results based on
-the two approaches thereby encouraging the use of the gradient-free approach in high-dimensional settings.
- In section~\ref{sec:app},
-the gradient-free approach is implemented to the 33-dimensional
-H$_2$/O$_2$ reaction kinetics application by accounting for the uncertainty in $E_{a,i}$'s as well. The
-resulting active subspace is found to be 1-dimensional indicating enormous potential for reduction in
-computational effort pertaining to UQ, in this case. 
-Finally, we summarize the results and include a discussion on this work in section~\ref{sec:conc}.
+ for computing the active subspace using an iterative gradient-based approach is provided. Numerical results
+ based on the gradient-free approach are compared with those obtained using the gradient-based approach.
+ The active subspace is computed for the H$_2$/O$_2$ reaction kinetics problem whereby only
+the $A_i$'s are considered as uncertain. In this case, the two approaches are observed to yield consistent
+results. In section~\ref{sec:app}, the gradient-based approach is further implemented to compute the
+active subspace for the high-dimensional H$_2$/O$_2$ reaction kinetics problem. The resulting active
+subspace is found to be 1-dimensional thereby indicating enormous potential for computational savings
+in this application. The active subspace is also computed using the gradient-free approach. It is observed
+that although the uncertainty in the ignition delay is reasonably approximated, the gradient-free approach
+fails to capture the global sensitivity of all the uncertain inputs. Hence, a trade-off between efficiency and
+accuracy between the two approaches is observed in this case. 
+Finally, a summary and discussion based on our findings is included in section~\ref{sec:conc}.

### 2018-08-13 Paper/method.tex

@@
 active subspace in the case of H$_2$/O$_2$ reaction kinetics with uncertain
-pre-exponent of individual reaction rates provided in Table~\ref{tab:kinetics}.
+pre-exponent of the rate law associated with individual reactions provided in
+Table~\ref{tab:kinetics}.
 Specifically, we employ two strategies: 
@@
 1.1, and the gradient-free approach is based on Algorithm 1.2
-in~\cite{Constantine:2015}. In this work, we implement the two
-strategies in an iterative manner and demonstrate potential for
-additional savings in both cases.  For the present application,
-our findings based on the two strategies are observed to be consistent.  
-Since,
-the gradient-free approach is more efficient, this observation
-encourages its use for high-dimensional applications as discussed later in
-Section~\ref{sec:app}. 
-%Note however, that the model output is required to be
-%differentiable throughout the input domain in both cases.
+in~\cite{Constantine:2015}. The two strategies are implemented in an
+iterative manner to ensure computational efficiency by avoiding 
+extraneous model evaluations once a converged active subspace
+is achieved.
 
@@
 eigenvectors between subsequent iterations is evaluated. The process is
-terminated and the resulting eigenspace is considered to be converged once
+terminated and the resulting eigenspace is considered to have converged once
 the maximum relative change at iteration $k$, $\delta \hat{\mat{W}}_{1,j}^{(k)}$
@@
 dominant eigenvectors are used to compute the activity scores, $\bm{\nu}_r(f)$,
-which provide insight into the relative importance of the uncertain inputs.
+which provide an insight into the relative importance of the uncertain inputs.
 Note that the index, $r$, corresponds to the number of dominant eigenvectors in
@@
 	$i=1,\ldots,N_p$.
+	\State Normalize $\nu_{i,r}(f)$ as $\tilde{\nu_{i,r}}(f)$ = \scalebox{1.25}{$\frac{\nu_{i,r}(f)}{\sum_i\nu_{i,r}(f)}$}.
 
@@
 smaller set of samples, $n$~=~$\{20,40,80,120\}$.
-Additionally, we compare relative L-2
-norm of the difference between the squared value of corresponding components
-of the eigenvectors, obtained using the large set, $N$, and the smaller sets, $n$.
-In  Figure~\ref{fig:eig_comp} (right), we plot the relative L-2 error norm for the
-first five eigenvectors.
+
+%Additionally, we compare relative L-2
+%norm of the difference between the squared value of corresponding components
+%of the eigenvectors, obtained using the large set, $N$, and the smaller sets, $n$.
+%In  Figure~\ref{fig:eig_comp} (right), the relative L-2 error norm for the
+%first five eigenvectors is plotted.
 
@@
   \includegraphics[width=0.45\textwidth]{./Figures/eig_comp}
-   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1_5}
+   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1}
 \caption{Left: A comparison of the normalized eigenvalue spectrum using $n$ = $\{20,40,80,120\}$ samples with that
@@
 norm of the difference between
-the square of individual components of the eigenvector evaluated using $N$=1000 and $n$ = $\{20,40,80,120\}$ samples.
-The comparison is illustrated for the first five eigenvectors.} 
+the square of individual components ($w_i$'s) of the dominant eigenvector evaluated using $N$=1000 and $n$ = $\{20,40,80,120\}$ samples.} 
 \label{fig:eig_comp}
@@
 reasonably well with just 20 samples. As expected, the accuracy of higher-order eigenvalues is observed
-to improve with the sample size. Similarly, the relative error associated with the dominant eigenvectors
-($i$ = 1,2) is found to range from $\mathcal{O}(10^{-2} - 10^{-1})$ and is relatively much smaller than 
-the higher-order eigenvectors. In this case, since 
+to improve with the sample size. Since 
 $\left(\frac{\lambda_1}{\lambda_2}\right)\gg 1$, we expect 
@@
 sufficiently captures the uncertainty in
-the  model output. Therefore, a sample size of 20 is adequate for computing the active subspace in this
+the  model output. To confirm this, we plot a relative L-2 norm of the difference between the 
+squared value of corresponding components of the dominant eigenvector, computed using $N$ = 1000
+and $n$ = $\{20,40,80,120\}$ samples in Figure~\ref{fig:eig_comp} (right). 
+The relative L-2 error norm is found to be $\mathcal{O}(10^{-2})$ even for the case, $n$ = 20.
+Hence, a sample size of 20 is adequate for computing the active subspace in this
 case. The iterative strategy therefore offers a significant potential for computational advantage. 
-However, estimating the gradient of the model output imposes a significant computational burden.
-Hence, we next explore the applicability of a so-called gradient-free approach
-as mentioned earlier.
+In the following section, we provide a brief description of the gradient-free approach and compare
+numerical results for the active subspace and the global sensitivity measures based on the two
+approaches.
 
@@
 in Algorithm~\ref{alg:grad}. Once again, we use an iterative strategy for
-computing the active subspace. The utility of this approach is demonstrated
-by constructing an efficient response surface for the 19-dimensional 
+computing the active subspace for the 19-dimensional 
 H$_2$/O$_2$ reaction kinetics problem considered in~\ref{sub:grad}.
@@
 those obtained using the gradient-based approach to assess consistency
-between the two approaches. Motivated by our findings, we compute the active 
-subspace for a 33-dimensional H$_2$/O$_2$ reaction kinetics application using the gradient-free
-approach in Section~\ref{sec:app}.
+between the two approaches.
 
@@
 Hence, the gradient-free approach essentially differs in the procedure for constructing the matrix, $\hat{\mat{C}}$.
+Instead of the model evaluations, a linear regression-based approximation is used to estimate the gradient of
+the model output. The suitability of this approach thus depends upon the accuracy of this approximation. 
 The active subspace computation involves the same sequence of steps as outlined previously for the gradient-based 
@@
 As mentioned earlier, the model output $f(\bm{\xi})$ predominantly varies in the active subspace. Hence, 
-$f(\bm{\xi})$ can be transformed as a function of the active variable as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
+$f(\bm{\xi})$ can be approximated as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
 the subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, regarded as the \textit{sufficient summary plot} (SSP) is 
@@
 The two SSPs are observed to be in excellent agreement with each other. Moreover, it is interesting to note
-that the variability in ignition
+that the response in ignition
 delay based on the considered prior marginals for $A_i$'s is observed to be approximately linear in the
-active subspace. This observation indicates the the active subspace is 1-dimensional in this case.
+active subspace. This observation further indicates that the active subspace is 1-dimensional in this case.
 
-We complement the comparative assessment of the two approaches by estimating
-the activity scores for individual uncertain inputs ($\nu_{i,r}$; $r$=1 since the
+We extend the comparative assessment of the two approaches by estimating
+the normalized activity scores for individual uncertain inputs ($\tilde{\nu_{i,r}}$; $r$=1 since the
 active subspace is 1-dimensional)

### 2018-08-13 Paper/method.tex
.

@@
   \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}
-\caption{Left: A bar-graph of activity scores ($\nu_{i,r}$'s) for the 19 uncertain pre-exponents ($A_i$'s); $r$
+\caption{Left: A bar-graph of normalized activity scores ($\tilde{\nu_{i,r}}$'s) for the 19 uncertain pre-exponents ($A_i$'s); $r$
 denotes the number of eigenvectors in the active subspace.
@@
 The above comparisons establish consistency between gradient-based and
-gradient-free approaches. Due to its reduced computational overhead, the
-gradient-free approach is our method of choice for the 33-dimensional problem
-in the next section.
+gradient-free approaches. This indicates that the local linear approximation
+of the model output resulted in a reasonably accurate estimation of the
+gradients in this case. In the following section, we will once again use both
+approaches to discover an active subspace in the high-dimensional 
+problem involving 33 uncertain inputs.

### 2018-08-14 Paper/app.tex
..

@@
 pre-exponents ($A_i$'s) as well as the activation energies ($E_{a,i}$'s) on ignition 
-delay during the H$_2$/O$_2$ reaction. The $A_i$'s are considered to be uniformly
+delay during the H$_2$/O$_2$ reaction. The $A_i$'s were considered to be uniformly
 distributed in the interval, $[0.9A_i^\ast, 1.1A_i^\ast]$. The $E_{a,i}$'s for all
-reactions except $\mathcal{R}_6$ -- $\mathcal{R}_9$ and $\mathcal{R}_{13}$
-are considered to be uncertain and uniformly distributed in the interval: 
+reactions except $\mathcal{R}_6$ -- $\mathcal{R}_9$ and $\mathcal{R}_{13}$ (due to 
+zero nominal values for $E_a$)
+were considered to be uncertain and uniformly distributed in the interval: 
 $[0.99E_{a,i}^\ast, 1.01E_{a,i}^\ast]$. The nominal values, $A_i^\ast$ and $E_{a,i}^\ast$
@@
 
-The gradient-free approach discussed earlier in~\ref{sub:gradfree} was used to compute the
+The gradient-based approach discussed earlier in~\ref{sub:grad} was used to compute the
 active subspace. Using the iterative procedure, the convergence of the eigenvectors
-was examined by tracking $\delta \hat{\mat{W}}_{1,j}^{(i)}$. In Figure~\ref{fig:conv_app} (left),
-components of the converged eigenvector based on 50 samples using $\tau$ = 0.2 are plotted.
-To illustrate convergence, we compare them with components of the eigenvector obtained
-using 100 samples (twice as many). The variation of maximum $\delta w_{1,j}^{(i)}$ between
-successive iterations during the course of the iterative procedure is illustrated in 
-Figure~\ref{fig:conv_app} (right). 
+was examined by tracking $\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$, plotted in 
+Figure~\ref{fig:conv_app} (right). In Figure~\ref{fig:conv_app} (left), individual
+components of the converged eigenvector are illustrated. The convergence was established using
+a $\tau$ value of 0.02. Finite difference was used to compute model gradients at 40 samples in
+the input domain. Hence, a total of 1360 model evaluations were required for this purpose.  
 %
@@
 using the two approaches discussed in~\ref{sub:grad} and~\ref{sub:gradfree}. Right: The quantity,  
-$\delta \hat{\mat{W}}_{1,j}^{(i)}$
+$\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$
 is plotted for successive iterations to illustrate the convergence behavior.}
@@
 %
-The dominant eigenvector is observed to converge with only 50 samples in the 33-dimensional space which
-is remarkably efficient.
 
-In Figure~\ref{fig:hd}, we illustrate the resulting eigenvalue spectrum and the SSP.
+In Figure~\ref{fig:hd} (left), we illustrate the resulting eigenvalue spectrum. The second eigenvalue is
+found to be roughly two orders of magnitude smaller than the first, indicating that the active subspace
+is 1-dimensional. This is further confirmed by the SSP plots in Figure~\ref{fig:hd} (right). Specifically
+in the case of gradient-based approach, the SSP exhibits a linear variation that is reasonably
+captured by its linear fit, $\tilde{G}$. Whereas, the SSP based on the gradient-free approach
+is relatively more scattered. Consequently, the linear fit exhibits a discrepancy with the corresponding
+fit for the gradient-based SSP. 
 %
@@
 %
-The second eigenvalue is observed to be roughly four orders of magnitude smaller than the first 
-indicating the existence of a 1-dimensional active subspace. This is further confirmed by the SSP
- which shows that the variability in $G$ can be approximated using a 1-dimensional regression
-fit ($\tilde{G}$) which could be used as a surrogate. A linear, quadratic, and a cubic fit to the SSP was performed.
-The linear fit illustrated in Figure~\ref{fig:hd} (right) was found to be the most accurate. In~\ref{sub:verify},
-we investigate the accuracy of the 1D surrogate based on the linear fit by estimating the relative L-2
-norm of its discrepancy with model evaluations at an independent set of samples, and by comparing the
-respective probability distributions.
-
-Components of the dominant eigenvector in Figure~\ref{fig:conv_app} (left) are used for estimating the
-activity scores, plotted below in Figure~\ref{fig:as_33D}.
+We plot the normalized activity scores as well as the total Sobol' indices for the 33 uncertain 
+rate-controlling parameters in Figure~\ref{fig:as_33D}. Note that the normalized activity scores
+as well as the Sobol' indices
+were evaluated using respective 1-dimensional surrogates for the two approaches.
 %
@@
  \begin{center}
-  \includegraphics[width=0.45\textwidth]{./Figures/as_A_33D}
-    \includegraphics[width=0.45\textwidth]{./Figures/as_E_33D}
+  \includegraphics[width=0.8\textwidth]{./Figures/as_33D_new}
 \caption{A bar-graph illustrating individual activity scores for the uncertain $A_i$' (left) and $E_{a,i}$'s (right).}
@@
 %
-From the above plot, it can be inferred that the ignition delay is predominantly sensitivity towards $A_1$, $A_9$, and
-$E_{a,1}$. Furthermore, it is moderately sensitivity towards $A_{15}$ and $A_{17}$ while sensitivity towards remaining
-inputs is found to be low or negligible. Note that the sensitivity results for the pre-exponents is found to be consistent
-with the 19-dimensional case presented in~\ref{sec:method}. This observation can also be exploited for reducing the 
-dimensionality of the problem from 33 to 3-5 depending upon the required level of accuracy in terms of characterizing 
-the uncertainty in the model output. The sensitivity-driven dimension reduction for uncertainty quantification has been
+
+Several useful inferences can be drawn from the above plot. Firstly, the total Sobol indices are found to be
+consistent
+with the normalized activity scores evaluated uing the gradient-based approach. Whereas, the gradient-free
+approach yields consistent results for the $A_i$'s, it does not capture the sensitivity associated with
+$E_{a,15}$. While this observation bolsters our confidence in the gradient-based approach, it clearly
+underscores the shortcoming of the gradient-free approach. Numerical errors are incurred when approximating
+the model gradients using the local linear approximation in this case. In~\ref{sub:verify}, we assess the
+suitability of the gradient-free approach in quantifying the uncertainty in the ignition delay. 
+Secondly, our findings indicate that the  
+the ignition delay is predominantly sensitivity towards $A_1$, $A_9$, $E_{a,1}$, and $E_{a,15}$ and 
+moderately sensitivity towards $A_{15}$ and $A_{17}$. Sensitivity towards the remaining uncertain inputs is
+found to be low or negligible. Note that his observation can also be exploited for reducing the 
+dimensionality of the problem from 33 inputs to 4-6 inputs depending upon the required level of accuracy.
+The sensitivity-driven dimension reduction for uncertainty quantification has been
 discussed in our earlier effort~\cite{Vohra:2018}. In this work, however, we focus on dimension reduction using the active 
-subspace approach which is shown to yield even greater scope for dimension reduction i.e. from 33 to 1. We verify
-the accuracy of the 1-d regression fit to $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in the following section.
+subspace approach which is shown to yield even greater scope for dimension reduction i.e. from 33 to 1. Furthermore, we have
+shown that the dominant eigenspace can be used to approximate global sensitivity measures with no additional effort.
+The accuracy of the 1-dimensional surrogate, $\tilde{G}$, obtained using the two approaches is assessed in the following section.
 

### 2018-08-14 Paper/app.tex
..

@@
 underscores the shortcoming of the gradient-free approach. Numerical errors are incurred when approximating
-the model gradients using the local linear approximation in this case. In~\ref{sub:verify}, we assess the
+the model gradients using the local linear approximation in this case. In~\ref{sub:verify}, we further assess the
 suitability of the gradient-free approach in quantifying the uncertainty in the ignition delay. 
-Secondly, our findings indicate that the  
-the ignition delay is predominantly sensitivity towards $A_1$, $A_9$, $E_{a,1}$, and $E_{a,15}$ and 
-moderately sensitivity towards $A_{15}$ and $A_{17}$. Sensitivity towards the remaining uncertain inputs is
-found to be low or negligible. Note that his observation can also be exploited for reducing the 
+Secondly, our findings indicate that   
+the ignition delay is predominantly sensitive towards $A_1$, $A_9$, $E_{a,1}$, and $E_{a,15}$ and 
+moderately sensitive towards $A_{15}$ and $A_{17}$. Sensitivity towards the remaining uncertain inputs is
+found to be low or negligible. Note that this observation can also be exploited for reducing the 
 dimensionality of the problem from 33 inputs to 4-6 inputs depending upon the required level of accuracy.
@@
 %
-The relative norm, $\varepsilon_d$ was estimated to be 8.50$\times10^{-2}$ based on a cross-validation test data
-comprising 10$^{4}$ model evaluations. Hence, in the norm-sense, it can be said that the 1-dimensional surrogate is 
-reasonably accurate.
+The relative norm, $\varepsilon_d$ was estimated to be 1.35$\times10^{-2}$ and 1.02$\times10^{-1}$
+using $\tilde{G}$ from the gradient-based approach and the gradient-free approach respectively. Hence,
+an order of magnitude increase in $\varepsilon_d$ was estimated in the case of gradient-free approach.
+Model evaluations ($G$) at 10$^{4}$ samples in the input domain were used for this purpose. 
+Hence, in the norm-sense, it can be said that the 1-dimensional surrogate in the active subspace, obtained
+using the gradient-based approach is remarkably accurate with a relative error of less than 1.5$\%$.
 
-Secondly, we verify the accuracy of the surrogate in a probabilistic setting. In particular, we compare a histogram
-plot of the model evaluations (in the cross-validation set) with a PDF evaluated on the same set of samples
-using the surrogate, $\tilde{G}$ in Figure~\ref{fig:pdf_33D}. The PDF is evaluated using surrogates constructed for
-a 1-dimensional as well as a 2-dimensional active subspace. 
+Secondly, we verify the accuracy of the two surrogates in a probabilistic setting. In particular, we compare 
+probability density functions (PDFs) obtained using the true set of model evaluations, and 1-dimensional
+surrogates from the two approaches, in Figure~\ref{fig:pdf_33D}. Note that the three PDFs were evaluated 
+using the same set of 10$^4$ samples in the cross-validation set. 
 %
@@
 \begin{center}
-  \includegraphics[width=0.45\textwidth]{./Figures/pdf_comp_id_1e4}
-\caption{A comparison of the histogram plot based on model evaluations and the PDF based on surrogate
-predictions for ignition delay. The two plots are based on samples in the validation set.}
+\begin{minipage}[htbp]{.25\linewidth}
+\vspace{0pt}
+%\centering
+\hspace{-25mm}
+\begin{tabular}{ccc}
+\toprule
+$\textbf{Distribution}$ & $\mu$ & $\sigma$ \\ 
+\bottomrule
+$G$~(Model) & 0.133 & 0.0198 \\
+$\tilde{G}$~(Gradient-based) & 0.133 & 0.0196 \\
+$\tilde{G}$~(Gradient-free) & 0.133 & 0.0167 \\
+\bottomrule
+\end{tabular}
+\end{minipage}
+\hspace{5mm}
+\begin{minipage}[htbp]{.25\linewidth}
+\vspace{0pt}
+%\centering
+\includegraphics[width=3.0in]{./Figures/pdf_comp_id_1e4}
+\end{minipage}%
+\end{center} 
+\caption{Left: Table providing the mean ($\mu$) and the standard deviation ($\sigma$) corresponding to
+the three PDFs using evaluations
+at 10$^4$ samples in the cross-validation set. Right: A comparison of the PDFs of ignition delay, obtained using model 
+evaluations (solid line) and 1-dimensional surrogates using the gradient-free approach (dashed line) and the gradient-based
+approach (dashed-dotted line). The same set of 10$^4$ samples in the cross-validation set were used in each case.}
 \label{fig:pdf_33D}
-\end{center}
 \end{figure}
 %
-In the above comparison, it is observed that the two PDFs are almost identical, thereby confirming a 1-dimensional
-active subspace. The modal estimate from either PDF is observed to be in agreement with that
-based on the histogram plot. The mean value based on the set of model evaluations used for verification as well
-as the 1-dimensional surrogate was found to be 0.133, whereas, the standard deviations were found to be 0.0198 and 0.0160
-respectively. Hence, while the surrogate yields an accurate estimation of the mean, the standard deviation is
-under-estimated. This can be understood from the SSP in Figure~\ref{fig:hd} (right) which shows that the model
-evaluations do not exactly lie on the linear 1-dimensional surrogate used to capture them. Therefore, the
-1-dimensional surrogate can be used to estimate the mean as well as the modal estimate of the ignition delay in
-the presence of uncertainty in the 33 parameter considered in this application. The uncertainty associated with the
-ignition delay though reasonably close to the true value, is under-estimated.
+%
+%\begin{figure}[htbp]
+% \begin{center}
+%  \includegraphics[width=0.45\textwidth]{./Figures/pdf_comp_id_1e4}
+%\caption{A comparison of the PDFs of ignition delay, obtained using model evaluations (solid line) and
+%1-dimensional surrogates using the gradient-free approach (dashed line) and the gradient-based approach
+%(dashed-dotted line). The same set of 10$^4$ samples in the cross-validation set were used in each case.}
+%\label{fig:pdf_33D}
+%\end{center}
+%\end{figure}
+%
+The PDFs based on the model evaluations and the gradient-based approach are nearly identical, confirming
+the existence of a 1-dimensional active subspace as well as the accuracy of the gradient-based approach. 
+Whereas, the PDF based on the gradient-free approach is observed to show consistency with the true PDF in the modal
+estimate; it underestimates the spread in the ignition delay. Specific values of the mean and the standard deviation
+in each case are provided in the table in Figure~\ref{fig:pdf_33D}. Hence, the three PDFs are found to be
+consistent in their prediction of mean value of the ignition delay.  The standard deviation estimates based on model
+evaluations and the gradient-based approach are found to be in close agreement, whereas, it  is accurately estimated
+by the gradient-free approach only upto the first significant figure.
 

### 2018-08-14 Paper/method.tex
..

@@
 with uncertain pre-exponents ($A_i$'s) also yielded a 1-dimensional active subspace. In Figure~\ref{fig:comp} (left),
-we provide a comparison of the components of the dominant eigenvector, obtained using the two approaches.
+we provide a comparison of the square of individual components of the dominant eigenvector, obtained using the two 
+approaches.
 Additionally, the quantity, $\delta \hat{\mat{W}}_{1,j}^{(r)}$ is also plotted in
@@
   \includegraphics[width=0.8\textwidth]{./Figures/eigv6}
-\caption{Left: An illustrative comparison of individual components of the converged dominant eigenvector obtained
+\caption{Left: An illustrative comparison of individual squared components of the converged dominant eigenvector obtained
 using the two approaches discussed in~\ref{sub:grad} and~\ref{sub:gradfree}. Right: The quantity,  $\delta \hat{\mat{W}}_{1,j}^{(r)}$
@@
  \begin{center}
-  \includegraphics[width=0.48\textwidth]{./Figures/comp_ssp}
+  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp}
 \caption{An illustrative comparison of the SSPs generated using the gradient-based and the gradient-free approaches.}
@@
 of the model output resulted in a reasonably accurate estimation of the
-gradients in this case. In the following section, we will once again use both
-approaches to discover an active subspace in the high-dimensional 
-problem involving 33 uncertain inputs.
+gradients in this case. In the following section, we shift our focus to 
+the high-dimensional H$_2$/O$_2$ reaction kinetics application involving
+33 uncertain inputs.

### 2018-08-14 Paper/conc.tex
..

@@
  a pre-exponent and an activation energy. Hence, in theory, the evolution of the chemical system
- depends upon 38 inputs which is reasonably high-dimensional. In such case, conventional means
+ depends upon 38 inputs which is reasonably high-dimensional. Conventional means
  for uncertainty quantification such as those involving a surrogate model construction as well as
- sensitivity analysis would be computationally challenging and even prohibitive in some cases. 
+ sensitivity analysis would thus be computationally challenging and even prohibitive in some cases. 
  Hence, we focused our efforts on reducing the dimensionality of the problem by identifying important
  directions (active subspaces) in the parameter space such that the variability in the model output is 
- predominantly captured along these directions. Additionally, we demonstrate that the activity scores,
+ predominantly captured along these directions. Additionally, we demonstrated that the activity scores,
  computed using the components of the dominant eigenvectors can be used to approximate the
- DGSMs. Furthermore, we establish mathematical linkages between the different global sensitivity measures:
+ DGSMs. Furthermore, we established generalized mathematical linkages between the different global sensitivity measures:
  activity scores, DGSMs, and total Sobol' index which could be exploited to reduce computational effort 
@@
 model output with respect to the input, whereas, the second strategy, referred to as gradient-free implements a local
-linear approximation to the model output in order to estimate the gradient. The two strategies are shown to yield
-consistent results for the 19-dimensional problem whereby only the pre-exponents are considered to be uncertain.
-Additionally, the activity scores are also shown to be consistent with the screening metric estimates based on
-DGSMs in~\cite{Vohra:2018}. Note that an iterative procedure is adopted for implementing the two strategies
-for enhancing computational efficiency. 
-Motivated by these findings, the gradient-free approach is used for computing the active subspace in the high-dimensional
-case involving 33 uncertain inputs. Based on the prior marginals of the uncertain inputs, our results
-indicate that the uncertainty in the ignition delay is largely dependent on $A_1$, $A_9$, and $E_{a,1}$, although
- contributions from $A_{15}$ and $A_{17}$ are also found to be significant.
+linear approximation to the model output in order to estimate the gradient using a regression-based fit.
+The two strategies were shown to yield
+consistent results for the 19-dimensional problem whereby only the pre-exponents were considered to be uncertain.
+Additionally, the activity scores were also shown to be consistent with the screening metric estimates based on
+DGSMs in~\cite{Vohra:2018}. Note that an iterative procedure was adopted to enhance the computational efficiency
+associated with both strategies.
+
+Both strategies were further implemented to discover and compute the active
+subspace for the 33-dimensional problem. The gradient-based approach yielded a 1-dimensional active subspace
+and the activity scores which revealed that the uncertainty in the ignition delay is largely dependent on $A_1$, $A_9$, 
+$E_{a,1}$ and $E_{a,15}$, although contributions from $A_{15}$ and $A_{17}$ were also found to be significant. The activity
+scores were verified against the total Sobol' indices, computed using the 1-dimensional surrogate fit to the model response in
+the subspace. The 1-dimensional surrogate was verified for accuracy by evaluating the relative L-2
+norm of the error between its predictions and model evaluations at $10^4$ samples in the 33-dimensional input domain.
+Additionally, the surrogate was verified by comparing probability distributions (based on kernel density estimation) for
+the ignition delay based on its predictions and model evaluations. The two PDFs were observed to be nearly identical
+and consequently, their mean and standard deviations were found to be in excellent agreement. On the other hand,
+the gradient-free approach failed to capture the sensitivity towards the most important parameter, $E_{a,15}$. 
+We attribute this discrepancy to two potential sources of numerical errors: (1) Local linear approximation of the
+gradient of the model output, and (2) Scatter in the corresponding SSP (see Figure~\ref{fig:hd}) leading to numerical
+error incurred by the linear regression-fit in this case. The PDF was observed to be consistent with the true PDF in its
+estimate for the mode and the mean value of the ignition delay. However, it was observed to underestimate the uncertainty 
+associated with the ignition delay. Hence, the gradient-free approach might prove to be valuable in applications 
+involving intensive simulations (rendering the gradient-based approach prohibitive), and focused on a reasonable 
+approximation of the mean value and the uncertainty associated with the quantity of interest. Based on our findings, the 
+estimate for the mean is expected to be relatively more accurate.
 
@@
  calibration of the rate-controlling parameters in a Bayesian setting. Moreover, the activity scores could help
- identify important stages in a complex reaction mechanism and thereby guide development of a reduced-order
- surrogate. Furthermore, dimension reduction using active subspaces could be useful in studies aimed at
+ identify important stages in a complex reaction mechanism and thereby guide development of a simplified
+ mechanism. Furthermore, dimension reduction using active subspaces could be useful in studies aimed at
  investigating the inadequacy or model error associated with the reaction rate laws for such systems.

### 2018-08-16 Paper/intro.tex
...

@@
 %Constributions:
-%\begin{enumerate}
-%\item Iterative computation of active subspaces for efficient surrogate modeling 
-%\item Comparison of gradient-free and gradient based approaches.
-%\item Perspectives on activity scores as means of approximating DGSMs.
-%\item Application to challenging chemical kinetics problem. 
-%\end{enumerate}
-%Insert introduction here.
-
-% Outline
-%1. Significance of the H2/O2 reaction. Include the mechanism and a discussion on why
-%   is it important to study the kinetics? 
-%2. Literature survey of studies aimed at quantifying the uncertainty associated with
-%reaction kinetics (including my own work) and specifically for the H2/O2 reaction.
-%3. Focus of this work and associated challenges (include a table).
-%4. Key contributions.
-%5. Organization
 
@@
 constants associated with individual reactions. The rate constants are
-typically assumed to exhibit a certain correlation with temperature (e.g.
+typically assumed to exhibit a certain correlation with temperature (e.g.,
 Arrhenius-type). Hence, accurate specification of the rate-controlling
 parameters is critical to the fidelity of simulations. However, in practical
-applications, the parameters are either specified using expert knowledge or
+applications, these parameters are either specified using expert knowledge or
 estimated based on a regression-fit to a set of sparse and noisy
 data~\cite{Burnham:1987, Burnham:1988, Vohra:2011, Sarathy:2012}.
-Consequently, the underlying uncertainty associated with the parameters and
-therefore the simulation is typically ignored. Nevertheless, the inclusion of
+%Consequently, the underlying uncertainty associated with the parameters and
+%therefore the simulation is typically ignored. 
+Consequently, the inclusion of
 uncertainty due to corrupted calibration data, parametric uncertainty,
 and the discrepancy between predictions and experiments
-due to model inadequacy has gained prominence in recent
-years~\cite{Vohra:2014, Vohra:2017, Morrison:2018, Hantouche:2018}. 
-However, in complex mechanisms involving a large number of reactions,
-characterizing the propagation of uncertainty from a large set of inputs to
-the model output becomes remarkably challenging due to associated
-computational effort. 
-A major focus of this paper pertains to the implementation of a robust
-framework that aims to identify `important' directions in the input space
-that predominantly captures the variability in the model output. These
-directions, referred to as the \textit{active subspaces}~\cite{Constantine:2015},
-constitute the dominant
-eigenspace of a matrix derived from the gradient information of the model
-output with respect to an input. The active subspace methodology thus
-focuses on reducing the dimensionality of the problem, and hence the 
-computational effort associated with uncertainty propagation.
+due to model inadequacy has been subject of increased research efforts in recent
+years~\cite{Vohra:2014, Vohra:2017, Morrison:2018, Hantouche:2018}. \alennote{Need more references,
+acknowledging research by others.}
+
+In complex mechanisms involving a large number of reactions, characterizing the
+propagation of uncertainty from a large set of inputs to the model output is
+challenging due to the associated computational effort.  A major focus of this
+article is the implementation of a robust framework that aims to identify
+\emph{important} directions in the input space that predominantly capture the
+variability in the model output. These directions, which span the so called
+\emph{active subspace}~\cite{Constantine:2015}, are the dominant eigenvectors
+of a matrix derived from the gradient information of the model output with
+respect to an input. The active subspace methodology thus focuses on reducing
+the dimensionality of the problem, and hence the computational effort
+associated with uncertainty propagation.
 
 The application problem considered in this work is the
-H$_2$/O$_2$ reaction mechanism~\cite{Yetter:1991}. This mechanism has gained
+H$_2$/O$_2$ reaction mechanism from~\cite{Yetter:1991}. This mechanism has gained
 a lot of attention as a potential source of clean energy for
@@
 reactions including chain reactions, dissociation/recombination reactions, and
-formation and consumption of intermediate species. For each reaction in
-table~\ref{tab:kinetics}, the reaction rate is assumed to show an Arrhenius
+formation and consumption of intermediate species; see Table~\ref{tab:kinetics}. 
+For each reaction, the reaction rate is assumed to follow an Arrhenius
 correlation with temperature:
@@
 universal gas constant.  
-
+%
 \begin{table}[htbp]
@@
 \end{table}
-
+%
 The global reaction associated with the H$_2$/O$_2$ mechanism can
@@
 
-We are interested in investigating the impact of uncertainty in the
+We seek to understand the impact of uncertainty in the
 rate-controlling parameters, pre-exponent~($A_i$) and the activation
@@
 in~\cite{Yetter:1991}.  The total number of uncertain inputs is 33 which makes
-the present problem high-dimensional and computationally challenging. 
-To address this challenge, we focus our efforts on reducing the dimensionality
-of the problem by computing the aforementioned active subspace using
-an approach that involves computation of gradients of a model output with
-respect to individual parameters. Estimating the
+the present problem computationally challenging due to the large number of 
+uncertain parameters.  
+To address this challenge, we focus on reducing the dimensionality
+of the problem by computing the aforementioned active subspace.
+This involves repeated evaluations of the gradient of a model output with
+respect to the input parameters. Estimating the
 gradients can however be computationally challenging in high-dimensional settings. 
@@
 approach that involves a regression-based local linear approximation of
-the model output. It must be noted that since the gradient-free approach
+the model output. \alennote{I don't understand what the next sentence is 
+really trying to say here; I mean whatever we do involves approximating the
+gradient, the question is how rough the approximation is, right?} 
+It must be noted that since the gradient-free approach
 relies on approximating the model gradients, it is expected to be less
@@
 
-An alternate approach to dimension reduction involves computing the global sensitivity
-measures associated with uncertain inputs of a model. Depending upon the estimates of
-the sensitivity measures, only the important inputs are varied for the purpose of
-uncertainty quantification (UQ). The
-Sobol' index is a commonly used global sensitivity measure~\cite{Sobol:2001}. However, obtaining
-converged estimates of the Sobol' indices typically requires tens of thousands of model
-runs and is therefore impractical in situations involving intensive model simulations and/or a
-large number of inputs.
+An alternate approach to dimension reduction involves computing the global
+sensitivity measures associated with uncertain inputs of a model. Depending
+upon the estimates of the sensitivity measures, only the important inputs are
+varied for the purpose of uncertainty quantification (UQ). Sobol' indices are
+commonly used as global sensitivity measures~\cite{Sobol:2001}. However,
+obtaining converged estimates of the Sobol' indices typically requires tens of
+thousands of model runs and is therefore impractical in situations involving
+expensive model simulations and/or high-dimensional input parameter spaces.
 Recently, the derivative-based global sensitivity measures~(DGSMs) were
-developed to approximate the upper bound on the Sobol' indices with much fewer
-computations~\cite{Sobol:2009, Lamboni:2013}. In Section~\ref{sec:assa}, we show
-that the DGSMs can in fact be approximated using the components of the eigenvectors
-in the active subspace by estimating the so-called activity 
-scores~\cite{Diaz:2016,Constantine:2017}.
+developed to approximate upper bound on the Sobol' indices with much fewer
+computations~\cite{Sobol:2009, Lamboni:2013}. It was noted
+in~\cite{Diaz:2016,Constantine:2017} that DGSMs can be approximated by
+exploiting their links with active subspaces. This led to definition of the so
+called \emph{activity scores}. In Section~\ref{sec:assa}, we build on these
+ideas to provide a complete analysis of links between Sobol indices, DGSMs, and
+activity scores for functions of independent random inputs whos distribution
+law belongs to the class of Boltzmann probability measures. 
+It is worth mentioning that computing global sensitivity measures provides 
+important information about a model that go beyond dimension reduction. By 
+identifying parameters with significant impact on model uncertainty global 
+sensitivity analysis can guide model calibration and risk assessment.
 
-Main contributions of this paper are as follows: (1) Mathematical linkages between the
-global sensitivity measures: Total Sobol' index, DGSMs, and the activity scores are
-established; (2) An initial comparison of the active subspace and the global sensitivity
-measures, computed using the gradient-based and the gradient-free approaches in
-a 19-dimensional H$_2$/O$_2$ kinetics problem is provided. In this case, only the $A_i$'s are
-considered to be uncertain. (3) Active subspace discovery in the high-dimensional
-H$_2$/O$_2$ kinetics problem involving 33 uncertain inputs.
+The main contributions of this paper are as follows: 
+\begin{itemize}
+\item 
+Active subspace
+discovery in the high-dimensional H$_2$/O$_2$ kinetics problem involving 33
+uncertain inputs: \alennote{Say some interesting conclusion of the analysis;
+e.g., we discover a 1D subspace for a 33-dimensional problem using a modest computational 
+budget ...} The presented
+analysis can also guide practitioners in area of chemical kinetics on using the
+method of active subspaces as an efficient means of uncertainty propagation in
+chemical systems.  
+\item Comprehensive numerical investigation of the gradient-based and gradient-free approach: 
+we investigate the suitability of both approach
+for computing active subspaces, surrogate modeling, and approximating global
+sensitivity measures through a comprehensive set of numerical experiments. Our
+results reveal insight into the merits of the methods approach as well as
+their shortcomings.  
+\item Analysis of the links between  
+global sensitivity measures: 
+By connecting the recent advances in theory of variance based and
+derivative based global sensitivity analysis to active subspaces, we provide a complete analysis
+of the links between total Sobol' indices, DGSMs, and activity scores for a broad
+class of probability distributions. Our analysis is concluded by a result quantifying
+approximation errors incurred due to fixing unimportant parameters, deemed so by 
+computing their activity scores.
 
- The remaining of this paper is organized as follows. In section~\ref{sub:ac}, a brief theoretical background on
- the active subspace methodology is provided. In section~\ref{sub:gsa}, it is shown that the activity
- scores provide a reasonable approximation to the DGSMs especially in a high-dimensional setting. 
- Additionally, a relationship between the three global sensitivity measures, namely, the activity scores,
- DGSMs, and the total Sobol' indices is established. In section~\ref{sec:method}, a systematic framework
- for computing the active subspace using an iterative gradient-based approach is provided. Numerical results
- based on the gradient-free approach are compared with those obtained using the gradient-based approach.
- The active subspace is computed for the H$_2$/O$_2$ reaction kinetics problem whereby only
-the $A_i$'s are considered as uncertain. In this case, the two approaches are observed to yield consistent
-results. In section~\ref{sec:app}, the gradient-based approach is further implemented to compute the
-active subspace for the high-dimensional H$_2$/O$_2$ reaction kinetics problem. The resulting active
-subspace is found to be 1-dimensional thereby indicating enormous potential for computational savings
-in this application. The active subspace is also computed using the gradient-free approach. It is observed
-that although the uncertainty in the ignition delay is reasonably approximated, the gradient-free approach
-fails to capture the global sensitivity of all the uncertain inputs. Hence, a trade-off between efficiency and
-accuracy between the two approaches is observed in this case. 
-Finally, a summary and discussion based on our findings is included in section~\ref{sec:conc}.
+\end{itemize}
+
+\alennote{The structure of the paper is mostly okay; I have one recommendation:
+(i) make section 2.1 a separate section called preliminaries on active subspaces (or 
+something like that); (ii) make section 2.2. a section of its own, with a more descriptive
+section heading. The way things stand, people might take our section 2.2. as just background materials}
+This article is organized as follows. In section~\ref{sub:ac}, a brief
+theoretical background on the active subspace methodology is provided. In
+section~\ref{sub:gsa}, it is shown that the activity scores provide a
+reasonable approximation to the DGSMs especially in a high-dimensional setting.
+Additionally, a relationship between the three global sensitivity measures,
+namely, the activity scores, DGSMs, and the total Sobol' indices is
+established. In section~\ref{sec:method}, a systematic framework for computing
+the active subspace using an iterative gradient-based approach is provided.
+Numerical results based on the gradient-free approach are compared with those
+obtained using the gradient-based approach.  The active subspace is computed
+for the H$_2$/O$_2$ reaction kinetics problem whereby only the $A_i$'s are
+considered as uncertain. In this case, the two approaches are observed to yield
+consistent results. In section~\ref{sec:app}, the gradient-based approach is
+further implemented to compute the active subspace for the high-dimensional
+H$_2$/O$_2$ reaction kinetics problem. The resulting active subspace is found
+to be 1-dimensional thereby indicating enormous potential for computational
+savings in this application. The active subspace is also computed using the
+gradient-free approach. It is observed that although the uncertainty in the
+ignition delay is reasonably approximated, the gradient-free approach fails to
+capture the global sensitivity of all the uncertain inputs. Hence, a trade-off
+between efficiency and accuracy between the two approaches is observed in this
+case.  Finally, a summary and discussion based on our findings is included in
+section~\ref{sec:conc}.

### 2018-08-17 Paper/assa.tex
minor ...
work in progress ...

@@
-\section{Active subspaces and sensitivity analysis}
-\label{sec:assa}
+%\section{Active subspaces and sensitivity analysis}
+%\label{sec:assa}
 %In this section, we provide key elements of the active subspace methodology and
@@
 
-\subsection{Active subspaces}
+\section{Active subspaces}
 \label{sub:ac}
@@
 \item For each $\bm{\xi}_i$, compute $\mat{W}_1^\top\bm{\xi}_i$. Note that
- $G(\mat{W}_1^\top \bm{\xi})$ $\approx$ $f(\bm{\xi}_i)$.
+ $G(\mat{W}_1^\top \bm{\xi}_i)$ $\approx$ $f(\bm{\xi}_i)$.
 \item Fit a regression surface, $\tilde{G}(\mat{W}_1^\top\bm{\xi}_i)$ $\approx$ 
-$G(\mat{W}_1^\top \bm{\xi})$.
+$G(\mat{W}_1^\top \bm{\xi}_i)$.
 \item Overall approximation, $f(\bm{\xi})$ $\approx$ $\tilde{G}(\mat{W}_1^\top\bm{\xi})$.
@@
 
-\subsection{Global sensitivity measures}
-
+\section{GSA measures and their links with active subspaces}
 \label{sub:gsa}
@@
 Here $\pi$ is the joint PDF of $\xi$. 
-Let $\mat{C} = \E{\nabla f \nabla f^\top}$ and note that
-\[
-   \nu_i(f) = \vec{e}_i^\top \mat{C} \vec{e}_i, 
-\]
-where $\vec{e}_i$ is the $i$th coordinate vector in $\R^\Np$.
-This also shows that $\sum_{j=1}^\Np \nu_i(f) = \trace(\mat{C})$.
-We consider the spectral decomposition of $\mat{C}$, given by  
+Let $\mat{C} = \E{\nabla f \nabla f^\top}$; consider the spectral decomposition,
 $\mat{C} = \sum_{k=1}^\Np \lambda_k \vec{u}_k \vec{u}_k^\top$, where
@@
 order, and
-$\vec{u}_k$ are the corresponding (orthonormal) eigenvectors.
-We note that,
+$\vec{u}_k$ are the corresponding (orthonormal) eigenvectors. The following
+result provides a representation of DGSMs in terms of the 
+spectral representation of $\mat{C}$: 
+\begin{lemma}
+We have
+$\nu_i(f) = \sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2$.
+\end{lemma}
+\begin{proof}
+Note that $\nu_i(f) = \vec{e}_i^\top \mat{C} \vec{e}_i$,  
+where $\vec{e}_i$ is the $i$th coordinate vector in $\R^\Np$, $i = 1, \ldots, \Np$.
+Therefore,
 \begin{equation}\label{equ:spectral_DGSM}
 \nu_i(f) = \vec{e}_i^T \Big(\sum_{k=1}^\Np \lambda_k \vec{u}_k \vec{u}_k^\top\Big) \vec{e}_i
- = \sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2, 
+ = \sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2. 
 \end{equation}
-which gives a spectral representation for the DGSMs. 
+\end{proof}
 In the case where the eigenvalues decay rapidly to zero, we can obtain
@@
 \begin{proposition}\label{prp:dgsm_bound} 
-For $1 \leq r \leq \Np$,
+For $1 \leq r < \Np$,
 \[
@@
 where for each $i$, $C_i$ is an appropriate \emph{Poincar\'{e}} constant
-that depends on the distribution of $\xi_i$.
+that depends on the distribution of $\xi_i$; see~\cite{Lamboni:2013}.
 This provides a strong theoretical basis for using DGSMs to identify 
-unimportant inputs. Moreover, 
-in~\cite{Vohra:2018} we introduced the normalized screening metric
-\[
-   \tilde{\nu}_i(f) = \frac{C_i \nu_i(f)}{\sum_{i=1}^\Np C_i \nu_i(f)},
-\]
-for detecting unimportant inputs. 
-Note that if $\xi_i$ are iid, the $C_i$'s will cancel and the normalized screening
-metric reduces to 
-\[
-    \tilde{\nu}_i(f) = \frac{\nu_i(f)}{\sum_{i=1}^\Np \nu_i(f)} 
-      = \frac{\vec{e}_i^T \mat{C} \vec{e}_i}{\trace(\mat{C})} 
-      = \frac{\sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2}{\sum_{k = 1}^\Np \lambda_k}.
-\]
-
-We can also bound the normalized DGSMs using activity scores; it is easy to see  
-\[
-\tilde{\nu}_i(f) \leq 
-\frac{ C_i \big(\act{f}{i}{r} + \lambda_{r+1}\big)}{\sum_{i=1}^\Np C_i \act{f}{i}{r}}
-=\frac{C_i \act{f}{i}{r}}{\sum_{i=1}^\Np C_i \act{f}{i}{r}} + \kappa_i \lambda_{r+1}, 
-\]
-with $\kappa_i = C_i / (\sum_i C_i \act{f}{i}{r})$. 
-In the case where where $\lambda_{r+1} \approx 0$, 
-this motivates definition of
-normalized activity scores
-\[
-   \actt{f}{i}{r} =  \frac{C_i \act{f}{i}{r}}{\sum_{i=1}^\Np C_i \act{f}{i}{r}}.
-\]
+unimportant inputs.
 
@@
 unimportant inputs at nominal values. Suppose $\xi_i$ is
-deemed unimportant as a result of a global sensitivity analysis
+deemed unimportant as a result of a global sensitivity analysis (GSA)
 based on the activity scores; we want to estimate
@@
 \begin{theorem}
-We have $\Ez{ \mathcal{E}(z)} \leq 2C_i\big(\act{f}{i}{r} + \lambda_{r+1}\big)/{\V(f)} $.
+We have $\Ez{ \mathcal{E}(z)} \leq 2C_i\big(\act{f}{i}{r} + \lambda_{r+1}\big)/{\V(f)}$, 
+for $1 \leq r < \Np$.
 \end{theorem}
@@
 \end{proof}
+
+In~\cite{Vohra:2018} the following normalized screening metric
+\[
+   \tilde{\nu}_i(f) = \frac{C_i \nu_i(f)}{\sum_{i=1}^\Np C_i \nu_i(f)},
+\]
+was shown to be useful for detecting unimportant inputs.
+
+We can also bound the normalized DGSMs using activity scores; it is easy to see  
+\[
+\tilde{\nu}_i(f) \leq 
+\frac{ C_i \big(\act{f}{i}{r} + \lambda_{r+1}\big)}{\sum_{i=1}^\Np C_i \act{f}{i}{r}}
+=\frac{C_i \act{f}{i}{r}}{\sum_{i=1}^\Np C_i \act{f}{i}{r}} + \kappa_i \lambda_{r+1}, 
+\]
+with $\kappa_i = C_i / (\sum_i C_i \act{f}{i}{r})$. 
+In the case where where $\lambda_{r+1} \approx 0$, 
+this motivates definition of
+normalized activity scores
+\[
+   \actt{f}{i}{r} =  \frac{C_i \act{f}{i}{r}}{\sum_{i=1}^\Np C_i \act{f}{i}{r}}.
+\]
+
+\begin{remark}
+Note that if $\xi_i$ are iid, the $C_i$'s in the definition of the 
+normalized screening metric will cancel and 
+\[
+    \tilde{\nu}_i(f) = \frac{\nu_i(f)}{\sum_{i=1}^\Np \nu_i(f)} 
+      = \frac{\vec{e}_i^T \mat{C} \vec{e}_i}{\trace(\mat{C})} 
+      = \frac{\sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2}{\sum_{k = 1}^\Np \lambda_k}.
+\]
+\end{remark}

### 2018-08-17 Paper/assa.tex
sec 2 ...

@@
-%\section{Active subspaces and sensitivity analysis}
-%\label{sec:assa}
-%In this section, we provide key elements of the active subspace methodology and
-%provide theoretical connections between the global sensitivity measures
-%(DGSMs), activity scores, and the total Sobol' indices.
-
 \section{Active subspaces}
 \label{sub:ac}
-As mentioned in the introduction, the active subspace is a low-dimensional subspace
+
+Herein, we use a random vector 
+$\vec\xi \in \Omega\in\mathbb{R}^{N_p}$ to parameterize model uncertainties, 
+where $N_p$ is number of uncertain inputs.
+In practical computations, the \emph{canonical} variables $\xi_i$ are  
+mapped to physical ranges meaningful in a given mathematical model. 
+As mentioned in the introduction, an active subspace is a low-dimensional subspace
 that constitutes important directions in a model's input
-space~\cite{Constantine:2015}. The effective variability in the model output, $f$
-due to uncertain inputs, $\bm{\theta}$ in the physical space, is predominantly captured
-along these directions. Each component of $\bm{\theta}$
-is mapped to a corresponding value in $\Omega\in\mathbb{R}^{N_p}$
-($N_p$: number of uncertain inputs)
-using a linear transformation.  The set of canonical random variables, corresponding
-to $\theta_i$'s, in $\Omega$ is referred to as $\bm{\xi}$.
-Hence, from this point onwards, we denote $f$ as $f(\bm{\xi})$.
-Mathematically, for a given $f$, and a set of inputs, $\bm{\xi}$,
-these directions are the eigenvectors of a positive
-semi-definite matrix, $\mat{C}$, defined as follows: 
+parameter space~\cite{Constantine:2015}. The effective variability in a model output $f$
+due to uncertain inputs is predominantly captured
+along these directions. 
+The directions spanning the active subspace are the eigenvectors of the positive
+semidefinite matrix 
 %
 \be
-\mat{C} = \int_\Omega (\nabla_{\bm{\xi}}f)(\nabla_{\bm{\xi}}f)^\top d\Pi_{\bm{\xi}}
+\mat{C} = \int_\Omega (\nabla_{\vec{\xi}}f)(\nabla_{\vec{\xi}}f)^\top \mu(d\vec\xi), 
 \label{eq:C}
@@
 %
-where 
-$d\Pi_{\bm{\xi}}$ = $\pi(\bm{\xi})d\bm{\xi}$, where $\pi(\bm{\xi})$ is the joint probability
-distribution of $\bm{\xi}$. Hence, $f$ is considered to be continuous 
-and differentiable in the input domain. 
-%Hence, it is possible that a given $f(\bm{\xi})$ might not admit an active
+with 
+$\mu(d\vec{\xi}) = \pi(\vec{\xi})d\vec{\xi}$, where $\pi(\vec{\xi})$ is the joint probability
+distribution of $\vec{\xi}$. Herein, $f$ is assumed to be a function with continuous partial 
+derivatives with respect to the input parameters. 
+%Hence, it is possible that a given $f(\vec{\xi})$ might not admit an active
 %subspace. However, it is of remarkable interest to investigate if one exists
@@
 %problem and hence the associated computational effort.
-Here,
-$\nabla_{\bm{\xi}}f$ denotes the gradient vector. 
+%Here,
+%$\nabla_{\vec{\xi}}f$ denotes the gradient vector. 
 %with individual components
@@
 Since $\mat{C}$ is symmetric and
-positive semi-definite, it admits a spectral decomposition:
+positive semidefinite, it admits a spectral decomposition:
 %
 \be
-\mat{C} = \mat{W}\mat{\Lambda}\mat{W}^\top
+\mat{C} = \mat{W}\mat{\Lambda}\mat{W}^\top.
 \ee
 %
-where $\mat{\Lambda}$ = diag($\lambda_1,\ldots,\lambda_{N_p}$) with
+Here $\mat{\Lambda}$ = diag($\lambda_1,\ldots,\lambda_{N_p}$) with the eigenvalues
 $\lambda_i$'s sorted in descending order
 \[
-     \lambda_1 \geq \lambda_2 \geq \cdots \geq \lambda_\Np \geq 0.
+     \lambda_1 \geq \lambda_2 \geq \cdots \geq \lambda_\Np \geq 0,
 \] 
+and $\mat{W}$ has the eigenvectors $\vec{w}_1, \ldots, \vec{w}_\Np$ as its columns.
 The eigenpairs are partitioned about the $i^{\text{th}}$ eigenvalue such that
-the ratio, $\left(\frac{\lambda_i}{\lambda_{i+1}}\right)$ $\gg$ 1 as follows:
+the ratio, $\lambda_i/\lambda_{i+1}\gg 1$ as follows:
 \be
  \mat{W} = [\mat{W}_1~\mat{W}_2],~~\mat{\Lambda} = \begin{bmatrix}\mat{\Lambda}_1 & \\  &
-  \mat{\Lambda}_2 \\
+  \mat{\Lambda}_2. 
 \end{bmatrix}
@@
  %
- where $\mat{W}_1$ comprises the dominant eigenspace of $\mat{C}$ regarded as
-the active subspace, and $\mat{\Lambda}_1$ is the corresponding set of
-eigenvalues. Once the active subspace is computed, dimension reduction is
-accomplished by transforming the set of canonical random variables, $\bm{\xi}$
-into a reduced set, $\mat{W}_1^\top \bm{\xi}$, referred to as the set of active
-variables. The number of active variables is equal to the number
-of dominant eigenvectors. The model output in the original space, $f(\bm{\xi})$,
-is approximated  as $G(\mat{W}_1^\top \bm{\xi})$ in the active subspace.
-Hence, $f$ is essentially evaluated at the projection of $\bm{\xi}$ on the
-column space of $\mat{W}_1$. We could confine UQ to the inputs in the
+The columns of $\mat{W}_1$ span the dominant eigenspace of $\mat{C}$ and
+defines the active subspace, and $\mat{\Lambda}_1$ is a diagonal matrix with
+the corresponding set of eigenvalues on its diagonal. Once the active subspace
+is computed, dimension reduction is accomplished by transforming the parameter
+vector $\vec\xi$ into 
+$\vec{y} = \mat{W}_1^\top\vec{\xi}$. The elements of $\vec{y}$ are 
+referred to as the set of active variables. The number of active
+variables is equal to the number of dominant eigenvectors. Consider the function
+\[
+    G(\vec{y}) = f(\mat{W}\vec{y}).
+\]
+The model output in the original space, $f(\vec{\xi})$,
+is approximated by $G(\mat{W}_1^\top \vec{\xi})$ in the active subspace~\cite{Constantine:2015}.
+%Hence, $f$ is essentially evaluated at the projection of $\vec{\xi}$ on the
+%column space of $\mat{W}_1$. 
+We could confine uncertainty analysis to the inputs in the
 active subspace whose dimension is typically much smaller (in applications that
 admit such a subspace) than the dimension of the original input space. To further
-expedite UQ, one could easily fit a regression-surface to $G$ using the 
-following sequence of steps outlined in~\cite{Constantine:2015} (see chapter 4).
+expedite uncertainty analysis, one could fit a regression surface to $G$ using the 
+following sequence of steps, as outlined in~\cite[chapter 4]{Constantine:2015}. 
 \begin{enumerate}
-\item Consider a set of input-output pairs, $\{\bm{\xi}_i, f(\bm{\xi}_i)\}$.
-\item For each $\bm{\xi}_i$, compute $\mat{W}_1^\top\bm{\xi}_i$. Note that
- $G(\mat{W}_1^\top \bm{\xi}_i)$ $\approx$ $f(\bm{\xi}_i)$.
-\item Fit a regression surface, $\tilde{G}(\mat{W}_1^\top\bm{\xi}_i)$ $\approx$ 
-$G(\mat{W}_1^\top \bm{\xi}_i)$.
-\item Overall approximation, $f(\bm{\xi})$ $\approx$ $\tilde{G}(\mat{W}_1^\top\bm{\xi})$.
+\item Consider a set of input-output pairs, $\{\vec{\xi}_i, f(\vec{\xi}_i)\}$, $i = 1, \ldots, N$.
+\item For each $\vec{\xi}_i$, compute $\vec{y}_i = \mat{W}_1^\top\vec{\xi}_i$. Note that
+ $G(\mat{W}_1^\top \vec{\xi}_i)$ $\approx$ $f(\vec{\xi}_i)$.
+\item Use $\{\vec{y}_i\}_{i =1}^N$ to compute a 
+regression surface, $\tilde{G}(\vec{y})\approx 
+G(\vec{y})$.
+\item Overall approximation, $f(\vec{\xi})$ $\approx$ $\tilde{G}(\mat{W}_1^\top\vec{\xi})$.
 \end{enumerate}
 
-In practice, the matrix, $\mat{C}$ is typically
-approximated using pseudo-random sampling techniques such as Monte Carlo and
-Latin hypercube sampling. The integral in~\eqref{eq:C} is replaced with a
-summation as follows:
+In practice, the matrix $\mat{C}$ defined in~\eqref{eq:C} is 
+approximated using pseudo-random sampling techniques such as Monte Carlo or
+Latin hypercube sampling:
+%The integral in~\eqref{eq:C} is replaced with a
+%summation as follows:
  %
@@
  \mat{C}\approx \hat{\mat{C}} = \frac{1}{N}\sum\limits_{i=1}^{N} 
- (\nabla_{\bm{\xi}}f(\bm{\xi}_i))(\nabla_{\bm{\xi}}f(\bm{\xi}_i))^\top
+ (\nabla_{\vec{\xi}}f(\vec{\xi}_i))(\nabla_{\vec{\xi}}f(\vec{\xi}_i))^\top
  = \hat{\mat{W}}\hat{\mat{\Lambda}}\hat{\mat{W}}^\top
@@
 $\hat{\mat{C}}$ scales with the number of samples, $N$. Hence, an iterative
-computational approach is adopted in this work to ensure that $N$ is sufficiently
-large so that the dominant eigenvectors in $\hat{\mat{W}}_1$ are approximated
-with reasonable accuracy as discussed later in Section~\ref{sec:method}.
+computational approach is adopted in this work to gradually increase  
+$N$ until the dominant eigenpairs are approximated
+with sufficient accuracy; see Section~\ref{sec:method}.
 
 %%It has been shown that the accuracy of approximated dominant eigenspace,
-%%$\hat{\bm{w}}_1$ is inversely proportional to the difference between the
-%%smallest eigenvalue in $\hat{\bm{\Lambda}}_1$ and the largest eigenvalue in
-%%$\hat{\bm{\Lambda}}_2$~\cite{Constantine:2014}. 
+%%$\hat{\vec{w}}_1$ is inversely proportional to the difference between the
+%%smallest eigenvalue in $\hat{\vec{\Lambda}}_1$ and the largest eigenvalue in
+%%$\hat{\vec{\Lambda}}_2$~\cite{Constantine:2014}. 
 %%Components of the eigenvectors in the active subspace could be used for
@@
 T_i(f) = 1 - 
-\frac{\V[\mathbb{E}(f|\bm{\xi}_{\sim i})]}{\V(f)},
+\frac{\V[\mathbb{E}(f|\vec{\xi}_{\sim i})]}{\V(f)},
 \label{eq:total}
@@
 %
-where $\bm{\xi}_{\sim i}$ is a vector of all the inputs $\xi_i$ except the  $i^\text{th}$ entry, and $\V$ denotes the variance. The total Sobol' index accounts
+where $\vec{\xi}_{\sim i}$ is the input parameter vector with the  
+$i^\text{th}$ entry removed, and $\V$ denotes the variance. The total Sobol' index accounts
 for the contribution of a given input to the variability in the output by itself
@@
 thousands of model runs and is therefore prohibitive in the case of
-compute-intensive applications. The DGSMs~\cite{Sobol:2009} provide a means for
-approximating the upper bound on $T_i(f)$ with fewer computations; see 
+compute-intensive applications. Derivative based 
+global sensitivity measures (DGSMs)~\cite{Sobol:2009} provide a means for
+approximating informative upper bounds on $T_i(f)$ with fewer computations; see 
 also~\cite{Vohra:2018}.
@@
 \]
-Here $\pi$ is the joint PDF of $\xi$. 
-Let $\mat{C} = \E{\nabla f \nabla f^\top}$; consider the spectral decomposition,
-$\mat{C} = \sum_{k=1}^\Np \lambda_k \vec{u}_k \vec{u}_k^\top$, where
-$\lambda_i$ are the (non-negative) eigenvalues of $\mat{C}$, in descending
-order, and
-$\vec{u}_k$ are the corresponding (orthonormal) eigenvectors. The following
+Here $\pi$ is the joint PDF of $\vec\xi$. 
+Let $\mat{C}$ be as in~\eqref{eq:C}, and consider the spectral decomposition written 
+as $\mat{C} = \sum_{k=1}^\Np \lambda_k \vec{w}_k \vec{w}_k^\top$. Herein, we use the notation 
+$\ip{\cdot}{\cdot}$ for the Euclidean inner product.
+% where
+%$\lambda_i$ are the (non-negative) eigenvalues of $\mat{C}$, in descending
+%order, and
+%$\vec{w}_k$ are the corresponding (orthonormal) eigenvectors. 
+The following
 result provides a representation of DGSMs in terms of the 
@@
 We have
-$\nu_i(f) = \sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2$.
+$\nu_i(f) = \sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2$.
 \end{lemma}
@@
 Therefore,
-\begin{equation}\label{equ:spectral_DGSM}
-\nu_i(f) = \vec{e}_i^T \Big(\sum_{k=1}^\Np \lambda_k \vec{u}_k \vec{u}_k^\top\Big) \vec{e}_i
- = \sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2. 
-\end{equation}
+\[
+\nu_i(f) = \vec{e}_i^T \Big(\sum_{k=1}^\Np \lambda_k \vec{w}_k \vec{w}_k^\top\Big) \vec{e}_i
+ = \sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2. \qedhere 
+\]
 \end{proof}
@@
 \[
-   \act{f}{i}{r} =  \sum_{k=1}^r \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2,
+   \act{f}{i}{r} =  \sum_{k=1}^r \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2,
    \quad j = 1, \ldots, \Np, \quad r \leq \Np.
@@
 %\[
-%\alpha_i(f;r) = \sum_{k=1}^{r}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2 \leq \sum_{k=1}^{N_p}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2 = \nu_i(f), \quad \quad i = 1,\ldots,N_p, \quad r \leq N_p
+%\alpha_i(f;r) = \sum_{k=1}^{r}\lambda_k \langle \vec{e}_i, \vec{w}_k \rangle^2 \leq \sum_{k=1}^{N_p}\lambda_k \langle \vec{e}_i, \vec{w}_k \rangle^2 = \nu_i(f), \quad \quad i = 1,\ldots,N_p, \quad r \leq N_p
 %\]
@@
 %\[
-%0 \leq  \sum_{k=1}^{N_p}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2 - \sum_{k=1}^{r}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2,  \quad \quad i = 1,\ldots,N_p
+%0 \leq  \sum_{k=1}^{N_p}\lambda_k \langle \vec{e}_i, \vec{w}_k \rangle^2 - \sum_{k=1}^{r}\lambda_k \langle \vec{e}_i, \vec{w}_k \rangle^2,  \quad \quad i = 1,\ldots,N_p
 %\]
@@
 %\begin{aligned}
-%\nu_i(f) = \sum_{k=1}^{N_p}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2 = \sum_{k=1}^{r}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2 + \sum_{k=r+1}^{N_p}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2  \\
-%= \alpha_i(f;r) + \sum_{k=r+1}^{N_p}\lambda_k \langle \vec{e}_i, \vec{u}_k \rangle^2 \leq \alpha_i(f;r) + \lambda_{r+1} \sum_{k=r+1}^{N_p} \langle \vec{e}_i, \vec{u}_k \rangle^2,  \quad \quad i = 1,\ldots,N_p
+%\nu_i(f) = \sum_{k=1}^{N_p}\lambda_k \langle \vec{e}_i, \vec{w}_k \rangle^2 = \sum_{k=1}^{r}\lambda_k \langle \vec{e}_i, \vec{w}_k \rangle^2 + \sum_{k=r+1}^{N_p}\lambda_k \langle \vec{e}_i, \vec{w}_k \rangle^2  \\
+%= \alpha_i(f;r) + \sum_{k=r+1}^{N_p}\lambda_k \langle \vec{e}_i, \vec{w}_k \rangle^2 \leq \alpha_i(f;r) + \lambda_{r+1} \sum_{k=r+1}^{N_p} \langle \vec{e}_i, \vec{w}_k \rangle^2,  \quad \quad i = 1,\ldots,N_p
 %\end{aligned}
 %\]
-%The eigenvectors $\vec{u}_k$ are orthonormal so they all have length 1.
-Also note that for every $x \in \R^n$ we have \[ \norm{x}^2 = \sum_{k=1}^{n} \langle \vec{x}, \vec{u}_k \rangle^2\] This is known as Parseval's identity.
+%The eigenvectors $\vec{w}_k$ are orthonormal so they all have length 1.
+Also note that for every $x \in \R^n$ we have \[ \norm{x}^2 = \sum_{k=1}^{n} \langle \vec{x}, \vec{w}_k \rangle^2\] This is known as Parseval's identity.
 %In particular in this case $\vec{x} = \vec{e}_i \in \R^{N_p}$ so
-%\[1 = \norm{\vec{e}_i}^2 = \sum_{k=r+1}^{N_p} \langle \vec{e}_i, \vec{u}_k \rangle^2\]
+%\[1 = \norm{\vec{e}_i}^2 = \sum_{k=r+1}^{N_p} \langle \vec{e}_i, \vec{w}_k \rangle^2\]
 %Finally we write:
@@
 
-Note that, $\nu_i(f) - \act{f}{i}{r}= \sum_{k=r+1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2 \geq 0$,
+Note that, $\nu_i(f) - \act{f}{i}{r}= \sum_{k=r+1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2 \geq 0$,
 which gives the first inequality. To see the upper bound, we note,
 \[
-   \sum_{k=r+1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2 \leq \lambda_{r+1} \sum_{k=r+1}^\Np \ip{\vec{e}_i}{\vec{u}_k}^2
+   \sum_{k=r+1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2 \leq \lambda_{r+1} \sum_{k=r+1}^\Np \ip{\vec{e}_i}{\vec{w}_k}^2
    \leq \lambda_{r+1}. 
@@
 $1 = \|\vec{e}_i\|_2^2 = 
-\sum_{k = 1}^\Np \ip{\vec{e}_i}{\vec{u}_k}^2 
-\geq \sum_{k=r+1}^\Np \ip{\vec{e}_i}{\vec{u}_k}^2$.
+\sum_{k = 1}^\Np \ip{\vec{e}_i}{\vec{w}_k}^2 
+\geq \sum_{k=r+1}^\Np \ip{\vec{e}_i}{\vec{w}_k}^2$.
 \end{proof} 
@@
 that depends on the distribution of $\xi_i$; see~\cite{Lamboni:2013}.
-This provides a strong theoretical basis for using DGSMs to identify 
+For instance, if $\xi_i$ is uniformly distributed on $[-1, 1]$, then $C_i = 2/\pi^2$; and in the 
+case $\xi_i$ is normally distributed with variance $\sigma_i^2$, then $C_i = \sigma_i^2$. 
+The bound~\eqref{equ:sobol_bound} provides a strong theoretical basis for using DGSMs to identify 
 unimportant inputs.
@@
 the approximation error that occurs once $\xi_i$ 
-is fixed at its nominal value.
+is fixed at a nominal value.
 %Let $\I \subset \{1, \ldots,
@@
 
-Let $\vec\xi$ be given and let $z$ be a nominal value for $\xi_i$;  we define 
-$\vec{y}^z(\vec\xi)$ as the vector with entries $y^z_j = \xi_j$ for $j \neq i$
-and $y^z_i = z_i$.
+Let $\vec\xi$ be given and let $z$ be a nominal value for $\xi_i$.  
+%we define 
+%$\vec{y}^z(\vec\xi)$ as the vector with entries $y^z_j = \xi_j$ for $j \neq i$
+%and $y^z_i = z$.
 We consider the \emph{reduced} model, 
@@
 \[
-f^{z}(\vec\xi) = f(\vec{y}^z(\vec\xi))
+f^{z}(\vec\xi) = f(\xi_1, \xi_2, \ldots, \xi_{i-1}, z, \xi_{i+1}, \ldots, \xi_\Np),
 \] 
-We consider the following error indicator.
+and consider the following error indicator:
 \[
@@
 was shown to be useful for detecting unimportant inputs. 
-
-We can also bound the normalized DGSMs using activity scores; it is easy to see  
+We can also bound the normalized DGSMs using activity scores as follows. It is straightforward to see  
+that
 \[
@@
       = \frac{\vec{e}_i^T \mat{C} \vec{e}_i}{\trace(\mat{C})} 
-      = \frac{\sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{u}_k}^2}{\sum_{k = 1}^\Np \lambda_k}.
+      = \frac{\sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2}{\sum_{k = 1}^\Np \lambda_k}.
+\]
+In this case, the normalized activity scores reduce to
+\[
+   \actt{f}{i}{r} =  \frac{\act{f}{i}{r}}{\sum_{i=1}^\Np \act{f}{i}{r}}.
 \]

### 2018-08-17 Paper/intro.tex
sec 2 ...

@@
 
-\alennote{The structure of the paper is mostly okay; I have one recommendation:
-(i) make section 2.1 a separate section called preliminaries on active subspaces (or 
-something like that); (ii) make section 2.2. a section of its own, with a more descriptive
-section heading. The way things stand, people might take our section 2.2. as just background materials}
+\alennote{Okay, I reorganized the sections a bit (new sec 2, 3); please read the following and
+make sure it reflects the contents of the sections correctly.}
 This article is organized as follows. In section~\ref{sub:ac}, a brief

### 2018-08-17 Paper/assa.tex
very minor edits

@@
 $\vec\xi \in \Omega\in\mathbb{R}^{N_p}$ to parameterize model uncertainties, 
-where $N_p$ is number of uncertain inputs.
-In practical computations, the \emph{canonical} variables $\xi_i$ are  
+where $N_p$ is the number of uncertain inputs.
+In practical computations, the \emph{canonical} variables $\xi_i, i=1,\ldots ,N_p$, are  
 mapped to physical ranges meaningful in a given mathematical model. 
@@
 \begin{enumerate}
-\item Consider a set of input-output pairs, $\{\vec{\xi}_i, f(\vec{\xi}_i)\}$, $i = 1, \ldots, N$.
+\item Consider a set of input-output pairs, $\{\vec{\xi}_i, f(\vec{\xi}_i)\}$, $i = 1, \ldots, N$. {\hayleynote {Do we need to define what N is?}}
 \item For each $\vec{\xi}_i$, compute $\vec{y}_i = \mat{W}_1^\top\vec{\xi}_i$. Note that
@@
 In the case where the eigenvalues decay rapidly to zero, we can obtain
-accurate approximations of $\nu_i(f)$ by Truncating the summation: 
+accurate approximations of $\nu_i(f)$ by truncating the summation: 
 \newcommand{\act}[3]{\nu_{{#2},{#3}}({#1})}
@@
 high-dimensional parameter spaces in which 
-the eigenvalues $\lambda_i$ decay rapidly to zero; in 
+the eigenvalues $\lambda_i, i=1,\ldots,N_p$, decay rapidly to zero; in 
 such cases, this result implies that  $\nu_i(f) \approx \act{f}{i}{r}$,

### 2018-08-17 Paper/assa.tex
very minor edits
sec 2, first pass ...

@@
 and $\mat{W}$ has the eigenvectors $\vec{w}_1, \ldots, \vec{w}_\Np$ as its columns.
-The eigenpairs are partitioned about the $i^{\text{th}}$ eigenvalue such that
-the ratio, $\lambda_i/\lambda_{i+1}\gg 1$ as follows:
+The eigenpairs are partitioned about the $r$th eigenvalue such that
+$\lambda_r/\lambda_{r+1}\gg 1$, 
 \be
@@
  %
-The columns of $\mat{W}_1$ span the dominant eigenspace of $\mat{C}$ and
-defines the active subspace, and $\mat{\Lambda}_1$ is a diagonal matrix with
-the corresponding set of eigenvalues on its diagonal. Once the active subspace
+The columns of $\mat{W}_1 = 
+\begin{bmatrix} \vec{w}_r \cdots \vec{w}_r\end{bmatrix}$ 
+span the dominant eigenspace of $\mat{C}$ and
+define the active subspace, and $\mat{\Lambda}_1$ is a diagonal matrix with
+the corresponding set of eigenvalues, $\lambda_1, \ldots, \lambda_r$, on its diagonal. 
+Once the active subspace
 is computed, dimension reduction is accomplished by transforming the parameter
 vector $\vec\xi$ into 
-$\vec{y} = \mat{W}_1^\top\vec{\xi}$. The elements of $\vec{y}$ are 
-referred to as the set of active variables. The number of active
-variables is equal to the number of dominant eigenvectors. Consider the function
+$\vec{y} = \mat{W}_1^\top\vec{\xi} \in \R^r$. The elements of $\vec{y}$ are 
+referred to as the set of active variables. 
+%The number $r$ of active
+%variables is equal to the number of dominant eigenvectors.
+
+Consider the function
 \[
-    G(\vec{y}) = f(\mat{W}\vec{y}).
+    G(\vec{y}) = f(\mat{W}_1\vec{y}), \quad \vec{y} \in \R^r.
 \]
-The model output in the original space, $f(\vec{\xi})$,
-is approximated by $G(\mat{W}_1^\top \vec{\xi})$ in the active subspace~\cite{Constantine:2015}.
+Following~\cite{Constantine:2015}, we use the approximation $f(\vec{\xi}) \approx 
+G(\mat{W}_1^\top \vec{\xi})$. 
+That is, the model output $f(\vec\xi)$, in the original parameter space,
+is approximated by $G(\mat{W}_1^\top \vec{\xi})$ in the active subspace.
 %Hence, $f$ is essentially evaluated at the projection of $\vec{\xi}$ on the
@@
 \begin{enumerate}
-\item Consider a set of input-output pairs, $\{\vec{\xi}_i, f(\vec{\xi}_i)\}$, $i = 1, \ldots, N$. {\hayleynote {Do we need to define what N is?}}
+\item Consider a given set of $N$ 
+data points, $\big(\vec{\xi}_i, f(\vec{\xi}_i)\big)$, $i = 1, \ldots, N$. 
 \item For each $\vec{\xi}_i$, compute $\vec{y}_i = \mat{W}_1^\top\vec{\xi}_i$. Note that
- $G(\mat{W}_1^\top \vec{\xi}_i)$ $\approx$ $f(\vec{\xi}_i)$.
-\item Use $\{\vec{y}_i\}_{i =1}^N$ to compute a 
-regression surface, $\tilde{G}(\vec{y})\approx 
+ $G(\vec{y}_i)$ $\approx$ $f(\vec{\xi}_i)$.
+\item Use data points $\big(\vec{y}_i, f(\vec\xi_i)\big)$, $i = 1, \ldots, N$, to compute a 
+regression surface $\tilde{G}(\vec{y})\approx 
 G(\vec{y})$.

### 2018-08-17 Paper/assa.tex
minor ...

@@
    \act{f}{i}{r} =  \sum_{k=1}^r \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2,
-   \quad j = 1, \ldots, \Np, \quad r \leq \Np.
+   \quad i = 1, \ldots, \Np, \quad r \leq \Np.
 \]
@@
 \begin{remark}
-Note that if $\xi_i$ are iid, the $C_i$'s in the definition of the 
-normalized screening metric will cancel and 
+If the random inputs $\xi_i$, $i = 1, \ldots, \Np$, are iid, then 
+the $C_i$'s in the definition of the normalized screening metric will cancel and 
 \[
     \tilde{\nu}_i(f) = \frac{\nu_i(f)}{\sum_{i=1}^\Np \nu_i(f)} 
-      = \frac{\vec{e}_i^T \mat{C} \vec{e}_i}{\trace(\mat{C})} 
+  %    = \frac{\vec{e}_i^T \mat{C} \vec{e}_i}{\trace(\mat{C})} 
       = \frac{\sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2}{\sum_{k = 1}^\Np \lambda_k}.
 \]
-In this case, the normalized activity scores reduce to
+The expression for the denominator follows, since 
+$\sum_{i=1}^\Np \nu_i(f) = \trace(\mat{C}) = \sum_{k = 1}^\Np \lambda_k$.
+Also, in the case of iid inputs, the normalized activity scores reduce to
 \[

### 2018-08-17 Paper/assa.tex
minor ...

@@
 \]
-Following~\cite{Constantine:2015}, we use the approximation $f(\vec{\xi}) \approx 
-G(\mat{W}_1^\top \vec{\xi})$. 
+Following~\cite{Constantine:2015}, we use the approximation 
+\[
+f(\vec{\xi}) \approx f(\mat{W}_1 \mat{W}_1^T \vec\xi) =  
+G(\mat{W}_1^\top \vec{\xi}).
+\] 
 That is, the model output $f(\vec\xi)$, in the original parameter space,

### 2018-08-17 Paper/assa.tex
sec 2, 3, almost there ...

@@
 
+We also record the following useful relations:
+\begin{lemma}\label{lem:sum}
+We have
+\begin{enumerate}[label=(\alph*)]
+\item $\sum_{i = 1}^\Np \act{f}{i}{r} = \sum_{k = 1}^r \lambda_k$. 
+\item $\sum_{i = 1}^\Np \nu_i(f) = \sum_{k = 1}^\Np \lambda_k$. 
+\end{enumerate}
+\end{lemma}
+\begin{proof}
+The first statement of the lemma holds, because
+\[
+\sum_{i=1}^\Np \act{f}{i}{r}    
+= \sum_{i=1}^\Np \sum_{k=1}^r \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2 
+= \sum_{k=1}^r \lambda_k \sum_{i=1}^\Np \ip{\vec{e}_i}{\vec{w}_k}^2 
+= \sum_{k=1}^r \lambda_k \| \vec{w}_k \|^2 = \sum_{k=1}^r \lambda_k.
+\]
+The statement (b) follows immediately from (a), because $\nu_i(f) = \act{f}{i}{\Np}$.
+%$\sum_{i=1}^\Np \nu_i(f) = \trace(\mat{C}) = \sum_{k = 1}^\Np \lambda_k$.
+\end{proof}
+
 It was shown in~\cite{Lamboni:2013} that the total Sobol' 
@@
 \]
-The expression for the denominator follows, since 
-$\sum_{i=1}^\Np \nu_i(f) = \trace(\mat{C}) = \sum_{k = 1}^\Np \lambda_k$.
+The expression for the denominator follows from Lemma~\ref{lem:sum}(b). 
 Also, in the case of iid inputs, the normalized activity scores reduce to
 \[
-   \actt{f}{i}{r} =  \frac{\act{f}{i}{r}}{\sum_{i=1}^\Np \act{f}{i}{r}}.
+   \actt{f}{i}{r} =  \frac{\act{f}{i}{r}}{\sum_{i=1}^\Np \act{f}{i}{r}} = 
+                     \frac{\sum_{k=1}^r \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2}
+                          {\sum_{k = 1}^r \lambda_k}.
 \]
+The expression for the denominator can be seen from Lemma~\ref{lem:sum}(a).
+
 \end{remark}

### 2018-08-18 Paper/assa.tex
minor ...

@@
+\newcommand{\act}[3]{\nu_{{#2},{#3}}({#1})}
+\newcommand{\actt}[3]{\tilde{\nu}_{{#2},{#3}}({#1})}
+\newcommand{\redf}[2]{f^{({#1})}(\vec\xi; {#2})}
 \section{Active subspaces}
@@
 $\mu(d\vec{\xi}) = \pi(\vec{\xi})d\vec{\xi}$, where $\pi(\vec{\xi})$ is the joint probability
-distribution of $\vec{\xi}$. Herein, $f$ is assumed to be a function with continuous partial 
-derivatives with respect to the input parameters. 
+distribution of $\vec{\xi}$. Herein, $f$ is assumed to be a square integrable 
+function with continuous partial 
+derivatives with respect to the input parameters; moreover, we assume the partial derivatives
+are square integrable. 
 %Hence, it is possible that a given $f(\vec{\xi})$ might not admit an active
@@
 \] 
-and $\mat{W}$ has the eigenvectors $\vec{w}_1, \ldots, \vec{w}_\Np$ as its columns.
+and $\mat{W}$ has the (orthonormal) eigenvectors $\vec{w}_1, \ldots, \vec{w}_\Np$ as its columns.
 The eigenpairs are partitioned about the $r$th eigenvalue such that
@@
  %
-Clearly, the amount of computational effort associated with constructing the matrix,
+Clearly, the computational effort associated with constructing the matrix,
 $\hat{\mat{C}}$ scales with the number of samples, $N$. Hence, an iterative
@@
 global sensitivity measures (DGSMs)~\cite{Sobol:2009} provide a means for
-approximating informative upper bounds on $T_i(f)$ with fewer computations; see 
+approximating informative upper bounds on $T_i(f)$ at a lower cost; see 
 also~\cite{Vohra:2018}.
@@
 Therefore,
-\[
-\nu_i(f) = \vec{e}_i^T \Big(\sum_{k=1}^\Np \lambda_k \vec{w}_k \vec{w}_k^\top\Big) \vec{e}_i
- = \sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2. \qedhere 
-\]
+$\nu_i(f) = \vec{e}_i^T \Big(\sum_{k=1}^\Np \lambda_k \vec{w}_k \vec{w}_k^\top\Big) \vec{e}_i
+ = \sum_{k=1}^\Np \lambda_k \ip{\vec{e}_i}{\vec{w}_k}^2$. 
 \end{proof}
@@
 accurate approximations of $\nu_i(f)$ by truncating the summation: 
-\newcommand{\act}[3]{\nu_{{#2},{#3}}({#1})}
-\newcommand{\actt}[3]{\tilde{\nu}_{{#2},{#3}}({#1})}
 \[
@@
 %$\xi_i$. 
-
+%
+To formalize this process, we proceed as follows.
 Let $\vec\xi$ be given and let $z$ be a nominal value for $\xi_i$.  
@@
 %and $y^z_i = z$.
-We consider the \emph{reduced} model, 
+Consider the \emph{reduced} model, 
 obtained by fixing $\xi_i$ at a nominal value $z \in \R$: 
 \[
-f^{z}(\vec\xi) = f(\xi_1, \xi_2, \ldots, \xi_{i-1}, z, \xi_{i+1}, \ldots, \xi_\Np),
+\redf{i}{z} = f(\xi_1, \xi_2, \ldots, \xi_{i-1}, z, \xi_{i+1}, \ldots, \xi_\Np),
 \] 
@@
 \mathcal{E}(z) =
-\frac{ \int_\Omega \big( f(\vec\xi) - f^{z}(\vec\xi)\big)^2 \, \mu(d\vec\xi) }
+\frac{ \int_\Omega \big( f(\vec\xi) - \redf{i}{z}\big)^2 \, \mu(d\vec\xi) }
           {\int_\Omega f(\vec\xi)^2 \, \mu(d\vec\xi)}.
@@
 \Ez{ \mathcal{E}(z)} \leq \frac{1}{\V(f)} \Ez{ 
-\int_\Omega \big( f(\vec\xi) - f^{z}(\vec\xi)\big)^2 \, \mu(d\vec\xi)}
+\int_\Omega \big( f(\vec\xi) - \redf{i}{z}\big)^2 \, \mu(d\vec\xi)}
 = 2 T_i(f), 
@@
 The expression for the denominator follows from Lemma~\ref{lem:sum}(b). 
-Also, in the case of iid inputs, the normalized activity scores reduce to
+Also, in the iid case, using Lemma~\ref{lem:sum}(a) we can simplify the normalized activity scores as follows. 
 \[
@@
 \]
-The expression for the denominator can be seen from Lemma~\ref{lem:sum}(a).
+%The expression for the denominator can be seen from Lemma~\ref{lem:sum}(a).
 

### 2018-08-18 Paper/assa.tex
fix typo ...

@@
 that depends on the distribution of $\xi_i$; see~\cite{Lamboni:2013}.
-For instance, if $\xi_i$ is uniformly distributed on $[-1, 1]$, then $C_i = 2/\pi^2$; and in the 
+For instance, if $\xi_i$ is uniformly distributed on $[-1, 1]$, then $C_i = 4/\pi^2$; and in the 
 case $\xi_i$ is normally distributed with variance $\sigma_i^2$, then $C_i = \sigma_i^2$. 

### 2018-08-20 Paper/app.tex
..

@@
 pre-exponents ($A_i$'s) as well as the activation energies ($E_{a,i}$'s) on ignition 
-delay during the H$_2$/O$_2$ reaction. The $A_i$'s were considered to be uniformly
-distributed in the interval, $[0.9A_i^\ast, 1.1A_i^\ast]$. The $E_{a,i}$'s for all
+delay during the H$_2$/O$_2$ reaction.  Prior distributions for $A_i$'s were considered
+to be the same as provided earlier in section~\ref{sec:method}. The $E_{a,i}$'s for all
 reactions except $\mathcal{R}_6$ -- $\mathcal{R}_9$ and $\mathcal{R}_{13}$ (due to 
@@
 were considered to be uncertain and uniformly distributed in the interval: 
-$[0.99E_{a,i}^\ast, 1.01E_{a,i}^\ast]$. The nominal values, $A_i^\ast$ and $E_{a,i}^\ast$
+$[0.99E_{a,i}^\ast, 1.01E_{a,i}^\ast]$. The nominal values, $E_{a,i}^\ast$
 corresponding to the different reaction rates are provided in~\cite{Yetter:1991}.
+
+\subsection{Computing the active subspace}
 
@@
 found to be roughly two orders of magnitude smaller than the first, indicating that the active subspace
-is 1-dimensional. This is further confirmed by the SSP plots in Figure~\ref{fig:hd} (right). Specifically
-in the case of gradient-based approach, the SSP exhibits a linear variation that is reasonably
-captured by its linear fit, $\tilde{G}$. Whereas, the SSP based on the gradient-free approach
-is relatively more scattered. Consequently, the linear fit exhibits a discrepancy with the corresponding
-fit for the gradient-based SSP. 
+is 1-dimensional. This is further confirmed by the SSP plots in Figure~\ref{fig:hd} (right). 
 %
@@
 %
-We plot the normalized activity scores as well as the total Sobol' indices for the 33 uncertain 
-rate-controlling parameters in Figure~\ref{fig:as_33D}. Note that the normalized activity scores
-as well as the Sobol' indices
-were evaluated using respective 1-dimensional surrogates for the two approaches.
-%
-\begin{figure}[htbp]
- \begin{center}
-  \includegraphics[width=0.8\textwidth]{./Figures/as_33D_new}
-\caption{A bar-graph illustrating individual activity scores for the uncertain $A_i$' (left) and $E_{a,i}$'s (right).}
-\label{fig:as_33D}
-\end{center}
-\end{figure}
-%
-
-Several useful inferences can be drawn from the above plot. Firstly, the total Sobol indices are found to be
-consistent
-with the normalized activity scores evaluated uing the gradient-based approach. Whereas, the gradient-free
-approach yields consistent results for the $A_i$'s, it does not capture the sensitivity associated with
-$E_{a,15}$. While this observation bolsters our confidence in the gradient-based approach, it clearly
-underscores the shortcoming of the gradient-free approach. Numerical errors are incurred when approximating
-the model gradients using the local linear approximation in this case. In~\ref{sub:verify}, we further assess the
-suitability of the gradient-free approach in quantifying the uncertainty in the ignition delay. 
-Secondly, our findings indicate that   
-the ignition delay is predominantly sensitive towards $A_1$, $A_9$, $E_{a,1}$, and $E_{a,15}$ and 
-moderately sensitive towards $A_{15}$ and $A_{17}$. Sensitivity towards the remaining uncertain inputs is
-found to be low or negligible. Note that this observation can also be exploited for reducing the 
-dimensionality of the problem from 33 inputs to 4-6 inputs depending upon the required level of accuracy.
-The sensitivity-driven dimension reduction for uncertainty quantification has been
-discussed in our earlier effort~\cite{Vohra:2018}. In this work, however, we focus on dimension reduction using the active 
-subspace approach which is shown to yield even greater scope for dimension reduction i.e. from 33 to 1. Furthermore, we have
-shown that the dominant eigenspace can be used to approximate global sensitivity measures with no additional effort.
-The accuracy of the 1-dimensional surrogate, $\tilde{G}$, obtained using the two approaches is assessed in the following section.
+Specifically
+in the case of gradient-based approach, the SSP exhibits a linear variation that is reasonably
+captured by its linear fit, $\tilde{G}$. Whereas, the SSP based on the gradient-free approach
+is relatively more scattered. Consequently, the linear fit exhibits a discrepancy with the corresponding
+fit for the gradient-based SSP.
 
@@
 
-%\subsection{Surrogate-induced risk analysis}
+\subsection{GSA consistency check}
+
+The normalized activity scores ($\tilde{\nu}_{i,r}$) based on the dominant eigenspace using the two
+approaches (gradient-based and gradient-free) are compared with the total Sobol' indices
+in Figure~\ref{fig:as_33D}. Note that the Sobol' indices were computed using the verified
+1-dimensional surrogate in the active subspace for the gradient-based approach. 
+%
+\begin{figure}[htbp]
+ \begin{center}
+  \includegraphics[width=0.8\textwidth]{./Figures/as_33D_new}
+\caption{A bar-graph illustrating individual activity scores for the uncertain $A_i$' (left) and $E_{a,i}$'s (right).}
+\label{fig:as_33D}
+\end{center}
+\end{figure}
+%
+Several useful inferences can be drawn from the above plot. Firstly, the total Sobol indices are found to be
+consistent
+with the normalized activity scores evaluated using the gradient-based approach. Whereas, the gradient-free
+approach yields consistent results for the $A_i$'s, it does not capture the sensitivity associated with
+$E_{a,15}$. While this observation bolsters our confidence in the gradient-based approach, it clearly
+underscores the shortcoming of the gradient-free approach. Numerical errors are incurred when approximating
+the model gradients using the local linear approximation in this case. 
+Secondly, our findings indicate that   
+the ignition delay is predominantly sensitive towards $A_1$, $A_9$, $E_{a,1}$, and $E_{a,15}$ and 
+moderately sensitive towards $A_{15}$ and $A_{17}$. Sensitivity towards the remaining uncertain inputs is
+found to be low or negligible. 
+%Note that this observation can also be exploited for reducing the 
+%dimensionality of the problem from 33 inputs to 4-6 inputs depending upon the required level of accuracy.
+%The sensitivity-driven dimension reduction for uncertainty quantification has been
+%discussed in our earlier effort~\cite{Vohra:2018}. In this work, however, we focus on dimension reduction using the active 
+%subspace approach which is shown to yield even greater scope for dimension reduction i.e. from 33 to 1. Furthermore, we have
+%shown that the dominant eigenspace can be used to approximate global sensitivity measures with no additional effort.

### 2018-08-20 Paper/intro.tex
..

@@
 approach that involves a regression-based local linear approximation of
-the model output. \alennote{I don't understand what the next sentence is 
-really trying to say here; I mean whatever we do involves approximating the
-gradient, the question is how rough the approximation is, right?} 
-It must be noted that since the gradient-free approach
-relies on approximating the model gradients, it is expected to be less
-robust than the gradient-based approach. The two approaches are based on
+the model output as discussed later in~\ref{sub:gradfree}. The two approaches
+ are based on
 algorithms 1.1 and 1.2 in~\cite{Constantine:2015}. Both strategies
@@
 exploiting their links with active subspaces. This led to definition of the so
-called \emph{activity scores}. In Section~\ref{sec:assa}, we build on these
+called \emph{activity scores}. In Section~\ref{sub:gsa}, we build on these
 ideas to provide a complete analysis of links between Sobol indices, DGSMs, and
-activity scores for functions of independent random inputs whos distribution
+activity scores for functions of independent random inputs whose distribution
 law belongs to the class of Boltzmann probability measures. 
@@
 Active subspace
-discovery in the high-dimensional H$_2$/O$_2$ kinetics problem involving 33
-uncertain inputs: \alennote{Say some interesting conclusion of the analysis;
-e.g., we discover a 1D subspace for a 33-dimensional problem using a modest computational 
-budget ...} The presented
+discovery in a high-dimensional H$_2$/O$_2$ kinetics problem involving 33
+uncertain inputs. The implementation of the methodology presented in this work
+resulted in a 1-dimensional active subspace, indicating immense potential
+for computational savings. The presented
 analysis can also guide practitioners in area of chemical kinetics on using the
-method of active subspaces as an efficient means of uncertainty propagation in
-chemical systems.  
+method of active subspaces as an efficient means for uncertainty propagation.  
 \item Comprehensive numerical investigation of the gradient-based and gradient-free approach: 
@@
 
-\alennote{Okay, I reorganized the sections a bit (new sec 2, 3); please read the following and
-make sure it reflects the contents of the sections correctly.}
 This article is organized as follows. In section~\ref{sub:ac}, a brief

### 2018-08-22 Paper/method.tex

@@
 
-In this section, we present the computational strategy used for computing the
-active subspace in the case of H$_2$/O$_2$ reaction kinetics with uncertain
-pre-exponent of the rate law associated with individual reactions provided in
-Table~\ref{tab:kinetics}.
+In this section, we present a generalized framework for computing the
+active subspace. The framework is employed to the H$_2$/O$_2$ reaction kinetics
+problem whereby the 
+pre-exponent ($A_i$) in the rate law associated with individual reactions provided in
+Table~\ref{tab:kinetics} is considered to be uniformly
+distributed in the interval, $[0.9A_i^\ast, 1.1A_i^\ast]$; $A_i^\ast$ is the nominal
+estimate provided in~\cite{Yetter:1991}.
 Specifically, we employ two strategies: 
@@
 $\hat{\mat{W}}_1$ and the set of corresponding eigenvalues, 
-$\hat{\mat{\Lambda}}_1$.
+$\hat{\mat{\Lambda}}_1$. Note that $\hat{\mat{W}}_1$ is obtained by
+partitioning the eigenspace around $\lambda_j$ such that the ratio of
+subsequent eigenvalues,
+$\left(\frac{\lambda_j}{\lambda_{j+1}}\right)>\mathcal{O}(10^2)$.
  At each subsequent iteration, model evaluations are
@@
 	\EndLoop
-	\State Compute $\nu_{i,r}(f) = \sum\limits_{j=1}^{r} \hat{\mat{\Lambda}}_{i,j}\hat{\mat{W}}_{i,j}^2$,
+	\State Compute $\nu_{i,r}(f) = \sum\limits_{j=1}^{r} \Lambda_j w_{i,j}^2$,
 	$i=1,\ldots,N_p$.
-	\State Normalize $\nu_{i,r}(f)$ as $\tilde{\nu_{i,r}}(f)$ = \scalebox{1.25}{$\frac{\nu_{i,r}(f)}{\sum_i\nu_{i,r}(f)}$}.
+	\State Normalize $\nu_{i,r}(f)$ as $\tilde{\nu}_{i,r}(f)$ = \scalebox{1.25}{$\frac{\nu_{i,r}(f)}{\sum_i\nu_{i,r}(f)}$}.
 
@@
 ($N$~=~1000) in the input domain and the corresponding active subspace is
-computed. In Figure~\ref{fig:eig_comp}~(left), we illustrate the comparison of
+computed. In Figure~\ref{fig:eig_comp}, we illustrate the comparison of
 the resulting normalized eigenvalue spectrum ($\lambda_1, \ldots, 
@@
 smaller set of samples, $n$~=~$\{20,40,80,120\}$.
-
-%Additionally, we compare relative L-2
-%norm of the difference between the squared value of corresponding components
-%of the eigenvectors, obtained using the large set, $N$, and the smaller sets, $n$.
-%In  Figure~\ref{fig:eig_comp} (right), the relative L-2 error norm for the
-%first five eigenvectors is plotted.
-
 %
@@
   \includegraphics[width=0.45\textwidth]{./Figures/eig_comp}
-   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1}
-\caption{Left: A comparison of the normalized eigenvalue spectrum using $n$ = $\{20,40,80,120\}$ samples with that
-obtained using a much larger sample size, $N$~=~1000. Right: Relative $L_2$ 
-norm of the difference between
-the square of individual components ($w_i$'s) of the dominant eigenvector evaluated using $N$=1000 and $n$ = $\{20,40,80,120\}$ samples.} 
+%   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1}
+\caption{A comparison of the normalized eigenvalue spectrum using $n$ = $\{20,40,80,120\}$ samples with that
+obtained using a much larger sample size, $N$~=~1000. 
+%Right: Relative $L_2$ 
+%norm of the difference between
+%the square of individual components ($w_i$'s) of the dominant eigenvector evaluated using 
+$N$=1000 and $n$ = $\{20,40,80,120\}$ samples.
+} 
 \label{fig:eig_comp}
@@
 sufficiently captures the uncertainty in
-the  model output. To confirm this, we plot a relative L-2 norm of the difference between the 
+the  model output. To further confirm this, we evaluate a relative L-2 norm of the difference
+ ($\varepsilon_\text{L-2}^{N-n}$) between the 
 squared value of corresponding components of the dominant eigenvector, computed using $N$ = 1000
-and $n$ = $\{20,40,80,120\}$ samples in Figure~\ref{fig:eig_comp} (right). 
-The relative L-2 error norm is found to be $\mathcal{O}(10^{-2})$ even for the case, $n$ = 20.
-Hence, a sample size of 20 is adequate for computing the active subspace in this
-case. The iterative strategy therefore offers a significant potential for computational advantage. 
+and $n$ = $\{20,40,80,120\}$.  The quantity, $\varepsilon_\text{L-2}^{N-n}$, was found to be 
+$\mathcal{O}(10^{-2})$ for $n$ = 20, 40, and 80; and $\mathcal{O}(10^{-3})$ for $n$ = 120.
+Thus, even a small sample size, $n$ = 20, seems to approximate the dominant eigenspace with
+reasonable accuracy in this case. 
+The iterative strategy therefore offers a significant potential for computational advantage. 
 In the following section, we provide a brief description of the gradient-free approach and compare
@@
   \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}
-\caption{Left: A bar-graph of normalized activity scores ($\tilde{\nu_{i,r}}$'s) for the 19 uncertain pre-exponents ($A_i$'s); $r$
+\caption{Left: A bar-graph of normalized activity scores ($\tilde{\nu_{i,r}}$'s) 
+for the 19 uncertain pre-exponents ($A_i$'s); $r$
 denotes the number of eigenvectors in the active subspace.
-Right: A plot of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$), defined as a normalized product of the
-Poincare constant and the DGSM ($\mu_i$) (adapted from~\cite{Vohra:2018}).}
+Right: A plot of the screening metric ($\tilde{\nu}_i$) for individual $A_i$'s 
+(adapted from~\cite{Vohra:2018}).}
 \label{fig:comp_as}

### 2018-08-22 Paper/app.tex
..

--- 2018-08-14 Paper/app.tex
@@
-\section{$\text{H}_2$/$\text{O}_2$ reaction kinetics: high-dimensional case}
-\label{sec:app}
+<!DOCTYPE html>
+<html lang="en">
+  <head>
+    <meta charset="utf-8">
+  <link rel="dns-prefetch" href="https://assets-cdn.github.com">
+  <link rel="dns-prefetch" href="https://avatars0.githubusercontent.com">
+  <link rel="dns-prefetch" href="https://avatars1.githubusercontent.com">
+  <link rel="dns-prefetch" href="https://avatars2.githubusercontent.com">
+  <link rel="dns-prefetch" href="https://avatars3.githubusercontent.com">
+  <link rel="dns-prefetch" href="https://github-cloud.s3.amazonaws.com">
+  <link rel="dns-prefetch" href="https://user-images.githubusercontent.com/">
 
-For the high-dimensional case, we aim to investigate the impact of the uncertain
-pre-exponents ($A_i$'s) as well as the activation energies ($E_{a,i}$'s) on ignition 
-delay during the H$_2$/O$_2$ reaction. The $A_i$'s were considered to be uniformly
-distributed in the interval, $[0.9A_i^\ast, 1.1A_i^\ast]$. The $E_{a,i}$'s for all
-reactions except $\mathcal{R}_6$ -- $\mathcal{R}_9$ and $\mathcal{R}_{13}$ (due to 
-zero nominal values for $E_a$)
-were considered to be uncertain and uniformly distributed in the interval: 
-$[0.99E_{a,i}^\ast, 1.01E_{a,i}^\ast]$. The nominal values, $A_i^\ast$ and $E_{a,i}^\ast$
-corresponding to the different reaction rates are provided in~\cite{Yetter:1991}.
+  <link crossorigin="anonymous" media="all" integrity="sha512-Z0JAar9+DkI1NjGVdZr3GivARUgJtA0o2eHlTv7Ou2gshR5awWVf8QGsq11Ns9ZxQLEs+G5/SuARmvpOLMzulw==" rel="stylesheet" href="https://assets-cdn.github.com/assets/frameworks-95aff0b550d3fe338b645a4deebdcb1b.css" />
+  <link crossorigin="anonymous" media="all" integrity="sha512-CKx3km+/sreL8TQnawck172wM/RpBNE60XhV8wwDbn+VZlleFETFHPsSnAdaNOUySA6yFi7Xc8vX6U9P0eNhXA==" rel="stylesheet" href="https://assets-cdn.github.com/assets/github-2e6b414f4a46ee13365df9e7fbfadb35.css" />
 
-The gradient-based approach discussed earlier in~\ref{sub:grad} was used to compute the
-active subspace. Using the iterative procedure, the convergence of the eigenvectors
-was examined by tracking $\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$, plotted in 
-Figure~\ref{fig:conv_app} (right). In Figure~\ref{fig:conv_app} (left), individual
-components of the converged eigenvector are illustrated. The convergence was established using
-a $\tau$ value of 0.02. Finite difference was used to compute model gradients at 40 samples in
-the input domain. Hence, a total of 1360 model evaluations were required for this purpose.  
-%
-\begin{figure}[htbp]
- \begin{center}
-  \includegraphics[width=0.8\textwidth]{./Figures/eigv10}
-\caption{Left: An illustrative comparison of individual components of the converged dominant eigenvector obtained
-using the two approaches discussed in~\ref{sub:grad} and~\ref{sub:gradfree}. Right: The quantity,  
-$\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$
-is plotted for successive iterations to illustrate the convergence behavior.}
-\label{fig:conv_app}
-\end{center}
-\end{figure}
-%
+  <meta name="viewport" content="width=device-width">
 
-In Figure~\ref{fig:hd} (left), we illustrate the resulting eigenvalue spectrum. The second eigenvalue is
-found to be roughly two orders of magnitude smaller than the first, indicating that the active subspace
-is 1-dimensional. This is further confirmed by the SSP plots in Figure~\ref{fig:hd} (right). Specifically
-in the case of gradient-based approach, the SSP exhibits a linear variation that is reasonably
-captured by its linear fit, $\tilde{G}$. Whereas, the SSP based on the gradient-free approach
-is relatively more scattered. Consequently, the linear fit exhibits a discrepancy with the corresponding
-fit for the gradient-based SSP. 
-%
-\begin{figure}[htbp]
- \begin{center}
-   \includegraphics[width=0.45\textwidth]{./Figures/eig_33D}
-   \includegraphics[width=0.45\textwidth]{./Figures/ssp_33D}
-\caption{Left: Eigenvalue spectrum of $\hat{\mat{C}}$. Right: SSP of the computed active subspace, and the
-corresponding 1-D linear fit.} 
-\label{fig:hd}
-\end{center}
-\end{figure}
-%
-We plot the normalized activity scores as well as the total Sobol' indices for the 33 uncertain 
-rate-controlling parameters in Figure~\ref{fig:as_33D}. Note that the normalized activity scores
-as well as the Sobol' indices
-were evaluated using respective 1-dimensional surrogates for the two approaches.
-%
-\begin{figure}[htbp]
- \begin{center}
-  \includegraphics[width=0.8\textwidth]{./Figures/as_33D_new}
-\caption{A bar-graph illustrating individual activity scores for the uncertain $A_i$' (left) and $E_{a,i}$'s (right).}
-\label{fig:as_33D}
-\end{center}
-\end{figure}
-%
+  <title>ActiveSubspace_Kinetics/app.tex at master · vohramanav/ActiveSubspace_Kinetics</title>
+    <meta name="description" content="GitHub is where people build software. More than 28 million people use GitHub to discover, fork, and contribute to over 85 million projects.">
+    <link rel="search" type="application/opensearchdescription+xml" href="/opensearch.xml" title="GitHub">
+  <link rel="fluid-icon" href="https://github.com/fluidicon.png" title="GitHub">
+  <meta property="fb:app_id" content="1401488693436528">
 
-Several useful inferences can be drawn from the above plot. Firstly, the total Sobol indices are found to be
-consistent
-with the normalized activity scores evaluated uing the gradient-based approach. Whereas, the gradient-free
-approach yields consistent results for the $A_i$'s, it does not capture the sensitivity associated with
-$E_{a,15}$. While this observation bolsters our confidence in the gradient-based approach, it clearly
-underscores the shortcoming of the gradient-free approach. Numerical errors are incurred when approximating
-the model gradients using the local linear approximation in this case. In~\ref{sub:verify}, we assess the
-suitability of the gradient-free approach in quantifying the uncertainty in the ignition delay. 
-Secondly, our findings indicate that the  
-the ignition delay is predominantly sensitivity towards $A_1$, $A_9$, $E_{a,1}$, and $E_{a,15}$ and 
-moderately sensitivity towards $A_{15}$ and $A_{17}$. Sensitivity towards the remaining uncertain inputs is
-found to be low or negligible. Note that his observation can also be exploited for reducing the 
-dimensionality of the problem from 33 inputs to 4-6 inputs depending upon the required level of accuracy.
-The sensitivity-driven dimension reduction for uncertainty quantification has been
-discussed in our earlier effort~\cite{Vohra:2018}. In this work, however, we focus on dimension reduction using the active 
-subspace approach which is shown to yield even greater scope for dimension reduction i.e. from 33 to 1. Furthermore, we have
-shown that the dominant eigenspace can be used to approximate global sensitivity measures with no additional effort.
-The accuracy of the 1-dimensional surrogate, $\tilde{G}$, obtained using the two approaches is assessed in the following section.
+    <meta property="og:image" content="https://avatars1.githubusercontent.com/u/22374862?s=400&amp;v=4" /><meta property="og:site_name" content="GitHub" /><meta property="og:type" content="object" /><meta property="og:title" content="vohramanav/ActiveSubspace_Kinetics" /><meta property="og:url" content="https://github.com/vohramanav/ActiveSubspace_Kinetics" /><meta property="og:description" content="Contribute to ActiveSubspace_Kinetics development by creating an account on Github." />
 
-\subsection{Surrogate Verification}
-\label{sub:verify}
+  <link rel="assets" href="https://assets-cdn.github.com/">
+  <link rel="web-socket" href="wss://live.github.com/_sockets/VjI6MzA5MjA1Njc0OmExZjFmODllNjEzOTRmODNiM2EzYjg2MjMxMjcwNTI2ODE5ZTg0OWE3YmRlMDYwMmRlZmZmNjFiMmMyNzI5MWY=--0d6991307d6fb87b7f9a3a145aeb29eb4bbf4bd4">
+  <meta name="pjax-timeout" content="1000">
+  <link rel="sudo-modal" href="/sessions/sudo_modal">
+  <meta name="request-id" content="C8D0:4A98:4461DB:7F3CB3:5B7D804B" data-pjax-transient>
 
-The 1-dimensional surrogate ($\tilde{G}$) is investigated for its accuracy as well as the ability to capture the 
-uncertainty in the
-model output in two ways. Firstly, we estimate the relative L-2 norm of the discrepancy~($\varepsilon_d$)
-between estimates of ignition delay in the case of H$_2$/O$_2$ reaction, obtained using the 
-model output and the surrogate in the following equation:
-%
-\be
-\varepsilon_d = \frac{\|G(\bm{\xi}) - \tilde{G}(\bm{\xi})\|_2}{\|G(\bm{\xi})\|_2}
-\ee
-%
-The relative norm, $\varepsilon_d$ was estimated to be 8.50$\times10^{-2}$ based on a cross-validation test data
-comprising 10$^{4}$ model evaluations. Hence, in the norm-sense, it can be said that the 1-dimensional surrogate is 
-reasonably accurate.
+  <meta name="selected-link" value="repo_source" data-pjax-transient>
 
-Secondly, we verify the accuracy of the surrogate in a probabilistic setting. In particular, we compare a histogram
-plot of the model evaluations (in the cross-validation set) with a PDF evaluated on the same set of samples
-using the surrogate, $\tilde{G}$ in Figure~\ref{fig:pdf_33D}. The PDF is evaluated using surrogates constructed for
-a 1-dimensional as well as a 2-dimensional active subspace. 
-%
-\begin{figure}[htbp]
- \begin{center}
-  \includegraphics[width=0.45\textwidth]{./Figures/pdf_comp_id_1e4}
-\caption{A comparison of the histogram plot based on model evaluations and the PDF based on surrogate
-predictions for ignition delay. The two plots are based on samples in the validation set.}
-\label{fig:pdf_33D}
-\end{center}
-\end{figure}
-%
-In the above comparison, it is observed that the two PDFs are almost identical, thereby confirming a 1-dimensional
-active subspace. The modal estimate from either PDF is observed to be in agreement with that
-based on the histogram plot. The mean value based on the set of model evaluations used for verification as well
-as the 1-dimensional surrogate was found to be 0.133, whereas, the standard deviations were found to be 0.0198 and 0.0160
-respectively. Hence, while the surrogate yields an accurate estimation of the mean, the standard deviation is
-under-estimated. This can be understood from the SSP in Figure~\ref{fig:hd} (right) which shows that the model
-evaluations do not exactly lie on the linear 1-dimensional surrogate used to capture them. Therefore, the
-1-dimensional surrogate can be used to estimate the mean as well as the modal estimate of the ignition delay in
-the presence of uncertainty in the 33 parameter considered in this application. The uncertainty associated with the
-ignition delay though reasonably close to the true value, is under-estimated.
+    <meta name="google-site-verification" content="KT5gs8h0wvaagLKAVWq8bbeNwnZZK1r1XQysX3xurLU">
+  <meta name="google-site-verification" content="ZzhVyEFwb7w3e0-uOTltm8Jsck2F5StVihD0exw2fsA">
+  <meta name="google-site-verification" content="GXs5KoUUkNCoaAZn7wPN-t01Pywp9M3sEjnt_3_ZWPc">
+    <meta name="google-analytics" content="UA-3769691-2">
 
-%\subsection{Surrogate-induced risk analysis}
+<meta name="octolytics-host" content="collector.githubapp.com" /><meta name="octolytics-app-id" content="github" /><meta name="octolytics-event-url" content="https://collector.githubapp.com/github-external/browser_event" /><meta name="octolytics-dimension-request_id" content="C8D0:4A98:4461DB:7F3CB3:5B7D804B" /><meta name="octolytics-dimension-region_edge" content="iad" /><meta name="octolytics-dimension-region_render" content="iad" /><meta name="octolytics-actor-id" content="22374862" /><meta name="octolytics-actor-login" content="vohramanav" /><meta name="octolytics-actor-hash" content="c3a22d3bae68f1b4588180c2c31f78aff3c5eee2ea5a5b23e0ab13a0fad0bd14" />
+<meta name="analytics-location" content="/&lt;user-name&gt;/&lt;repo-name&gt;/blob/show" data-pjax-transient="true" />
+
+  <meta class="js-ga-set" name="userId" content="d42e8fff36f702a794b24d5e9a4a8c36" %>
+
+<meta class="js-ga-set" name="dimension1" content="Logged In">
+
+      <meta name="hostname" content="github.com">
+    <meta name="user-login" content="vohramanav">
+
+      <meta name="expected-hostname" content="github.com">
+    <meta name="js-proxy-site-detection-payload" content="MjU5NmUzMWQwMmNmN2RlNDU4N2Y2MGM4NjBiYTM2YTFlNGU4YzNkMDdiYWU2YzYyMzFlZjRkNDhhNTZmY2Q1OXx7InJlbW90ZV9hZGRyZXNzIjoiMTI5LjU5Ljc2LjQwIiwicmVxdWVzdF9pZCI6IkM4RDA6NEE5ODo0NDYxREI6N0YzQ0IzOjVCN0Q4MDRCIiwidGltZXN0YW1wIjoxNTM0OTUxNTAzLCJob3N0IjoiZ2l0aHViLmNvbSJ9">
+
+    <meta name="enabled-features" content="DASHBOARD_V2_LAYOUT_OPT_IN,EXPLORE_DISCOVER_REPOSITORIES,UNIVERSE_BANNER,FREE_TRIALS,MARKETPLACE_INSIGHTS,MARKETPLACE_DOCKERFILE_CI_CTA,MARKETPLACE_PLAN_RESTRICTION_EDITOR,MARKETPLACE_SEARCH,MARKETPLACE_INSIGHTS_CONVERSION_PERCENTAGES,MARKETPLACE_RETARGETING">
+
+  <meta name="html-safe-nonce" content="d1d04259e83798f31bc24be67464ccae1ab86a27">
+
+  <meta http-equiv="x-pjax-version" content="6e18d0560e8611a207cfb2a7bb2a5b18">
+
+      <link href="https://github.com/vohramanav/ActiveSubspace_Kinetics/commits/master.atom" rel="alternate" title="Recent Commits to ActiveSubspace_Kinetics:master" type="application/atom+xml">
+
+  <meta name="go-import" content="github.com/vohramanav/ActiveSubspace_Kinetics git https://github.com/vohramanav/ActiveSubspace_Kinetics.git">
+
+  <meta name="octolytics-dimension-user_id" content="22374862" /><meta name="octolytics-dimension-user_login" content="vohramanav" /><meta name="octolytics-dimension-repository_id" content="139043874" /><meta name="octolytics-dimension-repository_nwo" content="vohramanav/ActiveSubspace_Kinetics" /><meta name="octolytics-dimension-repository_public" content="true" /><meta name="octolytics-dimension-repository_is_fork" content="false" /><meta name="octolytics-dimension-repository_network_root_id" content="139043874" /><meta name="octolytics-dimension-repository_network_root_nwo" content="vohramanav/ActiveSubspace_Kinetics" /><meta name="octolytics-dimension-repository_explore_github_marketplace_ci_cta_shown" content="true" />
+
+    <link rel="canonical" href="https://github.com/vohramanav/ActiveSubspace_Kinetics/blob/master/Paper/app.tex" data-pjax-transient>
+
+  <meta name="browser-stats-url" content="https://api.github.com/_private/browser/stats">
+
+  <meta name="browser-errors-url" content="https://api.github.com/_private/browser/errors">
+
+  <link rel="mask-icon" href="https://assets-cdn.github.com/pinned-octocat.svg" color="#000000">
+  <link rel="icon" type="image/x-icon" class="js-site-favicon" href="https://assets-cdn.github.com/favicon.ico">
+
+<meta name="theme-color" content="#1e2327">
+
+  <link rel="manifest" href="/manifest.json" crossOrigin="use-credentials">
+
+  </head>
+
+  <body class="logged-in env-production page-blob">
+
+  <div class="position-relative js-header-wrapper ">
+    <a href="#start-of-content" tabindex="1" class="p-3 bg-blue text-white show-on-focus js-skip-to-content">Skip to content</a>
+    <div id="js-pjax-loader-bar" class="pjax-loader-bar"><div class="progress"></div></div>
+
+<header class="Header  f5" role="banner">
+  <div class="d-flex flex-justify-between px-3 container-lg">
+    <div class="d-flex flex-justify-between ">
+      <div class="">
+        <a class="header-logo-invertocat" href="https://github.com/" data-hotkey="g d" aria-label="Homepage" data-ga-click="Header, go to dashboard, icon:logo">
+  <svg height="32" class="octicon octicon-mark-github" viewBox="0 0 16 16" version="1.1" width="32" aria-hidden="true"><path fill-rule="evenodd" d="M8 0C3.58 0 0 3.58 0 8c0 3.54 2.29 6.53 5.47 7.59.4.07.55-.17.55-.38 0-.19-.01-.82-.01-1.49-2.01.37-2.53-.49-2.69-.94-.09-.23-.48-.94-.82-1.13-.28-.15-.68-.52-.01-.53.63-.01 1.08.58 1.23.82.72 1.21 1.87.87 2.33.66.07-.52.28-.87.51-1.07-1.78-.2-3.64-.89-3.64-3.95 0-.87.31-1.59.82-2.15-.08-.2-.36-1.02.08-2.12 0 0 .67-.21 2.2.82.64-.18 1.32-.27 2-.27.68 0 1.36.09 2 .27 1.53-1.04 2.2-.82 2.2-.82.44 1.1.16 1.92.08 2.12.51.56.82 1.27.82 2.15 0 3.07-1.87 3.75-3.65 3.95.29.25.54.73.54 1.48 0 1.07-.01 1.93-.01 2.2 0 .21.15.46.55.38A8.013 8.013 0 0 0 16 8c0-4.42-3.58-8-8-8z"/></svg>
+</a>
+
+      </div>
+
+    </div>
+
+    <div class="HeaderMenu d-flex flex-justify-between flex-auto">
+      <div class="d-flex">
+            <div class="">
+              <div class="header-search scoped-search site-scoped-search js-site-search position-relative js-jump-to"
+  role="search combobox"
+  aria-owns="jump-to-results"
+  aria-label="Search or jump to"
+  aria-haspopup="listbox"
+  aria-expanded="true"
+>
+  <div class="position-relative">
+    <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="js-site-search-form" data-scope-type="Repository" data-scope-id="139043874" data-scoped-search-url="/vohramanav/ActiveSubspace_Kinetics/search" data-unscoped-search-url="/search" action="/vohramanav/ActiveSubspace_Kinetics/search" accept-charset="UTF-8" method="get"><input name="utf8" type="hidden" value="&#x2713;" />
+      <label class="form-control header-search-wrapper header-search-wrapper-jump-to position-relative d-flex flex-justify-between flex-items-center js-chromeless-input-container">
+        <input type="text"
+          class="form-control header-search-input jump-to-field js-jump-to-field js-site-search-focus js-site-search-field is-clearable"
+          data-hotkey="s,/"
+          name="q"
+          value=""
+          placeholder="Search or jump to…"
+          data-unscoped-placeholder="Search or jump to…"
+          data-scoped-placeholder="Search or jump to…"
+          autocapitalize="off"
+          aria-autocomplete="list"
+          aria-controls="jump-to-results"
+          data-jump-to-suggestions-path="/_graphql/GetSuggestedNavigationDestinations#csrf-token=Khu8Dh2eDlXUD2jr0BVZmOGTJeKujLMjOqtMwqFy3ShhK9UEv9eUmb6LqTqrYEPMr7se6xSia2vnRPEXKRS76A=="
+          spellcheck="false"
+          autocomplete="off"
+          >
+          <input type="hidden" class="js-site-search-type-field" name="type" >
+            <img src="https://assets-cdn.github.com/images/search-shortcut-hint.svg" alt="" class="mr-2 header-search-key-slash">
+
+            <div class="Box position-absolute overflow-hidden d-none jump-to-suggestions js-jump-to-suggestions-container">
+              <ul class="d-none js-jump-to-suggestions-template-container">
+                <li class="d-flex flex-justify-start flex-items-center p-0 f5 navigation-item js-navigation-item">
+                  <a tabindex="-1" class="no-underline d-flex flex-auto flex-items-center p-2 jump-to-suggestions-path js-jump-to-suggestion-path js-navigation-open" href="">
+                    <div class="jump-to-octicon js-jump-to-octicon mr-2 text-center d-none"></div>
+                    <img class="avatar mr-2 flex-shrink-0 js-jump-to-suggestion-avatar" alt="" aria-label="Team" src="" width="28" height="28">
+
+                    <div class="jump-to-suggestion-name js-jump-to-suggestion-name flex-auto overflow-hidden text-left no-wrap css-truncate css-truncate-target">
+                    </div>
+
+                    <div class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none js-jump-to-badge-search">
+                      <span class="js-jump-to-badge-search-text-default d-none" aria-label="in this repository">
+                        In this repository
+                      </span>
+                      <span class="js-jump-to-badge-search-text-global d-none" aria-label="in all of GitHub">
+                        All GitHub
+                      </span>
+                      <span aria-hidden="true" class="d-inline-block ml-1 v-align-middle">↵</span>
+                    </div>
+
+                    <div aria-hidden="true" class="border rounded-1 flex-shrink-0 bg-gray px-1 text-gray-light ml-1 f6 d-none d-on-nav-focus js-jump-to-badge-jump">
+                      Jump to
+                      <span class="d-inline-block ml-1 v-align-middle">↵</span>
+                    </div>
+                  </a>
+                </li>
+                <svg height="16" width="16" class="octicon octicon-repo flex-shrink-0 js-jump-to-repo-octicon-template" title="Repository" aria-label="Repository" viewBox="0 0 12 16" version="1.1" role="img"><path fill-rule="evenodd" d="M4 9H3V8h1v1zm0-3H3v1h1V6zm0-2H3v1h1V4zm0-2H3v1h1V2zm8-1v12c0 .55-.45 1-1 1H6v2l-1.5-1.5L3 16v-2H1c-.55 0-1-.45-1-1V1c0-.55.45-1 1-1h10c.55 0 1 .45 1 1zm-1 10H1v2h2v-1h3v1h5v-2zm0-10H2v9h9V1z"/></svg>
+                <svg height="16" width="16" class="octicon octicon-project flex-shrink-0 js-jump-to-project-octicon-template" title="Project" aria-label="Project" viewBox="0 0 15 16" version="1.1" role="img"><path fill-rule="evenodd" d="M10 12h3V2h-3v10zm-4-2h3V2H6v8zm-4 4h3V2H2v12zm-1 1h13V1H1v14zM14 0H1a1 1 0 0 0-1 1v14a1 1 0 0 0 1 1h13a1 1 0 0 0 1-1V1a1 1 0 0 0-1-1z"/></svg>
+                <svg height="16" width="16" class="octicon octicon-search flex-shrink-0 js-jump-to-search-octicon-template" title="Search" aria-label="Search" viewBox="0 0 16 16" version="1.1" role="img"><path fill-rule="evenodd" d="M15.7 13.3l-3.81-3.83A5.93 5.93 0 0 0 13 6c0-3.31-2.69-6-6-6S1 2.69 1 6s2.69 6 6 6c1.3 0 2.48-.41 3.47-1.11l3.83 3.81c.19.2.45.3.7.3.25 0 .52-.09.7-.3a.996.996 0 0 0 0-1.41v.01zM7 10.7c-2.59 0-4.7-2.11-4.7-4.7 0-2.59 2.11-4.7 4.7-4.7 2.59 0 4.7 2.11 4.7 4.7 0 2.59-2.11 4.7-4.7 4.7z"/></svg>
+              </ul>
+              <ul class="d-none js-jump-to-no-results-template-container">
+                <li class="d-flex flex-justify-center flex-items-center p-3 f5 d-none">
+                  <span class="text-gray">No suggested jump to results</span>
+                </li>
+              </ul>
+
+              <ul id="jump-to-results" class="js-navigation-container jump-to-suggestions-results-container js-jump-to-suggestions-results-container" >
+                <li class="d-flex flex-justify-center flex-items-center p-0 f5">
+                  <img src="https://assets-cdn.github.com/images/spinners/octocat-spinner-128.gif" alt="Octocat Spinner Icon" class="m-2" width="28">
+                </li>
+              </ul>
+            </div>
+      </label>
+</form>  </div>
+</div>
+
+            </div>
+
+          <ul class="d-flex pl-2 flex-items-center text-bold list-style-none" role="navigation">
+            <li>
+              <a class="js-selected-navigation-item HeaderNavlink px-2" data-hotkey="g p" data-ga-click="Header, click, Nav menu - item:pulls context:user" aria-label="Pull requests you created" data-selected-links="/pulls /pulls/assigned /pulls/mentioned /pulls" href="/pulls">
+                Pull requests
+</a>            </li>
+            <li>
+              <a class="js-selected-navigation-item HeaderNavlink px-2" data-hotkey="g i" data-ga-click="Header, click, Nav menu - item:issues context:user" aria-label="Issues you created" data-selected-links="/issues /issues/assigned /issues/mentioned /issues" href="/issues">
+                Issues
+</a>            </li>
+              <li>
+                <a class="js-selected-navigation-item HeaderNavlink px-2" data-ga-click="Header, click, Nav menu - item:marketplace context:user" data-octo-click="marketplace_click" data-octo-dimensions="location:nav_bar" data-selected-links=" /marketplace" href="/marketplace">
+                   Marketplace
+</a>              </li>
+            <li>
+              <a class="js-selected-navigation-item HeaderNavlink px-2" data-ga-click="Header, click, Nav menu - item:explore" data-selected-links="/explore /trending /trending/developers /integrations /integrations/feature/code /integrations/feature/collaborate /integrations/feature/ship showcases showcases_search showcases_landing /explore" href="/explore">
+                Explore
+</a>            </li>
+          </ul>
+      </div>
+
+      <div class="d-flex">
+
+<ul class="user-nav d-flex flex-items-center list-style-none" id="user-links">
+  <li class="dropdown">
+    <span class="d-inline-block  px-2">
+
+    </span>
+  </li>
+
+  <li class="dropdown">
+    <details class="details-overlay details-reset d-flex px-2 flex-items-center">
+      <summary class="HeaderNavlink"
+         aria-label="Create new…"
+         data-ga-click="Header, create new, icon:add">
+        <svg class="octicon octicon-plus float-left mr-1 mt-1" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 9H7v5H5V9H0V7h5V2h2v5h5v2z"/></svg>
+        <span class="dropdown-caret mt-1"></span>
+      </summary>
+      <details-menu class="dropdown-menu dropdown-menu-sw">
+
+<a role="menuitem" class="dropdown-item" href="/new" data-ga-click="Header, create new repository">
+  New repository
+</a>
+
+  <a role="menuitem" class="dropdown-item" href="/new/import" data-ga-click="Header, import a repository">
+    Import repository
+  </a>
+
+<a role="menuitem" class="dropdown-item" href="https://gist.github.com/" data-ga-click="Header, create new gist">
+  New gist
+</a>
+
+  <a role="menuitem" class="dropdown-item" href="/organizations/new" data-ga-click="Header, create new organization">
+    New organization
+  </a>
+
+  <div class="dropdown-divider"></div>
+  <div class="dropdown-header">
+    <span title="vohramanav/ActiveSubspace_Kinetics">This repository</span>
+  </div>
+    <a role="menuitem" class="dropdown-item" href="/vohramanav/ActiveSubspace_Kinetics/issues/new" data-ga-click="Header, create new issue">
+      New issue
+    </a>
+
+      </details-menu>
+    </details>
+  </li>
+
+  <li class="dropdown">
+
+    <details class="details-overlay details-reset d-flex pl-2 flex-items-center">
+      <summary class="HeaderNavlink name mt-1"
+        aria-label="View profile and more"
+        data-ga-click="Header, show menu, icon:avatar">
+        <img alt="@vohramanav" class="avatar float-left mr-1" src="https://avatars2.githubusercontent.com/u/22374862?s=40&amp;v=4" height="20" width="20">
+        <span class="dropdown-caret"></span>
+      </summary>
+      <details-menu class="dropdown-menu dropdown-menu-sw">
+        <ul>
+          <li class="header-nav-current-user css-truncate"><a role="menuitem" class="no-underline user-profile-link px-3 pt-2 pb-2 mb-n2 mt-n1 d-block" href="/vohramanav" data-ga-click="Header, go to profile, text:Signed in as">Signed in as <strong class="css-truncate-target">vohramanav</strong></a></li>
+          <li class="dropdown-divider"></li>
+          <li><a role="menuitem" class="dropdown-item" href="/vohramanav" data-ga-click="Header, go to profile, text:your profile">Your profile</a></li>
+          <li><a role="menuitem" class="dropdown-item" href="/vohramanav?tab=repositories" data-ga-click="Header, go to repositories, text:your repositories">Your repositories</a></li>
+          <li><a role="menuitem" class="dropdown-item" href="/vohramanav?tab=stars" data-ga-click="Header, go to starred repos, text:your stars">Your stars</a></li>
+            <li><a role="menuitem" class="dropdown-item" href="https://gist.github.com/" data-ga-click="Header, your gists, text:your gists">Your gists</a></li>
+          <li class="dropdown-divider"></li>
+          <li><a role="menuitem" class="dropdown-item" href="https://help.github.com" data-ga-click="Header, go to help, text:help">Help</a></li>
+          <li><a role="menuitem" class="dropdown-item" href="/settings/profile" data-ga-click="Header, go to settings, icon:settings">Settings</a></li>
+          <li>
+            <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="logout-form" action="/logout" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="3FaIm4CNM+9PIkyx5NRmLAxirPlwSyU57JNLpTDpqd2yz8pM/zAEZMKa+Utzum5LjQrgvK9fti0XWce6BUIaTw==" />
+              <button type="submit" class="dropdown-item dropdown-signout" data-ga-click="Header, sign out, icon:logout" role="menuitem">
+                Sign out
+              </button>
+</form>          </li>
+        </ul>
+      </details-menu>
+    </details>
+  </li>
+</ul>
+
+        <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="sr-only right-0" action="/logout" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="sdl7U6uC/YnIEBDe0m1Ffqw8ZpCWgyX9StrMRnOb1xTfQDmE1D/KAkWopSRFA00ZLVQq1UmXtumxEEBZRjBkhg==" />
+          <button type="submit" class="dropdown-item dropdown-signout" data-ga-click="Header, sign out, icon:logout">
+            Sign out
+          </button>
+</form>      </div>
+    </div>
+  </div>
+</header>
+
+  </div>
+
+  <div id="start-of-content" class="show-on-focus"></div>
+
+    <div id="js-flash-container">
+
+</div>
+
+  <div role="main" class="application-main ">
+        <div itemscope itemtype="http://schema.org/SoftwareSourceCode" class="">
+    <div id="js-repo-pjax-container" data-pjax-container >
+
+  <div class="pagehead repohead instapaper_ignore readability-menu experiment-repo-nav  ">
+    <div class="repohead-details-container clearfix container">
+
+      <ul class="pagehead-actions">
+  <li>
+        <!-- '"` --><!-- </textarea></xmp> --></option></form><form data-autosubmit="true" data-remote="true" class="js-social-container" action="/notifications/subscribe" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="hJC2MQ3NdavI/Csfx3TSs4s2QA4FpmgBs4G2/t++I+CnF84lI3fOU89Q7x8ng6qvvNfpBemZdOJN+oksI+YsXg==" />      <input type="hidden" name="repository_id" id="repository_id" value="139043874" class="form-control" />
+
+        <div class="select-menu js-menu-container js-select-menu">
+          <a href="/vohramanav/ActiveSubspace_Kinetics/subscription"
+            class="btn btn-sm btn-with-count select-menu-button js-menu-target"
+            role="button"
+            aria-haspopup="true"
+            aria-expanded="false"
+            aria-label="Toggle repository notifications menu"
+            data-ga-click="Repository, click Watch settings, action:blob#show">
+            <span class="js-select-button">
+                <svg class="octicon octicon-eye v-align-text-bottom" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.06 2C3 2 0 8 0 8s3 6 8.06 6C13 14 16 8 16 8s-3-6-7.94-6zM8 12c-2.2 0-4-1.78-4-4 0-2.2 1.8-4 4-4 2.22 0 4 1.8 4 4 0 2.22-1.78 4-4 4zm2-4c0 1.11-.89 2-2 2-1.11 0-2-.89-2-2 0-1.11.89-2 2-2 1.11 0 2 .89 2 2z"/></svg>
+                Watch
+            </span>
+          </a>
+          <a class="social-count js-social-count"
+            href="/vohramanav/ActiveSubspace_Kinetics/watchers"
+            aria-label="1 user is watching this repository">
+            1
+          </a>
+
+        <div class="select-menu-modal-holder">
+          <div class="select-menu-modal subscription-menu-modal js-menu-content">
+            <div class="select-menu-header js-navigation-enable" tabindex="-1">
+              <svg class="octicon octicon-x js-menu-close" role="img" aria-label="Close" viewBox="0 0 12 16" version="1.1" width="12" height="16"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
+              <span class="select-menu-title">Notifications</span>
+            </div>
+
+              <div class="select-menu-list js-navigation-container" role="menu">
+
+                <div class="select-menu-item js-navigation-item selected" role="menuitem" tabindex="0">
+                  <svg class="octicon octicon-check select-menu-item-icon" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5L12 5z"/></svg>
+                  <div class="select-menu-item-text">
+                    <input type="radio" name="do" id="do_included" value="included" checked="checked" />
+                    <span class="select-menu-item-heading">Not watching</span>
+                    <span class="description">Be notified when participating or @mentioned.</span>
+                    <span class="js-select-button-text hidden-select-button-text">
+                      <svg class="octicon octicon-eye v-align-text-bottom" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.06 2C3 2 0 8 0 8s3 6 8.06 6C13 14 16 8 16 8s-3-6-7.94-6zM8 12c-2.2 0-4-1.78-4-4 0-2.2 1.8-4 4-4 2.22 0 4 1.8 4 4 0 2.22-1.78 4-4 4zm2-4c0 1.11-.89 2-2 2-1.11 0-2-.89-2-2 0-1.11.89-2 2-2 1.11 0 2 .89 2 2z"/></svg>
+                      Watch
+                    </span>
+                  </div>
+                </div>
+
+                <div class="select-menu-item js-navigation-item " role="menuitem" tabindex="0">
+                  <svg class="octicon octicon-check select-menu-item-icon" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5L12 5z"/></svg>
+                  <div class="select-menu-item-text">
+                    <input type="radio" name="do" id="do_subscribed" value="subscribed" />
+                    <span class="select-menu-item-heading">Watching</span>
+                    <span class="description">Be notified of all conversations.</span>
+                    <span class="js-select-button-text hidden-select-button-text">
+                      <svg class="octicon octicon-eye v-align-text-bottom" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.06 2C3 2 0 8 0 8s3 6 8.06 6C13 14 16 8 16 8s-3-6-7.94-6zM8 12c-2.2 0-4-1.78-4-4 0-2.2 1.8-4 4-4 2.22 0 4 1.8 4 4 0 2.22-1.78 4-4 4zm2-4c0 1.11-.89 2-2 2-1.11 0-2-.89-2-2 0-1.11.89-2 2-2 1.11 0 2 .89 2 2z"/></svg>
+                        Unwatch
+                    </span>
+                  </div>
+                </div>
+
+                <div class="select-menu-item js-navigation-item " role="menuitem" tabindex="0">
+                  <svg class="octicon octicon-check select-menu-item-icon" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5L12 5z"/></svg>
+                  <div class="select-menu-item-text">
+                    <input type="radio" name="do" id="do_ignore" value="ignore" />
+                    <span class="select-menu-item-heading">Ignoring</span>
+                    <span class="description">Never be notified.</span>
+                    <span class="js-select-button-text hidden-select-button-text">
+                      <svg class="octicon octicon-mute v-align-text-bottom" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8 2.81v10.38c0 .67-.81 1-1.28.53L3 10H1c-.55 0-1-.45-1-1V7c0-.55.45-1 1-1h2l3.72-3.72C7.19 1.81 8 2.14 8 2.81zm7.53 3.22l-1.06-1.06-1.97 1.97-1.97-1.97-1.06 1.06L11.44 8 9.47 9.97l1.06 1.06 1.97-1.97 1.97 1.97 1.06-1.06L13.56 8l1.97-1.97z"/></svg>
+                        Stop ignoring
+                    </span>
+                  </div>
+                </div>
+
+              </div>
+
+            </div>
+          </div>
+        </div>
+</form>
+  </li>
+
+  <li>
+
+  <div class="js-toggler-container js-social-container starring-container ">
+    <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="starred js-social-form" action="/vohramanav/ActiveSubspace_Kinetics/unstar" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="nlQEgDZqIDarSTmYWZRyabcRQTOKN2kHMJDtdWpsDFkcvUvB9w+isACt0PdhXndQeYd1o5R4xxsREC3j4rQGrw==" />
+      <input type="hidden" name="context" value="repository"></input>
+      <button
+        type="submit"
+        class="btn btn-sm btn-with-count js-toggler-target"
+        aria-label="Unstar this repository" title="Unstar vohramanav/ActiveSubspace_Kinetics"
+        data-ga-click="Repository, click unstar button, action:blob#show; text:Unstar">
+        <svg class="octicon octicon-star v-align-text-bottom" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M14 6l-4.9-.64L7 1 4.9 5.36 0 6l3.6 3.26L2.67 14 7 11.67 11.33 14l-.93-4.74L14 6z"/></svg>
+        Unstar
+      </button>
+        <a class="social-count js-social-count" href="/vohramanav/ActiveSubspace_Kinetics/stargazers"
+           aria-label="0 users starred this repository">
+          0
+        </a>
+</form>
+    <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="unstarred js-social-form" action="/vohramanav/ActiveSubspace_Kinetics/star" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="c/RaxS+3o1dBydJ8SDGs6EgAoUEG/xq+9x5RUNJiDAmxQi21ayI25gQHGus4d0L2Jr/jPRTN7nvkgq+EYI9HdA==" />
+      <input type="hidden" name="context" value="repository"></input>
+      <button
+        type="submit"
+        class="btn btn-sm btn-with-count js-toggler-target"
+        aria-label="Star this repository" title="Star vohramanav/ActiveSubspace_Kinetics"
+        data-ga-click="Repository, click star button, action:blob#show; text:Star">
+        <svg class="octicon octicon-star v-align-text-bottom" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M14 6l-4.9-.64L7 1 4.9 5.36 0 6l3.6 3.26L2.67 14 7 11.67 11.33 14l-.93-4.74L14 6z"/></svg>
+        Star
+      </button>
+        <a class="social-count js-social-count" href="/vohramanav/ActiveSubspace_Kinetics/stargazers"
+           aria-label="0 users starred this repository">
+          0
+        </a>
+</form>  </div>
+
+  </li>
+
+  <li>
+        <span class="btn btn-sm btn-with-count disabled tooltipped tooltipped-sw" aria-label="Cannot fork because you own this repository and are not a member of any organizations.">
+          <svg class="octicon octicon-repo-forked v-align-text-bottom" viewBox="0 0 10 16" version="1.1" width="10" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8 1a1.993 1.993 0 0 0-1 3.72V6L5 8 3 6V4.72A1.993 1.993 0 0 0 2 1a1.993 1.993 0 0 0-1 3.72V6.5l3 3v1.78A1.993 1.993 0 0 0 5 15a1.993 1.993 0 0 0 1-3.72V9.5l3-3V4.72A1.993 1.993 0 0 0 8 1zM2 4.2C1.34 4.2.8 3.65.8 3c0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2zm3 10c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2zm3-10c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2z"/></svg>
+          Fork
+</span>
+    <a href="/vohramanav/ActiveSubspace_Kinetics/network/members" class="social-count"
+       aria-label="0 users forked this repository">
+      0
+    </a>
+  </li>
+</ul>
+
+      <h1 class="public ">
+  <svg class="octicon octicon-repo" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M4 9H3V8h1v1zm0-3H3v1h1V6zm0-2H3v1h1V4zm0-2H3v1h1V2zm8-1v12c0 .55-.45 1-1 1H6v2l-1.5-1.5L3 16v-2H1c-.55 0-1-.45-1-1V1c0-.55.45-1 1-1h10c.55 0 1 .45 1 1zm-1 10H1v2h2v-1h3v1h5v-2zm0-10H2v9h9V1z"/></svg>
+  <span class="author" itemprop="author"><a class="url fn" rel="author" href="/vohramanav">vohramanav</a></span><!--
+--><span class="path-divider">/</span><!--
+--><strong itemprop="name"><a data-pjax="#js-repo-pjax-container" href="/vohramanav/ActiveSubspace_Kinetics">ActiveSubspace_Kinetics</a></strong>
+
+</h1>
+
+    </div>
+
+<nav class="reponav js-repo-nav js-sidenav-container-pjax container"
+     itemscope
+     itemtype="http://schema.org/BreadcrumbList"
+     role="navigation"
+     data-pjax="#js-repo-pjax-container">
+
+  <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
+    <a class="js-selected-navigation-item selected reponav-item" itemprop="url" data-hotkey="g c" data-selected-links="repo_source repo_downloads repo_commits repo_releases repo_tags repo_branches repo_packages /vohramanav/ActiveSubspace_Kinetics" href="/vohramanav/ActiveSubspace_Kinetics">
+      <svg class="octicon octicon-code" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M9.5 3L8 4.5 11.5 8 8 11.5 9.5 13 14 8 9.5 3zm-5 0L0 8l4.5 5L6 11.5 2.5 8 6 4.5 4.5 3z"/></svg>
+      <span itemprop="name">Code</span>
+      <meta itemprop="position" content="1">
+</a>  </span>
+
+    <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
+      <a itemprop="url" data-hotkey="g i" class="js-selected-navigation-item reponav-item" data-selected-links="repo_issues repo_labels repo_milestones /vohramanav/ActiveSubspace_Kinetics/issues" href="/vohramanav/ActiveSubspace_Kinetics/issues">
+        <svg class="octicon octicon-issue-opened" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7 2.3c3.14 0 5.7 2.56 5.7 5.7s-2.56 5.7-5.7 5.7A5.71 5.71 0 0 1 1.3 8c0-3.14 2.56-5.7 5.7-5.7zM7 1C3.14 1 0 4.14 0 8s3.14 7 7 7 7-3.14 7-7-3.14-7-7-7zm1 3H6v5h2V4zm0 6H6v2h2v-2z"/></svg>
+        <span itemprop="name">Issues</span>
+        <span class="Counter">0</span>
+        <meta itemprop="position" content="2">
+</a>    </span>
+
+  <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
+    <a data-hotkey="g p" itemprop="url" class="js-selected-navigation-item reponav-item" data-selected-links="repo_pulls checks /vohramanav/ActiveSubspace_Kinetics/pulls" href="/vohramanav/ActiveSubspace_Kinetics/pulls">
+      <svg class="octicon octicon-git-pull-request" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M11 11.28V5c-.03-.78-.34-1.47-.94-2.06C9.46 2.35 8.78 2.03 8 2H7V0L4 3l3 3V4h1c.27.02.48.11.69.31.21.2.3.42.31.69v6.28A1.993 1.993 0 0 0 10 15a1.993 1.993 0 0 0 1-3.72zm-1 2.92c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2zM4 3c0-1.11-.89-2-2-2a1.993 1.993 0 0 0-1 3.72v6.56A1.993 1.993 0 0 0 2 15a1.993 1.993 0 0 0 1-3.72V4.72c.59-.34 1-.98 1-1.72zm-.8 10c0 .66-.55 1.2-1.2 1.2-.65 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2zM2 4.2C1.34 4.2.8 3.65.8 3c0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2z"/></svg>
+      <span itemprop="name">Pull requests</span>
+      <span class="Counter">0</span>
+      <meta itemprop="position" content="3">
+</a>  </span>
+
+    <a data-hotkey="g b" class="js-selected-navigation-item reponav-item" data-selected-links="repo_projects new_repo_project repo_project /vohramanav/ActiveSubspace_Kinetics/projects" href="/vohramanav/ActiveSubspace_Kinetics/projects">
+      <svg class="octicon octicon-project" viewBox="0 0 15 16" version="1.1" width="15" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M10 12h3V2h-3v10zm-4-2h3V2H6v8zm-4 4h3V2H2v12zm-1 1h13V1H1v14zM14 0H1a1 1 0 0 0-1 1v14a1 1 0 0 0 1 1h13a1 1 0 0 0 1-1V1a1 1 0 0 0-1-1z"/></svg>
+      Projects
+      <span class="Counter" >0</span>
+</a>
+
+    <a class="js-selected-navigation-item reponav-item" data-hotkey="g w" data-selected-links="repo_wiki /vohramanav/ActiveSubspace_Kinetics/wiki" href="/vohramanav/ActiveSubspace_Kinetics/wiki">
+      <svg class="octicon octicon-book" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M3 5h4v1H3V5zm0 3h4V7H3v1zm0 2h4V9H3v1zm11-5h-4v1h4V5zm0 2h-4v1h4V7zm0 2h-4v1h4V9zm2-6v9c0 .55-.45 1-1 1H9.5l-1 1-1-1H2c-.55 0-1-.45-1-1V3c0-.55.45-1 1-1h5.5l1 1 1-1H15c.55 0 1 .45 1 1zm-8 .5L7.5 3H2v9h6V3.5zm7-.5H9.5l-.5.5V12h6V3z"/></svg>
+      Wiki
+</a>
+
+  <a class="js-selected-navigation-item reponav-item" data-selected-links="repo_graphs repo_contributors dependency_graph pulse /vohramanav/ActiveSubspace_Kinetics/pulse" href="/vohramanav/ActiveSubspace_Kinetics/pulse">
+    <svg class="octicon octicon-graph" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M16 14v1H0V0h1v14h15zM5 13H3V8h2v5zm4 0H7V3h2v10zm4 0h-2V6h2v7z"/></svg>
+    Insights
+</a>
+    <a class="js-selected-navigation-item reponav-item" data-selected-links="repo_settings repo_branch_settings hooks integration_installations repo_keys_settings issue_template_editor /vohramanav/ActiveSubspace_Kinetics/settings" href="/vohramanav/ActiveSubspace_Kinetics/settings">
+      <svg class="octicon octicon-gear" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M14 8.77v-1.6l-1.94-.64-.45-1.09.88-1.84-1.13-1.13-1.81.91-1.09-.45-.69-1.92h-1.6l-.63 1.94-1.11.45-1.84-.88-1.13 1.13.91 1.81-.45 1.09L0 7.23v1.59l1.94.64.45 1.09-.88 1.84 1.13 1.13 1.81-.91 1.09.45.69 1.92h1.59l.63-1.94 1.11-.45 1.84.88 1.13-1.13-.92-1.81.47-1.09L14 8.75v.02zM7 11c-1.66 0-3-1.34-3-3s1.34-3 3-3 3 1.34 3 3-1.34 3-3 3z"/></svg>
+      Settings
+</a>
+</nav>
+
+  </div>
+
+<div class="container new-discussion-timeline experiment-repo-nav  ">
+  <div class="repository-content ">
+
+  <a class="d-none js-permalink-shortcut" data-hotkey="y" href="/vohramanav/ActiveSubspace_Kinetics/blob/de781a8c95a3c79c7d21d73d7c02fb74af163b10/Paper/app.tex">Permalink</a>
+
+  <!-- blob contrib key: blob_contributors:v21:b164a66d7003a37e5d351877d3770537 -->
+
+  <div class="file-navigation">
+
+<div class="select-menu branch-select-menu js-menu-container js-select-menu float-left">
+  <button class=" btn btn-sm select-menu-button js-menu-target css-truncate" data-hotkey="w"
+
+    type="button" aria-label="Switch branches or tags" aria-expanded="false" aria-haspopup="true">
+      <i>Branch:</i>
+      <span class="js-select-button css-truncate-target">master</span>
+  </button>
+
+  <div class="select-menu-modal-holder js-menu-content js-navigation-container" data-pjax>
+
+    <div class="select-menu-modal">
+      <div class="select-menu-header">
+        <svg class="octicon octicon-x js-menu-close" role="img" aria-label="Close" viewBox="0 0 12 16" version="1.1" width="12" height="16"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
+        <span class="select-menu-title">Switch branches/tags</span>
+      </div>
+
+      <div class="select-menu-filters">
+        <div class="select-menu-text-filter">
+          <input type="text" aria-label="Find or create a branch…" id="context-commitish-filter-field" class="form-control js-filterable-field js-navigation-enable" placeholder="Find or create a branch…">
+        </div>
+        <div class="select-menu-tabs">
+          <ul>
+            <li class="select-menu-tab">
+              <a href="#" data-tab-filter="branches" data-filter-placeholder="Find or create a branch…" class="js-select-menu-tab" role="tab">Branches</a>
+            </li>
+            <li class="select-menu-tab">
+              <a href="#" data-tab-filter="tags" data-filter-placeholder="Find a tag…" class="js-select-menu-tab" role="tab">Tags</a>
+            </li>
+          </ul>
+        </div>
+      </div>
+
+      <div class="select-menu-list select-menu-tab-bucket js-select-menu-tab-bucket" data-tab-filter="branches" role="menu">
+
+        <div data-filterable-for="context-commitish-filter-field" data-filterable-type="substring">
+
+            <a class="select-menu-item js-navigation-item js-navigation-open selected"
+               href="/vohramanav/ActiveSubspace_Kinetics/blob/master/Paper/app.tex"
+               data-name="master"
+               data-skip-pjax="true"
+               rel="nofollow">
+              <svg class="octicon octicon-check select-menu-item-icon" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5L12 5z"/></svg>
+              <span class="select-menu-item-text css-truncate-target js-select-menu-filter-text">
+                master
+              </span>
+            </a>
+        </div>
+
+          <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="js-create-branch select-menu-item select-menu-new-item-form js-navigation-item js-new-item-form" action="/vohramanav/ActiveSubspace_Kinetics/branches" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="Hqg2uRrsxvbZqMWPMWculEYvJcvnUljWYRsfJNiUGjzu88KqLa7MHSIm4IzPeIe/3TDui4Y6K/ivCRpqhQ3FLA==" />
+          <svg class="octicon octicon-git-branch select-menu-item-icon" viewBox="0 0 10 16" version="1.1" width="10" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M10 5c0-1.11-.89-2-2-2a1.993 1.993 0 0 0-1 3.72v.3c-.02.52-.23.98-.63 1.38-.4.4-.86.61-1.38.63-.83.02-1.48.16-2 .45V4.72a1.993 1.993 0 0 0-1-3.72C.88 1 0 1.89 0 3a2 2 0 0 0 1 1.72v6.56c-.59.35-1 .99-1 1.72 0 1.11.89 2 2 2 1.11 0 2-.89 2-2 0-.53-.2-1-.53-1.36.09-.06.48-.41.59-.47.25-.11.56-.17.94-.17 1.05-.05 1.95-.45 2.75-1.25S8.95 7.77 9 6.73h-.02C9.59 6.37 10 5.73 10 5zM2 1.8c.66 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2C1.35 4.2.8 3.65.8 3c0-.65.55-1.2 1.2-1.2zm0 12.41c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2zm6-8c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2z"/></svg>
+            <div class="select-menu-item-text">
+              <span class="select-menu-item-heading">Create branch: <span class="js-new-item-name"></span></span>
+              <span class="description">from ‘master’</span>
+            </div>
+            <input type="hidden" name="name" id="name" class="js-new-item-value">
+            <input type="hidden" name="branch" id="branch" value="master">
+            <input type="hidden" name="path" id="path" value="Paper/app.tex">
+</form>
+      </div>
+
+      <div class="select-menu-list select-menu-tab-bucket js-select-menu-tab-bucket" data-tab-filter="tags">
+        <div data-filterable-for="context-commitish-filter-field" data-filterable-type="substring">
+
+        </div>
+
+        <div class="select-menu-no-results">Nothing to show</div>
+      </div>
+
+    </div>
+  </div>
+</div>
+
+    <div class="BtnGroup float-right">
+      <a href="/vohramanav/ActiveSubspace_Kinetics/find/master"
+            class="js-pjax-capture-input btn btn-sm BtnGroup-item"
+            data-pjax
+            data-hotkey="t">
+        Find file
+      </a>
+      <clipboard-copy for="blob-path" class="btn btn-sm BtnGroup-item">
+        Copy path
+      </clipboard-copy>
+    </div>
+    <div id="blob-path" class="breadcrumb">
+      <span class="repo-root js-repo-root"><span class="js-path-segment"><a data-pjax="true" href="/vohramanav/ActiveSubspace_Kinetics"><span>ActiveSubspace_Kinetics</span></a></span></span><span class="separator">/</span><span class="js-path-segment"><a data-pjax="true" href="/vohramanav/ActiveSubspace_Kinetics/tree/master/Paper"><span>Paper</span></a></span><span class="separator">/</span><strong class="final-path">app.tex</strong>
+    </div>
+  </div>
+
+  <include-fragment src="/vohramanav/ActiveSubspace_Kinetics/contributors/master/Paper/app.tex" class="commit-tease commit-loader">
+    <div>
+      Fetching contributors&hellip;
+    </div>
+
+    <div class="commit-tease-contributors">
+        <img alt="" class="loader-loading float-left" src="https://assets-cdn.github.com/images/spinners/octocat-spinner-32-EAF2F5.gif" width="16" height="16" />
+      <span class="loader-error">Cannot retrieve contributors at this time</span>
+    </div>
+</include-fragment>
+
+  <div class="file">
+    <div class="file-header">
+  <div class="file-actions">
+
+    <div class="BtnGroup">
+      <a id="raw-url" class="btn btn-sm BtnGroup-item" href="/vohramanav/ActiveSubspace_Kinetics/raw/master/Paper/app.tex">Raw</a>
+        <a class="btn btn-sm js-update-url-with-hash BtnGroup-item" data-hotkey="b" href="/vohramanav/ActiveSubspace_Kinetics/blame/master/Paper/app.tex">Blame</a>
+      <a rel="nofollow" class="btn btn-sm BtnGroup-item" href="/vohramanav/ActiveSubspace_Kinetics/commits/master/Paper/app.tex">History</a>
+    </div>
+
+          <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="inline-form js-update-url-with-hash" action="/vohramanav/ActiveSubspace_Kinetics/edit/master/Paper/app.tex" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="ZclOij/AHtI5EL4dB5xh3b45JZkHqxOIBRanVfE3usFgSq86Mzn05lB2dE1IF2Hp9fA40gmcF54iyVQ0XRXCww==" />
+            <button class="btn-octicon tooltipped tooltipped-nw" type="submit"
+              aria-label="Edit this file" data-hotkey="e" data-disable-with>
+              <svg class="octicon octicon-pencil" viewBox="0 0 14 16" version="1.1" width="14" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M0 12v3h3l8-8-3-3-8 8zm3 2H1v-2h1v1h1v1zm10.3-9.3L12 6 9 3l1.3-1.3a.996.996 0 0 1 1.41 0l1.59 1.59c.39.39.39 1.02 0 1.41z"/></svg>
+            </button>
+</form>
+        <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="inline-form" action="/vohramanav/ActiveSubspace_Kinetics/delete/master/Paper/app.tex" accept-charset="UTF-8" method="post"><input name="utf8" type="hidden" value="&#x2713;" /><input type="hidden" name="authenticity_token" value="jq2W0gkiU13iPoX9HYL7sJjfuyOvQ95T4lcgMOuXHsLA4Ze1W50KlNwoOC0AWxrs3DnNpI3k/tZ9TxFdEW+6zA==" />
+          <button class="btn-octicon btn-octicon-danger tooltipped tooltipped-nw" type="submit"
+            aria-label="Delete this file" data-disable-with>
+            <svg class="octicon octicon-trashcan" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M11 2H9c0-.55-.45-1-1-1H5c-.55 0-1 .45-1 1H2c-.55 0-1 .45-1 1v1c0 .55.45 1 1 1v9c0 .55.45 1 1 1h7c.55 0 1-.45 1-1V5c.55 0 1-.45 1-1V3c0-.55-.45-1-1-1zm-1 12H3V5h1v8h1V5h1v8h1V5h1v8h1V5h1v9zm1-10H2V3h9v1z"/></svg>
+          </button>
+</form>  </div>
+
+  <div class="file-info">
+      196 lines (154 sloc)
+      <span class="file-info-divider"></span>
+    8.85 KB
+  </div>
+</div>
+
+  <div itemprop="text" class="blob-wrapper data type-tex">
+      <table class="highlight tab-size js-file-line-container" data-tab-size="8">
+      <tr>
+        <td id="L1" class="blob-num js-line-number" data-line-number="1"></td>
+        <td id="LC1" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\section</span>{<span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\text</span>{H}_<span class="pl-c1">2</span><span class="pl-pds">$</span></span>/<span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\text</span>{O}_<span class="pl-c1">2</span><span class="pl-pds">$</span></span> reaction kinetics: high-dimensional case}</td>
+      </tr>
+      <tr>
+        <td id="L2" class="blob-num js-line-number" data-line-number="2"></td>
+        <td id="LC2" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\label</span>{sec:app}</td>
+      </tr>
+      <tr>
+        <td id="L3" class="blob-num js-line-number" data-line-number="3"></td>
+        <td id="LC3" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L4" class="blob-num js-line-number" data-line-number="4"></td>
+        <td id="LC4" class="blob-code blob-code-inner js-file-line">For the high-dimensional case, we aim to investigate the impact of the uncertain</td>
+      </tr>
+      <tr>
+        <td id="L5" class="blob-num js-line-number" data-line-number="5"></td>
+        <td id="LC5" class="blob-code blob-code-inner js-file-line">pre-exponents (<span class="pl-s"><span class="pl-pds">$</span>A_i<span class="pl-pds">$</span></span>&#39;s) as well as the activation energies (<span class="pl-s"><span class="pl-pds">$</span>E_{a,i}<span class="pl-pds">$</span></span>&#39;s) on ignition </td>
+      </tr>
+      <tr>
+        <td id="L6" class="blob-num js-line-number" data-line-number="6"></td>
+        <td id="LC6" class="blob-code blob-code-inner js-file-line">delay during the H<span class="pl-s"><span class="pl-pds">$</span>_<span class="pl-c1">2</span><span class="pl-pds">$</span></span>/O<span class="pl-s"><span class="pl-pds">$</span>_<span class="pl-c1">2</span><span class="pl-pds">$</span></span> reaction.
+Prior distributions for <span class="pl-s"><span class="pl-pds">$</span>A_i<span class="pl-pds">$</span></span>&#39;s were considered</td>
+      </tr>
+      <tr>
+        <td id="L7" class="blob-num js-line-number" data-line-number="7"></td>
+        <td id="LC7" class="blob-code blob-code-inner js-file-line">to be the same as provided earlier in section~<span class="pl-c1">\ref</span>{sec:method}.
+The <span class="pl-s"><span class="pl-pds">$</span>E_{a,i}<span class="pl-pds">$</span></span>&#39;s for all</td>
+      </tr>
+      <tr>
+        <td id="L8" class="blob-num js-line-number" data-line-number="8"></td>
+        <td id="LC8" class="blob-code blob-code-inner js-file-line">reactions except <span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\mathcal</span>{R}_<span class="pl-c1">6</span><span class="pl-pds">$</span></span> -- <span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\mathcal</span>{R}_<span class="pl-c1">9</span><span class="pl-pds">$</span></span> and <span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\mathcal</span>{R}_{13}<span class="pl-pds">$</span></span> (due to </td>
+      </tr>
+      <tr>
+        <td id="L9" class="blob-num js-line-number" data-line-number="9"></td>
+        <td id="LC9" class="blob-code blob-code-inner js-file-line">zero nominal values for <span class="pl-s"><span class="pl-pds">$</span>E_a<span class="pl-pds">$</span></span>)</td>
+      </tr>
+      <tr>
+        <td id="L10" class="blob-num js-line-number" data-line-number="10"></td>
+        <td id="LC10" class="blob-code blob-code-inner js-file-line">were considered to be uncertain and uniformly distributed in the interval: </td>
+      </tr>
+      <tr>
+        <td id="L11" class="blob-num js-line-number" data-line-number="11"></td>
+        <td id="LC11" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>[<span class="pl-c1">0.99</span>E_{a,i}^<span class="pl-c1">\ast</span>, <span class="pl-c1">1.01</span>E_{a,i}^<span class="pl-c1">\ast</span>]<span class="pl-pds">$</span></span>.
+The nominal values, <span class="pl-s"><span class="pl-pds">$</span>E_{a,i}^<span class="pl-c1">\ast</span><span class="pl-pds">$</span></span></td>
+      </tr>
+      <tr>
+        <td id="L12" class="blob-num js-line-number" data-line-number="12"></td>
+        <td id="LC12" class="blob-code blob-code-inner js-file-line">corresponding to the different reaction rates are provided in~<span class="pl-c1">\cite</span>{Yetter:1991}. </td>
+      </tr>
+      <tr>
+        <td id="L13" class="blob-num js-line-number" data-line-number="13"></td>
+        <td id="LC13" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L14" class="blob-num js-line-number" data-line-number="14"></td>
+        <td id="LC14" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\subsection</span>{Computing the active subspace}</td>
+      </tr>
+      <tr>
+        <td id="L15" class="blob-num js-line-number" data-line-number="15"></td>
+        <td id="LC15" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L16" class="blob-num js-line-number" data-line-number="16"></td>
+        <td id="LC16" class="blob-code blob-code-inner js-file-line">The gradient-based approach discussed earlier in~<span class="pl-c1">\ref</span>{sub:grad} was used to compute the</td>
+      </tr>
+      <tr>
+        <td id="L17" class="blob-num js-line-number" data-line-number="17"></td>
+        <td id="LC17" class="blob-code blob-code-inner js-file-line">active subspace. Using the iterative procedure, the convergence of the eigenvectors</td>
+      </tr>
+      <tr>
+        <td id="L18" class="blob-num js-line-number" data-line-number="18"></td>
+        <td id="LC18" class="blob-code blob-code-inner js-file-line">was examined by tracking <span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\max</span>(<span class="pl-c1">\delta</span> <span class="pl-c1">\hat</span>{<span class="pl-c1">\mat</span>{W}}_{1,j}^{(i)})<span class="pl-pds">$</span></span>, plotted in </td>
+      </tr>
+      <tr>
+        <td id="L19" class="blob-num js-line-number" data-line-number="19"></td>
+        <td id="LC19" class="blob-code blob-code-inner js-file-line">Figure~<span class="pl-c1">\ref</span>{fig:conv_app} (right).
+In Figure~<span class="pl-c1">\ref</span>{fig:conv_app} (left), individual</td>
+      </tr>
+      <tr>
+        <td id="L20" class="blob-num js-line-number" data-line-number="20"></td>
+        <td id="LC20" class="blob-code blob-code-inner js-file-line">components of the converged eigenvector are illustrated. The convergence was established using</td>
+      </tr>
+      <tr>
+        <td id="L21" class="blob-num js-line-number" data-line-number="21"></td>
+        <td id="LC21" class="blob-code blob-code-inner js-file-line">a <span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\tau</span><span class="pl-pds">$</span></span> value of 0.02.
+Finite difference was used to compute model gradients at 40 samples in</td>
+      </tr>
+      <tr>
+        <td id="L22" class="blob-num js-line-number" data-line-number="22"></td>
+        <td id="LC22" class="blob-code blob-code-inner js-file-line">the input domain. Hence, a total of 1360 model evaluations were required for this purpose.  </td>
+      </tr>
+      <tr>
+        <td id="L23" class="blob-num js-line-number" data-line-number="23"></td>
+        <td id="LC23" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span></span></td>
+      </tr>
+      <tr>
+        <td id="L24" class="blob-num js-line-number" data-line-number="24"></td>
+        <td id="LC24" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\begin</span>{figure}[htbp]</td>
+      </tr>
+      <tr>
+        <td id="L25" class="blob-num js-line-number" data-line-number="25"></td>
+        <td id="LC25" class="blob-code blob-code-inner js-file-line"> <span class="pl-c1">\begin</span>{center}</td>
+      </tr>
+      <tr>
+        <td id="L26" class="blob-num js-line-number" data-line-number="26"></td>
+        <td id="LC26" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\includegraphics</span>[width=0.8<span class="pl-c1">\textwidth</span>]{./Figures/eigv10}</td>
+      </tr>
+      <tr>
+        <td id="L27" class="blob-num js-line-number" data-line-number="27"></td>
+        <td id="LC27" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\caption</span>{Left: An illustrative comparison of individual components of the converged dominant eigenvector obtained</td>
+      </tr>
+      <tr>
+        <td id="L28" class="blob-num js-line-number" data-line-number="28"></td>
+        <td id="LC28" class="blob-code blob-code-inner js-file-line">using the two approaches discussed in~<span class="pl-c1">\ref</span>{sub:grad} and~<span class="pl-c1">\ref</span>{sub:gradfree}.
+Right: The quantity,  </td>
+      </tr>
+      <tr>
+        <td id="L29" class="blob-num js-line-number" data-line-number="29"></td>
+        <td id="LC29" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\max</span>(<span class="pl-c1">\delta</span> <span class="pl-c1">\hat</span>{<span class="pl-c1">\mat</span>{W}}_{1,j}^{(i)})<span class="pl-pds">$</span></span></td>
+      </tr>
+      <tr>
+        <td id="L30" class="blob-num js-line-number" data-line-number="30"></td>
+        <td id="LC30" class="blob-code blob-code-inner js-file-line">is plotted for successive iterations to illustrate the convergence behavior.}</td>
+      </tr>
+      <tr>
+        <td id="L31" class="blob-num js-line-number" data-line-number="31"></td>
+        <td id="LC31" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\label</span>{fig:conv_app}</td>
+      </tr>
+      <tr>
+        <td id="L32" class="blob-num js-line-number" data-line-number="32"></td>
+        <td id="LC32" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\end</span>{center}</td>
+      </tr>
+      <tr>
+        <td id="L33" class="blob-num js-line-number" data-line-number="33"></td>
+        <td id="LC33" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\end</span>{figure}</td>
+      </tr>
+      <tr>
+        <td id="L34" class="blob-num js-line-number" data-line-number="34"></td>
+        <td id="LC34" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span></span></td>
+      </tr>
+      <tr>
+        <td id="L35" class="blob-num js-line-number" data-line-number="35"></td>
+        <td id="LC35" class="blob-code blob-code-inner js-file-line"> </td>
+      </tr>
+      <tr>
+        <td id="L36" class="blob-num js-line-number" data-line-number="36"></td>
+        <td id="LC36" class="blob-code blob-code-inner js-file-line">In Figure~<span class="pl-c1">\ref</span>{fig:hd} (left), we illustrate the resulting eigenvalue spectrum.
+The second eigenvalue is</td>
+      </tr>
+      <tr>
+        <td id="L37" class="blob-num js-line-number" data-line-number="37"></td>
+        <td id="LC37" class="blob-code blob-code-inner js-file-line">found to be roughly two orders of magnitude smaller than the first, indicating that the active subspace</td>
+      </tr>
+      <tr>
+        <td id="L38" class="blob-num js-line-number" data-line-number="38"></td>
+        <td id="LC38" class="blob-code blob-code-inner js-file-line">is 1-dimensional. This is further confirmed by the SSP plots in Figure~<span class="pl-c1">\ref</span>{fig:hd} (right). </td>
+      </tr>
+      <tr>
+        <td id="L39" class="blob-num js-line-number" data-line-number="39"></td>
+        <td id="LC39" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span></span></td>
+      </tr>
+      <tr>
+        <td id="L40" class="blob-num js-line-number" data-line-number="40"></td>
+        <td id="LC40" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\begin</span>{figure}[htbp]</td>
+      </tr>
+      <tr>
+        <td id="L41" class="blob-num js-line-number" data-line-number="41"></td>
+        <td id="LC41" class="blob-code blob-code-inner js-file-line"> <span class="pl-c1">\begin</span>{center}</td>
+      </tr>
+      <tr>
+        <td id="L42" class="blob-num js-line-number" data-line-number="42"></td>
+        <td id="LC42" class="blob-code blob-code-inner js-file-line">   <span class="pl-c1">\includegraphics</span>[width=0.45<span class="pl-c1">\textwidth</span>]{./Figures/eig_33D}</td>
+      </tr>
+      <tr>
+        <td id="L43" class="blob-num js-line-number" data-line-number="43"></td>
+        <td id="LC43" class="blob-code blob-code-inner js-file-line">   <span class="pl-c1">\includegraphics</span>[width=0.45<span class="pl-c1">\textwidth</span>]{./Figures/ssp_33D}</td>
+      </tr>
+      <tr>
+        <td id="L44" class="blob-num js-line-number" data-line-number="44"></td>
+        <td id="LC44" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\caption</span>{Left: Eigenvalue spectrum of <span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\hat</span>{<span class="pl-c1">\mat</span>{C}}<span class="pl-pds">$</span></span>.
+Right: SSP of the computed active subspace, and the</td>
+      </tr>
+      <tr>
+        <td id="L45" class="blob-num js-line-number" data-line-number="45"></td>
+        <td id="LC45" class="blob-code blob-code-inner js-file-line">corresponding 1-D linear fit.} </td>
+      </tr>
+      <tr>
+        <td id="L46" class="blob-num js-line-number" data-line-number="46"></td>
+        <td id="LC46" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\label</span>{fig:hd}</td>
+      </tr>
+      <tr>
+        <td id="L47" class="blob-num js-line-number" data-line-number="47"></td>
+        <td id="LC47" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\end</span>{center}</td>
+      </tr>
+      <tr>
+        <td id="L48" class="blob-num js-line-number" data-line-number="48"></td>
+        <td id="LC48" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\end</span>{figure}</td>
+      </tr>
+      <tr>
+        <td id="L49" class="blob-num js-line-number" data-line-number="49"></td>
+        <td id="LC49" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span></span></td>
+      </tr>
+      <tr>
+        <td id="L50" class="blob-num js-line-number" data-line-number="50"></td>
+        <td id="LC50" class="blob-code blob-code-inner js-file-line">Specifically</td>
+      </tr>
+      <tr>
+        <td id="L51" class="blob-num js-line-number" data-line-number="51"></td>
+        <td id="LC51" class="blob-code blob-code-inner js-file-line">in the case of gradient-based approach, the SSP exhibits a linear variation that is reasonably</td>
+      </tr>
+      <tr>
+        <td id="L52" class="blob-num js-line-number" data-line-number="52"></td>
+        <td id="LC52" class="blob-code blob-code-inner js-file-line">captured by its linear fit, <span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\tilde</span>{G}<span class="pl-pds">$</span></span>.
+Whereas, the SSP based on the gradient-free approach</td>
+      </tr>
+      <tr>
+        <td id="L53" class="blob-num js-line-number" data-line-number="53"></td>
+        <td id="LC53" class="blob-code blob-code-inner js-file-line">is relatively more scattered. Consequently, the linear fit exhibits a discrepancy with the corresponding</td>
+      </tr>
+      <tr>
+        <td id="L54" class="blob-num js-line-number" data-line-number="54"></td>
+        <td id="LC54" class="blob-code blob-code-inner js-file-line">fit for the gradient-based SSP. </td>
+      </tr>
+      <tr>
+        <td id="L55" class="blob-num js-line-number" data-line-number="55"></td>
+        <td id="LC55" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L56" class="blob-num js-line-number" data-line-number="56"></td>
+        <td id="LC56" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\subsection</span>{Surrogate Verification}</td>
+      </tr>
+      <tr>
+        <td id="L57" class="blob-num js-line-number" data-line-number="57"></td>
+        <td id="LC57" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\label</span>{sub:verify}</td>
+      </tr>
+      <tr>
+        <td id="L58" class="blob-num js-line-number" data-line-number="58"></td>
+        <td id="LC58" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L59" class="blob-num js-line-number" data-line-number="59"></td>
+        <td id="LC59" class="blob-code blob-code-inner js-file-line">The 1-dimensional surrogate (<span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\tilde</span>{G}<span class="pl-pds">$</span></span>) is investigated for its accuracy as well as the ability to capture the </td>
+      </tr>
+      <tr>
+        <td id="L60" class="blob-num js-line-number" data-line-number="60"></td>
+        <td id="LC60" class="blob-code blob-code-inner js-file-line">uncertainty in the</td>
+      </tr>
+      <tr>
+        <td id="L61" class="blob-num js-line-number" data-line-number="61"></td>
+        <td id="LC61" class="blob-code blob-code-inner js-file-line">model output in two ways.
+Firstly, we estimate the relative L-2 norm of the discrepancy~(<span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\varepsilon</span>_d<span class="pl-pds">$</span></span>)</td>
+      </tr>
+      <tr>
+        <td id="L62" class="blob-num js-line-number" data-line-number="62"></td>
+        <td id="LC62" class="blob-code blob-code-inner js-file-line">between estimates of ignition delay in the case of H<span class="pl-s"><span class="pl-pds">$</span>_<span class="pl-c1">2</span><span class="pl-pds">$</span></span>/O<span class="pl-s"><span class="pl-pds">$</span>_<span class="pl-c1">2</span><span class="pl-pds">$</span></span> reaction, obtained using the </td>
+      </tr>
+      <tr>
+        <td id="L63" class="blob-num js-line-number" data-line-number="63"></td>
+        <td id="LC63" class="blob-code blob-code-inner js-file-line">model output and the surrogate in the following equation:</td>
+      </tr>
+      <tr>
+        <td id="L64" class="blob-num js-line-number" data-line-number="64"></td>
+        <td id="LC64" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span></span></td>
+      </tr>
+      <tr>
+        <td id="L65" class="blob-num js-line-number" data-line-number="65"></td>
+        <td id="LC65" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\be</span></td>
+      </tr>
+      <tr>
+        <td id="L66" class="blob-num js-line-number" data-line-number="66"></td>
+        <td id="LC66" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\varepsilon</span>_d = <span class="pl-c1">\frac</span>{<span class="pl-cce">\|</span>G(<span class="pl-c1">\bm</span>{<span class="pl-c1">\xi</span>}) - <span class="pl-c1">\tilde</span>{G}(<span class="pl-c1">\bm</span>{<span class="pl-c1">\xi</span>})<span class="pl-cce">\|</span>_2}{<span class="pl-cce">\|</span>G(<span class="pl-c1">\bm</span>{<span class="pl-c1">\xi</span>})<span class="pl-cce">\|</span>_2}</td>
+      </tr>
+      <tr>
+        <td id="L67" class="blob-num js-line-number" data-line-number="67"></td>
+        <td id="LC67" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\ee</span></td>
+      </tr>
+      <tr>
+        <td id="L68" class="blob-num js-line-number" data-line-number="68"></td>
+        <td id="LC68" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span></span></td>
+      </tr>
+      <tr>
+        <td id="L69" class="blob-num js-line-number" data-line-number="69"></td>
+        <td id="LC69" class="blob-code blob-code-inner js-file-line">The relative norm, <span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\varepsilon</span>_d<span class="pl-pds">$</span></span> was estimated to be 1.35<span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\times</span><span class="pl-c1">10</span>^{-2}<span class="pl-pds">$</span></span> and 1.02<span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\times</span><span class="pl-c1">10</span>^{-1}<span class="pl-pds">$</span></span></td>
+      </tr>
+      <tr>
+        <td id="L70" class="blob-num js-line-number" data-line-number="70"></td>
+        <td id="LC70" class="blob-code blob-code-inner js-file-line">using <span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\tilde</span>{G}<span class="pl-pds">$</span></span> from the gradient-based approach and the gradient-free approach respectively.
+Hence,</td>
+      </tr>
+      <tr>
+        <td id="L71" class="blob-num js-line-number" data-line-number="71"></td>
+        <td id="LC71" class="blob-code blob-code-inner js-file-line">an order of magnitude increase in <span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\varepsilon</span>_d<span class="pl-pds">$</span></span> was estimated in the case of gradient-free approach.</td>
+      </tr>
+      <tr>
+        <td id="L72" class="blob-num js-line-number" data-line-number="72"></td>
+        <td id="LC72" class="blob-code blob-code-inner js-file-line">Model evaluations (<span class="pl-s"><span class="pl-pds">$</span>G<span class="pl-pds">$</span></span>) at 10<span class="pl-s"><span class="pl-pds">$</span>^{4}<span class="pl-pds">$</span></span> samples in the input domain were used for this purpose.
+</td>
+      </tr>
+      <tr>
+        <td id="L73" class="blob-num js-line-number" data-line-number="73"></td>
+        <td id="LC73" class="blob-code blob-code-inner js-file-line">Hence, in the norm-sense, it can be said that the 1-dimensional surrogate in the active subspace, obtained</td>
+      </tr>
+      <tr>
+        <td id="L74" class="blob-num js-line-number" data-line-number="74"></td>
+        <td id="LC74" class="blob-code blob-code-inner js-file-line">using the gradient-based approach is remarkably accurate with a relative error of less than 1.5<span class="pl-s"><span class="pl-pds">$</span><span class="pl-cce">\%</span><span class="pl-pds">$</span></span>.</td>
+      </tr>
+      <tr>
+        <td id="L75" class="blob-num js-line-number" data-line-number="75"></td>
+        <td id="LC75" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L76" class="blob-num js-line-number" data-line-number="76"></td>
+        <td id="LC76" class="blob-code blob-code-inner js-file-line">Secondly, we verify the accuracy of the two surrogates in a probabilistic setting. In particular, we compare </td>
+      </tr>
+      <tr>
+        <td id="L77" class="blob-num js-line-number" data-line-number="77"></td>
+        <td id="LC77" class="blob-code blob-code-inner js-file-line">probability density functions (PDFs) obtained using the true set of model evaluations, and 1-dimensional</td>
+      </tr>
+      <tr>
+        <td id="L78" class="blob-num js-line-number" data-line-number="78"></td>
+        <td id="LC78" class="blob-code blob-code-inner js-file-line">surrogates from the two approaches, in Figure~<span class="pl-c1">\ref</span>{fig:pdf_33D}.
+Note that the three PDFs were evaluated </td>
+      </tr>
+      <tr>
+        <td id="L79" class="blob-num js-line-number" data-line-number="79"></td>
+        <td id="LC79" class="blob-code blob-code-inner js-file-line">using the same set of 10<span class="pl-s"><span class="pl-pds">$</span>^<span class="pl-c1">4</span><span class="pl-pds">$</span></span> samples in the cross-validation set.
+</td>
+      </tr>
+      <tr>
+        <td id="L80" class="blob-num js-line-number" data-line-number="80"></td>
+        <td id="LC80" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span></span></td>
+      </tr>
+      <tr>
+        <td id="L81" class="blob-num js-line-number" data-line-number="81"></td>
+        <td id="LC81" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\begin</span>{figure}[htbp]</td>
+      </tr>
+      <tr>
+        <td id="L82" class="blob-num js-line-number" data-line-number="82"></td>
+        <td id="LC82" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\begin</span>{center}</td>
+      </tr>
+      <tr>
+        <td id="L83" class="blob-num js-line-number" data-line-number="83"></td>
+        <td id="LC83" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\begin</span>{minipage}[htbp]{.25<span class="pl-c1">\linewidth</span>}</td>
+      </tr>
+      <tr>
+        <td id="L84" class="blob-num js-line-number" data-line-number="84"></td>
+        <td id="LC84" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\vspace</span>{0pt}</td>
+      </tr>
+      <tr>
+        <td id="L85" class="blob-num js-line-number" data-line-number="85"></td>
+        <td id="LC85" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>\centering</span></td>
+      </tr>
+      <tr>
+        <td id="L86" class="blob-num js-line-number" data-line-number="86"></td>
+        <td id="LC86" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\hspace</span>{-25mm}</td>
+      </tr>
+      <tr>
+        <td id="L87" class="blob-num js-line-number" data-line-number="87"></td>
+        <td id="LC87" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\begin</span>{tabular}{ccc}</td>
+      </tr>
+      <tr>
+        <td id="L88" class="blob-num js-line-number" data-line-number="88"></td>
+        <td id="LC88" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\toprule</span></td>
+      </tr>
+      <tr>
+        <td id="L89" class="blob-num js-line-number" data-line-number="89"></td>
+        <td id="LC89" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\textbf</span>{Distribution}<span class="pl-pds">$</span></span> &amp; <span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\mu</span><span class="pl-pds">$</span></span> &amp; <span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\sigma</span><span class="pl-pds">$</span></span> <span class="pl-c1">\\</span> </td>
+      </tr>
+      <tr>
+        <td id="L90" class="blob-num js-line-number" data-line-number="90"></td>
+        <td id="LC90" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\bottomrule</span></td>
+      </tr>
+      <tr>
+        <td id="L91" class="blob-num js-line-number" data-line-number="91"></td>
+        <td id="LC91" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>G<span class="pl-pds">$</span></span>~(Model) &amp; 0.133 &amp; 0.0198 <span class="pl-c1">\\</span></td>
+      </tr>
+      <tr>
+        <td id="L92" class="blob-num js-line-number" data-line-number="92"></td>
+        <td id="LC92" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\tilde</span>{G}<span class="pl-pds">$</span></span>~(Gradient-based) &amp; 0.133 &amp; 0.0196 <span class="pl-c1">\\</span></td>
+      </tr>
+      <tr>
+        <td id="L93" class="blob-num js-line-number" data-line-number="93"></td>
+        <td id="LC93" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\tilde</span>{G}<span class="pl-pds">$</span></span>~(Gradient-free) &amp; 0.133 &amp; 0.0167 <span class="pl-c1">\\</span></td>
+      </tr>
+      <tr>
+        <td id="L94" class="blob-num js-line-number" data-line-number="94"></td>
+        <td id="LC94" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\bottomrule</span></td>
+      </tr>
+      <tr>
+        <td id="L95" class="blob-num js-line-number" data-line-number="95"></td>
+        <td id="LC95" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\end</span>{tabular}</td>
+      </tr>
+      <tr>
+        <td id="L96" class="blob-num js-line-number" data-line-number="96"></td>
+        <td id="LC96" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\end</span>{minipage}</td>
+      </tr>
+      <tr>
+        <td id="L97" class="blob-num js-line-number" data-line-number="97"></td>
+        <td id="LC97" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\hspace</span>{5mm}</td>
+      </tr>
+      <tr>
+        <td id="L98" class="blob-num js-line-number" data-line-number="98"></td>
+        <td id="LC98" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\begin</span>{minipage}[htbp]{.25<span class="pl-c1">\linewidth</span>}</td>
+      </tr>
+      <tr>
+        <td id="L99" class="blob-num js-line-number" data-line-number="99"></td>
+        <td id="LC99" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\vspace</span>{0pt}</td>
+      </tr>
+      <tr>
+        <td id="L100" class="blob-num js-line-number" data-line-number="100"></td>
+        <td id="LC100" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>\centering</span></td>
+      </tr>
+      <tr>
+        <td id="L101" class="blob-num js-line-number" data-line-number="101"></td>
+        <td id="LC101" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\includegraphics</span>[width=3.0in]{./Figures/pdf_comp_id_1e4}</td>
+      </tr>
+      <tr>
+        <td id="L102" class="blob-num js-line-number" data-line-number="102"></td>
+        <td id="LC102" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\end</span>{minipage}<span class="pl-c"><span class="pl-c">%</span></span></td>
+      </tr>
+      <tr>
+        <td id="L103" class="blob-num js-line-number" data-line-number="103"></td>
+        <td id="LC103" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\end</span>{center} </td>
+      </tr>
+      <tr>
+        <td id="L104" class="blob-num js-line-number" data-line-number="104"></td>
+        <td id="LC104" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\caption</span>{Left: Table providing the mean (<span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\mu</span><span class="pl-pds">$</span></span>) and the standard deviation (<span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\sigma</span><span class="pl-pds">$</span></span>) corresponding to</td>
+      </tr>
+      <tr>
+        <td id="L105" class="blob-num js-line-number" data-line-number="105"></td>
+        <td id="LC105" class="blob-code blob-code-inner js-file-line">the three PDFs using evaluations</td>
+      </tr>
+      <tr>
+        <td id="L106" class="blob-num js-line-number" data-line-number="106"></td>
+        <td id="LC106" class="blob-code blob-code-inner js-file-line">at 10<span class="pl-s"><span class="pl-pds">$</span>^<span class="pl-c1">4</span><span class="pl-pds">$</span></span> samples in the cross-validation set.
+Right: A comparison of the PDFs of ignition delay, obtained using model </td>
+      </tr>
+      <tr>
+        <td id="L107" class="blob-num js-line-number" data-line-number="107"></td>
+        <td id="LC107" class="blob-code blob-code-inner js-file-line">evaluations (solid line) and 1-dimensional surrogates using the gradient-free approach (dashed line) and the gradient-based</td>
+      </tr>
+      <tr>
+        <td id="L108" class="blob-num js-line-number" data-line-number="108"></td>
+        <td id="LC108" class="blob-code blob-code-inner js-file-line">approach (dashed-dotted line).
+The same set of 10<span class="pl-s"><span class="pl-pds">$</span>^<span class="pl-c1">4</span><span class="pl-pds">$</span></span> samples in the cross-validation set were used in each case.}</td>
+      </tr>
+      <tr>
+        <td id="L109" class="blob-num js-line-number" data-line-number="109"></td>
+        <td id="LC109" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\label</span>{fig:pdf_33D}</td>
+      </tr>
+      <tr>
+        <td id="L110" class="blob-num js-line-number" data-line-number="110"></td>
+        <td id="LC110" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\end</span>{figure}</td>
+      </tr>
+      <tr>
+        <td id="L111" class="blob-num js-line-number" data-line-number="111"></td>
+        <td id="LC111" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span></span></td>
+      </tr>
+      <tr>
+        <td id="L112" class="blob-num js-line-number" data-line-number="112"></td>
+        <td id="LC112" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span></span></td>
+      </tr>
+      <tr>
+        <td id="L113" class="blob-num js-line-number" data-line-number="113"></td>
+        <td id="LC113" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>\begin{figure}[htbp]</span></td>
+      </tr>
+      <tr>
+        <td id="L114" class="blob-num js-line-number" data-line-number="114"></td>
+        <td id="LC114" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span> \begin{center}</span></td>
+      </tr>
+      <tr>
+        <td id="L115" class="blob-num js-line-number" data-line-number="115"></td>
+        <td id="LC115" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>  \includegraphics[width=0.45\textwidth]{./Figures/pdf_comp_id_1e4}</span></td>
+      </tr>
+      <tr>
+        <td id="L116" class="blob-num js-line-number" data-line-number="116"></td>
+        <td id="LC116" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>\caption{A comparison of the PDFs of ignition delay, obtained using model evaluations (solid line) and</span></td>
+      </tr>
+      <tr>
+        <td id="L117" class="blob-num js-line-number" data-line-number="117"></td>
+        <td id="LC117" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>1-dimensional surrogates using the gradient-free approach (dashed line) and the gradient-based approach</span></td>
+      </tr>
+      <tr>
+        <td id="L118" class="blob-num js-line-number" data-line-number="118"></td>
+        <td id="LC118" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>(dashed-dotted line).
+The same set of 10$^4$ samples in the cross-validation set were used in each case.}</span></td>
+      </tr>
+      <tr>
+        <td id="L119" class="blob-num js-line-number" data-line-number="119"></td>
+        <td id="LC119" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>\label{fig:pdf_33D}</span></td>
+      </tr>
+      <tr>
+        <td id="L120" class="blob-num js-line-number" data-line-number="120"></td>
+        <td id="LC120" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>\end{center}</span></td>
+      </tr>
+      <tr>
+        <td id="L121" class="blob-num js-line-number" data-line-number="121"></td>
+        <td id="LC121" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>\end{figure}</span></td>
+      </tr>
+      <tr>
+        <td id="L122" class="blob-num js-line-number" data-line-number="122"></td>
+        <td id="LC122" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span></span></td>
+      </tr>
+      <tr>
+        <td id="L123" class="blob-num js-line-number" data-line-number="123"></td>
+        <td id="LC123" class="blob-code blob-code-inner js-file-line">The PDFs based on the model evaluations and the gradient-based approach are nearly identical, confirming</td>
+      </tr>
+      <tr>
+        <td id="L124" class="blob-num js-line-number" data-line-number="124"></td>
+        <td id="LC124" class="blob-code blob-code-inner js-file-line">the existence of a 1-dimensional active subspace as well as the accuracy of the gradient-based approach. </td>
+      </tr>
+      <tr>
+        <td id="L125" class="blob-num js-line-number" data-line-number="125"></td>
+        <td id="LC125" class="blob-code blob-code-inner js-file-line">Whereas, the PDF based on the gradient-free approach is observed to show consistency with the true PDF in the modal</td>
+      </tr>
+      <tr>
+        <td id="L126" class="blob-num js-line-number" data-line-number="126"></td>
+        <td id="LC126" class="blob-code blob-code-inner js-file-line">estimate; it underestimates the spread in the ignition delay. Specific values of the mean and the standard deviation</td>
+      </tr>
+      <tr>
+        <td id="L127" class="blob-num js-line-number" data-line-number="127"></td>
+        <td id="LC127" class="blob-code blob-code-inner js-file-line">in each case are provided in the table in Figure~<span class="pl-c1">\ref</span>{fig:pdf_33D}.
+Hence, the three PDFs are found to be</td>
+      </tr>
+      <tr>
+        <td id="L128" class="blob-num js-line-number" data-line-number="128"></td>
+        <td id="LC128" class="blob-code blob-code-inner js-file-line">consistent in their prediction of mean value of the ignition delay.  The standard deviation estimates based on model</td>
+      </tr>
+      <tr>
+        <td id="L129" class="blob-num js-line-number" data-line-number="129"></td>
+        <td id="LC129" class="blob-code blob-code-inner js-file-line">evaluations and the gradient-based approach are found to be in close agreement, whereas, it  is accurately estimated</td>
+      </tr>
+      <tr>
+        <td id="L130" class="blob-num js-line-number" data-line-number="130"></td>
+        <td id="LC130" class="blob-code blob-code-inner js-file-line">by the gradient-free approach only upto the first significant figure. </td>
+      </tr>
+      <tr>
+        <td id="L131" class="blob-num js-line-number" data-line-number="131"></td>
+        <td id="LC131" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L132" class="blob-num js-line-number" data-line-number="132"></td>
+        <td id="LC132" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\subsection</span>{GSA consistency check}</td>
+      </tr>
+      <tr>
+        <td id="L133" class="blob-num js-line-number" data-line-number="133"></td>
+        <td id="LC133" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L134" class="blob-num js-line-number" data-line-number="134"></td>
+        <td id="LC134" class="blob-code blob-code-inner js-file-line">The normalized activity scores (<span class="pl-s"><span class="pl-pds">$</span><span class="pl-c1">\tilde</span>{<span class="pl-c1">\nu</span>}_{i,r}<span class="pl-pds">$</span></span>) based on the dominant eigenspace using the two</td>
+      </tr>
+      <tr>
+        <td id="L135" class="blob-num js-line-number" data-line-number="135"></td>
+        <td id="LC135" class="blob-code blob-code-inner js-file-line">approaches (gradient-based and gradient-free) are compared with the total Sobol&#39; indices</td>
+      </tr>
+      <tr>
+        <td id="L136" class="blob-num js-line-number" data-line-number="136"></td>
+        <td id="LC136" class="blob-code blob-code-inner js-file-line">in Figure~<span class="pl-c1">\ref</span>{fig:as_33D}. Note that the Sobol&#39; indices were computed using the verified</td>
+      </tr>
+      <tr>
+        <td id="L137" class="blob-num js-line-number" data-line-number="137"></td>
+        <td id="LC137" class="blob-code blob-code-inner js-file-line">1-dimensional surrogate in the active subspace for the gradient-based approach. </td>
+      </tr>
+      <tr>
+        <td id="L138" class="blob-num js-line-number" data-line-number="138"></td>
+        <td id="LC138" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span></span></td>
+      </tr>
+      <tr>
+        <td id="L139" class="blob-num js-line-number" data-line-number="139"></td>
+        <td id="LC139" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\begin</span>{figure}[htbp]</td>
+      </tr>
+      <tr>
+        <td id="L140" class="blob-num js-line-number" data-line-number="140"></td>
+        <td id="LC140" class="blob-code blob-code-inner js-file-line"> <span class="pl-c1">\begin</span>{center}</td>
+      </tr>
+      <tr>
+        <td id="L141" class="blob-num js-line-number" data-line-number="141"></td>
+        <td id="LC141" class="blob-code blob-code-inner js-file-line">  <span class="pl-c1">\includegraphics</span>[width=0.8<span class="pl-c1">\textwidth</span>]{./Figures/as_33D_new}</td>
+      </tr>
+      <tr>
+        <td id="L142" class="blob-num js-line-number" data-line-number="142"></td>
+        <td id="LC142" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\caption</span>{A bar-graph illustrating individual activity scores for the uncertain <span class="pl-s"><span class="pl-pds">$</span>A_i<span class="pl-pds">$</span></span>&#39; (left) and <span class="pl-s"><span class="pl-pds">$</span>E_{a,i}<span class="pl-pds">$</span></span>&#39;s (right).}</td>
+      </tr>
+      <tr>
+        <td id="L143" class="blob-num js-line-number" data-line-number="143"></td>
+        <td id="LC143" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\label</span>{fig:as_33D}</td>
+      </tr>
+      <tr>
+        <td id="L144" class="blob-num js-line-number" data-line-number="144"></td>
+        <td id="LC144" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\end</span>{center}</td>
+      </tr>
+      <tr>
+        <td id="L145" class="blob-num js-line-number" data-line-number="145"></td>
+        <td id="LC145" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\end</span>{figure}</td>
+      </tr>
+      <tr>
+        <td id="L146" class="blob-num js-line-number" data-line-number="146"></td>
+        <td id="LC146" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span></span></td>
+      </tr>
+      <tr>
+        <td id="L147" class="blob-num js-line-number" data-line-number="147"></td>
+        <td id="LC147" class="blob-code blob-code-inner js-file-line">Several useful inferences can be drawn from the above plot. Firstly, the total Sobol indices are found to be</td>
+      </tr>
+      <tr>
+        <td id="L148" class="blob-num js-line-number" data-line-number="148"></td>
+        <td id="LC148" class="blob-code blob-code-inner js-file-line">consistent</td>
+      </tr>
+      <tr>
+        <td id="L149" class="blob-num js-line-number" data-line-number="149"></td>
+        <td id="LC149" class="blob-code blob-code-inner js-file-line">with the normalized activity scores evaluated using the gradient-based approach. Whereas, the gradient-free</td>
+      </tr>
+      <tr>
+        <td id="L150" class="blob-num js-line-number" data-line-number="150"></td>
+        <td id="LC150" class="blob-code blob-code-inner js-file-line">approach yields consistent results for the <span class="pl-s"><span class="pl-pds">$</span>A_i<span class="pl-pds">$</span></span>&#39;s, it does not capture the sensitivity associated with</td>
+      </tr>
+      <tr>
+        <td id="L151" class="blob-num js-line-number" data-line-number="151"></td>
+        <td id="LC151" class="blob-code blob-code-inner js-file-line"><span class="pl-s"><span class="pl-pds">$</span>E_{a,15}<span class="pl-pds">$</span></span>.
+While this observation bolsters our confidence in the gradient-based approach, it clearly</td>
+      </tr>
+      <tr>
+        <td id="L152" class="blob-num js-line-number" data-line-number="152"></td>
+        <td id="LC152" class="blob-code blob-code-inner js-file-line">underscores the shortcoming of the gradient-free approach. Numerical errors are incurred when approximating</td>
+      </tr>
+      <tr>
+        <td id="L153" class="blob-num js-line-number" data-line-number="153"></td>
+        <td id="LC153" class="blob-code blob-code-inner js-file-line">the model gradients using the local linear approximation in this case. </td>
+      </tr>
+      <tr>
+        <td id="L154" class="blob-num js-line-number" data-line-number="154"></td>
+        <td id="LC154" class="blob-code blob-code-inner js-file-line">Secondly, our findings indicate that   </td>
+      </tr>
+      <tr>
+        <td id="L155" class="blob-num js-line-number" data-line-number="155"></td>
+        <td id="LC155" class="blob-code blob-code-inner js-file-line">the ignition delay is predominantly sensitive towards <span class="pl-s"><span class="pl-pds">$</span>A_<span class="pl-c1">1</span><span class="pl-pds">$</span></span>, <span class="pl-s"><span class="pl-pds">$</span>A_<span class="pl-c1">9</span><span class="pl-pds">$</span></span>, <span class="pl-s"><span class="pl-pds">$</span>E_{a,1}<span class="pl-pds">$</span></span>, and <span class="pl-s"><span class="pl-pds">$</span>E_{a,15}<span class="pl-pds">$</span></span> and </td>
+      </tr>
+      <tr>
+        <td id="L156" class="blob-num js-line-number" data-line-number="156"></td>
+        <td id="LC156" class="blob-code blob-code-inner js-file-line">moderately sensitive towards <span class="pl-s"><span class="pl-pds">$</span>A_{15}<span class="pl-pds">$</span></span> and <span class="pl-s"><span class="pl-pds">$</span>A_{17}<span class="pl-pds">$</span></span>.
+Sensitivity towards the remaining uncertain inputs is</td>
+      </tr>
+      <tr>
+        <td id="L157" class="blob-num js-line-number" data-line-number="157"></td>
+        <td id="LC157" class="blob-code blob-code-inner js-file-line">found to be low or negligible. </td>
+      </tr>
+      <tr>
+        <td id="L158" class="blob-num js-line-number" data-line-number="158"></td>
+        <td id="LC158" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>Note that this observation can also be exploited for reducing the </span></td>
+      </tr>
+      <tr>
+        <td id="L159" class="blob-num js-line-number" data-line-number="159"></td>
+        <td id="LC159" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>dimensionality of the problem from 33 inputs to 4-6 inputs depending upon the required level of accuracy.</span></td>
+      </tr>
+      <tr>
+        <td id="L160" class="blob-num js-line-number" data-line-number="160"></td>
+        <td id="LC160" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>The sensitivity-driven dimension reduction for uncertainty quantification has been</span></td>
+      </tr>
+      <tr>
+        <td id="L161" class="blob-num js-line-number" data-line-number="161"></td>
+        <td id="LC161" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>discussed in our earlier effort~\cite{Vohra:2018}.
+In this work, however, we focus on dimension reduction using the active </span></td>
+      </tr>
+      <tr>
+        <td id="L162" class="blob-num js-line-number" data-line-number="162"></td>
+        <td id="LC162" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>subspace approach which is shown to yield even greater scope for dimension reduction i.e.
+from 33 to 1.
+Furthermore, we have</span></td>
+      </tr>
+      <tr>
+        <td id="L163" class="blob-num js-line-number" data-line-number="163"></td>
+        <td id="LC163" class="blob-code blob-code-inner js-file-line"><span class="pl-c"><span class="pl-c">%</span>shown that the dominant eigenspace can be used to approximate global sensitivity measures with no additional effort.</span></td>
+      </tr>
+      <tr>
+        <td id="L164" class="blob-num js-line-number" data-line-number="164"></td>
+        <td id="LC164" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L165" class="blob-num js-line-number" data-line-number="165"></td>
+        <td id="LC165" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L166" class="blob-num js-line-number" data-line-number="166"></td>
+        <td id="LC166" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L167" class="blob-num js-line-number" data-line-number="167"></td>
+        <td id="LC167" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L168" class="blob-num js-line-number" data-line-number="168"></td>
+        <td id="LC168" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L169" class="blob-num js-line-number" data-line-number="169"></td>
+        <td id="LC169" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L170" class="blob-num js-line-number" data-line-number="170"></td>
+        <td id="LC170" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L171" class="blob-num js-line-number" data-line-number="171"></td>
+        <td id="LC171" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L172" class="blob-num js-line-number" data-line-number="172"></td>
+        <td id="LC172" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L173" class="blob-num js-line-number" data-line-number="173"></td>
+        <td id="LC173" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L174" class="blob-num js-line-number" data-line-number="174"></td>
+        <td id="LC174" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L175" class="blob-num js-line-number" data-line-number="175"></td>
+        <td id="LC175" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L176" class="blob-num js-line-number" data-line-number="176"></td>
+        <td id="LC176" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L177" class="blob-num js-line-number" data-line-number="177"></td>
+        <td id="LC177" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L178" class="blob-num js-line-number" data-line-number="178"></td>
+        <td id="LC178" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L179" class="blob-num js-line-number" data-line-number="179"></td>
+        <td id="LC179" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L180" class="blob-num js-line-number" data-line-number="180"></td>
+        <td id="LC180" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L181" class="blob-num js-line-number" data-line-number="181"></td>
+        <td id="LC181" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L182" class="blob-num js-line-number" data-line-number="182"></td>
+        <td id="LC182" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L183" class="blob-num js-line-number" data-line-number="183"></td>
+        <td id="LC183" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L184" class="blob-num js-line-number" data-line-number="184"></td>
+        <td id="LC184" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L185" class="blob-num js-line-number" data-line-number="185"></td>
+        <td id="LC185" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L186" class="blob-num js-line-number" data-line-number="186"></td>
+        <td id="LC186" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L187" class="blob-num js-line-number" data-line-number="187"></td>
+        <td id="LC187" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L188" class="blob-num js-line-number" data-line-number="188"></td>
+        <td id="LC188" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L189" class="blob-num js-line-number" data-line-number="189"></td>
+        <td id="LC189" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L190" class="blob-num js-line-number" data-line-number="190"></td>
+        <td id="LC190" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L191" class="blob-num js-line-number" data-line-number="191"></td>
+        <td id="LC191" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L192" class="blob-num js-line-number" data-line-number="192"></td>
+        <td id="LC192" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L193" class="blob-num js-line-number" data-line-number="193"></td>
+        <td id="LC193" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L194" class="blob-num js-line-number" data-line-number="194"></td>
+        <td id="LC194" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+      <tr>
+        <td id="L195" class="blob-num js-line-number" data-line-number="195"></td>
+        <td id="LC195" class="blob-code blob-code-inner js-file-line">
+</td>
+      </tr>
+</table>
+
+  <details class="details-reset details-overlay BlobToolbar position-absolute js-file-line-actions dropdown d-none" aria-hidden="true">
+    <summary class="btn-octicon ml-0 px-2 p-0 bg-white border border-gray-dark rounded-1" aria-label="Inline file action toolbar">
+      <svg class="octicon octicon-kebab-horizontal" viewBox="0 0 13 16" version="1.1" width="13" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M1.5 9a1.5 1.5 0 1 0 0-3 1.5 1.5 0 0 0 0 3zm5 0a1.5 1.5 0 1 0 0-3 1.5 1.5 0 0 0 0 3zM13 7.5a1.5 1.5 0 1 1-3 0 1.5 1.5 0 0 1 3 0z"/></svg>
+    </summary>
+    <details-menu>
+      <ul class="BlobToolbar-dropdown dropdown-menu dropdown-menu-se mt-2">
+        <li><clipboard-copy role="menuitem" class="dropdown-item" id="js-copy-lines" style="cursor:pointer;" data-original-text="Copy lines">Copy lines</clipboard-copy></li>
+        <li><clipboard-copy role="menuitem" class="dropdown-item" id="js-copy-permalink" style="cursor:pointer;" data-original-text="Copy permalink">Copy permalink</clipboard-copy></li>
+        <li><a class="dropdown-item js-update-url-with-hash" id="js-view-git-blame" role="menuitem" href="/vohramanav/ActiveSubspace_Kinetics/blame/de781a8c95a3c79c7d21d73d7c02fb74af163b10/Paper/app.tex">View git blame</a></li>
+          <li><a class="dropdown-item" id="js-new-issue" role="menuitem" href="/vohramanav/ActiveSubspace_Kinetics/issues/new">Open new issue</a></li>
+      </ul>
+    </details-menu>
+  </details>
+
+  </div>
+
+  </div>
+
+  <details class="details-reset details-overlay details-overlay-dark">
+    <summary data-hotkey="l" aria-label="Jump to line"></summary>
+    <details-dialog class="Box Box--overlay d-flex flex-column anim-fade-in fast linejump" aria-label="Jump to line">
+      <!-- '"` --><!-- </textarea></xmp> --></option></form><form class="js-jump-to-line-form Box-body d-flex" action="" accept-charset="UTF-8" method="get"><input name="utf8" type="hidden" value="&#x2713;" />
+        <input class="form-control flex-auto mr-3 linejump-input js-jump-to-line-field" type="text" placeholder="Jump to line&hellip;" aria-label="Jump to line" autofocus>
+        <button type="submit" class="btn" data-close-dialog>Go</button>
+</form>    </details-dialog>
+  </details>
+
+  </div>
+  <div class="modal-backdrop js-touch-events"></div>
+</div>
+
+    </div>
+  </div>
+
+  </div>
+
+<div class="footer container-lg px-3" role="contentinfo">
+  <div class="position-relative d-flex flex-justify-between pt-6 pb-2 mt-6 f6 text-gray border-top border-gray-light ">
+    <ul class="list-style-none d-flex flex-wrap ">
+      <li class="mr-3">&copy; 2018 <span title="0.29776s from unicorn-5d658c9fd8-qzbnt">GitHub</span>, Inc.</li>
+        <li class="mr-3"><a data-ga-click="Footer, go to terms, text:terms" href="https://github.com/site/terms">Terms</a></li>
+        <li class="mr-3"><a data-ga-click="Footer, go to privacy, text:privacy" href="https://github.com/site/privacy">Privacy</a></li>
+        <li class="mr-3"><a href="https://help.github.com/articles/github-security/" data-ga-click="Footer, go to security, text:security">Security</a></li>
+        <li class="mr-3"><a href="https://status.github.com/" data-ga-click="Footer, go to status, text:status">Status</a></li>
+        <li><a data-ga-click="Footer, go to help, text:help" href="https://help.github.com">Help</a></li>
+    </ul>
+
+    <a aria-label="Homepage" title="GitHub" class="footer-octicon" href="https://github.com">
+      <svg height="24" class="octicon octicon-mark-github" viewBox="0 0 16 16" version="1.1" width="24" aria-hidden="true"><path fill-rule="evenodd" d="M8 0C3.58 0 0 3.58 0 8c0 3.54 2.29 6.53 5.47 7.59.4.07.55-.17.55-.38 0-.19-.01-.82-.01-1.49-2.01.37-2.53-.49-2.69-.94-.09-.23-.48-.94-.82-1.13-.28-.15-.68-.52-.01-.53.63-.01 1.08.58 1.23.82.72 1.21 1.87.87 2.33.66.07-.52.28-.87.51-1.07-1.78-.2-3.64-.89-3.64-3.95 0-.87.31-1.59.82-2.15-.08-.2-.36-1.02.08-2.12 0 0 .67-.21 2.2.82.64-.18 1.32-.27 2-.27.68 0 1.36.09 2 .27 1.53-1.04 2.2-.82 2.2-.82.44 1.1.16 1.92.08 2.12.51.56.82 1.27.82 2.15 0 3.07-1.87 3.75-3.65 3.95.29.25.54.73.54 1.48 0 1.07-.01 1.93-.01 2.2 0 .21.15.46.55.38A8.013 8.013 0 0 0 16 8c0-4.42-3.58-8-8-8z"/></svg>
+</a>
+   <ul class="list-style-none d-flex flex-wrap ">
+        <li class="mr-3"><a data-ga-click="Footer, go to contact, text:contact" href="https://github.com/contact">Contact GitHub</a></li>
+        <li class="mr-3"><a href="https://github.com/pricing" data-ga-click="Footer, go to Pricing, text:Pricing">Pricing</a></li>
+      <li class="mr-3"><a href="https://developer.github.com" data-ga-click="Footer, go to api, text:api">API</a></li>
+      <li class="mr-3"><a href="https://training.github.com" data-ga-click="Footer, go to training, text:training">Training</a></li>
+        <li class="mr-3"><a href="https://blog.github.com" data-ga-click="Footer, go to blog, text:blog">Blog</a></li>
+        <li><a data-ga-click="Footer, go to about, text:about" href="https://github.com/about">About</a></li>
+
+    </ul>
+  </div>
+  <div class="d-flex flex-justify-center pb-6">
+    <span class="f6 text-gray-light"></span>
+  </div>
+</div>
+
+  <div id="ajax-error-message" class="ajax-error-message flash flash-error">
+    <svg class="octicon octicon-alert" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.893 1.5c-.183-.31-.52-.5-.887-.5s-.703.19-.886.5L.138 13.499a.98.98 0 0 0 0 1.001c.193.31.53.501.886.501h13.964c.367 0 .704-.19.877-.5a1.03 1.03 0 0 0 .01-1.002L8.893 1.5zm.133 11.497H6.987v-2.003h2.039v2.003zm0-3.004H6.987V5.987h2.039v4.006z"/></svg>
+    <button type="button" class="flash-close js-ajax-error-dismiss" aria-label="Dismiss error">
+      <svg class="octicon octicon-x" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
+    </button>
+    You can’t perform that action at this time.
+  </div>
+
+    <script crossorigin="anonymous" integrity="sha512-2VdGgXQE8W5ONZ4OsrbEo/noennUZaqXkevD9R8juTHiCsjT0HFTTF6MoBfySUc8G+eFqUcDgd7v+CAu8Gjxlg==" type="application/javascript" src="https://assets-cdn.github.com/assets/compat-f849c975b0ffaa01d6ca305e48417d08.js"></script>
+    <script crossorigin="anonymous" integrity="sha512-xY0/FVA4OIId4Rg/UZwnuRt3NDzYfohwWB8mPrflYnyu+RSIdK8VxsCY3VkctCpeCW5BSB331jAUlVEbupUlGQ==" type="application/javascript" src="https://assets-cdn.github.com/assets/frameworks-caa0bdcb0cd0270ade93e4f0358e158e.js"></script>
+
+    <script crossorigin="anonymous" async="async" integrity="sha512-/gomoh+vz4Wtp3+45P6w1H96O3k/a09vbwD6g3etmqocBu12fflwXz0oFmWYELM1Kh85wnXuBytcmGv4QSca6w==" type="application/javascript" src="https://assets-cdn.github.com/assets/github-69f027cf810a992224f91b6f505ea246.js"></script>
+
+  <div class="js-stale-session-flash stale-session-flash flash flash-warn flash-banner d-none">
+    <svg class="octicon octicon-alert" viewBox="0 0 16 16" version="1.1" width="16" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M8.893 1.5c-.183-.31-.52-.5-.887-.5s-.703.19-.886.5L.138 13.499a.98.98 0 0 0 0 1.001c.193.31.53.501.886.501h13.964c.367 0 .704-.19.877-.5a1.03 1.03 0 0 0 .01-1.002L8.893 1.5zm.133 11.497H6.987v-2.003h2.039v2.003zm0-3.004H6.987V5.987h2.039v4.006z"/></svg>
+    <span class="signed-in-tab-flash">You signed in with another tab or window. <a href="">Reload</a> to refresh your session.</span>
+    <span class="signed-out-tab-flash">You signed out in another tab or window. <a href="">Reload</a> to refresh your session.</span>
+  </div>
+  <div class="facebox" id="facebox" style="display:none;">
+  <div class="facebox-popup">
+    <div class="facebox-content" role="dialog" aria-labelledby="facebox-header" aria-describedby="facebox-description">
+    </div>
+    <button type="button" class="facebox-close js-facebox-close" aria-label="Close modal">
+      <svg class="octicon octicon-x" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
+    </button>
+  </div>
+</div>
+
+  <template id="site-details-dialog">
+  <details class="details-reset details-overlay details-overlay-dark lh-default text-gray-dark" open>
+    <summary aria-haspopup="dialog" aria-label="Close dialog"></summary>
+    <details-dialog class="Box Box--overlay d-flex flex-column anim-fade-in fast">
+      <button class="m-3 btn-octicon position-absolute right-0 top-0" type="button" aria-label="Close dialog" data-close-dialog>
+        <svg class="octicon octicon-x" viewBox="0 0 12 16" version="1.1" width="12" height="16" aria-hidden="true"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48L7.48 8z"/></svg>
+      </button>
+      <div class="octocat-spinner my-6 js-details-dialog-spinner"></div>
+    </details-dialog>
+  </details>
+</template>
+
+  <div class="Popover js-hovercard-content position-absolute" style="display: none; outline: none;" tabindex="0">
+  <div class="Popover-message Popover-message--bottom-left Popover-message--large Box box-shadow-large" style="width:360px;">
+  </div>
+</div>
+
+<div id="hovercard-aria-description" class="sr-only">
+  Press h to open a hovercard with more details.
+</div>
+
+  </body>
+</html>

### 2018-08-28 Paper/abstract.tex
..

\section*{Abstract}
Time evolution of species' concentration in a chemical system is tightly coupled
with the underlying mechanism and the governing rate laws for individual reactions.
Oftentimes, the rate-controlling parameters cannot be specified precisely 
due to sparse calibration data and lack of knowledge about them, regarded as
the \textit{epistemic} uncertainty. Computer simulations for such systems must
therefore account for the presence of the epistemic uncertainty in these parameters
when making predictions. However, complex mechanisms typically involve a large number of
reactions and consequently, a large set of uncertain rate parameters. Propagation of
the uncertainty from parameters to predictions can be remarkably challenging for
such high-dimensional problems. In this study, we focus on enabling an efficient
approach for uncertainty propagation from inputs to the model output,
through dimension reduction by computing the
so-called
\textit{active subspace}. The active subspace
 predominantly captures the variability in the quantity of
interest (QoI) in terms of fewer independent variables. 
It is computed using an iterative, gradient-based
approach and a computationally less intensive gradient-free approach for the 
H$_2$/O$_2$ mechanism involving 19 reactions. An initial investigation considers
only the 19 pre-exponents of the Arrhenius rate laws as uncertain. In this case, both
approaches yield consistent results wherein the active subspace is found to be 
1-dimensional. The two approaches are further implemented to a high-dimensional
problem involving 33 uncertain rate-parameters: pre-exponent and the activation
energy. Once again, a 1-dimensional active subspace is obtained, indicating
enormous scope for dimension reduction in such applications. However, the
gradient-free approach though reasonably captured the mean, mode, and the uncertainty
in the QoI; failed to capture the sensitivity towards the most
important parameter according to the gradient-based analysis for the high-dimensional
application.

### 2018-09-05 Paper/abstract.tex
..

@@
 with the underlying mechanism and the governing rate laws for individual reactions.
-Oftentimes, the rate-controlling parameters cannot be specified precisely 
-due to sparse calibration data and lack of knowledge about them, regarded as
-the \textit{epistemic} uncertainty. Computer simulations for such systems must
-therefore account for the presence of the epistemic uncertainty in these parameters
+Often, the rate-controlling parameters cannot be specified precisely 
+due to \textit{epistemic} uncertainty, i.e., lack of knowledge and 
+sparse calibration data. Computer simulations of such systems must
+therefore account for the epistemic uncertainty regarding these parameters
 when making predictions. However, complex mechanisms typically involve a large number of
@@
 the uncertainty from parameters to predictions can be remarkably challenging for
-such high-dimensional problems. In this study, we focus on enabling an efficient
+such high-dimensional problems. In this study, we focus on developing an efficient
 approach for uncertainty propagation from inputs to the model output,
@@
 interest (QoI) in terms of fewer independent variables. 
-It is computed using an iterative, gradient-based
-approach and a computationally less intensive gradient-free approach for the 
-H$_2$/O$_2$ mechanism involving 19 reactions. An initial investigation considers
-only the 19 pre-exponents of the Arrhenius rate laws as uncertain. In this case, both
-approaches yield consistent results wherein the active subspace is found to be 
-1-dimensional. The two approaches are further implemented to a high-dimensional
-problem involving 33 uncertain rate-parameters: pre-exponent and the activation
-energy. Once again, a 1-dimensional active subspace is obtained, indicating
-enormous scope for dimension reduction in such applications. However, the
-gradient-free approach though reasonably captured the mean, mode, and the uncertainty
-in the QoI; failed to capture the sensitivity towards the most
-important parameter according to the gradient-based analysis for the high-dimensional
-application.
+However, its computation requires gradient estimation of
+the model output with respect to its inputs which can be
+remarkably challenging. In this study, we compute the active subspace
+for the H$_2$/O$_2$ mechanism that involves 19 chemical reactions,
+using two iterative strategies. The first strategy uses finite difference 
+to estimate the gradient, whereas, the second strategy approximates the
+gradient using a regression-based linear approximation of the model output.
+The finite difference approach requires additional model evaluations  and 
+is therefore computationally expensive. Both strategies yield consistent
+results for the case when only the pre-exponents in the Arrhenius rate laws
+for the 19 reactions are considered as uncertain. However, implementation 
+to a higher-dimensional problem involving 33 uncertain rate parameters:
+pre-exponents and activation energies reveals the trade-off between 
+accuracy and effort. Specifically, using the second strategy,
+the mean, mode, and the uncertainty
+in the QoI is reasonably approximated; the sensitivity towards the most
+important rate parameter is not captured. Hence, both approaches are
+useful depending upon the complexity of the mechanism and the QoI.

### 2018-09-05 Paper/assa.tex
..

@@
 As mentioned in the introduction, an active subspace is a low-dimensional subspace
-that constitutes important directions in a model's input
+that consists of important directions in a model's input
 parameter space~\cite{Constantine:2015}. The effective variability in a model output $f$
@@
 along these directions. 
-The directions spanning the active subspace are the eigenvectors of the positive
+The directions constituting the active subspace are the eigenvectors of the positive
 semidefinite matrix 
@@
  %
-Clearly, the computational effort associated with constructing the matrix,
+Clearly, the computational effort associated with constructing the matrix
 $\hat{\mat{C}}$ scales with the number of samples, $N$. Hence, an iterative
@@
 
-In~\cite{Vohra:2018} the following normalized screening metric
+In~\cite{Vohra:2018} the following screening metric
 \[

### 2018-09-05 Paper/intro.tex
..

@@
 applications, these parameters are either specified using expert knowledge or
-estimated based on a regression-fit to a set of sparse and noisy
+estimated based on a regression fit to a set of sparse and noisy
 data~\cite{Burnham:1987, Burnham:1988, Vohra:2011, Sarathy:2012}.
@@
 %therefore the simulation is typically ignored. 
-Consequently, the inclusion of
-uncertainty due to corrupted calibration data, parametric uncertainty,
-and the discrepancy between predictions and experiments
-due to model inadequacy has been subject of increased research efforts in recent
-years~\cite{Vohra:2014, Vohra:2017, Morrison:2018, Hantouche:2018}. \alennote{Need more references,
-acknowledging research by others.}
+Intensive research efforts in recent years within the field of uncertainty quantification (UQ)
+address the quantification and propagation of uncertainty in system models due to 
+inadequate data, parametric uncertainty, and model errors~\cite{Vohra:2014, 
+Vohra:2017, Morrison:2018, Hantouche:2018, Nannapaneni:2016, Sankararaman:2015}.
 
@@
 \emph{important} directions in the input space that predominantly capture the
-variability in the model output. These directions, which span the so called
+variability in the model output. These directions, which constitute the so called
 \emph{active subspace}~\cite{Constantine:2015}, are the dominant eigenvectors
@@
 gradients can however be computationally challenging in high-dimensional settings. 
-Hence, in this work, we further explore the suitability of a gradient-free
+Hence, in this work, we also explore the suitability of a gradient-free
 approach that involves a regression-based local linear approximation of
@@
 An alternate approach to dimension reduction involves computing the global
-sensitivity measures associated with uncertain inputs of a model. Depending
+sensitivity measures associated with the uncertain inputs of a model. Depending
 upon the estimates of the sensitivity measures, only the important inputs are
 varied for the purpose of uncertainty quantification (UQ). Sobol' indices are
-commonly used as global sensitivity measures~\cite{Sobol:2001}. However,
-obtaining converged estimates of the Sobol' indices typically requires tens of
-thousands of model runs and is therefore impractical in situations involving
-expensive model simulations and/or high-dimensional input parameter spaces.
-Recently, the derivative-based global sensitivity measures~(DGSMs) were
-developed to approximate upper bound on the Sobol' indices with much fewer
+commonly used as global sensitivity measures~\cite{Sobol:2001}. They are
+used to quantify the relative contributions of the uncertain inputs to the variance
+in the output, either individually, or in combination with other inputs. 
+Multiple efforts have focused on efficient computation of the Sobol' 
+indices~\cite{Sudret:2008,Plischke:2013,Tissot:2015,Li:2016} including the 
+derivative-based global sensitivity measures~(DGSMs), developed to
+compute approximate upper bounds for the Sobol' indices with much fewer
 computations~\cite{Sobol:2009, Lamboni:2013}. It was noted
@@
 important information about a model that go beyond dimension reduction. By 
-identifying parameters with significant impact on model uncertainty global 
-sensitivity analysis can guide model calibration and risk assessment.
+identifying parameters with significant impact on the model output, we can assess
+regimes of validity of the model formulation, and gain critical insight into the
+underlying physics in many cases.
 
@@
 discovery in a high-dimensional H$_2$/O$_2$ kinetics problem involving 33
-uncertain inputs. The implementation of the methodology presented in this work
+uncertain inputs: The methodology presented in this work
 resulted in a 1-dimensional active subspace, indicating immense potential
 for computational savings. The presented
-analysis can also guide practitioners in area of chemical kinetics on using the
-method of active subspaces as an efficient means for uncertainty propagation.  
+analysis can also guide practitioners in other problems of chemical kinetics on using the
+method of active subspaces to achieve efficiency in uncertainty propagation.  
 \item Comprehensive numerical investigation of the gradient-based and gradient-free approach: 
-we investigate the suitability of both approach
+We investigate the suitability of both approaches
 for computing active subspaces, surrogate modeling, and approximating global
@@
 global sensitivity measures: 
-By connecting the recent advances in theory of variance based and
-derivative based global sensitivity analysis to active subspaces, we provide a complete analysis
+By connecting the recent theoretical advances in variance-based and
+derivative-based global sensitivity analysis to active subspaces, we provide a complete analysis
 of the links between total Sobol' indices, DGSMs, and activity scores for a broad
@@
 consistent results. In section~\ref{sec:app}, the gradient-based approach is
-further implemented to compute the active subspace for the high-dimensional
-H$_2$/O$_2$ reaction kinetics problem. The resulting active subspace is found
-to be 1-dimensional thereby indicating enormous potential for computational
+further implemented to compute the active subspace for the higher-dimensional
+H$_2$/O$_2$ reaction kinetics problem. The resulting active subspace is once again
+found to be 1-dimensional, thereby indicating enormous potential for computational
 savings in this application. The active subspace is also computed using the

### 2018-09-05 Paper/app.tex
..

--- 2018-08-20 Paper/app.tex
@@
 were considered to be uncertain and uniformly distributed in the interval: 
-$[0.99E_{a,i}^\ast, 1.01E_{a,i}^\ast]$. The nominal values, $E_{a,i}^\ast$
+$[0.99E_{a,i}^\ast, 1.01E_{a,i}^\ast]$. Hence, the total number of uncertain rate parameters
+is 33. The nominal values, $E_{a,i}^\ast$
 corresponding to the different reaction rates are provided in~\cite{Yetter:1991}.
@@
 in the case of gradient-based approach, the SSP exhibits a linear variation that is reasonably
-captured by its linear fit, $\tilde{G}$. Whereas, the SSP based on the gradient-free approach
+captured by its linear fit, $\tilde{G}$, constructed using the sequence of steps outlined in Section~\ref{sub:ac}.
+ Whereas, the SSP based on the gradient-free approach
 is relatively more scattered. Consequently, the linear fit exhibits a discrepancy with the corresponding
@@
 %
-The relative norm, $\varepsilon_d$ was estimated to be 1.35$\times10^{-2}$ and 1.02$\times10^{-1}$
+Model evaluations ($G$) at 10$^{4}$ samples in the input domain were used to evaluate $\varepsilon_d$. 
+It was estimated to be 1.35$\times10^{-2}$ and 1.02$\times10^{-1}$
 using $\tilde{G}$ from the gradient-based approach and the gradient-free approach respectively. Hence,
 an order of magnitude increase in $\varepsilon_d$ was estimated in the case of gradient-free approach.
-Model evaluations ($G$) at 10$^{4}$ samples in the input domain were used for this purpose. 
 Hence, in the norm-sense, it can be said that the 1-dimensional surrogate in the active subspace, obtained
-using the gradient-based approach is remarkably accurate with a relative error of less than 1.5$\%$.
+using the gradient-based approach is remarkably accurate with a relative error of less than 1.5$\%$, whereas
+the gradient-free approach resulted in a relative error of about 10$\%$.
 
@@
 \begin{center}
-\begin{minipage}[htbp]{.25\linewidth}
-\vspace{0pt}
-%\centering
-\hspace{-25mm}
-\begin{tabular}{ccc}
-\toprule
-$\textbf{Distribution}$ & $\mu$ & $\sigma$ \\ 
-\bottomrule
-$G$~(Model) & 0.133 & 0.0198 \\
-$\tilde{G}$~(Gradient-based) & 0.133 & 0.0196 \\
-$\tilde{G}$~(Gradient-free) & 0.133 & 0.0167 \\
-\bottomrule
-\end{tabular}
-\end{minipage}
-\hspace{5mm}
-\begin{minipage}[htbp]{.25\linewidth}
-\vspace{0pt}
-%\centering
 \includegraphics[width=3.0in]{./Figures/pdf_comp_id_1e4}
-\end{minipage}%
 \end{center} 
-\caption{Left: Table providing the mean ($\mu$) and the standard deviation ($\sigma$) corresponding to
-the three PDFs using evaluations
-at 10$^4$ samples in the cross-validation set. Right: A comparison of the PDFs of ignition delay, obtained using model 
+\caption{A comparison of the PDFs of ignition delay, obtained using model 
 evaluations (solid line) and 1-dimensional surrogates using the gradient-free approach (dashed line) and the gradient-based
@@
 the existence of a 1-dimensional active subspace as well as the accuracy of the gradient-based approach. 
-Whereas, the PDF based on the gradient-free approach is observed to show consistency with the true PDF in the modal
+However, though the PDF based on the gradient-free approach is observed to show consistency with the true 
+PDF in the modal
 estimate; it underestimates the spread in the ignition delay. Specific values of the mean and the standard deviation
-in each case are provided in the table in Figure~\ref{fig:pdf_33D}. Hence, the three PDFs are found to be
+in each case are provided in Table~\ref{tab:stats}. 
+%
+\begin{table}[htbp]
+\begin{center}
+\begin{tabular}{ccc}
+\toprule
+$\textbf{Distribution}$ & $\mu$ & $\sigma$ \\ 
+\bottomrule
+$G$~(Model) & 0.133 & 0.0198 \\
+$\tilde{G}$~(Gradient-based) & 0.133 & 0.0196 \\
+$\tilde{G}$~(Gradient-free) & 0.133 & 0.0167 \\
+\bottomrule
+\end{tabular}
+\caption{The mean ($\mu$), and the standard deviation ($\sigma$), computed using the model ($G$), and
+the surrogate ($\tilde{G}$) based on the two approaches at 10$^4$ samples in the cross-validation
+set.}
+\label{tab:stats}
+\end{center}
+\end{table}
+%
+The three PDFs are found to be
 consistent in their prediction of mean value of the ignition delay.  The standard deviation estimates based on model
@@
 consistent
-with the normalized activity scores evaluated using the gradient-based approach. Whereas, the gradient-free
-approach yields consistent results for the $A_i$'s, it does not capture the sensitivity associated with
-$E_{a,15}$. While this observation bolsters our confidence in the gradient-based approach, it clearly
-underscores the shortcoming of the gradient-free approach. Numerical errors are incurred when approximating
-the model gradients using the local linear approximation in this case. 
+with the normalized activity scores evaluated using the gradient-based approach. The gradient-free
+approach yields consistent results for the $A_i$'s, but it does not capture the sensitivity associated with
+$E_{a,15}$. This is because numerical errors are incurred when approximating
+the model gradients using the regression-based approximation in the gradient-free case. 
 Secondly, our findings indicate that   

### 2018-09-05 Paper/conc.tex
..

@@
 the gradient-free approach failed to capture the sensitivity towards the most important parameter, $E_{a,15}$. 
-We attribute this discrepancy to two potential sources of numerical errors: (1) Local linear approximation of the
+We attribute this discrepancy to two potential sources of numerical errors: (1) Regression-based approximation of the
 gradient of the model output, and (2) Scatter in the corresponding SSP (see Figure~\ref{fig:hd}) leading to numerical
-error incurred by the linear regression-fit in this case. The PDF was observed to be consistent with the true PDF in its
+error incurred by the linear regression fit in this case. The PDF was observed to be consistent with the true PDF in its
 estimate for the mode and the mean value of the ignition delay. However, it was observed to underestimate the uncertainty 

### 2018-09-05 Paper/method.tex
..

@@
 estimate provided in~\cite{Yetter:1991}.
-Specifically, we employ two strategies: 
-A gradient-based strategy that involves computation of model
+Specifically, we employ two strategies --
+a gradient-based strategy that involves computation of model
 gradients in order to construct the matrix $\hat{\mat{C}}$
@@
 iterative manner to ensure computational efficiency by avoiding 
-extraneous model evaluations once a converged active subspace
+unnecessary model evaluations once a converged active subspace
 is achieved.
@@
 		\State Draw $n_k =  \lceil\beta n_{k-1}\rceil$  new random samples 
-                $\{\bm{\xi}_i\}_{i=1}^{n_k}$  $\beta\in[0,1]$  $k = n_{k-1}+1,\ldots,n_{k-1}+n_k$.
+                $\{\bm{\xi}_i\}_{i=1}^{n_k}$  $\beta\in[0,1]$
 
@@
 		 \hat{\mat{W}}^{(k)\top}$
+		\State Partition the eigenspace of $\hat{\mat{C}}$ as shown in Step 7
 		\State Compute $\delta \hat{\mat{W}}_{1,j}^{(k)}$ = 
@@
 For each sample in $M$, $p$ nearest neighbors in the $n_0$ set are identified. A least-squares fit to model
-evaluations at the $p$ samples is performed. The slope vector ($\bm{b}_i$) of the regression-fit is stored
+evaluations at the $p$ samples is performed. The slope vector ($\bm{b}_i$) of the regression fit is stored
 separately and the process is repeated for each sample in $M$. The symmetric and positive semi-definite matrix,
@@
 \renewcommand{\algorithmicensure}{\textbf{Output:}}
-  \caption{Algorithm for constructing the matrix, $\mat{C}$}
+  \caption{Algorithm for constructing the matrix, $\hat{\mat{C}} in~\eqref{eq:chat}$}
   \begin{algorithmic}[1]

### 2018-09-06 Paper/assa.tex
addressing comments on secs 2, 3 ...

@@
 along these directions. 
-The directions constituting the active subspace are the eigenvectors of the positive
+The directions constituting the active subspace are the dominant eigenvectors of the positive
 semidefinite matrix 
@@
 active subspace whose dimension is typically much smaller (in applications that
-admit such a subspace) than the dimension of the original input space. To further
+admit such a subspace) than the dimension of the original input parameter. To further
 expedite uncertainty analysis, one could fit a regression surface to $G$ using the 
@@
  %
-Clearly, the computational effort associated with constructing the matrix
+Clearly the computational effort associated with constructing the matrix
 $\hat{\mat{C}}$ scales with the number of samples, $N$. Hence, an iterative
@@
 as well as due to its interaction or coupling with other inputs. 
-Determining accurate estimates of $T_i(f)$ typically involves tens of
-thousands of model runs and is therefore prohibitive in the case of
+Determining accurate estimates of $T_i(f)$ typically involves a large 
+number of 
+model runs and is therefore can be prohibitive in the case of
 compute-intensive applications. Derivative based 
@@
 The quantities $\act{f}{i}{r}$ are called activity scores
-in~\cite{Diaz:2016,Constantine:2017}.
+in~\cite{Diaz:2016,Constantine:2017}, where links between GSA measures and
+active subspaces is explored.
 %The activity scores connect ideas from active subspaces and global sensitivity
@@
 One can attempt to reduce parameter dimension by fixing
-unimportant inputs at nominal values. Suppose $\xi_i$ is
-deemed unimportant as a result of a global sensitivity analysis (GSA)
-based on the activity scores; we want to estimate
-the approximation error that occurs once $\xi_i$ 
-is fixed at a nominal value.
+unimportant inputs at nominal values.
+
+Suppose activity scores
+are used to approximate DGSMs, and suppose
+$\xi_i$ is
+deemed unimportant as a result, due to a small activity score. 
+We want to estimate
+the approximation error that occurs once $\xi_i$ is fixed at a nominal value.
 %Let $\I \subset \{1, \ldots,
@@
 \] 
-and consider the following error indicator:
+and consider the following relative error indicator:
 \[
@@
 \] 
-\begin{theorem}
+\begin{theorem}\label{thm:error_estimate}
 We have $\Ez{ \mathcal{E}(z)} \leq 2C_i\big(\act{f}{i}{r} + \lambda_{r+1}\big)/{\V(f)}$, 
@@
 \]
-%The expression for the denominator can be seen from Lemma~\ref{lem:sum}(a).
+\end{remark}
 
-\end{remark}
+The significance of the developments in this section are as follows.
+Theorem~\ref{thm:error_estimate} provides a theoretical basis for parameter dimension
+reduction using activity scores. This is done by providing an estimate of the
+error between the reduced model and the original model. If a precise ranking of
+parameter importance based on total Sobol' indices is desired, one
+can first identify unimportant inputs by computing activity scores, and then
+perform a detailed variance based GSA of the remaining model parameters. This approach will 
+provide great computational savings as variance based GSA has to done for a small number of
+inputs deemed important based on their activity scores. Moreover, the
+presented result covers a broad class of input distributions coming from 
+Boltzmann family of distributions.
+Additionally, the normalized activity scores discussed above provide practical
+screening metrics that require only computing the activity scores. This is in
+contrast to the bound in Theorem~\ref{thm:error_estimate} that requires the
+variance $\V(f)$ of the model output.

### 2018-09-06 Paper/abstract.tex
.

@@
 for the H$_2$/O$_2$ mechanism that involves 19 chemical reactions,
-using two iterative strategies. The first strategy uses finite difference 
+using two iterative strategies. The first strategy uses perturbation methods
+such as finite differences 
 to estimate the gradient, whereas, the second strategy approximates the
 gradient using a regression-based linear approximation of the model output.
-The finite difference approach requires additional model evaluations  and 
+Perturbation-based strategies typically require additional model 
+evaluations  and 
 is therefore computationally expensive. Both strategies yield consistent

### 2018-09-06 Paper/method.tex
.

@@
 estimate provided in~\cite{Yetter:1991}.
-Specifically, we employ two strategies --
-a gradient-based strategy that involves computation of model
-gradients in order to construct the matrix $\hat{\mat{C}}$
-in~\eqref{eq:chat}, and a gradient-free approach 
-that involves a regression-based local linear approximation of
-the model output to circumvent the computational cost associated with gradient
-computation. 
-The gradient-based approach is based on Algorithm
-1.1, and the gradient-free approach is based on Algorithm 1.2
-in~\cite{Constantine:2015}. The two strategies are implemented in an
+Specifically, we employed two strategies --
+a perturbation-based strategy that involves computation of model
+gradients using finite difference in order to construct the matrix $\hat{\mat{C}}$
+in~\eqref{eq:chat}, and a regression-based approach 
+that involves a linear approximation of
+the model output to approximate the gradient. 
+The two approaches are based on Algorithms~1.1 and 1.2 in~\cite{Constantine:2015} and
+essentially differ in the way the gradient is estimated. They were implemented in an
 iterative manner to ensure computational efficiency by avoiding 
 unnecessary model evaluations once a converged active subspace
-is achieved.
+was achieved.
 
-\subsection{Gradient-based approach}
-\label{sub:grad}
+%\subsection{Gradient-based approach}
+%\label{sub:grad}
 
-As discussed earlier, in the gradient-based approach, the elements of
-$\hat{\mat{C}}$ are estimated using model gradients.  In situations where the
-gradient is not available analytically, one could consider using finite
-difference and other relatively more efficient techniques such as automatic
-differentiation~\cite{Kiparissides:2009} and adjoint-based gradient
-computation~\cite{Jameson:1988,Gunzburger:2003,Borzi:2011,Alexanderian:2017}.
-Model evaluations at neighboring points are required if finite difference is
-used. Hence, for $N$ samples in a $d$-dimensional parameter space, $N(d+1)$
-model evaluations are needed.
+As discussed earlier, gradient estimation using finite difference in the 
+perturbation-based strategy requires additional model evaluations at the
+neighboring points in the input domain. 
+Hence, for $N$ samples in a $d$-dimensional parameter space, $N(d+1)$
+model evaluations are needed. On the other hand, the regression-based
+strategy uses the available set of $N$ model evaluations to approximate
+the gradient. The computational effort in this case is therefore 
+reduced by a factor $(d+1)$ in this case. The specific sequence of steps
+for computing the active subspace is discussed as follows.
 
@@
 Note that the index, $r$, corresponds to the number of dominant eigenvectors in
-$\hat{\mat{W}}_1$. The sequence of steps for the gradient-based approach are
-provided in Algorithm~\ref{alg:grad}.
+$\hat{\mat{W}}_1$. The sequence of steps as discussed are outlined
+in Algorithm~\ref{alg:grad}.
 
@@
 \Ensure $\hat{\mat{\Lambda}}$, $\hat{\mat{W}}$, $\bm{\nu}_r(f)$ %$\eta$. %
-    \Procedure{Gradient-based}{}
+    \Procedure{Active Subspace Computation}{}
     \State Set $k$ = 0
@@
 
-To motivate our iterative strategy, we implement
-Algorithm~\ref{alg:grad} to the 19-dimensional H$_2$/O$_2$ reaction kinetics
+To assess its feasibility and suitability, we implement
+Algorithm~\ref{alg:grad} to compute the active subspace for 
+the 19-dimensional H$_2$/O$_2$ reaction kinetics
 problem with uncertain $A_i$'s. For the purpose of verification,
-$\hat{\mat{C}}$ is initially constructed using a large set of samples
-($N$~=~1000) in the input domain and the corresponding active subspace is
-computed. In Figure~\ref{fig:eig_comp}, we illustrate the comparison of
+$\hat{\mat{C}}$ was initially constructed using a large set of samples
+($N$~=~1000) in the input domain. The gradient was estimated using
+finite difference, and hence, a total of 20,000 model runs were performed. 
+In Figure~\ref{fig:eig_comp}, we illustrate the comparison of
 the resulting normalized eigenvalue spectrum ($\lambda_1, \ldots, 
@@
  ($\varepsilon_\text{L-2}^{N-n}$) between the 
-squared value of corresponding components of the dominant eigenvector, computed using $N$ = 1000
-and $n$ = $\{20,40,80,120\}$.  The quantity, $\varepsilon_\text{L-2}^{N-n}$, was found to be 
+squared value of corresponding components of the dominant eigenvector, computed using $N$~=~1000
+and $n$~=~$\{20,40,80,120\}$.  The quantity, $\varepsilon_\text{L-2}^{N-n}$, was found to be 
 $\mathcal{O}(10^{-2})$ for $n$ = 20, 40, and 80; and $\mathcal{O}(10^{-3})$ for $n$ = 120.
@@
 The iterative strategy therefore offers a significant potential for computational advantage.
-In the following section, we provide a brief description of the gradient-free approach and compare
-numerical results for the active subspace and the global sensitivity measures based on the two
-approaches.
 
-\subsection{Gradient-free approach}
-\label{sub:gradfree}
+%\subsection{Gradient-free approach}
+%\label{sub:gradfree}
 
-The gradient-free approach exploits a regression-based local linear
-approximation to the model output for estimating the gradients, $\bm{g}^i$
-in Algorithm~\ref{alg:grad}. Once again, we use an iterative strategy for
-computing the active subspace for the 19-dimensional 
-H$_2$/O$_2$ reaction kinetics problem considered in~\ref{sub:grad}.
-Moreover, we compare our findings with
-those obtained using the gradient-based approach to assess consistency
-between the two approaches.
-
-In the gradient-free approach, the model output is first computed at an initial set of $n_0$ samples. 
-An independent set of $M$ samples is also drawn according to the joint probability distribution of the inputs.
-For each sample in $M$, $p$ nearest neighbors in the $n_0$ set are identified. A least-squares fit to model
-evaluations at the $p$ samples is performed. The slope vector ($\bm{b}_i$) of the regression fit is stored
-separately and the process is repeated for each sample in $M$. The symmetric and positive semi-definite matrix,
-$\hat{\mat{C}}$ is constructed using the slope vectors as follows:
+The active subspace for the 19-dimensional problem was also computed using the regression-based
+approach. As discussed earlier, a linear regression fit to the model output was used to
+approximate the gradient using the sequence of steps outlined in Algorithm 1.2 
+in~\cite{Constantine:2015}. 
 %
-\be
-\hat{\mat{C}} = \frac{1}{M}\sum\limits_{i=1}^{M}\bm{b}_i\bm{b}_i^\top = \hat{\bm{W}}\hat{\bm{\Lambda}}\hat{\bm{W}}^\top
-\ee
+%The gradient-free approach exploits a regression-based local linear
+%approximation to the model output for estimating the gradients, $\bm{g}^i$
+%in Algorithm~\ref{alg:grad}. Once again, we use an iterative strategy for
+%computing the active subspace for the 19-dimensional 
+%H$_2$/O$_2$ reaction kinetics problem considered in~\ref{sub:grad}.
+%Moreover, we compare our findings with
+%those obtained using the gradient-based approach to assess consistency
+%between the two approaches. 
 %
-Hence, the gradient-free approach essentially differs in the procedure for constructing the matrix, $\hat{\mat{C}}$.
-Instead of the model evaluations, a linear regression-based approximation is used to estimate the gradient of
-the model output. The suitability of this approach thus depends upon the accuracy of this approximation. 
-The active subspace computation involves the same sequence of steps as outlined previously for the gradient-based 
-approach in Algorithm~\ref{alg:grad}. The procedure for constructing $\hat{\mat{C}}$ using a regression-based
-local linear approximation
-of the model output is provided in Algorithm~\ref{alg:free2}.
-
+%In the gradient-free approach, the model output is first computed at an initial set of $n_0$ samples. 
+%An independent set of $M$ samples is also drawn according to the joint probability distribution of the inputs.
+%For each sample in $M$, $p$ nearest neighbors in the $n_0$ set are identified. A least-squares fit to model
+%evaluations at the $p$ samples is performed. The slope vector ($\bm{b}_i$) of the regression fit is stored
+%separately and the process is repeated for each sample in $M$. The symmetric and positive semi-definite matrix,
+%$\hat{\mat{C}}$ is constructed using the slope vectors as follows:
+%%
+%\be
+%\hat{\mat{C}} = \frac{1}{M}\sum\limits_{i=1}^{M}\bm{b}_i\bm{b}_i^\top = \hat{\bm{W}}\hat{\bm{\Lambda}}\hat{\bm{W}}^\top
+%\ee
+%%
+%Hence, the gradient-free approach essentially differs in the procedure for constructing the matrix, $\hat{\mat{C}}$.
+%Instead of the model evaluations, a linear regression-based approximation is used to estimate the gradient of
+%the model output. The suitability of this approach thus depends upon the accuracy of this approximation. 
+%The active subspace computation involves the same sequence of steps as outlined previously for the gradient-based 
+%approach in Algorithm~\ref{alg:grad}. The procedure for constructing $\hat{\mat{C}}$ using a regression-based
+%local linear approximation
+%of the model output is provided in Algorithm~\ref{alg:free2}.
+%
 %\bigskip
@@
 %
-\bigskip
-\begin{breakablealgorithm}
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
-  \caption{Algorithm for constructing the matrix, $\hat{\mat{C}} in~\eqref{eq:chat}$}
-  \begin{algorithmic}[1]
-	\Procedure{Local Linear Approximation}{} 
-	\State Draw $N$ random samples, $\{\bm{\xi}_j\}_{j=1}^{N}$ 
-	according to $\pi_{\bm{\xi}}$.
-	\iffalse \State Project to the physical space:
-	$\{\bm{\theta}_k\}_{k=1}^{N}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{N}$ \fi
-	\State Compute $f(\bm\xi_j)$, $j=1, \ldots, N$.
-	\State Draw $M$ random samples, $\{\bm{\zeta}_i\}_{i=1}^{M}$
-	according to $\pi_{\bm{\xi}}$.
-	\State Choose an integer $p \leq N$ 
-	\State For each $i=1, \ldots, M$, compute 
-	\[
-	\begin{aligned}
-	\Phi_i &= \{ p \text{ nearest points in } \{\bm{\xi}_j\}_{j=1}^{N} \text{ to } \bm{\zeta}_i\}\\
-	\vspace{-2mm}
-	\Psi_i &= \text{subset of } f(\bm\xi_j) \text{ corresponding to the points in } \Phi_i
-	\end{aligned}
-	\]
-	\State Compute linear least square fits 
-	 $f(\bm\xi_j) \approx c_i + b_i^T\bm{\xi}_j$, \  $\bm{\xi}_j \in \Phi_i, \ f(\bm\xi_j) \in \Psi_i$
-	\State Compute the matrix 
-	\[
-	\hat{\mat{C}} \approx \frac{1}{M} \sum_{i=1}^{M} b_i b_i^T = \hat{\mat{W}}\hat{\mat{\Lambda}}\hat{\mat{W}}^\top
-	\]
-	\EndProcedure
-  \end{algorithmic}
-  \label{alg:free2}
-\end{breakablealgorithm}
-\bigskip
-
-The gradient-free approach when implemented to the 19-dimensional H$_2$/O$_2$ reaction kinetics problem
-with uncertain pre-exponents ($A_i$'s) also yielded a 1-dimensional active subspace. In Figure~\ref{fig:comp} (left),
+%\bigskip
+%\begin{breakablealgorithm}
+%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+%\renewcommand{\algorithmicensure}{\textbf{Output:}}
+%  \caption{Algorithm for constructing the matrix, $\hat{\mat{C}} in~\eqref{eq:chat}$}
+%  \begin{algorithmic}[1]
+%	\Procedure{Local Linear Approximation}{} 
+%	\State Draw $N$ random samples, $\{\bm{\xi}_j\}_{j=1}^{N}$ 
+%	according to $\pi_{\bm{\xi}}$.
+%	\iffalse \State Project to the physical space:
+%	$\{\bm{\theta}_k\}_{k=1}^{N}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{N}$ \fi
+%	\State Compute $f(\bm\xi_j)$, $j=1, \ldots, N$.
+%	\State Draw $M$ random samples, $\{\bm{\zeta}_i\}_{i=1}^{M}$
+%	according to $\pi_{\bm{\xi}}$.
+%	\State Choose an integer $p \leq N$ 
+%	\State For each $i=1, \ldots, M$, compute 
+%	\[
+%	\begin{aligned}
+%	\Phi_i &= \{ p \text{ nearest points in } \{\bm{\xi}_j\}_{j=1}^{N} \text{ to } \bm{\zeta}_i\}\\
+%	\vspace{-2mm}
+%	\Psi_i &= \text{subset of } f(\bm\xi_j) \text{ corresponding to the points in } \Phi_i
+%	\end{aligned}
+%	\]
+%	\State Compute linear least square fits 
+%	 $f(\bm\xi_j) \approx c_i + b_i^T\bm{\xi}_j$, \  $\bm{\xi}_j \in \Phi_i, \ f(\bm\xi_j) \in \Psi_i$
+%	\State Compute the matrix 
+%	\[
+%	\hat{\mat{C}} \approx \frac{1}{M} \sum_{i=1}^{M} b_i b_i^T = \hat{\mat{W}}\hat{\mat{\Lambda}}\hat{\mat{W}}^\top
+%	\]
+%	\EndProcedure
+%  \end{algorithmic}
+%  \label{alg:free2}
+%\end{breakablealgorithm}
+%\bigskip
+%
+In Figure~\ref{fig:comp} (left),
 we provide a comparison of the square of individual components of the dominant eigenvector, obtained using the two 
-approaches.
-Additionally, the quantity, $\delta \hat{\mat{W}}_{1,j}^{(r)}$ is also plotted in
-Figure~\ref{fig:comp} (right) to illustrate convergence characteristics for the gradient-free approach.
+strategies. Additionally, the quantity, $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ is also plotted in
+Figure~\ref{fig:comp} (right) to illustrate convergence characteristics for the regression-based strategy.
 %
@@
   \includegraphics[width=0.8\textwidth]{./Figures/eigv6}
-\caption{Left: An illustrative comparison of individual squared components of the converged dominant eigenvector obtained
-using the two approaches discussed in~\ref{sub:grad} and~\ref{sub:gradfree}. Right: The quantity,  $\delta \hat{\mat{W}}_{1,j}^{(r)}$
-is plotted for successive iterations to illustrate the convergence behavior.}
+\caption{Left: An illustrative comparison of individual squared components of the converged dominant
+eigenvector obtained using the two strategies. Right: The quantity,  $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$
+is plotted for successive iterations to illustrate convergence behavior for the regression-based strategy.}
 \label{fig:comp}
@@
 %
-Individual components of the dominant eigenvector from the two approaches are observed to be in excellent
-agreement with each other. Moreover, as expected,  $\delta \hat{\mat{W}}_{1,j}^{(r)}$ is observed to decrease with iterations, and
-the eigenvector is considered to have converged if its value is found to be smaller than 0.1. Note that only 30 samples
-were found to be sufficient for approximating the active subspace in the 19-dimensional input domain.
+Individual components of the dominant eigenvector from the two strategies are observed to be in excellent
+agreement with each other. Moreover, as expected,  $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ is observed 
+to decrease with iterations, and required model evaluations at only 30 samples to converge using $\tau$~=~0.1.
 
@@
 $f(\bm{\xi})$ can be approximated as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
-the subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, regarded as the \textit{sufficient summary plot} (SSP) is 
-compared for the two approaches in Figure~\ref{fig:comp_ssp}.
+the subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, 
+regarded as the \textit{sufficient summary plot} (SSP), obtained using the two strategies is 
+compared in Figure~\ref{fig:comp_ssp}.
 %
@@
   \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp}
-\caption{An illustrative comparison of the SSPs generated using the gradient-based and the gradient-free approaches.}
+\caption{An illustrative comparison of the SSPs generated using the 
+perturbation-based and the regression-based strategies for computing the active subspace.}
 \label{fig:comp_ssp}
@@
 We extend the comparative assessment of the two approaches by estimating
-the normalized activity scores for individual uncertain inputs ($\tilde{\nu_{i,r}}$; $r$=1 since the
+the normalized activity scores for individual uncertain inputs ($\tilde{\nu}_{i,r}$; $r$=1 since the
 active subspace is 1-dimensional)
@@
 verification, we compare these estimates with those obtained using the
-corresponding DGSMs, as reported in~\cite{Vohra:2018}, in
+normalized DGSMs in~\eqref{eq:ndgsm}, as reported in~\cite{Vohra:2018}, in
 Figure~\ref{fig:comp_as} (right). 
-
 %
@@
   \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}
-\caption{Left: A bar-graph of normalized activity scores ($\tilde{\nu_{i,r}}$'s) 
+\caption{Left: A bar-graph of normalized activity scores ($\tilde{\nu}_{i,r}$'s) 
 for the 19 uncertain pre-exponents ($A_i$'s); $r$

### 2018-09-06 Paper/method.tex
..

@@
 
-The above comparisons establish consistency between gradient-based and
-gradient-free approaches. This indicates that the local linear approximation
-of the model output resulted in a reasonably accurate estimation of the
-gradients in this case. In the following section, we shift our focus to 
-the high-dimensional H$_2$/O$_2$ reaction kinetics application involving
-33 uncertain inputs.
+The above comparisons indicate that the two strategies yield consistent results for the 19-dimensional
+H$_2$/O$_2$ reaction kinetics problem. In other words, the gradient estimates using pertubation and
+regression are reasonably close, and therefore, using the regression-based strategy would be computationally
+advantageous in this case. In the following section, we shift our focus to 
+the higher-dimensional H$_2$/O$_2$ reaction kinetics application involving 33 uncertain inputs.

### 2018-09-06 Paper/app.tex
..

@@
-\section{$\text{H}_2$/$\text{O}_2$ reaction kinetics: high-dimensional case}
+\section{$\text{H}_2$/$\text{O}_2$ reaction kinetics: higher-dimensional case}
 \label{sec:app}
@@
 
-The gradient-based approach discussed earlier in~\ref{sub:grad} was used to compute the
-active subspace. Using the iterative procedure, the convergence of the eigenvectors
+The active subspace was computed using the iterative procedure outlined in Algorithm~\ref{alg:grad}.
+The convergence of the eigenvectors
 was examined by tracking $\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$, plotted in 
@@
   \includegraphics[width=0.8\textwidth]{./Figures/eigv10}
-\caption{Left: An illustrative comparison of individual components of the converged dominant eigenvector obtained
-using the two approaches discussed in~\ref{sub:grad} and~\ref{sub:gradfree}. Right: The quantity,  
+\caption{Left: An illustrative comparison of individual components of the converged dominant eigenvector,
+obtained using the perturbation-based strategy. Right: The quantity,  
 $\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$
@@
 found to be roughly two orders of magnitude smaller than the first, indicating that the active subspace
-is 1-dimensional. This is further confirmed by the SSP plots in Figure~\ref{fig:hd} (right). 
+is 1-dimensional. This is further confirmed by the SSP plots in Figure~\ref{fig:hd} (right), generated
+using both, perturbation-based and regression-based strategies. 
 %
@@
 Specifically
-in the case of gradient-based approach, the SSP exhibits a linear variation that is reasonably
-captured by its linear fit, $\tilde{G}$, constructed using the sequence of steps outlined in Section~\ref{sub:ac}.
- Whereas, the SSP based on the gradient-free approach
-is relatively more scattered. Consequently, the linear fit exhibits a discrepancy with the corresponding
-fit for the gradient-based SSP.
+in the case of perturbation-based strategy, the SSP is observed to exhibit a linear variation that
+is reasonably
+captured by the corresponding straight-line fit, $\tilde{G}$, constructed using the sequence of steps
+outlined in Section~\ref{sub:ac}.
+On the other hand, the SSP constructed using the regression-based strategy
+is relatively more scattered. The two surrogates (straight-line fits) are observed to exhibit a
+discrepancy unlike our earlier observation for the 19-dimensional problem in Figure~\ref{fig:comp_ssp}.
 

### 2018-09-06 Paper/method.tex
..

@@
 H$_2$/O$_2$ reaction kinetics problem. In other words, the gradient estimates using pertubation and
-regression are reasonably close, and therefore, using the regression-based strategy would be computationally
-advantageous in this case. In the following section, we shift our focus to 
+regression are reasonably close, and therefore, using the regression-based strategy being computationally
+advantageous, would be preferred in this case. In the following section, we shift our focus to 
 the higher-dimensional H$_2$/O$_2$ reaction kinetics application involving 33 uncertain inputs.

### 2018-09-07 Paper/app.tex

@@
 On the other hand, the SSP constructed using the regression-based strategy
-is relatively more scattered. The two surrogates (straight-line fits) are observed to exhibit a
+is relatively more scattered. Consequently, the two surrogates (straight-line fits) are observed to exhibit a
 discrepancy unlike our earlier observation for the 19-dimensional problem in Figure~\ref{fig:comp_ssp}.
@@
 
-The 1-dimensional surrogate ($\tilde{G}$) is investigated for its accuracy as well as the ability to capture the 
-uncertainty in the
+The 1-dimensional surrogate ($\tilde{G}$) shown in Figure~\ref{fig:hd}~(right) is investigated
+for its accuracy as well as the ability to capture the uncertainty in the
 model output in two ways. Firstly, we estimate the relative L-2 norm of the discrepancy~($\varepsilon_d$)
@@
 It was estimated to be 1.35$\times10^{-2}$ and 1.02$\times10^{-1}$
-using $\tilde{G}$ from the gradient-based approach and the gradient-free approach respectively. Hence,
-an order of magnitude increase in $\varepsilon_d$ was estimated in the case of gradient-free approach.
-Hence, in the norm-sense, it can be said that the 1-dimensional surrogate in the active subspace, obtained
-using the gradient-based approach is remarkably accurate with a relative error of less than 1.5$\%$, whereas
-the gradient-free approach resulted in a relative error of about 10$\%$.
+using $\tilde{G}$ based on the perturbation-based and the regression-based strategies respectively. Hence,
+$\varepsilon_d$ was found to increase by an order of magnitude when using the regression-based strategy.
+Therefore, in the norm-sense, it can be said that the 1-dimensional surrogate in the active subspace, obtained
+using the perturbation-based approach is remarkably accurate with a relative error of less than 1.5$\%$, whereas
+the regression-based approach resulted in a relative error of about 10$\%$.
 
-Secondly, we verify the accuracy of the two surrogates in a probabilistic setting. In particular, we compare 
+Secondly, we verified the accuracy of the two surrogates in a probabilistic setting. In particular, we compared 
 probability density functions (PDFs) obtained using the true set of model evaluations, and 1-dimensional
-surrogates from the two approaches, in Figure~\ref{fig:pdf_33D}. Note that the three PDFs were evaluated 
+surrogates ($\tilde{G}$'s) from the two strategies, in Figure~\ref{fig:pdf_33D}. Note that the three PDFs were evaluated 
 using the same set of 10$^4$ samples in the cross-validation set. 
@@
 \caption{A comparison of the PDFs of ignition delay, obtained using model 
-evaluations (solid line) and 1-dimensional surrogates using the gradient-free approach (dashed line) and the gradient-based
-approach (dashed-dotted line). The same set of 10$^4$ samples in the cross-validation set were used in each case.}
+evaluations (solid line) and 1-dimensional surrogates using the regression-based strategy (dashed line) and the
+perturbation-based strategy (dashed-dotted line). The same set of 10$^4$ samples in the cross-validation set were 
+used in each case.}
 \label{fig:pdf_33D}
@@
 %
-The PDFs based on the model evaluations and the gradient-based approach are nearly identical, confirming
-the existence of a 1-dimensional active subspace as well as the accuracy of the gradient-based approach. 
-However, though the PDF based on the gradient-free approach is observed to show consistency with the true 
-PDF in the modal
-estimate; it underestimates the spread in the ignition delay. Specific values of the mean and the standard deviation
-in each case are provided in Table~\ref{tab:stats}. 
+The PDFs based on the model evaluations and the perturbation-based strategy are nearly identical, confirming
+the existence of a 1-dimensional active subspace as well as the accuracy of this strategy. 
+On the other hand, although the PDF obtained using the regression-based strategy is observed to show
+consistency with the true PDF in the modal
+estimate, it clearly underestimates the spread in the ignition delay. Specific values of the mean and the standard 
+deviation in each case are provided in Table~\ref{tab:stats}. 
 %
@@
 $G$~(Model) & 0.133 & 0.0198 \\
-$\tilde{G}$~(Gradient-based) & 0.133 & 0.0196 \\
-$\tilde{G}$~(Gradient-free) & 0.133 & 0.0167 \\
+$\tilde{G}$~(Perturbation-based) & 0.133 & 0.0196 \\
+$\tilde{G}$~(Regression-based) & 0.133 & 0.0167 \\
 \bottomrule
@@
 \caption{The mean ($\mu$), and the standard deviation ($\sigma$), computed using the model ($G$), and
-the surrogate ($\tilde{G}$) based on the two approaches at 10$^4$ samples in the cross-validation
+the surrogate ($\tilde{G}$) based on the two strategies at 10$^4$ samples in the cross-validation
 set.}
@@
 consistent in their prediction of mean value of the ignition delay.  The standard deviation estimates based on model
-evaluations and the gradient-based approach are found to be in close agreement, whereas, it  is accurately estimated
-by the gradient-free approach only upto the first significant figure.
+evaluations and the perturbation-based strategy are found to be in close agreement, whereas, it  is accurately
+ estimated in the case of regression-based strategy only upto the first significant digit.
 

### 2018-09-07 Paper/intro.tex

@@
 This involves repeated evaluations of the gradient of a model output with
-respect to the input parameters. Estimating the
-gradients can however be computationally challenging in high-dimensional settings. 
-Hence, in this work, we also explore the suitability of a gradient-free
-approach that involves a regression-based local linear approximation of
-the model output as discussed later in~\ref{sub:gradfree}. The two approaches
- are based on
-algorithms 1.1 and 1.2 in~\cite{Constantine:2015}. Both strategies
-are implemented in an iterative manner i.e. a new set of model evaluations are
-generated, and gradient evaluations are performed at each iteration until 
-a reasonably accurate approximation to the active subspace is achieved, as
-discussed later in Section~\ref{sec:method}.
+respect to the input parameters. Several numerical techniques are available
+for computing the gradient. Perturbation-based techniques
+include finite differences and more advanced methods involving
+adjoints~\cite{Jameson:1988,Gunzburger:2003,Borzi:2011,Alexanderian:2017}. The
+adjoint-based method requires a solution of the state equation (forward solve)
+and the corresponding adjoint equation. Hence, it is 
+limited by the availability of an adjoint solver. Additional model evaluations
+at neighboring points are required if finite difference is used which increases
+the computational effort. Regression-based techniques aim to estimate the
+gradient by approximating the model output using a regression fit. These
+are computationally less intensive than the former. However, as expected,
+there is a trade-off between computational effort and accuracy in the two
+approaches for estimating the gradient.
+
+In this work, we implement two strategies for computing the active subspace for the
+H$_2$/O$_2$ mechanism, and refer to them as \textit{perturbation-based} and 
+\textit{regression-based} to highlight the difference in the approach for estimating
+the gradient of the model output as discussed.  
+Other than the difference in gradient estimation, the two strategies follow the same
+sequence of steps provided in~\cite{Constantine:2015}. However, 
+in order to reduce computational effort, the two strategies are implemented in an
+iterative manner in this work as discussed later in Section~\ref{sec:method}.
 
@@
 in~\cite{Diaz:2016,Constantine:2017} that DGSMs can be approximated by
-exploiting their links with active subspaces. This led to definition of the so
-called \emph{activity scores}. In Section~\ref{sub:gsa}, we build on these
+exploiting their links with active subspaces. This led to the definition of the 
+so-called \emph{activity scores}. In Section~\ref{sub:gsa}, we build on these
 ideas to provide a complete analysis of links between Sobol indices, DGSMs, and
@@
 method of active subspaces to achieve efficiency in uncertainty propagation.  
-\item Comprehensive numerical investigation of the gradient-based and gradient-free approach: 
+\item Comprehensive numerical investigation of the perturbation-based and the regression-based approaches: 
 We investigate the suitability of both approaches
@@
 sensitivity measures through a comprehensive set of numerical experiments. Our
-results reveal insight into the merits of the methods approach as well as
+results reveal insight into the merits of the methods as well as
 their shortcomings.  
@@
 established. In section~\ref{sec:method}, a systematic framework for computing
-the active subspace using an iterative gradient-based approach is provided.
-Numerical results based on the gradient-free approach are compared with those
-obtained using the gradient-based approach.  The active subspace is computed
-for the H$_2$/O$_2$ reaction kinetics problem whereby only the $A_i$'s are
+the active subspace is provided. 
+Numerical results based on the perturbation-based approach are compared with those
+obtained using the regression-based approach.  For our initial investigation aimed
+as assessing the feasibility and suitability of the two strategies, the active 
+subspace is computed for a 19-dimensional 
+H$_2$/O$_2$ reaction kinetics problem whereby only the $A_i$'s are
 considered as uncertain. In this case, the two approaches are observed to yield
-consistent results. In section~\ref{sec:app}, the gradient-based approach is
-further implemented to compute the active subspace for the higher-dimensional
-H$_2$/O$_2$ reaction kinetics problem. The resulting active subspace is once again
-found to be 1-dimensional, thereby indicating enormous potential for computational
-savings in this application. The active subspace is also computed using the
-gradient-free approach. It is observed that although the uncertainty in the
-ignition delay is reasonably approximated, the gradient-free approach fails to
-capture the global sensitivity of all the uncertain inputs. Hence, a trade-off
-between efficiency and accuracy between the two approaches is observed in this
-case.  Finally, a summary and discussion based on our findings is included in
+consistent results. Motivated by these results, we further implement the two
+strategies to compute the active subspace for a 33-dimensional H$_2$/O$_2$
+reaction kinetics problem in section~\ref{sec:app}. Once again, the two
+strategies reveal the existence of a 1-dimensional active subspace. However,
+it is observed that although the regression-based strategy results in a 
+reasonable approximation of the mean, mode, and the uncertainty in the ignition
+delay, it fails to capture the global sensitivity of all the uncertain inputs. 
+Hence, a trade-off between efficiency and accuracy between the two approaches is
+apparent in this case.  Finally, a summary and discussion based on our findings is included in
 section~\ref{sec:conc}.

### 2018-09-07 Paper/assa.tex
.
..

@@
 In~\cite{Vohra:2018} the following screening metric
-\[
+\be
    \tilde{\nu}_i(f) = \frac{C_i \nu_i(f)}{\sum_{i=1}^\Np C_i \nu_i(f)},
-\]
+\label{eq:ndgsm}
+\ee
 was shown to be useful for detecting unimportant inputs. 
@@
 perform a detailed variance based GSA of the remaining model parameters. This approach will 
-provide great computational savings as variance based GSA has to done for a small number of
+provide great computational savings as variance based GSA will now be performed only
+for a small number of
 inputs deemed important based on their activity scores. Moreover, the

### 2018-09-07 Paper/method.tex
..

@@
 strategy uses the available set of $N$ model evaluations to approximate
-the gradient. The computational effort in this case is therefore 
+the gradient. The computational effort is therefore 
 reduced by a factor $(d+1)$ in this case. The specific sequence of steps
@@
 \renewcommand{\algorithmicensure}{\textbf{Output:}}
-  \caption{An iterative gradient-based approach for discovering the active subspace}
+  \caption{An iterative strategy for discovering the active subspace}
   \begin{algorithmic}[1]

### 2018-09-07 Paper/app.tex
..

@@
 
-The normalized activity scores ($\tilde{\nu}_{i,r}$) based on the dominant eigenspace using the two
-approaches (gradient-based and gradient-free) are compared with the total Sobol' indices
+The normalized activity scores ($\tilde{\nu}_{i,r}$) based on the dominant eigenspace, obtained
+using the two strategies (perturbation-based and regression-based) are compared with the 
+total Sobol' indices
 in Figure~\ref{fig:as_33D}. Note that the Sobol' indices were computed using the verified
-1-dimensional surrogate in the active subspace for the gradient-based approach. 
+1-dimensional surrogate ($\tilde{G}$) in the active subspace, computed using the 
+perturbation-based strategy. 
 %
@@
 consistent
-with the normalized activity scores evaluated using the gradient-based approach. The gradient-free
-approach yields consistent results for the $A_i$'s, but it does not capture the sensitivity associated with
-$E_{a,15}$. This is because numerical errors are incurred when approximating
-the model gradients using the regression-based approximation in the gradient-free case. 
+with the normalized activity scores evaluated using the perturbation-based strategy. The regression-based
+strategy yields consistent results for the $A_i$'s, but it does not capture the sensitivity associated with
+the most important parameter, $E_{a,15}$. This is because numerical errors are incurred when approximating
+the model gradients using a linear regression fit in this case. 
 Secondly, our findings indicate that   

### 2018-09-07 Paper/conc.tex
..

@@
  Hence, we focused our efforts on reducing the dimensionality of the problem by identifying important
- directions (active subspaces) in the parameter space such that the variability in the model output is 
+ directions (active subspace) in the parameter space such that the variability in the model output is 
  predominantly captured along these directions. Additionally, we demonstrated that the activity scores,
@@
 Two computational strategies were explored to compute the active subspace for the H$_2$/O$_2$
-kinetics application in this work. The first strategy referred to as gradient-based requires computation of gradient of the 
-model output with respect to the input, whereas, the second strategy, referred to as gradient-free implements a local
-linear approximation to the model output in order to estimate the gradient using a regression-based fit.
+kinetics application in this work. The first strategy referred to as perturbation-based
+used finite difference to estimate the gradient of the model output with respect to the uncertain
+inputs. The second strategy referred to as regression-based involved gradient estimation by
+means of approximating the available set of model evaluations using a linear regression fit. 
 The two strategies were shown to yield
@@
 
-Both strategies were further implemented to discover and compute the active
-subspace for the 33-dimensional problem. The gradient-based approach yielded a 1-dimensional active subspace
-and the activity scores which revealed that the uncertainty in the ignition delay is largely dependent on $A_1$, $A_9$, 
-$E_{a,1}$ and $E_{a,15}$, although contributions from $A_{15}$ and $A_{17}$ were also found to be significant. The activity
-scores were verified against the total Sobol' indices, computed using the 1-dimensional surrogate fit to the model response in
-the subspace. The 1-dimensional surrogate was verified for accuracy by evaluating the relative L-2
-norm of the error between its predictions and model evaluations at $10^4$ samples in the 33-dimensional input domain.
-Additionally, the surrogate was verified by comparing probability distributions (based on kernel density estimation) for
-the ignition delay based on its predictions and model evaluations. The two PDFs were observed to be nearly identical
-and consequently, their mean and standard deviations were found to be in excellent agreement. On the other hand,
-the gradient-free approach failed to capture the sensitivity towards the most important parameter, $E_{a,15}$. 
+The two strategies were further implemented to discover and compute the active
+subspace for the 33-dimensional problem. Both strategies revealed the existence of a 1-dimensional
+active subspace. The activity scores based on the dominant eigenvector and the total Sobol indices
+based on the 1-dimensional surrogate in the subspace, computed using the 
+perturbation-based strategy were found to be consistent with each other. Both sensitivity measures
+ indicated that the uncertainty in the ignition delay is largely dependent on $A_1$, $A_9$, 
+$E_{a,1}$ and $E_{a,15}$, although contributions from $A_{15}$ and $A_{17}$ were also found to be significant.
+
+Computational gains using the proposed strategy in this work are essentially realized by using the
+1-dimensional surrogate in the active subspace that essentially casts a mutlivariate input in the full space
+into a univariate input in the reduced space. The surrogate can be used to accelerate forward propagation
+of the uncertainty as well as parameter estimation in a Bayesian setting. Hence, we focused our efforts to
+verify the accuracy of the 1-dimensional surrogates, obtained using both strategies.  
+The 1-dimensional surrogates were assessed for accuracy by first evaluating the relative L-2
+norm of the error between its predictions and model evaluations at $10^4$ independent MC samples in the
+33-dimensional input domain. The relative error based on predictions from the surrogate, computed
+using the regression-based strategy were found to be an order of magnitude larger than that estimated
+using the surrogate from the perturbation-based strategy.
+
+Additionally, the surrogates were assessed by comparing probability distributions (based on kernel density estimation)
+for the ignition delay, constructed using surrogate predictions and model evaluations at the same set of 10$^4$
+random samples. The PDFs based on model evaluations and the 1-dimensional surrogate from the perturbation-based
+strategy were found to be almost identical. However, the PDF based on 1-dimensional surrogate from the
+regression-based strategy although captured the modal estimate with reasonable accuracy, was observed to
+under-estimate the uncertainty in the ignition delay. Mean estimates of the ignition delay using model evaluations
+and predictions using the two surrogates at 10$^4$ samples were found to be in agreement. However, as expected, the
+standard deviation, computed using predictions from the surrogate based on the regression-based strategy was
+found to be accurate only upto the first significant digit. Moreover, it was observed that the activity scores,
+computed using the regression-based strategy failed to capture the sensitivity towards the 
+most important parameter, $E_{a,15}$. 
 We attribute this discrepancy to two potential sources of numerical errors: (1) Regression-based approximation of the
 gradient of the model output, and (2) Scatter in the corresponding SSP (see Figure~\ref{fig:hd}) leading to numerical
-error incurred by the linear regression fit in this case. The PDF was observed to be consistent with the true PDF in its
-estimate for the mode and the mean value of the ignition delay. However, it was observed to underestimate the uncertainty 
-associated with the ignition delay. Hence, the gradient-free approach might prove to be valuable in applications 
-involving intensive simulations (rendering the gradient-based approach prohibitive), and focused on a reasonable 
-approximation of the mean value and the uncertainty associated with the quantity of interest. Based on our findings, the 
-estimate for the mean is expected to be relatively more accurate.
+error incurred by the linear regression fit in this case. Based on our findings in this work, it can be said that
+the regression-based strategy would be useful in situations involving complex mechanisms with a large number of
+uncertain inputs thereby rendering the UQ analysis intensive, and in applications where the focus is on a
+reasonable approximation of the statistics of the QoI.
 
- The authors would like to highlight that the computational strategy presented in this work is agnostic to the choice
+ The authors would like to highlight that the computational framework presented in this work is agnostic to the choice
  of the chemical system and could thus be easily adapted for other systems. However, it is important to note that

### 2018-09-22 Paper/assa.tex
..

--- 2018-09-05 Paper/assa.tex
@@
 either uniformly distributed or
-distributed according to a Boltzmann probability distribution.
+distributed according to the Boltzmann probability distribution.
 Note that a measure $\mu$

### 2018-09-22 Paper/method.tex

@@
 
-In this section, we present a generalized framework for computing the
+In this section, we present an efficient framework for computing the
 active subspace. The framework is employed to the H$_2$/O$_2$ reaction kinetics
@@
 estimate provided in~\cite{Yetter:1991}.
-Specifically, we employed two strategies --
+Specifically, we employed two strategies: 
 a perturbation-based strategy that involves computation of model
@@
 in~\eqref{eq:chat}, and a regression-based approach 
-that involves a linear approximation of
+that involves a local linear approximation of
 the model output to approximate the gradient. 
@@
 
-As discussed earlier, gradient estimation using finite difference in the 
+As discussed earlier, gradient estimation using finite differences in the 
 perturbation-based strategy requires additional model evaluations at the

### 2018-09-22 Paper/assa.tex
sync ...

--- 2018-09-07 Paper/assa.tex
@@
 The utility of this result is realized in problems with 
-high-dimensional parameter spaces in which 
+high-dimensional parameters in which 
 the eigenvalues $\lambda_i, i=1,\ldots,N_p$, decay rapidly to zero; in 
@@
 
-We also record the following useful relations:
+The relations recored in the following lemma will be useful in the discussion 
+that follows.
 \begin{lemma}\label{lem:sum}
@@
 where for each $i$, $C_i$ is an appropriate \emph{Poincar\'{e}} constant
-that depends on the distribution of $\xi_i$; see~\cite{Lamboni:2013}.
+that depends on the distribution of $\xi_i$.
 For instance, if $\xi_i$ is uniformly distributed on $[-1, 1]$, then $C_i = 4/\pi^2$; and in the 
 case $\xi_i$ is normally distributed with variance $\sigma_i^2$, then $C_i = \sigma_i^2$. 
+Note that~\eqref{equ:sobol_bound} for the special cases of 
+uniformly distributed or normally distributed inputs was established first in~\cite{Sobol:2009}.
 The bound~\eqref{equ:sobol_bound} provides a strong theoretical basis for using DGSMs to identify 
@@
 inputs.  
-One can attempt to reduce parameter dimension by fixing
+Subsequently, 
+one can attempt to reduce parameter dimension by fixing
 unimportant inputs at nominal values.
@@
 the approximation error that occurs once $\xi_i$ is fixed at a nominal value.
-%Let $\I \subset \{1, \ldots,
-%Np\}$ be an index set that indexes the unimportant parameters in the parameter
-%vector $\vec{\xi}$.  Letting $|\I|$ denote the number of elements of $\I$, 
-%we denote by $\vec{z} \in \R^{|\I|}$ a generic vector of nominal values for
-%the unimportant parameters $\vec\xi_\I$, and let $\vec{y}$ be a generic vector of
-%important variables, $\vec\xi_{\I^c}$,
-%where $\I^c$ denotes the complement of $\I$ in $\{1, \ldots, \Np\}$. 
-%For a vector $\xi$, we let $\xin$ denote the vector obtained by removing
-%$\xi_i$. 
-%
 To formalize this process, we proceed as follows.
@@
 Consider the \emph{reduced} model, 
-obtained by fixing $\xi_i$ at a nominal value $z \in \R$: 
+obtained by fixing $\xi_i$ at the nominal value: 
 \[
@@
 \] 
+This error indicator is a function of $z$ with $z$ distributed 
+according to the distribution of $\xi_i$.
 \begin{theorem}\label{thm:error_estimate}
@@
 \end{theorem}
-%
-%\[
-%       T_i(f) \leq \frac{C_i}{\V(f)}(\act{f}{i}{r} + \lambda_{r+1}).
-%\]
-%we obtain the following useful inequality, which
-%links activity scores and Sobol' indices:
-%for $1 \leq r \leq \Np$, and $i = 1, \ldots, \Np$, we have
-%\[
-%       T_i(f) \leq \frac{C_i}{\V(f)}(\act{f}{i}{r} + \lambda_{r+1}).
-%\]
-%This was noted in~\cite{Diaz:2016} for the specific case of all $\xi_i$'s being
-%iid $U(0, 1)$, in which case $C_i = 1/\pi^2$, $i = 1, \ldots, \Np$.
 \begin{proof} 
@@
 
-In~\cite{Vohra:2018} the following screening metric
+In~\cite{Vohra:2018} the screening metric
 \be

### 2018-09-22 Paper/app.tex
..

@@
 Figure~\ref{fig:conv_app} (right). In Figure~\ref{fig:conv_app} (left), individual
-components of the converged eigenvector are illustrated. The convergence was established using
+components of the converged eigenvector that constitutes the active subspace,
+are illustrated. The convergence was established using
 a $\tau$ value of 0.02. Finite difference was used to compute model gradients at 40 samples in
@@
   \includegraphics[width=0.8\textwidth]{./Figures/eigv10}
-\caption{Left: An illustrative comparison of individual components of the converged dominant eigenvector,
+\caption{Left: Individual components of the dominant eigenvector in the
+active subspace,
 obtained using the perturbation-based strategy. Right: The quantity,  
@@
 
-In Figure~\ref{fig:hd} (left), we illustrate the resulting eigenvalue spectrum. The second eigenvalue is
+In Figure~\ref{fig:hd} (left), we illustrate converged estimates of the eigenvalues.
+ The second eigenvalue is
 found to be roughly two orders of magnitude smaller than the first, indicating that the active subspace
@@
 
-\subsection{Surrogate Verification}
+\subsection{Surrogate Assessment}
 \label{sub:verify}
@@
 for its accuracy as well as the ability to capture the uncertainty in the
-model output in two ways. Firstly, we estimate the relative L-2 norm of the discrepancy~($\varepsilon_d$)
+model output in two ways. Firstly, we estimate the relative L$_2$ norm of the discrepancy~($\varepsilon_d$)
 between estimates of ignition delay in the case of H$_2$/O$_2$ reaction, obtained using the 
@@
 %
-Several useful inferences can be drawn from the above plot. Firstly, the total Sobol indices are found to be
+Several useful inferences can be drawn. Firstly, the total Sobol indices are found to be
 consistent
 with the normalized activity scores evaluated using the perturbation-based strategy. The regression-based
-strategy yields consistent results for the $A_i$'s, but it does not capture the sensitivity associated with
-the most important parameter, $E_{a,15}$. This is because numerical errors are incurred when approximating
+strategy yields consistent results for the $A_i$'s, but it does not capture the sensitivity associated with one of the most important parameters, $E_{a,15}$.
+This is because numerical errors are incurred when approximating
 the model gradients using a linear regression fit in this case. 

### 2018-09-22 Paper/conc.tex
..

@@
 gradient of the model output, and (2) Scatter in the corresponding SSP (see Figure~\ref{fig:hd}) leading to numerical
-error incurred by the linear regression fit in this case. Based on our findings in this work, it can be said that
-the regression-based strategy would be useful in situations involving complex mechanisms with a large number of
-uncertain inputs thereby rendering the UQ analysis intensive, and in applications where the focus is on a
-reasonable approximation of the statistics of the QoI.
+error incurred by the linear regression fit in this case. Hence, the regression-based approximation of the gradient 
+of the model output although reduces computational effort, the present investigations clearly reveal a potential pitfall associated
+with this approach. Therefore, this approach should only be preferred in situations involving intensive simulations, and 
+where the focus is essentially on estimating the statistics of the QoI as opposed to a more detailed analysis such as
+parametric sensitivity analysis.
 
- The authors would like to highlight that the computational framework presented in this work is agnostic to the choice
- of the chemical system and could thus be easily adapted for other systems. However, it is important to note that
- its applicability depends upon the nature of the relationship between the model output and the inputs. Specifically,
- the model output is required to be differentiable throughout the considered input domain to be able to evaluate
- the active subspace with reasonable accuracy. Hence, it is a realistic possibility that an active subspace might not
- exist for a given input-output relationship. While the active subspace could be exploited for a forward propagation of
- the uncertainty from inputs to the output, the low-dimensional surrogate could further be used to accelerate 
- calibration of the rate-controlling parameters in a Bayesian setting. Moreover, the activity scores could help
- identify important stages in a complex reaction mechanism and thereby guide development of a simplified
- mechanism. Furthermore, dimension reduction using active subspaces could be useful in studies aimed at
- investigating the inadequacy or model error associated with the reaction rate laws for such systems.
+ We would like to highlight that the computational framework presented in this work is agnostic to the choice
+ of the chemical system and could thus be easily adapted for other systems as long as the quantity of interest
+is continuously differentiable in the considered domain of the inputs.
+We have demonstrated that the active subspace could be exploited for a forward propagation of
+ the uncertainty from inputs to the output. However, the activity scores and the low-dimensional surrogate could further guide optimal allocation of computational resources for  
+ calibration of the important rate-controlling parameters in a Bayesian setting. 
+Additionally, dimension reduction using active subspaces could assist in developing robust formulations
+for predicting discrepancy between simulations and measurements due to epistemic uncertainty in the model inputs.

### 2018-09-22 Paper/cover.tex
..
minor ...

@@
 \begin{center}
-\textsc{Efficient UQ for chemical systems using active subspaces}
+%\textsc{Efficient UQ for chemical kinetics using active subspaces}
+\textsc{Subspace-based dimension reduction for chemical kinetics applications with epistemic uncertainty}
 
@@
 \begin{tabbing}
-Corresponding Author: \hspace{5mm} \= Sankaran Mahadevan\\
+Corresponding Author: \hspace{5mm} \= Manav Vohra\\
        \>  Department of Civil and Environmental Engineering\\
        \>  Vanderbilt University\\
-       \>  272 Jacobs Hall, VU Mailbox: PMB 351831 \\
+       \>  279 Jacobs Hall, VU Mailbox: PMB 351831 \\
        \>  Nashville, TN 37235 \\
@@
 Fax:   \> (615) 343-3773 \\
-Email: \>  sankaran.mahadevan@vanderbilt.edu   \\
+Email: \>  manav.vohra@vanderbilt.edu   \\
 \\
 Submitted to: \> \textit{The Journal of Chemical Physics} \\
-\>  August 2018\\
+\>  September 2018\\
 

### 2018-09-22 Paper/method.tex
..

@@
 Hence, for $N$ samples in a $d$-dimensional parameter space, $N(d+1)$
-model evaluations are needed. On the other hand, the regression-based
-strategy uses the available set of $N$ model evaluations to approximate
-the gradient. The computational effort is therefore 
-reduced by a factor $(d+1)$ in this case. The specific sequence of steps
+model evaluations are needed. On the other hand, gradient estimation
+using the regression-based approach involves a series of linear regression
+fits to subsets of available evaluations as discussed 
+in~\cite{Constantine:2015}. Hence, the computational effort is  
+reduced by a factor $(d+1)$ when using the regression-based approach.
+The specific sequence of steps
 for computing the active subspace is discussed as follows.
@@
 	\EndLoop
-	\State Compute $\nu_{i,r}(f) = \sum\limits_{j=1}^{r} \Lambda_j w_{i,j}^2$,
+	\State Compute $\nu_{i,r}(f) = \sum\limits_{j=1}^{r} \lambda_j w_{i,j}^2$,
 	$i=1,\ldots,N_p$.
@@
 to improve with the sample size. Since 
-$\left(\frac{\lambda_1}{\lambda_2}\right)\gg 1$, we expect 
+$\lambda_1$ is roughly three orders of magnitude larger than $\lambda_2$, we expect 
 a 1-dimensional active subspace. Hence, the first eigenvector 
 sufficiently captures the uncertainty in
-the  model output. To further confirm this, we evaluate a relative L-2 norm of the difference
- ($\varepsilon_\text{L-2}^{N-n}$) between the 
-squared value of corresponding components of the dominant eigenvector, computed using $N$~=~1000
-and $n$~=~$\{20,40,80,120\}$.  The quantity, $\varepsilon_\text{L-2}^{N-n}$, was found to be 
+the  model output. To further confirm this, we evaluate a relative L$^2$ norm of the difference
+ ($\varepsilon_{\text{L}^{2}}^{N-n}$) between the 
+squared value of corresponding components of the dominant eigenvector, computed using $N$~=~1000 ($\vec{w}_{1,N}$)
+and $n$~=~$\{20,40,80,120\}$ ($\vec{w}_{1,n}$) as follows:
+%
+\be
+\varepsilon_{\text{L}^{2}}^{N-n} = 
+\frac{\|\vec{w}_{1,N}^2-\vec{w}_{1,n}^2\|_2}{\|\vec{w}_{1,N}^2\|_2}
+\ee
+%
+The quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$, was found to be 
 $\mathcal{O}(10^{-2})$ for $n$ = 20, 40, and 80; and $\mathcal{O}(10^{-3})$ for $n$ = 120.
@@
 reasonable accuracy in this case. 
-The iterative strategy therefore offers a significant potential for computational advantage.
+The iterative strategy therefore offers a significant potential for computational gains.
 
@@
 
-As mentioned earlier, the model output $f(\bm{\xi})$ predominantly varies in the active subspace. Hence, 
+As mentioned earlier, the model output $f(\bm{\xi})$ i.e. the igntion delay in the
+H$_2$/O$_2$ reaction in this case, varies
+ predominantly in the active subspace. Hence, 
 $f(\bm{\xi})$ can be approximated as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
@@
 that the response in ignition
-delay based on the considered prior marginals for $A_i$'s is observed to be approximately linear in the
+delay based on the considered probability distributions for $A_i$'s is observed to be approximately linear in the
 active subspace. This observation further indicates that the active subspace is 1-dimensional in this case.

### 2018-09-22 Paper/conc.tex
..

@@
 with this approach. Therefore, this approach should only be preferred in situations involving intensive simulations, and 
-where the focus is essentially on estimating the statistics of the QoI as opposed to a more detailed analysis such as
-parametric sensitivity analysis.
+where the focus is essentially on estimating the statistics of the QoI as opposed to a more detailed analysis such as those involving
+parametric sensitivity.
 

### 2018-09-22 Paper/conc.tex
..

--- 2018-09-22 Paper/conc.tex
@@
 error incurred by the linear regression fit in this case. Hence, the regression-based approximation of the gradient 
-of the model output although reduces computational effort, the present investigations clearly reveal a potential pitfall associated
-with this approach. Therefore, this approach should only be preferred in situations involving intensive simulations, and 
-where the focus is essentially on estimating the statistics of the QoI as opposed to a more detailed analysis such as
-parametric sensitivity analysis.
+of the model output although reduces computational effort, the present investigations clearly 
+reveals potential pitfalls associated with this approach. More specifically:
+\begin{itemize}
+\item The desired statistics (e.g. mean and variance) of the QoI might not be approximated
+to a desired level of accuracy.
+\item The approach might fail to capture the sensitivity  of the QoI towards all inputs. 
+\end{itemize}
+Therefore, the regression-based approach should only be preferred in situations involving intensive simulations such that the perturbation-based methods are intractable, and the objective is
+essentially on obtaining rough estimates of the statistics of the QoI as opposed to a more detailed analysis such as those involving
+parametric sensitivity.
 

### 2018-09-22 Paper/conc.tex
conc ...

--- 2018-09-22 Paper/conc.tex
@@
 
- In this work, we focused on the uncertainty associated with the rate-controlling parameters in
- the H$_2$/O$_2$ reaction mechanism and its impact on ignition delay predictions. The mechanism
- involves 19 different reactions and in each case, the reaction rate depends upon the choice of
- a pre-exponent and an activation energy. Hence, in theory, the evolution of the chemical system
- depends upon 38 inputs which is reasonably high-dimensional. Conventional means
- for uncertainty quantification such as those involving a surrogate model construction as well as
- sensitivity analysis would thus be computationally challenging and even prohibitive in some cases. 
- Hence, we focused our efforts on reducing the dimensionality of the problem by identifying important
- directions (active subspace) in the parameter space such that the variability in the model output is 
- predominantly captured along these directions. Additionally, we demonstrated that the activity scores,
- computed using the components of the dominant eigenvectors can be used to approximate the
- DGSMs. Furthermore, we established generalized mathematical linkages between the different global sensitivity measures:
- activity scores, DGSMs, and total Sobol' index which could be exploited to reduce computational effort 
- associated with global sensitivity analysis.
+In this work, we focused on the uncertainty associated with the
+rate-controlling parameters in the H$_2$/O$_2$ reaction mechanism and its
+impact on ignition delay predictions. The mechanism involves 19 different
+reactions and in each case, the reaction rate depends upon the choice of a
+pre-exponent and an activation energy. Hence, in theory, the evolution of the
+chemical system depends upon 38 inputs. 
+\alennote{Maybe say something about how we went from 38 to 33?} 
+%Conventional means for uncertainty quantification such as those involving
+%surrogate model construction as well as sensitivity analysis would thus be
+%computationally challenging.
+To facilitate efficient uncertainty analysis, we focused our efforts on
+reducing the dimensionality of the problem by identifying important directions
+in the parameter space such that the variability in the model output is
+predominantly captured along these directions. This was done through active
+subspace construction.  Additionally, we demonstrated that the activity scores,
+computed using the components of the dominant eigenvectors provide an efficient
+means for approximating derivative based global sensitivity measures (DGSMs).
+Furthermore, we established generalized mathematical linkages between the
+different global sensitivity measures: activity scores, DGSMs, and total Sobol'
+index which could be exploited to reduce computational effort associated with
+global sensitivity analysis.
 
-Two computational strategies were explored to compute the active subspace for the H$_2$/O$_2$
-kinetics application in this work. The first strategy referred to as perturbation-based
-used finite difference to estimate the gradient of the model output with respect to the uncertain
-inputs. The second strategy referred to as regression-based involved gradient estimation by
-means of approximating the available set of model evaluations using a linear regression fit. 
-The two strategies were shown to yield
-consistent results for the 19-dimensional problem whereby only the pre-exponents were considered to be uncertain.
-Additionally, the activity scores were also shown to be consistent with the screening metric estimates based on
-DGSMs in~\cite{Vohra:2018}. Note that an iterative procedure was adopted to enhance the computational efficiency
+Two computational strategies were explored to compute the active subspace for
+the H$_2$/O$_2$ kinetics application in this work. The first strategy, referred
+to as perturbation-based, used finite differences to estimate the gradient of the
+model output with respect to the uncertain inputs. The second strategy, referred
+to as regression-based, involved gradient estimation by means of approximating
+the available set of model evaluations using a linear regression fit.  The two
+strategies were shown to yield consistent results for the 19-dimensional
+problem whereby only the pre-exponents were considered to be uncertain.
+Additionally, the activity scores were also shown to be consistent with the
+screening metric estimates based on DGSMs in~\cite{Vohra:2018}. An
+iterative procedure was adopted to enhance the computational efficiency
 associated with both strategies.
@@
 The two strategies were further implemented to discover and compute the active
-subspace for the 33-dimensional problem. Both strategies revealed the existence of a 1-dimensional
-active subspace. The activity scores based on the dominant eigenvector and the total Sobol indices
-based on the 1-dimensional surrogate in the subspace, computed using the 
-perturbation-based strategy were found to be consistent with each other. Both sensitivity measures
- indicated that the uncertainty in the ignition delay is largely dependent on $A_1$, $A_9$, 
-$E_{a,1}$ and $E_{a,15}$, although contributions from $A_{15}$ and $A_{17}$ were also found to be significant.
+subspace for the 33-dimensional problem. 
+Both strategies revealed the existence
+of a 1-dimensional active subspace. The activity scores based on the dominant
+eigenvector and the total Sobol' indices based on the 1-dimensional surrogate in
+the subspace, computed using the perturbation-based strategy were found to be
+consistent with each other. Both sensitivity measures indicated that the
+uncertainty in the ignition delay is largely dependent on $A_1$, $A_9$,
+$E_{a,1}$ and $E_{a,15}$, although contributions from $A_{15}$ and $A_{17}$
+were also found to be significant.
 
-Computational gains using the proposed strategy in this work are essentially realized by using the
-1-dimensional surrogate in the active subspace that essentially casts a mutlivariate input in the full space
-into a univariate input in the reduced space. The surrogate can be used to accelerate forward propagation
-of the uncertainty as well as parameter estimation in a Bayesian setting. Hence, we focused our efforts to
-verify the accuracy of the 1-dimensional surrogates, obtained using both strategies.  
-The 1-dimensional surrogates were assessed for accuracy by first evaluating the relative L-2
-norm of the error between its predictions and model evaluations at $10^4$ independent MC samples in the
-33-dimensional input domain. The relative error based on predictions from the surrogate, computed
-using the regression-based strategy were found to be an order of magnitude larger than that estimated
+Computational gains using the proposed strategy in this work are essentially
+realized by using the 1-dimensional surrogate in the active subspace that
+essentially casts a mutlivariate input in the full space into a univariate
+input in the reduced space. The surrogate can be used to accelerate forward
+propagation of the uncertainty as well as parameter estimation in a Bayesian
+setting. Hence, we focused our efforts to verify the accuracy of the
+1-dimensional surrogates, obtained using both strategies.  The 1-dimensional
+surrogates were assessed for accuracy by first evaluating the relative $L^2$-norm
+of the error between its predictions and model evaluations at $10^4$
+independent MC samples in the 33-dimensional input domain. The relative error
+based on predictions from the surrogate, computed using the regression-based
+strategy were found to be an order of magnitude larger than that estimated
 using the surrogate from the perturbation-based strategy.
 
-Additionally, the surrogates were assessed by comparing probability distributions (based on kernel density estimation)
-for the ignition delay, constructed using surrogate predictions and model evaluations at the same set of 10$^4$
-random samples. The PDFs based on model evaluations and the 1-dimensional surrogate from the perturbation-based
-strategy were found to be almost identical. However, the PDF based on 1-dimensional surrogate from the
-regression-based strategy although captured the modal estimate with reasonable accuracy, was observed to
-under-estimate the uncertainty in the ignition delay. Mean estimates of the ignition delay using model evaluations
-and predictions using the two surrogates at 10$^4$ samples were found to be in agreement. However, as expected, the
-standard deviation, computed using predictions from the surrogate based on the regression-based strategy was
-found to be accurate only upto the first significant digit. Moreover, it was observed that the activity scores,
-computed using the regression-based strategy failed to capture the sensitivity towards the 
-most important parameter, $E_{a,15}$. 
-We attribute this discrepancy to two potential sources of numerical errors: (1) Regression-based approximation of the
-gradient of the model output, and (2) Scatter in the corresponding SSP (see Figure~\ref{fig:hd}) leading to numerical
-error incurred by the linear regression fit in this case. Hence, the regression-based approximation of the gradient 
-of the model output although reduces computational effort, the present investigations clearly reveal a potential pitfall associated
-with this approach. Therefore, this approach should only be preferred in situations involving intensive simulations, and 
-where the focus is essentially on estimating the statistics of the QoI as opposed to a more detailed analysis such as
-parametric sensitivity analysis.
+Additionally, the surrogates were assessed by comparing probability
+distributions for the ignition delay,
+constructed using surrogate predictions and model evaluations at the same set
+of 10$^4$ random samples. The PDFs based on model evaluations and the
+1-dimensional surrogate from the perturbation-based strategy were found to be
+almost identical. However, the PDF based on 1-dimensional surrogate from the
+regression-based strategy although captured the modal estimate with reasonable
+accuracy, was observed to underestimate the uncertainty in the ignition delay.
+Mean estimates of the ignition delay using model evaluations and predictions
+using the two surrogates at 10$^4$ samples were found to be in agreement.
+However, the standard deviation, computed using predictions from
+the surrogate based on the regression-based strategy was found to be accurate
+only upto the first significant digit. Moreover, it was observed that the
+activity scores, computed using the regression-based strategy failed to capture
+the sensitivity towards one of the most important parameters, $E_{a,15}$.  We attribute
+this discrepancy to two potential sources of numerical errors: (1)
+Regression-based approximation of the gradient of the model output, and (2)
+Scatter in the corresponding SSP (see Figure~\ref{fig:hd}) leading to numerical
+error incurred by the linear regression fit in this case. Hence, while the
+regression-based approach 
+reduces computational effort, the present investigations clearly reveal
+potential pitfalls associated with this approach. 
+%
+%More specifically, using
+%the regression-based approach, 
+%\begin{itemize} 
+%\item The desired statistics (e.g. mean and variance) of the
+%QoI might not be approximated to a desired level of accuracy.  
+%\item The
+%approach might fail to capture the sensitivity  of the QoI towards all inputs.
+%\end{itemize} 
+Therefore, in general perturbation-based approaches are preferred. The
+regression-based approach can be explored in situations involving intensive
+simulations where the perturbation-based methods are intractable, and the goal
+is obtaining rough estimates of the statistics of the QoI as opposed to a more
+detailed analysis such as those involving parametric sensitivity.  We also
+mention that alternate regression based approaches such as ones based on
+computing a global quadratic model have been proposed and used in the
+literature; see e.g.,~\cite{ConstantineDoostan17}. The applicability of such an
+approach in the context of high-dimensional chemical reaction networks is
+subject to future work.
 
- We would like to highlight that the computational framework presented in this work is agnostic to the choice
- of the chemical system and could thus be easily adapted for other systems as long as the quantity of interest
-is continuously differentiable in the considered domain of the inputs.
-We have demonstrated that the active subspace could be exploited for a forward propagation of
- the uncertainty from inputs to the output. However, the activity scores and the low-dimensional surrogate could further guide optimal allocation of computational resources for  
+The computational framework presented in this work is agnostic to the choice of
+the chemical system and can be easily adapted for other systems as long
+as the quantity of interest is continuously differentiable in the considered
+domain of the inputs.  We have demonstrated that the active subspace could be
+exploited for efficient forward propagation of the uncertainty from inputs to
+the output. However, the activity scores and the low-dimensional surrogate
+could further guide optimal allocation of computational resources for
 calibration of the important rate-controlling parameters in a Bayesian setting.
-Additionally, dimension reduction using active subspaces could assist in developing robust formulations
-for predicting discrepancy between simulations and measurements due to epistemic uncertainty in the model inputs.
+Additionally, dimension reduction using active subspaces could assist in
+developing robust formulations for predicting discrepancy between simulations
+and measurements due to epistemic uncertainty in the model inputs.

### 2018-09-24 Paper/conc.tex
..

@@
 pre-exponent and an activation energy. Hence, in theory, the evolution of the
-chemical system depends upon 38 inputs. 
-\alennote{Maybe say something about how we went from 38 to 33?} 
+chemical system depends upon 38 inputs. In this work, we have considered 
+epistemic uncertainty in all pre-exponents and activation energies with non-zero
+nominal values. Hence, the input space dimensionality is 33 instead of 38.  
 %Conventional means for uncertainty quantification such as those involving
@@
 computing a global quadratic model have been proposed and used in the
-literature; see e.g.,~\cite{ConstantineDoostan17}. The applicability of such an
+literature; see e.g.,~\cite{Constantine:2017a}. The applicability of such an
 approach in the context of high-dimensional chemical reaction networks is

### 2018-09-24 Paper/intro.tex
..

@@
 
-In this work, we implement two strategies for computing the active subspace for the
-H$_2$/O$_2$ mechanism, and refer to them as \textit{perturbation-based} and 
-\textit{regression-based} to highlight the difference in the approach for estimating
-the gradient of the model output as discussed.  
-Other than the difference in gradient estimation, the two strategies follow the same
-sequence of steps provided in~\cite{Constantine:2015}. However, 
-in order to reduce computational effort, the two strategies are implemented in an
-iterative manner in this work as discussed later in Section~\ref{sec:method}.
+In this work, we adopt an iterative strategy to reduce the computational effort
+associated with active subspace computation. Moreover, we explore two approaches
+for estimating the gradient of the ignition delay with respect to the uncertain
+rate-controlling parameters: pre-exponent, $A_i$ and the activation energy, $E_{a,i}$,
+for $i^{\text{th}}$ reaction in the H$_2$/O$_2$ mechanism. As mentioned above, the
+first approach uses finite difference to estimate the gradient and is regarded as
+the perturbation-based approach, whereas, the second approach involves a linear
+regression-fit to the available set of model evaluations and is regarded as
+the regression-based approach.
 
-An alternate approach to dimension reduction involves computing the global
+An alternate strategy to dimension reduction involves computing the global
 sensitivity measures associated with the uncertain inputs of a model. Depending
@@
 We investigate the suitability of both approaches
-for computing active subspaces, surrogate modeling, and approximating global
+for estimating the gradient of ignition delay in the H$_2$/O$_2$ mechanism.
+Specifically, we compare resulting
+active subspaces, surrogate models, and the ability to approximate global
 sensitivity measures through a comprehensive set of numerical experiments. Our
@@
 Numerical results based on the perturbation-based approach are compared with those
-obtained using the regression-based approach.  For our initial investigation aimed
-as assessing the feasibility and suitability of the two strategies, the active 
-subspace is computed for a 19-dimensional 
-H$_2$/O$_2$ reaction kinetics problem whereby only the $A_i$'s are
+obtained using the regression-based approach.  The active 
+subspace is initially computed for a 19-dimensional 
+H$_2$/O$_2$ reaction kinetics problem wherein only the $A_i$'s are
 considered as uncertain. In this case, the two approaches are observed to yield
-consistent results. Motivated by these results, we further implement the two
-strategies to compute the active subspace for a 33-dimensional H$_2$/O$_2$
-reaction kinetics problem in section~\ref{sec:app}. Once again, the two
-strategies reveal the existence of a 1-dimensional active subspace. However,
-it is observed that although the regression-based strategy results in a 
-reasonable approximation of the mean, mode, and the uncertainty in the ignition
-delay, it fails to capture the global sensitivity of all the uncertain inputs. 
+consistent results. We further compute the active  for a 33-dimensional H$_2$/O$_2$
+reaction kinetics problem in section~\ref{sec:app}. Once again, 
+active subspace is found to be 1-dimensional. However, in this case,
+it is observed that although the regression-based estimation of the gradient results in a 
+reasonable approximation of the mean and mode, the variability in the ignition delay
+due to uncertain rate-controlling parameters is under-estimated. Additionally,  
+the sensitivity towards one of the most important parameter is not captured.
 Hence, a trade-off between efficiency and accuracy between the two approaches is
-apparent in this case.  Finally, a summary and discussion based on our findings is included in
+apparent in the higher-dimensional problem. 
+Finally, a summary and discussion based on our findings is included in
 section~\ref{sec:conc}.

### 2018-09-24 Paper/method.tex
..

@@
 estimate provided in~\cite{Yetter:1991}.
-Specifically, we employed two strategies: 
-a perturbation-based strategy that involves computation of model
+Two approaches are explored for estimating the gradient of ignition delay with
+respect to uncertain $A_i$'s: 
+a perturbation-based approach that involves computation of model
 gradients using finite difference in order to construct the matrix $\hat{\mat{C}}$
 in~\eqref{eq:chat}, and a regression-based approach 
-that involves a local linear approximation of
-the model output to approximate the gradient. 
-The two approaches are based on Algorithms~1.1 and 1.2 in~\cite{Constantine:2015} and
-essentially differ in the way the gradient is estimated. They were implemented in an
-iterative manner to ensure computational efficiency by avoiding 
-unnecessary model evaluations once a converged active subspace
-was achieved.
+that involves a linear regression fit to the available set of
+model evaluations in order to approximate the gradient. 
+The active subspace is computed in an iterative manner to avoid
+unnecessary model evaluations once converged is established.
 
@@
 
-As discussed earlier, gradient estimation using finite differences in the 
-perturbation-based strategy requires additional model evaluations at the
+As discussed earlier, gradient estimation using finite differences 
+requires additional model evaluations at the
 neighboring points in the input domain. 
@@
 fits to subsets of available evaluations as discussed 
-in~\cite{Constantine:2015}. Hence, the computational effort is  
+in~\cite{Constantine:2015} (see Algorithm 1.2). Hence, the computational effort is  
 reduced by a factor $(d+1)$ when using the regression-based approach.
@@
 We begin by evaluating the gradient of the model output, $\nabla_{\bm{\xi}}f$,
-at an initial set of $n_0$ samples denoted by $\bm{\xi}_i$, $i$ = $1,\ldots,n_0$.
+at an initial set of $n_0$ samples (generated using Monte Carlo sampling)
+denoted by $\bm{\xi}_i$, $i$ = $1,\ldots,n_0$.
 Using the gradient
@@
 In Figure~\ref{fig:eig_comp}, we illustrate the comparison of
-the resulting normalized eigenvalue spectrum ($\lambda_1, \ldots, 
-\lambda_{19}$) with the same using a much 
+the resulting normalized eigenvalue spectrum ($\lambda_i/\lambda_0,~i = 1,\ldots,19$)
+with the same using a much 
 smaller set of samples, $n$~=~$\{20,40,80,120\}$.
@@
 %   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1}
-\caption{A comparison of the normalized eigenvalue spectrum using $n$ = $\{20,40,80,120\}$ samples with that
+\caption{A comparison of the normalized eigenvalue spectrum ($\lambda_i/\lambda_0$)
+using $n$~=~$\{20,40,80,120\}$ samples with that
 obtained using a much larger sample size, $N$~=~1000. 
@@
 %the square of individual components ($w_i$'s) of the dominant eigenvector evaluated using 
-$N$=1000 and $n$ = $\{20,40,80,120\}$ samples.
+%$N$=1000 and $n$~=~$\{20,40,80,120\}$ samples.
 } 

### 2018-09-24 Paper/method.tex
..

@@
 
-The active subspace for the 19-dimensional problem was also computed using the regression-based
-approach. As discussed earlier, a linear regression fit to the model output was used to
-approximate the gradient using the sequence of steps outlined in Algorithm 1.2 
-in~\cite{Constantine:2015}. 
+The active subspace for the 19-dimensional problem was also computed using regression-based
+estimates of the gradient that does not require model evaluations at neighboring points as
+discussed earlier.
 %
@@
 In Figure~\ref{fig:comp} (left),
-we provide a comparison of the square of individual components of the dominant eigenvector, obtained using the two 
-strategies. Additionally, the quantity, $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ is also plotted in
-Figure~\ref{fig:comp} (right) to illustrate convergence characteristics for the regression-based strategy.
+we provide a comparison of the square of individual components of the dominant eigenvector, obtained using both, perturbation-based and regression-based estimates of the gradient of the ignition 
+delay. Additionally, the quantity, $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ is plotted in
+Figure~\ref{fig:comp} (right) to illustrate convergence characteristics of Algorithm~\ref{alg:grad}.
 %
@@
 %
-Individual components of the dominant eigenvector from the two strategies are observed to be in excellent
+Individual components of the dominant eigenvector from the two approaches are observed to be in excellent
 agreement with each other. Moreover, as expected,  $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ is observed 
-to decrease with iterations, and required model evaluations at only 30 samples to converge using $\tau$~=~0.1.
+to decrease with iterations, and requires model evaluations at only 30 samples to converge using $\tau$~=~0.1.
 
@@
 the subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, 
-regarded as the \textit{sufficient summary plot} (SSP), obtained using the two strategies is 
+regarded as the \textit{sufficient summary plot} (SSP), obtained using the perturbation-based and
+regression-based gradient estimates are
 compared in Figure~\ref{fig:comp_ssp}.
@@
 
-We extend the comparative assessment of the two approaches by estimating
+We further estimate
 the normalized activity scores for individual uncertain inputs ($\tilde{\nu}_{i,r}$; $r$=1 since the
@@
 The activity scores for the 19 uncertain pre-exponents ($A_i$'s), estimated
-using the gradient-based and gradient-free approaches are plotted in
+using the perturbation-based and regression-based approaches are plotted in
 Figure~\ref{fig:comp_as} (left).  Additionally, for the purpose of
@@
 %
-The activity scores estimated using the two approaches agree favorably with each other as well as those
+The activity scores based on the two approaches for gradient estimation agree favorably 
+with each other as well as those
 based on the screening metric involving the DGSMs from~\cite{Vohra:2018}. It is observed that the uncertainty
@@
 
-The above comparisons indicate that the two strategies yield consistent results for the 19-dimensional
-H$_2$/O$_2$ reaction kinetics problem. In other words, the gradient estimates using pertubation and
-regression are reasonably close, and therefore, using the regression-based strategy being computationally
-advantageous, would be preferred in this case. In the following section, we shift our focus to 
+The above comparisons indicate that the gradient of the ignition delay with respect to the
+uncertain $A_i$'s is reasonably approximated using both, perturbation-based and regression-based
+approaches in this case. Since the number of model evaluations required in the case of
+regression are much fewer, this approach seems favorable for the 19-dimensional problem. 
+In the following section, we shift our focus to 
 the higher-dimensional H$_2$/O$_2$ reaction kinetics application involving 33 uncertain inputs.

### 2018-09-24 Paper/conc.tex
..

@@
 pre-exponent and an activation energy. Hence, in theory, the evolution of the
-chemical system depends upon 38 inputs. In this work, we have considered 
+chemical system depends upon 38 inputs. However, we considered 
 epistemic uncertainty in all pre-exponents and activation energies with non-zero
-nominal values. Hence, the input space dimensionality is 33 instead of 38.  
+nominal values i.e. a total of 33 parameters instead of 38.  
 %Conventional means for uncertainty quantification such as those involving
@@
 
-Two computational strategies were explored to compute the active subspace for
-the H$_2$/O$_2$ kinetics application in this work. The first strategy, referred
-to as perturbation-based, used finite differences to estimate the gradient of the
-model output with respect to the uncertain inputs. The second strategy, referred
-to as regression-based, involved gradient estimation by means of approximating
-the available set of model evaluations using a linear regression fit.  The two
-strategies were shown to yield consistent results for the 19-dimensional
-problem whereby only the pre-exponents were considered to be uncertain.
+Active subspace computation requires repeated evaluations of the gradient of
+the QoI i.e. the ignition delay. For this purpose, we explored two approaches,
+namely, perturbation-based (e.g. finite difference - used in this work, adjoint-based
+methods, etc.) and regression-based. Both approaches were shown to yield consistent
+results for the 19-dimensional
+problem wherein only the pre-exponents were considered to be uncertain.
 Additionally, the activity scores were also shown to be consistent with the
 screening metric estimates based on DGSMs in~\cite{Vohra:2018}. An
-iterative procedure was adopted to enhance the computational efficiency
-associated with both strategies.
+iterative procedure was adopted to enhance the computational efficiency.
 
-The two strategies were further implemented to discover and compute the active
-subspace for the 33-dimensional problem. 
-Both strategies revealed the existence
-of a 1-dimensional active subspace. The activity scores based on the dominant
-eigenvector and the total Sobol' indices based on the 1-dimensional surrogate in
-the subspace, computed using the perturbation-based strategy were found to be
-consistent with each other. Both sensitivity measures indicated that the
-uncertainty in the ignition delay is largely dependent on $A_1$, $A_9$,
-$E_{a,1}$ and $E_{a,15}$, although contributions from $A_{15}$ and $A_{17}$
-were also found to be significant.
-
-Computational gains using the proposed strategy in this work are essentially
-realized by using the 1-dimensional surrogate in the active subspace that
-essentially casts a mutlivariate input in the full space into a univariate
-input in the reduced space. The surrogate can be used to accelerate forward
-propagation of the uncertainty as well as parameter estimation in a Bayesian
-setting. Hence, we focused our efforts to verify the accuracy of the
-1-dimensional surrogates, obtained using both strategies.  The 1-dimensional
-surrogates were assessed for accuracy by first evaluating the relative $L^2$-norm
-of the error between its predictions and model evaluations at $10^4$
-independent MC samples in the 33-dimensional input domain. The relative error
-based on predictions from the surrogate, computed using the regression-based
-strategy were found to be an order of magnitude larger than that estimated
-using the surrogate from the perturbation-based strategy.
-
-Additionally, the surrogates were assessed by comparing probability
-distributions for the ignition delay,
-constructed using surrogate predictions and model evaluations at the same set
-of 10$^4$ random samples. The PDFs based on model evaluations and the
-1-dimensional surrogate from the perturbation-based strategy were found to be
-almost identical. However, the PDF based on 1-dimensional surrogate from the
-regression-based strategy although captured the modal estimate with reasonable
-accuracy, was observed to underestimate the uncertainty in the ignition delay.
-Mean estimates of the ignition delay using model evaluations and predictions
-using the two surrogates at 10$^4$ samples were found to be in agreement.
-However, the standard deviation, computed using predictions from
-the surrogate based on the regression-based strategy was found to be accurate
-only upto the first significant digit. Moreover, it was observed that the
-activity scores, computed using the regression-based strategy failed to capture
-the sensitivity towards one of the most important parameters, $E_{a,15}$.  We attribute
+The active subspace was further computed for a 33-dimensional problem
+wherein all pre-exponents and activation energies with non-zero nominal
+estimates were considered uncertain. Once again, the active subspace
+was found to be 1-dimensional. However, unlike the 19-dimensional case,
+gradient estimation using the regression-based approach resulted in an
+under-estimation of the variability in the ignition delay and also failed
+to capture the sensitivity towards one of the most important parameters, $E_{a,15}$.
+We attribute
 this discrepancy to two potential sources of numerical errors: (1)
@@
 Scatter in the corresponding SSP (see Figure~\ref{fig:hd}) leading to numerical
-error incurred by the linear regression fit in this case. Hence, while the
+error incurred by the linear regression fit in this case. 
+Global sensitivity estimates using the perturbation-based
+approximation of the gradient indicated that the variability in the ignition
+delay is  largely dependent on $A_1$, $A_9$,
+$E_{a,1}$ and $E_{a,15}$, although contributions from $A_{15}$ and $A_{17}$
+were also found to be significant. 
+Hence, while the
 regression-based approach 
@@
 potential pitfalls associated with this approach.
+
+%Computational gains using the proposed framework in this work are essentially
+%realized by using the 1-dimensional surrogate in the active subspace that
+%essentially casts a mutlivariate input in the full space into a univariate
+%input in the reduced space. The surrogate can be used to accelerate forward
+%propagation of the uncertainty as well as parameter estimation in a Bayesian
+%setting. Hence, we focused our efforts to verify the accuracy of the
+%1-dimensional surrogates, obtained using both strategies.  The 1-dimensional
+%surrogates were assessed for accuracy by first evaluating the relative $L^2$-norm
+%of the error between its predictions and model evaluations at $10^4$
+%independent MC samples in the 33-dimensional input domain. The relative error
+%computed using predictions from the surrogate resulting from the regression-based
+%approach were found to be an order of magnitude larger than that estimated
+%using surrogate resulting from the perturbation-based approximation of the gradient.
+
+%Additionally, the surrogates were assessed by comparing probability
+%distributions for the ignition delay,
+%constructed using surrogate predictions and model evaluations at the same set
+%of 10$^4$ random samples. The PDFs based on model evaluations and the
+%1-dimensional surrogate from the perturbation-based approach were found to be
+%almost identical. However, the PDF based on 1-dimensional surrogate from the
+%regression-based approach, although captured the modal estimate with reasonable
+%accuracy, was observed to underestimate the uncertainty in the ignition delay.
+%Mean estimates of the ignition delay using model evaluations and predictions
+%using the two surrogates at 10$^4$ samples were found to be in agreement.
+%However, the standard deviation, computed using predictions from
+%the surrogate resulting from the regression-based approach was found to be accurate
+%only upto the first significant digit. Moreover, it was observed that the
+%activity scores, computed using the regression-based strategy failed to capture
+%the sensitivity towards one of the most important parameters, $E_{a,15}$.  
 %
@@
 %\end{itemize} 
-Therefore, in general perturbation-based approaches are preferred. The
+Therefore, in general, perturbation-based approaches should be preferred. The
 regression-based approach can be explored in situations involving intensive

### 2018-09-24 Paper/assa.tex
..

@@
 
-The relations recored in the following lemma will be useful in the discussion 
+The relations recorded in the following lemma will be useful in the discussion 
 that follows.
@@
 presented result covers a broad class of input distributions coming from 
-Boltzmann family of distributions.
+the Boltzmann family of distributions.
 Additionally, the normalized activity scores discussed above provide practical

### 2018-09-24 Paper/abstract.tex
abstract ...

--- 2018-08-28 Paper/abstract.tex
@@
 \section*{Abstract}
-Time evolution of species' concentration in a chemical system is tightly coupled
-with the underlying mechanism and the governing rate laws for individual reactions.
-Oftentimes, the rate-controlling parameters cannot be specified precisely 
-due to sparse calibration data and lack of knowledge about them, regarded as
-the \textit{epistemic} uncertainty. Computer simulations for such systems must
-therefore account for the presence of the epistemic uncertainty in these parameters
-when making predictions. However, complex mechanisms typically involve a large number of
-reactions and consequently, a large set of uncertain rate parameters. Propagation of
-the uncertainty from parameters to predictions can be remarkably challenging for
-such high-dimensional problems. In this study, we focus on enabling an efficient
-approach for uncertainty propagation from inputs to the model output,
-through dimension reduction by computing the
-so-called
-\textit{active subspace}. The active subspace
- predominantly captures the variability in the quantity of
-interest (QoI) in terms of fewer independent variables. 
-It is computed using an iterative, gradient-based
-approach and a computationally less intensive gradient-free approach for the 
-H$_2$/O$_2$ mechanism involving 19 reactions. An initial investigation considers
-only the 19 pre-exponents of the Arrhenius rate laws as uncertain. In this case, both
-approaches yield consistent results wherein the active subspace is found to be 
-1-dimensional. The two approaches are further implemented to a high-dimensional
-problem involving 33 uncertain rate-parameters: pre-exponent and the activation
-energy. Once again, a 1-dimensional active subspace is obtained, indicating
-enormous scope for dimension reduction in such applications. However, the
-gradient-free approach though reasonably captured the mean, mode, and the uncertainty
-in the QoI; failed to capture the sensitivity towards the most
-important parameter according to the gradient-based analysis for the high-dimensional
-application.
+We focus on efficient quantification of uncertainty in complex chemical
+reaction networks with a large number of uncertain parameters.  Parameter
+dimension reduction is accomplished by computing an~\emph{active subspace} that
+predominantly captures the variability in the quantity of interest (QoI).  In
+this study, we compute the active subspace for the H$_2$/O$_2$ mechanism that
+involves 19 chemical reactions, using an efficient iterative strategy.  The
+active subspace is first computed for a 19-parameter problem wherein only the
+uncertainty in the pre-exponents of the individual reaction rates is
+considered. This is followed by the analysis of a 33-dimensional case wherein
+the activation energies are also considered uncertain.  In both cases, a
+one-dimensional active subspace is identified, which indicates
+enormous potential for efficient statistical analysis of complex 
+chemical systems. In addition, we explore links between active subspaces and
+global sensitivity analysis measures, and exploit these links for 
+indentification of key contributors to the variability in the model response.
+
+%In this case, the two approaches are observed to yield consistent results.
+%However, when computing the active subspace for a 33-dimensional problem
+%involving uncertain activation energies in addition to the pre-exponents, a
+%trade-off between accuracy and effort is observed.
+
+%Specifically, when estimating the gradient using
+%the regression-based approach, the variability of the QoI (in the subspace),
+%due to uncertainty in the rate parameters, is under-estimated. Moreover, the
+%sensitivity towards one of the most important parameter is not captured. Hence,
+%we carefully recommend using the regression-based approach for estimating the
+%gradient only in situations involving intensive computations of the model 
+%(e.g. complex mechanisms involving a large number of reactions and species), and where the
+%goal is to obtain rough estimates of the statistics of the QoI.

### 2018-09-24 Paper/old_abstract.tex
abstract ...

\section*{Abstract}
Time evolution of species' concentration in a chemical system is tightly coupled
with the underlying mechanism and the governing rate laws for individual reactions.
Often, the rate-controlling parameters cannot be specified precisely 
due to \textit{epistemic} uncertainty, i.e., lack of knowledge and 
sparse calibration data. Computer simulations of such systems must
therefore account for the epistemic uncertainty regarding these parameters
when making predictions. However, complex mechanisms typically involve a large number of
reactions and consequently, a large set of uncertain rate parameters. Propagation of
the uncertainty from parameters to predictions can be remarkably challenging for
such high-dimensional problems. In this study, we focus on developing an efficient
approach for uncertainty propagation from inputs to the model output,
through dimension reduction by computing the
so-called
\textit{active subspace}~\cite{Constantine:2015}. The active subspace
 predominantly captures the variability in the quantity of
interest (QoI) in terms of fewer independent variables. 
However, its computation requires gradient estimation of
the model output with respect to its inputs at multiple points in the
input domain, and can therefore be
remarkably challenging. In this study, we compute the active subspace
for the H$_2$/O$_2$ mechanism that involves 19 chemical reactions,
using an iterative strategy that is expected to be relatively more efficient
than the conventional approach. 
The gradient is estimated using finite difference, regarded as a perturbation-based
approach, and using a linear approximation to the available set of model
evaluations, regarded as a regression-based approach. Since finite difference
requires model evaluations at neighboring points, the perturbation-based approach
is expected to be computationally expensive and relatively more accurate.
The active subspace is first
computed for a 19-dimensional problem wherein only the uncertainty in the pre-exponents 
is considered. In this case, the two approaches are observed to yield consistent results.
However, when computing the active subspace for a 33-dimensional problem involving
uncertain activation energies in addition to the pre-exponents, a trade-off between
accuracy and effort is observed. Specifically, when estimating the gradient using
the regression-based approach, the variability of the QoI (in the subspace),
due to uncertainty in the rate parameters, is under-estimated. Moreover, the
sensitivity towards one of the most important parameter is not captured. Hence,
we carefully recommend using the regression-based approach for estimating the
gradient only in situations involving intensive computations of the model 
(e.g. complex mechanisms involving a large number of reactions and species), and where the
goal is to obtain rough estimates of the statistics of the QoI.

### 2018-09-24 Paper/abstract.tex
update ...

@@
 predominantly captures the variability in the quantity of interest (QoI).  In
-this study, we compute the active subspace for the H$_2$/O$_2$ mechanism that
-involves 19 chemical reactions, using an efficient iterative strategy.  The
+the present work, we compute the active subspace for a H$_2$/O$_2$ mechanism 
+that involves 19 chemical reactions, using an efficient iterative strategy.  The
 active subspace is first computed for a 19-parameter problem wherein only the
@@
 the activation energies are also considered uncertain.  In both cases, a
-one-dimensional active subspace is identified, which indicates
-enormous potential for efficient statistical analysis of complex 
-chemical systems. In addition, we explore links between active subspaces and
-global sensitivity analysis measures, and exploit these links for 
-indentification of key contributors to the variability in the model response.
-
-%In this case, the two approaches are observed to yield consistent results.
-%However, when computing the active subspace for a 33-dimensional problem
-%involving uncertain activation energies in addition to the pre-exponents, a
-%trade-off between accuracy and effort is observed.
-
-%Specifically, when estimating the gradient using
-%the regression-based approach, the variability of the QoI (in the subspace),
-%due to uncertainty in the rate parameters, is under-estimated. Moreover, the
-%sensitivity towards one of the most important parameter is not captured. Hence,
-%we carefully recommend using the regression-based approach for estimating the
-%gradient only in situations involving intensive computations of the model 
-%(e.g. complex mechanisms involving a large number of reactions and species), and where the
-%goal is to obtain rough estimates of the statistics of the QoI.
+one-dimensional active subspace is identified, which indicates enormous
+potential for efficient statistical analysis of complex chemical systems. In
+addition, we explore links between active subspaces and global sensitivity
+analysis, and exploit these links for identification of key
+contributors to the variability in the model response.

### 2018-09-24 Paper/conc.tex
update ...

@@
 %\end{itemize} 
-Therefore, in general, perturbation-based approaches should be preferred. The
-regression-based approach can be explored in situations involving intensive
-simulations where the perturbation-based methods are intractable, and the goal
-is obtaining rough estimates of the statistics of the QoI as opposed to a more
-detailed analysis such as those involving parametric sensitivity.  We also
-mention that alternate regression based approaches such as ones based on
-computing a global quadratic model have been proposed and used in the
+Therefore, in general, perturbation-based approaches should be preferred for 
+active subspace computation.
+The regression-based approach can be explored in situations involving intensive
+simulations where the perturbation-based methods are intractable.
+%
+%and the goal is obtaining rough estimates of the statistics of the QoI as
+%opposed to a more detailed analysis such as those involving parametric
+%sensitivity.  
+%
+We also mention that alternate regression based approaches such as ones based
+on computing a global quadratic model have been proposed and used in the
 literature; see e.g.,~\cite{Constantine:2017a}.  The applicability of such an
@@
 exploited for efficient forward propagation of the uncertainty from inputs to
-the output. However, the activity scores and the low-dimensional surrogate
+the output. Moreover, the activity scores and the low-dimensional surrogate
 could further guide optimal allocation of computational resources for

### 2018-09-25 Paper/abstract.tex
..

@@
 \section*{Abstract}
-We focus on efficient quantification of uncertainty in complex chemical
+We focus on an efficient approach for quantification of uncertainty in complex chemical
 reaction networks with a large number of uncertain parameters.  Parameter
@@
 the activation energies are also considered uncertain.  In both cases, a
-one-dimensional active subspace is identified, which indicates enormous
+1-dimensional active subspace is identified, which indicates enormous
 potential for efficient statistical analysis of complex chemical systems. In

### 2018-09-25 Paper/conc.tex
..

@@
 reducing the dimensionality of the problem by identifying important directions
-in the parameter space such that the variability in the model output is
-predominantly captured along these directions. This was done through active
-subspace construction.  Additionally, we demonstrated that the activity scores,
+in the parameter space such that the model output 
+predominantly varies along these directions. These important directions
+constitute the active subspace. Additionally, we demonstrated that the activity scores,
 computed using the components of the dominant eigenvectors provide an efficient
@@
 exploited for efficient forward propagation of the uncertainty from inputs to
-the output. Moreover, the activity scores and the low-dimensional surrogate
+the output. The resulting activity scores and the low-dimensional surrogate
 could further guide optimal allocation of computational resources for

### 2018-09-25 Paper/method.tex
..

@@
 
-As mentioned earlier, the model output $f(\bm{\xi})$ i.e. the igntion delay in the
+As mentioned earlier, the model output $f(\bm{\xi})$ i.e. the ignition delay in the
 H$_2$/O$_2$ reaction in this case, varies
@@
 using the perturbation-based and regression-based approaches are plotted in
-Figure~\ref{fig:comp_as} (left).  Additionally, for the purpose of
-verification, we compare these estimates with those obtained using the
-normalized DGSMs in~\eqref{eq:ndgsm}, as reported in~\cite{Vohra:2018}, in
-Figure~\ref{fig:comp_as} (right). 
+Figure~\ref{fig:comp_as}.  
 %
@@
   \includegraphics[width=0.45\textwidth]{./Figures/comp_as}
-  \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}
+%  \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}
 \caption{Left: A bar-graph of normalized activity scores ($\tilde{\nu}_{i,r}$'s) 
 for the 19 uncertain pre-exponents ($A_i$'s); $r$
-denotes the number of eigenvectors in the active subspace.
-Right: A plot of the screening metric ($\tilde{\nu}_i$) for individual $A_i$'s 
-(adapted from~\cite{Vohra:2018}).}
+denotes the number of eigenvectors in the active subspace.}
 \label{fig:comp_as}
@@
 with each other as well as those
-based on the screening metric involving the DGSMs from~\cite{Vohra:2018}. It is observed that the uncertainty
+based on the screening metric involving the DGSMs in~\cite{Vohra:2018}. It is observed that the uncertainty
 associated with the ignition delay is largely due to the uncertainty in $A_9$ and $A_1$. Sensitivity towards

### 2018-09-28 Paper/intro.tex
..

@@
 inadequate data, parametric uncertainty, and model errors~\cite{Vohra:2014, 
-Vohra:2017, Morrison:2018, Hantouche:2018, Nannapaneni:2016, Sankararaman:2015}.
+Vohra:2017, Morrison:2018, Hantouche:2018, Nannapaneni:2016, Sankararaman:2015,
+Reagana:2003}.
 

### 2018-09-28 Paper/intro.tex
..

@@
 include finite differences and more advanced methods involving
-adjoints~\cite{Jameson:1988,Gunzburger:2003,Borzi:2011,Alexanderian:2017}. The
+adjoints~\cite{Jameson:1988,Borzi:2011,Alexanderian:2017}. The
 adjoint-based method requires a solution of the state equation (forward solve)

### 2018-11-21 Revision/Paper/app.tex
..

--- 2018-09-22 Paper/app.tex
@@
 is 1-dimensional. This is further confirmed by the SSP plots in Figure~\ref{fig:hd} (right), generated
-using both, perturbation-based and regression-based strategies. 
+using both, perturbation-based and regression-based approaches for estimating
+the gradient. 
 %
@@
 Specifically
-in the case of perturbation-based strategy, the SSP is observed to exhibit a linear variation that
+in the case of perturbation-based approach, the SSP is observed to exhibit a linear variation that
 is reasonably
@@
 outlined in Section~\ref{sub:ac}.
-On the other hand, the SSP constructed using the regression-based strategy
+On the other hand, the SSP obtained by estimating the gradient using regression,
 is relatively more scattered. Consequently, the two surrogates (straight-line fits) are observed to exhibit a
@@
 It was estimated to be 1.35$\times10^{-2}$ and 1.02$\times10^{-1}$
-using $\tilde{G}$ based on the perturbation-based and the regression-based strategies respectively. Hence,
-$\varepsilon_d$ was found to increase by an order of magnitude when using the regression-based strategy.
+using $\tilde{G}$ based on the perturbation-based and the regression-based approaches respectively. Hence,
+$\varepsilon_d$ was found to increase by an order of magnitude when the gradient is estimated using
+the regression-based approach.
 Therefore, in the norm-sense, it can be said that the 1-dimensional surrogate in the active subspace, obtained
-using the perturbation-based approach is remarkably accurate with a relative error of less than 1.5$\%$, whereas
+using the perturbation-based approach is remarkably accurate with a relative error of less than 1.5$\%$, whereas,
 the regression-based approach resulted in a relative error of about 10$\%$.
@@
 probability density functions (PDFs) obtained using the true set of model evaluations, and 1-dimensional
-surrogates ($\tilde{G}$'s) from the two strategies, in Figure~\ref{fig:pdf_33D}. Note that the three PDFs were evaluated 
-using the same set of 10$^4$ samples in the cross-validation set. 
+surrogates ($\tilde{G}$'s) based on the two approaches for gradient estimation as shown in Figure~\ref{fig:pdf_33D}.
+Note that the three PDFs were evaluated using the same set of 10$^4$ samples in the cross-validation set. 
 %
@@
 %
-The PDFs based on the model evaluations and the perturbation-based strategy are nearly identical, confirming
-the existence of a 1-dimensional active subspace as well as the accuracy of this strategy. 
-On the other hand, although the PDF obtained using the regression-based strategy is observed to show
+The PDFs based on the model evaluations and the perturbation-based approach are nearly identical, confirming
+the existence of a 1-dimensional active subspace as well as the accuracy of gradient estimates in this case. 
+On the other hand, although the PDF obtained using the regression-based approach  is observed to show
 consistency with the true PDF in the modal
@@
 The three PDFs are found to be
-consistent in their prediction of mean value of the ignition delay.  The standard deviation estimates based on model
-evaluations and the perturbation-based strategy are found to be in close agreement, whereas, it  is accurately
- estimated in the case of regression-based strategy only upto the first significant digit.
+consistent in their prediction of mean value of the ignition delay.  The standard deviation estimates using model
+evaluations and the perturbation-based approach are found to be in close agreement, whereas, it  is accurately
+ estimated in the case of regression-based approach only upto the first significant digit.
 
@@
 The normalized activity scores ($\tilde{\nu}_{i,r}$) based on the dominant eigenspace, obtained
-using the two strategies (perturbation-based and regression-based) are compared with the 
+using the two approaches for gradient estimation (perturbation-based and regression-based),
+are compared with the 
 total Sobol' indices
 in Figure~\ref{fig:as_33D}. Note that the Sobol' indices were computed using the verified
-1-dimensional surrogate ($\tilde{G}$) in the active subspace, computed using the 
-perturbation-based strategy. 
+1-dimensional surrogate ($\tilde{G}$) in the active subspace, obtained using the 
+perturbation-based approach. 
 %
@@
  \begin{center}
-  \includegraphics[width=0.8\textwidth]{./Figures/as_33D_new}
+  \includegraphics[width=0.8\textwidth]{./Figures/as_33D_new2}
 \caption{A bar-graph illustrating individual activity scores for the uncertain $A_i$' (left) and $E_{a,i}$'s (right).}
@@
 consistent
-with the normalized activity scores evaluated using the perturbation-based strategy. The regression-based
-strategy yields consistent results for the $A_i$'s, but it does not capture the sensitivity associated with one of the most important parameters, $E_{a,15}$.
-This is because numerical errors are incurred when approximating
+with the normalized activity scores as obtained when the gradient is estimated using a perturbation-based approach.
+The regression-based
+approach, however, yields consistent results for the $A_i$'s but fails capture the sensitivity associated with one
+of the most important parameters, $E_{a,15}$. This is because numerical errors are incurred when approximating
 the model gradients using a linear regression fit in this case. 

### 2018-11-21 Revision/Paper/assa.tex
..

--- 2018-09-24 Paper/assa.tex
@@
 are statistically independent. While extensions to the cases 
-of correlated inputs exist~\cite{Borgonovo:2007,Li:2010,Jacques:2006,Xu:2007,Hart:2017},
+of correlated inputs exist~\cite{Borgonovo:2007,Li:2010,Jacques:2006,Xu:2007},
  we limit the discussion in this section to the

### 2018-11-21 Revision/Paper/method.tex
..

--- 2018-09-25 Paper/method.tex
@@
 active subspace. The framework is employed to the H$_2$/O$_2$ reaction kinetics
-problem whereby the 
+problem whereby the \rebut{logarithm of the} 
 pre-exponent ($A_i$) in the rate law associated with individual reactions provided in
 Table~\ref{tab:kinetics} is considered to be uniformly
-distributed in the interval, $[0.9A_i^\ast, 1.1A_i^\ast]$; $A_i^\ast$ is the nominal
+distributed in the interval, \rebut{$[0.97\log(A_i^\ast), 1.03\log(A_i^\ast)]$};
+$A_i^\ast$ is the nominal
 estimate provided in~\cite{Yetter:1991}.
 Two approaches are explored for estimating the gradient of ignition delay with
-respect to uncertain $A_i$'s: 
+respect to \rebut{$\log(A_i)$}: 
 a perturbation-based approach that involves computation of model
@@
 subsequent eigenvalues,
-$\left(\frac{\lambda_j}{\lambda_{j+1}}\right)>\mathcal{O}(10^2)$.
+\rebut{$\left(\frac{\lambda_j}{\lambda_{j+1}}\right)\ge\mathcal{O}(10^1)$}.
  At each subsequent iteration, model evaluations are
@@
 terminated and the resulting eigenspace is considered to have converged once
-the maximum relative change at iteration $k$, $\delta \hat{\mat{W}}_{1,j}^{(k)}$
+the maximum relative change at iteration $k$, $\max(\delta \hat{\mat{W}}_{1,j}^{(k)})$
 ($j$ is used as an index for the eigenvectors),
@@
 which provide an insight into the relative importance of the uncertain inputs.
-Note that the index, $r$, corresponds to the number of dominant eigenvectors in
+Note that the index, $r$, corresponds to the number of eigenvectors in
 $\hat{\mat{W}}_1$. The sequence of steps as discussed are outlined
@@
 the 19-dimensional H$_2$/O$_2$ reaction kinetics
-problem with uncertain $A_i$'s. For the purpose of verification,
+problem with uncertain \rebut{$\log(A_i)$ values as discussed earlier}.
+For the purpose of verification,
 $\hat{\mat{C}}$ was initially constructed using a large set of samples
@@
 In Figure~\ref{fig:eig_comp}, we illustrate the comparison of
-the resulting normalized eigenvalue spectrum ($\lambda_i/\lambda_0,~i = 1,\ldots,19$)
-with the same using a much 
-smaller set of samples, $n$~=~$\{20,40,80,120\}$.
+the resulting normalized eigenvalue spectrum \rebut{by plotting $\log(\lambda_i/\lambda_0)$} ($i = 1,\ldots,19$) corresponding to $N$~=~1000 and the same quantity corresponding to
+a much smaller set of samples, $n$~=~$\{20,40,80,120\}$.
 %
@@
  \begin{center}
-  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp}
+  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp_p3}
 %   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1}
-\caption{A comparison of the normalized eigenvalue spectrum ($\lambda_i/\lambda_0$)
+\caption{A comparison of the normalized eigenvalue spectrum, $\log(\lambda_i/\lambda_0)$
 using $n$~=~$\{20,40,80,120\}$ samples with that
@@
 to improve with the sample size. Since 
-$\lambda_1$ is roughly three orders of magnitude larger than $\lambda_2$, we expect 
-a 1-dimensional active subspace. Hence, the first eigenvector 
-sufficiently captures the uncertainty in
-the  model output. To further confirm this, we evaluate a relative L$^2$ norm of the difference
+$\lambda_1$ is roughly \rebut{an order of magnitude} larger than $\lambda_2$, we expect 
+a 1-dimensional active subspace \rebut{to reasonably approximate the uncertainty
+in the ignition delay}. 
+To further confirm this, we evaluate a relative L$^2$ norm of the difference
  ($\varepsilon_{\text{L}^{2}}^{N-n}$) between the 
@@
 The quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$, was found to be 
-$\mathcal{O}(10^{-2})$ for $n$ = 20, 40, and 80; and $\mathcal{O}(10^{-3})$ for $n$ = 120.
+$\mathcal{O}(10^{-2})$ in all cases.
 Thus, even a small sample size, $n$ = 20, seems to approximate the dominant eigenspace with
@@
 delay. Additionally, the quantity, $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ is plotted in
-Figure~\ref{fig:comp} (right) to illustrate convergence characteristics of Algorithm~\ref{alg:grad}.
+Figure~\ref{fig:comp} (right) to illustrate convergence characteristics of Algorithm~\ref{alg:grad} \rebut{in both cases}. 
 %
@@
  \begin{center}
-  \includegraphics[width=0.8\textwidth]{./Figures/eigv6}
+  \includegraphics[width=0.8\textwidth]{./Figures/eig_convD19}
 \caption{Left: An illustrative comparison of individual squared components of the converged dominant
@@
 %
+\rebut{Convergence was established once $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ was found
+to be smaller than a set tolerance, $\tau$ = 0.05. As shown in Figure~\ref{fig:comp}, 8
+iterations were required in both cases. To compare convergence characteristics or the
+}
+
 Individual components of the dominant eigenvector from the two approaches are observed to be in excellent

### 2018-11-21 Revision/Paper/method.tex
..

@@
 \frac{\|\vec{w}_{1,N}^2-\vec{w}_{1,n}^2\|_2}{\|\vec{w}_{1,N}^2\|_2}
+\label{eq:accu}
 \ee
@@
 %
-In Figure~\ref{fig:comp} (left),
-we provide a comparison of the square of individual components of the dominant eigenvector, obtained using both, perturbation-based and regression-based estimates of the gradient of the ignition 
-delay. Additionally, the quantity, $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ is plotted in
-Figure~\ref{fig:comp} (right) to illustrate convergence characteristics of Algorithm~\ref{alg:grad} \rebut{in both cases}. 
+\rebut{The quantity, $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ is plotted in
+Figure~\ref{fig:comp} (center) to illustrate convergence characteristics based on Algorithm~\ref{alg:grad} 
+for the two approaches. As observed for a set tolerance $\tau$~=~0.05, 8 iterations are needed 
+to converge in both cases. 
+At each iteration, 5 new random samples were added for the perturbation approach. As discussed earlier, gradient 
+computation (using finite difference) at 5 samples requires 100 model runs. Hence, in order to compare the
+two approaches for the same computational effort, we generate 100 random samples in the 19D input domain for 
+the regression approach. Therefore, 8 iterations require 800 model runs leading to a total of 900 model runs 
+including the initial step in both cases.  The accuracy of the two approaches is compared for the same 
+computational effort by computing $\varepsilon_{\text{L}^{2}}^{N-n}$ using~\eqref{eq:accu} in 
+Figure~\ref{fig:comp} (right). Once again, $N$
+corresponds to the case of active subspace computation using the perturbation approach using 1000 samples
+in the 19D domain; $n$ ranges from 5 to 45 for the perturbation approach, and 100 to 900 for the regression 
+approach corresponding to iterations 1 to 8. The accuracy of the two approaches is found to be comparable for
+the same amount of computational effort. However, the perturbation approach is observed to be slightly more
+accurate.}
 %
@@
   \includegraphics[width=0.8\textwidth]{./Figures/eig_convD19}
-\caption{Left: An illustrative comparison of individual squared components of the converged dominant
-eigenvector obtained using the two strategies. Right: The quantity,  $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$
-is plotted for successive iterations to illustrate convergence behavior for the regression-based strategy.}
+\caption{\rebut{Left: An illustrative comparison of individual squared components of the converged dominant
+eigenvector obtained using perturbation and regression strategies using $M$~=~900 model runs in each case. 
+Additionally, the dominant eigenvector components obtained using $M$~=~20000 model runs (corresponding to 
+$N$~=~1000 samples)
+in the perturbation strategy (test case), used to assess the accuracy of the two strategies are also plotted.
+Center: The quantity,  $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$
+is plotted for successive iterations to illustrate convergence behavior for the two strategies. Right: The
+quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$ is plotted as a function of iterations to compare the accuracy
+of the two strategies. Note that the above plots compare the convergence and accuracy of the two
+strategies for the same amount of computational effort, quantified by the number of model runs, $M$.
+The legend provides the number of samples, $N$ and correspondingly, the
+model runs, $M$ required for obtaining a converged dominant eigenvector using the perturbation and
+regression strategies. Corresponding value of $N$ and $M$ for the test case are also provided.}}
 \label{fig:comp}
@@
 %
-\rebut{Convergence was established once $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ was found
-to be smaller than a set tolerance, $\tau$ = 0.05. As shown in Figure~\ref{fig:comp}, 8
-iterations were required in both cases. To compare convergence characteristics or the
-}
-
-Individual components of the dominant eigenvector from the two approaches are observed to be in excellent
-agreement with each other. Moreover, as expected,  $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ is observed 
-to decrease with iterations, and requires model evaluations at only 30 samples to converge using $\tau$~=~0.1.
+\rebut{In Figure~\ref{fig:comp} (left), we compare the
+squared values of individual components of the dominant eigenvector obtained using the two approaches
+at the end of 8 iterations in each case. Additionally, the dominant eigenvector components obtained
+for a test case involving $N$~=~1000 samples using the perturbation approach are also plotted. 
+The set of eigenvector components for the three cases are found to be in excellent agreement with
+each other, indicating that both
+approaches have sufficiently converged and are reasonably accurate for this setup.}
 
@@
 H$_2$/O$_2$ reaction in this case, varies
- predominantly in the active subspace. Hence, 
+ predominantly in \rebut{a 1-dimensional} active subspace. Hence, 
 $f(\bm{\xi})$ can be approximated as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
-the subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, 
+the \rebut{1-dimensional active} subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, 
 regarded as the \textit{sufficient summary plot} (SSP), obtained using the perturbation-based and
@@
  \begin{center}
-  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp}
-\caption{An illustrative comparison of the SSPs generated using the 
-perturbation-based and the regression-based strategies for computing the active subspace.}
+  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp19D}
+\caption{\rebut{An illustrative comparison of the SSPs generated using the 
+perturbation and the regression strategies for computing the active subspace.}}
 \label{fig:comp_ssp}
@@
 %
-The two SSPs are observed to be in excellent agreement with each other. Moreover, it is interesting to note
+\rebut{The dominant eigenvector obtained using perturbation is based on $N$~=~45 samples which requires
+$M$~=~900 model runs. For the same amount of computational effort, we can afford $N$~=~900 samples
+when using regression. Hence, the SSP from regression is based on 900
+points: ($\hat{\mat{W}}_1^\top\bm{\xi}$,$G(\hat{\mat{W}}_1^\top\bm{\xi})$). On the other hand, the SSP
+from perturbation is plotted using only 45 points as mentioned earlier.
+Nevertheless, the illustrative comparison clearly indicates that the two SSPs
+are in excellent agreement. Moreover, it is interesting to note
 that the response in ignition
-delay based on the considered probability distributions for $A_i$'s is observed to be approximately linear in the
-active subspace. This observation further indicates that the active subspace is 1-dimensional in this case.
+delay based on the considered probability distributions for $\log(A_i)$ although non-linear, can be approximated by
+a 1-dimensional active subspace.}
 
 We further estimate
-the normalized activity scores for individual uncertain inputs ($\tilde{\nu}_{i,r}$; $r$=1 since the
-active subspace is 1-dimensional)
-using the components of the dominant eigenvector as shown in Algorithm~\ref{alg:grad}.
+the normalized activity scores for individual uncertain inputs ($\tilde{\nu}_{i,r}$; $r$=1 \rebut{since a
+1-dimensional active subspace seems reasonably accurate})
+using the components of the dominant eigenvector as shown in Algorithm~\ref{alg:grad} (steps 21 and 22).
 The activity scores for the 19 uncertain pre-exponents ($A_i$'s), estimated
-using the perturbation-based and regression-based approaches are plotted in
-Figure~\ref{fig:comp_as}.  
+using the perturbation and regression strategies are plotted in Figure~\ref{fig:comp_as}.  
 %
@@
  \begin{center}
-  \includegraphics[width=0.45\textwidth]{./Figures/comp_as}
+  \includegraphics[width=0.45\textwidth]{./Figures/comp_as19D}
 %  \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}
-\caption{Left: A bar-graph of normalized activity scores ($\tilde{\nu}_{i,r}$'s) 
+\caption{\rebut{Left: A bar-graph of normalized activity scores ($\tilde{\nu}_{i,r}$'s) 
 for the 19 uncertain pre-exponents ($A_i$'s); $r$
-denotes the number of eigenvectors in the active subspace.}
+denotes the number of eigenvectors in the dominant eigenspace.}}
 \label{fig:comp_as}
@@
 based on the screening metric involving the DGSMs in~\cite{Vohra:2018}. It is observed that the uncertainty
-associated with the ignition delay is largely due to the uncertainty in $A_9$ and $A_1$. Sensitivity towards
-$A_{15}$ and $A_{17}$ is also found to be significant.
+associated with the ignition delay is largely due to the uncertainty in $A_9$ \rebut{while $A_1$, $A_{15}$, and
+$A_{17}$ are also observed to contribute significantly towards its variance.}
 
@@
 uncertain $A_i$'s is reasonably approximated using both, perturbation-based and regression-based
-approaches in this case. Since the number of model evaluations required in the case of
-regression are much fewer, this approach seems favorable for the 19-dimensional problem. 
+approaches in this case. Since the regression approach requires fewer model evaluations for
+gradient estimation, it might be a preferred approach for this setting.
 In the following section, we shift our focus to 
-the higher-dimensional H$_2$/O$_2$ reaction kinetics application involving 33 uncertain inputs.
+the higher-dimensional H$_2$/O$_2$ reaction kinetics application \rebut{wherein the activation
+energies in the rate law as well as initial pressure, temperature, and stoichiometric conditions
+are also considered to be uncertain.}

### 2018-11-29 Revision/Paper/app.tex
..

@@
 For the high-dimensional case, we aim to investigate the impact of the uncertain
-pre-exponents ($A_i$'s) as well as the activation energies ($E_{a,i}$'s) on ignition 
-delay during the H$_2$/O$_2$ reaction.  Prior distributions for $A_i$'s were considered
-to be the same as provided earlier in section~\ref{sec:method}. The $E_{a,i}$'s for all
-reactions except $\mathcal{R}_6$ -- $\mathcal{R}_9$ and $\mathcal{R}_{13}$ (due to 
-zero nominal values for $E_a$)
-were considered to be uncertain and uniformly distributed in the interval: 
-$[0.99E_{a,i}^\ast, 1.01E_{a,i}^\ast]$. Hence, the total number of uncertain rate parameters
-is 33. The nominal values, $E_{a,i}^\ast$
-corresponding to the different reaction rates are provided in~\cite{Yetter:1991}.
+pre-exponents ($A_i$'s) as well as the activation energies ($E_{a,i}$'s) \rebut{and
+the initial conditions: pressure~($P_0$), temperature~($T_0$), and stoichiometry based on
+the equivalence ratio~($\Phi_0$)} on ignition delay during the H$_2$/O$_2$ reaction. 
+\rebut{The $\log(A_i)$'s, $E_{a,i}$'s for all reactions except 
+$\mathcal{R}_6$ -- $\mathcal{R}_9$, $\mathcal{R}_{13}$ (due to 
+zero nominal values for $E_a$), and the initial conditions were perturbed by
+2$\%$ about their nominal values. Note that the mangnitude of the perturbation
+was selected such that the ignition delay remains positive throughout the input
+domain. The nominal values of the rate parameters, $A_i$'s and $E_{a,i}$'s 
+were taken from~\cite{Yetter:1991}. The nominal values of $P_0$, $T_0$, and
+$\Phi_0$ were considered to be 1.0 atm, 900~K, and 2.0 respectively.}
 
@@
 
-The active subspace was computed using the iterative procedure outlined in Algorithm~\ref{alg:grad}.
-The convergence of the eigenvectors
-was examined by tracking $\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$, plotted in 
-Figure~\ref{fig:conv_app} (right). In Figure~\ref{fig:conv_app} (left), individual
-components of the converged eigenvector that constitutes the active subspace,
-are illustrated. The convergence was established using
-a $\tau$ value of 0.02. Finite difference was used to compute model gradients at 40 samples in
-the input domain. Hence, a total of 1360 model evaluations were required for this purpose.  
+The active subspace was computed using the iterative procedure outlined in 
+Algorithm~\ref{alg:grad}. The convergence of the eigenvectors was examined
+by tracking the quantity `$\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$'. 
+\rebut{In Figure~\ref{fig:conv_app}, we examine $\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$
+with increasing iterations for the perturbation and the regression approaches 
+discussed earlier in Section~\ref{sec:method}}. 
 %
@@
  \begin{center}
-  \includegraphics[width=0.8\textwidth]{./Figures/eigv10}
-\caption{Left: Individual components of the dominant eigenvector in the
-active subspace,
-obtained using the perturbation-based strategy. Right: The quantity,  
-$\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$
-is plotted for successive iterations to illustrate the convergence behavior.}
+  \includegraphics[width=0.8\textwidth]{./Figures/eig_conv36Dp2}
+\caption{\rebut{Left: An illustrative comparison of individual components of the 
+dominant eigenvector in the converged active subspace i.e., at the end
+of 4 iterations in the perturbation approach and 9 iterations in the
+regression approach. Right: A comparison of the convergence behavior
+of the perturbation and the
+regression approaches. Convergence is accomplished once 
+$\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$ assumes a value smaller
+than 0.05.}}
 \label{fig:conv_app}

### 2018-11-30 Revision/Response/response.tex
..

\documentclass[11pt,final]{article}
\renewcommand*\familydefault{\sfdefault}
%\usepackage{amssymb,amsmath,amsfonts,comment}
%\usepackage{amsmath,amssymb,graphicx,subfigure,psfrag}
\usepackage{amsmath,amssymb,graphicx,subfigure,psfrag,upgreek}
\usepackage{algorithm,algorithmic}
\usepackage{amssymb,mathrsfs}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{color,pdfcolmk}
\newcommand{\todo}[1]{\noindent\emph{\textcolor{red}{Todo: #1\:}}}
%\newcommand{\alennote}[1]{\noindent\emph{\vspace{1ex}\textcolor{cyan}{Alen: #1\:}}\\[1ex]}
\newcommand{\referee}[1]{\vspace{.1ex}\noindent{\textcolor{blue}{#1}}}

\begin{document}

%We thank the reviewers for their careful reading of our article 
%and the helpful comments and suggestions.
%Please find below point-by-point replies (in black) to your comments and
%questions (which are reprinted in blue). To give you an overview of all the
%changes in the paper, we also provide a diff-document that highlights the
%changes between the initial submission and this re-submission.\\[1ex]
\begin{center}
{\bf Summary of Modifications to CNF-D-18-00757}\\[6pt]
{\bf Subspace-based dimension reduction for chemical kinetics applications with epistemic uncertainty}\\[6pt]
By \\
Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan 
\end{center}

\baselineskip=22pt

\vspace*{1in}

%We thank the reviewers for their assessment of our manuscript. Please find
%below point-by-point replies (in black) to your comments and questions
%(reprinted in blue). Where possible, key modifications have been highlighted
%in blue in the revised manuscript. We sincerely hope that with these
%modifications, the paper is found suitable for publication in the
%{\it Journal of Scientific Computing}.

\clearpage

\section{Replies to reviewer \#1}
%\referee{This draft proposed a systematic approach for surrogate model construction in
%reduced input parameter spaces in order to reduce the number of model
%evaluations. The key idea is to approximate the screening parameter iteratively
%in order to identify the unimportant inputs. The screening procedure also
%integrates with the adaptive construction of a surrogate in the reduced space
%to improve the efficiency of the methods. Several numerical examples are used
%to demonstrate the efficiency and accuracy of the proposed framework.}
%
%Comments and Questions:
%\begin{enumerate}
%
%\item \referee{P16, Figure.2,  the estimated screening parameters eq (5) based on eq(1)
%quickly converge with a small number of samples (5-10 samples). It seems that
%if $\mu_i$ is based on eq(1), it will still require a reasonable amount of
%samples, can authors comment on why this example converges with a small number
%of samples? The same question applies to Figure 5 and Figure 7.}
%
%Convergence behavior in this case depends upon the variability in the gradient of
%$G(\theta)$ with respect to individual parameters, $\theta_i$ in their respective
%domains. For the numerical tests and the application considered in this work, the
%gradient is not observed to vary significantly in the considered domain of the
%parameters. In other words, the variance of the estimator is low. 
%Hence, a small number of Monte Carlo samples are able to estimate
%the screening parameters. This is a common observation in many practical
%applications, and therefore the proposed methodology involving an iterative
%approach can yield significant computational savings. 
%
%These points have been added in the discussion in Section~6 (Summary and
%Conclusion, Page 34) in the revised manuscript. 
%
%\end{enumerate}

\section{Replies to reviewer \#2}

\end{document}

### 2018-11-30 Revision/Paper/abstract.tex
..

--- 2018-09-24 Paper/abstract.tex
@@
 \section*{Abstract}
-We focus on efficient quantification of uncertainty in complex chemical
-reaction networks with a large number of uncertain parameters.  Parameter
+We focus on an efficient approach for quantification of uncertainty in complex chemical
+reaction networks with a large number of uncertain parameters \rebut{and input conditions}.  Parameter
 dimension reduction is accomplished by computing an~\emph{active subspace} that
@@
 uncertainty in the pre-exponents of the individual reaction rates is
-considered. This is followed by the analysis of a 33-dimensional case wherein
-the activation energies are also considered uncertain.  In both cases, a
-one-dimensional active subspace is identified, which indicates enormous
+considered. This is followed by the analysis of a \rebut{36}-dimensional case wherein
+the activation energies \rebut{and initial conditions} are also considered uncertain.  In both cases, a
+1-dimensional active subspace is observed to capture the uncertainty in the QoI,
+ which indicates enormous
 potential for efficient statistical analysis of complex chemical systems. In
@@
 contributors to the variability in the model response.
+\bigskip
+
+\noindent \textbf{Keywords}: Chemical kinetics; epistemic uncertainty; active subspace; dimension reduction; surrogate

### 2018-11-30 Revision/Paper/app.tex
..

@@
 by tracking the quantity `$\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$'. 
-\rebut{In Figure~\ref{fig:conv_app}, we examine $\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$
+\rebut{In Figure~\ref{fig:conv_app}~(right), we examine $\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$
 with increasing iterations for the perturbation and the regression approaches 
@@
 %
+\rebut{At each iteration, we improve our estimates of the matrix $\hat{\mat{C}}$ by
+estimating the gradient of the ignition delay at 5 new randomly generated samples 
+in the 36-dimensional input space. However, gradient computation at these
+5 samples requires 185 (=5$\times$(36+1)) model runs when using perturbation. 
+For the same computational effort, the regression approach can afford 185 new
+samples at each iteration. It is observed that using $\tau$ = 0.05, the 
+active subspace requires 4 iterations (925 model runs) to converge in the case of perturbation, and
+9 iterations (1850 model runs) to converge in the case of regression. Hence, the computational effort
+required to obtain a converged active subspace is doubled when using
+regression to approximate the gradient. Moreover, gradient estimation in the perturbation approach
+can be made more efficient by using techniques such as automatic differentiation~\cite{Kiparissides:2009}
+and adjoint computation~\cite{Jameson:1988}. These techniques although not pursued here are 
+promising directions for
+future efforts pertaining to this work. In Figure~\ref{fig:conv_app}~(right), we compare
+individual components of the dominant eigenvector in the converged active subspace
+obtained using the two approaches. The components are observed to be in excellent
+agreement with each other.}
 
-In Figure~\ref{fig:hd} (left), we illustrate converged estimates of the eigenvalues.
- The second eigenvalue is
-found to be roughly two orders of magnitude smaller than the first, indicating that the active subspace
-is 1-dimensional. This is further confirmed by the SSP plots in Figure~\ref{fig:hd} (right), generated
-using both, perturbation-based and regression-based approaches for estimating
-the gradient. 
+\rebut{In Figure~\ref{fig:hd}, we plot the SSP for the perturbation approach (left) and the regression
+approach (center) in a 1-dimensional active subspace. A 1-dimensional polynomial fit is also
+illustrated in both cases. Moreover, the two surrogates are shown to be consistent with each other (right).
+From these results, it is clear that a 1-dimensional active subspace captures the variability in the
+ignition delay with reasonable accuracy, and that the two approaches yield consistent results.}
 %
@@
  \begin{center}
-   \includegraphics[width=0.45\textwidth]{./Figures/eig_33D}
-   \includegraphics[width=0.45\textwidth]{./Figures/ssp_33D}
-\caption{Left: Eigenvalue spectrum of $\hat{\mat{C}}$. Right: SSP of the computed active subspace, and the
-corresponding 1-D linear fit.} 
+   \includegraphics[width=0.9\textwidth]{./Figures/SSP_plot36Dp2}
+\caption{\rebut{Sufficient summary plots (SSPs) for the case of perturbation (left) and regression (center).
+A polynomial fit of degree 2 and 3 as shown in the plots is used as a surrogate in the 
+perturbation and regression approaches respectively. An illustrative comparison of the two
+surrogates is also provided (right).}}
 \label{fig:hd}
@@
 %
-Specifically
-in the case of perturbation-based approach, the SSP is observed to exhibit a linear variation that
-is reasonably
-captured by the corresponding straight-line fit, $\tilde{G}$, constructed using the sequence of steps
-outlined in Section~\ref{sub:ac}.
-On the other hand, the SSP obtained by estimating the gradient using regression,
-is relatively more scattered. Consequently, the two surrogates (straight-line fits) are observed to exhibit a
-discrepancy unlike our earlier observation for the 19-dimensional problem in Figure~\ref{fig:comp_ssp}.
 
@@
 
-The 1-dimensional surrogate ($\tilde{G}$) shown in Figure~\ref{fig:hd}~(right) is investigated
+The 1-dimensional surrogate ($\hat{G}$) shown in Figure~\ref{fig:hd} is investigated
 for its accuracy as well as the ability to capture the uncertainty in the
-model output in two ways. Firstly, we estimate the relative L$_2$ norm of the discrepancy~($\varepsilon_d$)
+model output in two ways. Firstly, we estimate the relative L$^2$ norm of the discrepancy~($\varepsilon_d$)
 between estimates of ignition delay in the case of H$_2$/O$_2$ reaction, obtained using the 
@@
 \be
-\varepsilon_d = \frac{\|G(\bm{\xi}) - \tilde{G}(\bm{\xi})\|_2}{\|G(\bm{\xi})\|_2}
+\varepsilon_d = \frac{\|G(\bm{\xi}) - \hat{G}(\bm{\xi})\|_2}{\|G(\bm{\xi})\|_2}
 \ee
 %
-Model evaluations ($G$) at 10$^{4}$ samples in the input domain were used to evaluate $\varepsilon_d$. 
-It was estimated to be 1.35$\times10^{-2}$ and 1.02$\times10^{-1}$
-using $\tilde{G}$ based on the perturbation-based and the regression-based approaches respectively. Hence,
-$\varepsilon_d$ was found to increase by an order of magnitude when the gradient is estimated using
-the regression-based approach.
-Therefore, in the norm-sense, it can be said that the 1-dimensional surrogate in the active subspace, obtained
-using the perturbation-based approach is remarkably accurate with a relative error of less than 1.5$\%$, whereas,
-the regression-based approach resulted in a relative error of about 10$\%$.
+Model evaluations ($G$) at 10$^{4}$ random samples in the input domain were used to evaluate $\varepsilon_d$. 
+\rebut{It was estimated to be approximately 0.17 for both approaches. Hence, the predictive accuracy of
+the two approaches is comparable with respect to the relative L$^2$ norm of the discrepancy.
+ However, as mentioned earlier, the computational effort required for
+obtaining a converged active subspace using perturbation was half of that required when using
+regression to estimate the model gradient at random samples in the 36-dimensional input domain.}
 
@@
 probability density functions (PDFs) obtained using the true set of model evaluations, and 1-dimensional
-surrogates ($\tilde{G}$'s) based on the two approaches for gradient estimation as shown in Figure~\ref{fig:pdf_33D}.
-Note that the three PDFs were evaluated using the same set of 10$^4$ samples in the cross-validation set. 
+surrogates ($\hat{G}$'s) based on the two approaches for gradient estimation as shown in 
+Figure~\ref{fig:pdf_36D}. Note that the three PDFs were evaluated using the same set of 10$^4$ samples in the 
+cross-validation set. 
 %
@@
 \begin{center}
-\includegraphics[width=3.0in]{./Figures/pdf_comp_id_1e4}
+\includegraphics[width=3.0in]{./Figures/pdf_plot36Dp2}
 \end{center} 
-\caption{A comparison of the PDFs of ignition delay, obtained using model 
+\caption{\rebut{A comparison of the PDFs of ignition delay, obtained using model 
 evaluations (solid line) and 1-dimensional surrogates using the regression-based strategy (dashed line) and the
 perturbation-based strategy (dashed-dotted line). The same set of 10$^4$ samples in the cross-validation set were 
-used in each case.}
-\label{fig:pdf_33D}
+used in each case.}}
+\label{fig:pdf_36D}
 \end{figure}
 %
-%
-%\begin{figure}[htbp]
-% \begin{center}
-%  \includegraphics[width=0.45\textwidth]{./Figures/pdf_comp_id_1e4}
-%\caption{A comparison of the PDFs of ignition delay, obtained using model evaluations (solid line) and
-%1-dimensional surrogates using the gradient-free approach (dashed line) and the gradient-based approach
-%(dashed-dotted line). The same set of 10$^4$ samples in the cross-validation set were used in each case.}
-%\label{fig:pdf_33D}
-%\end{center}
-%\end{figure}
-%
-The PDFs based on the model evaluations and the perturbation-based approach are nearly identical, confirming
-the existence of a 1-dimensional active subspace as well as the accuracy of gradient estimates in this case. 
-On the other hand, although the PDF obtained using the regression-based approach  is observed to show
-consistency with the true PDF in the modal
-estimate, it clearly underestimates the spread in the ignition delay. Specific values of the mean and the standard 
-deviation in each case are provided in Table~\ref{tab:stats}. 
+\rebut{The PDFs are observed to be in close agreement with each other. Specifically, the modal
+estimate and the uncertainty (quantified by the spread in the distributions) is found to be consistent for the three 
+cases. To confirm this, we compute the first-order (mean) and the second-order (standard deviation) statistics
+of the estimates of the ignition delay obtained using the model, 1-dimensional surrogate from perturbation, and
+1-dimensional surrogate from regression at the cross-validation sample set. The mean and standard deviation
+estimates are provided in Table~\ref{tab:stats}.}
 %
@@
 \bottomrule
-$G$~(Model) & 0.133 & 0.0198 \\
-$\tilde{G}$~(Perturbation-based) & 0.133 & 0.0196 \\
-$\tilde{G}$~(Regression-based) & 0.133 & 0.0167 \\
+$G$~(Model) & \rebut{0.1510} & \rebut{0.1373} \\
+$\hat{G}$~(Perturbation-based) & \rebut{0.1514} & \rebut{0.1275} \\
+$\hat{G}$~(Regression-based) & \rebut{0.1510} & \rebut{0.1330} \\
 \bottomrule
@@
 \caption{The mean ($\mu$), and the standard deviation ($\sigma$), computed using the model ($G$), and
-the surrogate ($\tilde{G}$) based on the two strategies at 10$^4$ samples in the cross-validation
+the surrogate ($\hat{G}$) based on the two strategies at 10$^4$ samples in the cross-validation
 set.}
@@
 %
-The three PDFs are found to be
-consistent in their prediction of mean value of the ignition delay.  The standard deviation estimates using model
-evaluations and the perturbation-based approach are found to be in close agreement, whereas, it  is accurately
- estimated in the case of regression-based approach only upto the first significant digit.
+\rebut{The mean and the standard deviation estimates obtained using the model and the 1-dimensional surrogates
+are found to be in close agreement. It is interesting to note that estimates from the regression approach are
+observed to be slightly more accurate. This is due to the fact that the converged active subspace in this case is 
+based on 1850 samples in the 36-dimensional input domain, whereas, the converged active subspace based the 
+perturbation approach is based on only 25 samples. As a result, the surrogate based on regression explores
+a much wider range of values of the ignition delay compared to the surrogate based on perturbation as shown in
+Figure~\ref{fig:hd}~(right).}
 
@@
 
-The normalized activity scores ($\tilde{\nu}_{i,r}$) based on the dominant eigenspace, obtained
-using the two approaches for gradient estimation (perturbation-based and regression-based),
+The normalized activity scores ($\tilde{\nu}_{i,r}$) based on the 1-dimensional active subspace, obtained
+using the two approaches for gradient estimation (perturbation and regression),
 are compared with the 
-total Sobol' indices
-in Figure~\ref{fig:as_33D}. Note that the Sobol' indices were computed using the verified
-1-dimensional surrogate ($\tilde{G}$) in the active subspace, obtained using the 
-perturbation-based approach. 
+total-effect Sobol' indices
+in Figure~\ref{fig:as_36D}. Note that the Sobol' indices were computed using the verified
+1-dimensional surrogate ($\hat{G}$) in the active subspace, obtained using the 
+perturbation approach. 
 %
@@
  \begin{center}
-  \includegraphics[width=0.8\textwidth]{./Figures/as_33D_new2}
-\caption{A bar-graph illustrating individual activity scores for the uncertain $A_i$' (left) and $E_{a,i}$'s (right).}
-\label{fig:as_33D}
+  \includegraphics[width=0.8\textwidth]{./Figures/as_36Dp2}
+\caption{\rebut{Bar graphs illustrating individual activity scores for the uncertain 
+rate parameters and the initial conditions for the H$_2$/O$_2$ reaction.}}
+\label{fig:as_36D}
 \end{center}
@@
 %
-Several useful inferences can be drawn. Firstly, the total Sobol indices are found to be
-consistent
-with the normalized activity scores as obtained when the gradient is estimated using a perturbation-based approach.
-The regression-based
-approach, however, yields consistent results for the $A_i$'s but fails capture the sensitivity associated with one
-of the most important parameters, $E_{a,15}$. This is because numerical errors are incurred when approximating
-the model gradients using a linear regression fit in this case. 
-Secondly, our findings indicate that   
-the ignition delay is predominantly sensitive towards $A_1$, $A_9$, $E_{a,1}$, and $E_{a,15}$ and 
-moderately sensitive towards $A_{15}$ and $A_{17}$. Sensitivity towards the remaining uncertain inputs is
-found to be low or negligible. 
-%Note that this observation can also be exploited for reducing the 
-%dimensionality of the problem from 33 inputs to 4-6 inputs depending upon the required level of accuracy.
-%The sensitivity-driven dimension reduction for uncertainty quantification has been
-%discussed in our earlier effort~\cite{Vohra:2018}. In this work, however, we focus on dimension reduction using the active 
-%subspace approach which is shown to yield even greater scope for dimension reduction i.e. from 33 to 1. Furthermore, we have
-%shown that the dominant eigenspace can be used to approximate global sensitivity measures with no additional effort.
+Several useful inferences can be drawn. \rebut{Firstly, the normalized activity scores from the two approaches
+and the total-effect Sobol' indices are found to be in close agreement with each other. Secondly, as expected, 
+$\tilde{\nu}_{i,r}$ based on perturbation exhibits a better agreement with the total-effect Sobol' indices since
+the 1-dimensional
+surrogate based on the same approach was used to evaluate the Sobol' indices. This observation demonstrates
+that the proposed framework is self-consistent. Thirdly, the variability in the ignition delay is predominantly due
+to the uncertainty in $A_1$, $A_9$, and $T_0$ while contributions from the uncertainty in $A_{15}$, $A_{17}$, and
+$E_{a,15}$, and $T_0$ are also found to be significant. The remaining rate parameters, initial pressure ($P_0$),
+and  stoichiometry ($\Phi_0$) do not seem to impact the ignition delay in their considered intervals. 
+Therefore, GSA has helped
+identify the important rate parameters i.e. key contributors to the uncertainty, and also demonstrated that among the 
+considered uncertain initial conditions, the ignition delay is mainly impacted by the perturbations in the initial 
+temperature in the considered interval.}

### 2018-11-30 Revision/Paper/assa.tex
..

@@
 \item Use data points $\big(\vec{y}_i, f(\vec\xi_i)\big)$, $i = 1, \ldots, N$, to compute a 
-regression surface $\tilde{G}(\vec{y})\approx 
+regression surface $\hat{G}(\vec{y})\approx 
 G(\vec{y})$.
-\item Overall approximation, $f(\vec{\xi})$ $\approx$ $\tilde{G}(\mat{W}_1^\top\vec{\xi})$.
+\item Overall approximation, $f(\vec{\xi})$ $\approx$ $\hat{G}(\mat{W}_1^\top\vec{\xi})$.
 \end{enumerate}
@@
 approximated using pseudo-random sampling techniques such as Monte Carlo or
-Latin hypercube sampling:
+Latin hypercube sampling (used in this work):
 %The integral in~\eqref{eq:C} is replaced with a
@@
 
-The total Sobol' index ($T_i(f)$) of a model output, $f(\vec\xi)$ quantifies
+The total-effect Sobol' index ($T_i(f)$) of a model output, $f(\vec\xi)$ quantifies
 the total contribution of the input, $\xi_i$ to the variance of the
@@
 where $\vec{\xi}_{\sim i}$ is the input parameter vector with the  
-$i^\text{th}$ entry removed, and $\V$ denotes the variance. The total Sobol' index accounts
+$i^\text{th}$ entry removed, and $\V$ denotes the variance. The total-effect Sobol' index accounts
 for the contribution of a given input to the variability in the output by itself
@@
 
-It was shown in~\cite{Lamboni:2013} that the total Sobol' 
+It was shown in~\cite{Lamboni:2013} that the total-effect Sobol' 
 index $T_i(f)$ can be bounded in terms of $\nu_i(f)$:
@@
 Combining Proposition~\ref{prp:dgsm_bound} and~\eqref{equ:sobol_bound}, shows
-an interesting link between the activity scores and total Sobol' indices.
+an interesting link between the activity scores and total-effect Sobol' indices.
 Specifically, by computing the activity scores, we can identify the unimportant
@@
 error between the reduced model and the original model. If a precise ranking of
-parameter importance based on total Sobol' indices is desired, one
+parameter importance based on total-effect Sobol' indices is desired, one
 can first identify unimportant inputs by computing activity scores, and then

### 2018-11-30 Revision/Paper/conc.tex
..

--- 2018-09-25 Paper/conc.tex
@@
 
-%%Brief summary
-%-Mathematical connections between Sobol, DGSMs, and activity scores exist and could
-%be exploited to perform GSA with reduced effort.
-%-Iterative implementation yields advantage
-%-Consistency between the two approaches with DGSM as well as among themselves
-%encourages the use of gradient-free.
-%-Grad-free was found to be suitable for a high-dimensional problem.
-%-Based on the results presented, the following parameter were found to be 
-%significant.
-%
-%discussion
-%- Methodology is agnostic to the choice of mechanism and could be applied to other
-%systems.
-%- However, constraints exist in terms of differentiability of the QoI. Not all
-%input/output relationships exhibit an active subspace. However, it is important to
-%check.
-%- Activity scores could be used to identify key reactions and thereby construct
-%reduced order mechanisms for reducing computational effort and investigating the
-%inadequacy associated with rate laws.
-
 In this work, we focused on the uncertainty associated with the
-rate-controlling parameters in the H$_2$/O$_2$ reaction mechanism and its
+rate-controlling parameters in the H$_2$/O$_2$ reaction mechanism
+\rebut{as well as the initial pressure, temperature, and stoichiometry} and its
 impact on ignition delay predictions. The mechanism involves 19 different
@@
 pre-exponent and an activation energy. Hence, in theory, the evolution of the
-chemical system depends upon 38 inputs. However, we considered 
+chemical system depends upon 38 rate parameters \rebut{and three initial conditions}.
+However, we considered 
 epistemic uncertainty in all pre-exponents and activation energies with non-zero
-nominal values i.e. a total of 33 parameters instead of 38.  
+nominal values i.e. a total of 33 rate parameters instead of 38 \rebut{in addition to
+the three initial conditions}.  
 %Conventional means for uncertainty quantification such as those involving
@@
 problem wherein only the pre-exponents were considered to be uncertain.
+\rebut{It was observed that the computational effort required to obtain a 
+converged active subspace was comparable for the two approaches. However,
+the predictive accuracy of the perturbation approach was observed to be
+slightly higher. Moreover, a 1-dimensional active subspace was shown to
+reasonably approximate the uncertainty in the ignition delay.}
 Additionally, the activity scores were also shown to be consistent with the
@@
 
-The active subspace was further computed for a 33-dimensional problem
+The active subspace was further computed for a 36-dimensional problem
 wherein all pre-exponents and activation energies with non-zero nominal
-estimates were considered uncertain. Once again, the active subspace
-was found to be 1-dimensional. However, unlike the 19-dimensional case,
-gradient estimation using the regression-based approach resulted in an
-under-estimation of the variability in the ignition delay and also failed
-to capture the sensitivity towards one of the most important parameters, $E_{a,15}$.
-We attribute
-this discrepancy to two potential sources of numerical errors: (1)
-Regression-based approximation of the gradient of the model output, and (2)
-Scatter in the corresponding SSP (see Figure~\ref{fig:hd}) leading to numerical
-error incurred by the linear regression fit in this case. 
-Global sensitivity estimates using the perturbation-based
-approximation of the gradient indicated that the variability in the ignition
-delay is  largely dependent on $A_1$, $A_9$,
-$E_{a,1}$ and $E_{a,15}$, although contributions from $A_{15}$ and $A_{17}$
-were also found to be significant. 
-Hence, while the
-regression-based approach 
-reduces computational effort, the present investigations clearly reveal
-potential pitfalls associated with this approach.
+estimates \rebut{as well as the initial conditions} were considered uncertain. 
+\rebut{Once again, consistent results were obtained using the two approaches.
+A 1-dimensional active subspace was shown to reasonably
+capture the uncertainty in the ignition delay in this case. However, the
+computational effort required to compute a converged active subspace
+using perturbation was found to be half of the effort required in the case
+of regression. Predictive accuracy of the two approaches was found to 
+be comparable. Hence, perturbation seems like a preferred approach
+for the higher-dimensional problem based on our findings. GSA results indicated
+that the variability in the ignition delay is predominantly due to the 
+uncertainty in the rate parameters, $A_1$ and $A_9$ with significant
+contributions from $A_{15}$, $A_{17}$, and $E_{a,15}$. Additionally, the
+ignition delay was found to be sensitive towards $T_0$.}
 
-%Computational gains using the proposed framework in this work are essentially
-%realized by using the 1-dimensional surrogate in the active subspace that
-%essentially casts a mutlivariate input in the full space into a univariate
-%input in the reduced space. The surrogate can be used to accelerate forward
-%propagation of the uncertainty as well as parameter estimation in a Bayesian
-%setting. Hence, we focused our efforts to verify the accuracy of the
-%1-dimensional surrogates, obtained using both strategies.  The 1-dimensional
-%surrogates were assessed for accuracy by first evaluating the relative $L^2$-norm
-%of the error between its predictions and model evaluations at $10^4$
-%independent MC samples in the 33-dimensional input domain. The relative error
-%computed using predictions from the surrogate resulting from the regression-based
-%approach were found to be an order of magnitude larger than that estimated
-%using surrogate resulting from the perturbation-based approximation of the gradient.
-
-%Additionally, the surrogates were assessed by comparing probability
-%distributions for the ignition delay,
-%constructed using surrogate predictions and model evaluations at the same set
-%of 10$^4$ random samples. The PDFs based on model evaluations and the
-%1-dimensional surrogate from the perturbation-based approach were found to be
-%almost identical. However, the PDF based on 1-dimensional surrogate from the
-%regression-based approach, although captured the modal estimate with reasonable
-%accuracy, was observed to underestimate the uncertainty in the ignition delay.
-%Mean estimates of the ignition delay using model evaluations and predictions
-%using the two surrogates at 10$^4$ samples were found to be in agreement.
-%However, the standard deviation, computed using predictions from
-%the surrogate resulting from the regression-based approach was found to be accurate
-%only upto the first significant digit. Moreover, it was observed that the
-%activity scores, computed using the regression-based strategy failed to capture
-%the sensitivity towards one of the most important parameters, $E_{a,15}$.  
-%
-%More specifically, using
-%the regression-based approach, 
-%\begin{itemize} 
-%\item The desired statistics (e.g. mean and variance) of the
-%QoI might not be approximated to a desired level of accuracy.  
-%\item The
-%approach might fail to capture the sensitivity  of the QoI towards all inputs.
-%\end{itemize} 
 Therefore, in general, perturbation-based approaches should be preferred for 
-active subspace computation.
-The regression-based approach can be explored in situations involving intensive
+active subspace computation especially in situations where efficient gradient
+computation techniques (adjoints-based, automatic differentiation, etc.) are
+feasible. The regression-based approach can be explored in situations involving intensive
 simulations where the perturbation-based methods are intractable.
@@
 could further guide optimal allocation of computational resources for
-calibration of the important rate-controlling parameters in a Bayesian setting.
+calibration of the important rate-controlling parameters \rebut{and input conditions}
+ in a Bayesian setting.
 Additionally, dimension reduction using active subspaces could assist in

### 2018-11-30 Revision/Paper/intro.tex
..

--- 2018-09-28 Paper/intro.tex
@@
 activation energy corresponding to the $i^{th}$ reaction, and $R$ is the
-universal gas constant.  
+universal gas constant.  \rebut{The Arrhenius rate law in~\eqref{eq:rate} 
+is often interpreted in a logarithmic form as follows:
+%
+\be
+\log(k_i) = \log(A_i) + n_i\log(T) - E_{a,i}/RT. 
+\label{eq:ratelog}
+\ee
+}
+%
 %
@@
 \ee 
-The equivalence ratio ($\phi$) is given as follows:
+The equivalence ratio ($\Phi$) is given as follows:
 %
 \be
-\phi = \frac{(M_{\text{H}_2}/M_{\text{O}_2})_\text{obs}}{(M_{\text{H}_2}/M_{\text{O}_2})_\text{st}}.
+\Phi = \frac{(M_{\text{H}_2}/M_{\text{O}_2})_\text{obs}}{(M_{\text{H}_2}/M_{\text{O}_2})_\text{st}}.
 \label{eq:phi}
@@
 conditions. In this study, computations were performed at fuel-rich conditions,
-$\phi$ = 2.0. Homogeneous ignition at constant pressure is simulated using the
-TChem software package~\cite{Safta:2011}, and the time required for the rate of
+$\Phi$ = 2.0. Homogeneous ignition at constant pressure is simulated using the
+TChem software package~\cite{Safta:2011} \rebut{using an initial pressure, $P_0$ = 1 atm and
+initial temperature, $T_0$ = 900~K.} The time required for the rate of
 temperature increase to exceed a given threshold, regarded as \emph{ignition delay}
@@
 rate-controlling parameters, pre-exponent~($A_i$) and the activation
-energy~($E_{a,i}$) on the ignition delay. The $A_i$'s associated with all
+energy~($E_{a,i}$) \rebut{as well as the initial pressure, temperature, and the
+equivalence ratio} on the ignition delay. The $A_i$'s associated with all
 reactions and the $E_{a,i}$'s with non-zero nominal estimates
 are considered to be uniformly distributed about their nominal estimates provided
-in~\cite{Yetter:1991}.  The total number of uncertain inputs is 33 which makes
+in~\cite{Yetter:1991}. \rebut{The initial conditions are also considered to be uniformly
+distributed about their respective aforementioned values.} 
+The total number of uncertain inputs is \rebut{36} which makes
 the present problem computationally challenging due to the large number of 
-uncertain parameters.  
+uncertain parameters \rebut{in addition to the initial conditions}.  
 To address this challenge, we focus on reducing the dimensionality
-of the problem by computing the aforementioned active subspace.
+of the problem by computing the active subspace.
 This involves repeated evaluations of the gradient of a model output with
@@
 at neighboring points are required if finite difference is used which increases
-the computational effort. Regression-based techniques aim to estimate the
+the computational effort. Regression-based techniques on the other hand,
+aim to estimate the
 gradient by approximating the model output using a regression fit. These
@@
 rate-controlling parameters: pre-exponent, $A_i$ and the activation energy, $E_{a,i}$,
-for $i^{\text{th}}$ reaction in the H$_2$/O$_2$ mechanism. As mentioned above, the
+for $i^{\text{th}}$ reaction in the H$_2$/O$_2$ mechanism \rebut{as well as
+the initial conditions: $P_0$, $T_0$, and $\Phi_0$. Note that the equivalence ratio
+corresponding to the initial molar ratios of $H_2$ and O$_2$ is denotes as $\Phi_0$.}
+As mentioned above, the
 first approach uses finite difference to estimate the gradient and is regarded as
-the perturbation-based approach, whereas, the second approach involves a linear
+the perturbation approach, whereas, the second approach involves a linear
 regression-fit to the available set of model evaluations and is regarded as
-the regression-based approach.
+the regression approach.
 
@@
 Active subspace
-discovery in a high-dimensional H$_2$/O$_2$ kinetics problem involving 33
+discovery in a high-dimensional H$_2$/O$_2$ kinetics problem involving \rebut{36}
 uncertain inputs: The methodology presented in this work
-resulted in a 1-dimensional active subspace, indicating immense potential
+\rebut{successfully demonstrated that a 1-dimensional active subspace can reasonably 
+approximate the uncertainty in the QoI}, indicating immense potential
 for computational savings. The presented
@@
 method of active subspaces to achieve efficiency in uncertainty propagation.  
-\item Comprehensive numerical investigation of the perturbation-based and the regression-based approaches: 
+\item Comprehensive numerical investigation of the perturbation and the regression approaches: 
 We investigate the suitability of both approaches
@@
 the active subspace is provided. 
-Numerical results based on the perturbation-based approach are compared with those
-obtained using the regression-based approach.  The active 
+Numerical results based on the perturbation approach are compared with those
+obtained using the regression approach.  The active 
 subspace is initially computed for a 19-dimensional 
 H$_2$/O$_2$ reaction kinetics problem wherein only the $A_i$'s are
-considered as uncertain. In this case, the two approaches are observed to yield
-consistent results. We further compute the active  for a 33-dimensional H$_2$/O$_2$
-reaction kinetics problem in section~\ref{sec:app}. Once again, 
-active subspace is found to be 1-dimensional. However, in this case,
-it is observed that although the regression-based estimation of the gradient results in a 
-reasonable approximation of the mean and mode, the variability in the ignition delay
-due to uncertain rate-controlling parameters is under-estimated. Additionally,  
-the sensitivity towards one of the most important parameter is not captured.
-Hence, a trade-off between efficiency and accuracy between the two approaches is
-apparent in the higher-dimensional problem. 
+considered as uncertain. We further compute the active subspace
+for a 36-dimensional H$_2$/O$_2$ reaction kinetics problem in section~\ref{sec:app}.
+\rebut{For both settings, the convergence characteristics and the predictive accuracy
+of the two approaches is compared for a given amount of computational effort. The two
+approaches are observed to yield consistent results, and a 1-dimensional active subspace
+is observed to capture the uncertainty in the ignition delay.}
 Finally, a summary and discussion based on our findings is included in

### 2018-11-30 Revision/Paper/method.tex
..

@@
 respect to \rebut{$\log(A_i)$}: 
-a perturbation-based approach that involves computation of model
+a perturbation  approach that involves computation of model
 gradients using finite difference in order to construct the matrix $\hat{\mat{C}}$
-in~\eqref{eq:chat}, and a regression-based approach 
+in~\eqref{eq:chat}, and a regression  approach 
 that involves a linear regression fit to the available set of
@@
 reduced by a factor $(d+1)$ when using the regression-based approach.
+In other words, for the same amount of computational effort, the regression
+approach can afford a sample size that is (d+1) times larger than that in the
+case of perturbation approach.
 The specific sequence of steps
@@
 quantify the uncertainty in the model output. Moreover, the components of the
-dominant eigenvectors are used to compute the activity scores, $\bm{\nu}_r(f)$,
+eigenvectors in the active subspace are used to compute the activity scores, $\bm{\nu}_r(f)$,
 which provide an insight into the relative importance of the uncertain inputs.
@@
 the 19-dimensional H$_2$/O$_2$ reaction kinetics
-problem with uncertain \rebut{$\log(A_i)$ values as discussed earlier}.
-For the purpose of verification,
+problem by perturbing \rebut{$\log(A_i)$ by 3$\%$ about its nominal
+value as discussed earlier}. For the purpose of verification,
 $\hat{\mat{C}}$ was initially constructed using a large set of samples
@@
 The active subspace for the 19-dimensional problem was also computed using regression-based
-estimates of the gradient that does not require model evaluations at neighboring points as
+estimates of the gradient that do not require model evaluations at neighboring points as
 discussed earlier.
@@
 of the two strategies. Note that the above plots compare the convergence and accuracy of the two
-strategies for the same amount of computational effort, quantified by the number of model runs, $M$.
-The legend provides the number of samples, $N$ and correspondingly, the
-model runs, $M$ required for obtaining a converged dominant eigenvector using the perturbation and
-regression strategies. Corresponding value of $N$ and $M$ for the test case are also provided.}}
+strategies for the same amount of computational effort, quantified by the number of model runs, $M$.}}
 \label{fig:comp}
@@
 at the end of 8 iterations in each case. Additionally, the dominant eigenvector components obtained
-for a test case involving $N$~=~1000 samples using the perturbation approach are also plotted. 
+for a test case involving $N$~=~1000 (and $M$ = 20,000 model runs)
+samples using the perturbation approach are also plotted. 
 The set of eigenvector components for the three cases are found to be in excellent agreement with
@@
 when using regression. Hence, the SSP from regression is based on 900
-points: ($\hat{\mat{W}}_1^\top\bm{\xi}$,$G(\hat{\mat{W}}_1^\top\bm{\xi})$). On the other hand, the SSP
+points: ($\hat{\mat{W}}_1^\top\bm{\xi}$,~$G(\hat{\mat{W}}_1^\top\bm{\xi})$). On the other hand, the SSP
 from perturbation is plotted using only 45 points as mentioned earlier.
@@
 The above comparisons indicate that the gradient of the ignition delay with respect to the
-uncertain $A_i$'s is reasonably approximated using both, perturbation-based and regression-based
-approaches in this case. Since the regression approach requires fewer model evaluations for
-gradient estimation, it might be a preferred approach for this setting.
+uncertain $A_i$'s is reasonably approximated using both, perturbation  and regression 
+approaches in this case. \rebut{Since both approaches yield consistent results and are
+comparable in terms of convergence and accuracy, we could use either for the purpose
+of active subspace computation for this setting.}
 In the following section, we shift our focus to 

### 2018-12-01 Revision/Paper/intro.tex
..

@@
 %
-where $A_i$ is the pre-exponent, $n_i$ is the index of $T$, $E_{a,i}$ is the
+where $A_i$ is the pre-exponent, $n_i$ is the \rebut{temperature
+exponent}, $E_{a,i}$ is the
 activation energy corresponding to the $i^{th}$ reaction, and $R$ is the

### 2018-12-01 Revision/Response/response.tex
..

@@
 \usepackage{color,pdfcolmk}
+\usepackage{enumitem,kantlipsum}
 \newcommand{\todo}[1]{\noindent\emph{\textcolor{red}{Todo: #1\:}}}
@@
 {\bf Summary of Modifications to CNF-D-18-00757}\\[6pt]
-{\bf Subspace-based dimension reduction for chemical kinetics applications with epistemic uncertainty}\\[6pt]
+{\bf Subspace-based dimension reduction for chemical kinetics applications with 
+epistemic uncertainty}\\[6pt]
 By \\
@@
 
-\section{Replies to reviewer \#1}
-%\referee{This draft proposed a systematic approach for surrogate model construction in
-%reduced input parameter spaces in order to reduce the number of model
-%evaluations. The key idea is to approximate the screening parameter iteratively
-%in order to identify the unimportant inputs. The screening procedure also
-%integrates with the adaptive construction of a surrogate in the reduced space
-%to improve the efficiency of the methods. Several numerical examples are used
-%to demonstrate the efficiency and accuracy of the proposed framework.}
-%
-%Comments and Questions:
-%\begin{enumerate}
-%
-%\item \referee{P16, Figure.2,  the estimated screening parameters eq (5) based on eq(1)
-%quickly converge with a small number of samples (5-10 samples). It seems that
-%if $\mu_i$ is based on eq(1), it will still require a reasonable amount of
-%samples, can authors comment on why this example converges with a small number
-%of samples? The same question applies to Figure 5 and Figure 7.}
-%
-%Convergence behavior in this case depends upon the variability in the gradient of
-%$G(\theta)$ with respect to individual parameters, $\theta_i$ in their respective
-%domains. For the numerical tests and the application considered in this work, the
-%gradient is not observed to vary significantly in the considered domain of the
-%parameters. In other words, the variance of the estimator is low. 
-%Hence, a small number of Monte Carlo samples are able to estimate
-%the screening parameters. This is a common observation in many practical
-%applications, and therefore the proposed methodology involving an iterative
-%approach can yield significant computational savings. 
-%
-%These points have been added in the discussion in Section~6 (Summary and
-%Conclusion, Page 34) in the revised manuscript. 
-%
-%\end{enumerate}
+\section*{Reviewer \#1}
 
-\section{Replies to reviewer \#2}
+\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
+\item \referee{The authors find that the uncertainty in the ignition delay time is governed
+by the uncertainty in H+O~2$<$=$>$~O+OH.
+The connection between this reaction and the ignition delay time is so overwhelming that it
+is probably responsible for the 1-dimensional active subspace reduction. Once the sensitivity
+indices are computed, the one associated with this key reaction is largest.
+To a combustion scientist, the fact that ignition delay time (and the uncertainty in the 
+ignition delay time when pre-exponentials or activation energies are changed) is governed 
+by H+O2~$<$=$>$~O+OH is obvious. This is THE key reaction in combustion kinetics with 10s 
+(if not 100s) of studies dedicated to it (from the beginning of the field…). This is textbook k knowledge.
+E.g. Figure 4 in Hong et al. (PCI 33, 2011, p. 309-316) shows that, indeed, this is the key 
+reaction in sensitivity (by a large margin…). That figure is a "sensitivity analysis" 
+which one can accomplish easily.
+So, if this is the case, is the authors' reduction framework working because the nonlinear system
+is really "simple" in its response to varying parameters}
+
+The complexity of the H$_2$/O$_2$ reaction kinetics application in the revised manuscript has been significantly
+enhanced by increasing the degree of uncertainty in the rate parameters, and considering additional 
+uncertainty in the
+initial pressure, temperature, and stoichiometry of the system. As a result, the non-linearity in the system
+response is observed to increase substantially as seen in Figures 3 and 6 unlike a straight line as 
+observed earlier. Our sensitivity analysis results in Figure 4 using the proposed framework based on the
+active subspace methodology are shown to be consistent with our earlier results using the derivative-based
+global sensitivity analysis (Ref. 35). The sensitivity results for the higher-dimensional problem in the
+revised manuscript reveal that the igntion delay is largely impacted by
+the uncertainty in the pre-exponents associated with reactions 1, 9, and the initial temperature. Additionally,
+we observe significant sensitivity towards the rate parameters associated with reaction 15 and the pre-exponent
+associated with the rate of reaction 17. 
+Therefore, depending upon the considered prior uncertainty in the individual rate parameters for the 19
+reactions, and the range of values of the initial conditions, the proposed framework helps identify the
+key contributors to the uncertainty in the ignition delay. The analysis accounts for individual 
+contributions of the parameters as well as due to their interactions with other parameters to the 
+output uncertainty in the system response. Through this relatively more complex
+problem, we have demonstrated the robustness of the proposed framework.
+
+As mentioned in Section 6
+(Summary and Discussion), the proposed framework is agnostic to the choice of the system and could be extended
+to other kinetics applications as long as the system response is continuously differentiable in the domain of
+the inputs and is therefore applicable to a wide variety of applications.
+
+\item \referee{Related to 1) above, it is of paramount importance that the authors consider, at least, 
+methane, for which I expect less of a spectral gap in the matrix C. Ideally, more complex fuels with 
+more nounced responses.
+In other words, if the method works well for a simple case, what is its value to the community?
+The authors must try this out on a much more complex hydrocarbon/oxygen system.}
+
+A more complex problem as suggested by the reviewer would manifest itself into an increased number of
+uncertain parameters and/or an increased amount of non-linearity in the system response. Therefore, while
+we did not implement the framework to a new system, we have significantly increased the non-linearity of the
+response by increasing the degree of uncertainty in the rate parameters. In addition to the rate parameters,
+we perturb the initial conditions to investigate their impact on the response individually as well as due to
+their interaction with the rate parameters. Hence, we have also increased the dimensionality of the problem.
+As expected by the reviewer, the resulting spectral gap between the first two eigenvalues is observed to
+decrease for this relative more complex scenario as shown in Figure~1 in the revised manuscript. 
+The proposed framework has been used effectively compute the active subspace, construct a low-dimensional
+surrogate, and perform a global sensitivity analysis in this case.
+
+\item \referee{The authors report ignition delay times of 0.1 s (?). See for example Fig. 7 and Tab. 2.
+Are these seconds (s)? If this is the case, what kind of conditions (temperature/pressure/stoichiometry)
+are the authors considering? They seem very unphysical if the yield an ignition delay time of O(0.1 s).
+I could not find any clarification of the initial conditions for the ignition calculations.
+This is again a symptom of the disconnect between the authors' aims (a novel method, potentially very useful to the combustion community) and the combustion community's aims (more understanding and tools that can be of help in combustion kinetics).}
+
+The nominal values of the initial conditions: pressure ($P_0$), temperature ($T_0$), and
+the equivalence ratio~($\Phi_0$) considered in the computations are 1~atm, 900~K, and 2.0
+respectively. These values have been reported in the revised manuscript in the begining of
+Section 5 on page 15.
+In this revision, we perturb the initial conditions to study their impact
+on ignition delay estimates. The mean estimate of the ignition delay using data at 10$^4$ random samples in
+the 36-dimensional space is found to be approximately 0.15 as reported in Table 2 on page 18 in the
+revised manuscript. As mentioned in the paper, we have used 
+the TChem software for model predictions and based on our recent communication with the developers at
+Sandia National Laboratories, these estimates seem reasonable.
+
+\noindent Minor comments by the reviewer have also been addressed in the revised manuscript.
+
+\end{enumerate}
 

### 2018-12-02 Revision/Paper/assa.tex
..

@@
 $\mu(d\vec{\xi}) = \pi(\vec{\xi})d\vec{\xi}$, where $\pi(\vec{\xi})$ is the joint probability
-distribution of $\vec{\xi}$. Herein, $f$ is assumed to be a square integrable 
+\rebut{density function} of $\vec{\xi}$. Herein, $f$ is assumed to be a square integrable 
 function with continuous partial 
@@
 The columns of $\mat{W}_1 = 
-\begin{bmatrix} \vec{w}_r \cdots \vec{w}_r\end{bmatrix}$ 
+\begin{bmatrix} \vec{w}_\rebut{1} \cdots \vec{w}_r\end{bmatrix}$ 
 span the dominant eigenspace of $\mat{C}$ and
@@
 T_i(f) = 1 - 
-\frac{\V[\mathbb{E}(f|\vec{\xi}_{\sim i})]}{\V(f)},
+\frac{\V_{\rebut{\vec{\xi}_{\sim i}}}[\mathbb{E}_{\rebut{\xi_i}}(f|\vec{\xi}_{\sim i})]}{\V(f)},
 \label{eq:total}
@@
 where $\vec{\xi}_{\sim i}$ is the input parameter vector with the  
-$i^\text{th}$ entry removed, and $\V$ denotes the variance. The total-effect Sobol' index accounts
+$i^\text{th}$ entry removed. \rebut{The expectation is taken over all possible
+values of $\xi_i$ and its variance is computed with respect to all parameters except $\xi_i$.}
+The quantity, $\V(f)$ denotes the total variance of the model output.
+ The total-effect Sobol' index accounts
 for the contribution of a given input to the variability in the output by itself

### 2018-12-02 Revision/Paper/intro.tex
..

@@
 conditions. In this study, computations were performed at fuel-rich conditions,
-$\Phi$ = 2.0. Homogeneous ignition at constant pressure is simulated using the
-TChem software package~\cite{Safta:2011} \rebut{using an initial pressure, $P_0$ = 1 atm and
-initial temperature, $T_0$ = 900~K.} The time required for the rate of
+$\Phi$~=~2.0. Homogeneous ignition at constant pressure is simulated using the
+TChem software package~\cite{Safta:2011} \rebut{using an initial pressure, $P_0$~=~1~atm and
+initial temperature, $T_0$~=~900~K.} The time required for the rate of
 temperature increase to exceed a given threshold, regarded as \emph{ignition delay}
@@
 are considered to be uniformly distributed about their nominal estimates provided
-in~\cite{Yetter:1991}. \rebut{The initial conditions are also considered to be uniformly
+in~\cite{Yetter:1991}. \rebut{Temperature exponent, $n_i$ for each reaction is fixed to
+its nominal value, also provided in~\cite{Yetter:1991}.
+The initial conditions are also considered to be uniformly
 distributed about their respective aforementioned values.} 

### 2018-12-02 Revision/Response/response.tex
..

@@
 
+\section*{Reviewer \#2}
+
+\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
+\item \referee{There is a lack of quantitative discussion of the expense of regression-vs-perturbation,
+in terms of the actual number of model simulations required. The work seems to suggest that perturbation
+works better than regression, but then it is a direct and clear consequence of the fact that local 
+approximation is supposed to be better than a global one. Ideally, much more quantitative comparison
+needs to be made here before reaching any meaningful conclusions.}
+
+As suggested by the reviewer, we have focused on a quantitative comparison in terms of the
+number of samples and associated model runs required to obtain a converged active subspace
+for the two approaches. In addition to convergence characteristics, we have also compared
+the two approaches for their predictive accuracy for the same computational effort for the
+19-dimensional problem as well as the 36-dimensional problem. In the revised work,
+we have increased the dimensionality of
+the higher-dimensional case from 33 to 36. Specifically, we have enhanced our analysis
+to study the impact of perturbations in the initial conditions: pressure, temperature, and
+stoichiometry on the system response i.e. ignition delay.
+
+\item \referee{Another fundamental flaw is related to the additive treatment of the uncertainty of 
+the pre-exponential factor A. In most works dealing with uncertainty in chemical kinetics, the
+uncertainty form is multiplicative or, equivalently, log(A) is the underlying fundamental parameter 
+with additive uncertainty range. In the current form of the uncertainty, the problem is too benign,
+and barely carries any practical usability for the practitioners.}
+
+The reviewer has made a valuable suggestion. Therefore, in the revised manuscript, we have
+updated the computations for the active subspace using perturbation and regression by considering
+the uncertainty in $\log(A)$ as opposed to $A$. As a result, the non-linearity of the system
+response is observed to increase significantly and the variability of the ignition delay in the
+active subspace is no longer a straight line as observed earlier. Hence, we have substantially
+increased the complexity of the problem by increasing the non-linearity in the system response
+as well as its dimensionality to illustrate the robustness and the applicability of the proposed framework
+to more complex scenarios.
+
+\item \referee{Why not use the word 'active' in the title, since active subspace is essentially
+what has been done here?}
+
+The title has been updated as ``Active subspace-based dimension reduction for chemical kinetics
+applications with epistemic uncertainty".
+
+\item \referee{3.36 $n_i$ is index or power? Is this an established term for the power of T? 
+Also, what are the values of $n_i$'s?}
+
+In the revised manuscript, we denote $n_i$ as the `temperature exponent' as suggested by
+the other reviewer. Yes, it is an established quantity and its value in addition to the
+rate constants for the 19 reactions
+in the H$_2$/O$_2$ mechanism is taken from [Yetter et. al., 1991] and implemented in the
+TChem software used for model predictions in this work. We have clarified this on Page 4
+in the revised manuscript.
+
+\item \referee{3.51 'rate of temperature' or 'temperature'?}
+
+The ignition delay is considered as the time required for the rate of temperature increase to
+exceed a threshold in this work.
+
+\item \referee{4.57 not clear where the number 33 comes from since we have 19 reactions. I see it is explained at the beginning of Section 5, but it should be noted here briefly, too.}
+
+We could not locate the number 33 in the location specified by the reviewer. Nevertheless, we would like
+to point out that on Page 4 (just below Table 1) in the introduction, 
+it is mentioned that the activation energies with a non-zero nominal value are considered as uncertain. 
+Since the uncertainty in the activation energy is considered only in the higher-dimensional case
+presented in Section 5, we have
+specified the reactions for which it is 0 in the begining of that section.
+
+\item \referee{6.19 distribution->density function}
+
+The revised manuscript has been updated to include the reviewer's suggestion.
+
+\item \referee{6.38: typo, $w_r$ should be $w_1$}
+
+The typo has been fixed in the revised manuscript.
+
+\item \referee{7.23-27: why 'statistical independence' and not just 'independence'? Is there any other 
+type of independence?}
+
+\item \referee{7.29-30: Any PDF can be written in a Boltzmann form, V(x) is just the negative log-PDF, 
+I am not sure if it is necessary to bring up a Boltzmann distribution language here.
+Also, it could be confused with Maxwell-Boltzmann distribution, which is a very specific PDF form.}
+
+\item \referee{7.40: Explain what E and V are with respect to in Eq (8).}
+
+The equation as well as the following descritption of the expectation and the variance
+has been updated to address the reviewer's comment in the revised manuscript.
+
+\end{enumerate}
+
 \end{document}

### 2018-12-02 Revision/Response/response.tex
..

@@
 
-The complexity of the H$_2$/O$_2$ reaction kinetics application in the revised manuscript has been significantly
+\noindent The complexity of the H$_2$/O$_2$ reaction kinetics application in the revised manuscript has been significantly
 enhanced by increasing the degree of uncertainty in the rate parameters, and considering additional 
@@
 
-A more complex problem as suggested by the reviewer would manifest itself into an increased number of
+\noindent  A more complex problem as suggested by the reviewer would manifest itself into an increased number of
 uncertain parameters and/or an increased amount of non-linearity in the system response. Therefore, while
@@
 
-The nominal values of the initial conditions: pressure ($P_0$), temperature ($T_0$), and
+\noindent  The nominal values of the initial conditions: pressure ($P_0$), temperature ($T_0$), and
 the equivalence ratio~($\Phi_0$) considered in the computations are 1~atm, 900~K, and 2.0
@@
 
-As suggested by the reviewer, we have focused on a quantitative comparison in terms of the
+\noindent  As suggested by the reviewer, we have focused on a quantitative comparison in terms of the
 number of samples and associated model runs required to obtain a converged active subspace
@@
 
-The reviewer has made a valuable suggestion. Therefore, in the revised manuscript, we have
+\noindent The reviewer has made a valuable suggestion. Therefore, in the revised manuscript, we have
 updated the computations for the active subspace using perturbation and regression by considering
@@
 
-The title has been updated as ``Active subspace-based dimension reduction for chemical kinetics
+\noindent  The title has been updated as ``Active subspace-based dimension reduction for chemical kinetics
 applications with epistemic uncertainty".
@@
 
-In the revised manuscript, we denote $n_i$ as the `temperature exponent' as suggested by
+\noindent  In the revised manuscript, we denote $n_i$ as the `temperature exponent' as suggested by
 the other reviewer. Yes, it is an established quantity and its value in addition to the
@@
 
-The ignition delay is considered as the time required for the rate of temperature increase to
+\noindent  The ignition delay is considered as the time required for the rate of temperature increase to
 exceed a threshold in this work.
@@
 
-We could not locate the number 33 in the location specified by the reviewer. Nevertheless, we would like
+\noindent  We could not locate the number 33 in the location specified by the reviewer. Nevertheless, we would like
 to point out that on Page 4 (just below Table 1) in the introduction, 
@@
 
-The revised manuscript has been updated to include the reviewer's suggestion.
+\noindent  The revised manuscript has been updated to include the reviewer's suggestion.
 
@@
 
-The typo has been fixed in the revised manuscript.
+\noindent  The typo has been fixed in the revised manuscript.
 
@@
 
-The equation as well as the following descritption of the expectation and the variance
-has been updated to address the reviewer's comment in the revised manuscript.
+\noindent  The equation as well as the following description of the expectation (E) and the variance (V)
+has been updated to address the reviewer's comment, in the revised manuscript.
+
+\item \referee{7.50: Just say in words that it is the i-th diagonal element of matrix C.}
+
+\noindent Following the reviewer's suggestion, we have updated the text in the revised manuscript to
+clearly note that the DGSM $\nu_i(f)$ is the $i^{\text{th}}$ diagonal element in the matrix, $\mathbf{C}$
+on Page 8.
+
+\item \referee{10.56: what if there is no eigenvalue gap of O(100)?}
+
+\noindent A large gap in the eigenvalue spectrum typically indicates that a low-dimensional (1--2) active subspace 
+would sufficiently capture the variability in the model output due to uncertainty in the inputs. However,
+in the revised manuscript, we have shown that the variability in the ignition delay is reasonably
+approximated with a 1-dimensional active subspace even though the gap in the first two eigenvalues
+is found to be O(10). Hence, a smaller gap in the eigenvalue spectrum might still lead to a reasonable
+accuracy with a low-dimensional active subspace depending upon the relationship between the
+system response and the inputs. However, it is also possible that we might need to a consider a higher 
+dimensional active subspace to be able to approximate the variability in the response with reasonable
+accuracy in situations where the gap is small.
+
+\item \referee{11.33: Alg. 1: should have \beta as input, too?}
+
+\noindent The reviewer is correct. Algorithm 1 in the revised manuscript has been updated to include $\beta$
+as an input.
+
+\noindent
 

### 2018-12-02 Revision/Paper/assa.tex
..

@@
 \]
-Here $\pi$ is the joint PDF of $\vec\xi$. 
-Let $\mat{C}$ be as in~\eqref{eq:C}, and consider the spectral decomposition written 
+Here $\pi$ is the joint PDF of $\vec\xi$. \rebut{Note that $\nu_i(f)$ is the 
+$i^{\text{th}}$ diagonal element of the matrix $\mat{C}$} as defined in~\eqref{eq:C}. 
+Consider the spectral decomposition written 
 as $\mat{C} = \sum_{k=1}^\Np \lambda_k \vec{w}_k \vec{w}_k^\top$. Herein, we use the notation 

### 2018-12-03 Revision/Response/response.tex
..

@@
 
-\item \referee{11.33: Alg. 1: should have \beta as input, too?}
+\item \referee{11.33: Alg. 1: should have $\beta$ as input, too?}
 
@@
 
-\noindent
+\item \referee{11.42: Which norm is used? Any intuition why squares of W are used, and not just W?}
+
+\noindent We have used the L$^2$ norm to estimate the relative change in the squared values of corresponding
+eigenvector components between successive iterations. The algorithm in the revised manuscript has been updated 
+to specify the norm. 
+We consider square of individual components due to the fact the both W and -W contain legitimate eigenvectors 
+of C. Alternatively, we could have also used the modulus value of the components to assess convergence.
+
+\item \referee{11.45: what is max over?}
+
+\noindent The max is over all eigenvectors in the active subspace denoted by $j$. This has been specified in
+the updated algorithm in the revised manuscript.
+
+\item \referee{12.29: 'higher-order' should be replaced with higher-index or smaller-magnitude, to avoid confusion.}
+
+\noindent The reviewer's suggestion has been incorporated in the revised manuscript.
 

### 2018-12-04 Revision/Response/response.tex
..

@@
 %{\it Journal of Scientific Computing}.
+
+We are extremely thankful to the reviewers for recognizing potential in this effort and 
+for their constructive feedback on the manuscript. A point-by-point
+response to the comments has been provided below. Corresponding modifications have been
+highlighted (in blue) in the revised manuscript. Major changes in the revised manuscript
+are summarized as follows:
+
+\begin{itemize}
+
+\item The complexity of the problem has been increased substantially in two ways: First, 
+the non-linearity of the system response has been signficantly increased by increasing
+the perturbations associated with the uncertain rate-controlling parameters: pre-exponent
+and the activation energy. Second, in addition to the rate parameters, the impact of
+perturbing the initial conditions: pressure, temperature, and stoichiometry pertaining to the
+H$_2$/O$_2$ reaction mechanism on the ignition delay is investigated. Such analysis using the
+proposed framework would provide useful insights to combustion scientists interested in
+kinetics applications. Moreover, it illustrates the applicability and robustness of the 
+framework to complex kinetics applications.
+
+\item The two approaches, perturbation and regression, for estimating the model gradient
+with respect to uncertain parameters and inputs are compared for convergence behavior and
+predictive accuracy using the same amount of
+computational effort in the revised manuscript, as suggested by one of the reviewers.
+
+\end{itemize}
+
+We sincerely hope that based on the modifications in the revised manuscript and our 
+response to reviewers' comments, the paper is found suitable for publication in
+{\it Combustion and Flame}.
 
@@
 
+\item \referee{Figs 2a, 3 and 4 demonstrate the regression-based gradient computation is working
+similarly to the perturbation based one. However, the comparison is really qualitative and does
+ not provide much insight. Clearly, regression will approximate the gradient less accurately then
+a local perturbation, but the latter has a price associated with it as new simulations are needed
+(a factor of d+1). It would have been great to see how accurate each method is for comparable
+expense, or vice-verse, how expensive each approach is for a comparable accuracy. I know there
+is a statement about this at the end of Section 4, but it is neither sufficient nor quantitative.
+Same goes for statements in Section 5.2 - it was not clear whether regression- and
+perturbation- based approaches had comparable expenses (in terms of model evaluations) or not.}
+
+\noindent Based on the reviewer's feedback, in the revised manuscript, the analysis has been
+updated to
+compare the perturbation and regression approaches for their convergence and predictive 
+accuracy using the same amount of computational effort. As a result, the comparative 
+analysis is much more conclusive with regards to the performance of the two approaches.
+
+\item \referee{For Figure 8b, the reasoning why E$_15$ was not captured by the regression approach,
+is not satisfactory. I think the authors could have gone a bit deeper in order to explain this
+glaring miss of the regression-based approach. There is some discussion in Section 6, but it is
+equally unsatisfactory.}
+
+\noindent Our latest analysis in the revised manuscript has revealed that the two approaches
+in fact yield consistent results
+for sensitivity analysis even for the 36-dimensional case with increased non-linearity
+when using the same amount of
+computational effort or the same number of model evaluations. Earlier, the two approaches were
+compared for the same number of samples, N. However, the number of model runs in the case of
+perturbation were N(d+1); d is the number of uncertain parameters and inputs to the model. Whereas,
+the number of model runs in the case of regression were only N. Therefore, it can be said that
+the accuracy of the regression-based approximation of the gradient can be increased to the 
+extent of perturbation-based approach by increasing the sample size.
+
+\item \referee{14.4-8: the near-linearity simply makes the problem extremely benign, nearly any
+reasonable method would have captured such behavior. It would have been great to consider more
+complicated case (more than simply going from 19 to 33 parameters), perhaps looking at logA 
+perturbation instead of A would help? Also consider more than 1$\%$ 
+perturbations to the input. Otherwise, the problem is too simple and does not do justice to the method.}
+
+\noindent We have made a sincere effort to incorporate these valuable suggestions by the 
+reviewer in the revised manuscript. 
+Specifically, we perturb $\log(A)$ instead of $A$ which is observed to significantly increase
+the non-linearity of the system response. In addition the rate parameters, we analyse the impact
+of perturbing the initial conditions. Moreover, as suggested by the reviewer, we have
+increased the mangnitude of the perturbations from 1$\%$ to 2$\%$ to illustrate the applicability
+and robustness of the proposed approach to more complex scenarios.
+
+\item \referee{14.8: here and in a few other places, I have an issue with statements like
+active subspace is 1-dimensional'. To me, active subspace is as many dimensional as one likes,
+so the statement should be changed to something like '1-dimensional active subspace captures
+most of the variance/uncertainty/dynamics etc..', or '1-dimensional active subspace provides
+very accurate approximation...'.}
+
+\noindent The reviewer's suggestion has been incorporated in the revised manuscript.
+
+\item \referee{14.57: even if nominal is zero, one could perturb them, right? 
+Or it is actually indicating the fact that there should be no exponential factor whatsoever.}
+
+\noindent Based on our conversations with the developers of the TChem software at Sandia
+National Laboratories, we decided not to perturb the activation energies with a 0
+nominal value.
+
+\item \referee{18.35: the remark in the parentheses did not make sense to me.}
+
+\noindent We have provided examples for perturbation-based techniques such as finite difference and
+adjoint-based methods for estimating the
+gradient of the model output with respect to uncertain parameters and inputs.
+
 \end{enumerate}

### 2018-12-04 Revision/Response/response.tex
..

@@
 key contributors to the uncertainty in the ignition delay. The analysis accounts for individual 
-contributions of the parameters as well as due to their interactions with other parameters to the 
+contributions of the parameters as well as contributions due to their interactions with other parameters to the 
 output uncertainty in the system response. Through this relatively more complex
@@
 decrease for this relative more complex scenario as shown in Figure~1 in the revised manuscript. 
-The proposed framework has been used effectively compute the active subspace, construct a low-dimensional
+The proposed framework has been used effectively to compute the active subspace, construct a low-dimensional
 surrogate, and perform a global sensitivity analysis in this case.
@@
 the uncertainty in $\log(A)$ as opposed to $A$. As a result, the non-linearity of the system
-response is observed to increase significantly and the variability of the ignition delay in the
+response is observed to increase significantly. As shown in Figures~3 and 6, 
+the variability of the ignition delay in the
 active subspace is no longer a straight line as observed earlier. Hence, we have substantially
@@
 as well as its dimensionality to illustrate the robustness and the applicability of the proposed framework
-to more complex scenarios.
+for a wide range of kinetics applications.
 
@@
 
-\noindent  The ignition delay is considered as the time required for the rate of temperature increase to
-exceed a threshold in this work.
+\noindent  The ignition delay is considered as the time required for the {\it rate of temperature}
+increase to exceed a threshold, in this work.
 
@@
 
-\item \referee{6.19 distribution->density function}
+\item \referee{6.19 distribution-$>$density function}
 
@@
 \noindent  The equation as well as the following description of the expectation (E) and the variance (V)
-has been updated to address the reviewer's comment, in the revised manuscript.
+has been updated to address the reviewer's comment on Page~7 in the revised manuscript.
 
@@
 Specifically, we perturb $\log(A)$ instead of $A$ which is observed to significantly increase
-the non-linearity of the system response. In addition the rate parameters, we analyse the impact
+the non-linearity of the system response. In addition to the rate parameters, we analyse the impact
 of perturbing the initial conditions. Moreover, as suggested by the reviewer, we have

### 2018-12-05 Revision/Response/response.tex
..

@@
 
-\noindent Minor comments by the reviewer have also been addressed in the revised manuscript.
+\item \referee{p.3, line 36. I would call this the ``temperature exponent".}
+
+\noindent The reviewer's suggestion has been incorporated in the revised manuscript on Page 3.
+
+\item \referee{p.6, line 32. This is a spectral gap. Need to give credit to the CSP community, 
+although they consider spectral gaps in the nonlinear system, while here the authors consider spectral 
+gaps in C, which is related to the system's response to varying model parameters.}
+
+\item \referee{p. 16, Fig. 6. y-axis label. Probably $\lambda_i/\lambda_0$ as goes to 1 for i=1? Typo?}
+
+It is in fact $\log(\lambda_i/\lambda_0)$. The typo has been corrected.
+
+%\noindent Minor comments by the reviewer have also been addressed in the revised manuscript.
 

### 2018-12-05 Revision/Response/response.tex
..

@@
 \item The complexity of the problem has been increased substantially in two ways: First, 
-the non-linearity of the system response has been signficantly increased by increasing
+the non-linearity of the system response has been significantly increased by increasing
 the perturbations associated with the uncertain rate-controlling parameters: pre-exponent
-and the activation energy. Second, in addition to the rate parameters, the impact of
+and the activation energy. Second, the dimensionality of the problem has also been increased
+significantly. Specifically, in addition to the rate parameters, the impact of
 perturbing the initial conditions: pressure, temperature, and stoichiometry pertaining to the
@@
 global sensitivity analysis (Ref. 35). The sensitivity results for the higher-dimensional problem in the
-revised manuscript reveal that the igntion delay is largely impacted by
+revised manuscript reveal that the ignition delay is largely impacted by
 the uncertainty in the pre-exponents associated with reactions 1, 9, and the initial temperature. Additionally,
@@
 methane, for which I expect less of a spectral gap in the matrix C. Ideally, more complex fuels with 
-more nounced responses.
+more nuanced responses.
 In other words, if the method works well for a simple case, what is its value to the community?
@@
 the equivalence ratio~($\Phi_0$) considered in the computations are 1~atm, 900~K, and 2.0
-respectively. These values have been reported in the revised manuscript in the begining of
+respectively. These values have been reported in the revised manuscript in the beginning of
 Section 5 on page 15.
@@
 presented in Section 5, we have
-specified the reactions for which it is 0 in the begining of that section.
+specified the reactions for which it is 0 in the beginning of that section.
 
@@
 Specifically, we perturb $\log(A)$ instead of $A$ which is observed to significantly increase
-the non-linearity of the system response. In addition to the rate parameters, we analyse the impact
+the non-linearity of the system response. In addition to the rate parameters, we analyze the impact
 of perturbing the initial conditions. Moreover, as suggested by the reviewer, we have
-increased the mangnitude of the perturbations from 1$\%$ to 2$\%$ to illustrate the applicability
+increased the magnitude of the perturbations from 1$\%$ to 2$\%$ to illustrate the applicability
 and robustness of the proposed approach to more complex scenarios.

### 2018-12-05 Revision/Response/response.tex
..

@@
 compared for the same number of samples, N. However, the number of model runs in the case of
-perturbation were N(d+1); d is the number of uncertain parameters and inputs to the model. Whereas,
-the number of model runs in the case of regression were only N. Therefore, it can be said that
+perturbation was N(d+1); d is the number of uncertain parameters and inputs to the model. Whereas,
+the number of model runs in the case of regression was only N. Therefore, it can be said that
 the accuracy of the regression-based approximation of the gradient can be increased to the 

### 2018-12-06 Revision/Paper/method.tex
..

@@
   \begin{algorithmic}[1]
-\Require $\theta_l$, $\theta_u$, $\tau$. 
+\Require $\theta_l$, $\theta_u$, \rebut{$\beta$}, $\tau$. 
 \Ensure $\hat{\mat{\Lambda}}$, $\hat{\mat{W}}$, $\bm{\nu}_r(f)$ %$\eta$. %
@@
                        \scalebox{1.25}{$\frac{\|(\hat{\mat{W}}_{1,j}^{k})^2 - 
-                       (\hat{\mat{W}}_{1,j}^{k-1})^2\|}{\|(\hat{\mat{W}}_{1,j}^{k-1})^2\|}$}, 
+                       (\hat{\mat{W}}_{1,j}^{k-1})^2\|_\rebut{2}}{\|(\hat{\mat{W}}_{1,j}^{k-1})^2\|_\rebut{2}}$}, 
                        $j = 1,\ldots,r$.
-		\If {$\max\left(\delta \hat{\mat{W}}_{1,j}^{(k)}\right)<\tau$}
+		\If {$\max\limits_{\rebut{j}}\left(\delta \hat{\mat{W}}_{1,j}^{(k)}\right)<\tau$}
 			\State break
@@
 are approximated 
-reasonably well with just 20 samples. As expected, the accuracy of higher-order eigenvalues is observed
+reasonably well with just 20 samples. As expected, the accuracy of higher-\rebut{index}
+ eigenvalues is observed
 to improve with the sample size. Since 
@@
 %
-\rebut{The quantity, $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ is plotted in
-Figure~\ref{fig:comp} (center) to illustrate convergence characteristics based on Algorithm~\ref{alg:grad} 
-for the two approaches. As observed for a set tolerance $\tau$~=~0.05, 8 iterations are needed 
-to converge in both cases. 
-At each iteration, 5 new random samples were added for the perturbation approach. As discussed earlier, gradient 
-computation (using finite difference) at 5 samples requires 100 model runs. Hence, in order to compare the
-two approaches for the same computational effort, we generate 100 random samples in the 19D input domain for 
-the regression approach. Therefore, 8 iterations require 800 model runs leading to a total of 900 model runs 
-including the initial step in both cases.  The accuracy of the two approaches is compared for the same 
-computational effort by computing $\varepsilon_{\text{L}^{2}}^{N-n}$ using~\eqref{eq:accu} in 
-Figure~\ref{fig:comp} (right). Once again, $N$
-corresponds to the case of active subspace computation using the perturbation approach using 1000 samples
-in the 19D domain; $n$ ranges from 5 to 45 for the perturbation approach, and 100 to 900 for the regression 
-approach corresponding to iterations 1 to 8. The accuracy of the two approaches is found to be comparable for
-the same amount of computational effort. However, the perturbation approach is observed to be slightly more
-accurate.}
+\rebut{The quantity, $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ defined in Algorithm~\ref{alg:grad} was
+used to assess the convergence behavior of the two approaches. Using a set tolerance, $\tau$~=~0.05,
+it was observed that both, perturbation and regression approaches took 8 iterations to converge. 
+Note that the computational effort at each iteration was considered to be the same in both cases.
+More specifically, 5 new random samples were added for the perturbation approach at each iteration.
+However, as discussed earlier, as total of 100 (=5$\times$(19+1)) model runs were needed to obtain the
+model prediction and its gradients at these newly generated samples. Hence, in the case of
+regression, 100 new random samples were generated at each iteration since gradient computation does
+not require additional model runs in this case. Thus, including the initial step, a total of 900
+model runs were required to obtained a converged active subspace in both cases.
+
+The accuracy of the two approaches was assessed by estimating $\varepsilon_{\text{L}^{2}}^{N-n}$ 
+using the components of the dominant eigenvector in the converged active subspace in each case
+in~\eqref{eq:accu}. The quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$ was estimated to be 0.0657
+and 0.1050 using perturbation and regression respectively. Hence, the regression approach was
+found to be relatively more accurate. Squared values of the individual components of the 
+dominant eigenvector from the
+two approaches and for the case using $N$~=~1000 in the perturbation approach are plotted in
+Figure~\ref{fig:comp}.}
+%
+%is plotted in
+%Figure~\ref{fig:comp} (center) to illustrate convergence characteristics based on Algorithm~\ref{alg:grad} 
+%for the two approaches. As observed for a set tolerance $\tau$~=~0.05, 8 iterations are needed 
+%to converge in both cases. 
+%At each iteration, 5 new random samples were added for the perturbation approach. As discussed earlier, gradient 
+%computation (using finite difference) at 5 samples requires 100 model runs. Hence, in order to compare the
+%two approaches for the same computational effort, we generate 100 random samples in the 19D input domain for 
+%the regression approach. Therefore, 8 iterations require 800 model runs leading to a total of 900 model runs 
+%including the initial step in both cases.  The accuracy of the two approaches is compared for the same 
+%computational effort by computing $\varepsilon_{\text{L}^{2}}^{N-n}$ using~\eqref{eq:accu} in 
+%Figure~\ref{fig:comp} (right). Once again, $N$
+%corresponds to the case of active subspace computation using the perturbation approach using 1000 samples
+%in the 19D domain; $n$ ranges from 5 to 45 for the perturbation approach, and 100 to 900 for the regression 
+%approach corresponding to iterations 1 to 8. The accuracy of the two approaches is found to be comparable for
+%the same amount of computational effort. However, the perturbation approach is observed to be slightly more
+%accurate.}
+%
+%\begin{figure}[htbp]
+% \begin{center}
+%  \includegraphics[width=0.8\textwidth]{./Figures/eig_convD19}
+%\caption{\rebut{Left: An illustrative comparison of individual squared components of the converged dominant
+%eigenvector obtained using perturbation and regression strategies using $M$~=~900 model runs in each case. 
+%Additionally, the dominant eigenvector components obtained using $M$~=~20000 model runs (corresponding to 
+%$N$~=~1000 samples)
+%in the perturbation strategy (test case), used to assess the accuracy of the two strategies are also plotted.
+%Center: The quantity,  $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$
+%is plotted for successive iterations to illustrate convergence behavior for the two strategies. Right: The
+%quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$ is plotted as a function of iterations to compare the accuracy
+%of the two strategies. Note that the above plots compare the convergence and accuracy of the two
+%strategies for the same amount of computational effort, quantified by the number of model runs, $M$.}}
+%\label{fig:comp}
+%\end{center}
+%\end{figure}
 %
@@
  \begin{center}
-  \includegraphics[width=0.8\textwidth]{./Figures/eig_convD19}
-\caption{\rebut{Left: An illustrative comparison of individual squared components of the converged dominant
+  \includegraphics[width=0.45\textwidth]{./Figures/eig_conv19D}
+\caption{\rebut{An illustrative comparison of individual squared components of the converged dominant
 eigenvector obtained using perturbation and regression strategies using $M$~=~900 model runs in each case. 
@@
 $N$~=~1000 samples)
-in the perturbation strategy (test case), used to assess the accuracy of the two strategies are also plotted.
-Center: The quantity,  $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$
-is plotted for successive iterations to illustrate convergence behavior for the two strategies. Right: The
-quantity, $\varepsilon_{\text{L}^{2}}^{N-n}$ is plotted as a function of iterations to compare the accuracy
-of the two strategies. Note that the above plots compare the convergence and accuracy of the two
-strategies for the same amount of computational effort, quantified by the number of model runs, $M$.}}
+in the perturbation strategy (test case), used to assess the accuracy of the two strategies are also plotted.}}
 \label{fig:comp}
@@
 %
-\rebut{In Figure~\ref{fig:comp} (left), we compare the
-squared values of individual components of the dominant eigenvector obtained using the two approaches
-at the end of 8 iterations in each case. Additionally, the dominant eigenvector components obtained
-for a test case involving $N$~=~1000 (and $M$ = 20,000 model runs)
-samples using the perturbation approach are also plotted. 
+\rebut{
 The set of eigenvector components for the three cases are found to be in excellent agreement with

### 2018-12-07 Revision/Response/response.tex

@@
 
-\baselineskip=22pt
+%\baselineskip=22pt
 
@@
 
-\noindent  The nominal values of the initial conditions: pressure ($P_0$), temperature ($T_0$), and
+\noindent  Depending upon the initial conditions, the ignition delay can range from $\mathcal{O}(10^{-3})$
+to $\mathcal{O}(10^3)$ seconds for hydrogen-air mixtures as shown by Starostin et al. in~\cite{Starostin:2011}.
+In this work, the initial pressure ($P_0$), temperature ($T_0$), and
 the equivalence ratio~($\Phi_0$) considered in the computations are 1~atm, 900~K, and 2.0
@@
 National Laboratories, we decided not to perturb the activation energies with a 0
-nominal value.
+nominal value since the exponential factor is not considered for these reactions, as mentioned
+by the reviewer.
 
@@
 
+\bibliographystyle{elsarticle-num}
+\bibliography{REFER}
+
 \end{document}

### 2018-12-08 Revision/Paper/app.tex
..
...

@@
 
-For the high-dimensional case, we aim to investigate the impact of the uncertain
-pre-exponents ($A_i$'s) as well as the activation energies ($E_{a,i}$'s) \rebut{and
-the initial conditions: pressure~($P_0$), temperature~($T_0$), and stoichiometry based on
-the equivalence ratio~($\Phi_0$)} on ignition delay during the H$_2$/O$_2$ reaction. 
-\rebut{The $\log(A_i)$'s, $E_{a,i}$'s for all reactions except 
-$\mathcal{R}_6$ -- $\mathcal{R}_9$, $\mathcal{R}_{13}$ (due to 
-zero nominal values for $E_a$), and the initial conditions were perturbed by
-2$\%$ about their nominal values. Note that the mangnitude of the perturbation
-was selected such that the ignition delay remains positive throughout the input
-domain. The nominal values of the rate parameters, $A_i$'s and $E_{a,i}$'s 
+For the high-dimensional case, we aim to investigate the impact of 
+uncertainty in the following problem parameters on 
+the ignition delay associated with the H$_2$/O$_2$ reaction:
+(i) pre-exponents ($A_i$'s); (ii) the activation energies
+($E_{a,i}$'s); and (iii) \rebut{the initial pressure~($P_0$),
+temperature~($T_0$), and stoichiometry~($\Phi_0$)}. 
+%
+\rebut{The $\log(A_i)$'s, $E_{a,i}$'s for all reactions except $\mathcal{R}_6$
+-- $\mathcal{R}_9$, $\mathcal{R}_{13}$ (due to zero nominal values for $E_a$),
+and the initial conditions were perturbed by 2$\%$ about their nominal values.
+Note that the mangnitude of the perturbation was selected such that the
+ignition delay assumes a physically meaningful value in the input domain. 
+The nominal values of the rate parameters, $A_i$'s and $E_{a,i}$'s 
 were taken from~\cite{Yetter:1991}. The nominal values of $P_0$, $T_0$, and

### 2018-12-08 Revision/Paper/intro.tex
...

@@
 activity scores for functions of independent random inputs whose distribution
-law belongs to the class of Boltzmann probability measures. 
+law belongs to a \rebut{broad class of probability measures.} 
 It is worth mentioning that computing global sensitivity measures provides 

### 2018-12-09 Revision/Response/response.tex

@@
 Also, it could be confused with Maxwell-Boltzmann distribution, which is a very specific PDF form.}
+Thanks for this comment. We followed the theory in ref [27] in the manuscript, 
+where the authors derive Poincare type inequalities for functions of 
+iid random inputs distributed according to Boltzmann probability laws. We point
+out that the uniform distribution does not belong to this class. That was the reason 
+for assuming the distribution laws of the
+random inputs are either \emph{uniform} or belong to 
+the Boltzmann class, in section 3. An important class of Boltzmann measures are 
+the so called log-concave distributions, such as normal, exponential, and beta, and gamma
+distributions.
+
+We share the reviewer's concern about the potential confusions some readers might face when 
+seeing the mention of Boltzmann. We have removed the mention of Boltzmann measures
+from the introduction, and now only refer to this in Section 3, where the theoretical
+results are discussed.
 

### 2018-12-09 Revision/Paper/app.tex
..

@@
 \rebut{The mean and the standard deviation estimates obtained using the model and the 1-dimensional surrogates
-are found to be in close agreement. It is interesting to note that estimates from the regression approach are
-observed to be slightly more accurate. This is due to the fact that the converged active subspace in this case is 
+are found to be in close agreement. Hence, although the relative L$^2$ norm of the discrepancy~($\varepsilon_d$)
+is found to be considerably large, the uncertainty in the ignition delay is captured by both 
+approaches with reasonable accuracy. 
+It is however interesting to note that estimates from the regression approach are
+observed to be slightly more accurate. This is potentially due to the fact that the 
+converged active subspace in this case is 
 based on 1850 samples in the 36-dimensional input domain, whereas, the converged active subspace based the 

### 2018-12-09 Revision/Paper/method.tex
..

@@
 two approaches and for the case using $N$~=~1000 in the perturbation approach are plotted in
-Figure~\ref{fig:comp}.}
+Figure~\ref{fig:comp}~(left).}
 %
@@
  \begin{center}
-  \includegraphics[width=0.45\textwidth]{./Figures/eig_conv19D}
-\caption{\rebut{An illustrative comparison of individual squared components of the converged dominant
+  \includegraphics[width=0.38\textwidth]{./Figures/eig_conv19D}
+  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp19D}
+\caption{\rebut{Left: An illustrative comparison of individual squared components of the converged dominant
 eigenvector obtained using perturbation and regression strategies using $M$~=~900 model runs in each case. 
@@
 $N$~=~1000 samples)
-in the perturbation strategy (test case), used to assess the accuracy of the two strategies are also plotted.}}
+in the perturbation strategy (test case), used to assess the accuracy of the two strategies are also plotted.
+Right: An illustrative comparison of the SSPs generated using the
+perturbation and the regression strategies for computing the active subspace. 
+}}
 \label{fig:comp}
@@
 regression-based gradient estimates are
-compared in Figure~\ref{fig:comp_ssp}.
+compared in Figure~\ref{fig:comp}~(right).
 %
-\begin{figure}[htbp]
- \begin{center}
-  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp19D}
-\caption{\rebut{An illustrative comparison of the SSPs generated using the 
-perturbation and the regression strategies for computing the active subspace.}}
-\label{fig:comp_ssp}
-\end{center}
-\end{figure}
+%\begin{figure}[htbp]
+% \begin{center}
+%  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp19D}
+%\caption{\rebut{An illustrative comparison of the SSPs generated using the 
+%perturbation and the regression strategies for computing the active subspace.}}
+%\label{fig:comp_ssp}
+%\end{center}
+%\end{figure}
 %

### 2018-12-09 Revision/Paper/intro.tex

@@
 first approach uses finite difference to estimate the gradient and is regarded as
-the perturbation approach, whereas, the second approach involves a linear
-regression-fit to the available set of model evaluations and is regarded as
-the regression approach.
+the perturbation approach. On the other hand, the second approach 
+follows from~\cite{Constantine:2015} and involves repeated
+linear regression-fits to a subset of available model evaluations, and is 
+regarded as the regression approach in this work.
 

### 2018-12-09 Revision/Response/response.tex
the two comments ...

--- 2018-12-07 Revision/Response/response.tex
@@
 
-\item \referee{p.6, line 32. This is a spectral gap. Need to give credit to the CSP community, 
-although they consider spectral gaps in the nonlinear system, while here the authors consider spectral 
-gaps in C, which is related to the system's response to varying model parameters.}
+\item \referee{p.6, line 32. This is a spectral gap. Need to give credit to the
+CSP community, although they consider spectral gaps in the nonlinear system,
+while here the authors consider spectral gaps in C, which is related to the
+system's response to varying model parameters.}
+
+Thanks for this comment. As the referee notes, while the spectral 
+truncations in active subspace methodology resemble aspects of CSP analysis,
+the goals and approaches of these methods are substantially different. However, 
+we are aware that CSP analysis is well-known in computational
+chemistry community, and therefore, have included a brief mention 
+of CSP in the introduction, clarifying the differences between 
+the active subspace method and CSP.
 
@@
 Also, it could be confused with Maxwell-Boltzmann distribution, which is a very specific PDF form.}
+Thanks for this comment. We followed the theory in ref [35] 
+in the manuscript, where the authors derive Poincare type 
+inequalities for functions of iid random
+inputs distributed according to Boltzmann probability laws. 
+We point out that the uniform distribution does not belong to this class. 
+The required Poincare type inequality for the case of uniform 
+distribution was derived in an earlier paper of Sobol' and Kucherenko (ref [34]). 
+That
+was the reason for assuming the distribution laws of the random inputs are
+either \emph{uniform} or belong to the Boltzmann class, in section 3. An
+important class of Boltzmann measures are the so called log-concave
+distributions such as Normal, Exponential, Beta, and Gamma distributions. These
+points have been clarified in the revised manuscript.
+
+Additionally, we share the reviewer's concern about the potential confusions some readers
+might face when seeing the mention of Boltzmann. We have removed the mention of
+Boltzmann measures from the introduction, and now only refer to Boltzman
+measures in Section
+3, where the theoretical results are discussed.
 

### 2018-12-09 Revision/Paper/assa.tex
...
minor ..
the two comments ...

@@
 where $V$ is a continuous function and $C$ a normalization 
-constant~\cite{Lamboni:2013}.
+constant~\cite{Lamboni:2013}. \rebut{
+An important class of Boltzmann distributions are the so called log-concave
+distributions, which include Normal, Exponential, Beta, Gamma, Gumbel, and
+Weibull distributions. Note also that the uniform distribution does not fall under
+the class of Boltzmanm distributions.~\cite{Lamboni:2013}.}
 

### 2018-12-09 Revision/Paper/intro.tex
the two comments ...

@@
 the dimensionality of the problem, and hence the computational effort
-associated with uncertainty propagation.
+associated with uncertainty propagation. The focus here is on input parameter
+dimension reduction. \rebut{This is different from 
+techniques such at Computational
+Singular Perturbation (CSP)~\cite{GoussisLam92,FotacheKreutzLaw97,
+Lam85,LamGoussis89,ValoraniCretaGoussisEtAl06,
+DebusschereMarzoukNajmEtAl12,
+SalloumAlexanderianLeMaitreEtAl12,
+GuptaImValorani13} that aim to reduce the
+complexity of stiff chemical systems by filtering out the 
+fast timescales from the system. The latter is done, for instance, using
+the eigenvectors of the system Jacobian 
+to decouple the fast and slow processes; see e.g.,~\cite{DebusschereMarzoukNajmEtAl12}.}
 

### 2018-12-09 Revision/Response/response.tex
..

@@
 
-We are extremely thankful to the reviewers for recognizing potential in this effort and 
+We are very thankful to the reviewers for recognizing potential in this effort and 
 for their constructive feedback on the manuscript. A point-by-point
@@
 the perturbations associated with the uncertain rate-controlling parameters: pre-exponent
-and the activation energy. Second, the dimensionality of the problem has also been increased
-significantly. Specifically, in addition to the rate parameters, the impact of
+and the activation energy. Second, the dimensionality of the problem for the higher
+dimensional case has been increased from 33 to 36.
+Specifically, in addition to the rate parameters, the impact of
 perturbing the initial conditions: pressure, temperature, and stoichiometry pertaining to the
-H$_2$/O$_2$ reaction mechanism on the ignition delay is investigated. Such analysis using the
+H$_2$/O$_2$ reaction mechanism on the ignition delay is investigated. Such analyses using the
 proposed framework would provide useful insights to combustion scientists interested in
@@
 we perturb the initial conditions to investigate their impact on the response individually as well as due to
-their interaction with the rate parameters. Hence, we have also increased the dimensionality of the problem.
+their interaction with the rate parameters. Hence, we have also increased the dimensionality of the problem
+for the higher dimensional case from 33 to 36. More importantly, the impact of initial conditions on the
+variability in the system response could provide useful insights to combustion scientists interested in 
+reaction kinetics for a range of applications.
+
 As expected by the reviewer, the resulting spectral gap between the first two eigenvalues is observed to

### 2018-12-09 Revision/Response/response.tex
mine ...

--- 2018-12-09 Revision/Response/response.tex
@@
 
-We are extremely thankful to the reviewers for recognizing potential in this effort and 
+We are thankful to the reviewers for recognizing potential in this effort and 
 for their constructive feedback on the manuscript. A point-by-point
@@
 
-We sincerely hope that based on the modifications in the revised manuscript and our 
-response to reviewers' comments, the paper is found suitable for publication in
-{\it Combustion and Flame}.
+%We sincerely hope that based on the modifications in the revised manuscript and our 
+%response to reviewers' comments, the paper is found suitable for publication in
+%{\it Combustion and Flame}.
 
@@
 active subspace methodology are shown to be consistent with our earlier results using the derivative-based
-global sensitivity analysis (Ref. 35). The sensitivity results for the higher-dimensional problem in the
+global sensitivity analysis (Ref. 35{\color{blue} Update}). The sensitivity results for the higher-dimensional problem in the
 revised manuscript reveal that the ignition delay is largely impacted by
@@
 the goals and approaches of these methods are substantially different. However, 
-we are aware that CSP analysis is well-known in computational
-chemistry community, and therefore, have included a brief mention 
+we are aware that CSP analysis is well-known in combustion 
+community, and therefore, have included a brief mention 
 of CSP in the introduction, clarifying the differences between 

### 2018-12-09 Revision/Response/response.tex

--- 2018-12-09 Revision/Response/response.tex
@@
 
-We are very thankful to the reviewers for recognizing potential in this effort and 
-for their constructive feedback on the manuscript. A point-by-point
-response to the comments has been provided below. Corresponding modifications have been
-highlighted (in blue) in the revised manuscript. Major changes in the revised manuscript
-are summarized as follows:
+We are very thankful to the reviewers for recognizing potential in this effort
+and for their constructive feedback on the manuscript. A point-by-point
+response to the comments has been provided below. Corresponding modifications
+have been highlighted (in blue) in the revised manuscript. Major changes in the
+revised manuscript are summarized as follows:
 

### 2018-12-09 Revision/Response/response.tex
update ...

@@
 \begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
-\item \referee{There is a lack of quantitative discussion of the expense of regression-vs-perturbation,
-in terms of the actual number of model simulations required. The work seems to suggest that perturbation
-works better than regression, but then it is a direct and clear consequence of the fact that local 
-approximation is supposed to be better than a global one. Ideally, much more quantitative comparison
-needs to be made here before reaching any meaningful conclusions.}
+\item \referee{There is a lack of quantitative discussion of the expense of
+regression-vs-perturbation, in terms of the actual number of model simulations
+required. The work seems to suggest that perturbation works better than
+regression, but then it is a direct and clear consequence of the fact that
+local approximation is supposed to be better than a global one. Ideally, much
+more quantitative comparison needs to be made here before reaching any
+meaningful conclusions.}
 
-\noindent  As suggested by the reviewer, we have focused on a quantitative comparison in terms of the
-number of samples and associated model runs required to obtain a converged active subspace
-for the two approaches. In addition to convergence characteristics, we have also compared
-the two approaches for their predictive accuracy for the same computational effort for the
-19-dimensional problem as well as the 36-dimensional problem. In the revised work,
-we have increased the dimensionality of
-the higher-dimensional case from 33 to 36. Specifically, we have enhanced our analysis
-to study the impact of perturbations in the initial conditions: pressure, temperature, and
+\noindent  As suggested by the reviewer, we have focused on a quantitative
+comparison in terms of the number of samples and associated model runs required
+to obtain a converged active subspace for the two approaches. In addition to
+convergence characteristics, we have also compared the two approaches for their
+predictive accuracy for the same computational effort for the 19-dimensional
+problem as well as the 36-dimensional problem. In the revised work, we have
+increased the dimensionality of the higher-dimensional case from 33 to 36.
+Specifically, we have enhanced our analysis to study the impact of
+perturbations in the initial conditions: pressure, temperature, and
 stoichiometry on the system response i.e. ignition delay.
 
-\item \referee{Another fundamental flaw is related to the additive treatment of the uncertainty of 
-the pre-exponential factor A. In most works dealing with uncertainty in chemical kinetics, the
-uncertainty form is multiplicative or, equivalently, log(A) is the underlying fundamental parameter 
-with additive uncertainty range. In the current form of the uncertainty, the problem is too benign,
-and barely carries any practical usability for the practitioners.}
+\item \referee{Another fundamental flaw is related to the additive treatment of
+the uncertainty of the pre-exponential factor A. In most works dealing with
+uncertainty in chemical kinetics, the uncertainty form is multiplicative or,
+equivalently, log(A) is the underlying fundamental parameter with additive
+uncertainty range. In the current form of the uncertainty, the problem is too
+benign, and barely carries any practical usability for the practitioners.}
 
-\noindent The reviewer has made a valuable suggestion. Therefore, in the revised manuscript, we have
-updated the computations for the active subspace using perturbation and regression by considering
-the uncertainty in $\log(A)$ as opposed to $A$. As a result, the non-linearity of the system
-response is observed to increase significantly. As shown in Figures~3 and 6, 
-the variability of the ignition delay in the
-active subspace is no longer a straight line as observed earlier. Hence, we have substantially
-increased the complexity of the problem by increasing the non-linearity in the system response
-as well as its dimensionality to illustrate the robustness and the applicability of the proposed framework
-for a wide range of kinetics applications.
+\noindent The reviewer has made a valuable suggestion. Therefore, in the
+revised manuscript, we have updated the computations for the active subspace
+using perturbation and regression by considering the uncertainty in $\log(A_i)$
+as opposed to $A_i$.  As a result, the non-linearity of the system response is
+observed to increase significantly. As shown in Figures~3 and 6, the
+variability of the ignition delay in the active subspace is no longer a
+straight line as observed earlier. Hence, we have substantially increased the
+complexity of the problem by increasing the non-linearity in the system
+response as well as its dimensionality to illustrate the robustness and the
+applicability of the proposed framework for a wide range of kinetics
+applications.
 
@@
 
-\noindent  The title has been updated as ``Active subspace-based dimension reduction for chemical kinetics
-applications with epistemic uncertainty".
+\noindent  The title has been updated as ``Active subspace-based dimension
+reduction for chemical kinetics applications with epistemic uncertainty".
 
-\item \referee{3.36 $n_i$ is index or power? Is this an established term for the power of T? 
-Also, what are the values of $n_i$'s?}
+\item \referee{3.36 $n_i$ is index or power? Is this an established term for
+the power of T?  Also, what are the values of $n_i$'s?}
 
-\noindent  In the revised manuscript, we denote $n_i$ as the `temperature exponent' as suggested by
-the other reviewer. Yes, it is an established quantity and its value in addition to the
-rate constants for the 19 reactions
-in the H$_2$/O$_2$ mechanism is taken from [Yetter et. al., 1991] and implemented in the
-TChem software used for model predictions in this work. We have clarified this on Page 4
-in the revised manuscript.
+\noindent  In the revised manuscript, we denote $n_i$ as the `temperature
+exponent' as suggested by the other reviewer. Yes, it is an established
+quantity and its value in addition to the rate constants for the 19 reactions
+in the H$_2$/O$_2$ mechanism is taken from [Yetter et. al., 1991] and
+implemented in the TChem software used for model predictions in this work. We
+have clarified this on Page 4 in the revised manuscript.
 
@@
 
-\item \referee{4.57 not clear where the number 33 comes from since we have 19 reactions. I see it is explained at the beginning of Section 5, but it should be noted here briefly, too.}
+\item \referee{4.57 not clear where the number 33 comes from since we have 19
+reactions. I see it is explained at the beginning of Section 5, but it should
+be noted here briefly, too.}
 
-\noindent  We could not locate the number 33 in the location specified by the reviewer. Nevertheless, we would like
-to point out that on Page 4 (just below Table 1) in the introduction, 
-it is mentioned that the activation energies with a non-zero nominal value are considered as uncertain. 
-Since the uncertainty in the activation energy is considered only in the higher-dimensional case
-presented in Section 5, we have
-specified the reactions for which it is 0 in the beginning of that section.
+\noindent  We could not locate the number 33 in the location specified by the
+reviewer. Nevertheless, we would like to point out that on Page 4 (just below
+Table 1) in the introduction, it is mentioned that the activation energies with
+a non-zero nominal value are considered as uncertain.  Since the uncertainty in
+the activation energy is considered only in the higher-dimensional case
+presented in Section 5, we have specified the reactions for which it is 0 in
+the beginning of that section.
 
@@
 
-\item \referee{7.29-30: Any PDF can be written in a Boltzmann form, V(x) is just the negative log-PDF, 
-I am not sure if it is necessary to bring up a Boltzmann distribution language here.
-Also, it could be confused with Maxwell-Boltzmann distribution, which is a very specific PDF form.}
-Thanks for this comment. We followed the theory in ref [35] 
-in the manuscript, where the authors derive Poincare type 
-inequalities for functions of iid random
-inputs distributed according to Boltzmann probability laws. 
-We point out that the uniform distribution does not belong to this class. 
-The required Poincare type inequality for the case of uniform 
-distribution was derived in an earlier paper of Sobol' and Kucherenko (ref [34]). 
-That
-was the reason for assuming the distribution laws of the random inputs are
-either \emph{uniform} or belong to the Boltzmann class, in section 3. An
-important class of Boltzmann measures are the so called log-concave
-distributions such as Normal, Exponential, Beta, and Gamma distributions. These
-points have been clarified in the revised manuscript.
+\item \referee{7.29-30: Any PDF can be written in a Boltzmann form, V(x) is
+just the negative log-PDF, I am not sure if it is necessary to bring up a
+Boltzmann distribution language here.  Also, it could be confused with
+Maxwell-Boltzmann distribution, which is a very specific PDF form.} Thanks for
+this comment. We followed the theory in ref [35] in the manuscript, where the
+authors derive Poincare type inequalities for functions of iid random inputs
+distributed according to Boltzmann probability laws.  We point out that the
+uniform distribution does not belong to this class.  The required Poincare type
+inequality for the case of uniform distribution was derived in an earlier paper
+of Sobol' and Kucherenko (ref [34]).  That was the reason for assuming the
+distribution laws of the random inputs are either \emph{uniform} or belong to
+the Boltzmann class, in section 3. An important class of Boltzmann measures are
+the so called log-concave distributions such as Normal, Exponential, Beta, and
+Gamma distributions. These points have been clarified in the revised
+manuscript.
 
-Additionally, we share the reviewer's concern about the potential confusions some readers
-might face when seeing the mention of Boltzmann. We have removed the mention of
-Boltzmann measures from the introduction, and now only refer to Boltzman
-measures in Section
-3, where the theoretical results are discussed.
+Additionally, we share the reviewer's concern about the potential confusions
+some readers might face when seeing the mention of Boltzmann. We have removed
+the mention of Boltzmann measures from the introduction, and now only refer to
+Boltzman measures in Section 3, where the theoretical results are discussed.
 
@@
 
-\noindent  The equation as well as the following description of the expectation (E) and the variance (V)
-has been updated to address the reviewer's comment on Page~7 in the revised manuscript.
+\noindent  
+That equation as well as the following description of the conditional 
+expectation and the variance has been updated to address the reviewer's comment.
 

### 2018-12-09 Revision/Paper/assa.tex
updates ...

@@
 T_i(f) = 1 - 
-\frac{\V_{\rebut{\vec{\xi}_{\sim i}}}[\mathbb{E}_{\rebut{\xi_i}}(f|\vec{\xi}_{\sim i})]}{\V(f)},
+\frac{\V_{\rebut{\vec{\xi}_{\sim i}}}\big[\mathbb{E}[f|\vec{\xi}_{\sim i}]\big]}{\V(f)},
 \label{eq:total}
@@
 where $\vec{\xi}_{\sim i}$ is the input parameter vector with the  
-$i^\text{th}$ entry removed. \rebut{The expectation is taken over all possible
-values of $\xi_i$ and its variance is computed with respect to all parameters except $\xi_i$.}
+$i^\text{th}$ entry removed. \rebut{
+Here $\mathbb{E}[f|\vec{\xi}_{\sim i}]$ denotes the conditional
+expectation of $f$ given $\vec{\xi}_{\sim i}$
+and its variance is computed with respect to $\vec{\xi}_{\sim i}$.}
 The quantity, $\V(f)$ denotes the total variance of the model output.

### 2018-12-09 Revision/Paper/intro.tex
updates ...

--- 2018-12-08 Revision/Paper/intro.tex
@@
 the dimensionality of the problem, and hence the computational effort
-associated with uncertainty propagation.
+associated with uncertainty propagation. The focus here is on input parameter
+dimension reduction. \rebut{This is different from 
+techniques such at Computational
+Singular Perturbation (CSP)~\cite{Lam85,
+LamGoussis89,ValoraniCretaGoussisEtAl06,
+DebusschereMarzoukNajmEtAl12,
+SalloumAlexanderianLeMaitreEtAl12} that aim to reduce the
+complexity of stiff chemical systems by filtering out the 
+fast timescales from the system. The latter is done, for instance, using
+the eigenvectors of the system Jacobian 
+to decouple the fast and slow processes; see e.g.,~\cite{DebusschereMarzoukNajmEtAl12}.}
 
@@
 We seek to understand the impact of uncertainty in the
-rate-controlling parameters, pre-exponent~($A_i$) and the activation
-energy~($E_{a,i}$) \rebut{as well as the initial pressure, temperature, and the
-equivalence ratio} on the ignition delay. The $A_i$'s associated with all
+rate-controlling parameters, pre-exponents~($A_i$'s) 
+and the activation energies~($E_{a,i}$'s) 
+\rebut{as well as the initial pressure, temperature, and the
+equivalence ratio} on the ignition delay. The \rebut{$\log(A_i)$'s} associated with all
 reactions and the $E_{a,i}$'s with non-zero nominal estimates
@@
 the initial conditions: $P_0$, $T_0$, and $\Phi_0$. Note that the equivalence ratio
-corresponding to the initial molar ratios of $H_2$ and O$_2$ is denotes as $\Phi_0$.}
+corresponding to the initial molar ratios of $H_2$ and O$_2$ is denoted as $\Phi_0$.}
 As mentioned above, the
 first approach uses finite difference to estimate the gradient and is regarded as
-the perturbation approach, whereas, the second approach involves a linear
-regression-fit to the available set of model evaluations and is regarded as
-the regression approach.
+the perturbation approach. On the other hand, the second approach 
+follows from~\cite{Constantine:2015} and involves repeated
+regression-fits to a subset of available model evaluations, and is 
+regarded as the regression approach in this work.
 

### 2018-12-09 Revision/Paper/method.tex
updates ...

@@
 In this section, we present an efficient framework for computing the
-active subspace. The framework is employed to the H$_2$/O$_2$ reaction kinetics
+active subspace. The framework is employed to analyze 
+the H$_2$/O$_2$ reaction kinetics
 problem whereby the \rebut{logarithm of the} 
@@
 fits to subsets of available evaluations as discussed 
-in~\cite{Constantine:2015} (see Algorithm 1.2). Hence, the computational effort is  
+in~\cite[Algorithm 1.2]{Constantine:2015}. Hence, the computational effort is  
 reduced by a factor $(d+1)$ when using the regression-based approach.
@@
 not require additional model runs in this case. Thus, including the initial step, a total of 900
-model runs were required to obtained a converged active subspace in both cases.
+model runs were required to obtain a converged active subspace in both cases.
 

### 2018-12-09 Revision/Response/response.tex
...

@@
 \usepackage[margin=1in]{geometry}
+\usepackage{parskip}
 \usepackage{graphicx}
@@
 
-\noindent The complexity of the H$_2$/O$_2$ reaction kinetics application in the revised manuscript has been significantly
+The complexity of the H$_2$/O$_2$ reaction kinetics application in the revised manuscript has been significantly
 enhanced by increasing the degree of uncertainty in the rate parameters, and considering additional 
@@
 
-\noindent  A more complex problem as suggested by the reviewer would manifest itself into an increased number of
+A more complex problem as suggested by the reviewer would manifest itself into an increased number of
 uncertain parameters and/or an increased amount of non-linearity in the system response. Therefore, while
@@
 
-\noindent  Depending upon the initial conditions, the ignition delay can range from $\mathcal{O}(10^{-3})$
+Depending upon the initial conditions, the ignition delay can range from $\mathcal{O}(10^{-3})$
 to $\mathcal{O}(10^3)$ seconds for hydrogen-air mixtures as shown by Starostin et al. in~\cite{Starostin:2011}.
@@
 
-\noindent The reviewer's suggestion has been incorporated in the revised manuscript on Page 3.
+The reviewer's suggestion has been incorporated in the revised manuscript on Page 3.
 
@@
 
-%\noindent Minor comments by the reviewer have also been addressed in the revised manuscript.
+%Minor comments by the reviewer have also been addressed in the revised manuscript.
 
@@
 
-\noindent  As suggested by the reviewer, we have focused on a quantitative
+As suggested by the reviewer, we have focused on a quantitative
 comparison in terms of the number of samples and associated model runs required
@@
 
-\noindent The reviewer has made a valuable suggestion. Therefore, in the
+The reviewer has made a valuable suggestion. Therefore, in the
 revised manuscript, we have updated the computations for the active subspace
@@
 
-\noindent  The title has been updated as ``Active subspace-based dimension
+The title has been updated as ``Active subspace-based dimension
 reduction for chemical kinetics applications with epistemic uncertainty".
@@
 
-\noindent  In the revised manuscript, we denote $n_i$ as the `temperature
+In the revised manuscript, we denote $n_i$ as the `temperature
 exponent' as suggested by the other reviewer. Yes, it is an established
@@
 
-\noindent  The ignition delay is considered as the time required for the {\it rate of temperature}
+The ignition delay is considered as the time required for the {\it rate of temperature}
 increase to exceed a threshold, in this work.
@@
 
-\noindent  We could not locate the number 33 in the location specified by the
-reviewer. Nevertheless, we would like to point out that on Page 4 (just below
-Table 1) in the introduction, it is mentioned that the activation energies with
-a non-zero nominal value are considered as uncertain.  Since the uncertainty in
-the activation energy is considered only in the higher-dimensional case
-presented in Section 5, we have specified the reactions for which it is 0 in
-the beginning of that section.
+We mention, in Page 4 (just below Table 1) in the introduction, that only the
+activation energies with a non-zero nominal value are considered as uncertain,
+in the higher-dimensional case.  Since the uncertainty in the activation energy
+is considered only in the higher-dimensional case presented in Section 5, we
+have specified the reactions for which it is 0 in the beginning of that
+section.
 
@@
 
-\noindent  The revised manuscript has been updated to include the reviewer's suggestion.
+The revised manuscript has been updated to include the reviewer's suggestion.
 
@@
 
-\noindent  The typo has been fixed in the revised manuscript.
+The typo has been fixed in the revised manuscript.
 
@@
 
+We agree; we have removed 'statistical' from that part.
+
 \item \referee{7.29-30: Any PDF can be written in a Boltzmann form, V(x) is
@@
 Boltzmann distribution language here.  Also, it could be confused with
-Maxwell-Boltzmann distribution, which is a very specific PDF form.} Thanks for
+Maxwell-Boltzmann distribution, which is a very specific PDF form.}
+
+Thanks for
 this comment. We followed the theory in ref [35] in the manuscript, where the
@@
 
-\noindent  
 That equation as well as the following description of the conditional 
@@
 
-\noindent Following the reviewer's suggestion, we have updated the text in the revised manuscript to
+Following the reviewer's suggestion, we have updated the text in the revised manuscript to
 clearly note that the DGSM $\nu_i(f)$ is the $i^{\text{th}}$ diagonal element in the matrix, $\mathbf{C}$
@@
 
-\noindent A large gap in the eigenvalue spectrum typically indicates that a low-dimensional (1--2) active subspace 
+A large gap in the eigenvalue spectrum typically indicates that a low-dimensional (1--2) active subspace 
 would sufficiently capture the variability in the model output due to uncertainty in the inputs. However,
@@
 
-\noindent The reviewer is correct. Algorithm 1 in the revised manuscript has been updated to include $\beta$
+The reviewer is correct. Algorithm 1 in the revised manuscript has been updated to include $\beta$
 as an input.
@@
 
-\noindent We have used the L$^2$ norm to estimate the relative change in the squared values of corresponding
+We have used the L$^2$ norm to estimate the relative change in the squared values of corresponding
 eigenvector components between successive iterations. The algorithm in the revised manuscript has been updated 
@@
 
-\noindent The max is over all eigenvectors in the active subspace denoted by $j$. This has been specified in
+The max is over all eigenvectors in the active subspace denoted by $j$. This has been specified in
 the updated algorithm in the revised manuscript.
@@
 
-\noindent The reviewer's suggestion has been incorporated in the revised manuscript.
+The reviewer's suggestion has been incorporated in the revised manuscript.
 
@@
 
-\noindent Based on the reviewer's feedback, in the revised manuscript, the analysis has been
+Based on the reviewer's feedback, in the revised manuscript, the analysis has been
 updated to
@@
 
-\item \referee{For Figure 8b, the reasoning why E$_15$ was not captured by the regression approach,
+\item \referee{For Figure 8b, the reasoning why E$_{15}$ was not captured by the regression approach,
 is not satisfactory. I think the authors could have gone a bit deeper in order to explain this
@@
 
-\noindent Our latest analysis in the revised manuscript has revealed that the two approaches
+Our latest analysis in the revised manuscript has revealed that the two approaches
 in fact yield consistent results
@@
 
-\noindent We have made a sincere effort to incorporate these valuable suggestions by the 
+We have made a sincere effort to incorporate these valuable suggestions by the 
 reviewer in the revised manuscript. 
-Specifically, we perturb $\log(A)$ instead of $A$ which is observed to significantly increase
+Specifically, we perturb $\log(A_i)$ instead of $A_i$ 
+which is observed to significantly increase
 the non-linearity of the system response. In addition to the rate parameters, we analyze the impact
@@
 
-\noindent Based on our conversations with the developers of the TChem software at Sandia
-National Laboratories, we decided not to perturb the activation energies with a 0
-nominal value since the exponential factor is not considered for these reactions, as mentioned
-by the reviewer.
+The latter statement is correct. We discussed this in our group extensively,
+and also consulted with the developers of the TChem software at Sandia National
+Laboratories, and decided not to perturb the activation energies with a 0
+nominal value since the exponential factor is not considered for these
+reactions, as mentioned by the reviewer.
 
@@
 
-\noindent We have provided examples for perturbation-based techniques such as finite difference and
-adjoint-based methods for estimating the
-gradient of the model output with respect to uncertain parameters and inputs.
+Our goal was to list some efficient alternatives to finite-difference gradient
+computations. We have modified that part to make things clearer.
 

### 2018-12-09 Revision/Paper/assa.tex
...

@@
 While the active subspace framework described above does not make any assumptions
-about statistical independence of the inputs $\xi_i$, $i = 1, \ldots, \Np$, 
+about independence of the inputs $\xi_i$, $i = 1, \ldots, \Np$, 
 the classical 

### 2018-12-09 Revision/Paper/conc.tex
..
...

@@
 the QoI i.e. the ignition delay. For this purpose, we explored two approaches,
-namely, perturbation-based (e.g. finite difference - used in this work, adjoint-based
-methods, etc.) and regression-based. Both approaches were shown to yield consistent
-results for the 19-dimensional
-problem wherein only the pre-exponents were considered to be uncertain.
-\rebut{It was observed that the computational effort required to obtain a 
-converged active subspace was comparable for the two approaches. However,
-the predictive accuracy of the perturbation approach was observed to be
-slightly higher. Moreover, a 1-dimensional active subspace was shown to
-reasonably approximate the uncertainty in the ignition delay.}
-Additionally, the activity scores were also shown to be consistent with the
-screening metric estimates based on DGSMs in~\cite{Vohra:2018}. An
-iterative procedure was adopted to enhance the computational efficiency.
+namely, perturbation and regression. Both approaches were shown to yield
+consistent results for the 19-dimensional problem wherein only the
+pre-exponents were considered to be uncertain.  \rebut{It was observed that the
+computational effort required to obtain a converged active subspace was
+comparable for the two approaches. However, the predictive accuracy of the
+perturbation approach was found to be relatively higher. Moreover, a
+1-dimensional active subspace was shown to reasonably approximate the
+uncertainty in the ignition delay.} Additionally, the activity scores were also
+shown to be consistent with the screening metric estimates based on DGSMs
+in~\cite{Vohra:2018}. An iterative procedure was adopted to enhance the
+computational efficiency.
 
@@
 
-Therefore, in general, perturbation-based approaches should be preferred for 
-active subspace computation especially in situations where efficient gradient
-computation techniques (adjoints-based, automatic differentiation, etc.) are
-feasible. The regression-based approach can be explored in situations involving intensive
-simulations where the perturbation-based methods are intractable.
+Based on our findings, the perturbation approach is preferable for active
+subspace computation; moreover, this can be made extremely efficient, when more
+efficient gradient computation techniques (adjoints-based, automatic
+differentiation, etc.) are feasible. The regression-based approach can be
+explored in situations involving intensive simulations where gradient
+computation is very challenging.
+
 %
@@
 The computational framework presented in this work is agnostic to the choice of
-the chemical system and can be easily adapted for other systems as long
-as the quantity of interest is continuously differentiable in the considered
-domain of the inputs.  We have demonstrated that the active subspace could be
-exploited for efficient forward propagation of the uncertainty from inputs to
-the output. The resulting activity scores and the low-dimensional surrogate
-could further guide optimal allocation of computational resources for
-calibration of the important rate-controlling parameters \rebut{and input conditions}
- in a Bayesian setting.
-Additionally, dimension reduction using active subspaces could assist in
-developing robust formulations for predicting discrepancy between simulations
-and measurements due to epistemic uncertainty in the model inputs.
+the chemical system and can be easily adapted for other systems as long as the
+quantity of interest is continuously differentiable in the considered domain of
+the inputs.  We have demonstrated that the active subspace could be exploited
+for efficient forward propagation of the uncertainty from inputs to the output.
+The resulting activity scores and the low-dimensional surrogate could further
+guide optimal allocation of computational resources for calibration of the
+important rate-controlling parameters \rebut{and input conditions} in a
+Bayesian setting.  Additionally, dimension reduction using active subspaces
+could assist in developing robust formulations for predicting discrepancy
+between simulations and measurements due to epistemic uncertainty in the model
+inputs.

### 2018-12-09 Revision/Paper/intro.tex
...

@@
 respect to the input parameters. Several numerical techniques are available
-for computing the gradient. Perturbation-based techniques
-include finite differences and more advanced methods involving
+for computing the gradient, such as 
+finite differences and more advanced methods involving
 adjoints~\cite{Jameson:1988,Borzi:2011,Alexanderian:2017}. The
@@
 at neighboring points are required if finite difference is used which increases
-the computational effort. Regression-based techniques on the other hand,
-aim to estimate the
-gradient by approximating the model output using a regression fit. These
-are computationally less intensive than the former. However, as expected,
-there is a trade-off between computational effort and accuracy in the two
-approaches for estimating the gradient.
+the computational effort. 
+Regression-based techniques, which can be suitable for active subspace computations,
+on the other hand, aim to estimate the gradient by approximating the model
+output using a regression fit.  
+These are computationally less intensive than
+the former. However, as expected, there is a trade-off between computational
+effort and accuracy in the two approaches for estimating the gradient.
 
 In this work, we adopt an iterative strategy to reduce the computational effort
-associated with active subspace computation. Moreover, we explore two approaches
-for estimating the gradient of the ignition delay with respect to the uncertain
-rate-controlling parameters: pre-exponent, $A_i$ and the activation energy, $E_{a,i}$,
-for $i^{\text{th}}$ reaction in the H$_2$/O$_2$ mechanism \rebut{as well as
-the initial conditions: $P_0$, $T_0$, and $\Phi_0$. Note that the equivalence ratio
-corresponding to the initial molar ratios of $H_2$ and O$_2$ is denoted as $\Phi_0$.}
-As mentioned above, the
-first approach uses finite difference to estimate the gradient and is regarded as
-the perturbation approach. On the other hand, the second approach 
-follows from~\cite{Constantine:2015} and involves repeated
-regression-fits to a subset of available model evaluations, and is 
+associated with active subspace computation. Moreover, we explore two
+approaches for estimating the gradient of the ignition delay with respect to
+the uncertain rate-controlling parameters: pre-exponent, $A_i$ and the
+activation energy, $E_{a,i}$, for $i^{\text{th}}$ reaction in the H$_2$/O$_2$
+mechanism \rebut{as well as the initial conditions: $P_0$, $T_0$, and $\Phi_0$.
+Note that the equivalence ratio corresponding to the initial molar ratios of
+$H_2$ and O$_2$ is denoted as $\Phi_0$.} The first approach uses finite
+difference to estimate the gradient and will be referred to as the
+\emph{perturbation} approach throughout the article.  The second approach
+is adapted from~\cite[Algorithm 1.2]{Constantine:2015} and involves
+repeated regression-fits to a subset of available model evaluations, and is
 regarded as the regression approach in this work.

### 2018-12-10 Revision/Paper/app.tex
..

@@
 and the initial conditions were perturbed by 2$\%$ about their nominal values.
-Note that the mangnitude of the perturbation was selected such that the
+Note that the magnitude of the perturbation was selected such that the
 ignition delay assumes a physically meaningful value in the input domain. 
@@
 were taken from~\cite{Yetter:1991}. The nominal values of $P_0$, $T_0$, and
-$\Phi_0$ were considered to be 1.0 atm, 900~K, and 2.0 respectively.}
+$\Phi_0$ were considered to be 1.0~atm, 900~K, and 2.0 respectively.}
 
@@
 
-The 1-dimensional surrogate ($\hat{G}$) shown in Figure~\ref{fig:hd} is investigated
-for its accuracy as well as the ability to capture the uncertainty in the
-model output in two ways. Firstly, we estimate the relative L$^2$ norm of the discrepancy~($\varepsilon_d$)
-between estimates of ignition delay in the case of H$_2$/O$_2$ reaction, obtained using the 
-model output and the surrogate in the following equation:
-%
-\be
-\varepsilon_d = \frac{\|G(\bm{\xi}) - \hat{G}(\bm{\xi})\|_2}{\|G(\bm{\xi})\|_2}
-\ee
-%
-Model evaluations ($G$) at 10$^{4}$ random samples in the input domain were used to evaluate $\varepsilon_d$. 
-\rebut{It was estimated to be approximately 0.17 for both approaches. Hence, the predictive accuracy of
-the two approaches is comparable with respect to the relative L$^2$ norm of the discrepancy.
- However, as mentioned earlier, the computational effort required for
-obtaining a converged active subspace using perturbation was half of that required when using
-regression to estimate the model gradient at random samples in the 36-dimensional input domain.}
-
-Secondly, we verified the accuracy of the two surrogates in a probabilistic setting. In particular, we compared 
-probability density functions (PDFs) obtained using the true set of model evaluations, and 1-dimensional
-surrogates ($\hat{G}$'s) based on the two approaches for gradient estimation as shown in 
+The 1-dimensional surrogate ($\hat{G}$) shown in Figure~\ref{fig:hd} for the perturbation
+and regression approaches is investigated for its ability to capture the uncertainty in the
+ignition delay. Specifically, we compare probability density functions (PDFs)
+obtained using the true set of model evaluations, and 1-dimensional
+surrogates ($\hat{G}$'s) based on the two approaches, as shown in 
 Figure~\ref{fig:pdf_36D}. Note that the three PDFs were evaluated using the same set of 10$^4$ samples in the 
@@
 estimate and the uncertainty (quantified by the spread in the distributions) is found to be consistent for the three 
-cases. To confirm this, we compute the first-order (mean) and the second-order (standard deviation) statistics
+cases. To confirm this, we further compute the first-order (mean) and the second-order (standard deviation) 
+statistics
 of the estimates of the ignition delay obtained using the model, 1-dimensional surrogate from perturbation, and
@@
 \rebut{The mean and the standard deviation estimates obtained using the model and the 1-dimensional surrogates
-are found to be in close agreement. Hence, although the relative L$^2$ norm of the discrepancy~($\varepsilon_d$)
-is found to be considerably large, the uncertainty in the ignition delay is captured by both 
-approaches with reasonable accuracy. 
+are found to be in close agreement. Hence, the uncertainty in the ignition delay is accurately captured 
+in both cases.
 It is however interesting to note that estimates from the regression approach are

### 2018-12-10 Revision/Response/response.tex
..

@@
 
-%We sincerely hope that based on the modifications in the revised manuscript and our 
-%response to reviewers' comments, the paper is found suitable for publication in
-%{\it Combustion and Flame}.
+We sincerely hope that based on the modifications in the revised manuscript and our 
+response to reviewers' comments, the paper is found suitable for publication in
+{\it Combustion and Flame}.
 
@@
 active subspace methodology are shown to be consistent with our earlier results using the derivative-based
-global sensitivity analysis (Ref. 35{\color{blue} Update}). The sensitivity results for the higher-dimensional problem in the
+global sensitivity analysis~\cite{Vohra:2018}. The sensitivity results for the higher-dimensional problem in the
 revised manuscript reveal that the ignition delay is largely impacted by
@@
 we did not implement the framework to a new system, we have significantly increased the non-linearity of the
-response by increasing the degree of uncertainty in the rate parameters. In addition to the rate parameters,
+response by increasing the degree of uncertainty in the rate parameters. Specifically, instead of
+the pre-exponent, $A_i$, we perturb $\log(A_i)$ as suggested by the other reviewer. Moreover, 
+the magnitude of the perturbation has been increased from 1$\%$ to $2\%$.
+In addition to the rate parameters,
 we perturb the initial conditions to investigate their impact on the response individually as well as due to
@@
 
+The model predictions for ignition delay in this work have been obtained using the 
+TChem software as mentioned in the paper.
 Depending upon the initial conditions, the ignition delay can range from $\mathcal{O}(10^{-3})$
@@
 respectively. These values have been reported in the revised manuscript in the beginning of
-Section 5 on page 15.
-In this revision, we perturb the initial conditions to study their impact
-on ignition delay estimates. The mean estimate of the ignition delay using data at 10$^4$ random samples in
-the 36-dimensional space is found to be approximately 0.15 as reported in Table 2 on page 18 in the
-revised manuscript. As mentioned in the paper, we have used 
-the TChem software for model predictions and based on our recent communication with the developers at
-Sandia National Laboratories, these estimates seem reasonable.
+Section 5 on page 15. Based on these initial conditions, we have verified with the developers of
+TChem at Sandia National Laboratories that our estimates for the ignition delay are reasonable.
 
@@
 
-Thanks for this comment. As the referee notes, while the spectral 
+As the reviewer notes, while the spectral 
 truncations in active subspace methodology resemble aspects of CSP analysis,
@@
 community, and therefore, have included a brief mention 
-of CSP in the introduction, clarifying the differences between 
+of CSP in the introduction along with references, and clarified the differences between 
 the active subspace method and CSP.

### 2018-12-10 Revision/Paper/conc.tex
...

@@
 Based on our findings, the perturbation approach is preferable for active
-subspace computation; moreover, this can be made extremely efficient, when more
-efficient gradient computation techniques (adjoints-based, automatic
-differentiation, etc.) are feasible. The regression-based approach can be
-explored in situations involving intensive simulations where gradient
-computation is very challenging.
+subspace computation; the computational cost of this approach can be reduced
+significantly, if more efficient gradient computation techniques (e.g.,
+adjoint-based approaches or automatic differentiation) are feasible. The
+regression-based approach can be explored in situations involving intensive
+simulations where gradient computation is very challenging.
 
@@
 %
-We also mention that alternate regression based approaches such as ones based
+We also mention that alternate regression-based approaches such as ones based
 on computing a global quadratic model have been proposed and used in the

### 2018-12-10 Revision/Paper/intro.tex
...

@@
 \be
-\Phi = \frac{(M_{\text{H}_2}/M_{\text{O}_2})_\text{obs}}{(M_{\text{H}_2}/M_{\text{O}_2})_\text{st}}.
+\Phi = \frac{(M_{\text{H}_2}/M_{\text{O}_2})_\text{obs}}{(M_{\text{H}_2}/M_{\text{O}_2})_\text{st}},
 \label{eq:phi}
@@
 approaches for estimating the gradient of the ignition delay with respect to
-the uncertain rate-controlling parameters: pre-exponent, $A_i$ and the
-activation energy, $E_{a,i}$, for $i^{\text{th}}$ reaction in the H$_2$/O$_2$
-mechanism \rebut{as well as the initial conditions: $P_0$, $T_0$, and $\Phi_0$.
+the uncertain rate-controlling parameters: pre-exponents ($A_i$'s), 
+the activation energies ($E_{a,i}$'s),
+\rebut{as well as the initial conditions: $P_0$, $T_0$, and $\Phi_0$.
 Note that the equivalence ratio corresponding to the initial molar ratios of
 $H_2$ and O$_2$ is denoted as $\Phi_0$.} The first approach uses finite
-difference to estimate the gradient and will be referred to as the
+differences to estimate the gradient and will be referred to as the
 \emph{perturbation} approach throughout the article.  The second approach

### 2018-12-10 Revision/Response/response.tex
..

@@
 kinetics applications. Moreover, it illustrates the applicability and robustness of the 
-framework to complex kinetics applications.
+proposed framework to complex kinetics applications.
 
@@
 the pre-exponent, $A_i$, we perturb $\log(A_i)$ as suggested by the other reviewer. Moreover, 
-the magnitude of the perturbation has been increased from 1$\%$ to $2\%$.
+the magnitude of the perturbation has been increased from 1$\%$ to $2\%$ in the
+higher-dimensional case.
 In addition to the rate parameters,

### 2018-12-10 Revision/Response/response.tex
..

@@
 the mention of Boltzmann measures from the introduction, and now only refer to
-Boltzman measures in Section 3, where the theoretical results are discussed.
+Boltzmann measures in Section 3, where the theoretical results are discussed.
 

### 2018-12-10 Revision/Paper/app.tex
..

@@
 -- $\mathcal{R}_9$, $\mathcal{R}_{13}$ (due to zero nominal values for $E_a$),
-and the initial conditions were perturbed by 2$\%$ about their nominal values.
+and the initial conditions were considered to be uniformly distributed, and
+perturbed by 2$\%$ about their nominal values.
 Note that the magnitude of the perturbation was selected such that the

### 2018-12-10 Revision/Paper/method.tex
...
..

@@
 
-In this section, we present an efficient framework for computing the
-active subspace. The framework is employed to analyze 
-the H$_2$/O$_2$ reaction kinetics
+In this section, we outline the methodology for computing the
+active subspace in an efficient manner. The proposed framework is employed to analyze 
+a 19-dimensional H$_2$/O$_2$ reaction kinetics
 problem whereby the \rebut{logarithm of the} 
@@
 %
-\rebut{The dominant eigenvector obtained using perturbation is based on $N$~=~45 samples which requires
-$M$~=~900 model runs. For the same amount of computational effort, we can afford $N$~=~900 samples
-when using regression. Hence, the SSP from regression is based on 900
-points: ($\hat{\mat{W}}_1^\top\bm{\xi}$,~$G(\hat{\mat{W}}_1^\top\bm{\xi})$). On the other hand, the SSP
-from perturbation is plotted using only 45 points as mentioned earlier.
-Nevertheless, the illustrative comparison clearly indicates that the two SSPs
-are in excellent agreement. Moreover, it is interesting to note
-that the response in ignition
-delay based on the considered probability distributions for $\log(A_i)$ although non-linear, can be approximated by
-a 1-dimensional active subspace.}
+\rebut{The dominant eigenvector obtained using perturbation is based on
+$N$~=~45 samples which requires $M$~=~900 model runs. For the same amount of
+computational effort, we can afford $N$~=~900 samples when using regression.
+Hence, the SSP from regression is based on 900 points:
+($\hat{\mat{W}}_1^\top\bm{\xi}_j$,~$G(\hat{\mat{W}}_1^\top\bm{\xi}_j)$),
+$j = 1, \ldots, 900$. On the
+other hand, the SSP from perturbation is plotted using only 45 points as
+mentioned earlier.  Nevertheless, the illustrative comparison clearly indicates
+that the two SSPs are in excellent agreement. Moreover, it is interesting to
+note that the response in ignition delay based on the considered probability
+distributions for $\log(A_i)$ although non-linear, can be approximated by a
+1-dimensional active subspace.}
 

### 2019-01-21 KineticsPaperArXiv/abstract.tex
..

--- 2018-11-30 Revision/Paper/abstract.tex
@@
 We focus on an efficient approach for quantification of uncertainty in complex chemical
-reaction networks with a large number of uncertain parameters \rebut{and input conditions}.  Parameter
+reaction networks with a large number of uncertain parameters and input conditions.  Parameter
 dimension reduction is accomplished by computing an~\emph{active subspace} that
@@
 uncertainty in the pre-exponents of the individual reaction rates is
-considered. This is followed by the analysis of a \rebut{36}-dimensional case wherein
-the activation energies \rebut{and initial conditions} are also considered uncertain.  In both cases, a
+considered. This is followed by the analysis of a 36-dimensional case wherein
+the activation energies and initial conditions are also considered uncertain.  In both cases, a
 1-dimensional active subspace is observed to capture the uncertainty in the QoI,

### 2019-01-21 KineticsPaperArXiv/ack.tex
..

\section*{Acknowledgment}

M.~Vohra and S.~Mahadevan gratefully acknowledge funding support from the
National Science Foundation (Grant No. 1404823, CDSE Program), and Sandia
National Laboratories (PO No. 1643376, Technical monitor: Dr. Joshua Mullins).
The research of A.~Alexanderian was partially supported by the
National Science Foundation through the grant DMS-1745654.  The authors also
thank Dr.~Cosmin Safta at Sandia National Laboratories for his guidance
pertaining to the usage of the TChem software package.

### 2019-01-21 KineticsPaperArXiv/app.tex
..

--- 2018-12-10 Revision/Paper/app.tex
@@
 (i) pre-exponents ($A_i$'s); (ii) the activation energies
-($E_{a,i}$'s); and (iii) \rebut{the initial pressure~($P_0$),
-temperature~($T_0$), and stoichiometry~($\Phi_0$)}. 
+($E_{a,i}$'s); and (iii) the initial pressure~($P_0$),
+temperature~($T_0$), and stoichiometry~($\Phi_0$). 
 %
-\rebut{The $\log(A_i)$'s, $E_{a,i}$'s for all reactions except $\mathcal{R}_6$
+The $\log(A_i)$'s, $E_{a,i}$'s for all reactions except $\mathcal{R}_6$
 -- $\mathcal{R}_9$, $\mathcal{R}_{13}$ (due to zero nominal values for $E_a$),
-and the initial conditions were perturbed by 2$\%$ about their nominal values.
+and the initial conditions were considered to be uniformly distributed, and
+perturbed by 2$\%$ about their nominal values.
 Note that the magnitude of the perturbation was selected such that the
@@
 were taken from~\cite{Yetter:1991}. The nominal values of $P_0$, $T_0$, and
-$\Phi_0$ were considered to be 1.0~atm, 900~K, and 2.0 respectively.}
+$\Phi_0$ were considered to be 1.0~atm, 900~K, and 2.0 respectively.
 
@@
 by tracking the quantity `$\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$'. 
-\rebut{In Figure~\ref{fig:conv_app}~(right), we examine $\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$
+In Figure~\ref{fig:conv_app}~(right), we examine $\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$
 with increasing iterations for the perturbation and the regression approaches 
-discussed earlier in Section~\ref{sec:method}}. 
+discussed earlier in Section~\ref{sec:method}. 
 %
@@
  \begin{center}
-  \includegraphics[width=0.8\textwidth]{./Figures/eig_conv36Dp2}
-\caption{\rebut{Left: An illustrative comparison of individual components of the 
+  \includegraphics[width=0.8\textwidth]{eig_conv36Dp2}
+\caption{Left: An illustrative comparison of individual components of the 
 dominant eigenvector in the converged active subspace i.e., at the end
@@
 $\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$ assumes a value smaller
-than 0.05.}}
+than 0.05.}
 \label{fig:conv_app}
@@
 %
-\rebut{At each iteration, we improve our estimates of the matrix $\hat{\mat{C}}$ by
+At each iteration, we improve our estimates of the matrix $\hat{\mat{C}}$ by
 estimating the gradient of the ignition delay at 5 new randomly generated samples 
@@
 obtained using the two approaches. The components are observed to be in excellent
-agreement with each other.}
+agreement with each other.
 
-\rebut{In Figure~\ref{fig:hd}, we plot the SSP for the perturbation approach (left) and the regression
+In Figure~\ref{fig:hd}, we plot the SSP for the perturbation approach (left) and the regression
 approach (center) in a 1-dimensional active subspace. A 1-dimensional polynomial fit is also
@@
 From these results, it is clear that a 1-dimensional active subspace captures the variability in the
-ignition delay with reasonable accuracy, and that the two approaches yield consistent results.}
+ignition delay with reasonable accuracy, and that the two approaches yield consistent results.
 %
@@
  \begin{center}
-   \includegraphics[width=0.9\textwidth]{./Figures/SSP_plot36Dp2}
-\caption{\rebut{Sufficient summary plots (SSPs) for the case of perturbation (left) and regression (center).
+   \includegraphics[width=0.9\textwidth]{SSP_plot36Dp2}
+\caption{Sufficient summary plots (SSPs) for the case of perturbation (left) and regression (center).
 A polynomial fit of degree 2 and 3 as shown in the plots is used as a surrogate in the 
 perturbation and regression approaches respectively. An illustrative comparison of the two
-surrogates is also provided (right).}}
+surrogates is also provided (right).}
 \label{fig:hd}
@@
 \begin{center}
-\includegraphics[width=3.0in]{./Figures/pdf_plot36Dp2}
+\includegraphics[width=3.0in]{pdf_plot36Dp2}
 \end{center} 
-\caption{\rebut{A comparison of the PDFs of ignition delay, obtained using model 
+\caption{A comparison of the PDFs of ignition delay, obtained using model 
 evaluations (solid line) and 1-dimensional surrogates using the regression-based strategy (dashed line) and the
 perturbation-based strategy (dashed-dotted line). The same set of 10$^4$ samples in the cross-validation set were 
-used in each case.}}
+used in each case.}
 \label{fig:pdf_36D}
@@
 %
-\rebut{The PDFs are observed to be in close agreement with each other. Specifically, the modal
+The PDFs are observed to be in close agreement with each other. Specifically, the modal
 estimate and the uncertainty (quantified by the spread in the distributions) is found to be consistent for the three 
@@
 1-dimensional surrogate from regression at the cross-validation sample set. The mean and standard deviation
-estimates are provided in Table~\ref{tab:stats}.}
+estimates are provided in Table~\ref{tab:stats}.
 %
@@
 \bottomrule
-$G$~(Model) & \rebut{0.1510} & \rebut{0.1373} \\
-$\hat{G}$~(Perturbation-based) & \rebut{0.1514} & \rebut{0.1275} \\
-$\hat{G}$~(Regression-based) & \rebut{0.1510} & \rebut{0.1330} \\
+$G$~(Model) & 0.1510 & 0.1373 \\
+$\hat{G}$~(Perturbation-based) & 0.1514 & 0.1275 \\
+$\hat{G}$~(Regression-based) & 0.1510 & 0.1330 \\
 \bottomrule
@@
 %
-\rebut{The mean and the standard deviation estimates obtained using the model and the 1-dimensional surrogates
+The mean and the standard deviation estimates obtained using the model and the 1-dimensional surrogates
 are found to be in close agreement. Hence, the uncertainty in the ignition delay is accurately captured 
@@
 a much wider range of values of the ignition delay compared to the surrogate based on perturbation as shown in
-Figure~\ref{fig:hd}~(right).}
+Figure~\ref{fig:hd}~(right).
 
@@
  \begin{center}
-  \includegraphics[width=0.8\textwidth]{./Figures/as_36Dp2}
-\caption{\rebut{Bar graphs illustrating individual activity scores for the uncertain 
-rate parameters and the initial conditions for the H$_2$/O$_2$ reaction.}}
+  \includegraphics[width=0.8\textwidth]{as_36Dp2}
+\caption{Bar graphs illustrating individual activity scores for the uncertain 
+rate parameters and the initial conditions for the H$_2$/O$_2$ reaction.}
 \label{fig:as_36D}
@@
 %
-Several useful inferences can be drawn. \rebut{Firstly, the normalized activity scores from the two approaches
+Several useful inferences can be drawn. Firstly, the normalized activity scores from the two approaches
 and the total-effect Sobol' indices are found to be in close agreement with each other. Secondly, as expected, 
@@
 considered uncertain initial conditions, the ignition delay is mainly impacted by the perturbations in the initial 
-temperature in the considered interval.}
+temperature in the considered interval.

### 2019-01-21 KineticsPaperArXiv/assa.tex
..

--- 2018-12-09 Revision/Paper/assa.tex
@@
 $\mu(d\vec{\xi}) = \pi(\vec{\xi})d\vec{\xi}$, where $\pi(\vec{\xi})$ is the joint probability
-\rebut{density function} of $\vec{\xi}$. Herein, $f$ is assumed to be a square integrable 
+density function of $\vec{\xi}$. Herein, $f$ is assumed to be a square integrable 
 function with continuous partial 
@@
 The columns of $\mat{W}_1 = 
-\begin{bmatrix} \vec{w}_\rebut{1} \cdots \vec{w}_r\end{bmatrix}$ 
+\begin{bmatrix} \vec{w}_1 \cdots \vec{w}_r\end{bmatrix}$ 
 span the dominant eigenspace of $\mat{C}$ and
@@
 where $V$ is a continuous function and $C$ a normalization 
-constant~\cite{Lamboni:2013}. \rebut{
+constant~\cite{Lamboni:2013}. 
 An important class of Boltzmann distributions are the so called log-concave
@@
 Weibull distributions. Note also that the uniform distribution does not fall under
-the class of Boltzmanm distributions.~\cite{Lamboni:2013}.}
+the class of Boltzmanm distributions~\cite{Lamboni:2013}.
 
@@
 T_i(f) = 1 - 
-\frac{\V_{\rebut{\vec{\xi}_{\sim i}}}\big[\mathbb{E}[f|\vec{\xi}_{\sim i}]\big]}{\V(f)},
+\frac{\V_{\vec{\xi}_{\sim i}}\big[\mathbb{E}[f|\vec{\xi}_{\sim i}]\big]}{\V(f)},
 \label{eq:total}
@@
 where $\vec{\xi}_{\sim i}$ is the input parameter vector with the  
-$i^\text{th}$ entry removed. \rebut{
+$i^\text{th}$ entry removed. 
 Here $\mathbb{E}[f|\vec{\xi}_{\sim i}]$ denotes the conditional
 expectation of $f$ given $\vec{\xi}_{\sim i}$
-and its variance is computed with respect to $\vec{\xi}_{\sim i}$.}
+and its variance is computed with respect to $\vec{\xi}_{\sim i}$.
 The quantity, $\V(f)$ denotes the total variance of the model output.
@@
 \]
-Here $\pi$ is the joint PDF of $\vec\xi$. \rebut{Note that $\nu_i(f)$ is the 
-$i^{\text{th}}$ diagonal element of the matrix $\mat{C}$} as defined in~\eqref{eq:C}. 
+Here $\pi$ is the joint PDF of $\vec\xi$. Note that $\nu_i(f)$ is the 
+$i^{\text{th}}$ diagonal element of the matrix $\mat{C}$ as defined in~\eqref{eq:C}. 
 Consider the spectral decomposition written 

### 2019-01-21 KineticsPaperArXiv/method.tex
..

--- 2018-12-10 Revision/Paper/method.tex
@@
 a 19-dimensional H$_2$/O$_2$ reaction kinetics
-problem whereby the \rebut{logarithm of the} 
+problem whereby the logarithm of the 
 pre-exponent ($A_i$) in the rate law associated with individual reactions provided in
 Table~\ref{tab:kinetics} is considered to be uniformly
-distributed in the interval, \rebut{$[0.97\log(A_i^\ast), 1.03\log(A_i^\ast)]$};
+distributed in the interval, $[0.97\log(A_i^\ast), 1.03\log(A_i^\ast)]$;
 $A_i^\ast$ is the nominal
@@
 Two approaches are explored for estimating the gradient of ignition delay with
-respect to \rebut{$\log(A_i)$}: 
+respect to $\log(A_i)$: 
 a perturbation  approach that involves computation of model
@@
 subsequent eigenvalues,
-\rebut{$\left(\frac{\lambda_j}{\lambda_{j+1}}\right)\ge\mathcal{O}(10^1)$}.
+$\left(\frac{\lambda_j}{\lambda_{j+1}}\right)\ge\mathcal{O}(10^1)$.
  At each subsequent iteration, model evaluations are
@@
   \begin{algorithmic}[1]
-\Require $\theta_l$, $\theta_u$, \rebut{$\beta$}, $\tau$. 
+\Require $\theta_l$, $\theta_u$, $\beta$, $\tau$. 
 \Ensure $\hat{\mat{\Lambda}}$, $\hat{\mat{W}}$, $\bm{\nu}_r(f)$ %$\eta$. %
@@
                        \scalebox{1.25}{$\frac{\|(\hat{\mat{W}}_{1,j}^{k})^2 - 
-                       (\hat{\mat{W}}_{1,j}^{k-1})^2\|_\rebut{2}}{\|(\hat{\mat{W}}_{1,j}^{k-1})^2\|_\rebut{2}}$}, 
+                       (\hat{\mat{W}}_{1,j}^{k-1})^2\|_2}{\|(\hat{\mat{W}}_{1,j}^{k-1})^2\|_2}$}, 
                        $j = 1,\ldots,r$.
-		\If {$\max\limits_{\rebut{j}}\left(\delta \hat{\mat{W}}_{1,j}^{(k)}\right)<\tau$}
+		\If {$\max\limits_{j}\left(\delta \hat{\mat{W}}_{1,j}^{(k)}\right)<\tau$}
 			\State break
@@
 the 19-dimensional H$_2$/O$_2$ reaction kinetics
-problem by perturbing \rebut{$\log(A_i)$ by 3$\%$ about its nominal
-value as discussed earlier}. For the purpose of verification,
+problem by perturbing $\log(A_i)$ by 3$\%$ about its nominal
+value as discussed earlier. For the purpose of verification,
 $\hat{\mat{C}}$ was initially constructed using a large set of samples
@@
 In Figure~\ref{fig:eig_comp}, we illustrate the comparison of
-the resulting normalized eigenvalue spectrum \rebut{by plotting $\log(\lambda_i/\lambda_0)$} ($i = 1,\ldots,19$) corresponding to $N$~=~1000 and the same quantity corresponding to
+the resulting normalized eigenvalue spectrum by plotting $\log(\lambda_i/\lambda_0)$ ($i = 1,\ldots,19$) corresponding to $N$~=~1000 and the same quantity corresponding to
 a much smaller set of samples, $n$~=~$\{20,40,80,120\}$.
@@
  \begin{center}
-  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp_p3}
+  \includegraphics[width=0.45\textwidth]{eig_comp_p3}
 %   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1}
@@
 are approximated 
-reasonably well with just 20 samples. As expected, the accuracy of higher-\rebut{index}
+reasonably well with just 20 samples. As expected, the accuracy of higher-index
  eigenvalues is observed
 to improve with the sample size. Since 
-$\lambda_1$ is roughly \rebut{an order of magnitude} larger than $\lambda_2$, we expect 
-a 1-dimensional active subspace \rebut{to reasonably approximate the uncertainty
-in the ignition delay}. 
+$\lambda_1$ is roughly an order of magnitude larger than $\lambda_2$, we expect 
+a 1-dimensional active subspace to reasonably approximate the uncertainty
+in the ignition delay. 
 To further confirm this, we evaluate a relative L$^2$ norm of the difference
@@
 %
-\rebut{The quantity, $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ defined in Algorithm~\ref{alg:grad} was
+The quantity, $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ defined in Algorithm~\ref{alg:grad} was
 used to assess the convergence behavior of the two approaches. Using a set tolerance, $\tau$~=~0.05,
@@
 two approaches and for the case using $N$~=~1000 in the perturbation approach are plotted in
-Figure~\ref{fig:comp}~(left).}
+Figure~\ref{fig:comp}~(left).
 %
@@
  \begin{center}
-  \includegraphics[width=0.38\textwidth]{./Figures/eig_conv19D}
-  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp19D}
-\caption{\rebut{Left: An illustrative comparison of individual squared components of the converged dominant
+  \includegraphics[width=0.38\textwidth]{eig_conv19D}
+  \includegraphics[width=0.45\textwidth]{comp_ssp19D}
+\caption{Left: An illustrative comparison of individual squared components of the converged dominant
 eigenvector obtained using perturbation and regression strategies using $M$~=~900 model runs in each case. 
@@
 perturbation and the regression strategies for computing the active subspace. 
-}}
+}
 \label{fig:comp}
@@
 %
-\rebut{
 The set of eigenvector components for the three cases are found to be in excellent agreement with
 each other, indicating that both
-approaches have sufficiently converged and are reasonably accurate for this setup.}
+approaches have sufficiently converged and are reasonably accurate for this setup.
 
@@
 H$_2$/O$_2$ reaction in this case, varies
- predominantly in \rebut{a 1-dimensional} active subspace. Hence, 
+ predominantly in a 1-dimensional active subspace. Hence, 
 $f(\bm{\xi})$ can be approximated as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
-the \rebut{1-dimensional active} subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, 
+the 1-dimensional active subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, 
 regarded as the \textit{sufficient summary plot} (SSP), obtained using the perturbation-based and
@@
 %
-\rebut{The dominant eigenvector obtained using perturbation is based on
+The dominant eigenvector obtained using perturbation is based on
 $N$~=~45 samples which requires $M$~=~900 model runs. For the same amount of
@@
 distributions for $\log(A_i)$ although non-linear, can be approximated by a
-1-dimensional active subspace.}
+1-dimensional active subspace.
 
 We further estimate
-the normalized activity scores for individual uncertain inputs ($\tilde{\nu}_{i,r}$; $r$=1 \rebut{since a
-1-dimensional active subspace seems reasonably accurate})
+the normalized activity scores for individual uncertain inputs ($\tilde{\nu}_{i,r}$; $r$=1 since a
+1-dimensional active subspace seems reasonably accurate)
 using the components of the dominant eigenvector as shown in Algorithm~\ref{alg:grad} (steps 21 and 22).
@@
  \begin{center}
-  \includegraphics[width=0.45\textwidth]{./Figures/comp_as19D}
+  \includegraphics[width=0.45\textwidth]{comp_as19D}
 %  \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}
-\caption{\rebut{Left: A bar-graph of normalized activity scores ($\tilde{\nu}_{i,r}$'s) 
+\caption{Left: A bar-graph of normalized activity scores ($\tilde{\nu}_{i,r}$'s) 
 for the 19 uncertain pre-exponents ($A_i$'s); $r$
-denotes the number of eigenvectors in the dominant eigenspace.}}
+denotes the number of eigenvectors in the dominant eigenspace.}
 \label{fig:comp_as}
@@
 based on the screening metric involving the DGSMs in~\cite{Vohra:2018}. It is observed that the uncertainty
-associated with the ignition delay is largely due to the uncertainty in $A_9$ \rebut{while $A_1$, $A_{15}$, and
-$A_{17}$ are also observed to contribute significantly towards its variance.}
+associated with the ignition delay is largely due to the uncertainty in $A_9$ while $A_1$, $A_{15}$, and
+$A_{17}$ are also observed to contribute significantly towards its variance.
 
@@
 uncertain $A_i$'s is reasonably approximated using both, perturbation  and regression 
-approaches in this case. \rebut{Since both approaches yield consistent results and are
+approaches in this case. Since both approaches yield consistent results and are
 comparable in terms of convergence and accuracy, we could use either for the purpose
-of active subspace computation for this setting.}
+of active subspace computation for this setting.
 In the following section, we shift our focus to 
-the higher-dimensional H$_2$/O$_2$ reaction kinetics application \rebut{wherein the activation
+the higher-dimensional H$_2$/O$_2$ reaction kinetics application wherein the activation
 energies in the rate law as well as initial pressure, temperature, and stoichiometric conditions
-are also considered to be uncertain.}
+are also considered to be uncertain.

### 2019-02-05 Revision2/Rebuttal/response.tex
..

\documentclass[11pt,final]{article}
\renewcommand*\familydefault{\sfdefault}
%\usepackage{amssymb,amsmath,amsfonts,comment}
%\usepackage{amsmath,amssymb,graphicx,subfigure,psfrag}
\usepackage{amsmath,amssymb,graphicx,subfigure,psfrag,upgreek}
\usepackage{algorithm,algorithmic}
\usepackage{amssymb,mathrsfs}
\usepackage[margin=1in]{geometry}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{color,pdfcolmk}
\usepackage{enumitem,kantlipsum}
\newcommand{\todo}[1]{\noindent\emph{\textcolor{red}{Todo: #1\:}}}
%\newcommand{\alennote}[1]{\noindent\emph{\vspace{1ex}\textcolor{cyan}{Alen: #1\:}}\\[1ex]}
\newcommand{\referee}[1]{\vspace{.1ex}\noindent{\textcolor{blue}{#1}}}

\begin{document}

%We thank the reviewers for their careful reading of our article 
%and the helpful comments and suggestions.
%Please find below point-by-point replies (in black) to your comments and
%questions (which are reprinted in blue). To give you an overview of all the
%changes in the paper, we also provide a diff-document that highlights the
%changes between the initial submission and this re-submission.\\[1ex]
\begin{center}
{\bf Summary of Modifications to CNF-D-18-00757}\\[6pt]
{\bf Subspace-based dimension reduction for chemical kinetics applications with 
epistemic uncertainty}\\[6pt]
By \\
Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan 
\end{center}

%\baselineskip=22pt

\vspace*{1in}

We are once again thankful to the reviewers for their positive feedback on the revised 
manuscript. Through this response, we sincerely hope to have addressed the
outstanding issues and concerns. 
A point-by-point
response to the comments has been provided as follows. Corresponding modifications
where appropriate, have been highlighted (in blue) in the revised manuscript.

\clearpage

\section*{Reviewer \#1}

\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\item \referee{I recommend accepting the manuscript, provided that the authors explore and provide evidence (to this 
Reviewer at least) that the applicability and conclusions on the method's performance do not change when the initial 
temperature and pressure are such that the ignition delay time falls in the 0.1 to 1 ms range, which is 2 to 3 orders of 
magnitude lower than the ignition delay time that they have presented.
It is my wish to see the data, perhaps just as a response to me facilitated by the editor, before recommending publication.
Perhaps there has been a misunderstanding. When I pointed to the O(0.1 s) delay time, I did not imply that there was an 
error in the calculation and/or a bug in TChem. What I meant is simply that the initial conditions were likely to be unrealistic.
In the revised revision, the authors mention that the conditions are 1 atm, 900 K, and equivalence ratio of 2.0. These 
conditions, giving an ignition delay time of O(0.1 s) are obviously highly unrealistic and impractical as there is no application 
that may rely on such long ignition delay times.
Conditions (higher pressures/higher temperatures) such that an ignition delay time O(0.1 - 1 ms) are practical and should be 
explored by the authors.
It may very well be that nothing changes in the analysis and conclusions, but then, again, why consider an unrealistic set of 
conditions in the first place$?$
}

We thank the reviewer for this comment and suggestions. In accordance with the reviewer's suggestion, the initial 
temperature~($T_0$) and pressure~($P_0$) was increased to ensure that the ignition delay is in the desired
range, i.e. 0.1--1~ms. Specifically, the nominal values of $T_0$ and $P_0$ were increased from 900~K to 1000~K, and
from 1~atm to 1.5~atm respectively. In keeping with the paper, a 2$\%$ perturbation about these nominal values was used
for the purpose of computing the active subspace. Prior distributions of the equivalence ratio and the rate-controlling
parameters were used as before. In Figure~\ref{fig:kde}, we illustrate the probability density function of the ignition delay
obtained using 10,000 pseudorandom samples in the input domain. 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.40\textwidth]{./Figures/pdf_ign_small}
  \includegraphics[width=0.38\textwidth]{./Figures/pdf_ign_small_zoom}
\caption{Left: PDF plots generated using kernel density estimation of the ignition delay computed at 10,000 samples.
Right: The same PDF is illustrated using a smaller range in the x-axis to highlight the region of interest, i.e. 0.1~ms--1~ms.
The range of interest is highlighted by the two vertical lines.} 
\label{fig:kde}
\end{center}
\end{figure}
%
As observed in the pdf plots in Figure~\ref{fig:kde}, the bulk of the high probability region is captured by the range of interest
for the ignition delay~(0.1~ms--1~ms), highlighted using the two vertical lines corresponding to the bounds of the interval.
Hence, we consider $P_0=1.5$~atm and $T_0=1000$~K as reasonable
choices for the purpose of this analysis.

The proposed methodology in the paper was implemented to this new set-up. The regression approach was used for
estimating the gradient using model evaluations at 500 samples. In Figure~\ref{fig:as}, we illustrate the
eigenvalue spectrum~(left), and the sufficient summary plot~(right).
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.4\textwidth]{./Figures/eig_plot_rev2}
  \includegraphics[width=0.4\textwidth]{./Figures/ssp_rev2}
\caption{Left: Eigenvalue spectrum of the matrix, $\hat{\bm{C}}$~(see~Eq.~8 in the paper). Right: Sufficient summary
plot~(SSP) in the active subspace. A polynomial fit (surrogate) is also shown.} 
\label{fig:as}
\end{center}
\end{figure}
%
The eigenvalue spectrum shows that $\frac{\lambda_1}{\lambda_2}\approx 10^{4}$ thereby indicating the presence of 
a 1-dimensional active subspace in this new regime as well. The SSP further confirms that the variability in ignition delay
is predominantly captured by a 1-dimensional active subspace. A polynomial fit of degree 3 is used as a surrogate for
estimating the global sensitivity measures of the individual rate-controlling parameters as well as the initial conditions:
$P_0$, $T_0$, and $\phi_0$. The sensitivity estimates based on activity scores are plotted in Figure~\ref{fig:as}.
 %
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.85\textwidth]{./Figures/as_36D_rev2}
\caption{} 
\label{fig:as}
\end{center}
\end{figure}
%
It is observed that the ignition delay is mainly sensitive towards the rate parameters, $A_1$ and $A_9$. Sensitivity
towards $A_{15}$ and $A_{17}$ is also found to be significant. Among the initial conditions, $T_0$ is found to be
the most important. These findings for the newly considered regime are in fact consistent with the sensitivity results
obtained for the regime considered in the paper. Therefore, it is interesting to note that parametric sensitivities have
not changed much as a result of faster ignition due to increased initial pressure and temperature.

Thus, the proposed methodology in the paper is successfully implemented in a regime of potential interest to
combustion scientists. It further highlights the robustness as well as the applicability of the framework to a wide
range of conditions for the kinetics application considered in this work. Moreover, the framework can be extended
to other systems
for the purpose of sensitivity analysis, uncertainty propagation, and parameter optimization.

\end{enumerate}

\section*{Reviewer \#2}

\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\item \referee{Fig 1: I don't think the why axis label is correct. There should be no 'log'. Same goes for the added blue text
on line 34 (page 12)}

The reviewer is correct in pointing out that there should be no $\log$ in the y axis label in
Figure~1 and the text in line 34 on Page~12.
The manuscript has been updated to reflect these corrections.

\item \referee{page 7, line 55: typo 'Boltzmanm'}

The typo has been corrected in the revised manuscript.

\item \referee{page 13, line 17: max is over j, it should be mentioned. Same comment for the beginning of Section 5.1, 
and for Figure 4.}

As suggested by the reviewer, the manuscript has been updated to indicate that the max is over `j' at the specified 
locations.

\item \referee{page 13, line 19: no comma after 'both'}

This suggestion has been incorporated in the revised manuscript.

\item \referee{page 13, line 23: typo 'as total of'$\rightarrow$'a total of'}

The typo has been rectified in the revised manuscript.

\item \referee{Table 2: since mean and st.dev. are computed using 10$^4$ samples, I would only report two significant digits in 
the table, since sqrt(10$^4$)=10$^2$; the rest of the digits are within the noise estimate due to sampling. Of course this is only an 
estimate, but the authors could easily check by repeating the procedure with different valudation sets. As an alternative, to 
differentiate between the three cases, I'd recommend using a larger validation set, if possible.}

As suggested by the reviewer, Table~2 has been updated to report only up to two significant digits.

\item \referee{page 17 (and perhaps elsewhere): why is it called cross-validation and not just validation?}

The term ``cross-validation" has been used as opposed to ``validation" since an independent set of model evaluations
were used to assess the predictive accuracy of its surrogate. The term ``validation" is typically used in
situations where a model is tested for its predictability against experimental test data i.e. data not used for the purpose
of calibrating the model.

\item \referee{Please improve the x-axis labels in Figure 7.}

Figure~7 has been updated to address the reviewer's concern.

\end{enumerate}

%\bibliographystyle{elsarticle-num}
%\bibliography{REFER}
\end{document}

### 2019-02-05 Revision2/Paper/app.tex
..

--- 2019-01-21 KineticsPaperArXiv/app.tex
@@
 Algorithm~\ref{alg:grad}. The convergence of the eigenvectors was examined
-by tracking the quantity `$\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$'. 
-In Figure~\ref{fig:conv_app}~(right), we examine $\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$
+by tracking the quantity `$\rebut{\max\limits_j}(\delta \hat{\mat{W}}_{1,j}^{(i)})$'. 
+In Figure~\ref{fig:conv_app}~(right), we examine $\rebut{\max\limits_j}(\delta \hat{\mat{W}}_{1,j}^{(i)})$
 with increasing iterations for the perturbation and the regression approaches 
@@
  \begin{center}
-  \includegraphics[width=0.8\textwidth]{eig_conv36Dp2}
+  \includegraphics[width=0.8\textwidth]{./Figures/eig_conv36Dp2}
 \caption{Left: An illustrative comparison of individual components of the 
@@
 regression approaches. Convergence is accomplished once 
-$\max(\delta \hat{\mat{W}}_{1,j}^{(i)})$ assumes a value smaller
+$\rebut{\max\limits_j}(\delta \hat{\mat{W}}_{1,j}^{(i)})$ assumes a value smaller
 than 0.05.}
@@
  \begin{center}
-   \includegraphics[width=0.9\textwidth]{SSP_plot36Dp2}
+   \includegraphics[width=0.9\textwidth]{./Figures/SSP_plot36Dp2}
 \caption{Sufficient summary plots (SSPs) for the case of perturbation (left) and regression (center).
@@
 \begin{center}
-\includegraphics[width=3.0in]{pdf_plot36Dp2}
+\includegraphics[width=3.0in]{./Figures/pdf_plot36Dp2}
 \end{center} 
@@
 \bottomrule
-$G$~(Model) & 0.1510 & 0.1373 \\
-$\hat{G}$~(Perturbation-based) & 0.1514 & 0.1275 \\
-$\hat{G}$~(Regression-based) & 0.1510 & 0.1330 \\
+$G$~(Model) & \rebut{0.15} & \rebut{0.14} \\
+$\hat{G}$~(Perturbation-based) & \rebut{0.15} & \rebut{0.13} \\
+$\hat{G}$~(Regression-based) & \rebut{0.15} & \rebut{0.13} \\
 \bottomrule
@@
 in both cases.
-It is however interesting to note that estimates from the regression approach are
-observed to be slightly more accurate. This is potentially due to the fact that the 
-converged active subspace in this case is 
-based on 1850 samples in the 36-dimensional input domain, whereas, the converged active subspace based the 
-perturbation approach is based on only 25 samples. As a result, the surrogate based on regression explores
-a much wider range of values of the ignition delay compared to the surrogate based on perturbation as shown in
-Figure~\ref{fig:hd}~(right).
 
@@
  \begin{center}
-  \includegraphics[width=0.8\textwidth]{as_36Dp2}
+  \includegraphics[width=0.8\textwidth]{./Figures/as_36Dp2_rev2}
 \caption{Bar graphs illustrating individual activity scores for the uncertain 

### 2019-02-05 Revision2/Paper/conc.tex
..

--- 2018-12-10 Revision/Paper/conc.tex
@@
 rate-controlling parameters in the H$_2$/O$_2$ reaction mechanism
-\rebut{as well as the initial pressure, temperature, and stoichiometry} and its
+as well as the initial pressure, temperature, and stoichiometry and its
 impact on ignition delay predictions. The mechanism involves 19 different
@@
 pre-exponent and an activation energy. Hence, in theory, the evolution of the
-chemical system depends upon 38 rate parameters \rebut{and three initial conditions}.
+chemical system depends upon 38 rate parameters and three initial conditions.
 However, we considered 
 epistemic uncertainty in all pre-exponents and activation energies with non-zero
-nominal values i.e. a total of 33 rate parameters instead of 38 \rebut{in addition to
-the three initial conditions}.  
+nominal values i.e. a total of 33 rate parameters instead of 38 in addition to
+the three initial conditions.  
 %Conventional means for uncertainty quantification such as those involving
@@
 consistent results for the 19-dimensional problem wherein only the
-pre-exponents were considered to be uncertain.  \rebut{It was observed that the
+pre-exponents were considered to be uncertain. It was observed that the
 computational effort required to obtain a converged active subspace was
@@
 1-dimensional active subspace was shown to reasonably approximate the
-uncertainty in the ignition delay.} Additionally, the activity scores were also
+uncertainty in the ignition delay. Additionally, the activity scores were also
 shown to be consistent with the screening metric estimates based on DGSMs
@@
 wherein all pre-exponents and activation energies with non-zero nominal
-estimates \rebut{as well as the initial conditions} were considered uncertain. 
-\rebut{Once again, consistent results were obtained using the two approaches.
+estimates as well as the initial conditions were considered uncertain. 
+Once again, consistent results were obtained using the two approaches.
 A 1-dimensional active subspace was shown to reasonably
@@
 contributions from $A_{15}$, $A_{17}$, and $E_{a,15}$. Additionally, the
-ignition delay was found to be sensitive towards $T_0$.}
+ignition delay was found to be sensitive towards $T_0$.
 
@@
 guide optimal allocation of computational resources for calibration of the
-important rate-controlling parameters \rebut{and input conditions} in a
+important rate-controlling parameters and input conditions in a
 Bayesian setting.  Additionally, dimension reduction using active subspaces

### 2019-02-05 Revision2/Paper/intro.tex
..

--- 2018-12-10 Revision/Paper/intro.tex
@@
 associated with uncertainty propagation. The focus here is on input parameter
-dimension reduction. \rebut{This is different from 
+dimension reduction. This is different from 
 techniques such at Computational
@@
 the eigenvectors of the system Jacobian 
-to decouple the fast and slow processes; see e.g.,~\cite{DebusschereMarzoukNajmEtAl12}.}
+to decouple the fast and slow processes; see e.g.,~\cite{DebusschereMarzoukNajmEtAl12}.
 
@@
 %
-where $A_i$ is the pre-exponent, $n_i$ is the \rebut{temperature
-exponent}, $E_{a,i}$ is the
+where $A_i$ is the pre-exponent, $n_i$ is the temperature
+exponent, $E_{a,i}$ is the
 activation energy corresponding to the $i^{th}$ reaction, and $R$ is the
-universal gas constant.  \rebut{The Arrhenius rate law in~\eqref{eq:rate} 
+universal gas constant.  The Arrhenius rate law in~\eqref{eq:rate} 
 is often interpreted in a logarithmic form as follows:
@@
 \ee
-}
 %
@@
 $\Phi$~=~2.0. Homogeneous ignition at constant pressure is simulated using the
-TChem software package~\cite{Safta:2011} \rebut{using an initial pressure, $P_0$~=~1~atm and
-initial temperature, $T_0$~=~900~K.} The time required for the rate of
+TChem software package~\cite{Safta:2011} using an initial pressure, $P_0$~=~1~atm and
+initial temperature, $T_0$~=~900~K. The time required for the rate of
 temperature increase to exceed a given threshold, regarded as \emph{ignition delay}
@@
 and the activation energies~($E_{a,i}$'s) 
-\rebut{as well as the initial pressure, temperature, and the
-equivalence ratio} on the ignition delay. The \rebut{$\log(A_i)$'s} associated with all
+as well as the initial pressure, temperature, and the
+equivalence ratio on the ignition delay. The $\log(A_i)$'s associated with all
 reactions and the $E_{a,i}$'s with non-zero nominal estimates
 are considered to be uniformly distributed about their nominal estimates provided
-in~\cite{Yetter:1991}. \rebut{Temperature exponent, $n_i$ for each reaction is fixed to
+in~\cite{Yetter:1991}. Temperature exponent, $n_i$ for each reaction is fixed to
 its nominal value, also provided in~\cite{Yetter:1991}.
 The initial conditions are also considered to be uniformly
-distributed about their respective aforementioned values.} 
-The total number of uncertain inputs is \rebut{36} which makes
+distributed about their respective aforementioned values. 
+The total number of uncertain inputs is 36 which makes
 the present problem computationally challenging due to the large number of 
-uncertain parameters \rebut{in addition to the initial conditions}.  
+uncertain parameters in addition to the initial conditions.  
 To address this challenge, we focus on reducing the dimensionality
@@
 the activation energies ($E_{a,i}$'s),
-\rebut{as well as the initial conditions: $P_0$, $T_0$, and $\Phi_0$.
+as well as the initial conditions: $P_0$, $T_0$, and $\Phi_0$.
 Note that the equivalence ratio corresponding to the initial molar ratios of
-$H_2$ and O$_2$ is denoted as $\Phi_0$.} The first approach uses finite
+$H_2$ and O$_2$ is denoted as $\Phi_0$. The first approach uses finite
 differences to estimate the gradient and will be referred to as the
@@
 activity scores for functions of independent random inputs whose distribution
-law belongs to a \rebut{broad class of probability measures.} 
+law belongs to a broad class of probability measures. 
 It is worth mentioning that computing global sensitivity measures provides 
@@
 Active subspace
-discovery in a high-dimensional H$_2$/O$_2$ kinetics problem involving \rebut{36}
+discovery in a high-dimensional H$_2$/O$_2$ kinetics problem involving 36
 uncertain inputs: The methodology presented in this work
-\rebut{successfully demonstrated that a 1-dimensional active subspace can reasonably 
-approximate the uncertainty in the QoI}, indicating immense potential
+successfully demonstrated that a 1-dimensional active subspace can reasonably 
+approximate the uncertainty in the QoI, indicating immense potential
 for computational savings. The presented
@@
 for a 36-dimensional H$_2$/O$_2$ reaction kinetics problem in section~\ref{sec:app}.
-\rebut{For both settings, the convergence characteristics and the predictive accuracy
+For both settings, the convergence characteristics and the predictive accuracy
 of the two approaches is compared for a given amount of computational effort. The two
 approaches are observed to yield consistent results, and a 1-dimensional active subspace
-is observed to capture the uncertainty in the ignition delay.}
+is observed to capture the uncertainty in the ignition delay.
 Finally, a summary and discussion based on our findings is included in

### 2019-02-05 Revision2/Paper/method.tex
..

--- 2019-01-21 KineticsPaperArXiv/method.tex
@@
 In Figure~\ref{fig:eig_comp}, we illustrate the comparison of
-the resulting normalized eigenvalue spectrum by plotting $\log(\lambda_i/\lambda_0)$ ($i = 1,\ldots,19$) corresponding to $N$~=~1000 and the same quantity corresponding to
+the resulting normalized eigenvalue spectrum by plotting 
+\rebut{$(\lambda_i/\lambda_0)$}
+($i = 1,\ldots,19$) corresponding to $N$~=~1000 and the same quantity corresponding to
 a much smaller set of samples, $n$~=~$\{20,40,80,120\}$.
@@
  \begin{center}
-  \includegraphics[width=0.45\textwidth]{eig_comp_p3}
+  \includegraphics[width=0.45\textwidth]{./Figures/eig_comp_p3}
 %   \includegraphics[width=0.48\textwidth]{./Figures/err_eigv_1}
-\caption{A comparison of the normalized eigenvalue spectrum, $\log(\lambda_i/\lambda_0)$
+\caption{A comparison of the normalized eigenvalue spectrum, 
+\rebut{($\lambda_i/\lambda_0)$}
 using $n$~=~$\{20,40,80,120\}$ samples with that
@@
 %
-The quantity, $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ defined in Algorithm~\ref{alg:grad} was
+The quantity, $\rebut{\max\limits_j}(\delta \hat{\mat{W}}_{1,j}^{(r)})$ defined in Algorithm~\ref{alg:grad} was
 used to assess the convergence behavior of the two approaches. Using a set tolerance, $\tau$~=~0.05,
-it was observed that both, perturbation and regression approaches took 8 iterations to converge. 
+it was observed that both perturbation and regression approaches took 8 iterations to converge. 
 Note that the computational effort at each iteration was considered to be the same in both cases.
 More specifically, 5 new random samples were added for the perturbation approach at each iteration.
-However, as discussed earlier, as total of 100 (=5$\times$(19+1)) model runs were needed to obtain the
+However, as discussed earlier, \rebut{a} total of 100 (=5$\times$(19+1)) model runs were needed to obtain the
 model prediction and its gradients at these newly generated samples. Hence, in the case of
@@
  \begin{center}
-  \includegraphics[width=0.38\textwidth]{eig_conv19D}
-  \includegraphics[width=0.45\textwidth]{comp_ssp19D}
+  \includegraphics[width=0.38\textwidth]{./Figures/eig_conv19D}
+  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp19D}
 \caption{Left: An illustrative comparison of individual squared components of the converged dominant
@@
  \begin{center}
-  \includegraphics[width=0.45\textwidth]{comp_as19D}
+  \includegraphics[width=0.45\textwidth]{./Figures/comp_as19D}
 %  \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}

### 2019-03-07 Revision2/Paper/app.tex
..

@@
 temperature in the considered interval.
+
+\rebut{As shown in the PDF plotted in Figure~\ref{fig:pdf_36D}, the ignition delay assumes a wide range of values, i.e.
+from about 2~ms to 400~ms. However, for many practical applications, a much smaller ignition delay (0.1~ms--1~ms) might
+be of interest. The authors would like to point out that the proposed framework was also implemented to such a regime
+by using a nominal value of the initial temperature, $T_0$ = 1000~K, and the initial pressure, $P_0$ = 1.5~atm. 
+Our analysis for this
+regime once again revealed that a 1-dimensional active subspace was able to capture the variability in the ingition delay due to
+the uncertainty in the rate-controlling parameters and the input conditions. The sensitivity trends were also found to
+be qualitatively similar to those presented in Figure~\ref{fig:as_36D}. We have not included these results in the interest
+of brevity. Therefore, the proposed methodology was tested
+for is robustness and applicability for a wide range of conditions pertaining to the considered application.}

### 2019-03-07 Revision2/Paper/app.tex
..

@@
 Algorithm~\ref{alg:grad}. The convergence of the eigenvectors was examined
-by tracking the quantity `$\rebut{\max\limits_j}(\delta \hat{\mat{W}}_{1,j}^{(i)})$'. 
-In Figure~\ref{fig:conv_app}~(right), we examine $\rebut{\max\limits_j}(\delta \hat{\mat{W}}_{1,j}^{(i)})$
+by tracking the quantity `$\max\limits_j(\delta \hat{\mat{W}}_{1,j}^{(i)})$'. 
+In Figure~\ref{fig:conv_app}~(right), we examine $\max\limits_j(\delta \hat{\mat{W}}_{1,j}^{(i)})$
 with increasing iterations for the perturbation and the regression approaches 
@@
 regression approaches. Convergence is accomplished once 
-$\rebut{\max\limits_j}(\delta \hat{\mat{W}}_{1,j}^{(i)})$ assumes a value smaller
+$\max\limits_j(\delta \hat{\mat{W}}_{1,j}^{(i)})$ assumes a value smaller
 than 0.05.}
@@
 \bottomrule
-$G$~(Model) & \rebut{0.15} & \rebut{0.14} \\
-$\hat{G}$~(Perturbation-based) & \rebut{0.15} & \rebut{0.13} \\
-$\hat{G}$~(Regression-based) & \rebut{0.15} & \rebut{0.13} \\
+$G$~(Model) & 0.15 & 0.14 \\
+$\hat{G}$~(Perturbation-based) & 0.15 & 0.13 \\
+$\hat{G}$~(Regression-based) & 0.15 & 0.13 \\
 \bottomrule
@@
 
-\rebut{As shown in the PDF plotted in Figure~\ref{fig:pdf_36D}, the ignition delay assumes a wide range of values, i.e.
+As shown in the PDF plotted in Figure~\ref{fig:pdf_36D}, the ignition delay assumes a wide range of values, i.e.
 from about 2~ms to 400~ms. However, for many practical applications, a much smaller ignition delay (0.1~ms--1~ms) might
@@
 Our analysis for this
-regime once again revealed that a 1-dimensional active subspace was able to capture the variability in the ingition delay due to
+regime once again revealed that a 1-dimensional active subspace was able to capture the variability in the ignition delay due to
 the uncertainty in the rate-controlling parameters and the input conditions. The sensitivity trends were also found to
@@
 of brevity. Therefore, the proposed methodology was tested
-for is robustness and applicability for a wide range of conditions pertaining to the considered application.}
+for is robustness and applicability for a wide range of conditions pertaining to the considered application.

### 2019-03-07 Revision2/Paper/method.tex
..

@@
 the resulting normalized eigenvalue spectrum by plotting 
-\rebut{$(\lambda_i/\lambda_0)$}
+$(\lambda_i/\lambda_0)$
 ($i = 1,\ldots,19$) corresponding to $N$~=~1000 and the same quantity corresponding to
@@
 \caption{A comparison of the normalized eigenvalue spectrum, 
-\rebut{($\lambda_i/\lambda_0)$}
+($\lambda_i/\lambda_0)$
 using $n$~=~$\{20,40,80,120\}$ samples with that
@@
 %
-The quantity, $\rebut{\max\limits_j}(\delta \hat{\mat{W}}_{1,j}^{(r)})$ defined in Algorithm~\ref{alg:grad} was
+The quantity, $\max\limits_j(\delta \hat{\mat{W}}_{1,j}^{(r)})$ defined in Algorithm~\ref{alg:grad} was
 used to assess the convergence behavior of the two approaches. Using a set tolerance, $\tau$~=~0.05,
@@
 More specifically, 5 new random samples were added for the perturbation approach at each iteration.
-However, as discussed earlier, \rebut{a} total of 100 (=5$\times$(19+1)) model runs were needed to obtain the
+However, as discussed earlier, a total of 100 (=5$\times$(19+1)) model runs were needed to obtain the
 model prediction and its gradients at these newly generated samples. Hence, in the case of
@@
 %  \includegraphics[width=0.8\textwidth]{./Figures/eig_convD19}
-%\caption{\rebut{Left: An illustrative comparison of individual squared components of the converged dominant
+%\caption{Left: An illustrative comparison of individual squared components of the converged dominant
 %eigenvector obtained using perturbation and regression strategies using $M$~=~900 model runs in each case. 
